### 2016-09-17 Rmd/JTB_ms.Rmd
Initial commit

---
title: A simplified dynamic bioenergetic model for coral-*Symbiodinium* symbioses
author:
  - name: Ross Cunning
    email: ross.cunning@gmail.com
    affiliation: University of Hawaii
    footnote: Corresponding Author
  - name: Erik B. Muller
    affiliation: University of California
  - name: Ruth D. Gates
    affiliation: University of Hawaii
  - name: Roger M. Nisbet
    affiliation: University of California
address:
  - code: University of Hawaii
    address: Hawaii Institute of Marine Biology, Kaneohe, HI 96744, USA
  - code: University of California
    address: Department of Ecology, Evolution, and Marine Biology, Santa Barbara, CA 93106, USA
abstract: |
  This is the abstract.

bibliography: ~/library.bib
output: 
  rticles::elsevier_article:
    fig_caption: yes

---

# Introduction

The nutritional exchange between corals and *Symbiodinium* directly underlies the capacity of corals to build coral reef ecosystems, worth trillions of US Dollars annually [@Costanza:2014ex]. However, the complex symbiotic metabolism of corals is vulnerable to disruption by numerous anthropogenic environmental perturbations [@Baker:2015kp], jeopardizing their future persistence. In order to understand and predict coral responses to complex changes in the environment, a mechanistic understanding of how multiple interacting factors drive the individual and emergent physiology of both symbiotic partners is necessary. Such a task is well suited for theoretical modeling frameworks such as Dynamic Energy Budget (DEB) theory [@Kooijman:2010vd], although the complexity of such theory makes these efforts inaccessible to many biologists [@Jager:2013bj]. In order to bridge this gap, we present here a simplified dynamic bioenergetic model for coral-*Symbiodinium* symbioses that aims to mechanistically integrate the impacts of complex environmental change on the physiological ecology of reef corals.

In reef coral symbioses, intracellular *Symbiodinium* translocate photosynthetically-fixed carbon to support coral metabolism, and utilize the animal’s metabolic waste products, including nitrogenous compounds and carbon dioxide, in return [@Muscatine:1977p4220]. Previous application of DEB theory to this syntrophic system [@Muller:2009io] demonstrated a stable symbiotic relationship and qualitatively realistic growth and biomass ratios across gradients of ambient irradiance, nutrients, and food. This model assumed that 1) *Symbiodinium* has priority access to carbon through photosynthesis, 2) the coral animal has priority access to dissolved nitrogen through contact with seawater, and 3) each partner shares with the other only what it cannot use itself. In its simplest form, this principle of sharing the surplus sufficiently describes diverse syntrophic interactions among organs and organisms (e.g., trees, duckweeds, corals), suggesting the mechanism is mathematically and evolutionarily robust (Nisbet et al., submitted).

While the work of @Muller:2009io applying DEB formalism to this system represents the most significant theoretical contribution in coral symbiosis research to date, we aim to strengthen the role of theory and broaden its potential application in three primary ways:

1. _Develop a detailed module of environmental stress._ Of primary interest to coral biologists and ecologists is symbiosis dysfunction under environmental stress, resulting in coral “bleaching”--the loss of algal symbionts from the association [@Jokiel:1977p7353]. Photooxidative stress in *Symbiodinium* is considered a primary trigger of bleaching in response to high temperature and/or light [@Weis:2008p944], and prolonged or severe bleaching can result in mortality, though corals sometimes recover their symbionts. Bleaching susceptibility, severity, and recovery may by influenced by interacting factors such as heterotrophy and nutrient availability [@Wooldridge:2014hc], and the genetic identity of *Symbiodinium* [@Glynn:2001p7571]. To simulate these bleaching-related phenomena, we develop a generalized framework linking overreduction of the photosynthetic light reactions to downstream impacts of photoinhibition and photodamage.

2. _Reduce theoretical and mathematical complexity._ Following the logic of @Jager:2013bj, we exclude certain features of formal DEB models in order to capture behaviors of interest with the simplest possible formulation. Here, we present a model without reserves, maturity, or reproduction [see @Kooijman:2010vd]. This formulation restricts the model's scope to the bioenergetics of growth and symbiosis dynamics in adult corals, but greatly reduces theoretical complexity and parameter numbers, which is advantageous given the relative paucity of data for corals. However, our primary motivation for reducing complexity was to increase accessibility and applicability for biologists and ecologists without requiring significant expertise in DEB theory.

3. _Provide well-documented, open-access code._ In order to facilitate the continued development and application of theoretical modeling tools for coral symbioses, we provide open access to the model in the form of detailed, commented code written in the R language [@RALanguageandEn:2014wf]. With an accessible and modular framework, we envision this as a resource for futher development by the scientific community to include additional complexity and problem-specific components. The R language was chosen because it is freely available and in common use by biologists and ecologists, to widen the audience for this work.

With these as our primary motivations, we describe a simplified approach to dynamic bioenergetic modeling of coral-algal symbioses that tracks carbon and nitrogen acquisition and sharing between partners. This theoretical framework dynamically integrates the influence of external irradiance, nutrients, and prey availability on coral growth and symbiosis dynamics (i.e., symbiont:host biomass ratios), allowing for the possibility of coral bleaching in the event of photooxidative stress. In the following sections, we describe the formulation of this model and justify its structure and parameter values based on relevant literature. We then demonstrate the model's behavior and discuss some of its major implications and outcomes, and the wide range of potential applications for this model in the study of cnidarian-algal symbioses.

# Model description
In this model of coral-algal symbiosis, carbon and nitrogen are acquired by each partner and used to construct biomass. A graphical representation of the model is presented in Fig. 1. We use C-moles as the unit of biomass for consistency with the rigorous mass balance of DEB theory: 1 C-mole is equivalent to the amount of biomass containing 1 mole of Carbon atoms. Coral biomass, symbiont biomass, and prey biomass have fixed, but different, molar N:C ratios (Table 1). Carbon and nitrogen are combined to produce biomass by synthesizing units (SU), which are mathematical specifications of the formation of a product from two substrates; we use the "parallel complementary" formulation of @Kooijman:2010vd to specify these fluxes. The two state variables in this dynamical system are symbiont biomass and coral biomass; because resources are acquired proportionally to surface area (and surface area is assumed proportional to volume for corals (i.e., they are "V1-morphs" in DEB terminology [@Kooijman:2010vd])), biomass increases exponentially during growth. The rate of increase in coral biomass (i.e. growth) and the ratio of symbiont to host biomass are the (i.e. symbiosis dynamics) are the responses of interest of the system. Below we describe the formulation of each flux involved in producing these responses.

```{r Fig1, fig.width=5, fig.height=5, echo=FALSE, fig.cap="Graphical representation of coral-algal symbiosis model."}
library(png)
library(grid)
img <- readPNG("../img/model.png")
grid.raster(img)
```

## Coral animal fluxes

The coral animal acquires both carbon and nitrogen from feeding on prey from the environment. Prey acquisition is specified by Michaelis-Menten kinetics using a maximum area-specific feeding rate and half-saturation constant:

\begin{equation} U_X = {{UX_m \cdot X} \over {X + X_{KX}}} \end{equation}

Additionally, the coral animal can acquire nitrogen from the surrounding seawater. This nitrogen source is assumed to represent ammonium, the primary form utilized by corals [@Wang:1998p128; @Yellowlees:2008p331]. The uptake of nitrogen from the environment is specified by Michaelis-Menten kinetics using a maximum area-specific uptake rate and half-saturation constant:

\begin{equation} U_N = {{UN_m \cdot N} \over {N + X_{KN}}} \end{equation}

Coral biomass formation is specified by a parallel complementary SU that combines carbon and nitrogen to form biomass. In addition to the carbon and nitrogen acquired by direct uptake (Eqs. 1 and 2), the coral recycles a portion of the nitrogen liberated by biomass turnover ($r_{NR}$), and receives surplus fixed carbon shared by the symbiont ($\rho_C$), such that the total biomass formation is specified as:

\begin{equation} Q_R = \bigg({1 \over Q_{Rm}} + {1 \over {\rho_C + U_X}} + {1 \over {(U_N + n_{NX}U_X + r_{NR})n_{NR}^{-1}}} - {1 \over {{\rho_C + U_X} + (U_N + n_{NX}U_X + r_{NR})n_{NR}^{-1}}}\bigg)^{-1} \end{equation}

where $Q_{Rm}$ is the maximum specific growth rate. Finally, a fixed portion of coral biomass is turned over to pay maintenance costs:

\begin{equation} T_R = \gamma_R \end{equation}

\begin{equation} {dR \over dt} = Q_R - \gamma_R \end{equation}

where $\gamma_R$ is the parameter specifying the turnover rate.

The amount of nitrogen input to the coral biomass SU in excess of what is actually consumed in biomass formation (i.e., surplus nitrogen, or the "rejection flux" in SU terminology) is then made available to the symbiont, and is specified as:

\begin{equation} \rho_N = U_N + n_{NX}U_X + r_{NR} - n_{NR}Q_R \end{equation}

Due to the inherent "leakiness" of the parallel complementary SU formulation, there will always be some nitrogen shared with the symbiont even when coral biomass formation is strongly nitrogen-limited. Likewise, there is always a non-zero rejection flux of carbon from the coral biomass SU, which is assumed to be lost to the environment.

## *Symbiodinium* fluxes

The symbiont produces fixed carbon through photosynthesis, a process represented here by a single SU with two substrates: light (photons) and inorganic carbon (CO~2~). The amount of light absorbed by the symbiont depends on the scalar irradiance at the site of light absorption, which is modified substantially relative to external downwelling irradiance owing to multiple scattering by the coral skeleton and self-shading by surrounding symbionts [@Enriquez:2005p142; Wangpraseurt et al., in prep.]. The ratio of internal scalar irradiance to external downwelling irradiance can take a theoretical maximum value of 3 for a "flat coral model" [@Marcelino:2013hz], and may be reduced below 1 at high symbiont densities due to self-shading. We used data from Wangpraseurt et al. (in prep) to empirically derive this ratio as a function of symbiont density (expressed as symbiont to host biomass ratio), and subsequently multiply this quantity by the external downwelling irradiance $L$ and the effective light-absorbing surface area of symbiont biomass $ds$ to specify the amount of light absorbed:

\begin{equation} U_L = 3 \cdot \exp(-4.905 \cdot S/R) \cdot L \cdot ds \end{equation}

We then specify two pathways for input of inorganic carbon to the photosynthesis SU: 1) passive diffusion of CO~2~ from the external environment, and 2) active delivery of CO~2~ to the symbiont by the host. The passive flux ensures that some CO~2~ is always available to photosynthesis, and the active flux encompasses the potentially diverse mechanisms by which the host may enhance CO~2~ availability for the symbiont, including active transport of bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to CO~2~ to promote diffusion toward the symbiont, and acidification of the symbiosome to increase localized CO~2~ concentrations. Since the host physically separates the symbiont from the external environment, both the passive and active flux rates are proportional to host surface area. A more rigorous, mechanistic model of inorganic carbon processing would require spatially explicit internal pools accounting for pH and carbon speciation, which is beyond the current scope of this work. Instead, the specification of CO~2~ delivery rates offers the user the opportunity to compare different rates of CO~2~ delivery that may characterize different coral species [@Wooldridge:2014di]. The input of CO~2~ to the photosynthesis SU is therefore specified as:

\begin{equation} UCP = UCP_p + UCP_a \end{equation}

In addition to the inputs CO~2~ specified in Eqs. 7, additional CO~2~ representing the metabolic production of CO~2~ in the host liberated from host biomass turnover (note: recylced C from symbiont was removed so that photodamage turnover does not make tons of C available), is made available to the photosynthesis SU. fixed carbon is produced by the photosynthesis SU according to:

\begin{equation} UCS = \bigg({1 \over UCS_m} + {1 \over {U_L \cdot n_{LC}}} + {1 \over {UCP + r_{CR}}} - {1 \over {U_L \cdot n_{LC} + UCP + r_{CR}}}\bigg)^{-1} \div ROS \end{equation}

where $UCS_m$ is the maximum specific rate of photosynthesis, and $ROS$ is the photooxidative stress multiplier (see below). Dividing by $ROS$ causes the rate of photosynthesis to decline in response to photo-oxidative stress, a phenomenon known as photoinhibition.

Light energy absorbed in excess of what is used to fix carbon is specified by the SU "rejection flux", according to:

\begin{equation} e_L = U_L - UCS \cdot n_{LC} \end{equation}

This excess light energy must be quenched by alternative pathways in order to prevent photo-oxidative damage. *Symbiodinium* may utilize a variety of pathways for non-photochemical quenching [NPQ; @Roth:2014wf], which we collect in a total capacity for NPQ as a parameter of the symbiont ($NPQ$). If light energy exceeds the total capacity of both carbon fixation and NPQ, then damaging reactive oxygen species (ROS) may be produced, which may have negative downstream consequences. We specify a generalized photo-oxidative stress flux $ROS$ which takes a value of 1 in the absence of stress, and increases as the amount of excess excitation energy increases.

\begin{equation} ROS = 1 + \bigg[\bigg({{e_L/S - NPQ} \over L50}\bigg)^k\bigg]_+ \end{equation}

where $NPQ$, $L50$, and $k$ are parameters of the symbiont that determine the overall susceptibility to photo-oxidative stress. Importantly, $ROS$ as specified here is not a function of absolute light absorption, but rather the amount of excess light energy $e_L$ after accounting for carbon fixation and NPQ. A direct consequence of this formulation is that carbon-limitation of photosynthesis can lead to photo-oxidative stress, a mechanism of biological importance [@Wooldridge:2009p7807] that was not captured by previous representations of photo-oxidative stress [@Eynaud:2011tv]. Moreover, this formulation allows functional diversity among symbiont types to be explored by changing the parameters $NPQ$, $L50$, and $k$.

Carbon fixed by photosynthesis ($UCS$) is then used in conjunction with nitrogen shared by the host ($\rho_N$) to build symbiont biomass, following the SU equation:

\begin{equation} Q_S = \bigg({1 \over Q_{Sm}} + {1 \over UCS} + {1 \over {(\rho_N + r_{NS})n_{NR}^{-1}}} - {1 \over {UCS + (\rho_N + r_{NS})n_{NR}^{-1}}}\bigg)^{-1} \end{equation}

The rejection flux of carbon from this SU represents the amount of fixed carbon produced by photosynthesis in excess of what can be used to produce symbiont biomass; this surplus $\rho_C$ is translocated to the coral host:

\begin{equation} rho_C = UCS - Q_S \end{equation}

The rejection flux of nitrogen from the symbiont biomass SU is lost to the environment.

Finally, symbiont biomass is lost at a rate of

\begin{equation} T_S = \gamma_S \cdot 5 \cdot ROS \end{equation}

# Model behavior evaluation

# Discussion  -  Potential applications/utility

References {#references .unnumbered}
==========

### 2016-09-17 Rmd/JTB_ms.tex
Initial commit

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}
\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\bibliographystyle{elsarticle-harv}
\biboptions{sort&compress} % For natbib
\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%% Redefines the elsarticle footer
%\makeatletter
%\def\ps@pprintTitle{%
% \let\@oddhead\@empty
% \let\@evenhead\@empty
% \def\@oddfoot{\it \hfill\today}%
% \let\@evenfoot\@oddfoot}
%\makeatother

% A modified page layout
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={A simplified dynamic bioenergetic model for coral-Symbiodinium symbioses},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

\usepackage[nomarkers]{endfloat}

\begin{document}
\begin{frontmatter}

  \title{A simplified dynamic bioenergetic model for coral-\emph{Symbiodinium}
symbioses}
    \author[University of Hawaii]{Ross Cunning\corref{c1}}
   \ead{ross.cunning@gmail.com} 
   \cortext[c1]{Corresponding Author}
    \author[University of California]{Erik B. Muller}

    \author[University of Hawaii]{Ruth D. Gates}

    \author[University of California]{Roger M. Nisbet}

      \address[University of Hawaii]{Hawaii Institute of Marine Biology, Kaneohe, HI 96744, USA}
    \address[University of California]{Department of Ecology, Evolution, and Marine Biology, Santa Barbara, CA
93106, USA}

  \begin{abstract}
  This is the abstract.
  \end{abstract}

 \end{frontmatter}

\section{Introduction}\label{introduction}

The nutritional exchange between corals and \emph{Symbiodinium} directly
underlies the capacity of corals to build coral reef ecosystems, worth
trillions of US Dollars annually (Costanza, Groot, and Sutton 2014).
However, the complex symbiotic metabolism of corals is vulnerable to
disruption by numerous anthropogenic environmental perturbations (Baker
and Cunning 2015), jeopardizing their future persistence. In order to
understand and predict coral responses to complex changes in the
environment, a mechanistic understanding of how multiple interacting
factors drive the individual and emergent physiology of both symbiotic
partners is necessary. Such a task is well suited for theoretical
modeling frameworks such as Dynamic Energy Budget (DEB) theory (Kooijman
2010), although the complexity of such theory makes these efforts
inaccessible to many biologists (Jager, Martin, and Zimmer 2013). In
order to bridge this gap, we present here a simplified dynamic
bioenergetic model for coral-\emph{Symbiodinium} symbioses that aims to
mechanistically integrate the impacts of complex environmental change on
the physiological ecology of reef corals.

In reef coral symbioses, intracellular \emph{Symbiodinium} translocate
photosynthetically-fixed carbon to support coral metabolism, and utilize
the animal's metabolic waste products, including nitrogenous compounds
and carbon dioxide, in return (Muscatine and Porter 1977). Previous
application of DEB theory to this syntrophic system (Muller et al. 2009)
demonstrated a stable symbiotic relationship and qualitatively realistic
growth and biomass ratios across gradients of ambient irradiance,
nutrients, and food. This model assumed that 1) \emph{Symbiodinium} has
priority access to carbon through photosynthesis, 2) the coral animal
has priority access to dissolved nitrogen through contact with seawater,
and 3) each partner shares with the other only what it cannot use
itself. In its simplest form, this principle of sharing the surplus
sufficiently describes diverse syntrophic interactions among organs and
organisms (e.g., trees, duckweeds, corals), suggesting the mechanism is
mathematically and evolutionarily robust (Nisbet et al., submitted).

While the work of Muller et al. (2009) applying DEB formalism to this
system represents the most significant theoretical contribution in coral
symbiosis research to date, we aim to strengthen the role of theory and
broaden its potential application in three primary ways:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \emph{Develop a detailed module of environmental stress.} Of primary
  interest to coral biologists and ecologists is symbiosis dysfunction
  under environmental stress, resulting in coral ``bleaching''--the loss
  of algal symbionts from the association (Jokiel and Coles 1977).
  Photooxidative stress in \emph{Symbiodinium} is considered a primary
  trigger of bleaching in response to high temperature and/or light
  (Weis 2008), and prolonged or severe bleaching can result in
  mortality, though corals sometimes recover their symbionts. Bleaching
  susceptibility, severity, and recovery may by influenced by
  interacting factors such as heterotrophy and nutrient availability
  (Wooldridge 2014b), and the genetic identity of \emph{Symbiodinium}
  (Glynn et al. 2001). To simulate these bleaching-related phenomena, we
  develop a generalized framework linking overreduction of the
  photosynthetic light reactions to downstream impacts of
  photoinhibition and photodamage.
\item
  \emph{Reduce theoretical and mathematical complexity.} Following the
  logic of Jager, Martin, and Zimmer (2013), we exclude certain features
  of formal DEB models in order to capture behaviors of interest with
  the simplest possible formulation. Here, we present a model without
  reserves, maturity, or reproduction (see Kooijman 2010). This
  formulation restricts the model's scope to the bioenergetics of growth
  and symbiosis dynamics in adult corals, but greatly reduces
  theoretical complexity and parameter numbers, which is advantageous
  given the relative paucity of data for corals. However, our primary
  motivation for reducing complexity was to increase accessibility and
  applicability for biologists and ecologists without requiring
  significant expertise in DEB theory.
\item
  \emph{Provide well-documented, open-access code.} In order to
  facilitate the continued development and application of theoretical
  modeling tools for coral symbioses, we provide open access to the
  model in the form of detailed, commented code written in the R
  language (R Core Team 2014). With an accessible and modular framework,
  we envision this as a resource for futher development by the
  scientific community to include additional complexity and
  problem-specific components. The R language was chosen because it is
  freely available and in common use by biologists and ecologists, to
  widen the audience for this work.
\end{enumerate}

With these as our primary motivations, we describe a simplified approach
to dynamic bioenergetic modeling of coral-algal symbioses that tracks
carbon and nitrogen acquisition and sharing between partners. This
theoretical framework dynamically integrates the influence of external
irradiance, nutrients, and prey availability on coral growth and
symbiosis dynamics (i.e., symbiont:host biomass ratios), allowing for
the possibility of coral bleaching in the event of photooxidative
stress. In the following sections, we describe the formulation of this
model and justify its structure and parameter values based on relevant
literature. We then demonstrate the model's behavior and discuss some of
its major implications and outcomes, and the wide range of potential
applications for this model in the study of cnidarian-algal symbioses.

\section{Model description}\label{model-description}

In this model of coral-algal symbiosis, carbon and nitrogen are acquired
by each partner and used to construct biomass. A graphical
representation of the model is presented in Fig. 1. We use C-moles as
the unit of biomass for consistency with the rigorous mass balance of
DEB theory: 1 C-mole is equivalent to the amount of biomass containing 1
mole of Carbon atoms. Coral biomass, symbiont biomass, and prey biomass
have fixed, but different, molar N:C ratios (Table 1). Carbon and
nitrogen are combined to produce biomass by synthesizing units (SU),
which are mathematical specifications of the formation of a product from
two substrates; we use the ``parallel complementary'' formulation of
Kooijman (2010) to specify these fluxes. The two state variables in this
dynamical system are symbiont biomass and coral biomass; because
resources are acquired proportionally to surface area (and surface area
is assumed proportional to volume for corals (i.e., they are
``V1-morphs'' in DEB terminology (Kooijman 2010))), biomass increases
exponentially during growth. The rate of increase in coral biomass
(i.e.~growth) and the ratio of symbiont to host biomass are the
(i.e.~symbiosis dynamics) are the responses of interest of the system.
Below we describe the formulation of each flux involved in producing
these responses.

\begin{figure}[htbp]
\centering
\includegraphics{JTB_ms_files/figure-latex/Fig1-1.pdf}
\caption{Graphical representation of coral-algal symbiosis model.}
\end{figure}

\subsection{Coral animal fluxes}\label{coral-animal-fluxes}

The coral animal acquires both carbon and nitrogen from feeding on prey
from the environment. Prey acquisition is specified by Michaelis-Menten
kinetics using a maximum area-specific feeding rate and half-saturation
constant:

\begin{equation} U_X = {{UX_m \cdot X} \over {X + X_{KX}}} \end{equation}

Additionally, the coral animal can acquire nitrogen from the surrounding
seawater. This nitrogen source is assumed to represent ammonium, the
primary form utilized by corals (Wang and Douglas 1998; Yellowlees,
Rees, and Leggat 2008). The uptake of nitrogen from the environment is
specified by Michaelis-Menten kinetics using a maximum area-specific
uptake rate and half-saturation constant:

\begin{equation} U_N = {{UN_m \cdot N} \over {N + X_{KN}}} \end{equation}

Coral biomass formation is specified by a parallel complementary SU that
combines carbon and nitrogen to form biomass. In addition to the carbon
and nitrogen acquired by direct uptake (Eqs. 1 and 2), the coral
recycles a portion of the nitrogen liberated by biomass turnover
(\(r_{NR}\)), and receives surplus fixed carbon shared by the symbiont
(\(\rho_C\)), such that the total biomass formation is specified as:

\begin{equation} Q_R = \bigg({1 \over Q_{Rm}} + {1 \over {\rho_C + U_X}} + {1 \over {(U_N + n_{NX}U_X + r_{NR})n_{NR}^{-1}}} - {1 \over {{\rho_C + U_X} + (U_N + n_{NX}U_X + r_{NR})n_{NR}^{-1}}}\bigg)^{-1} \end{equation}

where \(Q_{Rm}\) is the maximum specific growth rate. Finally, a fixed
portion of coral biomass is turned over to pay maintenance costs:

\begin{equation} T_R = \gamma_R \end{equation}

\begin{equation} {dR \over dt} = Q_R - \gamma_R \end{equation}

where \(\gamma_R\) is the parameter specifying the turnover rate.

The amount of nitrogen input to the coral biomass SU in excess of what
is actually consumed in biomass formation (i.e., surplus nitrogen, or
the ``rejection flux'' in SU terminology) is then made available to the
symbiont, and is specified as:

\begin{equation} \rho_N = U_N + n_{NX}U_X + r_{NR} - n_{NR}Q_R \end{equation}

Due to the inherent ``leakiness'' of the parallel complementary SU
formulation, there will always be some nitrogen shared with the symbiont
even when coral biomass formation is strongly nitrogen-limited.
Likewise, there is always a non-zero rejection flux of carbon from the
coral biomass SU, which is assumed to be lost to the environment.

\subsection{\texorpdfstring{\emph{Symbiodinium}
fluxes}{Symbiodinium fluxes}}\label{symbiodinium-fluxes}

The symbiont produces fixed carbon through photosynthesis, a process
represented here by a single SU with two substrates: light (photons) and
inorganic carbon (CO\textsubscript{2}). The amount of light absorbed by
the symbiont depends on the scalar irradiance at the site of light
absorption, which is modified substantially relative to external
downwelling irradiance owing to multiple scattering by the coral
skeleton and self-shading by surrounding symbionts {[}Enríquez, Méndez,
and Iglesias-Prieto (2005); Wangpraseurt et al., in prep.{]}. The ratio
of internal scalar irradiance to external downwelling irradiance can
take a theoretical maximum value of 3 for a ``flat coral model''
(Marcelino et al. 2013), and may be reduced below 1 at high symbiont
densities due to self-shading. We used data from Wangpraseurt et al. (in
prep) to empirically derive this ratio as a function of symbiont density
(expressed as symbiont to host biomass ratio), and subsequently multiply
this quantity by the external downwelling irradiance \(L\) and the
effective light-absorbing surface area of symbiont biomass \(ds\) to
specify the amount of light absorbed:

\begin{equation} U_L = 3 \cdot \exp(-4.905 \cdot S/R) \cdot L \cdot ds \end{equation}

We then specify two pathways for input of inorganic carbon to the
photosynthesis SU: 1) passive diffusion of CO\textsubscript{2} from the
external environment, and 2) active delivery of CO\textsubscript{2} to
the symbiont by the host. The passive flux ensures that some
CO\textsubscript{2} is always available to photosynthesis, and the
active flux encompasses the potentially diverse mechanisms by which the
host may enhance CO\textsubscript{2} availability for the symbiont,
including active transport of bicarbonate, carbonic anhydrase-catalyzed
conversion of bicarbonate to CO\textsubscript{2} to promote diffusion
toward the symbiont, and acidification of the symbiosome to increase
localized CO\textsubscript{2} concentrations. Since the host physically
separates the symbiont from the external environment, both the passive
and active flux rates are proportional to host surface area. A more
rigorous, mechanistic model of inorganic carbon processing would require
spatially explicit internal pools accounting for pH and carbon
speciation, which is beyond the current scope of this work. Instead, the
specification of CO\textsubscript{2} delivery rates offers the user the
opportunity to compare different rates of CO\textsubscript{2} delivery
that may characterize different coral species (Wooldridge 2014a). The
input of CO\textsubscript{2} to the photosynthesis SU is therefore
specified as:

\begin{equation} UCP = UCP_p + UCP_a \end{equation}

In addition to the inputs CO\textsubscript{2} specified in Eqs. 7,
additional CO\textsubscript{2} representing the metabolic production of
CO\textsubscript{2} in the host liberated from host biomass turnover
(note: recylced C from symbiont was removed so that photodamage turnover
does not make tons of C available), is made available to the
photosynthesis SU. fixed carbon is produced by the photosynthesis SU
according to:

\begin{equation} UCS = \bigg({1 \over UCS_m} + {1 \over {U_L \cdot n_{LC}}} + {1 \over {UCP + r_{CR}}} - {1 \over {U_L \cdot n_{LC} + UCP + r_{CR}}}\bigg)^{-1} \div ROS \end{equation}

where \(UCS_m\) is the maximum specific rate of photosynthesis, and
\(ROS\) is the photooxidative stress multiplier (see below). Dividing by
\(ROS\) causes the rate of photosynthesis to decline in response to
photo-oxidative stress, a phenomenon known as photoinhibition.

Light energy absorbed in excess of what is used to fix carbon is
specified by the SU ``rejection flux'', according to:

\begin{equation} e_L = U_L - UCS \cdot n_{LC} \end{equation}

This excess light energy must be quenched by alternative pathways in
order to prevent photo-oxidative damage. \emph{Symbiodinium} may utilize
a variety of pathways for non-photochemical quenching (NPQ; Roth 2014),
which we collect in a total capacity for NPQ as a parameter of the
symbiont (\(NPQ\)). If light energy exceeds the total capacity of both
carbon fixation and NPQ, then damaging reactive oxygen species (ROS) may
be produced, which may have negative downstream consequences. We specify
a generalized photo-oxidative stress flux \(ROS\) which takes a value of
1 in the absence of stress, and increases as the amount of excess
excitation energy increases.

\begin{equation} ROS = 1 + \bigg[\bigg({{e_L/S - NPQ} \over L50}\bigg)^k\bigg]_+ \end{equation}

where \(NPQ\), \(L50\), and \(k\) are parameters of the symbiont that
determine the overall susceptibility to photo-oxidative stress.
Importantly, \(ROS\) as specified here is not a function of absolute
light absorption, but rather the amount of excess light energy \(e_L\)
after accounting for carbon fixation and NPQ. A direct consequence of
this formulation is that carbon-limitation of photosynthesis can lead to
photo-oxidative stress, a mechanism of biological importance (Wooldridge
2009) that was not captured by previous representations of
photo-oxidative stress (Eynaud, Nisbet, and Muller 2011). Moreover, this
formulation allows functional diversity among symbiont types to be
explored by changing the parameters \(NPQ\), \(L50\), and \(k\).

Carbon fixed by photosynthesis (\(UCS\)) is then used in conjunction
with nitrogen shared by the host (\(\rho_N\)) to build symbiont biomass,
following the SU equation:

\begin{equation} Q_S = \bigg({1 \over Q_{Sm}} + {1 \over UCS} + {1 \over {(\rho_N + r_{NS})n_{NR}^{-1}}} - {1 \over {UCS + (\rho_N + r_{NS})n_{NR}^{-1}}}\bigg)^{-1} \end{equation}

The rejection flux of carbon from this SU represents the amount of fixed
carbon produced by photosynthesis in excess of what can be used to
produce symbiont biomass; this surplus \(\rho_C\) is translocated to the
coral host:

\begin{equation} rho_C = UCS - Q_S \end{equation}

The rejection flux of nitrogen from the symbiont biomass SU is lost to
the environment.

Finally, symbiont biomass is lost at a rate of

\begin{equation} T_S = \gamma_S \cdot 5 \cdot ROS \end{equation}

\section{Model behavior evaluation}\label{model-behavior-evaluation}

\section{Discussion - Potential
applications/utility}\label{discussion---potential-applicationsutility}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-Baker:2015kp}{}
Baker, Andrew C, and Ross Cunning. 2015. ``Coral 'Bleaching' as a
Generalized Stress Response to Environmental Disturbance.'' In
\emph{Diseases of Coral}, 396--409. Hoboken, NJ: John Wiley \& Sons,
Inc.
doi:\href{https://doi.org/10.1002/9781118828502.ch30}{10.1002/9781118828502.ch30}.

\hypertarget{ref-Costanza:2014ex}{}
Costanza, R, R de Groot, and P Sutton. 2014. ``Changes in the global
value of ecosystem services.'' \emph{Global Environmental \ldots{}} 26:
152--58.
doi:\href{https://doi.org/10.1016/j.gloenvcha.2014.04.002}{10.1016/j.gloenvcha.2014.04.002}.

\hypertarget{ref-Enriquez:2005p142}{}
Enríquez, Susana, Eugenio R Méndez, and Roberto Iglesias-Prieto. 2005.
``Multiple scattering on coral skeletons enhances light absorption by
symbiotic algae.'' \emph{Limnology and Oceanography} 50 (4): 1025--32.

\hypertarget{ref-Eynaud:2011tv}{}
Eynaud, Yoan, Roger M Nisbet, and Erik B Muller. 2011. ``Impact of
excess and harmful radiation on energy budgets in scleractinian
corals.'' \emph{Ecological Modelling} 222 (7). Elsevier: 1315--22.
\url{http://www.sciencedirect.com/science/article/pii/S0304380011000263}.

\hypertarget{ref-Glynn:2001p7571}{}
Glynn, Peter W, Juan L Maté, Andrew C Baker, and MO Calderón. 2001.
``Coral bleaching and mortality in Panama and Ecuador during the
1997-1998 El Niño-Southern Oscillation event: Spatial/temporal patterns
and comparisons with the 1982-1983 event.'' \emph{Bulletin of Marine
Science} 69 (1): 79--109.
\url{http://www.scopus.com/record/display.url?fedsrfIntegrator=MEKPAPERS-SCOCIT\&origin=fedsrf\&view=basic\&eid=2-s2.0-0034761826}.

\hypertarget{ref-Jager:2013bj}{}
Jager, Tjalling, Benjamin T Martin, and Elke I Zimmer. 2013. ``DEBkiss
or the quest for the simplest generic model of animal life history.''
\emph{Journal of Theoretical Biology} 328: 9--18.
doi:\href{https://doi.org/10.1016/j.jtbi.2013.03.011}{10.1016/j.jtbi.2013.03.011}.

\hypertarget{ref-Jokiel:1977p7353}{}
Jokiel, Paul L, and SL Coles. 1977. ``Effects of temperature on the
mortality and growth of Hawaiian reef corals.'' \emph{Marine Biology} 43
(3): 201--8.
\url{http://www.scopus.com/record/display.url?fedsrfIntegrator=MEKPAPERS-SCOCIT\&origin=fedsrf\&view=basic\&eid=2-s2.0-0000237488}.

\hypertarget{ref-Kooijman:2010vd}{}
Kooijman, SALM. 2010. \emph{Dynamic Energy Budget Theory for Metabolic
Organization}. 3rd ed. Cambridge University Press.

\hypertarget{ref-Marcelino:2013hz}{}
Marcelino, Luisa A, Mark W Westneat, Valentina Stoyneva, Jillian Henss,
Jeremy D Rogers, Andrew Radosevich, Vladimir Turzhitsky, et al. 2013.
``Modulation of Light-Enhancement to Symbiotic Algae by Light-Scattering
in Corals and Evolutionary Trends in Bleaching.'' \emph{PLoS ONE} 8 (4):
e61492.
doi:\href{https://doi.org/10.1371/journal.pone.0061492.s008}{10.1371/journal.pone.0061492.s008}.

\hypertarget{ref-Muller:2009io}{}
Muller, Erik B, Sebastiaan A L M Kooijman, Peter J Edmunds, Francis J
Doyle, and Roger M Nisbet. 2009. ``Dynamic energy budgets in syntrophic
symbiotic relationships between heterotrophic hosts and photoautotrophic
symbionts.'' \emph{Journal of Theoretical Biology} 259 (1): 44--57.
doi:\href{https://doi.org/10.1016/j.jtbi.2009.03.004}{10.1016/j.jtbi.2009.03.004}.

\hypertarget{ref-Muscatine:1977p4220}{}
Muscatine, Leonard, and James W Porter. 1977. ``Reef corals: mutualistic
symbioses adapted to nutrient-poor environments.'' \emph{Bioscience} 27
(7): 454--60.

\hypertarget{ref-RALanguageandEn:2014wf}{}
R Core Team. 2014. ``R: A Language and Environment for Statistical
Computing.'' Vienna, Austria: R Foundation for Statistical Computing.
\url{http://www.R-project.org/}.

\hypertarget{ref-Roth:2014wf}{}
Roth, M S. 2014. ``The engine of the reef: Photobiology of the
coral-algal symbiosis.'' \emph{Frontiers in Microbiology}.
\url{http://journal.frontiersin.org/Journal/10.3389/fmicb.2014.00422/pdf}.

\hypertarget{ref-Wang:1998p128}{}
Wang, J, and Angela E Douglas. 1998. ``Nitrogen recycling or nitrogen
conservation in an alga-invertebrate symbiosis?'' \emph{The Journal of
Experimental Biology} 201: 2445--53.

\hypertarget{ref-Weis:2008p944}{}
Weis, Virginia M. 2008. ``Cellular mechanisms of Cnidarian bleaching:
stress causes the collapse of symbiosis.'' \emph{The Journal of
Experimental Biology} 211 (Pt 19): 3059--66.
doi:\href{https://doi.org/10.1242/jeb.009597}{10.1242/jeb.009597}.

\hypertarget{ref-Wooldridge:2009p7807}{}
Wooldridge, Scott A. 2009. ``A new conceptual model for the warm-water
breakdown of the coral-algae endosymbiosis.'' \emph{Marine and
Freshwater Research} 60 (June): 483--96.

\hypertarget{ref-Wooldridge:2014di}{}
---------. 2014a. ``Differential thermal bleaching susceptibilities
amongst coral taxa: re-posing the role of the host.'' \emph{Coral Reefs}
33 (1). Springer Berlin Heidelberg: 15--27.
doi:\href{https://doi.org/10.1007/s00338-013-1111-4}{10.1007/s00338-013-1111-4}.

\hypertarget{ref-Wooldridge:2014hc}{}
---------. 2014b. ``Formalising a mechanistic linkage between
heterotrophic feeding and thermal bleaching resistance.'' \emph{Coral
Reefs}. Springer Berlin Heidelberg, 1--6.
doi:\href{https://doi.org/10.1007/s00338-014-1193-7}{10.1007/s00338-014-1193-7}.

\hypertarget{ref-Yellowlees:2008p331}{}
Yellowlees, David, T A V Rees, and William Leggat. 2008. ``Metabolic
interactions between algal symbionts and invertebrate hosts.''
\emph{Plant, Cell and Environment} 31: 679--94.

\end{document}

### 2016-09-17 Rmd/analyze.Rmd
Initial commit

---
title: "Untitled"
author: "Ross Cunning, Erik B. Muller, Ruth D. Gates, Roger R. Nisbet"
date: "September 15, 2016"
output: html_document
---
```{r setup, include=FALSE}

# Initialize SU function
synth <- function(x, y, m) 1/((1/m)+(1/x)+(1/y)-(1/(x+y)))
# Source default parameters and functions
sapply(list.files(path="../R", pattern="*.R", full.names=T), source, .GlobalEnv)
```

```{r lightamp, fig.width=5, fig.height=5, eval=F}
source("conversions.R")  # import conversion factors for transforming data
# Import data from Marcelino et al. 2013
md <- read.csv("marcelino2013_fig2a.csv")
md$S.R <- md$density * cm2.m2 / CmolR.m2 * CmolS.cellS  # Convert units of symbiont density to CmolS/CmolR
md$A <- md$E + 1  # Convert "E"=excess light metric of Marcelino to "A"=amplification factor
# Plot data
plot(md$A ~ md$S.R, pch=c(21,23,22)[factor(md$spp)], bg=c("white", "black", "gray")[factor(md$spp)], 
     ylab="internal/external light", ylim=c(1,3.2),
     xlab="Symbiont density (CmolS/CmolR)",
     main="Light Amplification")
legend("topright", legend=levels(md$spp), pch=c(21,23,22), pt.bg=c("white", "black", "gray"), inset=0.02)
# Fit exponential decay model and plot fitted values
mod <- nls(A ~ x + y * exp(z * S.R), data=md, start=list(x=1, y=2, z=-3))
fit <- predict(mod, list(S.R=seq(0, 0.5, 0.01)))
lines(fit ~ seq(0, 0.5, 0.01))
# Get parameters of fit
formula(mod); coef(mod)
```

### Run 1
```{r run1}
# Set run time vector
time <- seq(1, 365, 1)  # if single model run, use time input

# Initialize environment
env <- init_env(time=time, L=c(10,40,2), N=c(1e-7,1e-7,0), X=c(0,0,0))

# Set parameters
pars <- def_pars()  # Get default parameters
#pars$rpars$gammaR <- 0.03  # Change any default parameters

# Run simulation
run <- run_coral(time=time, env=env, pars=pars)
```

##### Plot run 1 symbiont PI curve
```{r run1_PI, fig.height=1}
sym_PI(pars)
```

##### Plot run 1 output
```{r run1_plots}
plot_run(run)
```

### Run 2 - tolerant symbiont
```{r run2, echo=F}
pars2 <- pars
pars2$spars$L50 <- 80
pars2$spars$UCSm <- 1.4

run2 <- run_coral(time=time, env=env, pars=pars2)
```

##### Plot symbiont PI curve
```{r run2_PI, fig.height=2}
sym_PI(pars2)
```

##### Plot run 2 output
```{r run2_plots}
plot_run(run2)
```

##### Increase light to bleach tolerant symbiont
```{r}
env <- init_env(time=time, L=c(10,60,2), N=c(1e-7,1e-7,0), X=c(0,0,0))
run <- run_coral(time=time, env=env, pars=pars2)
plot_run(run)
```

##### Back to Run 1 with sensitive symbiont
Does carbon limitation affect bleaching susceptibility?
```{r clim}
plot_PSU(run)
pars$rpars$UCPaf <- 10
run3 <- run_coral(time=time, env=env, pars=pars)
plot_PSU(run3)
plot_sh(run)
plot_sh(run3)
```

### 2016-09-17 Rmd/parameter_notes.Rmd
Initial commit

---
output: 
  html_document:
    mathjax: default
---
```{r setup}
source("../R/conversions.R")
```

### Parameter values

#### Host parameters

##### Maximum host growth rate (QRm)

* [Jacobson et al. 2016](dx.doi.org/10.1242/jeb.136044):
    + Parameterized coral symbiosis DEB model from [Muller et al. (2009)](dx.doi.org/10.1016/j.jtbi.2009.03.004) using a resource-limitation experiment. Maximum growth occurs when reserve density _m~E~_ approaches infinity; the growth rate then approaches the reserve turnover rate _k~E~_ (reported value: 0.28) minus the specific maintenance rate _j~D~_ (reported value: 0.03).
    + Reported value: `r 0.28 - 0.03` CmolR • CmolR^-1^ • day^-1^

##### Turnover rate (gammaR)
*This is the maintenance rate of the host, i.e., the amount of biomass turned over each day to pay maintenance costs.*

* [Jacobson et al. 2016](dx.doi.org/10.1242/jeb.136044):
    + Parameterized coral symbiosis DEB model (see above). The specific maintenance rate is analagous to the turnover rate in the current model.
    + Reported value: 0.03 CmolR • CmolR^-1^ • d^-1^

##### Recycled N (sigmaNR) and recycled C (sigmaCR)
* Biomass turnover generates nitrogen and carbon waste (ammonia and carbon dioxide) that can be recycled for use in biomass formation. Both host and symbiont have the capacity to assimilate ammonia, so N-waste generated by each partner is made available to their specific biomass formation synthesizing units. Carbon dioxide can only be used in photosynthesis, so the C-waste generated by both partners is made available to the symbiont photosynthesis synthesizing unit. 
    + Default values used: 0.9

##### Host N:C (nNR)
*The general heterotroph N:C ratio used in DEB modeling is 0.2 (Muller et al. 2009).*

* [Horwitz et al. 2015](dx.doi.org/10.1038/srep08779):
    + Measured N:C ratio in *Anemonia viridis*, control *p*CO~2~ value reported below.
    + Reported value: 0.18083183 [inverse of C:N=5.53, see Results section]
* [Muller-Parker et al. 1994](http://hdl.handle.net/10125/2232):
    + Measured N:C ratio in *Pocillopora damicornis*
    + Reported value: 0.19230769 [inverse of C:N=5.2, see Abstract]

##### Maximum specific feeding rate (UXm)

* [Wijgerde et al. 2011](dx.doi.org/10.1242/jeb.058354)
    + Measured depletion of C and N from captured prey in *Galaxea fascicularis*, including both intracoelenteric and extracoelenteric feeding.
    + Reported value: 77.6 µg C • cm^-2^ • 6h^-1^ [sum of intra- and extracoelenteric feeding; Table 2]
    + Converted value: `r 77.6 / µg.g / gC.molC * cm2.m2 / CmolR.m2_Gfas * 4` CmolX • CmolR^-1^ • d^-1^

##### Half-saturation constant for feeding (XKX)

* [Purser et al. 2010](dx.doi.org/10.1016/j.jembe.2010.08.013)
    + Measured feeding rates of the coral *____* at different concentrations of *Artemia*
    + Reported value: ~345 *Artemia* • L^-1^
    + Converted value: `r 345 * gC.artemia / gC.molC * 10^6` µCmolX • L^-1^

##### Maximum host DIN uptake rate (UNm)

* [Muscatine & D'Elia 1978](dx.doi.org/10.4319/lo.1978.23.4.0725):
    + Measured ammonium uptake in *Pocillopora damicornis*
    + Reported value: 7.14 µmol NH~4~^+^-N • mg chl *a*^-1^ • h^-1^
    + Converted value: `r 7.14 * h.d * nmolchlA.cm2 * gchlA.molchlA * mg.g / nmol.mol / CmolR.m2_Spis * cm2.m2 / µmol.mol` mol NH~4~^+^-N • CmolR^-1^ • d^-1^

* [Hoegh-Guldberg & Williamson 1999](dx.doi.org/10.1007/s002270050496):
    + Measured ammonium and dissolved free amino acid uptake in *Pocillopora damicornis*
    + Reported value: 91.8 nmol N • cm^-2^ • h^-1^
    + Converted value: `r 91.8 * h.d * cm2.m2 / CmolR.m2_Spis / nmol.mol` mol N • CmolR^-1^ • d

* [Godinot et al. 2011 PLoS ONE](dx.doi.org/10.1371/journal.pone.0025024):
    + Measured nitrate and ammonium uptake rates in *Stylophora pistillata* under high nutrient concentrations
    + Reported value: 3.5 µmol N • mg chl^-1^ • d^-1^ [addition of bars in Fig. 4D,E @ 29°C]
    + Converted value: `r 3.5 * h.d * 22 / CmolR.m2_Spis * cm2.m2 / µmol.mol / 1000` mol N • CmolR^-1^ • d^-1^

* [Godinot et al. 2011 JEB](dx.doi.org/10.1242/jeb.054239):
    + Measured nitrate and ammonium uptake kinetics in *Stylophora pistillata*
    + Estimated V~max~ and K~1/2~ for each nutrient
    + Reported value (V~max~): 5.42 mol N • mg chl^-1^ • h^-1^
    + Converted value: `r 5.42 * h.d * µgchl.cm2_Spis / CmolR.m2_Spis * cm2.m2 / µmol.mol / 1000` mol N • CmolR^-1^ • d^-1^

##### Half-saturation constant for DIN uptake

* [Godinot et al. 2011 JEB](dx.doi.org/10.1242/jeb.054239):
    + Measured nitrate and ammonium uptake kinetics in *Stylophora pistillata*
    + Reported value (K~1/2~): 0.46 µmol • L^-1^

##### Host carbon concentrating for photosynthesis
*This is the amount of carbon the host takes up from the environment and makes available to symbiont photosynthesis. We assume that host delivery of DIC is the limiting step, rather than symbiont DIC uptake, because inhibiting carbonic anhydrases on the coral surface decreases symbiont photosynthesis (Tansik et al. 2015).*

* [Tansik et al. 2015](dx.doi.org/10.1007/s00338-015-1289-8):
    + Measured external carbonic anhydrase activity to estimate the amount of CO2 generated at the surface of the colony, which could then enter by diffusion and ultimately support symbiont photosynthesis.
    + Reported values: 5e-6 -- 3e-4 mol CO~2~ • cm^-2^ • h^-1^  
    + Converted values: `r 5e-6 * h.d * cm2.m2 / CmolR.m2` -- `r 3e-4 * h.d * cm2.m2 / CmolR.m2` mol CO~2~ • CmolR^-1^ • d^-1^
    + Lower value used in case downstream steps may further limit DIC available to photosynthesis.
    + In the absence of eCA, a maximum rate of passive CO~2~ influx was estimated:
        + 6.4e-8 CO~2~ • cm^-2^ • h^-1^
        + Converted value: `r 6.4e-8 * h.d * cm2.m2 / CmolR.m2` mol CO~2~ • CmolR^-1^ • d^-1^

#### Symbiont parameters

##### Maximum symbiont growth rate

* [Sugget et al. 2015](10.1111/nph.13483):
    + Measured growth rates of 18 *Symbiodinium* types in culture across 5 clades using FRR fluorometry.
    + Reported values: 0.19-0.26 day^-1^
* Change et al. 2983
    + Measured specific growth rates of *Symbiodinium* in culture based on cell counting at different growth irradiances.
    + Reported values: 0.22 -- 0.43 day^-1^ [Table 9]

##### Symbiont biomass turnover rate

* This value is set to equal the turnover rate for the host: 0.03 day^-1^

##### Recycled N and recycled C

* Biomass turnover generates nitrogen and carbon waste (ammonia and carbon dioxide) that can be recycled for use in biomass formation. Both host and symbiont have the capacity to assimilate ammonia, so N-waste generated by each partner is made available to their specific biomass formation synthesizing units. Carbon dioxide can only be used in photosynthesis, so the C-waste generated by both partners is made available to the symbiont photosynthesis synthesizing unit. 
    + Default values used: 0.9

##### Symbiont biomass N:C ratio (nNS)

*Theoretically should be close to Redfield ratio at 0.15.*

* [Horwitz et al. 2015](dx.doi.org/10.1038/srep08779):
    + Measured N:C ratio in *Symbiodinium* from *Anemonia viridis*.
    + Reported value: 0.137 [inverse of C:N in Fig. 5]
* [Muscatine et al. 1989](dx.doi.org/10.1098/rspb.1989.0025):
    + Reported value: 0.12987013 [inverse of C:N=7.7 in Table 2]

##### Maximum symbiont photosynthetic rate (UCSm)
*In theory, this is the maximum carboxylation rate of* Symbiodinium *Rubisco (RuBP).*

* [Leggat et al. 2004](https://digitalcollections.anu.edu.au/handle/1885/87271):
    + Estimated RuBP carboxylation at 29°C.
    + Reported value: 140 CO~2~ • RuBP^-1^ • s^-1^ [Max. value, p. 143]
    + Converted value: `r 140 * s.h * h.d * RuBP.CmolS` mol CO~2~ • CmolS^-1^ • d^-1^

*Other maximum photosynthesis rates from literature:*

* [Oakley et al. 2014](dx.doi.org/10.1007/s00338-014-1130-9):
    + Measured light- and DIC-saturated rates of O~2~ evolution in various types of cultured *Symbiodinium*
    + Reported values: 6.11 - 21.85 nmol O~2~ • mm^-3^ • min^-1^ [Fig. 2C; note: incorrect units reported in publication, corrected data and units obtained from C. Oakley]
    + Converted values: `r 6.11 * min.d / nmol.mol * mm3.µm3 / CmolS.µm3` - `r 21.85 * min.d / nmol.mol * mm3.µm3 / CmolS.µm3` mol O~2~ • CmolS^-1^ • d^-1^

* [Brading et al. 2013](dx.doi.org/10.1111/nph.12379):
    + Measured light- and DIC-saturated C-fixation in cultured *Symbiodinium* A13
    + Reported value: 259 mmol C • g chl *a*^-1^ • h^-1^ [from Table 4]
    + Converted value: `r 259 * gchlA.CmolS * h.d / mmol.mol` mol C • CmolS^-1^ • d^-1^

* Tansik unpublished:
    + Ofav Pmax: 9.94e-7 mol O~2~ • cm^-2^ • h^-1^
    + Converted: `r 9.94e-7 * h.d * cm2.m2 / CmolR.m2 / 0.1` mol O~2~ • CmolS^-1^ • d^-1^
    + Sym B1 Pmax: 2.69e-17 mol O2 • cell^-1^ • s^-1^
    + Converted: `r 2.69e-17 * s.h * h.d / pgC.cellS * pg.g * gC.molC` mol O~2~ • CmolS^-1^ • d^-1^
    + Past Pmax: 1.82e-6 mol O~2~ • cm^-2^ • h^-1^
    + Converted: `r 1.82e-6 * h.d * cm2.m2 / CmolR.m2 / 0.1` mol O~2~ • CmolS^-1^ • d^-1^
    + Sym A4a Pmax: 7.18e-17 mol O2 • cell^-1^ • s^-1^
    + Converted: `r 7.18e-17 * s.h * h.d / pgC.cellS * pg.g * gC.molC` mol O~2~ • CmolS^-1^ • d^-1^
    + Srad Pmax: 1.26e-6 mol O~2~ • cm^-2^ • h^-1^
    + Converted: `r 1.26e-6 * h.d * cm2.m2 / CmolR.m2 / 0.1` mol O~2~ • CmolS^-1^ • d^-1^
    + Sym B5 Pmax: 6.08e-17 mol O2 • cell^-1^ • s^-1^
    + Converted: `r 6.08e-17 * s.h * h.d / pgC.cellS * pg.g * gC.molC` mol O~2~ • CmolS^-1^ • d^-1^

##### Symbiont light absorption cross section (ds)
*The effective light-absorbing surface area of symbiont biomass.*

* [Hennige et al. 2009](dx.doi.org/10.1007/s00338-008-0444-x):
    + Measured spectra-weighted absoprtion coefficients from 400 to 700nm *ā\**~(400-700)~ of eight *Symbiodinium* types in culture at two different light levels (grand mean reported here).
    + Reported value: 0.01095 m^2^ • mg chl *a*^-1^ [from Table 6]
    + Converted value: `r 0.01095 * mg.g * gchlA.gC * gC.molC` m^2^ • CmolS^-1^

##### Quantum efficiency
*The number of O~2~ molecules evolved (or C fixed) per photon absorbed. The maximum theoretical value is 0.125.*

* [Brodersen et al. 2014](dx.doi.org/10.1098/rsif.2013.0997):
    + Present balanced light budget based on microenvironmental photobiology of *Montastrea curta*.
    + Reported value: 0.102 O~2~ photon^-1^

##### L50
*The amount of excitation energy in excess of both photochemical and non-photochemical quenching (NPQ) that leads to a doubling in ROS production relative to baseline values.* Units: mol photons • CmolS^-1^ • d^-1^. Values used here were determined by matching the "bleaching" response of the model to observed values of irradiance where this response is expected to occur (i.e., 40-60 mol photons • m^-2^ • d^-1^).

##### NPQ
*The flux of excitation energy in excess of photochemical quenching that can be quenched non-photochemically without producing ROS. This can be conceptually considered as the capacity for non-photochemical quenching (NPQ) in the symbiont.* Units: mol photons • CmolS^-1^ • d^-1^.

* [Gorbunov et al. 2001](dx.doi.org/10.4319/lo.2001.46.1.0075):
    + Measured the daily integrated amount of photons quenched by both photosynthetic electron transport and non-photochemical quenching (=thermal dissipation, or NPQ) in _Montastraea faveolata_ on a sunny day.
    + Reported NPQ value: 15.3 mol _e_ • m^-2^ • d^-1^.
    + Assume ~10% of total biomass is symbiont
    + Calculated value: `r 15.3 / CmolR.m2 / 0.1` mol photons • CmolS^-1^ • d^-1^

##### k
*The exponent determining the rate of increase in ROS production as a function of excess excitation energy.* Dimensionless. A default value of 1 is used here such that ROS production scales linearly with excess excitation energy.

### 2016-09-17 R/conversions.R

@@ -1,0 +1,29 @@
+# Conversion values used in parameter value calculations
+CmolR.m2 <- 3.8  # Edmunds et al. 2011
+cm2.m2 <- 10000
+mg.g <- 1000
+h.d <- 24
+s.h <- 3600
+min.d <- 1440
+cm2.m2 <- 10000
+RuBP.CmolS <- 2.31e-7  # Brading et al. 2013
+nmol.mol <- 1e9
+mmol.mol <- 1e3
+µmol.mol <- 1e6
+CmolS.µm3 <- 2.29144e-14  # Brading et al. 2011
+mm3.µm3 <- 1e-9
+gchlA.CmolS <- 0.111  # Brading et al. 2013
+gC.molC <- 12.011
+mg.g <- 1000
+µg.g <- 1e6
+pg.g <- 1e12
+gchlA.gC <- 0.01020408  # Brading et al. 2013
+gchlA.molchlA <- 893.5
+CmolR.m2_Spis <- 0.6  # Muscatine et al. 1989 in Edmunds et al. 2011 -- also 0.5 in Grover 2002
+nmolchlA.cm2 <- 9.66  # Edmunds et al. 2011
+µgchl.cm2_Spis <- 22  # Godinot et al. 2011 PLoS ONE
+CmolR.m2_Gfas <- 2  # Schutter et al. 2010 [based on reported 6 mg afdw cm-2 (Fig. 9) and 41.4% Carbon by weight (Results section)]
+gC.artemia <- 0.678e-6  # Wijgerde et al. 2011 JEB
+pgC.cellS <- 234.5 # Average of two clade A phylotypes in Brading et al. 2013
+#pgC.cellS <- 100 # Average of two types from Cassiopeia in Verde & McCloskey 1998
+CmolS.cellS <- pgC.cellS / pg.g / gC.molC

### 2016-09-17 R/def_pars.R

@@ -1,0 +1,42 @@
+def_pars <- function() {
+
+  # Default host parameters
+  defrpars <- list(
+    gammaR=0.03,  # Host specific biomass turnover rate (d-1)
+    nNR=0.19,  # N:C ratio in host biomass (-)
+    nNX=0.2,  # N:C ratio in prey biomass (-)
+    sigmaNR=0.9,  # Proportion of host nitrogen turnover recycled (-)
+    sigmaCR=0.9,  # Proportion of host carbon turnover recycled (-)
+    UXm=0.1292,  # Maximum specific host feeding rate (molX/C-molR/d)
+    UNm=0.048,  # Maximum specific host DIN uptake rate (molN/C-molR/d)
+    #UCPm=2,  # Maximum specific host CO2 uptake/delivery rate (molCO2/C-molS/d)
+    UCPpf=4.04e-3,  # Specific flux of CO2 into tissue due to passive reaciton-diffusion (Tansik et al. 2015)
+    UCPaf=0.32,  # Specific delivery of CO2 to photosynthesis by coral host (Tansik et al. 2015)
+    QRm=1,  # Maximum specific host growth rate (C-molR/C-molR/d)
+    XKN=0.46e-6,  # Half-saturation constant for host DIN uptake (molN/L)
+    XKX=20e-6,  # Half-saturation constant for host feeding (molX/L)
+    #XKC=4e-4,  # Half-saturation constant for CO2 uptake (molCO2/L)
+    initR=1  # Initial host biomass (C-molR)
+  )  
+  
+  # Default symbiont parameters
+  defspars <- list(
+    gammaS=0.03,  # Symbiont specific biomass turnover rate (d-1),
+    nNS=0.13,  # N:C ratio in symbiont biomass (-)
+    nLC=0.1,  # L:C ratio in fixed carbon (=quantum yield) (mol C/mol photons)
+    NPQ=40,  # no effect concentration of ROS per symbiont
+    L50=40,  # amount of (eL-NPQ) that doubles ROS production
+    k=1,  # exponent on ROS production
+    ds=1.34,  # Symbiont specific cross-sectional area (m^2/C-molS)
+    sigmaNS=0.9,  # Proportion of symbiont nitrogen turnover recylced (-)
+    sigmaCS=0.9,  # Proportion of symbiont carbon turnover recycled (-)
[…]

### 2016-09-17 R/init_coral.R

@@ -1,0 +1,42 @@
+init_coral <- function(pars) {
+  require(dplyr)
+  
+  # Host initialization
+  rpars <- pars$rpars
+  R <- data_frame(
+    time=time,
+    R=rpars$initR,
+    UX=(rpars$UXm * env$X[1] / (env$X[1] + rpars$XKX)),
+    UCPa=rpars$UCPaf,
+    UCPp=rpars$UCPpf,
+    UCPt=UCPa+UCPp,
+    UN=(rpars$UNm * env$N[1] / (env$N[1] + rpars$XKN)),
+    rhoN=UN,
+    wC=0,
+    QR=0.25,
+    TR=rpars$gammaR,
+    rNR=TR * rpars$nNR * rpars$sigmaNR,
+    rCR=TR * rpars$sigmaCR
+  )
+  R[2:nrow(R), 2:ncol(R)] <- NA
+  
+  # Symbiont initialization
+  spars <- pars$spars
+  S <- data_frame(
+    time=time,
+    S=spars$initS,
+    UL=env$L[1] * spars$ds,
+    UCS=max(0, synth(UL * spars$nLC, (R$UCPp[1] + R$UCPa[1])*R$R/S, spars$UCSm), na.rm=T),
+    eL=max(UL - UCS/spars$nLC, 0),
+    eC=R$UCPt*R$R/S - UCS,
+    QS=spars$QSm,
+    rhoC=UCS, 
+    wN=0,
+    TS=spars$gammaS,
+    rNS=TS * spars$nNS * spars$sigmaNS,
+    rCS=TS * spars$sigmaCS,
+    dp=1)
+  S[2:nrow(S), 2:ncol(S)] <- NA
+  
+  return(list(R=R, S=S))
+}

### 2016-09-17 R/init_env.R

@@ -1,0 +1,40 @@
+# Environment initialization
+
+init_env <- function(time, L, N, X) {
+  require(scales)
+  # Light
+  L <- if (L[3]==0) {
+    seq(L[1], L[2], along=time)
+  } else if (L[3]==1) {
+    seq(L[2], L[1], along=time)
+  } else if (L[3]==2) {
+    0.5 * (L[2] - L[1]) * sin(0.0172*time) + L[1] + 0.5 * (L[2] - L[1])
+  } else if (L[3]==3) {
+    f <- 0.5 * (L[2] - L[1]) * sin(0.0172*time)
+    f <- rescale(f, to=c(L[1]/0.4167, L[2]))
+    ff <- rescale(1 + (0.01 / (1 + exp(0.05*(time-250)))), to=c(0.4167,1))
+    f * ff
+  }
+  
+  # DIN
+  N <- if (N[3]==0) {
+    seq(N[1], N[2], along=time)
+  } else if (N[3]==1) {
+    seq(N[2], N[1], along=time)
+  } else {
+    0.5 * (N[2] - N[1]) * sin(0.0172*time) + N[1] + 0.5 * (N[2] - N[1])
+  }
+  # Prey
+  X <- if (X[3]==0) {
+    seq(X[1], X[2], along=time)
+  } else if (X[3]==1) {
+    seq(X[2], X[1], along=time)
+  } else {
+    0.5 * (X[2] - X[1]) * sin(0.0172*time) + X[1]*10^-6 + 0.5 * (X[2] - X[1])
+  }
+  # Set environment specifications
+  env <- list(L=L, N=N, X=X)
+  return(env)
+}
+
+

### 2016-09-17 R/plot_results.R

@@ -1,0 +1,166 @@
+# Plot environment for single model run
+plot_L <- function(run) with(run, {
+  plot(time, env$L, type="l", col="yellow", ylim=c(0,60), lwd=3, xlab="", ylab="mol/m2/d", main="Irradiance")
+})
+
+plot_DIN <- function(run) with(run, {
+  plot(time, env$N * 10^6, type="l", col="blue", ylim=c(0,4), lwd=3, xlab="", ylab="µmol/L", main="DIN")
+})
+
+plot_gr <- function(run) with(run, {
+  Sgr <- S$QS - S$TS
+  Rgr <- R$QR - R$TR
+  Rgrf <- Rgr[length(Rgr)]
+  plot(time, Rgr, type="l", ylim=c(min(0, min(c(Rgr[-(1:10)], Sgr[-(1:10)]))), max(c(Rgr[-(1:10)], Sgr[-(1:10)]))),
+       xlab="", ylab="d-1", lwd=3, cex=1, cex.lab=1, 
+       main="Specific growth rate")
+  if(any(c(Rgr[-(1:10)], Sgr[-(1:10)])<0)) abline(h=0, lwd=0.5, lty=3)
+  text(time[0.95*length(time)], Rgrf, labels=as.character(round(Rgrf, 3)), pos=3, xpd=T, cex=0.75)
+  lines(time, Sgr, col="black", lwd=1)
+  legend("topright", legend=c("Host", "Sym"), lwd=c(3,1), col="black", bty="n", y.intersp=1)
+})
+
+plot_sh <- function(run) with(run, {
+  totSH <- S$S / R$R
+  totSHf <- totSH[length(totSH)]
+  plot(time, totSH, type="l", col="black", ylim=c(0, max(totSH, na.rm=T)), ylab="C-molS/C-molR", xlab="", lwd=3,
+       main="Symbiont:host biomass")
+  text(time[0.95*length(time)], totSHf, labels=as.character(round(totSHf, 3)), pos=3, xpd=T, cex=0.75)
+})
+
+plot_Lq <- function(run) with(run, {
+  plot(NA, xlim=range(time), xlab="", ylim=c(0, max(c(S$UL, S$eL))), ylab="mol photons/C-molS/d",
+       main="Light quenching")
+  lines(time, S$UL, col="yellow", lty=3, lwd=3) # total amount absorbed
+  lines(time, S$UL - S$eL, col="yellow", lwd=1) # amt. used in photosynthesis
[…]

### 2016-09-17 R/run_coral.R

@@ -1,0 +1,98 @@
+run_coral <- function(time, env, pars) {
+  
+  require(dplyr)
+  
+  # Get initial values
+  # Host initialization
+  rpars <- pars$rpars
+  R <- data_frame(
+    time=time,
+    R=rpars$initR,
+    UX=(rpars$UXm * env$X[1] / (env$X[1] + rpars$XKX)),
+    UCPa=rpars$UCPaf,
+    UCPp=rpars$UCPpf,
+    UCPt=UCPa+UCPp,
+    UN=(rpars$UNm * env$N[1] / (env$N[1] + rpars$XKN)),
+    rhoN=UN,
+    wC=0,
+    QR=0.25,
+    TR=rpars$gammaR,
+    rNR=TR * rpars$nNR * rpars$sigmaNR,
+    rCR=TR * rpars$sigmaCR
+  )
+  R[2:nrow(R), 2:ncol(R)] <- NA
+  
+  # Symbiont initialization
+  spars <- pars$spars
+  S <- data_frame(
+    time=time,
+    S=spars$initS,
+    UL=env$L[1] * spars$ds,
+    UCS=max(0, synth(UL * spars$nLC, (R$UCPp[1] + R$UCPa[1])*R$R/S, spars$UCSm), na.rm=T),
+    eL=max(UL - UCS/spars$nLC, 0),
+    eC=R$UCPt*R$R/S - UCS,
+    QS=spars$QSm,
+    rhoC=UCS, 
+    wN=0,
+    TS=spars$gammaS,
+    rNS=TS * spars$nNS * spars$sigmaNS,
+    rCS=TS * spars$sigmaCS,
+    dp=1)
+  S[2:nrow(S), 2:ncol(S)] <- NA
+  
+  # Run simulation by updating
+  for (t in 2:length(time)) {
+    # Symbiont biomass loss
+    # Make turnover rate increase twice the dp (ROS) flux? 5 times!!!!
+    S$TS[t] <- spars$gammaS + spars$gammaS * 5 * (S$dp[t-1] - 1) #+ 5 * (S$dp[t-1] - 1) # REMOVE TIMES 5 -- should be 5 * (dp - 1)
+    # Coral biomass loss
+    R$TR[t] <- rpars$gammaR
+    # Biomass recycling
+    S$rNS[t] <- S$TS[t] * spars$nNS * spars$sigmaNS
[…]

### 2016-09-17 R/sym_PI.R

@@ -1,0 +1,23 @@
+sym_PI <- function(pars) {
+  with(pars, {
+    UL <- seq(0,200,1)
+    # Set carbon delivery rate (mol C / CmolS / d)
+    UCP <- 100  # Extremely high carbon delivery - no carbon limitation
+    # Calculate photosynthesis rate
+    UCS <- synth(UCP, UL*spars$nLC, spars$UCSm)
+    # Calculate light in excess of photosynthetic quenching
+    eL <- pmax(0, UL - UCS/spars$nLC)
+    # Calculate ROS (dp) produced due to excess light
+    dp <- pmax(1, 1 + (pmax(0, eL-spars$NPQ)/(spars$L50))^spars$k)
+    # Calculate photosynthesis rate accounting for ROS-induced photoinhibition
+    UCS <- UCS / dp
+    # Return ROS and photosynthesis rate for plotting
+    par(mfrow=c(1,2), mar=c(2,2,1,1), mgp=c(1.2,0.2,0), cex=0.7, tck=0.025, xaxs="i")
+    plot(UL, UCS, xlab="Light (mol photons/C-molS/d)", ylab="Carbon fixation (mol C/C-molS/d)",
+         main="Photosynthesis", type="l", lwd=3, col="red", ylim=c(0,3))
+    plot(UL, dp, xlab="Light (mol photons/C-molS/d)", ylab="ROS (relative to baseline)",
+         main="ROS production", type="l", lwd=3, col="orange", ylim=c(1,5))
+  })
+}
+
+  

### 2016-09-18 Rmd/JTB_ms.Rmd
update notation

@@
     fig_caption: yes
-
+header-includes:
+  - \usepackage{tabularx}
 ---
@@
 The nutritional exchange between corals and *Symbiodinium* directly underlies the capacity of corals to build coral reef ecosystems, worth trillions of US Dollars annually [@Costanza:2014ex].
-However, the complex symbiotic metabolism of corals is vulnerable to disruption by numerous anthropogenic environmental perturbations [@Baker:2015kp], jeopardizing their future persistence.
+However, the complex symbiotic metabolism of corals is vulnerable to disruption by numerous anthropogenic environmental perturbations, jeopardizing their future persistence.
 In order to understand and predict coral responses to complex changes in the environment, a mechanistic understanding of how multiple interacting factors drive the individual and emergent physiology of both symbiotic partners is necessary.
@@
 
-While the work of @Muller:2009io applying DEB formalism to this system represents the most significant theoretical contribution in coral symbiosis research to date, we aim to strengthen the role of theory and broaden its potential application in three primary ways:
+While the work of @Muller:2009io applying DEB formalism to this system represents the most significant theoretical contribution in coral symbiosis research to date, we aim to strengthen the role of theory and broaden its potential application to corals in three primary ways:
 
@@
 We use C-moles as the unit of biomass for consistency with the rigorous mass balance of DEB theory: 1 C-mole is equivalent to the amount of biomass containing 1 mole of Carbon atoms.
-Coral biomass, symbiont biomass, and prey biomass have fixed, but different, molar N:C ratios (Table 1).
+Host biomass ($H$), symbiont biomass ($S$), and prey biomass ($X$) have fixed, but different, molar N:C ratios (Table 1).
 Carbon and nitrogen are combined to produce biomass by synthesizing units (SU), which are mathematical specifications of the formation of a product from two substrates; we use the "parallel complementary" formulation of @Kooijman:2010vd to specify these fluxes.
@@
 Below we describe the formulation of each flux involved in producing these responses.
+
+Table 1
+
+| Symbol  | Description                            | Value  | Units              |
+| ------- | -------------------------------------- | ------ | ------------------ |
+| $n_{NH}$ | N:C molar ratio in host biomass | 0.19 | -- |
+| $n_{NS}$ | N:C molar ratio in symbiont biomass | 0.2 | -- |
+| $n_{NX}$ | N:C molar ratio in prey biomass | 0.13 | -- |
+| $j_{HT}^0$ | Specific turnover rate of host biomass | 0.03 | $d^{-1}$ |
+| $j_{ST}^0$ | Specific turnover rate of symbiont biomass | 0.03 | $d^{-1}$ |
+| $\sigma_{NH}$ | Proportion host nitrogen turnover recycled | 0.9 | -- |
+| $\sigma_{CH}$ | Proportion host carbon turnover recycled | 0.9 | -- |
+| $\sigma_{NS}$ | Proportion symbiont nitrogen turnover recycled | 0.9 | -- |
+| $\sigma_{CS}$ | Proportion symbiont carbon turnover recycled | 0.9 | -- |
+| $j_{Xm}$ | Maximum specific feeding rate of host | 0.1292 | $molX \cdot CmolH^{-1} \cdot d^{-1}$ |
+| $K_X$ | Half-saturation constant for prey uptake by host | 20e-6 | $molX \cdot L^{-1}$ | 
+| $j_{Nm}$ | Maximum specific DIN uptake rate by host | 0.048 | $molN \cdot CmolH^{-1} \cdot d^{-1}$ |
+| $K_N$ | Half-saturation constant for DIN uptake by host | 0.46e-6 | $molN \cdot L^{-1}$ | 
+| $j_{CO_2}^p$ | Passive CO~2~ delivery to symbiont | 4.04e-3 | $molC \cdot CmolH^{-1} \cdot d^{-1}$ |
+| $j_{CO_2}^a$ | Active CO~2~ delivery to symbiont | 0.32--18.0 | $molC \cdot CmolH^{-1} \cdot d^{-1}$ |
+| $j_{HGm}$ | Maximum specific growth rate of host | 1 | $d^{-1}$ |
+| $n_{LC}$ | Quantum yield of photosynthesis | 0.1 | $molC \cdot mol ph^{-1}$ |
+| $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | 1.34 | $m^2 \cdot CmolS^{-1}$ |
+| $j_{NPQ}$ | Non-photochemical quenching capacity of symbiont | 40 | $mol ph \cdot CmolS^{-1} \cdot d^{-1}$|
+| $k_{ROS}$ | Excess photon energy that doubles ROS prod. | 40-80 | $mol ph \cdot CmolS^{-1} \cdot d^{-1}$|
+| $k$ | Exponent on ROS production rate | 1 | -- |
+| $j_{CPm}$ | Maximum specific photosynthesis rate of symbiont | 2.8 | $molC \cdot CmolS^{-1} \cdot d^{-1}$ |
+| $j_{SGm}$ | Maximum specific growth rate of symbiont | 0.25 | $d^{-1}$ |
+
+Also include tables of fluxes and environmental inputs?
 
@@
 
-\begin{equation} U_X = {{UX_m \cdot X} \over {X + X_{KX}}} \end{equation}
+\begin{equation} j_X = {{j_{Xm} \cdot X} \over {X + K_X}} \end{equation}
 
@@
 
-\begin{equation} U_N = {{UN_m \cdot N} \over {N + X_{KN}}} \end{equation}
+\begin{equation} j_N = {{j_{Nm} \cdot N} \over {N + K_N}} \end{equation}
 
@@
 In addition to the carbon and nitrogen acquired by direct uptake (Eqs.
-1 and 2), the coral recycles a portion of the nitrogen liberated by biomass turnover ($r_{NR}$), and receives surplus fixed carbon shared by the symbiont ($\rho_C$), such that the total biomass formation is specified as:
+1 and 2), the coral recycles a portion of the nitrogen liberated by biomass turnover ($r_{NH}=\sigma_{NH}j_{HT}^0$), and receives surplus fixed carbon shared by the symbiont ($\rho_C$), such that the total biomass formation is specified as:
 
-\begin{equation} Q_R = \bigg({1 \over Q_{Rm}} + {1 \over {\rho_C + U_X}} + {1 \over {(U_N + n_{NX}U_X + r_{NR})n_{NR}^{-1}}} - {1 \over {{\rho_C + U_X} + (U_N + n_{NX}U_X + r_{NR})n_{NR}^{-1}}}\bigg)^{-1} \end{equation}
+\begin{equation} j_{HG} = \bigg({1 \over j_{HGm}} + {1 \over {\rho_C{S \over H} + j_X}} + {1 \over {(j_N + n_{NX}j_X + r_{NH}) / n_{NH}}} - {1 \over {{\rho_C{S \over H} + j_X} + (j_N + n_{NX}j_X + r_{NH}) / n_{NH}}}\bigg)^{-1} \end{equation}
 
-where $Q_{Rm}$ is the maximum specific growth rate. Finally, a fixed portion of coral biomass is turned over to pay maintenance costs:
+where $j_{HGm}$ is the maximum specific growth rate.
+Finally, the rate of change in coral biomass is calculated as the specific growth rate minum the specific biomass turnover rate (i.e., maintenance costs), according to:
 
-\begin{equation} T_R = \gamma_R \end{equation}
-
-\begin{equation} {dR \over dt} = Q_R - \gamma_R \end{equation}
-
-where $\gamma_R$ is the parameter specifying the turnover rate.
+\begin{equation} {dH \over dt} = j_{HG} - j_{HT}^0 \end{equation}
 
@@
 
-\begin{equation} \rho_N = U_N + n_{NX}U_X + r_{NR} - n_{NR}Q_R \end{equation}
+\begin{equation} \rho_N = j_N + n_{NX}j_X + r_{NH} - n_{NH}j_{HG} \end{equation}
 
-Due to the inherent "leakiness" of the parallel complementary SU formulation, there will always be some nitrogen shared with the symbiont even when coral biomass formation is strongly nitrogen-limited.
+Due to the inherent inefficiency of the parallel complementary SU formulation, there will always be some nitrogen shared with the symbiont even when coral biomass formation is strongly nitrogen-limited.
 Likewise, there is always a non-zero rejection flux of carbon from the coral biomass SU, which is assumed to be lost to the environment.
@@
 The symbiont produces fixed carbon through photosynthesis, a process represented here by a single SU with two substrates: light (photons) and inorganic carbon (CO~2~).
-The amount of light absorbed by the symbiont depends on the scalar irradiance at the site of light absorption, which is modified substantially relative to external downwelling irradiance owing to multiple scattering by the coral skeleton and self-shading by surrounding symbionts [@Enriquez:2005p142; Wangpraseurt et al., in prep.].
-The ratio of internal scalar irradiance to external downwelling irradiance can take a theoretical maximum value of 3 for a "flat coral model" [@Marcelino:2013hz], and may be reduced below 1 at high symbiont densities due to self-shading.
-We used data from Wangpraseurt et al.
-(in prep) to empirically derive this ratio as a function of symbiont density (expressed as symbiont to host biomass ratio), and subsequently multiply this quantity by the external downwelling irradiance $L$ and the effective light-absorbing surface area of symbiont biomass $ds$ to specify the amount of light absorbed:
+The amount of light absorbed by the symbiont depends on the scalar irradiance at the site of light absorption, which is modified substantially relative to external downwelling irradiance owing to multiple scattering by the coral skeleton and self-shading by surrounding symbionts [@Enriquez:2005p142; @Marcelino:2013hz].
+We used data from @Marcelino:2013hz to empirically derive the ratio of internal scalar irradiance to external downwelling irradiance as a function of symbiont density (expressed as symbiont to host biomass ratio), and subsequently multiply this quantity by the external downwelling irradiance $L$ and the effective light-absorbing surface area of symbiont biomass $ds$ to specify the amount of light absorbed:
 
-\begin{equation} U_L = 3 \cdot \exp(-4.905 \cdot S/R) \cdot L \cdot ds \end{equation}
+\begin{equation} j_L = \big[1.26 + 1.39 \cdot \exp(-6.48 \cdot {S \over H})\big] \cdot L \cdot \bar{a}^* \end{equation}
 
@@
 
-\begin{equation} UCP = UCP_p + UCP_a \end{equation}
+\begin{equation} j_{CO_2} = j_{CO_2}^p + j_{CO_2}^a \end{equation}
 
 In addition to the inputs CO~2~ specified in Eqs.
-7, additional CO~2~ representing the metabolic production of CO~2~ in the host liberated from host biomass turnover (note: recylced C from symbiont was removed so that photodamage turnover does not make tons of C available), is made available to the photosynthesis SU.
+7, additional CO~2~ representing the metabolic production of CO~2~ from host and symbiont biomass turnover ($r_{CH}=\sigma_{CH}j_{HT}^0; r_{CS}=\sigma_{CS}j_{ST}^0$) is made available to the photosynthesis SU.
 fixed carbon is produced by the photosynthesis SU according to:
 
-\begin{equation} UCS = \bigg({1 \over UCS_m} + {1 \over {U_L \cdot n_{LC}}} + {1 \over {UCP + r_{CR}}} - {1 \over {U_L \cdot n_{LC} + UCP + r_{CR}}}\bigg)^{-1} \div ROS \end{equation}
+\begin{equation} j_{CP} = \bigg({1 \over j_{CPm}} + {1 \over {n_{LC} j_L }} + {1 \over {(j_{CO_2} + r_{CH}){H \over S} + r_{CS}}} - {1 \over {n_{LC} j_L + (j_{CO_2} + r_{CH}){H \over S} + r_{CS}}}\bigg)^{-1} \div c_{ROS} \end{equation}
 
-where $UCS_m$ is the maximum specific rate of photosynthesis, and $ROS$ is the photooxidative stress multiplier (see below).
-Dividing by $ROS$ causes the rate of photosynthesis to decline in response to photo-oxidative stress, a phenomenon known as photoinhibition.
+where $j_{CPm}$ is the maximum specific rate of photosynthesis, and $c_{ROS}$ is the photooxidative stress multiplier (see below).
+Dividing by $c_{ROS}$ causes the rate of photosynthesis to decline in response to photo-oxidative stress, a phenomenon known as photoinhibition.
 
@@
 
-\begin{equation} e_L = U_L - UCS \cdot n_{LC} \end{equation}
+\begin{equation} j_{eL} = j_L - j_{CP} / n_{LC} \end{equation}
 
 This excess light energy must be quenched by alternative pathways in order to prevent photo-oxidative damage.
-*Symbiodinium* may utilize a variety of pathways for non-photochemical quenching [NPQ; @Roth:2014wf], which we collect in a total capacity for NPQ as a parameter of the symbiont ($NPQ$).
-If light energy exceeds the total capacity of both carbon fixation and NPQ, then damaging reactive oxygen species (ROS) may be produced, which may have negative downstream consequences.
-We specify a generalized photo-oxidative stress flux $ROS$ which takes a value of 1 in the absence of stress, and increases as the amount of excess excitation energy increases.
+*Symbiodinium* may utilize a variety of pathways for non-photochemical quenching [NPQ; @Roth:2014wf], which we collect in a total capacity for NPQ as a parameter of the symbiont ($j_{NPQ}$).
+If light energy exceeds the total capacity of both carbon fixation and NPQ, then damaging reactive oxygen species (ROS) may be produced.
+We represent this as a scaled flux of ROS $c_{ROS}$, which takes a value of 1 when all light absorbed is quenched by photochemistry and NPQ, and increases as the amount of excess excitation energy increases.
 
-\begin{equation} ROS = 1 + \bigg[\bigg({{e_L/S - NPQ} \over L50}\bigg)^k\bigg]_+ \end{equation}
+\begin{equation} c_{ROS} = 1 + \bigg[\bigg({{j_{eL} - j_{NPQ}} \over k_{ROS}}\bigg)^k\bigg]_+ \end{equation}
 
-where $NPQ$, $L50$, and $k$ are parameters of the symbiont that determine the overall susceptibility to photo-oxidative stress.
-Importantly, $ROS$ as specified here is not a function of absolute light absorption, but rather the amount of excess light energy $e_L$ after accounting for carbon fixation and NPQ.
+where $j_{NPQ}$, $k_{ROS}$, and $k$ are parameters of the symbiont that determine the onset and rate of ROS production.
+Importantly, $c_{ROS}$ as specified here is not a function of absolute light absorption, but rather the amount of excess light energy $j_{eL}$ after accounting for carbon fixation and NPQ.
 A direct consequence of this formulation is that carbon-limitation of photosynthesis can lead to photo-oxidative stress, a mechanism of biological importance [@Wooldridge:2009p7807] that was not captured by previous representations of photo-oxidative stress [@Eynaud:2011tv].
-Moreover, this formulation allows functional diversity among symbiont types to be explored by changing the parameters $NPQ$, $L50$, and $k$.
+Moreover, this formulation allows functional diversity among symbiont types to be explored by changing the parameters $j_{NPQ}$, $k_{ROS}$, and $k$.
 
-Carbon fixed by photosynthesis ($UCS$) is then used in conjunction with nitrogen shared by the host ($\rho_N$) to build symbiont biomass, following the SU equation:
+Carbon fixed by photosynthesis ($j_{CP}$) is then used in conjunction with nitrogen shared by the host ($\rho_N$) and a proportion of nitrogen recycled from symbiont biomass turnover ($r_{NS}=\sigma_{NS}j_{ST}^0$) to build new symbiont biomass, following the SU equation:
 
-\begin{equation} Q_S = \bigg({1 \over Q_{Sm}} + {1 \over UCS} + {1 \over {(\rho_N + r_{NS})n_{NR}^{-1}}} - {1 \over {UCS + (\rho_N + r_{NS})n_{NR}^{-1}}}\bigg)^{-1} \end{equation}
+\begin{equation} j_{SG} = \bigg({1 \over j_{SGm}} + {1 \over j_{CP}} + {1 \over {(\rho_N{H \over S} + r_{NS}) / n_{NH}}} - {1 \over {j_{CP} + (\rho_N{H \over S} + r_{NS}) / n_{NH}}}\bigg)^{-1} \end{equation}
 
@@
 
-\begin{equation} rho_C = UCS - Q_S \end{equation}
+\begin{equation} \rho_C = j_{CP} - j_{SG} \end{equation}
 
@@
 
-Finally, symbiont biomass is lost at a rate of
+Symbiont biomass turnover includes a component of constant turnover specified by the parameter $j_{ST}^0$, representing fixed maintenance costs, plus a component that scales with the magnitude of ROS production.
 
-\begin{equation} T_S = \gamma_S \cdot 5 \cdot ROS \end{equation}
+\begin{equation} j_{ST} = j_{ST}^0(1 + 5 \cdot (c_{ROS}-1)) \end{equation}
+
+This second component of symbiont biomass loss can represent both photodamage and/or symbiont expulsion (i.e., bleaching), both of which happen in response to higher levels or ROS production.
+The constant 5 is included to increase biomass loss in response to ROS.
+(Note that recycling of symbiont biomass turnover ($rNS$ and $rCS$) only occurs based on the basal maintenance related turnover (i.e., $j_{ST}^0$), and not the bleaching-related biomass loss, as this loss represents biomass being damaged or being expelled from the holobiont).
+Finally, the balance equation representing the change in symbiont biomass over time can be specified as the growth rate minus the turnover rate:
+
+\begin{equation} {dS \over dt} = j_{SG} - j_{ST} \end{equation}
+
+## Numerical analysis
+
+The differential equations specified in 7 and 14 were solved using an implicit time-stepping method because of the implicit variables of surplus sharing. Euler method?
 

### 2016-09-18 Rmd/JTB_ms.tex
update notation

@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\usepackage{longtable}
 \usepackage{graphicx}
@@
 \setcounter{secnumdepth}{0}
+\usepackage{tabularx}
 % Pandoc toggle for numbering sections (defaults to be off)
@@
 % Pandoc header
+\usepackage{tabularx}
 
@@
 However, the complex symbiotic metabolism of corals is vulnerable to
-disruption by numerous anthropogenic environmental perturbations (Baker
-and Cunning 2015), jeopardizing their future persistence. In order to
-understand and predict coral responses to complex changes in the
-environment, a mechanistic understanding of how multiple interacting
-factors drive the individual and emergent physiology of both symbiotic
-partners is necessary. Such a task is well suited for theoretical
-modeling frameworks such as Dynamic Energy Budget (DEB) theory (Kooijman
-2010), although the complexity of such theory makes these efforts
-inaccessible to many biologists (Jager, Martin, and Zimmer 2013). In
-order to bridge this gap, we present here a simplified dynamic
-bioenergetic model for coral-\emph{Symbiodinium} symbioses that aims to
-mechanistically integrate the impacts of complex environmental change on
-the physiological ecology of reef corals.
+disruption by numerous anthropogenic environmental perturbations,
+jeopardizing their future persistence. In order to understand and
+predict coral responses to complex changes in the environment, a
+mechanistic understanding of how multiple interacting factors drive the
+individual and emergent physiology of both symbiotic partners is
+necessary. Such a task is well suited for theoretical modeling
+frameworks such as Dynamic Energy Budget (DEB) theory (Kooijman 2010),
+although the complexity of such theory makes these efforts inaccessible
+to many biologists (Jager, Martin, and Zimmer 2013). In order to bridge
+this gap, we present here a simplified dynamic bioenergetic model for
+coral-\emph{Symbiodinium} symbioses that aims to mechanistically
+integrate the impacts of complex environmental change on the
+physiological ecology of reef corals.
 
@@
 symbiosis research to date, we aim to strengthen the role of theory and
-broaden its potential application in three primary ways:
+broaden its potential application to corals in three primary ways:
 
@@
 DEB theory: 1 C-mole is equivalent to the amount of biomass containing 1
-mole of Carbon atoms. Coral biomass, symbiont biomass, and prey biomass
-have fixed, but different, molar N:C ratios (Table 1). Carbon and
-nitrogen are combined to produce biomass by synthesizing units (SU),
-which are mathematical specifications of the formation of a product from
-two substrates; we use the ``parallel complementary'' formulation of
-Kooijman (2010) to specify these fluxes. The two state variables in this
-dynamical system are symbiont biomass and coral biomass; because
-resources are acquired proportionally to surface area (and surface area
-is assumed proportional to volume for corals (i.e., they are
-``V1-morphs'' in DEB terminology (Kooijman 2010))), biomass increases
-exponentially during growth. The rate of increase in coral biomass
-(i.e.~growth) and the ratio of symbiont to host biomass are the
-(i.e.~symbiosis dynamics) are the responses of interest of the system.
-Below we describe the formulation of each flux involved in producing
-these responses.
+mole of Carbon atoms. Host biomass (\(H\)), symbiont biomass (\(S\)),
+and prey biomass (\(X\)) have fixed, but different, molar N:C ratios
+(Table 1). Carbon and nitrogen are combined to produce biomass by
+synthesizing units (SU), which are mathematical specifications of the
+formation of a product from two substrates; we use the ``parallel
+complementary'' formulation of Kooijman (2010) to specify these fluxes.
+The two state variables in this dynamical system are symbiont biomass
+and coral biomass; because resources are acquired proportionally to
+surface area (and surface area is assumed proportional to volume for
+corals (i.e., they are ``V1-morphs'' in DEB terminology (Kooijman
+2010))), biomass increases exponentially during growth. The rate of
+increase in coral biomass (i.e.~growth) and the ratio of symbiont to
+host biomass are the (i.e.~symbiosis dynamics) are the responses of
+interest of the system. Below we describe the formulation of each flux
+involved in producing these responses.
+
+Table 1
+
+\begin{longtable}[c]{@{}llll@{}}
+\toprule
+\begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
+Symbol
+\strut\end{minipage} &
+\begin{minipage}[b]{0.48\columnwidth}\raggedright\strut
+Description
+\strut\end{minipage} &
+\begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
+Value
+\strut\end{minipage} &
+\begin{minipage}[b]{0.23\columnwidth}\raggedright\strut
+Units
+\strut\end{minipage}\tabularnewline
+\midrule
+\endhead
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(n_{NH}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+N:C molar ratio in host biomass
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+0.19
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+--
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(n_{NS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+N:C molar ratio in symbiont biomass
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+0.2
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+--
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(n_{NX}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+N:C molar ratio in prey biomass
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+0.13
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+--
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{HT}^0\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Specific turnover rate of host biomass
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+0.03
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\(d^{-1}\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{ST}^0\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Specific turnover rate of symbiont biomass
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+0.03
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\(d^{-1}\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(\sigma_{NH}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Proportion host nitrogen turnover recycled
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+0.9
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+--
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(\sigma_{CH}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Proportion host carbon turnover recycled
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+0.9
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+--
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(\sigma_{NS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Proportion symbiont nitrogen turnover recycled
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+0.9
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+--
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(\sigma_{CS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Proportion symbiont carbon turnover recycled
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+0.9
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+--
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{Xm}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Maximum specific feeding rate of host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+0.1292
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\(molX \cdot CmolH^{-1} \cdot d^{-1}\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(K_X\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Half-saturation constant for prey uptake by host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+20e-6
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\(molX \cdot L^{-1}\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{Nm}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Maximum specific DIN uptake rate by host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+0.048
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\(molN \cdot CmolH^{-1} \cdot d^{-1}\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(K_N\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Half-saturation constant for DIN uptake by host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+0.46e-6
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\(molN \cdot L^{-1}\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{CO_2}^p\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Passive CO\textsubscript{2} delivery to symbiont
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+4.04e-3
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\(molC \cdot CmolH^{-1} \cdot d^{-1}\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{CO_2}^a\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Active CO\textsubscript{2} delivery to symbiont
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+0.32--18.0
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\(molC \cdot CmolH^{-1} \cdot d^{-1}\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{HGm}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Maximum specific growth rate of host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+1
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\(d^{-1}\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(n_{LC}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Quantum yield of photosynthesis
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+0.1
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\(molC \cdot mol ph^{-1}\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(\bar{a}^*\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Effective light-absorbing cross-section of symbiont
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+1.34
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\(m^2 \cdot CmolS^{-1}\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{NPQ}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Non-photochemical quenching capacity of symbiont
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+40
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\(mol ph \cdot CmolS^{-1} \cdot d^{-1}\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(k_{ROS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Excess photon energy that doubles ROS prod.
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+40-80
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\(mol ph \cdot CmolS^{-1} \cdot d^{-1}\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(k\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Exponent on ROS production rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+1
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+--
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{CPm}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Maximum specific photosynthesis rate of symbiont
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+2.8
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\(molC \cdot CmolS^{-1} \cdot d^{-1}\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{SGm}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Maximum specific growth rate of symbiont
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+0.25
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\(d^{-1}\)
+\strut\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
+
+Also include tables of fluxes and environmental inputs?
 
@@
 recycles a portion of the nitrogen liberated by biomass turnover
-(\(r_{NR}\)), and receives surplus fixed carbon shared by the symbiont
-(\(\rho_C\)), such that the total biomass formation is specified as:
+(\(r_{NH}=\sigma_{NH}j_{HT}^0\)), and receives surplus fixed carbon
+shared by the symbiont (\(\rho_C\)), such that the total biomass
+formation is specified as:
 
-\begin{equation} Q_R = \bigg({1 \over Q_{Rm}} + {1 \over {\rho_C + U_X}} + {1 \over {(U_N + n_{NX}U_X + r_{NR})n_{NR}^{-1}}} - {1 \over {{\rho_C + U_X} + (U_N + n_{NX}U_X + r_{NR})n_{NR}^{-1}}}\bigg)^{-1} \end{equation}
+\begin{equation} j_{HG} = \bigg({1 \over j_{HGm}} + {1 \over {\rho_C{S \over H} + j_X}} + {1 \over {(j_N + n_{NX}j_X + r_{NH}) / n_{NH}}} - {1 \over {{\rho_C{S \over H} + j_X} + (j_N + n_{NX}j_X + r_{NH}) / n_{NH}}}\bigg)^{-1} \end{equation}
 
-where \(Q_{Rm}\) is the maximum specific growth rate. Finally, a fixed
-portion of coral biomass is turned over to pay maintenance costs:
+where \(j_{HGm}\) is the maximum specific growth rate. Finally, the rate
+of change in coral biomass is calculated as the specific growth rate
+minum the specific biomass turnover rate (i.e., maintenance costs),
+according to:
 
-\begin{equation} T_R = \gamma_R \end{equation}
-
-\begin{equation} {dR \over dt} = Q_R - \gamma_R \end{equation}
-
-where \(\gamma_R\) is the parameter specifying the turnover rate.
+\begin{equation} {dH \over dt} = j_{HG} - j_{HT}^0 \end{equation}
 
@@
 
-\begin{equation} \rho_N = U_N + n_{NX}U_X + r_{NR} - n_{NR}Q_R \end{equation}
+\begin{equation} \rho_N = j_N + n_{NX}j_X + r_{NH} - n_{NH}j_{HG} \end{equation}
 
-Due to the inherent ``leakiness'' of the parallel complementary SU
+Due to the inherent inefficiency of the parallel complementary SU
 formulation, there will always be some nitrogen shared with the symbiont
@@
 downwelling irradiance owing to multiple scattering by the coral
-skeleton and self-shading by surrounding symbionts {[}Enríquez, Méndez,
-and Iglesias-Prieto (2005); Wangpraseurt et al., in prep.{]}. The ratio
-of internal scalar irradiance to external downwelling irradiance can
-take a theoretical maximum value of 3 for a ``flat coral model''
-(Marcelino et al. 2013), and may be reduced below 1 at high symbiont
-densities due to self-shading. We used data from Wangpraseurt et al. (in
-prep) to empirically derive this ratio as a function of symbiont density
-(expressed as symbiont to host biomass ratio), and subsequently multiply
-this quantity by the external downwelling irradiance \(L\) and the
-effective light-absorbing surface area of symbiont biomass \(ds\) to
-specify the amount of light absorbed:
+skeleton and self-shading by surrounding symbionts (Enríquez, Méndez,
+and Iglesias-Prieto 2005; Marcelino et al. 2013). We used data from
+Marcelino et al. (2013) to empirically derive the ratio of internal
+scalar irradiance to external downwelling irradiance as a function of
+symbiont density (expressed as symbiont to host biomass ratio), and
+subsequently multiply this quantity by the external downwelling
+irradiance \(L\) and the effective light-absorbing surface area of
+symbiont biomass \(ds\) to specify the amount of light absorbed:
 
-\begin{equation} U_L = 3 \cdot \exp(-4.905 \cdot S/R) \cdot L \cdot ds \end{equation}
+\begin{equation} j_L = \big[1.26 + 1.39 \cdot \exp(-6.48 \cdot {S \over H})\big] \cdot L \cdot \bar{a}^* \end{equation}
 
@@
 
-\begin{equation} UCP = UCP_p + UCP_a \end{equation}
+\begin{equation} j_{CO_2} = j_{CO_2}^p + j_{CO_2}^a \end{equation}
 
@@
 additional CO\textsubscript{2} representing the metabolic production of
-CO\textsubscript{2} in the host liberated from host biomass turnover
-(note: recylced C from symbiont was removed so that photodamage turnover
-does not make tons of C available), is made available to the
-photosynthesis SU. fixed carbon is produced by the photosynthesis SU
-according to:
+CO\textsubscript{2} from host and symbiont biomass turnover
+(\(r_{CH}=\sigma_{CH}j_{HT}^0; r_{CS}=\sigma_{CS}j_{ST}^0\)) is made
+available to the photosynthesis SU. fixed carbon is produced by the
+photosynthesis SU according to:
 
-\begin{equation} UCS = \bigg({1 \over UCS_m} + {1 \over {U_L \cdot n_{LC}}} + {1 \over {UCP + r_{CR}}} - {1 \over {U_L \cdot n_{LC} + UCP + r_{CR}}}\bigg)^{-1} \div ROS \end{equation}
+\begin{equation} j_{CP} = \bigg({1 \over j_{CPm}} + {1 \over {n_{LC} j_L }} + {1 \over {(j_{CO_2} + r_{CH}){H \over S} + r_{CS}}} - {1 \over {n_{LC} j_L + (j_{CO_2} + r_{CH}){H \over S} + r_{CS}}}\bigg)^{-1} \div c_{ROS} \end{equation}
 
-where \(UCS_m\) is the maximum specific rate of photosynthesis, and
-\(ROS\) is the photooxidative stress multiplier (see below). Dividing by
-\(ROS\) causes the rate of photosynthesis to decline in response to
-photo-oxidative stress, a phenomenon known as photoinhibition.
+where \(j_{CPm}\) is the maximum specific rate of photosynthesis, and
+\(c_{ROS}\) is the photooxidative stress multiplier (see below).
+Dividing by \(c_{ROS}\) causes the rate of photosynthesis to decline in
+response to photo-oxidative stress, a phenomenon known as
+photoinhibition.
 
@@
 
-\begin{equation} e_L = U_L - UCS \cdot n_{LC} \end{equation}
+\begin{equation} j_{eL} = j_L - j_{CP} / n_{LC} \end{equation}
 
@@
 which we collect in a total capacity for NPQ as a parameter of the
-symbiont (\(NPQ\)). If light energy exceeds the total capacity of both
-carbon fixation and NPQ, then damaging reactive oxygen species (ROS) may
-be produced, which may have negative downstream consequences. We specify
-a generalized photo-oxidative stress flux \(ROS\) which takes a value of
-1 in the absence of stress, and increases as the amount of excess
-excitation energy increases.
+symbiont (\(j_{NPQ}\)). If light energy exceeds the total capacity of
+both carbon fixation and NPQ, then damaging reactive oxygen species
+(ROS) may be produced. We represent this as a scaled flux of ROS
+\(c_{ROS}\), which takes a value of 1 when all light absorbed is
+quenched by photochemistry and NPQ, and increases as the amount of
+excess excitation energy increases.
 
-\begin{equation} ROS = 1 + \bigg[\bigg({{e_L/S - NPQ} \over L50}\bigg)^k\bigg]_+ \end{equation}
+\begin{equation} c_{ROS} = 1 + \bigg[\bigg({{j_{eL} - j_{NPQ}} \over k_{ROS}}\bigg)^k\bigg]_+ \end{equation}
 
-where \(NPQ\), \(L50\), and \(k\) are parameters of the symbiont that
-determine the overall susceptibility to photo-oxidative stress.
-Importantly, \(ROS\) as specified here is not a function of absolute
-light absorption, but rather the amount of excess light energy \(e_L\)
+where \(j_{NPQ}\), \(k_{ROS}\), and \(k\) are parameters of the symbiont
+that determine the onset and rate of ROS production. Importantly,
+\(c_{ROS}\) as specified here is not a function of absolute light
+absorption, but rather the amount of excess light energy \(j_{eL}\)
 after accounting for carbon fixation and NPQ. A direct consequence of
@@
 formulation allows functional diversity among symbiont types to be
-explored by changing the parameters \(NPQ\), \(L50\), and \(k\).
+explored by changing the parameters \(j_{NPQ}\), \(k_{ROS}\), and \(k\).
 
-Carbon fixed by photosynthesis (\(UCS\)) is then used in conjunction
-with nitrogen shared by the host (\(\rho_N\)) to build symbiont biomass,
+Carbon fixed by photosynthesis (\(j_{CP}\)) is then used in conjunction
+with nitrogen shared by the host (\(\rho_N\)) and a proportion of
+nitrogen recycled from symbiont biomass turnover
+(\(r_{NS}=\sigma_{NS}j_{ST}^0\)) to build new symbiont biomass,
 following the SU equation:
 
-\begin{equation} Q_S = \bigg({1 \over Q_{Sm}} + {1 \over UCS} + {1 \over {(\rho_N + r_{NS})n_{NR}^{-1}}} - {1 \over {UCS + (\rho_N + r_{NS})n_{NR}^{-1}}}\bigg)^{-1} \end{equation}
+\begin{equation} j_{SG} = \bigg({1 \over j_{SGm}} + {1 \over j_{CP}} + {1 \over {(\rho_N{H \over S} + r_{NS}) / n_{NH}}} - {1 \over {j_{CP} + (\rho_N{H \over S} + r_{NS}) / n_{NH}}}\bigg)^{-1} \end{equation}
 
@@
 
-Finally, symbiont biomass is lost at a rate of
+Symbiont biomass turnover includes a component of constant turnover
+specified by the parameter \(j_{ST}^0\), representing fixed maintenance
+costs, plus a component that scales with the magnitude of ROS
+production.
 
-\begin{equation} T_S = \gamma_S \cdot 5 \cdot ROS \end{equation}
+\begin{equation} j_{ST} = j_{ST}^0(1 + 5 \cdot (c_{ROS}-1)) \end{equation}
+
+This second component of symbiont biomass loss can represent both
+photodamage and/or symbiont expulsion (i.e., bleaching), both of which
+happen in response to higher levels or ROS production. The constant 5 is
+included to increase biomass loss in response to ROS. (Note that
+recycling of symbiont biomass turnover (\(rNS\) and \(rCS\)) only occurs
+based on the basal maintenance related turnover (i.e., \(j_{ST}^0\)),
+and not the bleaching-related biomass loss, as this loss represents
+biomass being damaged or being expelled from the holobiont). Finally,
+the balance equation representing the change in symbiont biomass over
+time can be specified as the growth rate minus the turnover rate:
+
+\begin{equation} {dS \over dt} = j_{SG} - j_{ST} \end{equation}
 
@@
 \hypertarget{refs}{}
-\hypertarget{ref-Baker:2015kp}{}
-Baker, Andrew C, and Ross Cunning. 2015. ``Coral 'Bleaching' as a
-Generalized Stress Response to Environmental Disturbance.'' In
-\emph{Diseases of Coral}, 396--409. Hoboken, NJ: John Wiley \& Sons,
-Inc.
-doi:\href{https://doi.org/10.1002/9781118828502.ch30}{10.1002/9781118828502.ch30}.
-
 \hypertarget{ref-Costanza:2014ex}{}

### 2016-09-18 Rmd/analyze.Rmd
update notation

@@
 ```{r setup, include=FALSE}
+knitr::opts_knit$set(root.dir = normalizePath(".."))
+```
 
+```{r initialize, include=FALSE}
 # Initialize SU function
@@
 # Source default parameters and functions
-sapply(list.files(path="../R", pattern="*.R", full.names=T), source, .GlobalEnv)
+sapply(list.files(path="R", pattern="*.R", full.names=T), source, .GlobalEnv)
 ```
@@
 ```{r lightamp, fig.width=5, fig.height=5, eval=F}
-source("conversions.R")  # import conversion factors for transforming data
 # Import data from Marcelino et al. 2013
-md <- read.csv("marcelino2013_fig2a.csv")
+md <- read.csv("data/marcelino2013_fig2a.csv")
 md$S.R <- md$density * cm2.m2 / CmolR.m2 * CmolS.cellS  # Convert units of symbiont density to CmolS/CmolR
@@
 # Set run time vector
-time <- seq(1, 365, 1)  # if single model run, use time input
+time <- seq(1, 365, 0.1)  # if single model run, use time input
 
@@
 plot_run(run)
+
+```
+
+#### Run 1 with no symbiont recycling
+```{r runNR_plots}
+parsNR <- pars
+parsNR$spars$sigmaNS <- 0
+parsNR$spars$sigmaCS <- 0
+runNR <- run_coral(time=time, env=env, pars=parsNR)
+plot_run(runNR)
+with(runNR, {
+
+})
+```
+
+#### Compare amount of recycled sym nitrogen between runs
+```{r compareN}
+proprNS <- run$S$rNS / (run$S$rNS + run$H$rhoN*run$H$H/run$S$S)
+par(mfrow=c(2,1))
+plot(time, proprNS)
+proprNS <- runNR$S$rNS / (runNR$S$rNS + runNR$H$rhoN*runNR$H$H/runNR$S$S)
+plot(time, proprNS)
+
+proprCS <- run$S$rCS / (run$S$rCS + run$H$jCO2*run$H$H/run$S$S)
+par(mfrow=c(2,1))
+plot(time, proprCS)
+proprCS <- runNR$S$rCS / (runNR$S$rCS + (runNR$H$rCH + runNR$H$jCO2)*runNR$H$H/runNR$S$S)
+plot(time, proprCS)
 ```
@@
 pars2 <- pars
-pars2$spars$L50 <- 80
-pars2$spars$UCSm <- 1.4
+pars2$spars$kROS <- 80
+pars2$spars$jCPm <- 1.4
 
@@
 ```
-
-##### Back to Run 1 with sensitive symbiont
-Does carbon limitation affect bleaching susceptibility?
-```{r clim}
-plot_PSU(run)
-pars$rpars$UCPaf <- 10
-run3 <- run_coral(time=time, env=env, pars=pars)
-plot_PSU(run3)
-plot_sh(run)
-plot_sh(run3)
-```

### 2016-09-18 R/def_pars.R

@@ -1,41 +1,40 @@
 def_pars <- function() {
 
   # Default host parameters
-  defrpars <- list(
-    gammaR=0.03,  # Host specific biomass turnover rate (d-1)
-    nNR=0.19,  # N:C ratio in host biomass (-)
+  defhpars <- list(
+    jHT0=0.03,  # Host specific biomass turnover rate (d-1)
+    nNH=0.19,  # N:C ratio in host biomass (-)
     nNX=0.2,  # N:C ratio in prey biomass (-)
-    sigmaNR=0.9,  # Proportion of host nitrogen turnover recycled (-)
-    sigmaCR=0.9,  # Proportion of host carbon turnover recycled (-)
-    UXm=0.1292,  # Maximum specific host feeding rate (molX/C-molR/d)
-    UNm=0.048,  # Maximum specific host DIN uptake rate (molN/C-molR/d)
-    #UCPm=2,  # Maximum specific host CO2 uptake/delivery rate (molCO2/C-molS/d)
-    UCPpf=4.04e-3,  # Specific flux of CO2 into tissue due to passive reaciton-diffusion (Tansik et al. 2015)
-    UCPaf=0.32,  # Specific delivery of CO2 to photosynthesis by coral host (Tansik et al. 2015)
-    QRm=1,  # Maximum specific host growth rate (C-molR/C-molR/d)
-    XKN=0.46e-6,  # Half-saturation constant for host DIN uptake (molN/L)
-    XKX=20e-6,  # Half-saturation constant for host feeding (molX/L)
+    sigmaNH=0.9,  # Proportion of host nitrogen turnover recycled (-)
+    sigmaCH=0.9,  # Proportion of host carbon turnover recycled (-)
+    jXm=0.1292,  # Maximum specific host feeding rate (molX/C-molH/d)
+    jNm=0.048,  # Maximum specific host DIN uptake rate (molN/C-molH/d)
+    jCO2p=4.04e-3,  # Specific flux of CO2 into tissue due to passive reaciton-diffusion (Tansik et al. 2015)
+    jCO2a=0.32,  # Specific delivery of CO2 to photosynthesis by coral host (Tansik et al. 2015)
+    jHGm=1,  # Maximum specific host growth rate (C-molH/C-molH/d)
+    KN=0.46e-6,  # Half-saturation constant for host DIN uptake (molN/L)
+    KX=20e-6,  # Half-saturation constant for host feeding (molX/L)
[…]

### 2016-09-18 R/init_coral.R

@@ -1,42 +1,0 @@
-init_coral <- function(pars) {
-  require(dplyr)
-  
-  # Host initialization
-  rpars <- pars$rpars
-  R <- data_frame(
-    time=time,
-    R=rpars$initR,
-    UX=(rpars$UXm * env$X[1] / (env$X[1] + rpars$XKX)),
-    UCPa=rpars$UCPaf,
-    UCPp=rpars$UCPpf,
-    UCPt=UCPa+UCPp,
-    UN=(rpars$UNm * env$N[1] / (env$N[1] + rpars$XKN)),
-    rhoN=UN,
-    wC=0,
-    QR=0.25,
-    TR=rpars$gammaR,
-    rNR=TR * rpars$nNR * rpars$sigmaNR,
-    rCR=TR * rpars$sigmaCR
-  )
-  R[2:nrow(R), 2:ncol(R)] <- NA
-  
-  # Symbiont initialization
-  spars <- pars$spars
-  S <- data_frame(
-    time=time,
-    S=spars$initS,
-    UL=env$L[1] * spars$ds,
-    UCS=max(0, synth(UL * spars$nLC, (R$UCPp[1] + R$UCPa[1])*R$R/S, spars$UCSm), na.rm=T),
-    eL=max(UL - UCS/spars$nLC, 0),
-    eC=R$UCPt*R$R/S - UCS,
-    QS=spars$QSm,
-    rhoC=UCS, 
-    wN=0,
-    TS=spars$gammaS,
-    rNS=TS * spars$nNS * spars$sigmaNS,
-    rCS=TS * spars$sigmaCS,
-    dp=1)
-  S[2:nrow(S), 2:ncol(S)] <- NA
-  
-  return(list(R=R, S=S))
-}

### 2016-09-18 R/plot_results.R

@@ -8,69 +8,69 @@ plot_DIN <- function(run) with(run, {
 })
 
 plot_gr <- function(run) with(run, {
-  Sgr <- S$QS - S$TS
-  Rgr <- R$QR - R$TR
-  Rgrf <- Rgr[length(Rgr)]
-  plot(time, Rgr, type="l", ylim=c(min(0, min(c(Rgr[-(1:10)], Sgr[-(1:10)]))), max(c(Rgr[-(1:10)], Sgr[-(1:10)]))),
+  Sgr <- S$jSG - S$ST
+  Hgr <- H$jHG - H$HT
+  Hgrf <- Hgr[length(Hgr)]
+  plot(time, Hgr, type="l", ylim=c(min(0, min(c(Hgr[-(1:10)], Sgr[-(1:10)]))), max(c(Hgr[-(1:10)], Sgr[-(1:10)]))),
        xlab="", ylab="d-1", lwd=3, cex=1, cex.lab=1, 
        main="Specific growth rate")
-  if(any(c(Rgr[-(1:10)], Sgr[-(1:10)])<0)) abline(h=0, lwd=0.5, lty=3)
-  text(time[0.95*length(time)], Rgrf, labels=as.character(round(Rgrf, 3)), pos=3, xpd=T, cex=0.75)
+  if(any(c(Hgr[-(1:10)], Sgr[-(1:10)])<0)) abline(h=0, lwd=0.5, lty=3)
+  text(time[0.95*length(time)], Hgrf, labels=as.character(round(Hgrf, 3)), pos=3, xpd=T, cex=0.75)
   lines(time, Sgr, col="black", lwd=1)
   legend("topright", legend=c("Host", "Sym"), lwd=c(3,1), col="black", bty="n", y.intersp=1)
 })
 
 plot_sh <- function(run) with(run, {
-  totSH <- S$S / R$R
+  totSH <- S$S / H$H
   totSHf <- totSH[length(totSH)]
-  plot(time, totSH, type="l", col="black", ylim=c(0, max(totSH, na.rm=T)), ylab="C-molS/C-molR", xlab="", lwd=3,
+  plot(time, totSH, type="l", col="black", ylim=c(0, max(totSH, na.rm=T)), ylab="C-molS/C-molH", xlab="", lwd=3,
        main="Symbiont:host biomass")
[…]

### 2016-09-18 R/run_coral.R

@@ -1,98 +1,97 @@
 run_coral <- function(time, env, pars) {
-  
   require(dplyr)
   
-  # Get initial values
+  # Get initial values 
   # Host initialization
-  rpars <- pars$rpars
-  R <- data_frame(
+  hpars <- pars$hpars
+  H <- data_frame(
     time=time,
-    R=rpars$initR,
-    UX=(rpars$UXm * env$X[1] / (env$X[1] + rpars$XKX)),
-    UCPa=rpars$UCPaf,
-    UCPp=rpars$UCPpf,
-    UCPt=UCPa+UCPp,
-    UN=(rpars$UNm * env$N[1] / (env$N[1] + rpars$XKN)),
-    rhoN=UN,
+    H=hpars$initH,
+    jX=(hpars$jXm * env$X[1] / (env$X[1] + hpars$KX)),
+    jCO2a=hpars$jCO2a,
+    jCO2p=hpars$jCO2p,
+    jCO2=jCO2a+jCO2p,
+    jN=(hpars$jNm * env$N[1] / (env$N[1] + hpars$KN)),
+    rhoN=jN,
     wC=0,
-    QR=0.25,
-    TR=rpars$gammaR,
-    rNR=TR * rpars$nNR * rpars$sigmaNR,
-    rCR=TR * rpars$sigmaCR
+    jHG=0.25,
+    HT=hpars$jHT0,
+    rNH=HT * hpars$nNH * hpars$sigmaNH,
+    rCH=HT * hpars$sigmaCH
   )
-  R[2:nrow(R), 2:ncol(R)] <- NA
+  H[2:nrow(H), 2:ncol(H)] <- NA
   
   # Symbiont initialization
   spars <- pars$spars
   S <- data_frame(
     time=time,
     S=spars$initS,
-    UL=env$L[1] * spars$ds,
-    UCS=max(0, synth(UL * spars$nLC, (R$UCPp[1] + R$UCPa[1])*R$R/S, spars$UCSm), na.rm=T),
-    eL=max(UL - UCS/spars$nLC, 0),
-    eC=R$UCPt*R$R/S - UCS,
-    QS=spars$QSm,
-    rhoC=UCS, 
+    jL=env$L[1] * spars$astar,
+    jCP=max(0, synth(jL * spars$nLC, (H$jCO2p[1] + H$jCO2a[1])*H$H/S, spars$jCPm), na.rm=T),
[…]

### 2016-09-18 R/sym_PI.R

@@ -1,21 +1,21 @@
 sym_PI <- function(pars) {
   with(pars, {
-    UL <- seq(0,200,1)
+    jL <- seq(0,200,1)
     # Set carbon delivery rate (mol C / CmolS / d)
-    UCP <- 100  # Extremely high carbon delivery - no carbon limitation
+    jCO2 <- 100  # Extremely high carbon delivery - no carbon limitation
     # Calculate photosynthesis rate
-    UCS <- synth(UCP, UL*spars$nLC, spars$UCSm)
+    jCP <- synth(jCO2, jL*spars$nLC, spars$jCPm)
     # Calculate light in excess of photosynthetic quenching
-    eL <- pmax(0, UL - UCS/spars$nLC)
-    # Calculate ROS (dp) produced due to excess light
-    dp <- pmax(1, 1 + (pmax(0, eL-spars$NPQ)/(spars$L50))^spars$k)
+    jeL <- pmax(0, jL - jCP/spars$nLC)
+    # Calculate ROS (cROS) produced due to excess light
+    cROS <- pmax(1, 1 + (pmax(0, jeL-spars$jNPQ)/(spars$kROS))^spars$k)
     # Calculate photosynthesis rate accounting for ROS-induced photoinhibition
-    UCS <- UCS / dp
+    jCP <- jCP / cROS
     # Return ROS and photosynthesis rate for plotting
     par(mfrow=c(1,2), mar=c(2,2,1,1), mgp=c(1.2,0.2,0), cex=0.7, tck=0.025, xaxs="i")
-    plot(UL, UCS, xlab="Light (mol photons/C-molS/d)", ylab="Carbon fixation (mol C/C-molS/d)",
+    plot(jL, jCP, xlab="Light (mol photons/C-molS/d)", ylab="Carbon fixation (mol C/C-molS/d)",
          main="Photosynthesis", type="l", lwd=3, col="red", ylim=c(0,3))
-    plot(UL, dp, xlab="Light (mol photons/C-molS/d)", ylab="ROS (relative to baseline)",
+    plot(jL, cROS, xlab="Light (mol photons/C-molS/d)", ylab="ROS (relative to baseline)",
          main="ROS production", type="l", lwd=3, col="orange", ylim=c(1,5))
   })
 }

### 2016-09-20 JTB_ms/JTB_ms.Rmd
move to subdir

--- 2016-09-18 Rmd/JTB_ms.Rmd
@@
 
-bibliography: ~/library.bib
+bibliography: library.bib
 output: 
@@
     fig_caption: yes
-header-includes:
-  - \usepackage{tabularx}
 ---
+```{r setup, include=FALSE}
+#knitr::opts_knit$set(root.dir = normalizePath(".."))
+```
+
+```{r load, include=FALSE}
+# Source default parameters and functions
+sapply(list.files(path="../../R", pattern="*.R", full.names=T), source, .GlobalEnv)
+```
 
@@
 
-```{r Fig1, fig.width=5, fig.height=5, echo=FALSE, fig.cap="Graphical representation of coral-algal symbiosis model."}
-library(png)
-library(grid)
-img <- readPNG("../img/model.png")
-grid.raster(img)
-```
+![Graphical representation of coral-algal symbiosis model.](../img/model.png)
 
@@
 
-where $j_{HGm}$ is the maximum specific growth rate.
-Finally, the rate of change in coral biomass is calculated as the specific growth rate minum the specific biomass turnover rate (i.e., maintenance costs), according to:
-
-\begin{equation} {dH \over dt} = j_{HG} - j_{HT}^0 \end{equation}
-
 The amount of nitrogen input to the coral biomass SU in excess of what is actually consumed in biomass formation (i.e., surplus nitrogen, or the "rejection flux" in SU terminology) is then made available to the symbiont, and is specified as:
 
-\begin{equation} \rho_N = j_N + n_{NX}j_X + r_{NH} - n_{NH}j_{HG} \end{equation}
+\begin{equation} \rho_N = (j_N + n_{NX}j_X + r_{NH} - n_{NH}j_{HG})_+ \end{equation}
 
@@
 
-\begin{equation} j_{CP} = \bigg({1 \over j_{CPm}} + {1 \over {n_{LC} j_L }} + {1 \over {(j_{CO_2} + r_{CH}){H \over S} + r_{CS}}} - {1 \over {n_{LC} j_L + (j_{CO_2} + r_{CH}){H \over S} + r_{CS}}}\bigg)^{-1} \div c_{ROS} \end{equation}
+\begin{equation} j_{CP} = \bigg({1 \over j_{CPm}} + {1 \over {n_{LC} j_L }} + {1 \over {(j_{CO_2} + r_{CH}){H \over S} + r_{CS}}} - {1 \over {n_{LC} j_L + (j_{CO_2} + r_{CH}){H \over S} + r_{CS}}}\bigg)^{-1} \cdot c_{ROS}^{-1} \end{equation}
 
@@
 
-\begin{equation} j_{eL} = j_L - j_{CP} / n_{LC} \end{equation}
+\begin{equation} j_{eL} = (j_L - j_{CP} / n_{LC})_+ \end{equation}
 
@@
 
-\begin{equation} \rho_C = j_{CP} - j_{SG} \end{equation}
+\begin{equation} \rho_C = (j_{CP} - j_{SG})_+ \end{equation}
 
@@
 (Note that recycling of symbiont biomass turnover ($rNS$ and $rCS$) only occurs based on the basal maintenance related turnover (i.e., $j_{ST}^0$), and not the bleaching-related biomass loss, as this loss represents biomass being damaged or being expelled from the holobiont).
-Finally, the balance equation representing the change in symbiont biomass over time can be specified as the growth rate minus the turnover rate:
 
-\begin{equation} {dS \over dt} = j_{SG} - j_{ST} \end{equation}
+## Model state equations
+Finally, the balance equations representing the specific growth rates of symbiont and host biomass over time can be expressed as:
+
+\begin{equation} {dS \over Sdt} = j_{SG} - j_{ST} \end{equation}
+
+\begin{equation} {dH \over Hdt} = j_{HG} - j_{HT}^0 \end{equation}
 
@@
 
-The differential equations specified in 7 and 14 were solved using an implicit time-stepping method because of the implicit variables of surplus sharing. Euler method?
+A time-stepping Euler method was used to solve the state equations since the production and rejection fluxes of the SUs are imiplicitly defined.
+Specifically, the rejection fluxes of carbon and nitrogen from the symbiont and host biomass SUs act as reciprocal input fluxes to the other SU.
+In addition, the rejection flux of excitation energy from the photosynthesis SU acts to reduce its own production flux (i.e., photoinhibition), and hence a discretized time-stepping procedure was necessary.
+A vector of time values was created for each simulation run, along which dynamic environmental forcing functions (irradiance, DIN, and prey abundance) can be designed.
+These vectors, along with initial values of symbiont and host biomass, then serve as input to the time-stepping function, which solves for the current system state using values of the previous system state where necessary.
+A default time step of 0.1 days was used for all simulations, which were performed using R code that is available in the data repository accompanying this article: [github.com/jrcunning/Rcoral](github.com/jrcunning/Rcoral).
 
 # Model behavior evaluation
+
+To analyze qualitative behavior, we ran the model to steady state across gradients of external irradiance and DIN, and plotted the specific growth rate (Fig.
+2A) and symbiont to host biomass ratio (Fig.
+2B).
+Results are consistent with observations in corals: specific growth rates
+
+![Steady state values of (A) specific growth ($Cmol \cdot Cmol^{-1} \cdot d^{-1}$) and (B) the symbiont to host biomass ratio ($CmolS \cdot CmolH^{-1}$) across gradients of external irradiance and dissolved inorganic nitrogen.
+Simulations for each combination of light and nutrients (21 points along each axis) were run for 100 days with a time step of 1 day.
+Negative steady state growth rates, and corresponding S:H ratios, were set to zero.](../img/Fig2.png)
 

### 2016-09-20 JTB_ms/JTB_ms.Rmd
update Fig2

@@
 ```{r setup, include=FALSE}
+library(reshape2); library(dplyr)
 #knitr::opts_knit$set(root.dir = normalizePath(".."))
@@
 2B).
-Results are consistent with observations in corals: specific growth rates
+Results are consistent with observations in corals: specific growth rates are low at very low levels of irradiance and DIN, as these scenarios limit the availability of carbon and nitrogen, respectively, to build biomass.
+As these resources increase, specific growth increases, reaching its highest levels above 2 µM DIN and ~15 mol photons m^-2^ d^-1^.
+Peak specific growth rates are around ~0.1 d^-1^, which is consistent with values reported in the literature (@Muscatine:1981jy).
 
-![Steady state values of (A) specific growth ($Cmol \cdot Cmol^{-1} \cdot d^{-1}$) and (B) the symbiont to host biomass ratio ($CmolS \cdot CmolH^{-1}$) across gradients of external irradiance and dissolved inorganic nitrogen.
+![Steady state values of (A) specific growth (Cmol Cmol^-1^ d^-1^) and (B) the symbiont to host biomass ratio (CmolS CmolH^-1^) across gradients of external irradiance and dissolved inorganic nitrogen.
 Simulations for each combination of light and nutrients (21 points along each axis) were run for 100 days with a time step of 1 day.

### 2016-09-20 JTB_ms/JTB_ms.tex
update Fig2

--- 2016-09-18 Rmd/JTB_ms.tex
@@
 \setcounter{secnumdepth}{0}
-\usepackage{tabularx}
 % Pandoc toggle for numbering sections (defaults to be off)
@@
 % Pandoc header
-\usepackage{tabularx}
 
@@
 \centering
-\includegraphics{JTB_ms_files/figure-latex/Fig1-1.pdf}
+\includegraphics{../img/model.png}
 \caption{Graphical representation of coral-algal symbiosis model.}
@@
 
-where \(j_{HGm}\) is the maximum specific growth rate. Finally, the rate
-of change in coral biomass is calculated as the specific growth rate
-minum the specific biomass turnover rate (i.e., maintenance costs),
-according to:
-
-\begin{equation} {dH \over dt} = j_{HG} - j_{HT}^0 \end{equation}
-
 The amount of nitrogen input to the coral biomass SU in excess of what
@@
 
-\begin{equation} \rho_N = j_N + n_{NX}j_X + r_{NH} - n_{NH}j_{HG} \end{equation}
+\begin{equation} \rho_N = (j_N + n_{NX}j_X + r_{NH} - n_{NH}j_{HG})_+ \end{equation}
 
@@
 and not the bleaching-related biomass loss, as this loss represents
-biomass being damaged or being expelled from the holobiont). Finally,
-the balance equation representing the change in symbiont biomass over
-time can be specified as the growth rate minus the turnover rate:
+biomass being damaged or being expelled from the holobiont).
 
-\begin{equation} {dS \over dt} = j_{SG} - j_{ST} \end{equation}
+\subsection{Model state equations}\label{model-state-equations}
+
+Finally, the balance equations representing the specific growth rates of
+symbiont and host biomass over time can be expressed as:
+
+\begin{equation} {dS \over Sdt} = j_{SG} - j_{ST} \end{equation}
+
+\begin{equation} {dH \over Hdt} = j_{HG} - j_{HT}^0 \end{equation}
+
+\subsection{Numerical analysis}\label{numerical-analysis}
+
+A time-stepping Euler method was used to solve the state equations since
+the production and rejection fluxes of the SUs are imiplicitly defined.
+Specifically, the rejection fluxes of carbon and nitrogen from the
+symbiont and host biomass SUs act as reciprocal input fluxes to the
+other SU. In addition, the rejection flux of excitation energy from the
+photosynthesis SU acts to reduce its own production flux (i.e.,
+photoinhibition), and hence a discretized time-stepping procedure was
+necessary. A vector of time values was created for each simulation run,
+along which dynamic environmental forcing functions (irradiance, DIN,
+and prey abundance) can be designed. These vectors, along with initial
+values of symbiont and host biomass, then serve as input to the
+time-stepping function, which solves for the current system state using
+values of the previous system state where necessary. A default time step
+of 0.1 days was used for all simulations, which were performed using R
+code that is available in the data repository accompanying this article:
+\url{github.com/jrcunning/Rcoral}.
 
 \section{Model behavior evaluation}\label{model-behavior-evaluation}
+
+To analyze qualitative behavior, we ran the model to steady state across
+gradients of external irradiance and DIN, and plotted the specific
+growth rate (Fig. 2A) and symbiont to host biomass ratio (Fig. 2B).
+Results are consistent with observations in corals: specific growth
+rates are low at very low levels of irradiance and DIN, as these
+scenarios limit the availability of carbon and nitrogen, respectively,
+to build biomass. As these resources increase, specific growth
+increases, reaching its highest levels above 2 µM DIN and
+\textasciitilde{}15 mol photons m\textsuperscript{-2}
+d\textsuperscript{-1}. Peak specific growth rates are around
+\textasciitilde{}0.1 d\textsuperscript{-1}, which is consistent with
+values reported in the literature (Muscatine, R McCloskey, and E Marian
+(1981)).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{../img/Fig2.png}
+\caption{Steady state values of (A) specific growth (Cmol
+Cmol\textsuperscript{-1} d\textsuperscript{-1}) and (B) the symbiont to
+host biomass ratio (CmolS CmolH\textsuperscript{-1}) across gradients of
+external irradiance and dissolved inorganic nitrogen. Simulations for
+each combination of light and nutrients (21 points along each axis) were
+run for 100 days with a time step of 1 day. Negative steady state growth
+rates, and corresponding S:H ratios, were set to zero.}
+\end{figure}
 
@@
 Costanza, R, R de Groot, and P Sutton. 2014. ``Changes in the global
-value of ecosystem services.'' \emph{Global Environmental \ldots{}} 26:
+value of ecosystem services.'' \emph{Global Environmental Change} 26:
 152--58.
@@
 Science} 69 (1): 79--109.
-\url{http://www.scopus.com/record/display.url?fedsrfIntegrator=MEKPAPERS-SCOCIT\&origin=fedsrf\&view=basic\&eid=2-s2.0-0034761826}.
 
@@
 (3): 201--8.
-\url{http://www.scopus.com/record/display.url?fedsrfIntegrator=MEKPAPERS-SCOCIT\&origin=fedsrf\&view=basic\&eid=2-s2.0-0000237488}.
 
@@
 (7): 454--60.
+
+\hypertarget{ref-Muscatine:1981jy}{}
+Muscatine, Leonard, L R McCloskey, and R E Marian. 1981. ``Estimating
+the daily contribution of carbon from zooxanthellae to coral animal
+respiration.'' \emph{Limnology and Oceanography} 26 (4): 601--11.
+doi:\href{https://doi.org/10.4319/lo.1981.26.4.0601}{10.4319/lo.1981.26.4.0601}.
 

### 2016-09-20 R/Fig2_plot.R

@@ -1,0 +1,80 @@
+library(reshape2)
+library(dplyr)
+load("output/ss.Rdata")
+
+# For runs that did NOT reach a steady state of positive growth, set all results to zero
+ss[which(ss$gr < 0), c("gr", "sh", "hl", "sl", "ee", "pl")] <- NA
+# For runs where steady state value of sh > 1, set sh to 1
+ss[which(ss$sh > 1), "sh"] <- 1
+
+
+# Reshape steady state results into matrices for plotting cont
+sh <- acast(ss, L~N, value.var="sh")
+gr <- acast(ss, L~N, value.var="gr")
+hl <- acast(ss, L~N, value.var="hl")
+sl <- acast(ss, L~N, value.var="sl")
+ee <- acast(ss, L~N, value.var="ee")
+pl <- acast(ss, L~N, value.var="pl")
+
+# Build plotting function
+imagef <- function(r, bin, main, col) {
+  par(mar=c(3,3,3,6), mgp=c(1,0.5,0))
+  if (any(r<0, na.rm=T) & any(r>0, na.rm=T)) {
+    nb <- rev(seq(0, -1*min(r[which(r!=0)], na.rm=T)+bin, bin))*-1  # negative portion of breaks
+    pb <- seq(0, max(r[which(r!=0)], na.rm=T)+bin, bin)
+    breaks <- c(nb, pb[-1])
+  } else {
+    breaks <- seq(min(r[which(r!=0)], na.rm=T), max(r, na.rm=T)+bin, bin)
+  }
+  ncolors <- length(breaks) - 1
+  if (col=="grayscale") {
+    colors <- gray.colors(n=ncolors, start=0.9, end=0.1)
+  } else {
+    dir <- which.max(table(breaks>0))
+    nn <- max(table(breaks>0))*2
+    rc1 <- colorRampPalette(colors=c("#CA0020", "gray90"), space="Lab")(nn/2)
+    rc1 <- rc1[-(length(rc1))]
+    rc2 <- colorRampPalette(colors=c("gray90", "gray30"), space="Lab")((nn/2)+1)[-1]
+    colors <- c(rc1, rc2)
+    if (dir==1) {
+      colors <- colors[1:ncolors]
+    } else {
+      colors <- colors[(length(colors)-ncolors+1):length(colors)]
+    }
+  }
+  image(t(r), breaks=breaks, col=colors, xaxt="n", yaxt="n", main=main, adj=0)
+  mtext(side=1, text="[DIN] (mol/L)", line=2)
[…]

### 2016-09-20 R/Fig2_steadystates.R

@@ -13,11 +13,11 @@ source("R/run_coral.R")
 # Set time vector and default parameters
 time <- seq(1,120,1)
 pars <- def_pars()
-#pars$jCO2a <- 1  # having higher carbon delivery from host reduces S:H ratios and causes photosynthesis to be light-limited more frequently.
+pars$jCO2a <- 1  # having higher carbon delivery from host reduces S:H ratios and causes photosynthesis to be light-limited more frequently.
 
 # Set values of L and N at which to get steady state values
-at <- expand.grid(L=seq(from=0, to=40, length.out=101),
-                  N=seq(from=0, to=2e-6, length.out=101))
+at <- expand.grid(L=seq(from=0, to=40, length.out=41),
+                  N=seq(from=0, to=2e-6, length.out=41))
 
 # Set up cluster for parallel processing
 cl <- makeCluster(detectCores())  # Initiate cluster
@@ -39,85 +39,11 @@ steady_states <- foreach(i=1:nrow(at), .combine='rbind', .packages='dplyr') %dop
   data.frame(gr=gr, sh=sh, hl=hl, sl=sl, ee=ee, pl=pl)
 }
 ss <- cbind(at, steady_states)
-save(ss, file = "data/ss.Rdata")
+save(ss, file = "output/ss.Rdata")
 
 stopCluster(cl)  # Stop cluster
 ptm - proc.time()  # 2601 simulations with 100 time steps each = 427 sec
                    # 10201 simulations with 100 time steps each = 1670 sec
-
-# For runs that did NOT reach a steady state of positive growth, set all results to zero
-ss[which(ss$gr < 0), c("gr", "sh", "hl", "sl", "ee", "pl")] <- NA
-# For runs where steady state value of sh > 1, set sh to 1
-ss[which(ss$sh > 1), "sh"] <- 1
-
-
-# Reshape steady state results into matrices for plotting cont
-sh <- acast(ss, L~N, value.var="sh")
-gr <- acast(ss, L~N, value.var="gr")
-hl <- acast(ss, L~N, value.var="hl")
-sl <- acast(ss, L~N, value.var="sl")
-ee <- acast(ss, L~N, value.var="ee")
-pl <- acast(ss, L~N, value.var="pl")
-
-# Build plotting function
[…]

### 2016-09-21 R/Fig2.R
create steady state plots

@@ -1,0 +1,50 @@
+# This script creates the surface plots used in Figure 2.
+library(png)
+source("R/def_pars.R")
+source("R/init_env.R")
+source("R/run_coral.R")
+source("R/coral_steady_states.R")
+
+
+# Define dimensions for surface plots
+dim <- list(
+  L=seq(from=0, to=40, length.out=21),
+  N=seq(from=0, to=4e-6, length.out=21)
+)
+
+# Set up and run simulations for surface plot using default parameters
+time <- seq(1,100,1)
+env <- init_env(time, L=c(10,10,0), N=c(1e-7,1e-7,0), X=c(0,0,0))
+pars <- def_pars()
+surf <- coral_steady_states(time, env, pars, dim)
+
+# Plot growth surface and save as png
+png("img/growth.png", width=5, height=4.2, units="in", res=300)
+par(mar=c(3,3,3,0), mgp=c(1.5,0.2,0), tck=0.02)
+filled.contour(x=dim[[2]]*10^6, y=dim[[1]], z=t(surf$gr), nlevels=max(surf$gr) %/% 0.01,
+               col=gray.colors(max(surf$gr) %/% 0.01+1, start=1, end=0.2),
+               xlab="[DIN] (µmol/L)", ylab="Irradiance (mol ph/m2/d)")
+title("A. Specific growth rate", adj=0)
+#mtext(side=4, text="Cmol/Cmol/d")
+dev.off()
+
+# Plot growth surface and save as png
+png("img/sh.png", width=5, height=4.2, units="in", res=300)
+par(mar=c(3,3,3,0), mgp=c(1.5,0.2,0), tck=0.02)
+filled.contour(x=dim[[2]]*10^6, y=dim[[1]], z=t(surf$sh), nlevels=max(surf$sh) %/% 0.05,
+               col=gray.colors(max(surf$sh) %/% 0.05+1, start=1, end=0.2),
+               xlab="[DIN] (µmol/L)", ylab="Irradiance (mol ph/m2/d)")
+title("B. Symbiont:host biomass", adj=0)
+dev.off()
+
+# Plot both pngs as a two-panel figure
+png1 <- readPNG("img/growth.png")
+png2 <- readPNG("img/sh.png")
+
+png("img/Fig2.png", width=7, height=2.94, units="in", res=300)
+par(mfrow=c(1,2), mar=c(0,0,0,0))
+plot.new()
[…]

### 2016-09-21 R/coral_steady_states.R

@@ -1,0 +1,34 @@
+coral_steady_states <- function(time, env, pars, at) {
+
+  # Create empty matrices to populate with steady state values (gr=growth, sh=S/H ratio)
+  gr <- matrix(NA, length(dim[[1]]), length(dim[[2]]), dimnames=dim)
+  sh <- matrix(NA, length(dim[[1]]), length(dim[[2]]), dimnames=dim)
+  
+  # Calculate steady states for each combination of points
+  for (x in 1:length(dim[[1]])) {
+    # Create constant vector for dimension 1 at each evaluation point
+    dimx <- seq(dim[[1]][x], dim[[1]][x], along=time)
+    env1 <- replace(env, match(names(dim)[1], names(env)), list(dimx))
+    for (y in 1:length(dim[[2]])) {
+      # Create constant vector for dimension 2 at each evaluation point
+      dimy <- seq(dim[[2]][y], dim[[2]][y], along=time)
+      env2 <- replace(env1, match(names(dim)[2], names(env1)), list(dimy))
+      # Run model to steady state
+      res <- run_coral(time, env2, pars)
+      # Calculate final growth rates and S/H ratios
+      gr[x,y] <- res$H$dH.Hdt[length(res$H$dH.Hdt)]
+      sh[x,y] <- (res$S$S/res$H$H)[length(res$S$S)]
+    }
+  }
+  
+  # Replace any negative growth rates with zero for both responses
+  sh[gr<0] <- 0
+  gr[gr<0] <- 0
+  # Set any NA values to zero
+  sh[is.na(sh)] <- 0
+  gr[is.na(gr)] <- 0
+  
+  # Return results
+  return(list(time=time, env=env, pars=pars, dim=dim, gr=gr, sh=sh))
+  
+}

### 2016-09-21 Rmd/JTB_ms.tex
Numerical analysis and model behavior text

@@
 and not the bleaching-related biomass loss, as this loss represents
-biomass being damaged or being expelled from the holobiont). Finally,
-the balance equation representing the change in symbiont biomass over
-time can be specified as the growth rate minus the turnover rate:
+biomass being damaged or being expelled from the holobiont).
 
-\begin{equation} {dS \over dt} = j_{SG} - j_{ST} \end{equation}
+\subsection{Model state equations}\label{model-state-equations}
+
+Finally, the balance equations representing the specific growth rates of
+symbiont and host biomass over time can be expressed as:
+
+\begin{equation} {dS \over Sdt} = j_{SG} - j_{ST} \end{equation}
+
+\begin{equation} {dH \over Hdt} = j_{HG} - j_{HT}^0 \end{equation}
+
+\subsection{Numerical analysis}\label{numerical-analysis}
+
+A time-stepping Euler method was used to solve the state equations since
+the production and rejection fluxes of the SUs are imiplicitly defined.
+Specifically, the rejection fluxes of carbon and nitrogen from the
+symbiont and host biomass SUs act as reciprocal input fluxes to the
+other SU. In addition, the rejection flux of excitation energy from the
+photosynthesis SU acts to reduce its own production flux (i.e.,
+photoinhibition), and hence a discretized time-stepping procedure was
+necessary. A vector of time values was created for each simulation run,
+along which dynamic environmental forcing functions (irradiance, DIN,
+and prey abundance) can be designed. These vectors, along with initial
+values of symbiont and host biomass, then serve as input to the
+time-stepping function, which solves for the current system state using
+values of the previous system state where necessary. A default time step
+of 0.1 days was used for all simulations, which were performed using R
+code that is available in the data repository accompanying this article:
+\url{github.com/jrcunning/Rcoral}.
 
 \section{Model behavior evaluation}\label{model-behavior-evaluation}
+
+To analyze qualitative behavior, we ran the model to steady state across
+gradients of external irradiance and DIN, and plotted the specific
+growth rate (Fig. 2A) and symbiont to host biomass ratio (Fig. 2B).
+Results are consistent with observations in corals: specific growth
+rates
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{../img/Fig2.png}
+\caption{Steady state values of (A) specific growth
+(\(Cmol \cdot Cmol^{-1} \cdot d^{-1}\)) and (B) the symbiont to host
+biomass ratio (\(CmolS \cdot CmolH^{-1}\)) across gradients of external
+irradiance and dissolved inorganic nitrogen. Simulations for each
+combination of light and nutrients (21 points along each axis) were run
+for 100 days with a time step of 1 day. Negative steady state growth
+rates, and corresponding S:H ratios, were set to zero.}
+\end{figure}
 

### 2016-09-23 Rmd/JTB_ms.tex
add some text

@@
 Results are consistent with observations in corals: specific growth
-rates
+rates are low at very low levels of irradiance and DIN, as these
+scenarios limit the availability of carbon and nitrogen, respectively,
+to build biomass. As these resources increase, specific growth
+increases, reaching its highest levels above 2 µM DIN and
+\textasciitilde{}15 mol photons m\textsuperscript{-2}
+d\textsuperscript{-1}.
 
@@
 \includegraphics{../img/Fig2.png}
-\caption{Steady state values of (A) specific growth
-(\(Cmol \cdot Cmol^{-1} \cdot d^{-1}\)) and (B) the symbiont to host
-biomass ratio (\(CmolS \cdot CmolH^{-1}\)) across gradients of external
-irradiance and dissolved inorganic nitrogen. Simulations for each
-combination of light and nutrients (21 points along each axis) were run
-for 100 days with a time step of 1 day. Negative steady state growth
+\caption{Steady state values of (A) specific growth (\$Cmol
+Cmol\textsuperscript{-1} d\textsuperscript{-1}) and (B) the symbiont to
+host biomass ratio (CmolS CmolH\textsuperscript{-1}) across gradients of
+external irradiance and dissolved inorganic nitrogen. Simulations for
+each combination of light and nutrients (21 points along each axis) were
+run for 100 days with a time step of 1 day. Negative steady state growth
 rates, and corresponding S:H ratios, were set to zero.}

### 2016-09-23 JTB_ms/JTB_ms.Rmd

@@
 ```{r setup, include=FALSE}
-library(reshape2); library(dplyr)
 #knitr::opts_knit$set(root.dir = normalizePath(".."))

### 2016-09-23 JTB_ms/JTB_ms.tex

@@
 Results are consistent with observations in corals: specific growth
-rates are low at very low levels of irradiance and DIN, as these
-scenarios limit the availability of carbon and nitrogen, respectively,
-to build biomass. As these resources increase, specific growth
-increases, reaching its highest levels above 2 µM DIN and
-\textasciitilde{}15 mol photons m\textsuperscript{-2}
-d\textsuperscript{-1}. Peak specific growth rates are around
-\textasciitilde{}0.1 d\textsuperscript{-1}, which is consistent with
-values reported in the literature (Muscatine, R McCloskey, and E Marian
-(1981)).
+rates
 
@@
 \includegraphics{../img/Fig2.png}
-\caption{Steady state values of (A) specific growth (Cmol
-Cmol\textsuperscript{-1} d\textsuperscript{-1}) and (B) the symbiont to
-host biomass ratio (CmolS CmolH\textsuperscript{-1}) across gradients of
-external irradiance and dissolved inorganic nitrogen. Simulations for
-each combination of light and nutrients (21 points along each axis) were
-run for 100 days with a time step of 1 day. Negative steady state growth
+\caption{Steady state values of (A) specific growth
+(\(Cmol \cdot Cmol^{-1} \cdot d^{-1}\)) and (B) the symbiont to host
+biomass ratio (\(CmolS \cdot CmolH^{-1}\)) across gradients of external
+irradiance and dissolved inorganic nitrogen. Simulations for each
+combination of light and nutrients (21 points along each axis) were run
+for 100 days with a time step of 1 day. Negative steady state growth
 rates, and corresponding S:H ratios, were set to zero.}
@@
 
-\hypertarget{ref-Muscatine:1981jy}{}
-Muscatine, Leonard, L R McCloskey, and R E Marian. 1981. ``Estimating
-the daily contribution of carbon from zooxanthellae to coral animal
-respiration.'' \emph{Limnology and Oceanography} 26 (4): 601--11.
-doi:\href{https://doi.org/10.4319/lo.1981.26.4.0601}{10.4319/lo.1981.26.4.0601}.
-
 \hypertarget{ref-RALanguageandEn:2014wf}{}

### 2016-09-23 Rmd/analyze.Rmd
link NPQ capacity to photochemistry

@@
 # Source default parameters and functions
-sapply(list.files(path="R", pattern="*.R", full.names=T), source, .GlobalEnv)
+sapply(c("R/conversions.R", "R/def_pars.R", "R/init_env.R", "R/plot_results.R",
+         "R/run_coral.R", "R/sym_PI.R"), source, .GlobalEnv)
 ```
@@
 # Initialize environment
-env <- init_env(time=time, L=c(10,40,2), N=c(1e-7,1e-7,0), X=c(0,0,0))
+env <- init_env(time=time, L=c(20,32,2), N=c(1e-6,1e-6,0), X=c(0,0,0))
 
@@
 #pars$rpars$gammaR <- 0.03  # Change any default parameters
+#pars$jCO2a <- 1
+#pars$jCO2a <- 0.1
 
@@
 ```{r run1_PI, fig.height=1}
-sym_PI(pars)
+#sym_PI(pars)
 ```
@@
 
-#### Run 1 with no symbiont recycling
-```{r runNR_plots}
-parsNR <- pars
-parsNR$spars$sigmaNS <- 0
-parsNR$spars$sigmaCS <- 0
-runNR <- run_coral(time=time, env=env, pars=parsNR)
-plot_run(runNR)
-with(runNR, {
+##### Run 2 -- increase jCO2a
+```{r run2_plots}
+pars2 <- pars
+pars2$jCO2a <- 10
+run2 <- run_coral(time=time, env=env, pars=pars2)
+plot_run(run2)
 
-})
 ```
-
-#### Compare amount of recycled sym nitrogen between runs
-```{r compareN}
-proprNS <- run$S$rNS / (run$S$rNS + run$H$rhoN*run$H$H/run$S$S)
-par(mfrow=c(2,1))
-plot(time, proprNS)
-proprNS <- runNR$S$rNS / (runNR$S$rNS + runNR$H$rhoN*runNR$H$H/runNR$S$S)
-plot(time, proprNS)
-
-proprCS <- run$S$rCS / (run$S$rCS + run$H$jCO2*run$H$H/run$S$S)
-par(mfrow=c(2,1))
-plot(time, proprCS)
-proprCS <- runNR$S$rCS / (runNR$S$rCS + (runNR$H$rCH + runNR$H$jCO2)*runNR$H$H/runNR$S$S)
-plot(time, proprCS)
-```
-
-### Run 2 - tolerant symbiont
-```{r run2, echo=F}
-pars2 <- pars
-pars2$spars$kROS <- 80
-pars2$spars$jCPm <- 1.4
-
-run2 <- run_coral(time=time, env=env, pars=pars2)
-```
-
-##### Plot symbiont PI curve
-```{r run2_PI, fig.height=2}
-sym_PI(pars2)
-```
-
-##### Plot run 2 output
-```{r run2_plots}
-plot_run(run2)
-```
-
-##### Increase light to bleach tolerant symbiont
-```{r}
-env <- init_env(time=time, L=c(10,60,2), N=c(1e-7,1e-7,0), X=c(0,0,0))
-run <- run_coral(time=time, env=env, pars=pars2)
-plot_run(run)
-```

### 2016-09-23 R/def_pars.R

@@ -18,7 +18,7 @@ def_pars <- function() {
     jST0=0.03,  # Symbiont specific biomass turnover rate (d^-1)
     nNS=0.13,  # N:C ratio in symbiont biomass (-)
     nLC=0.1,  # L:C ratio in fixed carbon (=quantum yield) (molC/mol ph)
-    jNPQ=40,  # capacity of non-photochemical quenching (mol ph/CmolS/d)
+    jNPQ=4,  # capacity of non-photochemical quenching (mol ph/CmolS/d)
     kROS=40,  # amount of (jeL-jNPQ) that doubles ROS production (mol ph/CmolS/d)
     k=1,  # exponent on ROS production (-)
     astar=1.34,  # Symbiont specific cross-sectional area (m^2/C-molS)

### 2016-09-23 R/plot_results.R

@@ -38,11 +38,12 @@ plot_sh <- function(run) with(run, {
 
 # Light quenching (carbon fixation, NPQ, and excess (=ROS producing))
 plot_Lq <- function(run) with(run, {
-  plot(NA, xlim=range(time), xlab="", ylim=c(0, max(c(S$jL, S$jeL))), ylab="mol photons/C-molS/d",
+  plot(NA, xlim=range(time), xlab="", ylim=c(0, max(c(S$jL, S$jeL), na.rm=T)), ylab="mol photons/C-molS/d",
        main="Light quenching")
   lines(time, S$jL, col="yellow", lty=3, lwd=3) # total amount absorbed
   lines(time, S$jL - S$jeL, col="yellow", lwd=1) # amt. used in photosynthesis
-  lines(time, (S$jL - S$jeL) + pmin(S$jeL, pars$jNPQ), col="yellow", lty=1, lwd=3) # amt. quenched by NPQ
+  #lines(time, (S$jL - S$jeL) + pmin(S$jeL, pars$jNPQ), col="yellow", lty=1, lwd=3) # amt. quenched by NPQ
+  lines(time, S$jNPQ, col="yellow", lty=1, lwd=3)
   legend("topright", legend=c("Excess", "NPQ", "Photo."), lty=c(3,1,1), lwd=c(3,3,1), col="yellow", bty="n")
 })
 
@@ -125,11 +126,12 @@ plot_run <- function(run) {
     title("Symbiont to host ratio", adj=0.05, line=0)
 
     # Light quenching (carbon fixation, NPQ, and excess (=ROS producing))
-    plot(NA, xlim=range(time), xlab="", ylim=c(0, max(c(S$jL, S$jeL))), ylab="mol photons/C-molS/d")
+    plot(NA, xlim=range(time), xlab="", ylim=c(0, max(c(S$jL, S$jeL), na.rm=T)), ylab="mol photons/C-molS/d")
     title("Light quenching", adj=0.05, line=0)
     lines(time, S$jL, col="yellow", lty=3, lwd=3) # total amount absorbed
     lines(time, S$jL - S$jeL, col="yellow", lwd=1) # amt. used in photosynthesis
-    lines(time, (S$jL - S$jeL) + pmin(S$jeL, pars$jNPQ), col="yellow", lty=1, lwd=3) # amt. quenched by NPQ
[…]

### 2016-09-23 R/run_coral.R

@@ -38,6 +38,7 @@ run_coral <- function(time, env, pars) {
     jL=env$L[1] * pars$astar,
     jCP=max(0, synth(jL * pars$nLC, (H$jCO2p[1] + H$jCO2a[1])*H$H/S, pars$jCPm), na.rm=T),
     jeL=max(jL - jCP/pars$nLC, 0),
+    jNPQ=pars$jNPQ/pars$nLC,
     jCO2w=H$jCO2*H$H/S - jCP,
     jSG=pars$jSGm,
     rhoC=jCP, 
@@ -69,8 +70,10 @@ run_coral <- function(time, env, pars) {
     S$jCO2w[t] <- max((H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t] - S$jCP[t], 0)
     # Rejection flux: Light (=excitation energy not quenched by carbon fixation)
     S$jeL[t] <- max(S$jL[t] - S$jCP[t]/pars$nLC, 0)
+    # Amount of excitation energy quenched by NPQ
+    S$jNPQ[t] <- (pars$jNPQ * S$jCP[t])/pars$nLC
     # Scaled ROS production due to excess excitation energy (=not quenched by carbon fixation AND NPQ)
-    S$cROS[t] <- 1 + (max(0, (S$jeL[t] - pars$jNPQ)) / pars$kROS) ^ pars$k
+    S$cROS[t] <- 1 + (max(0, (S$jeL[t] - S$jNPQ[t])) / pars$kROS) ^ pars$k
     
     # Symbiont biomass SU
     # ===================

### 2016-09-26 R/Fig2.R
reorganize code making figures

@@ -1,0 +1,16 @@
+# Load functions
+sapply(c("R/def_pars.R",
+         "R/run_steady_states.R", 
+         "R/plot_steady_states.R"), 
+       source, .GlobalEnv)
+
+# Set up simulations
+time <- seq(1,100,1)  # Set time vector
+pars <- def_pars()  # Get default parameters
+# Set values of L and N at which to get steady state values
+at <- expand.grid(L=seq(from=0, to=40, length.out=41),
+                  N=seq(from=0, to=4e-6, length.out=41))
+
+# Get steady state values and plot results to png
+ssdat <- run_steady_states(time=time, pars=pars, at=at)
+plot_steady_states(ssdat, png="img/Fig2.png")

### 2016-09-26 R/Fig2_plot.R

@@ -1,84 +1,0 @@
-library(reshape2)
-library(dplyr)
-load("output/ss.Rdata")
-
-# For runs that did NOT reach a steady state of positive growth, set all results to zero
-ss[which(ss$gr < 0), c("gr", "sh", "hl", "sl", "ee", "pl")] <- NA
-# For runs where steady state value of sh > 1, set sh to 1
-ss[which(ss$sh > 1), "sh"] <- 1
-
-
-# Reshape steady state results into matrices for plotting cont
-sh <- acast(ss, L~N, value.var="sh")
-gr <- acast(ss, L~N, value.var="gr")
-hl <- acast(ss, L~N, value.var="hl")
-sl <- acast(ss, L~N, value.var="sl")
-ee <- acast(ss, L~N, value.var="ee")
-pl <- acast(ss, L~N, value.var="pl")
-
-# Build plotting function
-imagef <- function(r, bin, main, col) {
-  par(tcl=-0.2, cex.main=1, cex.axis=0.75, mar=c(2.3,2.3,2,3.5), mgp=c(0,0.2,0))
-  if (any(r<0, na.rm=T) & any(r>0, na.rm=T)) {
-    nb <- rev(seq(0, -1*min(r[which(r!=0)], na.rm=T)+bin, bin))*-1  # negative portion of breaks
-    pb <- seq(0, max(r[which(r!=0)], na.rm=T)+bin, bin)
-    breaks <- c(nb, pb[-1])
-  } else {
-    breaks <- seq(min(r[which(r!=0)], na.rm=T), max(r, na.rm=T)+bin, bin)
-  }
-  ncolors <- length(breaks) - 1
-  if (col=="grayscale") {
-    colors <- gray.colors(n=ncolors, start=0.9, end=0.1)
-  } else {
-    dir <- which.max(table(breaks>0))
-    nn <- max(table(breaks>0))*2
-    rc1 <- colorRampPalette(colors=c(col[[1]], "gray90"), space="Lab")(nn/2)
-    rc1 <- rc1[-(length(rc1))]
-    rc2 <- colorRampPalette(colors=c("gray90", col[[2]]), space="Lab")((nn/2)+1)[-1]
-    colors <- c(rc1, rc2)
-    if (dir==1) {
-      colors <- colors[1:ncolors]
-    } else {
-      colors <- colors[(length(colors)-ncolors+1):length(colors)]
-    }
-  }
-  image(t(r), breaks=breaks, col=colors, xaxt="n", yaxt="n")
[…]

### 2016-09-26 R/Fig2_steadystates.R

@@ -1,52 +1,0 @@
-# This script creates the surface plots used in Figure 2.
-library(foreach)
-library(parallel)
-library(doParallel)
-library(reshape2)
-library(dplyr)
-
-source("R/def_pars.R")
-source("R/init_env.R")
-source("R/run_coral.R")
-
-
-# Set time vector and default parameters
-time <- seq(1,120,1)
-pars <- def_pars()
-#pars$jCO2a <- 1  # having higher carbon delivery from host reduces S:H ratios and causes photosynthesis to be light-limited more frequently.
-
-# Set values of L and N at which to get steady state values
-at <- expand.grid(L=seq(from=0, to=40, length.out=41),
-                  N=seq(from=0, to=2e-6, length.out=41))
-
-# Set up cluster for parallel processing
-cl <- makeCluster(detectCores())  # Initiate cluster
-registerDoParallel(cl)
-
-ptm <- proc.time()
-# Run simulations in parallel for each combination of L and N values
-steady_states <- foreach(i=1:nrow(at), .combine='rbind', .packages='dplyr') %dopar% {
-  L <- at[i,]$L; N <- at[i,]$N
-  env <- init_env(time=time, L=c(L,L,0), N=c(N,N,0), X=c(0,0,0))
-  run <- run_coral(time=time, env=env, pars=pars)
-  ss <- lapply(run[c("H", "S")], function(x) x[length(time), ])
-  gr <- ss$H$dH.Hdt
-  sh <- ss$S$S/ss$H$H
-  hl <- with(ss, log(  pmin(S$rhoC*S$S/H$H + H$jX, pars$jHGm) / pmin((H$jN + pars$nNX*H$jX + H$rNH) / pars$nNH, pars$jHGm)  ))
-  sl <- with(ss, log(  pmin(S$jCP, pars$jSGm)   /   pmin((H$rhoN*H$H/S$S + S$rNS)/pars$nNS, pars$jSGm)  ))
-  ee <- with(ss, max(0, S$jL - (S$jCP/run$pars$nLC + run$pars$jNPQ)))
-  pl <- with(ss, log(   pmin((H$jCO2 + H$rCH)*H$H/S$S + S$rCS, pars$jCPm)   /   pmin(S$jL * pars$nLC, pars$jCPm)    ))
-  data.frame(gr=gr, sh=sh, hl=hl, sl=sl, ee=ee, pl=pl)
-}
-ss <- cbind(at, steady_states)
[…]

### 2016-09-26 R/Fig3.R

@@ -7,7 +7,7 @@ source("R/run_coral.R")
 source("R/plot_results.R")
 
 # Set run time vector
-time <- seq(1, 365, 0.5)  # if single model run, use time input
+time <- seq(1, 365, 0.1)  # if single model run, use time input
 
 # Initialize environment
 env <- init_env(time=time, 

### 2016-09-26 R/plot_steady_states.R

@@ -1,0 +1,89 @@
+plot_steady_states <- function(ss, png=NULL) {
+  
+  # Load libraries
+  require(reshape2)
+  require(dplyr)
+  
+  # For runs that did NOT reach a steady state of positive growth, set all results to zero
+  ss[which(ss$gr < 0), c("gr", "sh", "hl", "sl", "ee", "pl")] <- NA
+  # For runs where steady state value of S:H > 0.7, set S:H to 0.7
+  ss[which(ss$sh > 0.7), "sh"] <- 0.7
+  
+  # Reshape steady state results into matrices for plotting cont
+  sh <- acast(ss, L~N, value.var="sh")
+  gr <- acast(ss, L~N, value.var="gr")
+  hl <- acast(ss, L~N, value.var="hl")
+  sl <- acast(ss, L~N, value.var="sl")
+  ee <- acast(ss, L~N, value.var="ee")
+  pl <- acast(ss, L~N, value.var="pl")
+  
+  # Build plotting function
+  imagef <- function(r, bin, main, col) {
+    par(tcl=-0.2, cex.main=1, cex.axis=0.75, mar=c(2.3,2.3,2,3.5), mgp=c(0,0.2,0))
+    if (any(r<0, na.rm=T) & any(r>0, na.rm=T)) {
+      nb <- rev(seq(0, -1*min(r[which(r!=0)], na.rm=T)+bin, bin))*-1  # negative portion of breaks
+      pb <- seq(0, max(r[which(r!=0)], na.rm=T)+bin, bin)
+      breaks <- c(nb, pb[-1])
+    } else {
+      breaks <- seq(min(r[which(r!=0)], na.rm=T), max(r, na.rm=T)+bin, bin)
+    }
+    ncolors <- length(breaks) - 1
+    if (col[[1]]=="grayscale") {
+      colors <- gray.colors(n=ncolors, start=0.9, end=0.1)
+    } else if (length(table(breaks>0))>1) {
+      dir <- which.max(table(breaks>0))
+      nn <- max(table(breaks>0))*2
+      rc1 <- colorRampPalette(colors=c(col[[1]], "gray90"), space="Lab")(nn/2)
+      rc1 <- rc1[-(length(rc1))]
+      rc2 <- colorRampPalette(colors=c("gray90", col[[2]]), space="Lab")((nn/2)+1)[-1]
+      colors <- c(rc1, rc2)
+      if (dir==1) {
+        colors <- colors[1:ncolors]
+      } else {
[…]

### 2016-09-26 R/run_steady_states.R

@@ -1,0 +1,45 @@
+run_steady_states <- function(time, pars, at, outfile=NULL, runtime=F) {
+  
+  # Load libraries
+  require(foreach)
+  require(parallel)
+  require(doParallel)
+  require(reshape2)
+  require(dplyr)
+  
+  # Load functions
+  source("R/init_env.R", local=TRUE)
+  source("R/run_coral.R", local=TRUE)
+  
+  # Set up cluster for parallel processing
+  cl <- makeCluster(detectCores())  # Initiate cluster
+  registerDoParallel(cl)
+  
+  # Run simulations in parallel for each combination of "at" values
+  start <- proc.time()  # Start timer...
+  # Calculate steady states...
+  steady_states <- foreach(i=1:nrow(at), .combine='rbind', .packages='dplyr') %dopar% {
+    L <- at[i,]$L; N <- at[i,]$N
+    env <- init_env(time=time, L=c(L,L,0), N=c(N,N,0), X=c(0,0,0))
+    run <- run_coral(time=time, env=env, pars=pars)
+    ss <- lapply(run[c("H", "S")], function(x) x[length(time), ])
+    gr <- ss$H$dH.Hdt
+    sh <- ss$S$S/ss$H$H
+    hl <- with(ss, log(  pmin(S$rhoC*S$S/H$H + H$jX, pars$jHGm) / pmin((H$jN + pars$nNX*H$jX + H$rNH) / pars$nNH, pars$jHGm)  ))
+    sl <- with(ss, log(  pmin(S$jCP, pars$jSGm)   /   pmin((H$rhoN*H$H/S$S + S$rNS)/pars$nNS, pars$jSGm)  ))
+    ee <- with(ss, max(0, S$jL - (S$jCP/run$pars$nLC + run$pars$jNPQ)))
+    pl <- with(ss, log(   pmin((H$jCO2 + H$rCH)*H$H/S$S + S$rCS, pars$jCPm)   /   pmin(S$jL * pars$nLC, pars$jCPm)    ))
+    data.frame(gr=gr, sh=sh, hl=hl, sl=sl, ee=ee, pl=pl)
+  }
+  stopCluster(cl)  # Stop cluster
+  if (runtime==T) print(proc.time() - start)  # Print time elapsed if runtime==T
+  
+  # Collect results in ss
+  ss <- cbind(at, steady_states)
+  
+  # If output filename give, save ss as Rdata
[…]

### 2016-09-26 R/def_pars.R
match KN to Muller 2009

@@ -12,7 +12,7 @@ def_pars <- function() {
     jCO2p=4.04e-3,  # Specific flux of CO2 into tissue due to passive reaction-diffusion (molC/CmolH/d)
     jCO2a=0.32,  # Specific delivery of CO2 to photosynthesis by coral host (molC/CmolH/d)
     jHGm=1,  # Maximum specific host growth rate (CmolH/CmolH/d)
-    KN=0.46e-6,  # Half-saturation constant for host DIN uptake (molN/L)
+    KN=1e-6,  # Half-saturation constant for host DIN uptake (molN/L)
     KX=20e-6,  # Half-saturation constant for host feeding (molX/L)
     initH=1,  # Initial host biomass (CmolH)
     jST0=0.03,  # Symbiont specific biomass turnover rate (d^-1)

### 2016-09-26 R/run_coral.R
decrease photodamage multiplier to 4

@@ -90,7 +90,7 @@ run_coral <- function(time, env, pars) {
     S$jNw[t] <- max(H$rhoN[t-1]*H$H[t-1]/S$S[t-1] + S$rNS[t] - pars$nNS * S$jSG[t], 0)
     
     # Symbiont biomass loss (turnover)
-    S$jST[t] <- pars$jST0 * (1 + 5 * (S$cROS[t] - 1))
+    S$jST[t] <- pars$jST0 * (1 + 4 * (S$cROS[t] - 1))
     
     # Host biomass SU
     # ===============

### 2016-09-26 JTB_ms/JTB_ms.Rmd
more text

--- 2016-09-20 JTB_ms/JTB_ms.Rmd
@@
 ```{r setup, include=FALSE}
-#knitr::opts_knit$set(root.dir = normalizePath(".."))
+library(reshape2); library(dplyr)
+knitr::opts_knit$set(root.dir = normalizePath(".."))
 ```
@@
 # Source default parameters and functions
-sapply(list.files(path="../../R", pattern="*.R", full.names=T), source, .GlobalEnv)
+sapply(list.files(path="../../R", pattern="^[^F].*.R", full.names=T), source, .GlobalEnv)
 ```
@@
 | $n_{NX}$ | N:C molar ratio in prey biomass | 0.13 | -- |
-| $j_{HT}^0$ | Specific turnover rate of host biomass | 0.03 | $d^{-1}$ |
-| $j_{ST}^0$ | Specific turnover rate of symbiont biomass | 0.03 | $d^{-1}$ |
+| $j_{HT}^0$ | Specific turnover rate of host biomass | 0.03 | d^-1^ |
+| $j_{ST}^0$ | Specific turnover rate of symbiont biomass | 0.03 | d^-1^ |
 | $\sigma_{NH}$ | Proportion host nitrogen turnover recycled | 0.9 | -- |
@@
 | $\sigma_{CS}$ | Proportion symbiont carbon turnover recycled | 0.9 | -- |
-| $j_{Xm}$ | Maximum specific feeding rate of host | 0.1292 | $molX \cdot CmolH^{-1} \cdot d^{-1}$ |
-| $K_X$ | Half-saturation constant for prey uptake by host | 20e-6 | $molX \cdot L^{-1}$ | 
-| $j_{Nm}$ | Maximum specific DIN uptake rate by host | 0.048 | $molN \cdot CmolH^{-1} \cdot d^{-1}$ |
-| $K_N$ | Half-saturation constant for DIN uptake by host | 0.46e-6 | $molN \cdot L^{-1}$ | 
-| $j_{CO_2}^p$ | Passive CO~2~ delivery to symbiont | 4.04e-3 | $molC \cdot CmolH^{-1} \cdot d^{-1}$ |
-| $j_{CO_2}^a$ | Active CO~2~ delivery to symbiont | 0.32--18.0 | $molC \cdot CmolH^{-1} \cdot d^{-1}$ |
-| $j_{HGm}$ | Maximum specific growth rate of host | 1 | $d^{-1}$ |
-| $n_{LC}$ | Quantum yield of photosynthesis | 0.1 | $molC \cdot mol ph^{-1}$ |
-| $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | 1.34 | $m^2 \cdot CmolS^{-1}$ |
-| $j_{NPQ}$ | Non-photochemical quenching capacity of symbiont | 40 | $mol ph \cdot CmolS^{-1} \cdot d^{-1}$|
-| $k_{ROS}$ | Excess photon energy that doubles ROS prod. | 40-80 | $mol ph \cdot CmolS^{-1} \cdot d^{-1}$|
+| $j_{Xm}$ | Maximum specific feeding rate of host | 0.1292 | molX CmolH^-1^ d^-1^ |
+| $K_X$ | Half-saturation constant for prey uptake by host | 20e-6 | molX L^-1^ | 
+| $j_{Nm}$ | Maximum specific DIN uptake rate by host | 0.048 | molN CmolH^-1^ d^-1^ |
+| $K_N$ | Half-saturation constant for DIN uptake by host | 0.46e-6 | molN L^-1^ | 
+| $j_{CO_2}^p$ | Passive CO~2~ delivery to symbiont | 4.04e-3 | molC CmolH^-1^ d^-1^ |
+| $j_{CO_2}^a$ | Active CO~2~ delivery to symbiont | 0.32--18.0 | molC CmolH^-1^ d^-1^ |
+| $j_{HGm}$ | Maximum specific growth rate of host | 1 | d^-1^ |
+| $n_{LC}$ | Quantum yield of photosynthesis | 0.1 | molC mol ph^-1^ |
+| $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | 1.34 | m^2^ CmolS^-1^ |
+| $j_{NPQ}$ | Non-photochemical quenching capacity of symbiont | 40 | mol ph CmolS^-1^ d^-1^ |
+| $k_{ROS}$ | Excess photon energy that doubles ROS prod. | 40-80 | mol ph CmolS^-1^ d^-1^ |
 | $k$ | Exponent on ROS production rate | 1 | -- |
-| $j_{CPm}$ | Maximum specific photosynthesis rate of symbiont | 2.8 | $molC \cdot CmolS^{-1} \cdot d^{-1}$ |
-| $j_{SGm}$ | Maximum specific growth rate of symbiont | 0.25 | $d^{-1}$ |
+| $j_{CPm}$ | Maximum specific photosynthesis rate of symbiont | 2.8 | molC CmolS^-1^ d^-1^ |
+| $j_{SGm}$ | Maximum specific growth rate of symbiont | 0.25 | d^-1^ |
 
@@
 
+To aid in visualizing model results, we calculated values to indicate the degree to which product formation at an SU was limited by availability of either of its two substrates
+
+\begin{equation} \log \bigg({{\min(j_{S1}, j_{Pm})} \over {\min(j_{S2}, j_{Pm})}}\bigg) \end{equation}
+
+where $j_{S1}$ and $j_{S2}$ are the specific input fluxes of the two substrates and $j_{Pm}$ is the maximum specific product formation rate, all with units of Cmol Cmol^-1^ d^-1^.
+
 # Model behavior evaluation
 
-To analyze qualitative behavior, we ran the model to steady state across gradients of external irradiance and DIN, and plotted the specific growth rate (Fig.
-2A) and symbiont to host biomass ratio (Fig.
+To analyze qualitative behavior, we ran the model to steady state across gradients of external irradiance and nutrients (Fig.
+2), which revealed patterns that are consistent with observed phenomena in corals.
+
+_Specific growth rates_ are low at low light and DIN (~0.01-0.03 d^-1^), and begin increasing as both of these factors increase (Fig.
+2A).
+Higher light alleviates light-limitation of photosynthesis (Fig.
+2D), and consequently carbon-limitation of host growth (Fig.
 2B).
-Results are consistent with observations in corals: specific growth rates
+Similarly, elevated DIN alleviates nitrogen-limitation of host growth (Fig.
+2B), a trend that was also predicted by the DEB model of @Muller:2009io, and has been observed experimentally [@MullerParker:1994uo; @Tanaka:2013dj].
+Very high nutrient levels may have negative direct and indirect impacts on growth [@Shantz:2015dz], though these are not likely to occur until beyond the range of concentrations considered here [@FerrierPages:2000p1950].
+The maximum specific growth rates of ~0.09 d^-1^, occurring between ~10-15 mol photons m^-2^ s^-1^ light and >1 µM DIN, are an order of magnitude higher than those typically observed for corals [@Osinga:2011kq].
+The primary reasons for this overestimation of host growth are 1) the model does not account for calcification, and 2) it does not account for size-dependent reductions in growth that likely occur as individual polyps and/or colonies grow (Sebens1980?, probably better refs)...
+(i.e., the system continues growing exponentially in this model).
+However, the estimated specific growth rates are consistent with rates of biomass increase in *Aiptasia*, measured at 0.04 d^-1^ [@ArmozaZvuloni:2014ju], suggesting they are biologically realistic.
+They also match the values predicted by the full DEB model of @Muller2009:io.
+Furthermore, the relative differences across environmental gradients are likely to be the same.
 
-![Steady state values of (A) specific growth ($Cmol \cdot Cmol^{-1} \cdot d^{-1}$) and (B) the symbiont to host biomass ratio ($CmolS \cdot CmolH^{-1}$) across gradients of external irradiance and dissolved inorganic nitrogen.
-Simulations for each combination of light and nutrients (21 points along each axis) were run for 100 days with a time step of 1 day.
+As irradiance continues to increase above ~15-20 mol photons m^-2^ d^-1^, growth rates decline until the symbiosis is no longer viable above ~30 µmol photons m^-2^ d^-1^.
+The mechanism underlying this breakdown is the production of ROS caused by excitation energy in excess of what can be quenched by symbiont photosynthesis and non-photochemical quenching.
+ROS production has the consequence of increasing symbiont biomass loss (photodamage) and decreasing photosynthesis rates (photoinhibition), which have the net effect of reducing fixed carbon input to the system below levels necessary to sustain the symbiosis.
+
+_Symbiont to host biomass ratios_ at low DIN (e.g.
+<1 µM) and high light (e.g., ~20 mol ph m^-2^ d^-2^) typical of coral reef environments are around ~0.1-0.3, which are consistent with values reported in the literature (@Muscatine:1981jy; @Edmunds:2011bv).
+S:H ratios increase under higher nutrients () and lower light ().
+
+_
+
+![Steady state values of **A)** specific growth (Cmol Cmol^-1^ d^-1^), **B)** the symbiont to host biomass ratio (CmolS CmolH^-1^), **C)** relative C- or N-limitation of host biomass formation, and **D)** relative CO~2~- or light-limitation of symbiont photosynthesis, across gradients of external irradiance and dissolved inorganic nitrogen.
+Simulations for each combination of light and nutrients (41 points along each axis) were run for 120 days with a time step of 1 day.
 Negative steady state growth rates, and corresponding S:H ratios, were set to zero.](../img/Fig2.png)
 
-# Discussion  -  Potential applications/utility
+# More things to show:
+## varying light: consequences of excess light and light-limitation
+## varying DIN: role of N-limitation and consequences of elevated N
+## elevated nutrients cause system to become C-limited (photo: CO2 limited; host: fixed C limited)
+## C-limitation increases susceptibility to bleaching
+## consequences of a less efficient but more tolerant symbiont
+## Rate of symbiont repopulation (Berner et al. 1993)
+## Bicarbonate addition promotes coral growth (Marubini and Thake 1999)
 

### 2016-09-26 JTB_ms/JTB_ms.tex
more text

@@
 \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-\(d^{-1}\)
+d\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-\(molX \cdot CmolH^{-1} \cdot d^{-1}\)
+molX CmolH\textsuperscript{-1} d\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-\(molX \cdot L^{-1}\)
+molX L\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-\(molN \cdot CmolH^{-1} \cdot d^{-1}\)
+molN CmolH\textsuperscript{-1} d\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-\(molN \cdot L^{-1}\)
+molN L\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-\(molC \cdot CmolH^{-1} \cdot d^{-1}\)
+molC CmolH\textsuperscript{-1} d\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-\(molC \cdot mol ph^{-1}\)
+molC mol ph\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-\(m^2 \cdot CmolS^{-1}\)
+m\textsuperscript{2} CmolS\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-\(mol ph \cdot CmolS^{-1} \cdot d^{-1}\)
+mol ph CmolS\textsuperscript{-1} d\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-\(molC \cdot CmolS^{-1} \cdot d^{-1}\)
+molC CmolS\textsuperscript{-1} d\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 
+To aid in visualizing model results, we calculated values to indicate
+the degree to which product formation at an SU was limited by
+availability of either of its two substrates
+
+\begin{equation} \log \bigg({{\min(j_{S1}, j_{Pm})} \over {\min(j_{S2}, j_{Pm})}}\bigg) \end{equation}
+
+where \(j_{S1}\) and \(j_{S2}\) are the specific input fluxes of the two
+substrates and \(j_{Pm}\) is the maximum specific product formation
+rate, all with units of Cmol Cmol\textsuperscript{-1}
+d\textsuperscript{-1}.
+
 \section{Model behavior evaluation}\label{model-behavior-evaluation}
@@
 To analyze qualitative behavior, we ran the model to steady state across
-gradients of external irradiance and DIN, and plotted the specific
-growth rate (Fig. 2A) and symbiont to host biomass ratio (Fig. 2B).
-Results are consistent with observations in corals: specific growth
-rates
+gradients of external irradiance and nutrients (Fig. 2), which revealed
+patterns that are consistent with observed phenomena in corals.
+
+\emph{Specific growth rates} are low at low light and DIN
+(\textasciitilde{}0.01-0.03 d\textsuperscript{-1}), and begin increasing
+as both of these factors increase (Fig. 2A). Higher light alleviates
+light-limitation of photosynthesis (Fig. 2D), and consequently
+carbon-limitation of host growth (Fig. 2B). Similarly, elevated DIN
+alleviates nitrogen-limitation of host growth (Fig. 2B), a trend that
+was also predicted by the DEB model of Muller et al. (2009), and has
+been observed experimentally (Muller-Parker et al. 1994; Tanaka et al.
+2013). Very high nutrient levels may have negative direct and indirect
+impacts on growth (\textbf{???}), though these are not likely to occur
+until beyond the range of concentrations considered here (Ferrier-Pagès
+et al. 2000). The maximum specific growth rates of \textasciitilde{}0.09
+d\textsuperscript{-1}, occurring between \textasciitilde{}10-15 mol
+photons m\textsuperscript{-2} s\textsuperscript{-1} light and
+\textgreater{}1 µM DIN, are an order of magnitude higher than those
+typically observed for corals (\textbf{???}). The primary reasons for
+this overestimation of host growth are 1) the model does not account for
+calcification, and 2) it does not account for size-dependent reductions
+in growth that likely occur as individual polyps and/or colonies grow
+(Sebens1980?, probably better refs)\ldots{} (i.e., the system continues
+growing exponentially in this model). However, the estimated specific
+growth rates are consistent with rates of biomass increase in
+\emph{Aiptasia}, measured at 0.04 d\textsuperscript{-1} (\textbf{???}),
+suggesting they are biologically realistic. They also match the values
+predicted by the full DEB model of (\textbf{???}). Furthermore, the
+relative differences across environmental gradients are likely to be the
+same.
+
+As irradiance continues to increase above \textasciitilde{}15-20 mol
+photons m\textsuperscript{-2} d\textsuperscript{-1}, growth rates
+decline until the symbiosis is no longer viable above
+\textasciitilde{}30 µmol photons m\textsuperscript{-2}
+d\textsuperscript{-1}. The mechanism underlying this breakdown is the
+production of ROS caused by excitation energy in excess of what can be
+quenched by symbiont photosynthesis and non-photochemical quenching. ROS
+production has the consequence of increasing symbiont biomass loss
+(photodamage) and decreasing photosynthesis rates (photoinhibition),
+which have the net effect of reducing fixed carbon input to the system
+below levels necessary to sustain the symbiosis.
+
+\emph{Symbiont to host biomass ratios} at low DIN (e.g. \textless{}1 µM)
+and high light (e.g., \textasciitilde{}20 mol ph m\textsuperscript{-2}
+d\textsuperscript{-2}) typical of coral reef environments are around
+\textasciitilde{}0.1-0.3, which are consistent with values reported in
+the literature (Muscatine, R McCloskey, and E Marian (1981); Edmunds et
+al. (2011)). S:H ratios increase under higher nutrients () and lower
+light ().
+
+\_
 
@@
 \includegraphics{../img/Fig2.png}
-\caption{Steady state values of (A) specific growth
-(\(Cmol \cdot Cmol^{-1} \cdot d^{-1}\)) and (B) the symbiont to host
-biomass ratio (\(CmolS \cdot CmolH^{-1}\)) across gradients of external
-irradiance and dissolved inorganic nitrogen. Simulations for each
-combination of light and nutrients (21 points along each axis) were run
-for 100 days with a time step of 1 day. Negative steady state growth
-rates, and corresponding S:H ratios, were set to zero.}
+\caption{Steady state values of \textbf{A)} specific growth (Cmol
+Cmol\textsuperscript{-1} d\textsuperscript{-1}), \textbf{B)} the
+symbiont to host biomass ratio (CmolS CmolH\textsuperscript{-1}),
+\textbf{C)} relative C- or N-limitation of host biomass formation, and
+\textbf{D)} relative CO\textsubscript{2}- or light-limitation of
+symbiont photosynthesis, across gradients of external irradiance and
+dissolved inorganic nitrogen. Simulations for each combination of light
+and nutrients (41 points along each axis) were run for 120 days with a
+time step of 1 day. Negative steady state growth rates, and
+corresponding S:H ratios, were set to zero.}
 \end{figure}
 
-\section{Discussion - Potential
-applications/utility}\label{discussion---potential-applicationsutility}
+\section{More things to show:}\label{more-things-to-show}
+
+\subsection{varying light: consequences of excess light and
+light-limitation}\label{varying-light-consequences-of-excess-light-and-light-limitation}
+
+\subsection{varying DIN: role of N-limitation and consequences of
+elevated
+N}\label{varying-din-role-of-n-limitation-and-consequences-of-elevated-n}
+
+\subsection{elevated nutrients cause system to become C-limited (photo:
+CO2 limited; host: fixed C
+limited)}\label{elevated-nutrients-cause-system-to-become-c-limited-photo-co2-limited-host-fixed-c-limited}
+
+\subsection{C-limitation increases susceptibility to
+bleaching}\label{c-limitation-increases-susceptibility-to-bleaching}
+
+\subsection{consequences of a less efficient but more tolerant
+symbiont}\label{consequences-of-a-less-efficient-but-more-tolerant-symbiont}
+
+\subsection{Rate of symbiont repopulation (Berner et al.
+1993)}\label{rate-of-symbiont-repopulation-berner-et-al.-1993}
+
+\subsection{Bicarbonate addition promotes coral growth (Marubini and
+Thake
+1999)}\label{bicarbonate-addition-promotes-coral-growth-marubini-and-thake-1999}
 
@@
 
+\hypertarget{ref-Edmunds:2011bv}{}
+Edmunds, Peter J, Hollie M Putnam, Roger M Nisbet, and Erik B Muller.
+2011. ``Benchmarks in organism performance and their use in comparative
+analyses.'' \emph{Oecologia} 167 (2): 379--90.
+doi:\href{https://doi.org/10.1007/s00442-011-2004-2}{10.1007/s00442-011-2004-2}.
+
 \hypertarget{ref-Enriquez:2005p142}{}
@@
 \url{http://www.sciencedirect.com/science/article/pii/S0304380011000263}.
+
+\hypertarget{ref-FerrierPages:2000p1950}{}
+Ferrier-Pagès, Christine, Jean-Pierre Gattuso, S Dallot, and Jean
+Jaubert. 2000. ``Effect of nutrient enrichment on growth and
+photosynthesis of the zooxanthellate coral Stylophora \ldots{}.''
+\emph{Coral Reefs}.
+\url{http://www.springerlink.com/index/3DRC3K1Q57TPC0UT.pdf}.
 
@@
 
+\hypertarget{ref-MullerParker:1994uo}{}
+Muller-Parker, Gisèle, Lawrence R McCloskey, Ove Hoegh-Guldberg, and PJ
+McAuley. 1994. ``Effect of ammonium enrichment on animal and algal
+biomass of the coral Pocillopora damicornis.'' \emph{Pacific Science} 48
+(3). \url{http://scholarspace.manoa.hawaii.edu/handle/10125/2236}.
+
 \hypertarget{ref-Muscatine:1977p4220}{}
@@
 \url{http://journal.frontiersin.org/Journal/10.3389/fmicb.2014.00422/pdf}.
+
+\hypertarget{ref-Tanaka:2013dj}{}
+Tanaka, Yasuaki, Akira Iguchi, Mayuri Inoue, Chiharu Mori, K Sakai,
+Atsushi Suzuki, Hodaka Kawahata, and Takashi Nakamura. 2013. ``Marine
+Pollution Bulletin.'' \emph{Marine Pollution Bulletin} 68 (1-2): 93--98.
+doi:\href{https://doi.org/10.1016/j.marpolbul.2012.12.017}{10.1016/j.marpolbul.2012.12.017}.
 

### 2016-09-26 JTB_ms/JTB_ms.Rmd
change parameter jNPQ to kNPQ

@@
 | $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | 1.34 | m^2^ CmolS^-1^ |
-| $j_{NPQ}$ | Non-photochemical quenching capacity of symbiont | 40 | mol ph CmolS^-1^ d^-1^ |
+| $k_{NPQ}$ | NPQ capacity of symbiont, relative to C-fixation | 4 | -- |
 | $k_{ROS}$ | Excess photon energy that doubles ROS prod. | 40-80 | mol ph CmolS^-1^ d^-1^ |
@@
 This excess light energy must be quenched by alternative pathways in order to prevent photo-oxidative damage.
-*Symbiodinium* may utilize a variety of pathways for non-photochemical quenching [NPQ; @Roth:2014wf], which we collect in a total capacity for NPQ as a parameter of the symbiont ($j_{NPQ}$).
+*Symbiodinium* may utilize a variety of pathways for non-photochemical quenching [NPQ; @Roth:2014wf], which we collect in a total capacity for NPQ as a parameter of the symbiont ($k_{NPQ}$).
 If light energy exceeds the total capacity of both carbon fixation and NPQ, then damaging reactive oxygen species (ROS) may be produced.
@@
 
-where $j_{NPQ}$, $k_{ROS}$, and $k$ are parameters of the symbiont that determine the onset and rate of ROS production.
+where $k_{NPQ}$, $k_{ROS}$, and $k$ are parameters of the symbiont that determine the onset and rate of ROS production.
 Importantly, $c_{ROS}$ as specified here is not a function of absolute light absorption, but rather the amount of excess light energy $j_{eL}$ after accounting for carbon fixation and NPQ.
 A direct consequence of this formulation is that carbon-limitation of photosynthesis can lead to photo-oxidative stress, a mechanism of biological importance [@Wooldridge:2009p7807] that was not captured by previous representations of photo-oxidative stress [@Eynaud:2011tv].
-Moreover, this formulation allows functional diversity among symbiont types to be explored by changing the parameters $j_{NPQ}$, $k_{ROS}$, and $k$.
+Moreover, this formulation allows functional diversity among symbiont types to be explored by changing the parameters $k_{NPQ}$, $k_{ROS}$, and $k$.
 

### 2016-09-26 R/def_pars.R

@@ -18,7 +18,7 @@ def_pars <- function() {
     jST0=0.03,  # Symbiont specific biomass turnover rate (d^-1)
     nNS=0.13,  # N:C ratio in symbiont biomass (-)
     nLC=0.1,  # L:C ratio in fixed carbon (=quantum yield) (molC/mol ph)
-    jNPQ=4,  # capacity of non-photochemical quenching (mol ph/CmolS/d)
+    kNPQ=4,  # capacity of non-photochemical quenching (mol ph/CmolS/d)
     kROS=40,  # amount of (jeL-jNPQ) that doubles ROS production (mol ph/CmolS/d)
     k=1,  # exponent on ROS production (-)
     astar=1.34,  # Symbiont specific cross-sectional area (m^2/C-molS)

### 2016-09-26 R/run_coral.R

@@ -38,7 +38,7 @@ run_coral <- function(time, env, pars) {
     jL=env$L[1] * pars$astar,
     jCP=max(0, synth(jL * pars$nLC, (H$jCO2p[1] + H$jCO2a[1])*H$H/S, pars$jCPm), na.rm=T),
     jeL=max(jL - jCP/pars$nLC, 0),
-    jNPQ=pars$jNPQ/pars$nLC,
+    jNPQ=pars$kNPQ/pars$nLC,
     jCO2w=H$jCO2*H$H/S - jCP,
     jSG=pars$jSGm/10,
     rhoC=jCP, 
@@ -68,10 +68,10 @@ run_coral <- function(time, env, pars) {
     S$jCP[t] <- synth(S$jL[t] * pars$nLC, (H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t], pars$jCPm) / S$cROS[t-1]
     # Rejection flux: CO2 (wasted to the environment)
     S$jCO2w[t] <- max((H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t] - S$jCP[t], 0)
-    # Rejection flux: Light (=excitation energy not quenched by carbon fixation)
+    # Rejection flux: excess light energy not quenched by carbon fixation
     S$jeL[t] <- max(S$jL[t] - S$jCP[t]/pars$nLC, 0)
-    # Amount of excitation energy quenched by NPQ
-    S$jNPQ[t] <- min(S$jeL[t], (pars$jNPQ * S$jCP[t])/pars$nLC)
+    # Amount of excess light energy quenched by NPQ
+    S$jNPQ[t] <- min(S$jeL[t], (S$jCP[t] * pars$kNPQ)/pars$nLC)
     # Scaled ROS production due to excess excitation energy (=not quenched by carbon fixation AND NPQ)
     S$cROS[t] <- 1 + (max(0, (S$jeL[t] - S$jNPQ[t])) / pars$kROS) ^ pars$k
     

### 2016-10-06 JTB_ms/JTB_ms.Rmd
updates from UCSB

--- 2016-09-23 JTB_ms/JTB_ms.Rmd
@@
 ```{r load, include=FALSE}
-# Source default parameters and functions
-sapply(list.files(path="../../R", pattern="*.R", full.names=T), source, .GlobalEnv)
+# Source metrics reported in text
+source("R/reported_metrics.R")
 ```
@@
 However, the complex symbiotic metabolism of corals is vulnerable to disruption by numerous anthropogenic environmental perturbations, jeopardizing their future persistence.
-In order to understand and predict coral responses to complex changes in the environment, a mechanistic understanding of how multiple interacting factors drive the individual and emergent physiology of both symbiotic partners is necessary.
+In order to understand and predict coral responses to complex changes in their environment, a mechanistic understanding of how multiple interacting factors drive the individual and emergent physiology of both symbiotic partners is necessary.
 Such a task is well suited for theoretical modeling frameworks such as Dynamic Energy Budget (DEB) theory [@Kooijman:2010vd], although the complexity of such theory makes these efforts inaccessible to many biologists [@Jager:2013bj].
-In order to bridge this gap, we present here a simplified dynamic bioenergetic model for coral-*Symbiodinium* symbioses that aims to mechanistically integrate the impacts of complex environmental change on the physiological ecology of reef corals.
+In order to bridge this gap, we present here a simplified dynamic bioenergetic model for coral-*Symbiodinium* symbioses that aims to mechanistically integrate the impacts of complex environmental change on the physiological performance of reef corals, including responses to environmental stress.
 
-In reef coral symbioses, intracellular *Symbiodinium* translocate photosynthetically-fixed carbon to support coral metabolism, and utilize the animal’s metabolic waste products, including nitrogenous compounds and carbon dioxide, in return [@Muscatine:1977p4220].
+In reef coral symbioses, intracellular *Symbiodinium* translocate photosynthetically fixed carbon to support coral metabolism, while the animal host provides access to inorganic nutrients and carbon dioxide [@Muscatine:1977p4220].
 Previous application of DEB theory to this syntrophic system [@Muller:2009io] demonstrated a stable symbiotic relationship and qualitatively realistic growth and biomass ratios across gradients of ambient irradiance, nutrients, and food.
-This model assumed that 1) *Symbiodinium* has priority access to carbon through photosynthesis, 2) the coral animal has priority access to dissolved nitrogen through contact with seawater, and 3) each partner shares with the other only what it cannot use itself.
-In its simplest form, this principle of sharing the surplus sufficiently describes diverse syntrophic interactions among organs and organisms (e.g., trees, duckweeds, corals), suggesting the mechanism is mathematically and evolutionarily robust (Nisbet et al., submitted).
+This model assumed that 1) *Symbiodinium* has priority access to fixed carbon through photosynthesis, 2) the coral animal has priority access to dissolved nitrogen through contact with seawater, and 3) each partner shares with the other only what it cannot use for its own growth.
+In its simplest form, this principle of sharing the surplus is sufficient to describe the dynamics of diverse syntrophic organs and organisms (e.g., trees, duckweeds, corals), suggesting the mechanism is mathematically and evolutionarily robust (Nisbet et al., in prep.).
 
-While the work of @Muller:2009io applying DEB formalism to this system represents the most significant theoretical contribution in coral symbiosis research to date, we aim to strengthen the role of theory and broaden its potential application to corals in three primary ways:
+While the formal DEB model of @Muller:2009io represents the most significant theoretical contribution in coral symbiosis research to date, we aim to strengthen the role of theory and broaden its potential application in three primary ways:
 
@@
 Here, we present a model without reserves, maturity, or reproduction [see @Kooijman:2010vd].
-This formulation restricts the model's scope to the bioenergetics of growth and symbiosis dynamics in adult corals, but greatly reduces theoretical complexity and parameter numbers, which is advantageous given the relative paucity of data for corals.
+This formulation restricts the model's scope to the bioenergetics of growth and symbiosis dynamics in adult corals (i.e., reproduction, larval stages, and metamorphosis are not considered), but greatly reduces theoretical complexity and parameter numbers, which is advantageous given the relative paucity of data for corals.
 However, our primary motivation for reducing complexity was to increase accessibility and applicability for biologists and ecologists without requiring significant expertise in DEB theory.
@@
 With an accessible and modular framework, we envision this as a resource for futher development by the scientific community to include additional complexity and problem-specific components.
-The R language was chosen because it is freely available and in common use by biologists and ecologists, to widen the audience for this work.
+To widen the audience for this work, we chose the R language because it is freely available and in common use by biologists and ecologists.
 
@@
 # Model description
-In this model of coral-algal symbiosis, carbon and nitrogen are acquired by each partner and used to construct biomass.
+In this dynamic bioenergetic model of coral-algal symbiosis, carbon and nitrogen are acquired by each partner and used to construct biomass.
 A graphical representation of the model is presented in Fig.
-1.
+1, and each model flux and parameter are defined in Tables 1 and 2, respectively.
 We use C-moles as the unit of biomass for consistency with the rigorous mass balance of DEB theory: 1 C-mole is equivalent to the amount of biomass containing 1 mole of Carbon atoms.
-Host biomass ($H$), symbiont biomass ($S$), and prey biomass ($X$) have fixed, but different, molar N:C ratios (Table 1).
-Carbon and nitrogen are combined to produce biomass by synthesizing units (SU), which are mathematical specifications of the formation of a product from two substrates; we use the "parallel complementary" formulation of @Kooijman:2010vd to specify these fluxes.
-The two state variables in this dynamical system are symbiont biomass and coral biomass; because resources are acquired proportionally to surface area (and surface area is assumed proportional to volume for corals (i.e., they are "V1-morphs" in DEB terminology [@Kooijman:2010vd])), biomass increases exponentially during growth.
-The rate of increase in coral biomass (i.e.
-growth) and the ratio of symbiont to host biomass are the (i.e.
-symbiosis dynamics) are the responses of interest of the system.
+Host biomass ($H$), symbiont biomass ($S$), and prey biomass ($X$) have fixed, but different, molar N:C ratios (Table 2).
+Carbon and nitrogen are combined to produce biomass by synthesizing units (SU), which are mathematical specifications of the formation of a product from two substrates; we use the parallel complementary formulation of @Kooijman:2010vd to specify these fluxes.
+The two state variables are symbiont biomass and coral biomass; because resources are acquired proportionally to surface area, and surface area is proportional to volume (i.e., corals are "V1-morphs" in DEB terminology [@Kooijman:2010vd]), biomass increases exponentially during growth (indeed, corals grow exponentially [@Bak:1976bv]).
+The specific growth rate and the ratio of symbiont to host biomass are the responses of interest of the system.
 Below we describe the formulation of each flux involved in producing these responses.
 
-Table 1
+Table 1. Model fluxes
+
+| Symbol    | Description                            | Units                | Eq. no. |
+| --------- | -------------------------------------- | -------------------- | ------- |
+| $j_X$     | Prey uptake rate                       | molX CmolH^-1^ d^-1^ | 3 |
+| $j_N$     | Nitrogen uptake rate                   | molN CmolH^-1^ d^-1^ | 4 |
+| $j_{HG}$  | Host biomass formation rate            | CmolH CmolH^-1^ d^-1^| 5 |
+| $r_{NH}$  | Recycled nitrogen from host turnover   | molN CmolH^-1^ d^-1^ | 6 |
+| $\rho_N$  | Nitrogen shared with the symbiont      | molN CmolH^-1^ d^-1^ | 7 |
+| $j_L$     | Light absorption rate                  | molph CmolS^-1^ d^-1^| 9 |
+| $j_{CO_2}$| CO~2~ input to photosynthesis          | molCO~2~ CmolH^-1^ d^-1^| 10 |
+| $r_{CH}$  | Recycled CO~2~ from host turnover      | molCO~2~ CmolH^-1^ d^-1^| 11 |
+| $r_{CS}$  | Recycled CO~2~ from symbiont turnover  | molCO~2~ CmolS^-1^ d^-1^| 12 |
+| $j_{CP}$  | Photosynthesis rate                    | molC CmolS^-1^ d^-1^ | 13 |
+| $j_{eL}$  | Light energy in excess of photochemistry| molph CmolS^-1^ d^-1^| 14 |
+| $j_{NPQ}$ | Total capacity of NPQ                  | molph CmolS^-1^ d^-1^ | 15 |
+| $c_{ROS}$ | ROS production proportional to baseline | -- | 16 |
+| $r_{NS}$  | Recycled nitrogen from symbiont turnover| molN CmolS^-1^ d^-1^| 17 |
+| $j_{SG}$  | Symbiont biomass formation rate        | CmolS CmolS^-1^ d^-1^| 18 |
+| $\rho_C$  | Fixed carbon shared with host          | molC CmolS^-1^ d^-1^ | 19 |
+| $j_{ST}$  | Symbiont turnover rate                 | CmolS CmolS^-1^ d^-1^| 20 |
+
+Table 2. Model parameters
 
@@
 | $n_{NX}$ | N:C molar ratio in prey biomass | 0.13 | -- |
-| $j_{HT}^0$ | Specific turnover rate of host biomass | 0.03 | $d^{-1}$ |
-| $j_{ST}^0$ | Specific turnover rate of symbiont biomass | 0.03 | $d^{-1}$ |
+| $j_{HT}^0$ | Specific turnover rate of host biomass | 0.03 | CmolH CmolH^-1^ d^-1^ |
+| $j_{ST}^0$ | Specific turnover rate of symbiont biomass | 0.03 | CmolS CmolS^-1^ d^-1^ |
 | $\sigma_{NH}$ | Proportion host nitrogen turnover recycled | 0.9 | -- |
@@
 | $\sigma_{CS}$ | Proportion symbiont carbon turnover recycled | 0.9 | -- |
-| $j_{Xm}$ | Maximum specific feeding rate of host | 0.1292 | $molX \cdot CmolH^{-1} \cdot d^{-1}$ |
-| $K_X$ | Half-saturation constant for prey uptake by host | 20e-6 | $molX \cdot L^{-1}$ | 
-| $j_{Nm}$ | Maximum specific DIN uptake rate by host | 0.048 | $molN \cdot CmolH^{-1} \cdot d^{-1}$ |
-| $K_N$ | Half-saturation constant for DIN uptake by host | 0.46e-6 | $molN \cdot L^{-1}$ | 
-| $j_{CO_2}^p$ | Passive CO~2~ delivery to symbiont | 4.04e-3 | $molC \cdot CmolH^{-1} \cdot d^{-1}$ |
-| $j_{CO_2}^a$ | Active CO~2~ delivery to symbiont | 0.32--18.0 | $molC \cdot CmolH^{-1} \cdot d^{-1}$ |
-| $j_{HGm}$ | Maximum specific growth rate of host | 1 | $d^{-1}$ |
-| $n_{LC}$ | Quantum yield of photosynthesis | 0.1 | $molC \cdot mol ph^{-1}$ |
-| $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | 1.34 | $m^2 \cdot CmolS^{-1}$ |
-| $j_{NPQ}$ | Non-photochemical quenching capacity of symbiont | 40 | $mol ph \cdot CmolS^{-1} \cdot d^{-1}$|
-| $k_{ROS}$ | Excess photon energy that doubles ROS prod. | 40-80 | $mol ph \cdot CmolS^{-1} \cdot d^{-1}$|
+| $j_{Xm}$ | Maximum specific feeding rate of host | 0.1292 | molX CmolH^-1^ d^-1^ |
+| $K_X$ | Half-saturation constant for prey uptake by host | 20e-6 | molX L^-1^ | 
+| $j_{Nm}$ | Maximum specific DIN uptake rate by host | 0.048 | molN CmolH^-1^ d^-1^ |
+| $K_N$ | Half-saturation constant for DIN uptake by host | 0.46e-6 | molN L^-1^ | 
+| $j_{CO_2}^p$ | Passive CO~2~ delivery to symbiont | 4.04e-3 | molC CmolH^-1^ d^-1^ |
+| $j_{CO_2}^a$ | Active CO~2~ delivery to symbiont | 0.32--18.0 | molC CmolH^-1^ d^-1^ |
+| $j_{HGm}$ | Maximum specific growth rate of host | 1 | CmolH CmolH^-1^ d^-1^ |
+| $n_{LC}$ | Quantum yield of photosynthesis | 0.1 | molC mol ph^-1^ |
+| $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | 1.34 | m^2^ CmolS^-1^ |
+| $k_{NPQ}$ | NPQ capacity of symbiont, relative to C-fixation | 4 | -- |
+| $k_{ROS}$ | Excess photon energy that doubles ROS production, relative to baseline levels | 40-80 | mol ph CmolS^-1^ d^-1^ |
 | $k$ | Exponent on ROS production rate | 1 | -- |
-| $j_{CPm}$ | Maximum specific photosynthesis rate of symbiont | 2.8 | $molC \cdot CmolS^{-1} \cdot d^{-1}$ |
-| $j_{SGm}$ | Maximum specific growth rate of symbiont | 0.25 | $d^{-1}$ |
+| $j_{CPm}$ | Maximum specific photosynthesis rate of symbiont | 2.8 | molC CmolS^-1^ d^-1^ |
+| $j_{SGm}$ | Maximum specific growth rate of symbiont | 0.25 | CmolS CmolS^-1^ d^-1^ |
+| $b$ | Scaling parameter for strength of bleaching response | 3 | -- |
 
-Also include tables of fluxes and environmental inputs?
+![Graphical representation of coral-algal symbiosis model.
+Light, CO~2~, prey, and DIN are acquired from the external environment proportional to the biomass of the partner indicated by the black box for uptake.
+Mass fluxes (see Table 1 for definitions) are represented by $j$'s with subscripts indicating the type of mass, and in some cases the process (e.g., $j_{CP}$ is the flux of carbon produced by photosynthesis), and $\rho$'s indicate fluxes that are shared by one partner with the other.
+Parallel complementary synthesizing units (SUs) are represented by large circles, and rejection fluxes from these SUs are indicated by dashed lines.
+$c_{ROS}$ is a proportional rate that impacts other model fluxes by inhibition or acceleration.
+Recycled mass fluxes from biomass turnover are not shown for clarity (but see Table 1 for definitions).](../img/Fig1.png)
 
-![Graphical representation of coral-algal symbiosis model.](../img/model.png)
+## State equations
+The balance equations for symbiont and host biomass are expressed as:
+
+\begin{equation} {dS \over Sdt} = j_{SG} - j_{ST} \end{equation}
+
+\begin{equation} {dH \over Hdt} = j_{HG} - j_{HT}^0 \end{equation}
+
+These are expressed per unit of symbiont and host biomass, respectively.
+The specific biomass growth and turnover rates that define these balance equations are produced by combinations of the individual model fluxes (see Table 1 for definitions and units), which are each expressed as mass-specific rates (e.g., per C-mole of symbiont or host biomass).
+When necessary, conversions between symbiont-mass-specific and host-mass-specific rates are accomplished by multiplying or dividing by the symbiont:host biomass ratio.
 
@@
 The coral animal acquires both carbon and nitrogen from feeding on prey from the environment.
-Prey acquisition is specified by Michaelis-Menten kinetics using a maximum area-specific feeding rate and half-saturation constant:
+Prey acquisition is specified by Michaelis-Menten kinetics (i.e., a Holling type II function) with a maximum feeding rate $j_{Xm}$ and half-saturation constant $K_X$:
 
@@
 
-Additionally, the coral animal can acquire nitrogen from the surrounding seawater.
-This nitrogen source is assumed to represent ammonium, the primary form utilized by corals [@Wang:1998p128; @Yellowlees:2008p331].
-The uptake of nitrogen from the environment is specified by Michaelis-Menten kinetics using a maximum area-specific uptake rate and half-saturation constant:
+Additionally, the coral animal acquires nitrogen dissolved in the surrounding seawater, which is assumed to represent ammonium, the primary form utilized by corals [@Yellowlees:2008p331].
+This gives the host (rather than the symbiont) priority in nitrogen utilization; this capacity is supported by experimental evidence [@Wang:1998p128] and is consistent with the physical arrangment of the partners, where the host is in direct contact with the external environment.
+The uptake of nitrogen from the environment is thus specified by Michaelis-Menten kinetics using a maximum uptake rate $j_{Nm}$ and half-saturation constant $K_N$:
 
@@
 
-Coral biomass formation is specified by a parallel complementary SU that combines carbon and nitrogen to form biomass.
-In addition to the carbon and nitrogen acquired by direct uptake (Eqs.
-1 and 2), the coral recycles a portion of the nitrogen liberated by biomass turnover ($r_{NH}=\sigma_{NH}j_{HT}^0$), and receives surplus fixed carbon shared by the symbiont ($\rho_C$), such that the total biomass formation is specified as:
+Coral biomass formation is then specified by a parallel complementary SU that combines carbon and nitrogen, according to:
 
@@
 
-The amount of nitrogen input to the coral biomass SU in excess of what is actually consumed in biomass formation (i.e., surplus nitrogen, or the "rejection flux" in SU terminology) is then made available to the symbiont, and is specified as:
+where $\rho_C$ is the surplus fixed carbon shared by the symbiont (see below), and $r_{NH}$ is the recycled portion of the nitrogen liberated by host biomass turnover:
+
+\begin{equation} r_{NH}=\sigma_{NH}j_{HT}^0 \end{equation}
+
+The amount of nitrogen input to the coral biomass SU in excess of what is actually consumed in biomass formation (i.e., surplus nitrogen, or the rejection flux[^1] of the SU) is then made available to the symbiont:
 
@@
 
-Due to the inherent inefficiency of the parallel complementary SU formulation, there will always be some nitrogen shared with the symbiont even when coral biomass formation is strongly nitrogen-limited.
+[^1]: Rejection fluxes must always be positive, and hence are specified with the notation $(x)_+$, which means $max(x, 0)$.
+
+Due to the inherent inefficiency of the parallel complementary SU formulation, there is always some nitrogen shared with the symbiont even when coral biomass formation is strongly nitrogen-limited.
 Likewise, there is always a non-zero rejection flux of carbon from the coral biomass SU, which is assumed to be lost to the environment.
@@
 The amount of light absorbed by the symbiont depends on the scalar irradiance at the site of light absorption, which is modified substantially relative to external downwelling irradiance owing to multiple scattering by the coral skeleton and self-shading by surrounding symbionts [@Enriquez:2005p142; @Marcelino:2013hz].
-We used data from @Marcelino:2013hz to empirically derive the ratio of internal scalar irradiance to external downwelling irradiance as a function of symbiont density (expressed as symbiont to host biomass ratio), and subsequently multiply this quantity by the external downwelling irradiance $L$ and the effective light-absorbing surface area of symbiont biomass $ds$ to specify the amount of light absorbed:
+We used data from @Marcelino:2013hz to empirically derive an "amplification factor" $A$ (the ratio of internal scalar irradiance to external downwelling irradiance) as a function of symbiont density (expressed as symbiont to host biomass ratio), which is specified as:
 
-\begin{equation} j_L = \big[1.26 + 1.39 \cdot \exp(-6.48 \cdot {S \over H})\big] \cdot L \cdot \bar{a}^* \end{equation}
+\begin{equation} A = 1.26 + 1.39 \cdot \exp(-6.48 \cdot {S \over H}) \end{equation}
 
-We then specify two pathways for input of inorganic carbon to the photosynthesis SU: 1) passive diffusion of CO~2~ from the external environment, and 2) active delivery of CO~2~ to the symbiont by the host.
-The passive flux ensures that some CO~2~ is always available to photosynthesis, and the active flux encompasses the potentially diverse mechanisms by which the host may enhance CO~2~ availability for the symbiont, including active transport of bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to CO~2~ to promote diffusion toward the symbiont, and acidification of the symbiosome to increase localized CO~2~ concentrations.
-Since the host physically separates the symbiont from the external environment, both the passive and active flux rates are proportional to host surface area.
+This amplification factor is then multiplied by the external downwelling irradiance $L$ and the effective light-absorbing surface area of symbiont biomass $\bar{a}^*$ to specify the total amount of light absorbed:
+
+\begin{equation} j_L =  A \cdot L \cdot \bar{a}^* \end{equation}
+
+We then specify two pathways for input of inorganic carbon to the photosynthesis SU: 1) passive diffusion of CO~2~ from the external environment ($j_{CO_2}^p$), and 2) active delivery of CO~2~ to the symbiont by the host ($j_{CO_2}^a$).
+The passive flux ensures that some CO~2~ is always available to photosynthesis, and the active flux encompasses potentially diverse mechanisms by which the host may enhance CO~2~ availability for the symbiont, including active transport of bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to CO~2~ to promote diffusion toward the symbiont [@Tansik:2015cm], and acidification of the symbiosome to increase localized CO~2~ concentrations around the symbiont [@Barott:2014gx].
+Since the host physically separates the symbiont from the external environment, both the passive and active flux rates are proportional to host biomass.
 A more rigorous, mechanistic model of inorganic carbon processing would require spatially explicit internal pools accounting for pH and carbon speciation, which is beyond the current scope of this work.
-Instead, the specification of CO~2~ delivery rates offers the user the opportunity to compare different rates of CO~2~ delivery that may characterize different coral species [@Wooldridge:2014di].
+Instead, the specification of fixed passive and active CO~2~ input rates, $j_{CO_2}^p$ and $j_{CO_2}^a$, offers the user the opportunity to compare different rates of CO~2~ delivery that may characterize different coral species [@Wooldridge:2014di].
 The input of CO~2~ to the photosynthesis SU is therefore specified as:
@@
 
-In addition to the inputs CO~2~ specified in Eqs.
-7, additional CO~2~ representing the metabolic production of CO~2~ from host and symbiont biomass turnover ($r_{CH}=\sigma_{CH}j_{HT}^0; r_{CS}=\sigma_{CS}j_{ST}^0$) is made available to the photosynthesis SU.
-fixed carbon is produced by the photosynthesis SU according to:
+Additional metabolic CO~2~ recycled from both the host and symbiont biomass turnover is also made available to the photosynthesis SU, according to:
+
+\begin{equation} r_{CH}=\sigma_{CH}j_{HT}^0 \end{equation}
+\begin{equation} r_{CS}=\sigma_{CS}j_{ST}^0 \end{equation}
+
+Fixed carbon is then produced by the photosynthesis SU according to:
 
@@
 
-where $j_{CPm}$ is the maximum specific rate of photosynthesis, and $c_{ROS}$ is the photooxidative stress multiplier (see below).
-Dividing by $c_{ROS}$ causes the rate of photosynthesis to decline in response to photo-oxidative stress, a phenomenon known as photoinhibition.
+where $j_{CPm}$ is the maximum specific rate of photosynthesis, and $c_{ROS}$ is the relative rate of reactive oxygen species production (see below).
+Dividing the photosynthetic rate by $c_{ROS}$ causes a decline in response to photooxidative stress (i.e., photoinhibition).
 
@@
 
-This excess light energy must be quenched by alternative pathways in order to prevent photo-oxidative damage.
-*Symbiodinium* may utilize a variety of pathways for non-photochemical quenching [NPQ; @Roth:2014wf], which we collect in a total capacity for NPQ as a parameter of the symbiont ($j_{NPQ}$).
-If light energy exceeds the total capacity of both carbon fixation and NPQ, then damaging reactive oxygen species (ROS) may be produced.
-We represent this as a scaled flux of ROS $c_{ROS}$, which takes a value of 1 when all light absorbed is quenched by photochemistry and NPQ, and increases as the amount of excess excitation energy increases.
+This excess light energy must be quenched by alternative pathways in order to prevent photooxidative damage [@Powles:1984tm].
+*Symbiodinium* utilize a variety of pathways for non-photochemical quenching [NPQ; @Roth:2014wf], which we collect in a relative NPQ capacity specified as a parameter of the symbiont ($k_{NPQ}$).
+NPQ capacity may also be susceptible to reduction by photoinhibition, so the total NPQ capacity is specified as:
 
-\begin{equation} c_{ROS} = 1 + \bigg[\bigg({{j_{eL} - j_{NPQ}} \over k_{ROS}}\bigg)^k\bigg]_+ \end{equation}
+\begin{equation} j_{NPQ}= k_{NPQ} \cdot c_{ROS}^{-1} \end{equation}
 
-where $j_{NPQ}$, $k_{ROS}$, and $k$ are parameters of the symbiont that determine the onset and rate of ROS production.
-Importantly, $c_{ROS}$ as specified here is not a function of absolute light absorption, but rather the amount of excess light energy $j_{eL}$ after accounting for carbon fixation and NPQ.
-A direct consequence of this formulation is that carbon-limitation of photosynthesis can lead to photo-oxidative stress, a mechanism of biological importance [@Wooldridge:2009p7807] that was not captured by previous representations of photo-oxidative stress [@Eynaud:2011tv].
-Moreover, this formulation allows functional diversity among symbiont types to be explored by changing the parameters $j_{NPQ}$, $k_{ROS}$, and $k$.
+If light energy further exceeds the total capacity of both carbon fixation and NPQ, then reactive oxygen species (ROS) are produced.
+We represent this as a relative quantity $c_{ROS}$, which takes a value of 1 when all light absorbed can be quenched by photochemistry and NPQ, and increases as the amount of excess excitation energy further increases:
 
-Carbon fixed by photosynthesis ($j_{CP}$) is then used in conjunction with nitrogen shared by the host ($\rho_N$) and a proportion of nitrogen recycled from symbiont biomass turnover ($r_{NS}=\sigma_{NS}j_{ST}^0$) to build new symbiont biomass, following the SU equation:
+\begin{equation} c_{ROS} = 1 + \bigg({{(j_{eL} - j_{NPQ})_+} \over k_{ROS}}\bigg)^k \end{equation}
+
+where $k_{NPQ}$, $k_{ROS}$, and $k$ are parameters of the symbiont that determine the onset and rate of ROS production.
+Importantly, $c_{ROS}$ as specified here is not a function of absolute light absorption, but rather the amount of excess light energy, $j_{eL}$, after accounting for carbon fixation and NPQ.
+A direct consequence of this formulation is that CO~2~-limitation of photosynthesis can lead to ROS production, an important mechanism [@Wooldridge:2009p7807] that was not captured by previous representations of photooxidative stress [@Eynaud:2011tv].
+Moreover, this formulation allows functional diversity among symbiont types to be explored by changing the parameters $k_{NPQ}$, $k_{ROS}$, and $k$.
+
+Carbon fixed by photosynthesis ($j_{CP}$) is then combined with nitrogen shared by the host ($\rho_N$) and nitrogen recycled from symbiont biomass turnover
+
+\begin{equation} r_{NS}=\sigma_{NS}j_{ST}^0 \end{equation}
+
+to build new symbiont biomass, following the SU equation:
 
@@
 
-The rejection flux of carbon from this SU represents the amount of fixed carbon produced by photosynthesis in excess of what can be used to produce symbiont biomass; this surplus $\rho_C$ is translocated to the coral host:
+The rejection flux of carbon from this SU represents the amount of fixed carbon produced by photosynthesis in excess of what can be used to produce symbiont biomass; this surplus, $\rho_C$, is translocated to the coral host:
 
@@
 
-\begin{equation} j_{ST} = j_{ST}^0(1 + 5 \cdot (c_{ROS}-1)) \end{equation}
+\begin{equation} j_{ST} = j_{ST}^0(1 + b(c_{ROS}-1)) \end{equation}
 
-This second component of symbiont biomass loss can represent both photodamage and/or symbiont expulsion (i.e., bleaching), both of which happen in response to higher levels or ROS production.
-The constant 5 is included to increase biomass loss in response to ROS.
-(Note that recycling of symbiont biomass turnover ($rNS$ and $rCS$) only occurs based on the basal maintenance related turnover (i.e., $j_{ST}^0$), and not the bleaching-related biomass loss, as this loss represents biomass being damaged or being expelled from the holobiont).
-
-## Model state equations
-Finally, the balance equations representing the specific growth rates of symbiont and host biomass over time can be expressed as:
-
-\begin{equation} {dS \over Sdt} = j_{SG} - j_{ST} \end{equation}
-
-\begin{equation} {dH \over Hdt} = j_{HG} - j_{HT}^0 \end{equation}
+This second component of symbiont biomass loss can represent both photodamage and/or symbiont expulsion (i.e., bleaching), both of which occur in response to high levels or ROS production.
+The parameter $b$ is included to scale biomass loss due to bleaching in response to ROS.
+(Note that recycling of symbiont biomass turnover ($rNS$ and $rCS$) only occurs based on the maintenance component of turnover (i.e., $j_{ST}^0$), and not the bleaching-related component, as this loss represents biomass that is damaged or expelled from the host).
 
@@
 These vectors, along with initial values of symbiont and host biomass, then serve as input to the time-stepping function, which solves for the current system state using values of the previous system state where necessary.
-A default time step of 0.1 days was used for all simulations, which were performed using R code that is available in the data repository accompanying this article: [github.com/jrcunning/Rcoral](github.com/jrcunning/Rcoral).
+A time step of 0.1 days was used for all simulations, which were performed using R code that is available in the data repository accompanying this article: [github.com/jrcunning/Rcoral](github.com/jrcunning/Rcoral).
 
-# Model behavior evaluation
+To aid in visualizing model results, we calculated values to indicate the degree to which product formation at an SU was limited by availability of either of its two substrates:
 
-To analyze qualitative behavior, we ran the model to steady state across gradients of external irradiance and DIN, and plotted the specific growth rate (Fig.
-2A) and symbiont to host biomass ratio (Fig.
+\begin{equation} \log \bigg({{\min(j_{S1}, j_{Pm})} \over {\min(j_{S2}, j_{Pm})}}\bigg) \end{equation}
+
+where $j_{S1}$ and $j_{S2}$ are the specific input fluxes of the two substrates and $j_{Pm}$ is the maximum specific product formation rate, all with units of Cmol Cmol^-1^ d^-1^.
+
+# Sensitivity analysis -- should come after the steady states bc describes changes in steady states
+
+Each default parameter value used in the model (Table 2) is based on carefully selected relevant literature.
+A supplementary document detailing the derivations of each of these default parameter values is provided.
+Nevertheless, it is important to evaluate the sensitivity of the model to changes in these parameter values.
+We measured fractional change in steady state values in response to fractional change in parameter values, relative to their default values (Table 2).
+
+![Sensitivity analysis.
+Plots show the fractional change in steady state values in response to fractional changes in default parameter values (see Table 2 for default values).
+Parameters are grouped by which processes they are involved in.
+Sensitivity in response to changes in each parameter was evaluated in each of four environments (indicated by line type): 1) LLLN=low light (2 mol ph m^-2^ d^-1^) and low DIN (1e-7 mol L^-1^), 2) LLHN=low light (2 mol ph m^-2^ d^-1^) and high DIN (4e-6 mol L^-1^), 3) HLLN=high light (30 mol ph m^-2^ d^-1^) and low DIN (1e-7 mol L^-1^), and 4) HLHN=high light (30 mol ph m^-2^ d^-1^) and high DIN (4e-6 mol L^-1^).](../img/Fig3.png)
+
+# Steady state behavior
+
+*emphasize that steady state is exponential growth, and fluxes are per unit biomass/specific rates*
+To analyze qualitative behavior, we ran the model to steady state across gradients of external irradiance and nutrients (Fig.
+2), which revealed patterns consistent with observed phenomena in corals.
+Predicted growth rates are low at low light and DIN, and begin increasing as both of these factors increase (Fig.
+2A).
+Low light is limiting to photosynthesis (Fig.
+2D), which leads to carbon-limitation of host growth (Fig.
+2B) and an associated increase in the symbiont to host biomass ratio (Fig.
+2C).
+In agreement with this trend are many observations of negative correlation between irradiance and symbiont density [@Stimson:1997p3837; @Brown:1999p3534; @Fitt:2000p3792; @Titlyanov:2001p3631].
+As higher light alleviates light-limitation of photosynthesis, growth becomes less carbon-limited.
+Similarly, elevating DIN up to ~2 µM alleviates nitrogen-limitation (Fig.
 2B).
-Results are consistent with observations in corals: specific growth rates are low at very low levels of irradiance and DIN, as these scenarios limit the availability of carbon and nitrogen, respectively, to build biomass.
-As these resources increase, specific growth increases, reaching its highest levels above 2 µM DIN and ~15 mol photons m^-2^ d^-1^.
-Peak specific growth rates are around ~0.1 d^-1^, which is consistent with values reported in the literature (@Muscatine:1981jy).
+Increased growth at higher DIN is predicted by the DEB model of @Muller:2009io, and has also been observed experimentally [@MullerParker:1994uo; @Tanaka:2007uv; @Tanaka:2013dj].
+However, DIN elevation beyond a certain point (e.g., 2-3 µM in these simulations) has little effect on growth as carbon becomes limiting (Fig.
+2B).
+Very high nutrient levels may even reduce growth [@Shantz:2015dz], although these impacts are not likely to occur within the range of concentrations considered here (<4 µM) [@FerrierPages:2000p1950].
 
-![Steady state values of (A) specific growth (Cmol Cmol^-1^ d^-1^) and (B) the symbiont to host biomass ratio (CmolS CmolH^-1^) across gradients of external irradiance and dissolved inorganic nitrogen.
-Simulations for each combination of light and nutrients (21 points along each axis) were run for 100 days with a time step of 1 day.
+In addition to increasing growth, DIN also increases the symbiont to host biomass ratio, a phenomenon also observed in reef corals [@Marubini:1996p8460].
+S:H ratios also increase under low light [@Stimson:1997p3837; @Fitt:2000p3792; @Anthony:2003p3632].
+At low DIN and intermediate light, more typical of coral reef environments, symbiont to host biomass ratios are around ~0.1-0.3, which is consistent with values reported in the literature [@Muscatine:1981jy; @Edmunds:2011bv].
+
+![Steady state values of **A)** specific growth (Cmol Cmol^-1^ d^-1^), **B)** the symbiont to host biomass ratio (CmolS CmolH^-1^), **C)** relative C- or N-limitation of host biomass formation, and **D)** relative CO~2~- or light-limitation of symbiont photosynthesis, across gradients of external irradiance and dissolved inorganic nitrogen.
+Note that typical conditions for reefs are ~1e-7 M DIN and 10-20 mol photons m^-2^ d^-1^.
+Simulations for each combination of light and nutrients (41 points along each axis) were run for 100 days with a time step of 1 day.
 Negative steady state growth rates, and corresponding S:H ratios, were set to zero.](../img/Fig2.png)
 
-# Discussion  -  Potential applications/utility
+The maximum predicted growth rates of ~0.1 d^-1^, occurring between ~10-15 mol photons m^-2^ s^-1^ light and >~3 µM DIN (Fig.
+2A), are comparable to the rate of `r Tanaka_gr` d^-1^ measured by @Tanaka:2007uv under similar N-enriched conditions.
+Growth under conditions more typical of reef environments (<0.5 µM DIN) are within the range of ~0.01-0.05 d^-1^; however, these values are still higher than most specific growth rates reported in the literature, which are typically based on skeletal mass, and fall near or below 0.01 d^-1^ [@Osinga:2011kq; @Osinga:2012is] (though values as high as 0.025 d^-1^ have been reported [@Schutter:2010p7758]).
+These generally low skeletal growth rates relative to predicted biomass growth may be explained by differential substrate-limitation (i.e., CO~2~ may limit calcification, which is not present in the model); indeed, tissue and skeletal growth may not always be well correlated [@Anthony:2002tc].
+Furthermore, bioerosion may reduce net skeletal growth, especially in the field [@Bak:1976bv].
+Nevertheless, when rates of increase in coral tissue biomass have been directly measured over short periods, values are consistent with model predictions (e.g., `r Tanaka_gr` d^-1^ in @Tanaka:2007uv), and similar rates (`r Armoza_gr` d^-1^) have been measured in *Aiptasia diaphana*, a non-calcifying symbiotic anemone [@ArmozaZvuloni:2014ju].
+
+As irradiance continues to increase above ~15-20 mol photons m^-2^ d^-1^, growth rates decline until positive growth ceases above ~30 µmol photons m^-2^ d^-1^.
+The mechanism underlying this decline is the increase in light energy beyond the capacity of photosynthesis and non-photochemical quenching: excess excitation energy generates reactive oxygen species (ROS) [@Weis:2008p944; @Roth:2014wf], which, in this model, have the phenomenological consequences of reducing the photosynthetic rate (representing photoinhibition) and increasing the symbiont biomass loss rate (representing photodamage and/or symbiont expulsion) [see @Eynaud:2011tv].
+Together, these impacts reduce the symbiont to host biomass ratio (Fig.
+2C), as occurs during coral bleaching.
+This reduction in symbionts consequently reduces the flux of fixed carbon to the host, resulting in increasing carbon-limitation (Fig.
+2B) and eventual cessation of growth (Fig.
+2A).
+Importantly, these consequences arise in the present model only when light energy exceeds the capacity of photochemistry and NPQ, which in turn depends on CO~2~ availability.
+While previous models framed photophysiological stress as a fixed response to absolute irradiance levels [@Eynaud:2011tv], the current implementation considers the dynamic balance of multiple energy sinks in the causation of stress, which is more consistent with current understanding of symbiosis dysfunction [@Wooldridge:2013tj], and establishes an important role of the host in providing CO~2~ for symbiont photosynthesis [@Tansik:2015cm; @Hopkinson:2015ev].
+
+While bleaching in response to high light alone has been observed experimentally [@Schutter:2011ie; @Downs:2013kc], mass coral bleaching events occur in response to high temperature [@HoeghGuldberg:1999p4222]; thus, it is important to justify our consideration of light as the primary stressor in the model.
+In reality, light and temperature interact synergistically [@Coles:1978p1124; @Jones:1998p3572], and in fact, any stressor that disrupts the quenching of light energy may lead to bleaching [@Wooldridge:2010p7809; @Baker:2015kp].
+This is because the proximate cause of coral bleaching is excess excitation energy, but the upstream events that lead to this situation may be diverse: indeed, elevated temperature inhibits Rubisco functioning [@Jones:1998p3572] and repair of the D1 protein in photosystem II [@Warner:1999p4239], which reduces the capacity of photochemical quenching and leads to an excess of light energy.
+In this way, elevated temperature serves to reduce the threshold above which light causes stress to the system [@HoeghGuldberg:1999p4222]; importantly, light is still the proximate stressor.
+Therefore, we omitted temperature from the model, which serves to maintain a desired level of simplicity, while still allowing photophysiological stress and bleaching to be simulated with biological realism in response to light.
+Thus, the specific light levels that cause stress here should be interpreted in this context, and the effects of high light are analagous to the effects of high temperature.
+
+The incorporation of light stress in the model sets an upper limit to the amount of light at which a stable symbiotic interaction can be maintained, but even below this threshold of breakdown, negative effects of high light reduce steady state growth and symbiont:host biomass (Fig.
+2A,C).
+This gradual decline is consistent with experimental results showing that high light levels decrease growth [@Schutter:2011ie], and field studies documenting optimum growth rates at intermediate depths [@Baker:1975ip; @Huston:1985ec].
+Consequently, the model predicts greater variation in state variables across light gradients than was predicted by the model of @Muller:2009io, which did not include photoinhibition and photodamage.
+It is important to recognize that these are steady state values predicted under a constant environment at the given light and nutrient levels; a dynamic system may temporarily cross this threshold and experience a period of symbiont loss (bleaching) and reduced growth, after which a return to benign conditions may restore symbiont biomass and positive growth.
+To explore this further and illustrate the behavior of the model in more detail, we evaluate a number of dynamic simulations below.
+
+# Dynamic behavior
+## varying light: consequences of excess light and light-limitation
+## varying DIN: role of N-limitation and consequences of elevated N
+## elevated nutrients cause system to become C-limited (photo: CO2 limited; host: fixed-C limited)
+## C-limitation increases susceptibility to bleaching
+## consequences of a less efficient but more tolerant symbiont
+## Rate of symbiont repopulation (Berner et al. 1993)
+## Bicarbonate addition promotes coral growth (Marubini and Thake 1999)
+## Seasonal change in sym. density is larger when nutrients are elevated (Stimson 1997). Such a phenomenon was predicted by Fitt et al. 2000.
+## Death rule: negative growth for a certain amount of time should be defensible.
 

### 2016-10-06 JTB_ms/JTB_ms.tex
updates from UCSB

@@
 jeopardizing their future persistence. In order to understand and
-predict coral responses to complex changes in the environment, a
+predict coral responses to complex changes in their environment, a
 mechanistic understanding of how multiple interacting factors drive the
@@
 integrate the impacts of complex environmental change on the
-physiological ecology of reef corals.
+physiological performance of reef corals, including responses to
+environmental stress.
 
 In reef coral symbioses, intracellular \emph{Symbiodinium} translocate
-photosynthetically-fixed carbon to support coral metabolism, and utilize
-the animal's metabolic waste products, including nitrogenous compounds
-and carbon dioxide, in return (Muscatine and Porter 1977). Previous
-application of DEB theory to this syntrophic system (Muller et al. 2009)
-demonstrated a stable symbiotic relationship and qualitatively realistic
-growth and biomass ratios across gradients of ambient irradiance,
-nutrients, and food. This model assumed that 1) \emph{Symbiodinium} has
-priority access to carbon through photosynthesis, 2) the coral animal
-has priority access to dissolved nitrogen through contact with seawater,
-and 3) each partner shares with the other only what it cannot use
-itself. In its simplest form, this principle of sharing the surplus
-sufficiently describes diverse syntrophic interactions among organs and
-organisms (e.g., trees, duckweeds, corals), suggesting the mechanism is
-mathematically and evolutionarily robust (Nisbet et al., submitted).
-
-While the work of Muller et al. (2009) applying DEB formalism to this
-system represents the most significant theoretical contribution in coral
-symbiosis research to date, we aim to strengthen the role of theory and
-broaden its potential application to corals in three primary ways:
+photosynthetically fixed carbon to support coral metabolism, while the
+animal host provides access to inorganic nutrients and carbon dioxide
+(Muscatine and Porter 1977). Previous application of DEB theory to this
+syntrophic system (Muller et al. 2009) demonstrated a stable symbiotic
+relationship and qualitatively realistic growth and biomass ratios
+across gradients of ambient irradiance, nutrients, and food. This model
+assumed that 1) \emph{Symbiodinium} has priority access to fixed carbon
+through photosynthesis, 2) the coral animal has priority access to
+dissolved nitrogen through contact with seawater, and 3) each partner
+shares with the other only what it cannot use for its own growth. In its
+simplest form, this principle of sharing the surplus is sufficient to
+describe the dynamics of diverse syntrophic organs and organisms (e.g.,
+trees, duckweeds, corals), suggesting the mechanism is mathematically
+and evolutionarily robust (Nisbet et al., in prep.).
+
+While the formal DEB model of Muller et al. (2009) represents the most
+significant theoretical contribution in coral symbiosis research to
+date, we aim to strengthen the role of theory and broaden its potential
+application in three primary ways:
 
@@
   formulation restricts the model's scope to the bioenergetics of growth
-  and symbiosis dynamics in adult corals, but greatly reduces
+  and symbiosis dynamics in adult corals (i.e., reproduction, larval
+  stages, and metamorphosis are not considered), but greatly reduces
   theoretical complexity and parameter numbers, which is advantageous
@@
   scientific community to include additional complexity and
-  problem-specific components. The R language was chosen because it is
-  freely available and in common use by biologists and ecologists, to
-  widen the audience for this work.
+  problem-specific components. To widen the audience for this work, we
+  chose the R language because it is freely available and in common use
+  by biologists and ecologists.
 \end{enumerate}
@@
 
-In this model of coral-algal symbiosis, carbon and nitrogen are acquired
-by each partner and used to construct biomass. A graphical
-representation of the model is presented in Fig. 1. We use C-moles as
-the unit of biomass for consistency with the rigorous mass balance of
-DEB theory: 1 C-mole is equivalent to the amount of biomass containing 1
-mole of Carbon atoms. Host biomass (\(H\)), symbiont biomass (\(S\)),
-and prey biomass (\(X\)) have fixed, but different, molar N:C ratios
-(Table 1). Carbon and nitrogen are combined to produce biomass by
-synthesizing units (SU), which are mathematical specifications of the
-formation of a product from two substrates; we use the ``parallel
-complementary'' formulation of Kooijman (2010) to specify these fluxes.
-The two state variables in this dynamical system are symbiont biomass
-and coral biomass; because resources are acquired proportionally to
-surface area (and surface area is assumed proportional to volume for
-corals (i.e., they are ``V1-morphs'' in DEB terminology (Kooijman
-2010))), biomass increases exponentially during growth. The rate of
-increase in coral biomass (i.e.~growth) and the ratio of symbiont to
-host biomass are the (i.e.~symbiosis dynamics) are the responses of
-interest of the system. Below we describe the formulation of each flux
-involved in producing these responses.
-
-Table 1
+In this dynamic bioenergetic model of coral-algal symbiosis, carbon and
+nitrogen are acquired by each partner and used to construct biomass. A
+graphical representation of the model is presented in Fig. 1, and each
+model flux and parameter are defined in Tables 1 and 2, respectively. We
+use C-moles as the unit of biomass for consistency with the rigorous
+mass balance of DEB theory: 1 C-mole is equivalent to the amount of
+biomass containing 1 mole of Carbon atoms. Host biomass (\(H\)),
+symbiont biomass (\(S\)), and prey biomass (\(X\)) have fixed, but
+different, molar N:C ratios (Table 2). Carbon and nitrogen are combined
+to produce biomass by synthesizing units (SU), which are mathematical
+specifications of the formation of a product from two substrates; we use
+the parallel complementary formulation of Kooijman (2010) to specify
+these fluxes. The two state variables are symbiont biomass and coral
+biomass; because resources are acquired proportionally to surface area,
+and surface area is proportional to volume (i.e., corals are
+``V1-morphs'' in DEB terminology (Kooijman 2010)), biomass increases
+exponentially during growth (indeed, corals grow exponentially (Bak
+1976)). The specific growth rate and the ratio of symbiont to host
+biomass are the responses of interest of the system. Below we describe
+the formulation of each flux involved in producing these responses.
+
+Table 1. Model fluxes
+
+\begin{longtable}[c]{@{}llll@{}}
+\toprule
+\begin{minipage}[b]{0.12\columnwidth}\raggedright\strut
+Symbol
+\strut\end{minipage} &
+\begin{minipage}[b]{0.48\columnwidth}\raggedright\strut
+Description
+\strut\end{minipage} &
+\begin{minipage}[b]{0.26\columnwidth}\raggedright\strut
+Units
+\strut\end{minipage} &
+\begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
+Eq. no.
+\strut\end{minipage}\tabularnewline
+\midrule
+\endhead
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(j_X\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Prey uptake rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+molX CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+3
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(j_N\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Nitrogen uptake rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+molN CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+4
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(j_{HG}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Host biomass formation rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+CmolH CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+5
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(r_{NH}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Recycled nitrogen from host turnover
+\strut\end{minipage} &
+\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+molN CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+6
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(\rho_N\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Nitrogen shared with the symbiont
+\strut\end{minipage} &
+\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+molN CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+7
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(j_L\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Light absorption rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+molph CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+9
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(j_{CO_2}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+CO\textsubscript{2} input to photosynthesis
+\strut\end{minipage} &
+\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+molCO\textsubscript{2} CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+10
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(r_{CH}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Recycled CO\textsubscript{2} from host turnover
+\strut\end{minipage} &
+\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+molCO\textsubscript{2} CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+11
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(r_{CS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Recycled CO\textsubscript{2} from symbiont turnover
+\strut\end{minipage} &
+\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+molCO\textsubscript{2} CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+12
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(j_{CP}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Photosynthesis rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+molC CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+13
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(j_{eL}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Light energy in excess of photochemistry
+\strut\end{minipage} &
+\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+molph CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+14
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(j_{NPQ}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Total capacity of NPQ
+\strut\end{minipage} &
+\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+molph CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+15
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(c_{ROS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+ROS production proportional to baseline
+\strut\end{minipage} &
+\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+--
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+16
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(r_{NS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Recycled nitrogen from symbiont turnover
+\strut\end{minipage} &
+\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+molN CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+17
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(j_{SG}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Symbiont biomass formation rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+18
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(\rho_C\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Fixed carbon shared with host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+molC CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+19
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(j_{ST}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Symbiont turnover rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+20
+\strut\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
+
+Table 2. Model parameters
 
@@
 \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-d\textsuperscript{-1}
+CmolH CmolH\textsuperscript{-1} d\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-d\textsuperscript{-1}
+CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{NPQ}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Non-photochemical quenching capacity of symbiont
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-40
+\(k_{NPQ}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+NPQ capacity of symbiont, relative to C-fixation
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+4
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+--
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(k_{ROS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Excess photon energy that doubles ROS production, relative to baseline
+levels
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+40-80
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(k_{ROS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Excess photon energy that doubles ROS prod.
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-40-80
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-mol ph CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 \(k\)
@@
 \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-d\textsuperscript{-1}
+CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(b\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Scaling parameter for strength of bleaching response
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+3
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+--
 \strut\end{minipage}\tabularnewline
@@
 
-Also include tables of fluxes and environmental inputs?
-
 \begin{figure}[htbp]
 \centering
-\includegraphics{../img/model.png}
-\caption{Graphical representation of coral-algal symbiosis model.}
+\includegraphics{../img/Fig1.png}
+\caption{Graphical representation of coral-algal symbiosis model. Light,
+CO\textsubscript{2}, prey, and DIN are acquired from the external
+environment proportional to the biomass of the partner indicated by the
+black box for uptake. Mass fluxes (see Table 1 for definitions) are
+represented by \(j\)'s with subscripts indicating the type of mass, and
+in some cases the process (e.g., \(j_{CP}\) is the flux of carbon
+produced by photosynthesis), and \(\rho\)'s indicate fluxes that are
+shared by one partner with the other. Parallel complementary
+synthesizing units (SUs) are represented by large circles, and rejection
+fluxes from these SUs are indicated by dashed lines. \(c_{ROS}\) is a
+proportional rate that impacts other model fluxes by inhibition or
+acceleration. Recycled mass fluxes from biomass turnover are not shown
+for clarity (but see Table 1 for definitions).}
 \end{figure}
+
+\subsection{State equations}\label{state-equations}
+
+The balance equations for symbiont and host biomass are expressed as:
+
+\begin{equation} {dS \over Sdt} = j_{SG} - j_{ST} \end{equation}
+
+\begin{equation} {dH \over Hdt} = j_{HG} - j_{HT}^0 \end{equation}
+
+These are expressed per unit of symbiont and host biomass, respectively.
+The specific biomass growth and turnover rates that define these balance
+equations are produced by combinations of the individual model fluxes
+(see Table 1 for definitions and units), which are each expressed as
+mass-specific rates (e.g., per C-mole of symbiont or host biomass). When
+necessary, conversions between symbiont-mass-specific and
+host-mass-specific rates are accomplished by multiplying or dividing by
+the symbiont:host biomass ratio.
 
@@
 from the environment. Prey acquisition is specified by Michaelis-Menten
-kinetics using a maximum area-specific feeding rate and half-saturation
-constant:
+kinetics (i.e., a Holling type II function) with a maximum feeding rate
+\(j_{Xm}\) and half-saturation constant \(K_X\):
 
@@
 
-Additionally, the coral animal can acquire nitrogen from the surrounding
-seawater. This nitrogen source is assumed to represent ammonium, the
-primary form utilized by corals (Wang and Douglas 1998; Yellowlees,
-Rees, and Leggat 2008). The uptake of nitrogen from the environment is
-specified by Michaelis-Menten kinetics using a maximum area-specific
-uptake rate and half-saturation constant:
+Additionally, the coral animal acquires nitrogen dissolved in the
+surrounding seawater, which is assumed to represent ammonium, the
+primary form utilized by corals (Yellowlees, Rees, and Leggat 2008).
+This gives the host (rather than the symbiont) priority in nitrogen
+utilization; this capacity is supported by experimental evidence (Wang
+and Douglas 1998) and is consistent with the physical arrangment of the
+partners, where the host is in direct contact with the external
+environment. The uptake of nitrogen from the environment is thus
+specified by Michaelis-Menten kinetics using a maximum uptake rate
+\(j_{Nm}\) and half-saturation constant \(K_N\):
 
@@
 
-Coral biomass formation is specified by a parallel complementary SU that
-combines carbon and nitrogen to form biomass. In addition to the carbon
-and nitrogen acquired by direct uptake (Eqs. 1 and 2), the coral
-recycles a portion of the nitrogen liberated by biomass turnover
-(\(r_{NH}=\sigma_{NH}j_{HT}^0\)), and receives surplus fixed carbon
-shared by the symbiont (\(\rho_C\)), such that the total biomass
-formation is specified as:
+Coral biomass formation is then specified by a parallel complementary SU
+that combines carbon and nitrogen, according to:
 
 \begin{equation} j_{HG} = \bigg({1 \over j_{HGm}} + {1 \over {\rho_C{S \over H} + j_X}} + {1 \over {(j_N + n_{NX}j_X + r_{NH}) / n_{NH}}} - {1 \over {{\rho_C{S \over H} + j_X} + (j_N + n_{NX}j_X + r_{NH}) / n_{NH}}}\bigg)^{-1} \end{equation}
+
+where \(\rho_C\) is the surplus fixed carbon shared by the symbiont (see
+below), and \(r_{NH}\) is the recycled portion of the nitrogen liberated
+by host biomass turnover:
+
+\begin{equation} r_{NH}=\sigma_{NH}j_{HT}^0 \end{equation}
 
@@
 is actually consumed in biomass formation (i.e., surplus nitrogen, or
-the ``rejection flux'' in SU terminology) is then made available to the
-symbiont, and is specified as:
+the rejection flux\footnote{Rejection fluxes must always be positive,
+  and hence are specified with the notation \((x)_+\), which means
+  \(max(x, 0)\).} of the SU) is then made available to the symbiont:
 
@@
 Due to the inherent inefficiency of the parallel complementary SU
-formulation, there will always be some nitrogen shared with the symbiont
-even when coral biomass formation is strongly nitrogen-limited.
-Likewise, there is always a non-zero rejection flux of carbon from the
-coral biomass SU, which is assumed to be lost to the environment.
+formulation, there is always some nitrogen shared with the symbiont even
+when coral biomass formation is strongly nitrogen-limited. Likewise,
+there is always a non-zero rejection flux of carbon from the coral
+biomass SU, which is assumed to be lost to the environment.
 
@@
 and Iglesias-Prieto 2005; Marcelino et al. 2013). We used data from
-Marcelino et al. (2013) to empirically derive the ratio of internal
-scalar irradiance to external downwelling irradiance as a function of
-symbiont density (expressed as symbiont to host biomass ratio), and
-subsequently multiply this quantity by the external downwelling
+Marcelino et al. (2013) to empirically derive an ``amplification
+factor'' \(A\) (the ratio of internal scalar irradiance to external
+downwelling irradiance) as a function of symbiont density (expressed as
+symbiont to host biomass ratio), which is specified as:
+
+\begin{equation} A = 1.26 + 1.39 \cdot \exp(-6.48 \cdot {S \over H}) \end{equation}
+
+This amplification factor is then multiplied by the external downwelling
 irradiance \(L\) and the effective light-absorbing surface area of
-symbiont biomass \(ds\) to specify the amount of light absorbed:
-
-\begin{equation} j_L = \big[1.26 + 1.39 \cdot \exp(-6.48 \cdot {S \over H})\big] \cdot L \cdot \bar{a}^* \end{equation}
+symbiont biomass \(\bar{a}^*\) to specify the total amount of light
+absorbed:
+
+\begin{equation} j_L =  A \cdot L \cdot \bar{a}^* \end{equation}
 
@@
 photosynthesis SU: 1) passive diffusion of CO\textsubscript{2} from the
-external environment, and 2) active delivery of CO\textsubscript{2} to
-the symbiont by the host. The passive flux ensures that some
-CO\textsubscript{2} is always available to photosynthesis, and the
-active flux encompasses the potentially diverse mechanisms by which the
-host may enhance CO\textsubscript{2} availability for the symbiont,
-including active transport of bicarbonate, carbonic anhydrase-catalyzed
-conversion of bicarbonate to CO\textsubscript{2} to promote diffusion
-toward the symbiont, and acidification of the symbiosome to increase
-localized CO\textsubscript{2} concentrations. Since the host physically
-separates the symbiont from the external environment, both the passive
-and active flux rates are proportional to host surface area. A more
-rigorous, mechanistic model of inorganic carbon processing would require
-spatially explicit internal pools accounting for pH and carbon
-speciation, which is beyond the current scope of this work. Instead, the
-specification of CO\textsubscript{2} delivery rates offers the user the
-opportunity to compare different rates of CO\textsubscript{2} delivery
-that may characterize different coral species (Wooldridge 2014a). The
-input of CO\textsubscript{2} to the photosynthesis SU is therefore
-specified as:
+external environment (\(j_{CO_2}^p\)), and 2) active delivery of
+CO\textsubscript{2} to the symbiont by the host (\(j_{CO_2}^a\)). The
+passive flux ensures that some CO\textsubscript{2} is always available
+to photosynthesis, and the active flux encompasses potentially diverse
+mechanisms by which the host may enhance CO\textsubscript{2}
+availability for the symbiont, including active transport of
+bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to
+CO\textsubscript{2} to promote diffusion toward the symbiont (Tansik,
+Fitt, and Hopkinson 2015), and acidification of the symbiosome to
+increase localized CO\textsubscript{2} concentrations around the
+symbiont (Barott et al. 2014). Since the host physically separates the
+symbiont from the external environment, both the passive and active flux
+rates are proportional to host biomass. A more rigorous, mechanistic
+model of inorganic carbon processing would require spatially explicit
+internal pools accounting for pH and carbon speciation, which is beyond
+the current scope of this work. Instead, the specification of fixed
+passive and active CO\textsubscript{2} input rates, \(j_{CO_2}^p\) and
+\(j_{CO_2}^a\), offers the user the opportunity to compare different
+rates of CO\textsubscript{2} delivery that may characterize different
+coral species (Wooldridge 2014a). The input of CO\textsubscript{2} to
+the photosynthesis SU is therefore specified as:
 
@@
 
-In addition to the inputs CO\textsubscript{2} specified in Eqs. 7,
-additional CO\textsubscript{2} representing the metabolic production of
-CO\textsubscript{2} from host and symbiont biomass turnover
-(\(r_{CH}=\sigma_{CH}j_{HT}^0; r_{CS}=\sigma_{CS}j_{ST}^0\)) is made
-available to the photosynthesis SU. fixed carbon is produced by the
-photosynthesis SU according to:
+Additional metabolic CO\textsubscript{2} recycled from both the host and
+symbiont biomass turnover is also made available to the photosynthesis
+SU, according to:
+
+\begin{equation} r_{CH}=\sigma_{CH}j_{HT}^0 \end{equation}\begin{equation} r_{CS}=\sigma_{CS}j_{ST}^0 \end{equation}
+
+Fixed carbon is then produced by the photosynthesis SU according to:
 
@@
 where \(j_{CPm}\) is the maximum specific rate of photosynthesis, and
-\(c_{ROS}\) is the photooxidative stress multiplier (see below).
-Dividing by \(c_{ROS}\) causes the rate of photosynthesis to decline in
-response to photo-oxidative stress, a phenomenon known as
-photoinhibition.
+\(c_{ROS}\) is the relative rate of reactive oxygen species production
+(see below). Dividing the photosynthetic rate by \(c_{ROS}\) causes a
+decline in response to photooxidative stress (i.e., photoinhibition).
 
@@
 This excess light energy must be quenched by alternative pathways in
-order to prevent photo-oxidative damage. \emph{Symbiodinium} may utilize
-a variety of pathways for non-photochemical quenching (NPQ; Roth 2014),
-which we collect in a total capacity for NPQ as a parameter of the
-symbiont (\(j_{NPQ}\)). If light energy exceeds the total capacity of
-both carbon fixation and NPQ, then damaging reactive oxygen species
-(ROS) may be produced. We represent this as a scaled flux of ROS
-\(c_{ROS}\), which takes a value of 1 when all light absorbed is
-quenched by photochemistry and NPQ, and increases as the amount of
-excess excitation energy increases.
-
-\begin{equation} c_{ROS} = 1 + \bigg[\bigg({{j_{eL} - j_{NPQ}} \over k_{ROS}}\bigg)^k\bigg]_+ \end{equation}
-
-where \(j_{NPQ}\), \(k_{ROS}\), and \(k\) are parameters of the symbiont
+order to prevent photooxidative damage (Powles 1984).
+\emph{Symbiodinium} utilize a variety of pathways for non-photochemical
+quenching (NPQ; Roth 2014), which we collect in a relative NPQ capacity
+specified as a parameter of the symbiont (\(k_{NPQ}\)). NPQ capacity may
+also be susceptible to reduction by photoinhibition, so the total NPQ
+capacity is specified as:
+
+\begin{equation} j_{NPQ}= k_{NPQ} \cdot c_{ROS}^{-1} \end{equation}
+
+If light energy further exceeds the total capacity of both carbon
+fixation and NPQ, then reactive oxygen species (ROS) are produced. We
+represent this as a relative quantity \(c_{ROS}\), which takes a value
+of 1 when all light absorbed can be quenched by photochemistry and NPQ,
+and increases as the amount of excess excitation energy further
+increases:
+
+\begin{equation} c_{ROS} = 1 + \bigg({{(j_{eL} - j_{NPQ})_+} \over k_{ROS}}\bigg)^k \end{equation}
+
+where \(k_{NPQ}\), \(k_{ROS}\), and \(k\) are parameters of the symbiont
 that determine the onset and rate of ROS production. Importantly,
 \(c_{ROS}\) as specified here is not a function of absolute light
-absorption, but rather the amount of excess light energy \(j_{eL}\)
+absorption, but rather the amount of excess light energy, \(j_{eL}\),
 after accounting for carbon fixation and NPQ. A direct consequence of
-this formulation is that carbon-limitation of photosynthesis can lead to
-photo-oxidative stress, a mechanism of biological importance (Wooldridge
-2009) that was not captured by previous representations of
-photo-oxidative stress (Eynaud, Nisbet, and Muller 2011). Moreover, this
+this formulation is that CO\textsubscript{2}-limitation of
+photosynthesis can lead to ROS production, an important mechanism
+(Wooldridge 2009) that was not captured by previous representations of
+photooxidative stress (Eynaud, Nisbet, and Muller 2011). Moreover, this
 formulation allows functional diversity among symbiont types to be
-explored by changing the parameters \(j_{NPQ}\), \(k_{ROS}\), and \(k\).
-
-Carbon fixed by photosynthesis (\(j_{CP}\)) is then used in conjunction
-with nitrogen shared by the host (\(\rho_N\)) and a proportion of
-nitrogen recycled from symbiont biomass turnover
-(\(r_{NS}=\sigma_{NS}j_{ST}^0\)) to build new symbiont biomass,
-following the SU equation:
+explored by changing the parameters \(k_{NPQ}\), \(k_{ROS}\), and \(k\).
+
+Carbon fixed by photosynthesis (\(j_{CP}\)) is then combined with
+nitrogen shared by the host (\(\rho_N\)) and nitrogen recycled from
+symbiont biomass turnover
+
+\begin{equation} r_{NS}=\sigma_{NS}j_{ST}^0 \end{equation}
+
+to build new symbiont biomass, following the SU equation:
 
@@
 carbon produced by photosynthesis in excess of what can be used to
-produce symbiont biomass; this surplus \(\rho_C\) is translocated to the
-coral host:
+produce symbiont biomass; this surplus, \(\rho_C\), is translocated to
+the coral host:
 
@@
 
-\begin{equation} j_{ST} = j_{ST}^0(1 + 5 \cdot (c_{ROS}-1)) \end{equation}
+\begin{equation} j_{ST} = j_{ST}^0(1 + b(c_{ROS}-1)) \end{equation}
 
@@
 photodamage and/or symbiont expulsion (i.e., bleaching), both of which
-happen in response to higher levels or ROS production. The constant 5 is
-included to increase biomass loss in response to ROS. (Note that
-recycling of symbiont biomass turnover (\(rNS\) and \(rCS\)) only occurs
-based on the basal maintenance related turnover (i.e., \(j_{ST}^0\)),
-and not the bleaching-related biomass loss, as this loss represents
-biomass being damaged or being expelled from the holobiont).
-
-\subsection{Model state equations}\label{model-state-equations}
-
-Finally, the balance equations representing the specific growth rates of
-symbiont and host biomass over time can be expressed as:
-
-\begin{equation} {dS \over Sdt} = j_{SG} - j_{ST} \end{equation}
-
-\begin{equation} {dH \over Hdt} = j_{HG} - j_{HT}^0 \end{equation}
+occur in response to high levels or ROS production. The parameter \(b\)
+is included to scale biomass loss due to bleaching in response to ROS.
+(Note that recycling of symbiont biomass turnover (\(rNS\) and \(rCS\))
+only occurs based on the maintenance component of turnover (i.e.,
+\(j_{ST}^0\)), and not the bleaching-related component, as this loss
+represents biomass that is damaged or expelled from the host).
 
@@
 time-stepping function, which solves for the current system state using
-values of the previous system state where necessary. A default time step
-of 0.1 days was used for all simulations, which were performed using R
-code that is available in the data repository accompanying this article:
+values of the previous system state where necessary. A time step of 0.1
+days was used for all simulations, which were performed using R code
+that is available in the data repository accompanying this article:
 \url{github.com/jrcunning/Rcoral}.
@@
 the degree to which product formation at an SU was limited by
-availability of either of its two substrates
+availability of either of its two substrates:
 
@@
 
-\section{Model behavior evaluation}\label{model-behavior-evaluation}
+\section{Sensitivity analysis}\label{sensitivity-analysis}
+
+Each default parameter value used in the model (Table 2) is based on
+carefully selected relevant literature. A supplementary document
+detailing the derivations of each of these default parameter values is
+provided. Nevertheless, it is important to evaluate the sensitivity of
+the model to changes in these parameter values. We measured fractional
+change in steady state values in response to fractional change in
+parameter values, relative to their default values (Table 2).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{../img/Fig3.png}
+\caption{Sensitivity analysis. Plots show the fractional change in
+steady state values in response to fractional changes in default
+parameter values (see Table 2 for default values). Parameters are
+grouped by which processes they are involved in. Sensitivity in response
+to changes in each parameter was evaluated in each of four environments
+(indicated by line type): 1) LLLN=low light (2 mol ph
+m\textsuperscript{-2} d\textsuperscript{-1}) and low DIN (1e-7 mol
+L\textsuperscript{-1}), 2) LLHN=low light (2 mol ph
+m\textsuperscript{-2} d\textsuperscript{-1}) and high DIN (4e-6 mol
+L\textsuperscript{-1}), 3) HLLN=high light (30 mol ph
+m\textsuperscript{-2} d\textsuperscript{-1}) and low DIN (1e-7 mol
+L\textsuperscript{-1}), and 4) HLHN=high light (30 mol ph
+m\textsuperscript{-2} d\textsuperscript{-1}) and high DIN (4e-6 mol
+L\textsuperscript{-1}).}
+\end{figure}
+
+\section{Steady state behavior}\label{steady-state-behavior}
 
@@
 gradients of external irradiance and nutrients (Fig. 2), which revealed
-patterns that are consistent with observed phenomena in corals.
-
-\emph{Specific growth rates} are low at low light and DIN
-(\textasciitilde{}0.01-0.03 d\textsuperscript{-1}), and begin increasing
-as both of these factors increase (Fig. 2A). Higher light alleviates
-light-limitation of photosynthesis (Fig. 2D), and consequently
-carbon-limitation of host growth (Fig. 2B). Similarly, elevated DIN
-alleviates nitrogen-limitation of host growth (Fig. 2B), a trend that
-was also predicted by the DEB model of Muller et al. (2009), and has
-been observed experimentally (Muller-Parker et al. 1994; Tanaka et al.
-2013). Very high nutrient levels may have negative direct and indirect
-impacts on growth (\textbf{???}), though these are not likely to occur
-until beyond the range of concentrations considered here (Ferrier-Pagès
-et al. 2000). The maximum specific growth rates of \textasciitilde{}0.09
-d\textsuperscript{-1}, occurring between \textasciitilde{}10-15 mol
-photons m\textsuperscript{-2} s\textsuperscript{-1} light and
-\textgreater{}1 µM DIN, are an order of magnitude higher than those
-typically observed for corals (\textbf{???}). The primary reasons for
-this overestimation of host growth are 1) the model does not account for
-calcification, and 2) it does not account for size-dependent reductions
-in growth that likely occur as individual polyps and/or colonies grow
-(Sebens1980?, probably better refs)\ldots{} (i.e., the system continues
-growing exponentially in this model). However, the estimated specific
-growth rates are consistent with rates of biomass increase in
-\emph{Aiptasia}, measured at 0.04 d\textsuperscript{-1} (\textbf{???}),
-suggesting they are biologically realistic. They also match the values
-predicted by the full DEB model of (\textbf{???}). Furthermore, the
-relative differences across environmental gradients are likely to be the
-same.
-
-As irradiance continues to increase above \textasciitilde{}15-20 mol
-photons m\textsuperscript{-2} d\textsuperscript{-1}, growth rates
-decline until the symbiosis is no longer viable above
-\textasciitilde{}30 µmol photons m\textsuperscript{-2}
-d\textsuperscript{-1}. The mechanism underlying this breakdown is the
-production of ROS caused by excitation energy in excess of what can be
-quenched by symbiont photosynthesis and non-photochemical quenching. ROS
-production has the consequence of increasing symbiont biomass loss
-(photodamage) and decreasing photosynthesis rates (photoinhibition),
-which have the net effect of reducing fixed carbon input to the system
-below levels necessary to sustain the symbiosis.
-
-\emph{Symbiont to host biomass ratios} at low DIN (e.g. \textless{}1 µM)
-and high light (e.g., \textasciitilde{}20 mol ph m\textsuperscript{-2}
-d\textsuperscript{-2}) typical of coral reef environments are around
-\textasciitilde{}0.1-0.3, which are consistent with values reported in
-the literature (Muscatine, R McCloskey, and E Marian (1981); Edmunds et
-al. (2011)). S:H ratios increase under higher nutrients () and lower
-light ().
-
-\_
+patterns consistent with observed phenomena in corals. Predicted growth
+rates are low at low light and DIN, and begin increasing as both of
+these factors increase (Fig. 2A). Low light is limiting to
+photosynthesis (Fig. 2D), which leads to carbon-limitation of host
+growth (Fig. 2B) and an associated increase in the symbiont to host
+biomass ratio (Fig. 2C). In agreement with this trend are many
+observations of negative correlation between irradiance and symbiont
+density (Stimson 1997; Brown et al. 1999; Fitt et al. 2000; Titlyanov et
+al. 2001). As higher light alleviates light-limitation of
+photosynthesis, growth becomes less carbon-limited. Similarly, elevating
+DIN up to \textasciitilde{}2 µM alleviates nitrogen-limitation (Fig.
+2B). Increased growth at higher DIN is predicted by the DEB model of
+Muller et al. (2009), and has also been observed experimentally
+(Muller-Parker et al. 1994; Tanaka et al. 2007; Tanaka et al. 2013).
+However, DIN elevation beyond a certain point (e.g., 2-3 µM in these
+simulations) has little effect on growth as carbon becomes limiting
+(Fig. 2B). Very high nutrient levels may even reduce growth (Shantz,
+Lemoine, and Burkepile 2015), although these impacts are not likely to
+occur within the range of concentrations considered here (\textless{}4
+µM) (Ferrier-Pagès et al. 2000).
+
+In addition to increasing growth, DIN also increases the symbiont to
+host biomass ratio, a phenomenon also observed in reef corals (Marubini
+and Davies 1996). S:H ratios also increase under low light (Stimson
+1997; Fitt et al. 2000; Anthony and Hoegh-Guldberg 2003). At low DIN and
+intermediate light, more typical of coral reef environments, symbiont to
+host biomass ratios are around \textasciitilde{}0.1-0.3, which is
+consistent with values reported in the literature (Muscatine, R
+McCloskey, and E Marian 1981; Edmunds et al. 2011).
 
@@
 symbiont photosynthesis, across gradients of external irradiance and
-dissolved inorganic nitrogen. Simulations for each combination of light
-and nutrients (41 points along each axis) were run for 120 days with a
-time step of 1 day. Negative steady state growth rates, and
-corresponding S:H ratios, were set to zero.}
+dissolved inorganic nitrogen. Note that typical conditions for reefs are
+\textasciitilde{}1e-7 M DIN and 10-20 mol photons m\textsuperscript{-2}
+d\textsuperscript{-1}. Simulations for each combination of light and
+nutrients (41 points along each axis) were run for 100 days with a time
+step of 1 day. Negative steady state growth rates, and corresponding S:H
+ratios, were set to zero.}
 \end{figure}
 
-\section{More things to show:}\label{more-things-to-show}
+The maximum predicted growth rates of \textasciitilde{}0.1
+d\textsuperscript{-1}, occurring between \textasciitilde{}10-15 mol
+photons m\textsuperscript{-2} s\textsuperscript{-1} light and
+\textgreater{}\textasciitilde{}3 µM DIN (Fig. 2A), are comparable to the
+rate of 0.07 d\textsuperscript{-1} measured by Tanaka et al. (2007)
+under similar N-enriched conditions. Growth under conditions more
+typical of reef environments (\textless{}0.5 µM DIN) are within the
+range of \textasciitilde{}0.01-0.05 d\textsuperscript{-1}; however,
+these values are still higher than most specific growth rates reported
+in the literature, which are typically based on skeletal mass, and fall
+near or below 0.01 d\textsuperscript{-1} (Osinga et al. 2011; Osinga et
+al. 2012) (though values as high as 0.025 d\textsuperscript{-1} have
+been reported (Schutter et al. 2010)). These generally low skeletal
+growth rates relative to predicted biomass growth may be explained by
+differential substrate-limitation (i.e., CO\textsubscript{2} may limit
+calcification, which is not present in the model); indeed, tissue and
+skeletal growth may not always be well correlated (Anthony 2002).
+Furthermore, bioerosion may reduce net skeletal growth, especially in
+the field (Bak 1976). Nevertheless, when rates of increase in coral
+tissue biomass have been directly measured over short periods, values
+are consistent with model predictions (e.g., 0.07 d\textsuperscript{-1}
+in Tanaka et al. (2007)), and similar rates (0.04 d\textsuperscript{-1})
+have been measured in \emph{Aiptasia diaphana}, a non-calcifying
+symbiotic anemone (Armoza-Zvuloni et al. 2014).
+
+As irradiance continues to increase above \textasciitilde{}15-20 mol
+photons m\textsuperscript{-2} d\textsuperscript{-1}, growth rates
+decline until positive growth ceases above \textasciitilde{}30 µmol
+photons m\textsuperscript{-2} d\textsuperscript{-1}. The mechanism
+underlying this decline is the increase in light energy beyond the
+capacity of photosynthesis and non-photochemical quenching: excess
+excitation energy generates reactive oxygen species (ROS) (Weis 2008;
+Roth 2014), which, in this model, have the phenomenological consequences
+of reducing the photosynthetic rate (representing photoinhibition) and
+increasing the symbiont biomass loss rate (representing photodamage
+and/or symbiont expulsion) (see Eynaud, Nisbet, and Muller 2011).
+Together, these impacts reduce the symbiont to host biomass ratio (Fig.
+2C), as occurs during coral bleaching. This reduction in symbionts
+consequently reduces the flux of fixed carbon to the host, resulting in
+increasing carbon-limitation (Fig. 2B) and eventual cessation of growth
+(Fig. 2A). Importantly, these consequences arise in the present model
+only when light energy exceeds the capacity of photochemistry and NPQ,
+which in turn depends on CO\textsubscript{2} availability. While
+previous models framed photophysiological stress as a fixed response to
+absolute irradiance levels (Eynaud, Nisbet, and Muller 2011), the
+current implementation considers the dynamic balance of multiple energy
+sinks in the causation of stress, which is more consistent with current
+understanding of symbiosis dysfunction (Wooldridge 2013), and
+establishes an important role of the host in providing
+CO\textsubscript{2} for symbiont photosynthesis (Tansik, Fitt, and
+Hopkinson 2015; Hopkinson, Tansik, and Fitt 2015).
+
+While bleaching in response to high light alone has been observed
+experimentally (Schutter et al. 2011; Downs et al. 2013), mass coral
+bleaching events occur in response to high temperature (Hoegh-Guldberg
+1999); thus, it is important to justify our consideration of light as
+the primary stressor in the model. In reality, light and temperature
+interact synergistically (Coles and Jokiel 1978; Jones et al. 1998), and
+in fact, any stressor that disrupts the quenching of light energy may
+lead to bleaching (Wooldridge 2010; Baker and Cunning 2015). This is
+because the proximate cause of coral bleaching is excess excitation
+energy, but the upstream events that lead to this situation may be
+diverse: indeed, elevated temperature inhibits Rubisco functioning
+(Jones et al. 1998) and repair of the D1 protein in photosystem II
+(Warner, Fitt, and Schmidt 1999), which reduces the capacity of
+photochemical quenching and leads to an excess of light energy. In this
+way, elevated temperature serves to reduce the threshold above which
+light causes stress to the system (Hoegh-Guldberg 1999); importantly,
+light is still the proximate stressor. Therefore, we omitted temperature
+from the model, which serves to maintain a desired level of simplicity,
+while still allowing photophysiological stress and bleaching to be
+simulated with biological realism in response to light. Thus, the
+specific light levels that cause stress here should be interpreted in
+this context, and the effects of high light are analagous to the effects
+of high temperature.
+
+The incorporation of light stress in the model sets an upper limit to
+the amount of light at which a stable symbiotic interaction can be
+maintained, but even below this threshold of breakdown, negative effects
+of high light reduce steady state growth and symbiont:host biomass (Fig.
+2A,C). This gradual decline is consistent with experimental results
+showing that high light levels decrease growth (Schutter et al. 2011),
+and field studies documenting optimum growth rates at intermediate
+depths (Baker and Weber 1975; Huston 1985). Consequently, the model
+predicts greater variation in state variables across light gradients
+than was predicted by the model of Muller et al. (2009), which did not
+include photoinhibition and photodamage. It is important to recognize
+that these are steady state values predicted under a constant
+environment at the given light and nutrient levels; a dynamic system may
+temporarily cross this threshold and experience a period of symbiont
+loss (bleaching) and reduced growth, after which a return to benign
+conditions may restore symbiont biomass and positive growth. To explore
+this further and illustrate the behavior of the model in more detail, we
+evaluate a number of dynamic simulations below.
+
+\section{Dynamic behavior}\label{dynamic-behavior}
 
@@
 \subsection{elevated nutrients cause system to become C-limited (photo:
-CO2 limited; host: fixed C
+CO2 limited; host: fixed-C
 limited)}\label{elevated-nutrients-cause-system-to-become-c-limited-photo-co2-limited-host-fixed-c-limited}
@@
 
+\subsection{Seasonal change in sym. density is larger when nutrients are
+elevated (Stimson 1997). Such a phenomenon was predicted by Fitt et al.
+2000.}\label{seasonal-change-in-sym.-density-is-larger-when-nutrients-are-elevated-stimson-1997.-such-a-phenomenon-was-predicted-by-fitt-et-al.-2000.}
+
+\subsection{Death rule: negative growth for a certain amount of time
+should be
+defensible.}\label{death-rule-negative-growth-for-a-certain-amount-of-time-should-be-defensible.}
+
 \section*{References}\label{references}
@@
 \hypertarget{refs}{}
+\hypertarget{ref-Anthony:2002tc}{}
+Anthony, K. 2002. ``Comparative analysis of energy allocation to tissue
+and skeletal growth in corals.'' \emph{Limnology and Oceanography} 47
+(5): 1417--29.
+\url{https://www.scopus.com/inward/record.uri?partnerID=HzOxMe3b\&scp=0036733455\&origin=inward}.
+
+\hypertarget{ref-Anthony:2003p3632}{}
+Anthony, Kenneth R N, and Ove Hoegh-Guldberg. 2003. ``Variation in coral
+photosynthesis, respiration and growth characteristics in contrasting
+light microhabitats: an analogue to plants in forest gaps and
+understoreys?'' \emph{Functional Ecology} 17 (April): 246--59.
+
+\hypertarget{ref-ArmozaZvuloni:2014ju}{}
+Armoza-Zvuloni, Rachel, Esti Kramarsky-Winter, Yossi Loya, Ami
+Schlesinger, and Hanna Rosenfeld. 2014. ``Trioecy, a unique breeding
+strategy in the sea anemone Aiptasia diaphana and its association with
+sex steroids.'' \emph{Biology of Reproduction} 90 (6). Society for the
+Study of Reproduction: 122--22.
+doi:\href{https://doi.org/10.1095/biolreprod.113.114116}{10.1095/biolreprod.113.114116}.
+
+\hypertarget{ref-Bak:1976bv}{}
+Bak, R. 1976. ``The growth of coral colonies and the importance of
+crustose coralline algae and burrowing sponges in relation with
+carbonate accumulation.'' \emph{Netherlands Journal of Sea Research} 10
+(3): 285--337.
+doi:\href{https://doi.org/10.1016/0077-7579(76)90009-0}{10.1016/0077-7579(76)90009-0}.
+
+\hypertarget{ref-Baker:2015kp}{}
+Baker, Andrew C, and Ross Cunning. 2015. ``Coral 'Bleaching' as a
+Generalized Stress Response to Environmental Disturbance.'' In
+\emph{Diseases of Coral}, 396--409. Hoboken, NJ: John Wiley \& Sons,
+Inc.
+doi:\href{https://doi.org/10.1002/9781118828502.ch30}{10.1002/9781118828502.ch30}.
+
+\hypertarget{ref-Baker:1975ip}{}
+Baker, P, and Jon N Weber. 1975. ``Coral growth rate: Variation with
+depth.'' \emph{Earth and Planetary Science Letters} 27 (1): 57--61.
+doi:\href{https://doi.org/10.1016/0012-821X(75)90160-0}{10.1016/0012-821X(75)90160-0}.
+
+\hypertarget{ref-Barott:2014gx}{}
+Barott, Katie L, Alexander A Venn, Sidney O Perez, Sylvie Tambutte, and
+Martin Tresguerres. 2014. ``Coral host cells acidify symbiotic algal
+microenvironment to promote photosynthesis.'' \emph{Proceedings Of The
+National Academy Of Sciences Of The United States Of America}, December.
+National Acad Sciences, 201413483.
+doi:\href{https://doi.org/10.1073/pnas.1413483112}{10.1073/pnas.1413483112}.
+
+\hypertarget{ref-Brown:1999p3534}{}
+Brown, Barbara E, R P Dunne, I Ambarsari, Martin Le Tissier, and U
+Satapoomin. 1999. ``Seasonal fluctuations in environmental factors and
+variations in symbiotic algae and chlorophyll pigments in four
+Indo-Pacific coral species.'' \emph{Marine Ecology Progress Series} 191:
+53--69.
+
+\hypertarget{ref-Coles:1978p1124}{}
+Coles, SL, and Paul L Jokiel. 1978. ``Synergistic effects of
+temperature, salinity and light on the hermatypic coral Montipora
+verrucosa.'' \emph{Marine Biology} 49: 187--95.
+\url{http://www.springerlink.com/index/LU582837924G7674.pdf}.
+
 \hypertarget{ref-Costanza:2014ex}{}
@@
 doi:\href{https://doi.org/10.1016/j.gloenvcha.2014.04.002}{10.1016/j.gloenvcha.2014.04.002}.
+
+\hypertarget{ref-Downs:2013kc}{}
+Downs, C A, Kathleen E McDougall, Cheryl M Woodley, John E Fauth, Robert
+H Richmond, Ariel Kushmaro, Stuart W Gibb, Yossi Loya, Gary K Ostrander,
+and Esti Kramarsky-Winter. 2013. ``Heat-Stress and Light-Stress Induce
+Different Cellular Pathologies in the Symbiotic Dinoflagellate during
+Coral Bleaching.'' \emph{PLoS ONE} 8 (12): e77173.
+doi:\href{https://doi.org/10.1371/journal.pone.0077173}{10.1371/journal.pone.0077173}.
 
@@
 
+\hypertarget{ref-Fitt:2000p3792}{}
+Fitt, William K, F K McFarland, Mark E Warner, and Geoff C Chilcoat.
+2000. ``Seasonal patterns of tissue biomass and densities of symbiotic
+dinoflagellates in reef corals and relation to coral bleaching.''
+\emph{Limnology and Oceanography} 45 (3): 677--85.
+
 \hypertarget{ref-Glynn:2001p7571}{}
@@
 
+\hypertarget{ref-HoeghGuldberg:1999p4222}{}
+Hoegh-Guldberg, Ove. 1999. ``Climate change, coral bleaching and the
+future of the world's coral reefs.'' \emph{Marine and Freshwater
+Research} 50 (November): 839--66.
+
+\hypertarget{ref-Hopkinson:2015ev}{}
+Hopkinson, Brian M, Anna L Tansik, and William K Fitt. 2015. ``Internal
+carbonic anhydrase activity in the tissue of scleractinian corals is
+sufficient to support proposed roles in photosynthesis and
+calcification.'' \emph{The Journal of Experimental Biology}, April. The
+Company of Biologists Ltd, jeb.118182.
+doi:\href{https://doi.org/10.1242/jeb.118182}{10.1242/jeb.118182}.
+
+\hypertarget{ref-Huston:1985ec}{}
+Huston, M. 1985. ``Variation in coral growth rates with depth at
+Discovery Bay, Jamaica.'' \emph{Coral Reefs} 4 (1): 19--25.
+doi:\href{https://doi.org/10.1007/BF00302200}{10.1007/BF00302200}.
+
 \hypertarget{ref-Jager:2013bj}{}
@@
 (3): 201--8.
+
+\hypertarget{ref-Jones:1998p3572}{}
+Jones, Ross J, Ove Hoegh-Guldberg, Anthony W D Larkum, and Ulrich
+Schreiber. 1998. ``Temperature-induced bleaching of corals begins with
+impairment of the CO2 fixation mechanism in zooxanthellae.''
+\emph{Plant, Cell and Environment} 21 (12): 1219--30.
 
@@
 
+\hypertarget{ref-Marubini:1996p8460}{}
+Marubini, F, and PS Davies. 1996. ``Nitrate increases zooxanthellae
+population density and reduces skeletogenesis in corals.'' \emph{Marine
+Biology} 127 (2): 319--28.
+\url{http://www.scopus.com/inward/record.url?partnerID=yv4JPVwI\&eid=2-s2.0-0030301761\&md5=0971551a6865f30046294b9e17fb4081}.
+
 \hypertarget{ref-Muller:2009io}{}
@@
 
+\hypertarget{ref-Osinga:2011kq}{}
+Osinga, Ronald, Miriam Schutter, Ben Griffioen, René H Wijffels, Johan A
+J Verreth, Shai Shafir, Stéphane Henard, Maura Taruffi, Claudia Gili,
+and Silvia Lavorano. 2011. ``The Biology and Economics of Coral
+Growth.'' \emph{Marine Biotechnology} 13 (4): 658--71.
+doi:\href{https://doi.org/10.1007/s10126-011-9382-7}{10.1007/s10126-011-9382-7}.
+
+\hypertarget{ref-Osinga:2012is}{}
+Osinga, Ronald, Miriam Schutter, Tim Wijgerde, Buki Rinkevich, Shai
+Shafir, Muki Shpigel, Gian Marco Luna, et al. 2012. ``The CORALZOO
+project: a synopsis of four years of public aquarium science.''
+\emph{Journal of the Marine Biological Association of the UK} 92 (04).
+Cambridge University Press: 753--68.
+doi:\href{https://doi.org/10.1017/S0025315411001779}{10.1017/S0025315411001779}.
+
+\hypertarget{ref-Powles:1984tm}{}
+Powles, Stephen B. 1984. ``Photoinhibition of photosynthesis induced by
+visible light.'' \emph{Annual Review of Plant Physiology} 35: 15--44.
+
 \hypertarget{ref-RALanguageandEn:2014wf}{}
@@
 \url{http://journal.frontiersin.org/Journal/10.3389/fmicb.2014.00422/pdf}.
+
+\hypertarget{ref-Schutter:2010p7758}{}
+Schutter, M, J Crocker, A Paijmans, M Janse, R Osinga, A J Verreth, and
+R H Wijffels. 2010. ``The effect of different flow regimes on the growth
+and metabolic rates of the scleractinian coral Galaxea fascicularis.''
+\emph{Coral Reefs}, April.
+doi:\href{https://doi.org/10.1007/s00338-010-0617-2}{10.1007/s00338-010-0617-2}.
+
+\hypertarget{ref-Schutter:2011ie}{}
+Schutter, Miriam, Rosa M van der Ven, Max Janse, Johan A J Verreth, René
+H Wijffels, and Ronald Osinga. 2011. ``Light intensity, photoperiod
+duration, daily light flux and coral growth of Galaxea fascicularis in
+an aquarium setting: a matter of photons?'' \emph{Journal of the Marine
+Biological Association of the UK} 92 (04). Cambridge University Press:
+703--12.
+doi:\href{https://doi.org/10.1017/S0025315411000920}{10.1017/S0025315411000920}.
+
+\hypertarget{ref-Shantz:2015dz}{}
+Shantz, Andrew A, Nathan P Lemoine, and Deron E Burkepile. 2015.
+``Nutrient loading alters the performance of key nutrient exchange
+mutualisms.'' \emph{Ecology Letters} 19 (1): 20--28.
+doi:\href{https://doi.org/10.1111/ele.12538}{10.1111/ele.12538}.
+
+\hypertarget{ref-Stimson:1997p3837}{}
+Stimson, J. 1997. ``The annual cycle of density of zooxanthellae in the
+tissues of field and laboratory-held Pocillopora damicornis
+(Linnaeus).'' \emph{Journal of Experimental Marine Biology and Ecology}
+214 (1-2): 35--48.
 
@@
 
+\hypertarget{ref-Tanaka:2007uv}{}
+Tanaka, Yasuaki, Toshihiro Miyajima, Isao Koike, Takeshi Hayashibara,
+and Hiroshi Ogawa. 2007. ``Imbalanced coral growth between organic
+tissue and carbonate skeleton caused by nutrient enrichment.''
+\emph{Limnology and Oceanography}.
+\url{http://www.aslo.org/lo/toc/vol_52/issue_3/1139.pdf}.
+
+\hypertarget{ref-Tansik:2015cm}{}
+Tansik, Anna L, William K Fitt, and Brian M Hopkinson. 2015. ``External
+carbonic anhydrase in three Caribbean corals: quantification of activity
+and role in CO2 uptake.'' \emph{Coral Reefs} 34 (3). Springer Berlin
+Heidelberg: 703--13.
+doi:\href{https://doi.org/10.1007/s00338-015-1289-8}{10.1007/s00338-015-1289-8}.
+
+\hypertarget{ref-Titlyanov:2001p3631}{}
+Titlyanov, E A, T V Titlyanova, K Yamazato, and Robert van Woesik. 2001.
+``Photo-acclimation dynamics of the coral Stylophora pistillata to low
+and extremely low light.'' \emph{Journal of Experimental Marine Biology
+and Ecology} 263 (2): 211--25.
+
 \hypertarget{ref-Wang:1998p128}{}
@@
 Experimental Biology} 201: 2445--53.
+
+\hypertarget{ref-Warner:1999p4239}{}
+Warner, Mark E, William K Fitt, and Gregory W Schmidt. 1999. ``Damage to
+photosystem II in symbiotic dinoflagellates: A determinant of coral
+bleaching.'' \emph{Proceedings Of The National Academy Of Sciences Of
+The United States Of America} 96 (14): 8007--12.
 
@@
 
+\hypertarget{ref-Wooldridge:2010p7809}{}
+---------. 2010. ``Is the coral-algae symbiosis really 'mutually
+beneficial' for the partners?'' \emph{BioEssays}, May.
+
+\hypertarget{ref-Wooldridge:2013tj}{}
+---------. 2013. ``Breakdown of the coral-algae symbiosis: towards
+formalising a linkage between warm-water bleaching thresholds and the
+growth rate of the intracellular zooxanthellae.'' \emph{Biogeosciences}
+10: 1647--58.
+\url{http://www.biogeosciences.net/10/1647/2013/bg-10-1647-2013.html}.
+
 \hypertarget{ref-Wooldridge:2014di}{}

### 2016-10-06 R/jCO2_analysis.Rmd
updates from UCSB

---
title: "jCO2 analysis"
author: "Ross Cunning"
date: "October 5, 2016"
output: html_document
---

```{r load_functions, include=F, message=F}
sapply(c("../R/def_pars.R", "../R/init_env.R", "../R/run_coral.R", "../R/plot_results.R"), source, .GlobalEnv)
```

How many CO2's can the host deliver to photosynthesis SU per 1 C-mol of fixed carbon? The host will use this carbon to generate ATP to energize the CO2-concentrating mechanisms that deliver CO2 to the photosynthesis SU.
```{r new_model, include=F}
run_coral_wC <- function(time, env, pars, NPQ="inhibited") {
  require(dplyr)
  # Define parallel complementary synthesizing unit function
  synth <- function(x, y, m) 1/((1/m)+(1/x)+(1/y)-(1/(x+y)))
  #synth <- function(x, y, m) (m^-2 + x^-2 + y^-2)^(-1/2)
  # Set initial values
  # ==================
  # Host
  H <- data_frame(
    time=time,
    H=pars$initH,
    jX=(pars$jXm * env$X[1] / (env$X[1] + pars$KX)),
    jCO2a=pars$jCO2a,
    jCO2p=pars$jCO2p,
    jCO2=jCO2a+jCO2p,
    jN=(pars$jNm * env$N[1] / (env$N[1] + pars$KN)),
    rhoN=jN,
    jCw=0,
    jHG=0.25,
    jHT=pars$jHT0,
    rNH=jHT * pars$nNH * pars$sigmaNH,
    rCH=jHT * pars$sigmaCH,
    dH.Hdt=0
  )
  H[2:nrow(H), 2:ncol(H)] <- NA

  # Symbiont
  S <- data_frame(
    time=time,
    S=pars$initS,
    jL=env$L[1] * pars$astar,
    jCP=max(0, synth(jL * pars$nLC, (H$jCO2p[1] + H$jCO2a[1])*H$H/S, pars$jCPm), na.rm=T),
    jeL=max(jL - jCP/pars$nLC, 0),
    jNPQ=pars$kNPQ/pars$nLC,
    jCO2w=H$jCO2*H$H/S - jCP,
    jSG=pars$jSGm/10,
    rhoC=jCP, 
    jNw=0,
    jST=pars$jST0,
    rNS=jST * pars$nNS * pars$sigmaNS,
    rCS=jST * pars$sigmaCS,
    cROS=1,
    dS.Sdt=1)
  S[2:nrow(S), 2:ncol(S)] <- NA

  # Run simulation by updating
  # ==========================
  for (t in 2:length(time)) {

    # Photosynthesis SU
    # =================
    # Light input flux
    S$jL[t] <- (1.256307 + 1.385969 * exp(-6.479055 * (S$S[t-1]/H$H[t-1]))) * env$L[t] * pars$astar
    # CO2 input flux
    H$jCO2p[t] <- pars$jCO2p # passive CO2 flux from the environment
    H$jCO2a[t] <- pars$jCO2a  # active delivery of CO2 from the environment (mediated by the host)
    H$jCO2[t] <- H$jCO2p[t] + H$jCO2a[t]  # Total CO2 input from the environment
    S$rCS[t] <- pars$jST0 * pars$sigmaCS  # metabolic CO2 recycled from symbiont biomass turnover
    H$rCH[t] <- H$jHT[t-1] * pars$sigmaCH  # metabolic CO2 recycled from host biomass turnover
    H$jCw[t-1] <- H$jCw[t-1]  # carbon rejected from host biomass SU, given back as CO2 to photosynthesis
    # Production flux (photosynthetic carbon fixation)
    S$jCP[t] <- synth(S$jL[t] * pars$nLC, (5*H$jCw[t-1] + H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t], pars$jCPm) / S$cROS[t-1]

    # Rejection flux: CO2 (wasted to the environment)
    S$jCO2w[t] <- max((H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t] - S$jCP[t], 0)
    # Rejection flux: excess light energy not quenched by carbon fixation
    S$jeL[t] <- max(S$jL[t] - S$jCP[t]/pars$nLC, 0)
    # Amount of excess light energy quenched by NPQ
    #S$jNPQ[t] <- min(S$jeL[t], (S$jCP[t] * pars$kNPQ)/pars$nLC)
    if (NPQ=="inhibited") S$jNPQ[t] <- min(S$jeL[t], pars$kNPQ / S$cROS[t-1])
    if (NPQ=="inhibited2") S$jNPQ[t] <- (pars$kNPQ^(-3)+S$jeL[t]^(-3))^(-1/3) / S$cROS[t-1]
    if (NPQ=="fixed") S$jNPQ[t] <- min(S$jeL[t], pars$kNPQ)
    if (NPQ=="fixed2") S$jNPQ[t] <- (pars$kNPQ^(-3)+S$jeL[t]^(-3))^(-1/3)
    # Scaled ROS production due to excess excitation energy (=not quenched by carbon fixation AND NPQ)
    S$cROS[t] <- 1 + ((S$jeL[t] - S$jNPQ[t]) / pars$kROS)^pars$k

    # Symbiont biomass SU
    # ===================
    # Nitrogen input flux
    S$rNS[t] <- pars$jST0 * pars$nNS * pars$sigmaNS  # Recylced N from symbiont biomass turnover.
    H$rhoN[t-1] <- H$rhoN[t-1]  # Nitrogen shared from the host (defined below, so previous time step used)
    # Carbon input flux
    S$jCP[t] <- S$jCP[t]  # Production of fixed carbon from photosynthesis SU
    # Production flux (symbiont biomass formation)
    S$jSG[t] <- synth(S$jCP[t], (H$rhoN[t-1]*H$H[t-1]/S$S[t-1] + S$rNS[t])/pars$nNS, pars$jSGm)
    # Rejection flux: carbon (surplus carbon shared with the host)
    S$rhoC[t] <- max(S$jCP[t] - S$jSG[t], 0)
    # Rejection flux: nitrogen (surplus nitrogen wasted to the environment)
    S$jNw[t] <- max(H$rhoN[t-1]*H$H[t-1]/S$S[t-1] + S$rNS[t] - pars$nNS * S$jSG[t], 0)

    # Symbiont biomass loss (turnover)
    S$jST[t] <- pars$jST0 * (1 + pars$b * (S$cROS[t] - 1))

    # Host biomass SU
    # ===============
    # Food input flux (prey=both carbon and nitrogen)
    H$jX[t] <- (pars$jXm * env$X[t] / (env$X[t] + pars$KX))  # Prey uptake from the environment
    # Nitrogen input flux
    H$jN[t] <- (pars$jNm * env$N[t] / (env$N[t] + pars$KN))  # N uptake from the environment
    H$rNH[t] <- H$jHT[t-1] * pars$nNH * pars$sigmaNH  # Recycled N from host biomass turnover
    # Carbon input flux
    S$rhoC[t] <- S$rhoC[t]  # Carbon shared by the symbiont (defined above)
    # Production flux (host biomass formation)
    H$jHG[t] <- synth(S$rhoC[t]*S$S[t-1]/H$H[t-1] + H$jX[t], (H$jN[t] + pars$nNX*H$jX[t] + H$rNH[t]) / pars$nNH, pars$jHGm)
    # Rejection flux: nitrogen (surplus nitrogen shared with the symbiont)
    H$rhoN[t] <- max(H$jN[t] + pars$nNX * H$jX[t] + H$rNH[t] - pars$nNH * H$jHG[t], 0)
    # Rejection flux: carbon -- given back to symbiont as CO2 input to photosynthesis
    H$jCw[t] <- max(H$jX[t] + S$rhoC[t]*S$S[t-1]/H$H[t-1] - H$jHG[t], 0)

    # Host biomass loss
    H$jHT[t] <- pars$jHT0

    # State equations
    # ===============
    # Specific growth rates (Cmol/Cmol/d)
    S$dS.Sdt[t] <- S$jSG[t] - S$jST[t]
    H$dH.Hdt[t] <- H$jHG[t] - H$jHT[t]
    # Biomass (Cmol)
    S$S[t] <- S$S[t-1] + S$dS.Sdt[t] * S$S[t-1] * (time[2] - time[1])
    H$H[t] <- H$H[t-1] + H$dH.Hdt[t] * H$H[t-1] * (time[2] - time[1])

  }

  # Return results
  # ==============
  return(list(time=time, env=env, pars=pars, H=H, S=S))
}
```

```{r run_models, cache=T, message=F, warning=F, include=F}
##### VARY NITROGEN
# RUN USING WASTED CARBON AND NO JCO2A
time <- seq(1,1230,1)
defpars <- def_pars()
env <- init_env(time=time, L=c(15,15,2), N=c(1e-7,1e-6,2), X=c(5e-6,5e-6,0))
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
                              c(40,       30,    1,   5,      0.9,         0.9, 0)),
                 NPQ="fixed2")
plot_run_wC(run)

# RUN USING JCO2A (original version of model)
run <- run_coral(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
                              c(40,       30,    1,   5,      0.9,         0.9, 0.32)),
                 NPQ="fixed2")
plot_run(run)

##### VARY LIGHT
# RUN USING WASTED CARBON AND NO JCO2A
time <- seq(1,1230,1)
env <- init_env(time=time, L=c(10,25,2), N=c(1e-7,1e-7,2), X=c(5e-6,5e-6,0))
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
                              c(40,       30,    1,   5,      0.9,         0.9, 0)),
                 NPQ="fixed2")
plot_run_wC(run)

# RUN USING JCO2A (original version of model)
run <- run_coral(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
                              c(40,       30,    1,   5,      0.9,         0.9, 0.32)),
                 NPQ="fixed2")
plot_run(run)
```

### Now remove feeding
```{r run_models_nofeeding, cache=T, message=F, warning=F}
##### VARY NITROGEN
# RUN USING WASTED CARBON AND NO JCO2A OR JCO2P
time <- seq(1,1230,1)
defpars <- def_pars()
env <- init_env(time=time, L=c(15,15,2), N=c(1e-7,1e-6,2), X=c(0,0,0))
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                              c(40,       30,    1,   5,      0.9,         0.9, 0, 0)),
                 NPQ="fixed2")
plot_run_wC(run)

# RUN USING JCO2A (original version of model)
run <- run_coral(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
                              c(40,       30,    1,   5,      0.9,         0.9, 0.32)),
                 NPQ="fixed2")
plot_run(run)

##### VARY LIGHT
# RUN USING WASTED CARBON AND NO JCO2A OR JCO2P
time <- seq(1,1230,1)
env <- init_env(time=time, L=c(10,25,2), N=c(1e-7,1e-7,2), X=c(0,0,0))
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                              c(40,       30,    1,   5,      0.9,         0.9, 0, 0)),
                 NPQ="fixed2")
plot_run_wC(run)

# RUN USING JCO2A (original version of model)
run <- run_coral(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
                              c(40,       30,    1,   5,      0.9,         0.9, 0.32)),
                 NPQ="fixed2")
plot_run(run)
```

### Now make jPNQ inhibited
```{r run_models_nofeeding_jNPQinhibited, message=F, warning=F, cache=T}
##### VARY NITROGEN
# RUN USING WASTED CARBON AND NO JCO2A OR JCO2P
time <- seq(1,1230,1)
defpars <- def_pars()
env <- init_env(time=time, L=c(15,15,2), N=c(1e-7,1e-6,2), X=c(0,0,0))
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                              c(60,       40,    1,   5,      0.9,         0.9, 0, 0)),
                 NPQ="inhibited2")
plot_run_wC(run)

# RUN USING JCO2A (original version of model)
run <- run_coral(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
                              c(60,       40,    1,   5,      0.9,         0.9, 0.32)),
                 NPQ="inhibited2")
plot_run(run)

##### VARY LIGHT
# RUN USING WASTED CARBON AND NO JCO2A OR JCO2P
time <- seq(1,1230,1)
env <- init_env(time=time, L=c(10,25,2), N=c(1e-7,1e-7,2), X=c(0,0,0))
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                              c(60,       40,    1,   5,      0.9,         0.9, 0, 0)),
                 NPQ="inhibited2")
plot_run_wC(run)

# RUN USING JCO2A (original version of model)
run <- run_coral(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
                              c(60,       40,    1,   5,      0.9,         0.9, 0.32)),
                 NPQ="inhibited2")
plot_run(run)
```

### Trying different stress parameter sets
```{r par_tests, cache=T, warning=F, message=F}
##### VARY LIGHT
# RUN USING WASTED CARBON AND NO JCO2A OR JCO2P
time <- seq(1,1230,0.5)
env <- init_env(time=time, L=c(10,40,2), N=c(1e-7,1e-7,2), X=c(0,0,0))
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                              c(80,       40,    1,   5,      0.9,         0.9, 0, 0)),
                 NPQ="fixed2")
plot_run_wC(run)

run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                              c(80,       80,    1,   5,      0.9,         0.9, 0, 0)),
                 NPQ="fixed2")
plot_run_wC(run)

run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
                 NPQ="inhibited2")
plot_run_wC(run)
with(run, {
  plot(time, 5*H$jCw)
})
```

### 2016-10-06 R/Fig2.R

@@ -8,7 +8,7 @@ sapply(c("R/def_pars.R",
 time <- seq(1,100,1)  # Set time vector
 pars <- def_pars()  # Get default parameters
 # Set values of L and N at which to get steady state values
-at <- expand.grid(L=seq(from=0, to=40, length.out=41),
+at <- expand.grid(L=seq(from=0, to=60, length.out=41),
                   N=seq(from=0, to=4e-6, length.out=41))
 
 # Get steady state values and plot results to png

### 2016-10-06 R/Fig3.R

@@ -1,27 +1,0 @@
-library(reshape2)
-library(dplyr)
-
-source("R/def_pars.R")
-source("R/init_env.R")
-source("R/run_coral.R")
-source("R/plot_results.R")
-
-# Set run time vector
-time <- seq(1, 365, 0.1)  # if single model run, use time input
-
-# Initialize environment
-env <- init_env(time=time, 
-                L=c(5,30,2),  # Varying light
-                N=c(1e-7,1e-7,0),  # Low external DIN
-                X=c(0,0,0))  # No feeding
-
-# Set parameters
-pars <- def_pars()  # Get default parameters 
-
-# Run simulation
-run <- run_coral(time=time, env=env, pars=pars)
-
-# Make figure
-png("img/Fig3.png", height=6, width=3, units="in", res=300)
-plot_run2(run)
-dev.off()

### 2016-10-06 R/Fig4.R

@@ -1,0 +1,27 @@
+library(reshape2)
+library(dplyr)
+
+source("R/def_pars.R")
+source("R/init_env.R")
+source("R/run_coral.R")
+source("R/plot_results.R")
+
+# Set run time vector
+time <- seq(1, 365, 0.1)  # if single model run, use time input
+
+# Initialize environment
+env <- init_env(time=time, 
+                L=c(5,30,2),  # Varying light
+                N=c(1e-7,1e-7,0),  # Low external DIN
+                X=c(0,0,0))  # No feeding
+
+# Set parameters
+pars <- def_pars()  # Get default parameters 
+
+# Run simulation
+run <- run_coral(time=time, env=env, pars=pars)
+
+# Make figure
+png("img/Fig4.png", height=6, width=3, units="in", res=300)
+plot_run2(run)
+dev.off()

### 2016-10-06 R/def_pars.R

@@ -11,22 +11,25 @@ def_pars <- function() {
     jNm=0.048,  # Maximum specific host DIN uptake rate (molN/CmolH/d)
     jCO2p=4.04e-3,  # Specific flux of CO2 into tissue due to passive reaction-diffusion (molC/CmolH/d)
     jCO2a=0.32,  # Specific delivery of CO2 to photosynthesis by coral host (molC/CmolH/d)
-    jHGm=1,  # Maximum specific host growth rate (CmolH/CmolH/d)
+    jHGm=1000,  # Maximum specific host growth rate (CmolH/CmolH/d)
     KN=1e-6,  # Half-saturation constant for host DIN uptake (molN/L)
     KX=20e-6,  # Half-saturation constant for host feeding (molX/L)
     initH=1,  # Initial host biomass (CmolH)
     jST0=0.03,  # Symbiont specific biomass turnover rate (d^-1)
     nNS=0.13,  # N:C ratio in symbiont biomass (-)
     nLC=0.1,  # L:C ratio in fixed carbon (=quantum yield) (molC/mol ph)
-    kNPQ=4,  # capacity of non-photochemical quenching (mol ph/CmolS/d)
-    kROS=40,  # amount of (jeL-jNPQ) that doubles ROS production (mol ph/CmolS/d)
+    kNPQ=56,  # capacity of non-photochemical quenching (mol ph/CmolS/d)
+    # calculated as twice the max photosynthesis rate, converted from C to photons...
+    # A value of 40 was calculated from Gorbunov et al. 2001
+    kROS=60,  # amount of excess light beyond NPQ capacity (e.g., jeL-jNPQ) that doubles ROS production relative to baseline (mol ph/CmolS/d)
     k=1,  # exponent on ROS production (-)
     astar=1.34,  # Symbiont specific cross-sectional area (m^2/C-molS)
     sigmaNS=0.9,  # Proportion of symbiont nitrogen turnover recylced (-)
     sigmaCS=0.9,  # Proportion of symbiont carbon turnover recycled (-)
     jCPm=2.8,  # Maximum specific photosynthate production rate (Cmol/CmolS/d)
     jSGm=0.25,  # Maximum specific symbiont growth rate (CmolS/CmolS/d)
-    initS=0.1  # Initial symbiont biomass (CmolS)
+    initS=0.1,  # Initial symbiont biomass (CmolS)
+    b=3  # Scaling parameter for bleaching response
   ))
 }
 

### 2016-10-06 R/explore_NPQ.R

@@ -1,0 +1,71 @@
+# This function plots the photosynthetic rate and relative ROS production as a function of irradiance
+# for a given set of parameters (i.e., jCPm, jNPQ, kROS, k). The argument to this function should be a 
+# full set of model parameters (i.e., object with structure matching output of "def_pars.R"). This script 
+# is meant to show the photosynthetic performance of a given symbiont independent from a host (i.e., in 
+# a state that is not CO2-limited), so the "jCO2" parameter is set at a very high value here.
+
+sym <- function(pars, SU="pc", jL, jCO2, NPQ, NEC=NULL, kROS=NULL, photoinhibition=T) {
+  if (SU=="pc") synth <- function(x, y, m) 1/((1/m)+(1/x)+(1/y)-(1/(x+y)))
+  if (SU=="he") synth <- function(x, y, m) (m^-2 + x^-2 + y^-2)^(-1/2)
+  with(pars, {
+    # Set light range for PI curve
+    time <- seq(1,100,0.1)
+    jCO2 <- seq(jCO2[1], jCO2[2], along=time)
+    jL <- seq(jL[1], jL[2], along=time)
+    # Initialize values...
+    jCP <- c(pars$jCPm, rep(NA, length(time)-1))
+    jeL <- c(0, rep(NA, length(time)-1))
+    jNPQ <- c(0, rep(NA, length(time)-1))
+    cROS <- c(1, rep(NA, length(time)-1))
+    # Time-stepping solution...
+    for (t in 2:length(time)) {
+      # Calculate photosynthesis rate
+      if (photoinhibition==T) jCP[t] <- synth(jCO2[t], jL[t]*pars$nLC, pars$jCPm)/cROS[t-1]
+      if (photoinhibition==F) jCP[t] <- synth(jCO2[t], jL[t]*pars$nLC, pars$jCPm)
+      
+      # Calculate light in excess of photosynthetic quenching
+      jeL[t] <- max(0, jL[t] - jCP[t]/pars$nLC)
+      # Calculate light energy quenched by NPQ
+      if (NPQ=="none") {
+        jNPQ[t] <- 0
+      } else if (NPQ=="inhibited") {  #
[…]

### 2016-10-06 R/explore_jCO2.R

@@ -1,0 +1,392 @@
+# This script initializes and runs a simulation using specified environmental inputs and model parameters.
+# The arguments to this function are:
+#   - time: a vector of time steps at which the model should be evaluated (units=days) (e.g., seq(0,365,0.1))
+#   - env: an object with structure matching the output of "init_env.R"
+#   - pars: an object with structure matching the output of "def_pars.R"
+
+
+run_coral_jCO2a <- function(time, env, pars, NPQ="inhibited") {
+  require(dplyr)
+  # Define parallel complementary synthesizing unit function
+  synth <- function(x, y, m) 1/((1/m)+(1/x)+(1/y)-(1/(x+y)))
+  synth <- function(x, y, m) (m^-2 + x^-2 + y^-2)^(-1/2)
+  jCO2 <- seq(jCO2a_vals[1], jCO2a_vals[2], along=time)
+  # Set initial values
+  # ==================
+  # Host
+  H <- data_frame(
+    time=time,
+    H=pars$initH,
+    jX=(pars$jXm * env$X[1] / (env$X[1] + pars$KX)),
+    jCO2a=pars$jCO2a,
+    jCO2p=pars$jCO2p,
+    jCO2=jCO2a_vals[1],
+    jN=(pars$jNm * env$N[1] / (env$N[1] + pars$KN)),
+    rhoN=jN,
+    jCw=0,
+    jHG=0.25,
+    jHT=pars$jHT0,
+    rNH=jHT * pars$nNH * pars$sigmaNH,
+    rCH=jHT * pars$sigmaCH,
+    dH.Hdt=0
+  )
+  H[2:nrow(H), 2:ncol(H)] <- NA
+  
+  # Symbiont
+  S <- data_frame(
+    time=time,
+    S=pars$initS,
+    jL=env$L[1] * pars$astar,
+    jCP=max(0, synth(jL * pars$nLC, (H$jCO2p[1] + H$jCO2a[1])*H$H/S, pars$jCPm), na.rm=T),
+    jeL=max(jL - jCP/pars$nLC, 0),
+    jNPQ=pars$kNPQ/pars$nLC,
+    jCO2w=H$jCO2*H$H/S - jCP,
+    jSG=pars$jSGm/10,
+    rhoC=jCP, 
+    jNw=0,
+    jST=pars$jST0,
[…]

### 2016-10-06 R/plot_results.R

@@ -282,3 +282,81 @@ plot_run2 <- function(run) {
     #lines(time, H$jHG*pars$nNH, lwd=2) # Nitrogen used by host
   })
 }
+
+
+
+
+plot_run_wC <- function(run) {
+  with(run, {
+    # Set up graphical parameters
+    par(mar=c(2,2,1,1), oma=c(0,0,0,0), mfrow=c(5,2), mgp=c(1.2,0.2,0), tck=0.025, lwd=1, xaxs="i", 
+        cex.main=1.5, cex.axis=1, cex.lab=1)
+    
+    # External irradiance
+    plot(time, env$L, type="l", col="yellow", ylim=c(0,60), lwd=2, xlab="", ylab="mol/m2/d")
+    title("Light", adj=0.05, line=0)
+    
+    # External DIN concentration
+    plot(time, env$N * 10^6, type="l", col="blue", ylim=c(0,4), ylab="µmol/L", xlab="", lwd=2)
+    title("DIN", adj=0.05, line=0)
+    
+    # Specific growth rates of host and symbiont
+    Hgrf <- H$dH.Hdt[length(H$dH.Hdt)]
+    plot(time, H$dH.Hdt, type="l", ylim=c(min(0, min(c(H$dH.Hdt[-(1:10)], S$dS.Sdt[-(1:10)]))), max(c(H$dH.Hdt[-(1:10)], S$dS.Sdt[-(1:10)]))), 
+         xlab="", ylab="d-1", lwd=2, cex=1, cex.lab=1)
+    if(any(c(H$dH.Hdt[-(1:10)], S$dS.Sdt[-(1:10)])<0)) abline(h=0, lwd=0.5, lty=3)
+    text(time[0.95*length(time)], Hgrf, labels=as.character(round(Hgrf, 3)), pos=3, xpd=T, cex=0.75)
+    #title("State variable dynamics", adj=0, cex.main=2, outer = T)
+    title("Specific growth rate", adj=0.05, line=0)
+    lines(time, S$dS.Sdt, col="black", lwd=1)
+    legend("topright", legend=c("Host", "Sym"), lwd=c(3,1), col="black", bty="n", y.intersp=1)
+    
+    # Symbiont to host ratio
+    totSH <- S$S / H$H
+    totSHf <- totSH[length(totSH)]
+    plot(time, totSH, type="l", col="black", ylim=c(0, max(totSH, na.rm=T)), ylab="C-molS/C-molH", xlab="", lwd=2)
[…]

### 2016-10-06 R/plot_steady_states.R

@@ -53,11 +53,11 @@ plot_steady_states <- function(ss, png=NULL) {
     text(x=par("usr")[1]-par("mai")[2]*conv, y=par("usr")[4]+par("mai")[3]/2.5*conv, adj=c(0,0), 
          labels=main, xpd=NA, font=2)
     mtext(side=1, text="[DIN] (mol/L)", line=1.2, cex=0.75)
-    axis(side=1, at=seq(0,1,0.1), 
+    axis(side=1, at=seq(0,1,0.125), 
          labels=quantile(seq(as.numeric(first(colnames(r))),
                              as.numeric(last(colnames(r))),
                              length.out=100),
-                         probs=seq(0,1,0.1)))
+                         probs=seq(0,1,0.125)))
     mtext(side=2, text="Light (mol ph/m2/d)", line=1.2, cex=0.75)
     axis(side=2, at=seq(0,1,0.1), 
          labels=quantile(seq(as.numeric(first(rownames(r))),

### 2016-10-06 R/reported_metrics.R

@@ -1,0 +1,17 @@
+# Metrics cited in this manuscript derived from literature
+
+# Specific growth rates
+# =====================
+# From Tanaka et al. 2007 Limnol Oceanogr: Table 1
+C0 <- 343  # tissue carbon at time zero (µmolC/cm2)
+C5 <- 495  # tissue carbon after 5 days of nutrient enrichment (µmolC/cm2)
+Tanaka_gr <- log(C5/C0) / 5  # specific growth of tissue carbon (d^-1)
+Tanaka_gr <- round(Tanaka_gr, 2)
+
+# From Armoza-Zvuloni et al. 2014
+Armoza_gr <- 0.2938  # Exponential growth rate reported on page 3 (wk^-1)
+Armoza_gr <- 0.2938 / 7  # Convert to d^-1
+Armoza_gr <- round(Armoza_gr, 2)
+
+# From Shafir et al. 20016
+Shafir_gr <- 0.0167  # d^-1; average of 5 species in a nursery setting, based on approximate volumes

### 2016-10-06 Rmd/analyze.Rmd
updates from UCSB

@@
 ---
-title: "Untitled"
+title: "Model behavior analysis"
 author: "Ross Cunning, Erik B. Muller, Ruth D. Gates, Roger R. Nisbet"
@@
 knitr::opts_knit$set(root.dir = normalizePath(".."))
+knitr::opts_chunk$set(message = F, warning = F)
 ```
@@
 ```{r initialize, include=FALSE}
-# Initialize SU function
-synth <- function(x, y, m) 1/((1/m)+(1/x)+(1/y)-(1/(x+y)))
 # Source default parameters and functions
 sapply(c("R/conversions.R", "R/def_pars.R", "R/init_env.R", "R/plot_results.R",
-         "R/run_coral.R", "R/sym_PI.R"), source, .GlobalEnv)
+         "R/run_coral.R", "R/sym_PI.R", "R/run_steady_states.R", "R/plot_steady_states.R",
+         "R/explore_NPQ.R"), source, .GlobalEnv)
 ```
 
-```{r lightamp, fig.width=5, fig.height=5, eval=F}
+```{r lightamp, fig.width=5, fig.height=5, eval=F, include=F}
 # Import data from Marcelino et al. 2013
@@
 
+### Steady states values with default parameters (jCO2a=0.32)
+```{r ss, cache=T, fig.width=5, fig.height=4.4}
+# Set time and parameters
+time <- seq(1,100,1)
+defpars <- def_pars()  # Get default parameters
+
+# Set values of L and N at which to get steady state values
+at <- expand.grid(L=seq(from=0, to=60, length.out=11),
+                  N=seq(from=0, to=4e-6, length.out=11))
+
+# Run to steady state and plot results
+ssdat <- run_steady_states(time=time, pars=defpars, at=at)
+plot_steady_states(ssdat)
+```
+
+With default parameters, steady state values of specific growth are higher than usually observed in corals (skeletal growth).
+
+### Change kROS from 60 to 100
+```{r, cache=T, fig.width=5, fig.height=4.4, eval=T, include=T}
+pars <- defpars
+pars$kROS <- 100
+# Run to steady state and plot results
+ssdat <- run_steady_states(time=time, pars=pars, at=at)
+plot_steady_states(ssdat)#
+```
+
+### Change kROS to 100 and b to 5
+```{r, cache=T, fig.width=5, fig.height=4.4, eval=T, include=T}
+pars <- defpars
+pars$kROS <- 100
+pars$b <- 5
+# Run to steady state and plot results
+ssdat <- run_steady_states(time=time, pars=pars, at=at)
+plot_steady_states(ssdat)#
+```
+
+### Take away recycling...
+```{r ss2, cache=T, fig.width=5, fig.height=4.4, eval=F, include=F}
+pars <- defpars
+pars$sigmaCH <- 0
+pars$sigmaCS <- 0
+pars$sigmaNH <- 0
+pars$sigmaNS <- 0
+# Run to steady state and plot results
+ssdat <- run_steady_states(time=time, pars=pars, at=at)
+plot_steady_states(ssdat)
+```
+
+Specific growth rates are reduced by a bit, and both growth and S:H do not increase as much in response in increasing DIN.
+
+### Increase CO2 delivery rate (jCO2a=0.5)
+```{r, cache=T, fig.width=5, fig.height=4.4, eval=F, include=F}
+ssdat <- run_steady_states(time=time, at=at,
+                            pars=replace(defpars, "jCO2a", 0.5))
+plot_steady_states(ssdat) 
+```
+
+CO~2~ limitation of photosynthesis is reduced, and S:H biomass does not increase as much in response to low light.
+
+### Increase CO2 delivery rate even more (jCO2a=1)
+```{r, cache=T, fig.width=5, fig.height=4.4, eval=F, include=F}
+ssdat <- run_steady_states(time=time, at=at,
+                            pars=replace(defpars, "jCO2a", 1))
+plot_steady_states(ssdat) 
+```
+
+Photosynthesis never becomes CO~2~-limited.
+
+### Decrease CO2 delivery rate (jCO2a=0.1)
+```{r, cache=T, fig.width=5, fig.height=4.4, eval=F, include=F}
+ssdat <- run_steady_states(time=time, at=at,
+                            pars=replace(defpars, "jCO2a", 0.1))
+plot_steady_states(ssdat) 
+```
+
+Photosynthesis is always CO~2~-limited, and host growth almost always carbon-limited. Upper light limit is reduced, and S:H biomass ratios are increased.
+
+### Decrease CO2 delivery rate even more (jCO2a=0.05)
+```{r, cache=T, fig.width=5, fig.height=4.4, eval=F, include=F}
+ssdat <- run_steady_states(time=time, at=at,
+                            pars=replace(defpars, "jCO2a", 0.05))
+plot_steady_states(ssdat) 
+```
+
+Increasing DIN now decreases growth rates. S:H ratios are uniformly high.
+
+### Default parameters with tolerant symbiont (kROS=100; jCPm=1.4)
+```{r, cache=T, fig.width=5, fig.height=4.4, eval=F, include=F}
+ssdat <- run_steady_states(time=time, at=at,
+                            pars=replace(defpars, c("kROS","jCPm"), c(100, 1.4)))
+plot_steady_states(ssdat) 
+```
+
+### Default parameters with only lower photosynthesis rate (jCPm=1.4)
+```{r, cache=T, fig.width=5, fig.height=4.4, eval=F, include=F}
+ssdat <- run_steady_states(time=time, at=at,
+                            pars=replace(defpars, "jCPm", 1.4))
+plot_steady_states(ssdat) 
+```
+
+### Default parameters with only lower photosynthesis rate (jCPm=2)
+```{r, cache=T, fig.width=5, fig.height=4.4, eval=F, include=F}
+ssdat <- run_steady_states(time=time, at=at,
+                            pars=replace(defpars, "jCPm", 2))
+plot_steady_states(ssdat) 
+```
+
 ### Run 1
@@
 # Set run time vector
-time <- seq(1, 365, 0.1)  # if single model run, use time input
+time <- seq(1, 365, 0.5)  # if single model run, use time input
 
 # Initialize environment
-env <- init_env(time=time, L=c(20,32,2), N=c(1e-6,1e-6,0), X=c(0,0,0))
+env <- init_env(time=time, L=c(5,40,2), N=c(1e-7,1e-7,0), X=c(0,0,0))
 
 # Set parameters
-pars <- def_pars()  # Get default parameters
-#pars$rpars$gammaR <- 0.03  # Change any default parameters
-#pars$jCO2a <- 1
-#pars$jCO2a <- 0.1
+defpars <- def_pars()  # Get default parameters
 
 # Run simulation
-run <- run_coral(time=time, env=env, pars=pars)
+run <- run_coral(time=time, env=env, pars=defpars)#
 ```
@@
 ##### Plot run 1 symbiont PI curve
-```{r run1_PI, fig.height=1}
-#sym_PI(pars)
+```{r run1_PI, cache=T, fig.height=3}
+sym(pars=replace(defpars, c("kNPQ", "kROS", "k"), c(40, 30, 1)), jL=c(0,250), jCO2=c(100,100), NPQ="fixed")
 ```
@@
 ##### Plot run 1 output
-```{r run1_plots}
+```{r abb, fig.height=8, fig.width=3, eval=T, include=T}
+time <- seq(1,365,0.5)
+env <- init_env(time=time, L=c(5,40,2), N=c(1e-7,1e-7,0), X=c(4e-6,4e-6,0))
+run <- run_coral(time=time, env=env, 
+                 pars=replace(defpars, c("kNPQ", "kROS", "k", "b", "jCO2a"), c(30, 10, 1, 5, 0.32)),
+                 NPQ="fixed2", photoinhibition=T)
+plot_run2(run)
+env <- init_env(time=time, L=c(10,40,2), N=c(1e-7,1e-7,0), X=c(4e-6,4e-6,0))
+run <- run_coral_jCO2a(time=time, env=env,
+                 pars=replace(defpars, c("kNPQ", "kROS", "k", "b"), c(80, 80, 1, 10)),
+                 NPQ="fixed2")
+run$H$jCO2a
+run$S$jCP * run$S$S / run$H$H
+plot_run2(run)
+
+env <- init_env(time=time, L=c(40,40,0), N=c(1e-7,1e-7,0), X=c(5e-5,5e-5,0))
+run <- run_coral_jCO2a_manual(time=time, env=env, jCO2a_vals=c(3,1),
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH"), 
+                              c(60,       40,    1,   5,      0,         0)),
+                 NPQ="fixed2")
+plot_run2(run)
+
+##### VARY NITROGEN
+# RUN USING WASTED CARBON AND NO JCO2A
+time <- seq(1,1230,1)
+env <- init_env(time=time, L=c(15,15,2), N=c(1e-7,1e-6,2), X=c(5e-6,5e-6,0))
+run <- run_coral_wC(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
+                              c(40,       30,    1,   5,      0.9,         0.9, 0)),
+                 NPQ="fixed2")
 plot_run(run)
+
+# RUN USING JCO2A (original version of model)
+run <- run_coral(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
+                              c(40,       30,    1,   5,      0.9,         0.9, 0.32)),
+                 NPQ="fixed2")
+plot_run(run)
+
+##### VARY LIGHT
+# RUN USING WASTED CARBON AND NO JCO2A
+time <- seq(1,1230,1)
+env <- init_env(time=time, L=c(10,25,2), N=c(1e-7,1e-7,2), X=c(5e-6,5e-6,0))
+run <- run_coral_wC(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
+                              c(40,       30,    1,   5,      0.9,         0.9, 0)),
+                 NPQ="fixed2")
+plot_run_wC(run)
+
+# RUN USING JCO2A (original version of model)
+run <- run_coral(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
+                              c(40,       30,    1,   5,      0.9,         0.9, 0.32)),
+                 NPQ="fixed2")
+plot_run(run)
+```
+
+##### Run 2 -- increase kROS to 100 and b to 5
+```{r, fig.height=7, fig.width=3, eval=T, include=T}
+run <- run_coral(time=time, env=env, pars=replace(defpars, c("kROS", "b"), c(100, 5)))
+plot_run2(run)#
+```
+
+##### Run 2 -- increase kROS to 100 and b to 10
+```{r, fig.height=7, fig.width=3, eval=T, include=T}
+run <- run_coral(time=time, env=env, pars=replace(defpars, c("kROS", "b"), c(100, 10)))
+plot_run2(run)#
+```
+
+##### Run 2 -- fluctuate DIN
+```{r run2_plot, fig.height=7, fig.width=3, eval=T, include=T}
+env2 <- init_env(time=time, L=c(20,20,0), N=c(1e-7,4e-6,2), X=c(0,0,0))
+run2 <- run_coral(time=time, env=env2, pars=pars)
+#plot_run2(run2)
+par(mar=c(2,2,1,1), oma=c(0,0,0,0), mfrow=c(6,1), mgp=c(1,0,0), tck=0.025, lwd=1, xaxs="i", 
+    cex.main=1.2, cex.axis=1, cex.lab=1)
+plot_DIN(run2)
+plot_sh(run2)
+plot_gr(run2)
+plot_Lq(run2)
+plot_pl(run2)
+plot_bl(run2)
+
+```
+
+##### Run 3 -- fluctuate light with higher DIN
+```{r run3_plot, fig.height=3, fig.width=3, eval=T, include=T}
+# Set run time vector
+#time <- seq(1, 365, 0.5)  # if single model run, use time input
+
+# Initialize environment
+env <- init_env(time=time, L=c(2,40,2), N=c(4e-6,4e-6,0), X=c(0,0,0))
+
+# Run simulation
+run3 <- run_coral(time=time, env=env, pars=defpars)
+par(mar=c(2,2,1,1), oma=c(0,0,0,0), mfrow=c(2,1), mgp=c(1,0,0), tck=0.025, lwd=1, xaxs="i", 
+    cex.main=1.2, cex.axis=1, cex.lab=1)
+plot_sh(run)  # run 1 -- lower DIN
+plot_sh(run3)  # run3 -- higher DIN
 
@@
 ##### Run 2 -- increase jCO2a
-```{r run2_plots}
+```{r run2_plots, eval=F, include=F}
 pars2 <- pars
-pars2$jCO2a <- 10
+pars2$jCO2a <- 1
 run2 <- run_coral(time=time, env=env, pars=pars2)
@@
 ```
+
+##### Run 3 -- increase DIN
+```{r run3_plots, eval=F, include=F}
+env3 <- init_env(time=time, L=c(5,30,2), N=c(4e-6,4e-6,0), X=c(0,0,0))
+pars3 <- pars2
+
+run3 <- run_coral(time=time, env=env3, pars=pars3)
+plot_run(run3)
+
+```
+
+#### RUn 4 - 2e-6 DIN and

### 2016-10-06 R/run_coral.R

@@ -5,11 +5,11 @@
 #   - pars: an object with structure matching the output of "def_pars.R"
 
 
-run_coral <- function(time, env, pars) {
+run_coral <- function(time, env, pars, NPQ="inhibited", photoinhibition=T) {
   require(dplyr)
   # Define parallel complementary synthesizing unit function
   synth <- function(x, y, m) 1/((1/m)+(1/x)+(1/y)-(1/(x+y)))
-  
+  #synth <- function(x, y, m) (m^-2 + x^-2 + y^-2)^(-1/2)
   # Set initial values
   # ==================
   # Host
@@ -65,15 +65,24 @@ run_coral <- function(time, env, pars) {
     S$rCS[t] <- pars$jST0 * pars$sigmaCS  # metabolic CO2 recycled from symbiont biomass turnover
     H$rCH[t] <- H$jHT[t-1] * pars$sigmaCH  # metabolic CO2 recycled from host biomass turnover
     # Production flux (photosynthetic carbon fixation)
-    S$jCP[t] <- synth(S$jL[t] * pars$nLC, (H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t], pars$jCPm) / S$cROS[t-1]
+    if (photoinhibition==T) {
+      S$jCP[t] <- synth(S$jL[t] * pars$nLC, (H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t], pars$jCPm) / S$cROS[t-1]
+    } else if (photoinhibition==F) {
+      S$jCP[t] <- synth(S$jL[t] * pars$nLC, (H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t], pars$jCPm)
+    }
+    
     # Rejection flux: CO2 (wasted to the environment)
     S$jCO2w[t] <- max((H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t] - S$jCP[t], 0)
     # Rejection flux: excess light energy not quenched by carbon fixation
     S$jeL[t] <- max(S$jL[t] - S$jCP[t]/pars$nLC, 0)
     # Amount of excess light energy quenched by NPQ
[…]

### 2016-10-06 R/run_steady_states.R

@@ -27,7 +27,7 @@ run_steady_states <- function(time, pars, at, outfile=NULL, runtime=F) {
     sh <- ss$S$S/ss$H$H
     hl <- with(ss, log(  pmin(S$rhoC*S$S/H$H + H$jX, pars$jHGm) / pmin((H$jN + pars$nNX*H$jX + H$rNH) / pars$nNH, pars$jHGm)  ))
     sl <- with(ss, log(  pmin(S$jCP, pars$jSGm)   /   pmin((H$rhoN*H$H/S$S + S$rNS)/pars$nNS, pars$jSGm)  ))
-    ee <- with(ss, max(0, S$jL - (S$jCP/run$pars$nLC + run$pars$jNPQ)))
+    ee <- with(ss, max(0, S$jL - (S$jCP/run$pars$nLC + run$pars$kNPQ)))
     pl <- with(ss, log(   pmin((H$jCO2 + H$rCH)*H$H/S$S + S$rCS, pars$jCPm)   /   pmin(S$jL * pars$nLC, pars$jCPm)    ))
     data.frame(gr=gr, sh=sh, hl=hl, sl=sl, ee=ee, pl=pl)
   }

### 2016-10-06 R/sensitivity.R

@@ -1,0 +1,386 @@
+library(parallel)
+library(doParallel)
+library(foreach)
+source("R/def_pars.R")
+source("R/run_coral.R")
+source("R/init_env.R")
+
+# Sensitivity analysis
+
+# Create function to measure sensitivity of responses to specified change in single parameter
+sens <- function(time, env, pars, par, change) {
+  # Run 1 
+  run1 <- run_coral(time=time, env=env, pars=pars)
+  run1ss <- lapply(run1[c("H", "S")], function(x) x[length(time), ])
+  run1gr <- run1ss$H$dH.Hdt
+  run1sh <- run1ss$S$S/run1ss$H$H
+  # Run 2 - with changed parameter
+  run2 <- run_coral(time=time, env=env, pars=replace(pars, par, with(pars, get(par))*change))
+  run2ss <- lapply(run2[c("H", "S")], function(x) x[length(time), ])
+  run2gr <- run2ss$H$dH.Hdt
+  run2sh <- run2ss$S$S/run2ss$H$H
+  # Calculate relative change in steady state response
+  grchange <- run2gr / run1gr
+  shchange <- run2sh / run1sh
+  cROSchange <- run2ss$S$cROS / run1ss$S$cROS
+  # Return relative changes
+  return(list(grchange=grchange, shchange=shchange, cROSchange=cROSchange))
+}
+
+# Define function to measure sensitivity across a range of changes in parameter value
+vsens <- function(par, levs, env) {
+  gr <- rep(NA, length(levs))
+  sh <- rep(NA, length(levs))
+  cROS <- rep(NA, length(levs))
+  for (i in 1:length(levs)) {
+    sa <- sens(time=time, env=env, pars=defpars, par=par, change=levs[i])
+    gr[i] <- sa[[1]]
+    sh[i] <- sa[[2]]
+    cROS[i] <- sa[[3]]
+  }
+  return(data.frame(lev=levs, gr=gr, sh=sh, cROS=cROS))
+}
+
[…]

### 2016-10-06 R/sym_PI.R

@@ -5,25 +5,39 @@
 # a state that is not CO2-limited), so the "jCO2" parameter is set at a very high value here.
 
 sym_PI <- function(pars) {
+  synth <- function(x, y, m) 1/((1/m)+(1/x)+(1/y)-(1/(x+y)))
   with(pars, {
-    jL <- seq(0,200,1)
+    # Set light range for PI curve
+    time <- seq(1,100,0.01)
+    jL <- seq(0,250,along=time)
     # Set carbon delivery rate (mol C / CmolS / d)
-    jCO2 <- 100  # Extremely high carbon delivery - no carbon limitation
-    # Calculate photosynthesis rate
-    jCP <- synth(jCO2, jL*pars$nLC, pars$jCPm)
-    # Calculate light in excess of photosynthetic quenching
-    jeL <- pmax(0, jL - jCP/pars$nLC)
-    # Calculate ROS (cROS) produced due to excess light
-    cROS <- pmax(1, 1 + (pmax(0, jeL-pars$jNPQ)/(pars$kROS))^pars$k)
-    # Calculate photosynthesis rate accounting for ROS-induced photoinhibition
-    jCP <- jCP / cROS
+    jCO2 <- 1000  # Extremely high carbon delivery - no carbon limitation
+    # Initialize values...
+    jCP <- c(pars$jCPm, rep(NA, length(time)-1))
+    jeL <- c(0, rep(NA, length(time)-1))
+    jNPQ <- c(0, rep(NA, length(time)-1))
+    cROS <- c(1, rep(NA, length(time)-1))
+    # Time-stepping solution...
+    for (t in 2:length(time)) {
+      # Calculate photosynthesis rate
+      jCP[t] <- synth(jCO2, jL[t]*pars$nLC, pars$jCPm)/cROS[t-1]
+      # Calculate light in excess of photosynthetic quenching
+      jeL[t] <- max(0, jL[t] - jCP[t]/pars$nLC)
+      # Calculate light energy quenched by NPQ
+      jNPQ[t] <- min(jeL[t], pars$kNPQ / cROS[t-1])
+      # Calculate ROS (cROS) produced due to excess light
+      cROS[t] <- 1 + ((jeL[t] - jNPQ[t]) / pars$kROS)^pars$k
+    }
     # Return ROS and photosynthesis rate for plotting
[…]

### 2016-10-07 R/jCO2_analysis.Rmd
new model structure: jCw = CO2 input

@@
 ```{r load_functions, include=F, message=F}
-sapply(c("../R/def_pars.R", "../R/init_env.R", "../R/run_coral.R", "../R/plot_results.R"), source, .GlobalEnv)
+sapply(c("../R/def_pars.R", "../R/init_env.R", "../R/plot_results.R", "../R/plot_steady_states.R"), source, .GlobalEnv)
 ```
 
-How many CO2's can the host deliver to photosynthesis SU per 1 C-mol of fixed carbon?
-The host will use this carbon to generate ATP to energize the CO2-concentrating mechanisms that deliver CO2 to the photosynthesis SU.
-```{r new_model, include=F}
-run_coral_wC <- function(time, env, pars, NPQ="inhibited") {
+```{r old_model_code, include=F}
+run_coral <- function(time, env, pars, NPQ="inhibited", photoinhibition=T) {
   require(dplyr)
@@
     rhoC=jCP, 
+    jNw=0,
+    jST=pars$jST0,
+    rNS=jST * pars$nNS * pars$sigmaNS,
+    rCS=jST * pars$sigmaCS,
+    cROS=1,
+    dS.Sdt=1)
+  S[2:nrow(S), 2:ncol(S)] <- NA
+
+  # Run simulation by updating
+  # ==========================
+  for (t in 2:length(time)) {
+
+    # Photosynthesis SU
+    # =================
+    # Light input flux
+    S$jL[t] <- (1.256307 + 1.385969 * exp(-6.479055 * (S$S[t-1]/H$H[t-1]))) * env$L[t] * pars$astar
+    # CO2 input flux
+    H$jCO2p[t] <- pars$jCO2p # passive CO2 flux from the environment
+    H$jCO2a[t] <- pars$jCO2a  # active delivery of CO2 from the environment (mediated by the host)
+    H$jCO2[t] <- H$jCO2p[t] + H$jCO2a[t]  # Total CO2 input from the environment
+    S$rCS[t] <- pars$jST0 * pars$sigmaCS  # metabolic CO2 recycled from symbiont biomass turnover
+    H$rCH[t] <- H$jHT[t-1] * pars$sigmaCH  # metabolic CO2 recycled from host biomass turnover
+    # Production flux (photosynthetic carbon fixation)
+    if (photoinhibition==T) {
+      S$jCP[t] <- synth(S$jL[t] * pars$nLC, (H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t], pars$jCPm) / S$cROS[t-1]
+    } else if (photoinhibition==F) {
+      S$jCP[t] <- synth(S$jL[t] * pars$nLC, (H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t], pars$jCPm)
+    }
+
+    # Rejection flux: CO2 (wasted to the environment)
+    S$jCO2w[t] <- max((H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t] - S$jCP[t], 0)
+    # Rejection flux: excess light energy not quenched by carbon fixation
+    S$jeL[t] <- max(S$jL[t] - S$jCP[t]/pars$nLC, 0)
+    # Amount of excess light energy quenched by NPQ
+    #S$jNPQ[t] <- min(S$jeL[t], (S$jCP[t] * pars$kNPQ)/pars$nLC)
+    if (NPQ=="inhibited") S$jNPQ[t] <- min(S$jeL[t], pars$kNPQ / S$cROS[t-1])
+    if (NPQ=="inhibited2") S$jNPQ[t] <- (pars$kNPQ^(-3)+S$jeL[t]^(-3))^(-1/3) / S$cROS[t-1]
+    if (NPQ=="fixed") S$jNPQ[t] <- min(S$jeL[t], pars$kNPQ)
+    if (NPQ=="fixed2") S$jNPQ[t] <- (pars$kNPQ^(-3)+S$jeL[t]^(-3))^(-1/3)
+    # Scaled ROS production due to excess excitation energy (=not quenched by carbon fixation AND NPQ)
+    S$cROS[t] <- 1 + ((S$jeL[t] - S$jNPQ[t]) / pars$kROS)^pars$k
+
+    # Symbiont biomass SU
+    # ===================
+    # Nitrogen input flux
+    S$rNS[t] <- pars$jST0 * pars$nNS * pars$sigmaNS  # Recylced N from symbiont biomass turnover.
+    H$rhoN[t-1] <- H$rhoN[t-1]  # Nitrogen shared from the host (defined below, so previous time step used)
+    # Carbon input flux
+    S$jCP[t] <- S$jCP[t]  # Production of fixed carbon from photosynthesis SU
+    # Production flux (symbiont biomass formation)
+    S$jSG[t] <- synth(S$jCP[t], (H$rhoN[t-1]*H$H[t-1]/S$S[t-1] + S$rNS[t])/pars$nNS, pars$jSGm)
+    # Rejection flux: carbon (surplus carbon shared with the host)
+    S$rhoC[t] <- max(S$jCP[t] - S$jSG[t], 0)
+    # Rejection flux: nitrogen (surplus nitrogen wasted to the environment)
+    S$jNw[t] <- max(H$rhoN[t-1]*H$H[t-1]/S$S[t-1] + S$rNS[t] - pars$nNS * S$jSG[t], 0)
+
+    # Symbiont biomass loss (turnover)
+    S$jST[t] <- pars$jST0 * (1 + pars$b * (S$cROS[t] - 1))
+
+    # Host biomass SU
+    # ===============
+    # Food input flux (prey=both carbon and nitrogen)
+    H$jX[t] <- (pars$jXm * env$X[t] / (env$X[t] + pars$KX))  # Prey uptake from the environment
+    # Nitrogen input flux
+    H$jN[t] <- (pars$jNm * env$N[t] / (env$N[t] + pars$KN))  # N uptake from the environment
+    H$rNH[t] <- H$jHT[t-1] * pars$nNH * pars$sigmaNH  # Recycled N from host biomass turnover
+    # Carbon input flux
+    S$rhoC[t] <- S$rhoC[t]  # Carbon shared by the symbiont (defined above)
+    # Production flux (host biomass formation)
+    H$jHG[t] <- synth(S$rhoC[t]*S$S[t-1]/H$H[t-1] + H$jX[t], (H$jN[t] + pars$nNX*H$jX[t] + H$rNH[t]) / pars$nNH, pars$jHGm)
+    # Rejection flux: nitrogen (surplus nitrogen shared with the symbiont)
+    H$rhoN[t] <- max(H$jN[t] + pars$nNX * H$jX[t] + H$rNH[t] - pars$nNH * H$jHG[t], 0)
+    # Rejection flux: carbon (surplus carbon lost to environment)
+    H$jCw[t] <- max(H$jX[t] + S$rhoC[t]*S$S[t-1]/H$H[t-1] - H$jHG[t], 0)
+
+    # Host biomass loss
+    H$jHT[t] <- pars$jHT0
+
+    # State equations
+    # ===============
+    # Specific growth rates (Cmol/Cmol/d)
+    S$dS.Sdt[t] <- S$jSG[t] - S$jST[t]
+    H$dH.Hdt[t] <- H$jHG[t] - H$jHT[t]
+    # Biomass (Cmol)
+    S$S[t] <- S$S[t-1] + S$dS.Sdt[t] * S$S[t-1] * (time[2] - time[1])
+    H$H[t] <- H$H[t-1] + H$dH.Hdt[t] * H$H[t-1] * (time[2] - time[1])
+
+  }
+
+  # Return results
+  # ==============
+  return(list(time=time, env=env, pars=pars, H=H, S=S))
+}
+```
+
+# Code for new model
+The rejection flux of carbon from host biomass SU, $jCw$, is used as input to photosynthesis SU.
+A multiplier is included indicating the number of moles of CO~2~ the host can deliver to the photosynthesis SU using one mole of fixed carbon.
+Mechanistically, the host will use this carbon to generate ATP to energize a variety of CO~2~-concentrating mechanisms (CCM's) that enhance CO~2~ availability to photosynthesis.
+```{r new_model, include=T}
+run_coral_wC <- function(time, env, pars, NPQ="inhibited", jDOC=0) {
+  require(dplyr)
+  # Define parallel complementary synthesizing unit function
+  synth <- function(x, y, m) 1/((1/m)+(1/x)+(1/y)-(1/(x+y)))
+  #synth <- function(x, y, m) (m^-2 + x^-2 + y^-2)^(-1/2)
+  # Set initial values
+  # ==================
+  # Host
+  H <- data_frame(
+    time=time,
+    H=pars$initH,
+    jX=(pars$jXm * env$X[1] / (env$X[1] + pars$KX)),
+    jCO2a=pars$jCO2a,
+    jCO2p=pars$jCO2p,
+    jCO2=jCO2a+jCO2p,
+    jN=(pars$jNm * env$N[1] / (env$N[1] + pars$KN)),
+    rhoN=jN/2,
+    jCw=1,
+    jHG=0.25,
+    jHT=pars$jHT0,
+    rNH=jHT * pars$nNH * pars$sigmaNH,
+    rCH=jHT * pars$sigmaCH,
+    dH.Hdt=0
+  )
+  H[2:nrow(H), 2:ncol(H)] <- NA
+
+  # Symbiont
+  S <- data_frame(
+    time=time,
+    S=.2,
+    jL=env$L[1] * pars$astar,
+    jCP=max(0, synth(jL * pars$nLC, (H$jCw[1] + H$jCO2p[1] + H$jCO2a[1])*H$H/S, pars$jCPm), na.rm=T),
+    jeL=max(jL - jCP/pars$nLC, 0),
+    jNPQ=pars$kNPQ/pars$nLC,
+    jCO2w=(H$jCw + H$jCO2)*H$H/S - jCP,
+    jSG=pars$jSGm/10,
+    rhoC=jCP/2, 
     jNw=0,
@@
     S$rhoC[t] <- S$rhoC[t]  # Carbon shared by the symbiont (defined above)
+    # Add dissolved organic carbon uptake
     # Production flux (host biomass formation)
-    H$jHG[t] <- synth(S$rhoC[t]*S$S[t-1]/H$H[t-1] + H$jX[t], (H$jN[t] + pars$nNX*H$jX[t] + H$rNH[t]) / pars$nNH, pars$jHGm)
+    H$jHG[t] <- synth(S$rhoC[t]*S$S[t-1]/H$H[t-1] + H$jX[t] + jDOC, (H$jN[t] + pars$nNX*H$jX[t] + H$rNH[t]) / pars$nNH, pars$jHGm)
     # Rejection flux: nitrogen (surplus nitrogen shared with the symbiont)
@@
 
-```{r run_models, cache=T, message=F, warning=F, include=F}
-##### VARY NITROGEN
-# RUN USING WASTED CARBON AND NO JCO2A
+# Comparison of new and old model behavior
+## Varying nitrogen
+### Old model
+```{r vary_DIN_old, cache=T, message=F, warning=F}
+# Set up simulation
 time <- seq(1,1230,1)
 defpars <- def_pars()
-env <- init_env(time=time, L=c(15,15,2), N=c(1e-7,1e-6,2), X=c(5e-6,5e-6,0))
-run <- run_coral_wC(time=time, env=env,
-                 pars=replace(defpars, 
-                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
-                              c(40,       30,    1,   5,      0.9,         0.9, 0)),
-                 NPQ="fixed2")
-plot_run_wC(run)
+defpars <- replace(defpars, "jHGm", 1)
+env <- init_env(time=time, L=c(15,15,2), N=c(1e-7,1e-6,2), X=c(0,0,0))
 
-# RUN USING JCO2A (original version of model)
+# Run old model (uses jCO2a and jCO2p to deliver CO2 to photosynthesis SU)
 run <- run_coral(time=time, env=env,
                  pars=replace(defpars, 
-                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
-                              c(40,       30,    1,   5,      0.9,         0.9, 0.32)),
-                 NPQ="fixed2")
-plot_run(run)
-
-##### VARY LIGHT
-# RUN USING WASTED CARBON AND NO JCO2A
-time <- seq(1,1230,1)
-env <- init_env(time=time, L=c(10,25,2), N=c(1e-7,1e-7,2), X=c(5e-6,5e-6,0))
-run <- run_coral_wC(time=time, env=env,
-                 pars=replace(defpars, 
-                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
-                              c(40,       30,    1,   5,      0.9,         0.9, 0)),
-                 NPQ="fixed2")
-plot_run_wC(run)
-
-# RUN USING JCO2A (original version of model)
-run <- run_coral(time=time, env=env,
-                 pars=replace(defpars, 
-                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
-                              c(40,       30,    1,   5,      0.9,         0.9, 0.32)),
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jHGm"), 
+                              c(60,       60,    1,   5,      0.9,      0.9,      0.32,     1)),
                  NPQ="fixed2")
@@
 
-### Now remove feeding
-```{r run_models_nofeeding, cache=T, message=F, warning=F}
-##### VARY NITROGEN
-# RUN USING WASTED CARBON AND NO JCO2A OR JCO2P
-time <- seq(1,1230,1)
-defpars <- def_pars()
-env <- init_env(time=time, L=c(15,15,2), N=c(1e-7,1e-6,2), X=c(0,0,0))
+### New model
+```{r vary_DIN_new, cache=T, message=F, warning=F}
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
 run <- run_coral_wC(time=time, env=env,
                  pars=replace(defpars, 
-                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
-                              c(40,       30,    1,   5,      0.9,         0.9, 0, 0)),
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "jHGm"), 
+                              c(60,       60,    1,   5,      0.9,      0.9,       0,       0,      1)),
                  NPQ="fixed2")
 plot_run_wC(run)
+```
 
-# RUN USING JCO2A (original version of model)
+## Varying light
+### Old model
+```{r vary_L_old, cache=T, message=F, warning=F}
+# Set up simulation
+time <- seq(1,1230,1)
+env <- init_env(time=time, L=c(10,25,2), N=c(1e-7,1e-7,2), X=c(0,0,0))
+
+# Run old model (uses jCO2a and jCO2p to deliver CO2 to photosynthesis SU)
 run <- run_coral(time=time, env=env,
@@
                               c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
-                              c(40,       30,    1,   5,      0.9,         0.9, 0.32)),
-                 NPQ="fixed2")
-plot_run(run)
-
-##### VARY LIGHT
-# RUN USING WASTED CARBON AND NO JCO2A OR JCO2P
-time <- seq(1,1230,1)
-env <- init_env(time=time, L=c(10,25,2), N=c(1e-7,1e-7,2), X=c(0,0,0))
-run <- run_coral_wC(time=time, env=env,
-                 pars=replace(defpars, 
-                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
-                              c(40,       30,    1,   5,      0.9,         0.9, 0, 0)),
-                 NPQ="fixed2")
-plot_run_wC(run)
-
-# RUN USING JCO2A (original version of model)
-run <- run_coral(time=time, env=env,
-                 pars=replace(defpars, 
-                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
-                              c(40,       30,    1,   5,      0.9,         0.9, 0.32)),
+                              c(60,       60,    1,   5,      0.9,     0.9,      0.32)),
                  NPQ="fixed2")
@@
 
-### Now make jPNQ inhibited
-```{r run_models_nofeeding_jNPQinhibited, message=F, warning=F, cache=T}
-##### VARY NITROGEN
-# RUN USING WASTED CARBON AND NO JCO2A OR JCO2P
-time <- seq(1,1230,1)
-defpars <- def_pars()
-env <- init_env(time=time, L=c(15,15,2), N=c(1e-7,1e-6,2), X=c(0,0,0))
-run <- run_coral_wC(time=time, env=env,
-                 pars=replace(defpars, 
-                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
-                              c(60,       40,    1,   5,      0.9,         0.9, 0, 0)),
-                 NPQ="inhibited2")
-plot_run_wC(run)
-
-# RUN USING JCO2A (original version of model)
-run <- run_coral(time=time, env=env,
-                 pars=replace(defpars, 
-                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
-                              c(60,       40,    1,   5,      0.9,         0.9, 0.32)),
-                 NPQ="inhibited2")
-plot_run(run)
-
-##### VARY LIGHT
-# RUN USING WASTED CARBON AND NO JCO2A OR JCO2P
-time <- seq(1,1230,1)
-env <- init_env(time=time, L=c(10,25,2), N=c(1e-7,1e-7,2), X=c(0,0,0))
-run <- run_coral_wC(time=time, env=env,
-                 pars=replace(defpars, 
-                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
-                              c(60,       40,    1,   5,      0.9,         0.9, 0, 0)),
-                 NPQ="inhibited2")
-plot_run_wC(run)
-
-# RUN USING JCO2A (original version of model)
-run <- run_coral(time=time, env=env,
-                 pars=replace(defpars, 
-                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
-                              c(60,       40,    1,   5,      0.9,         0.9, 0.32)),
-                 NPQ="inhibited2")
-plot_run(run)
-```
-
-### Trying different stress parameter sets
-```{r par_tests, cache=T, warning=F, message=F}
-##### VARY LIGHT
-# RUN USING WASTED CARBON AND NO JCO2A OR JCO2P
-time <- seq(1,1230,0.5)
-env <- init_env(time=time, L=c(10,40,2), N=c(1e-7,1e-7,2), X=c(0,0,0))
-run <- run_coral_wC(time=time, env=env,
-                 pars=replace(defpars, 
-                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
-                              c(80,       40,    1,   5,      0.9,         0.9, 0, 0)),
-                 NPQ="fixed2")
-plot_run_wC(run)
-
-run <- run_coral_wC(time=time, env=env,
-                 pars=replace(defpars, 
-                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
-                              c(80,       80,    1,   5,      0.9,         0.9, 0, 0)),
-                 NPQ="fixed2")
-plot_run_wC(run)
-
+### New model
+```{r vary_L_new, cache=T, message=F, warning=F}
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
 run <- run_coral_wC(time=time, env=env,
@@
                               c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
-                 NPQ="inhibited2")
+                 NPQ="fixed2")
 plot_run_wC(run)
+```
+
+## Light stress
+### old model: Increasing Light in a non-feeding host
+```{r incr_L_old_nofeeding, cache=T, message=F, warning=F}
+# Set up simulation
+env <- init_env(time=time, L=c(2,60,0), N=c(1e-7,1e-7,0), X=c(0,0,0))
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
+run <- run_coral(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0.32)),
+                 NPQ="fixed2")
+plot_run(run)
+```
+
+### new model: Increasing Light in a non-feeding host
+```{r incr_L_nofeeding, cache=T, message=F, warning=F}
+# Set up simulation
+env <- init_env(time=time, L=c(2,60,0), N=c(1e-7,1e-7,0), X=c(0,0,0))
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
+run <- run_coral_wC(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
+                 NPQ="fixed2")
+plot_run_wC(run)
+```
+
+### new model: Increasing Light in a feeding host
+```{r incr_L, cache=T, message=F, warning=F}
+# Set up simulation
+env <- init_env(time=time, L=c(2,60,0), N=c(1e-7,1e-7,0), X=c(5e-6,5e-6,0))
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
+run <- run_coral_wC(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
+                 NPQ="fixed2")
+plot_run_wC(run)
+```
+
+#### possibility of DOC uptake?
+```{r incr_L_DOC, cache=T, message=F, warning=F, include=F, eval=F}
+# Set up simulation
+env <- init_env(time=time, L=c(2,60,0), N=c(1e-7,1e-7,0), X=c(0,0,0))
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
+run <- run_coral_wC(time=time, env=env, jDOC=0.005,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
+                 NPQ="fixed2")
+plot_run_wC(run)
+```
+
+### Light stress with old model
+```{r Lstress_old, cache=T, message=F, warning=F}
+# Set up simulation
+time <- seq(1,1230,1)
+env <- init_env(time=time, L=c(10,40,2), N=c(1e-7,1e-7,2), X=c(0,0,0))
+
+# Run old model (uses jCO2a and jCO2p to deliver CO2 to photosynthesis SU)
+run <- run_coral(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
+                              c(60,       60,    1,   5,      0.9,     0.9,      0.32)),
+                 NPQ="fixed2")
+plot_run(run)
+```
+
+### Light stress with new model
+```{r Lstress_new, cache=T, message=F, warning=F}
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
+run <- run_coral_wC(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
+                 NPQ="fixed2")
+plot_run_wC(run)
+```
+
+#### Examine sources of CO~2~ for photosynthesis in new model
+```{r CO2_sources, cache=T, message=F, warning=F, fig.width=3.5, fig.height=6}
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw and metabolic CO2)
+run <- run_coral_wC(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
+                 NPQ="fixed2")
+par(mfrow=c(4,1), mar=c(2,2,1,1))
+# Plot light
+plot_L(run)
+# Plot symbiont to host ratio
+plot_sh(run)
+# Plot symbiont-specific CO2 delivery sources
 with(run, {
-  plot(time, 5*H$jCw)
+  totalCO2 <- (H$jCw+H$rCH)*H$H/S$S + S$rCS
+  plot(time, totalCO2, type="l", col="gray")
+  lines(time, H$jCw*H$H/S$S, lty=1, col="black")
+  lines(time, H$rCH*H$H/S$S, lty=2, col="brown")
+  lines(time, S$rCS, lty=2, col="green")
+})
+# Plot proportions of each CO2 source
+with(run, {
+  totalCO2 <- (H$jCw+H$rCH)*H$H/S$S + S$rCS
+  plot(time, H$jCw*H$H/S$S/totalCO2, type="l", ylim=c(0,1))
+  lines(time, H$rCH*H$H/S$S/totalCO2, lty=2, col="brown")
+  lines(time, S$rCS/totalCO2, lty=2, col="green")
 })
 ```
+
+### Light stress with new model, +feeding
+```{r Lstress_new_feeding, cache=T, message=F, warning=F}
+# Set up simulation
+env <- init_env(time=time, L=c(10,40,2), N=c(1e-7,1e-7,2), X=c(1e-6,1e-6,0))
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
+run <- run_coral_wC(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
+                 NPQ="fixed2")
+plot_run_wC(run)
+```
+
+### Light stress with new model, +feeding, reduced DIN
+```{r Lstress_new_feeding_lowerDIN, cache=T, message=F, warning=F}
+# Set up simulation
+env <- init_env(time=time, L=c(10,40,2), N=c(0.5e-7,0.5e-7,2), X=c(1e-6,1e-6,0))
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
+run <- run_coral_wC(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
+                 NPQ="fixed2")
+plot_run_wC(run)
+```
+
+### Light stress with new model, +tolerant symbiont
+```{r Lstress_new_tolerant, cache=T, message=F, warning=F}
+# Set up simulation
+env <- init_env(time=time, L=c(10,40,2), N=c(1e-7,1e-7,2), X=c(1e-6,1e-6,0))
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
+run <- run_coral_wC(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
+                              c(60,       120,    1,   5,      0.9,         0.9, 0, 0)),
+                 NPQ="fixed2")
+plot_run_wC(run)
+```
+
+## Increasing DIN
+### Old model
+```{r incr_DIN_old, cache=T, message=F, warning=F}
+# Set up simulation
+env <- init_env(time=time, L=c(30,30,0), N=c(1e-8,4e-6,0), X=c(0,0,0))
+# Run old model
+run <- run_coral(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0.32)),
+                 NPQ="fixed2")
+plot_run(run)
+```
+
+### New model
+```{r incr_DIN_new, cache=T, message=F, warning=F}
+# Set up simulation
+env <- init_env(time=time, L=c(30,30,0), N=c(1e-8,4e-6,0), X=c(0,0,0))
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
+run <- run_coral_wC(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
+                 NPQ="fixed2")
+plot_run_wC(run)
+```
+
+# Steady state behavior of new model
+```{r new_steadystates_code, message=F, warning=F, include=F}
+run_steady_states <- function(time, pars, at, outfile=NULL, runtime=F, food=0, jDOC=0) {
+  #
+  run_coral_wC <- get("run_coral_wC")
+  # Load libraries
+  require(foreach)
+  require(parallel)
+  require(doParallel)
+  require(reshape2)
+  require(dplyr)
+
+  # Load functions
+  source("../R/init_env.R", local=TRUE)
+  #source("R/run_coral.R", local=TRUE)
+
+  # Set up cluster for parallel processing
+  cl <- makeCluster(detectCores())  # Initiate cluster
+  registerDoParallel(cl)
+
+  # Run simulations in parallel for each combination of "at" values
+  start <- proc.time()  # Start timer...
+  # Calculate steady states...
+  steady_states <- foreach(i=1:nrow(at), .combine='rbind', .packages='dplyr') %dopar% {
+    L <- at[i,]$L; N <- at[i,]$N
+    env <- init_env(time=time, L=c(L,L,0), N=c(N,N,0), X=c(food,food,0))
+    run <- run_coral_wC(time=time, env=env, pars=pars, jDOC=jDOC)
+    ss <- lapply(run[c("H", "S")], function(x) x[length(time), ])
+    gr <- ss$H$dH.Hdt
+    sh <- ss$S$S/ss$H$H
+    hl <- with(ss, log(  pmin(S$rhoC*S$S/H$H + H$jX, pars$jHGm) / pmin((H$jN + pars$nNX*H$jX + H$rNH) / pars$nNH, pars$jHGm)  ))
+    sl <- with(ss, log(  pmin(S$jCP, pars$jSGm)   /   pmin((H$rhoN*H$H/S$S + S$rNS)/pars$nNS, pars$jSGm)  ))
+    ee <- with(ss, max(0, S$jL - (S$jCP/run$pars$nLC + run$pars$kNPQ)))
+    pl <- with(ss, log(   pmin((H$jCw + H$jCO2 + H$rCH)*H$H/S$S + S$rCS, pars$jCPm)   /   pmin(S$jL * pars$nLC, pars$jCPm)    ))
+    data.frame(gr=gr, sh=sh, hl=hl, sl=sl, ee=ee, pl=pl)
+  }
+  stopCluster(cl)  # Stop cluster
+  if (runtime==T) print(proc.time() - start)  # Print time elapsed if runtime==T
+
+  # Collect results in ss
+  ss <- cbind(at, steady_states)
+
+  # If output filename give, save ss as Rdata
+  if (!is.null(outfile)) save(ss, file=outfile)
+
+  # Return steady state values
+  return(ss)
+}
+```
+
+## No feeding
+```{r run_new_steadystates, cache=T, message=F, warning=F}
+# Set time and parameters 
+time <- seq(1,100,1) 
+pars <- replace(defpars, c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
+                          c(60,       60,    1,   5,      0.9,     0.9,       0,       0))
+# Set values of L and N at which to get steady state values 
+at <- expand.grid(L=seq(from=0, to=60, length.out=41),
+                  N=seq(from=0, to=4e-6, length.out=41))
+
+# Run to steady state and plot results
+ssdat <- run_steady_states(time=time, pars=pars, at=at, food=0)
+plot_steady_states(ssdat)
+```
+
+## + feeding
+```{r run_new_steadystates_feeding, cache=T, message=F, warning=F}
+# Set time and parameters 
+time <- seq(1,100,1)
+pars <- replace(defpars, c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
+                          c(60,       60,    1,   5,      0.9,     0.9,       0,       0))
+# Set values of L and N at which to get steady state values 
+at <- expand.grid(L=seq(from=0, to=60, length.out=41),
+                  N=seq(from=0, to=4e-6, length.out=41))
+
+# Run to steady state and plot results
+ssdat <- run_steady_states(time=time, pars=pars, at=at, food=5e-6)
+plot_steady_states(ssdat)
+```
+
+## + feeding + tolerant symbiont
+```{r run_new_steadystates_feeding_tolerant, cache=T, message=F, warning=F}
+# Set time and parameters 
+time <- seq(1,100,1)
+pars <- replace(defpars, c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
+                          c(60,       200,    1,   5,      0.9,     0.9,       0,       0))
+# Set values of L and N at which to get steady state values 
+at <- expand.grid(L=seq(from=0, to=60, length.out=41),
+                  N=seq(from=0, to=4e-6, length.out=41))
+
+# Run to steady state and plot results
+ssdat <- run_steady_states(time=time, pars=pars, at=at, food=5e-6)
+plot_steady_states(ssdat)
+```
+
+# Stability at high nitrogen...?
+## When system starts at 2 µM DIN and increases to 4 µM DIN, outcome is negative growth.
+```{r incrN1, cache=T, message=F, warning=F}
+# Set up simulation
+env <- init_env(time=time, L=c(10,10,0), N=c(2e-6,4e-6,0), X=c(0,0,0))
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
+run <- run_coral_wC(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "jHGm"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0, 1)),
+                 NPQ="fixed2")
+plot_run_wC(run)
+```
+
+## When system starts at 1 µM DIN and increases to 4 µM DIN, outcome is positive growth.
+```{r incrN2, cache=T, message=F, warning=F}
+# Set up simulation
+env <- init_env(time=time, L=c(10,10,0), N=c(1e-6,4e-6,0), X=c(0,0,0))
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
+run <- run_coral_wC(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "jHGm"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0, 1)),
+                 NPQ="fixed2")
+plot_run_wC(run)
+```

### 2016-10-09 R/plot_results.R
adjust plotting functions

@@ -137,6 +137,16 @@ plot_bl <- function(run) with(run, {
 # ========================================================
 
 plot_run <- function(run) {
+  # If run was a steady state run, set up time and environment vectors for plotting
+  if (is.null(run$time)) {
+    run <- within(run, {
+      time <- seq(1, nrow(S))
+      env$L <- rep(env$L, last(time))
+      env$N <- rep(env$N, last(time))
+      env$X <- rep(env$X, last(time))
+    })
+  }
+  # Plotting
   with(run, {
     # Set up graphical parameters
     par(mar=c(2,2,1,1), oma=c(0,0,0,0), mfrow=c(5,2), mgp=c(1.2,0.2,0), tck=0.025, lwd=1, xaxs="i", 
@@ -213,7 +223,8 @@ plot_run <- function(run) {
 
 
 
-plot_run2 <- function(run) {
+plot_run_v2 <- function(run) {
+  if (is.null(run$time)) time <- seq(1, nrow(run$S))
   with(run, {
     # Set up graphical parameters
     par(mar=c(2,2,1,1), oma=c(0,0,0,0), mfrow=c(6,1), mgp=c(1,0,0), tck=0.025, lwd=1, xaxs="i", 
@@ -280,83 +291,5 @@ plot_run2 <- function(run) {
     #plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(0, max(c(H$jN+H$rNH))))
     #lines(time, H$jN + H$rNH + H$jX*pars$nNX, lwd=2, lty=2) # Total nitrogen available to host
     #lines(time, H$jHG*pars$nNH, lwd=2) # Nitrogen used by host
-  })
-}
-
-
-
-
-plot_run_wC <- function(run) {
-  with(run, {
-    # Set up graphical parameters
-    par(mar=c(2,2,1,1), oma=c(0,0,0,0), mfrow=c(5,2), mgp=c(1.2,0.2,0), tck=0.025, lwd=1, xaxs="i", 
-        cex.main=1.5, cex.axis=1, cex.lab=1)
-    
-    # External irradiance
-    plot(time, env$L, type="l", col="yellow", ylim=c(0,60), lwd=2, xlab="", ylab="mol/m2/d")
-    title("Light", adj=0.05, line=0)
-    
-    # External DIN concentration
[…]

### 2016-10-09 R/plot_steady_states.R

@@ -83,7 +83,7 @@ plot_steady_states <- function(ss, png=NULL) {
   par(mfrow=c(2,2))
   imagef(gr, bin=0.01, main="A. Specific growth", col="grayscale")
   imagef(hl, bin=0.25, main="B. Host growth limitation", col=list("#67001F","#053061"))
-  imagef(sh, bin=0.1, main="C. Symbiont:host biomass", col="grayscale")
-  imagef(pl, bin=0.3, main="D. Photosynthesis limitation", col=list("#67001F", "gray30"))
+  imagef(sh, bin=0.05, main="C. Symbiont:host biomass", col="grayscale")
+  imagef(pl, bin=0.25, main="D. Photosynthesis limitation", col=list("#67001F", "gray30"))
   if (!is.null(png)) dev.off()
 }

### 2016-10-09 R/jCO2_analysis.Rmd
finish comparison of jeC/jCO2 changes

@@
 ```{r new_model, include=T}
-run_coral_wC <- function(time, env, pars, NPQ="inhibited", jDOC=0) {
+run_coral_wC <- function(time, env, pars, NPQ="inhibited", kCO2=5) {
   require(dplyr)
@@
     jN=(pars$jNm * env$N[1] / (env$N[1] + pars$KN)),
-    rhoN=jN/2,
-    jCw=1,
+    rhoN=jN,
+    jCw=10,
     jHG=0.25,
@@
     time=time,
-    S=.2,
+    S=pars$initS,
     jL=env$L[1] * pars$astar,
@@
     jSG=pars$jSGm/10,
-    rhoC=jCP/2, 
+    rhoC=jCP, 
     jNw=0,
@@
     # Production flux (photosynthetic carbon fixation)
-    S$jCP[t] <- synth(S$jL[t] * pars$nLC, (5*H$jCw[t-1] + H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t], pars$jCPm) / S$cROS[t-1]
+    S$jCP[t] <- synth(S$jL[t] * pars$nLC, (kCO2*H$jCw[t-1] + H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t], pars$jCPm) / S$cROS[t-1]
 
@@
     # Production flux (host biomass formation)
-    H$jHG[t] <- synth(S$rhoC[t]*S$S[t-1]/H$H[t-1] + H$jX[t] + jDOC, (H$jN[t] + pars$nNX*H$jX[t] + H$rNH[t]) / pars$nNH, pars$jHGm)
+    H$jHG[t] <- synth(S$rhoC[t]*S$S[t-1]/H$H[t-1] + H$jX[t], (H$jN[t] + pars$nNX*H$jX[t] + H$rNH[t]) / pars$nNH, pars$jHGm)
     # Rejection flux: nitrogen (surplus nitrogen shared with the symbiont)
@@
 
-#### possibility of DOC uptake?
-```{r incr_L_DOC, cache=T, message=F, warning=F, include=F, eval=F}
-# Set up simulation
-env <- init_env(time=time, L=c(2,60,0), N=c(1e-7,1e-7,0), X=c(0,0,0))
-# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
-run <- run_coral_wC(time=time, env=env, jDOC=0.005,
-                 pars=replace(defpars, 
-                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
-                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
-                 NPQ="fixed2")
-plot_run_wC(run)
-```
-
 ### Light stress with old model
@@
 ```{r Lstress_new, cache=T, message=F, warning=F}
+# Set up simulation
+time <- seq(1,1230,1)
+env <- init_env(time=time, L=c(10,40,2), N=c(1e-7,1e-7,2), X=c(0,0,0))
 # Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
-run <- run_coral_wC(time=time, env=env,
+run <- run_coral_wC(time=time, env=env, kCO2=2,
                  pars=replace(defpars, 
@@
 ```{r new_steadystates_code, message=F, warning=F, include=F}
-run_steady_states <- function(time, pars, at, outfile=NULL, runtime=F, food=0, jDOC=0) {
+run_steady_states <- function(time, pars, at, outfile=NULL, runtime=F, food=0) {
   #
@@
     env <- init_env(time=time, L=c(L,L,0), N=c(N,N,0), X=c(food,food,0))
-    run <- run_coral_wC(time=time, env=env, pars=pars, jDOC=jDOC)
+    run <- run_coral_wC(time=time, env=env, pars=pars, NPQ="fixed2")
     ss <- lapply(run[c("H", "S")], function(x) x[length(time), ])
@@
 
-## No feeding
+## Starting biomass = 0.01
 ```{r run_new_steadystates, cache=T, message=F, warning=F}
 # Set time and parameters 
-time <- seq(1,100,1) 
-pars <- replace(defpars, c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
-                          c(60,       60,    1,   5,      0.9,     0.9,       0,       0))
+time <- seq(1,200,0.5) 
+defpars <- def_pars()
+defpars <- replace(defpars, "jHGm", 1)
+pars <- replace(defpars, c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "initS"), 
+                          c(60,       60,    1,   5,      0.9,     0.9,       0,       0,      0.01))
 # Set values of L and N at which to get steady state values 
-at <- expand.grid(L=seq(from=0, to=60, length.out=41),
-                  N=seq(from=0, to=4e-6, length.out=41))
+at <- expand.grid(L=seq(from=0, to=60, length.out=21),
+                  N=seq(from=0, to=4e-6, length.out=21))
+
+# Run to steady state and plot results
+ssdat <- run_steady_states(time=time, pars=pars, at=at, food=0)
+plot_steady_states(ssdat)
+```
+
+## Starting biomass = 0.1
+```{r run_new_steadystates.1, cache=T, message=F, warning=F}
+# Set time and parameters 
+time <- seq(1,200,0.5) 
+defpars <- def_pars()
+defpars <- replace(defpars, "jHGm", 1)
+pars <- replace(defpars, c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "initS"), 
+                          c(60,       60,    1,   5,      0.9,     0.9,       0,       0,      0.1))
+# Set values of L and N at which to get steady state values 
+at <- expand.grid(L=seq(from=0, to=60, length.out=21),
+                  N=seq(from=0, to=4e-6, length.out=21))
+
+# Run to steady state and plot results
+ssdat <- run_steady_states(time=time, pars=pars, at=at, food=0)
+plot_steady_states(ssdat)
+```
+
+## Starting biomass = 0.5
+```{r run_new_steadystates.5, cache=T, message=F, warning=F}
+# Set time and parameters 
+time <- seq(1,200,0.5) 
+defpars <- def_pars()
+defpars <- replace(defpars, "jHGm", 1)
+pars <- replace(defpars, c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "initS"), 
+                          c(60,       60,    1,   5,      0.9,     0.9,       0,       0,      0.5))
+# Set values of L and N at which to get steady state values 
+at <- expand.grid(L=seq(from=0, to=60, length.out=21),
+                  N=seq(from=0, to=4e-6, length.out=21))
+
+# Run to steady state and plot results
+ssdat <- run_steady_states(time=time, pars=pars, at=at, food=0)
+plot_steady_states(ssdat)
+```
+
+```{r testing, cache=T, message=F, warning=F, eval=F, include=F}
+# Set up simulation
+time <- seq(1,200,0.1) 
+env <- init_env(time=time, L=c(2,2,0), N=c(1e-7,3.5e-6,0), X=c(0,0,0))
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
+run <- run_coral_wC(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "initS"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0, 1)),
+                 NPQ="fixed2")
+plot_run_wC(run)
+# Do elevated nutrients prevent coral from recovering from bleaching?
+Start with 0.01 initS, constant light (10), lower DIN (1e-7) --> recovery and pos.
+growth; higher DIN (1e-6) --> no recovery, neg.
+growth...
+```
+
+## Starting symbiont biomass = 1
+```{r run_new_steadystates1, cache=T, message=F, warning=F}
+# Set time and parameters 
+time <- seq(1,200,0.5) 
+pars <- replace(defpars, c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "initS"), 
+                          c(60,       60,    1,   5,      0.9,     0.9,       0,       0,       1))
+# Set values of L and N at which to get steady state values 
+at <- expand.grid(L=seq(from=0, to=60, length.out=21),
+                  N=seq(from=0, to=4e-6, length.out=21))
 
@@
 # Stability at high nitrogen...?
-## When system starts at 2 µM DIN and increases to 4 µM DIN, outcome is negative growth.
+## When system starts at 2 µM DIN and increases to 4 µM DIN, switch is btw. 0.15 and 0.2 initS.
 ```{r incrN1, cache=T, message=F, warning=F}
-# Set up simulation
-env <- init_env(time=time, L=c(10,10,0), N=c(2e-6,4e-6,0), X=c(0,0,0))
-# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
-run <- run_coral_wC(time=time, env=env,
-                 pars=replace(defpars, 
-                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "jHGm"), 
-                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0, 1)),
-                 NPQ="fixed2")
-plot_run_wC(run)
-```
-
-## When system starts at 1 µM DIN and increases to 4 µM DIN, outcome is positive growth.
-```{r incrN2, cache=T, message=F, warning=F}
 # Set up simulation
@@
                  pars=replace(defpars, 
-                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "jHGm"), 
-                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0, 1)),
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "initS"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0, 0.15)),
+                 NPQ="fixed2")
+plot_run_wC(run)
+
+# Set up simulation
+env <- init_env(time=time, L=c(10,10,0), N=c(1e-6,4e-6,0), X=c(0,0,0))
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
+run <- run_coral_wC(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "initS"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0, 0.2)),
                  NPQ="fixed2")
@@
 ```
+
+## When system starts at 2 µM DIN and increases to 4 µM DIN, switch is btw. 0.2 and 0.3 initS.
+```{r incrN2, cache=T, message=F, warning=F}
+# Set up simulation
+env <- init_env(time=time, L=c(10,10,0), N=c(2e-6,4e-6,0), X=c(0,0,0))
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
+run <- run_coral_wC(time=time, env=env, kCO2=5,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "jHGm", "initS"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0, 1, 0.2)),
+                 NPQ="fixed2")
+plot_run_wC(run)
+
+# Set up simulation
+env <- init_env(time=time, L=c(9,9,0), N=c(2e-6,4e-6,0), X=c(0,0,0))
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
+run <- run_coral_wC(time=time, env=env, kCO2=5,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "jHGm", "initS"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0, 1, 0.3)),
+                 NPQ="fixed2")
+plot_run_wC(run)
+```
+
+```{r}
+# Set up simulation
+env <- init_env(time=time, L=c(10,10,0), N=c(4e-6,4e-6,0), X=c(0,0,0))
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
+run <- run_coral_wC(time=time, env=env, kCO2=5,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "jHGm", "initS"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0, 1, 0.5)),
+                 NPQ="fixed2")
+plot_run_wC(run)
+
+# Set up simulation
+env <- init_env(time=time, L=c(10,10,0), N=c(4e-6,4e-6,0), X=c(0,0,0))
+# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
+run <- run_coral_wC(time=time, env=env, kCO2=5,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "jHGm", "initS"), 
+                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0, 1, 0.01)),
+                 NPQ="fixed2")
+plot_run_wC(run)
+with(run, {
+  plot(S$S/H$H, S$jSG)
+})
+
+```

### 2016-10-09 R/def_pars.R
replace jCO2a/jCO2p with kCO2

@@ -9,10 +9,8 @@ def_pars <- function() {
     sigmaCH=0.9,  # Proportion of host carbon turnover recycled (-)
     jXm=0.1292,  # Maximum specific host feeding rate (molX/CmolH/d)
     jNm=0.048,  # Maximum specific host DIN uptake rate (molN/CmolH/d)
-    jCO2p=4.04e-3,  # Specific flux of CO2 into tissue due to passive reaction-diffusion (molC/CmolH/d)
-    jCO2a=0.32,  # Specific delivery of CO2 to photosynthesis by coral host (molC/CmolH/d)
     jHGm=1,  # Maximum specific host growth rate (CmolH/CmolH/d)
-    kCO2=5,
+    kCO2=5,  # Rate of host CCM's (molCO2/molC/d)
     KN=1e-6,  # Half-saturation constant for host DIN uptake (molN/L)
     KX=20e-6,  # Half-saturation constant for host feeding (molX/L)
     initH=1,  # Initial host biomass (CmolH)

### 2016-10-13 R/run_coral_ss.R
steady state checks both growth and S/H

@@ -49,10 +49,11 @@ run_coral_ss <- function(env, pars) {
   
   # Run simulation by updating
   # ==========================
-  ss <- FALSE
+  grss <- FALSE
+  shss <- FALSE
   t <- 2
   
-  while (ss==FALSE) {
+  while (grss==FALSE | shss==FALSE) {
     
     # Photosynthesis
     # ==============
@@ -118,8 +119,13 @@ run_coral_ss <- function(env, pars) {
    
     # Test if steady state has been reached
     if (t > 21) {
-      ss <- ifelse(abs(H$dH.Hdt[t] - H$dH.Hdt[t-20]) < 0.00001, T, F)
+      grss <- ifelse(abs(H$dH.Hdt[t] - H$dH.Hdt[t-20]) < 0.00001, T, F)
+      shss <- ifelse(abs(S$S[t]/H$H[t] - S$S[t-20]/H$H[t-20]) < 0.00001, T, F)
     }
+    # Test if system is oscillating but with negative growth (e.g., L=30 & N=0)
+    if (t > 1000) {
+      if (!any(H$dH.Hdt[(t-1000):t] > 0)) grss <- T; shss <- T
+    } 
     
     # Increment time
     t <- t + 1

### 2016-10-13 R/plot_steady_states.R
upper S/H = 0.5

@@ -6,8 +6,8 @@ plot_steady_states <- function(ss, png=NULL) {
   
   # For runs that did NOT reach a steady state of positive growth, set all results to zero
   ss[which(ss$gr < 0), c("gr", "sh", "hl", "sl", "ee", "pl")] <- NA
-  # For runs where steady state value of S:H > 0.7, set S:H to 0.7
-  ss[which(ss$sh > 0.7), "sh"] <- 0.7
+  # For runs where steady state value of S:H > 0.5, set S:H to 0.5
+  ss[which(ss$sh > 0.5), "sh"] <- 0.5
   
   # Reshape steady state results into matrices for plotting cont
   sh <- acast(ss, L~N, value.var="sh")
@@ -84,6 +84,6 @@ plot_steady_states <- function(ss, png=NULL) {
   imagef(gr, bin=0.01, main="A. Specific growth", col="grayscale")
   imagef(hl, bin=0.25, main="B. Host growth limitation", col=list("#67001F","#053061"))
   imagef(sh, bin=0.05, main="C. Symbiont:host biomass", col="grayscale")
-  imagef(pl, bin=0.25, main="D. Photosynthesis limitation", col=list("#67001F", "gray30"))
+  imagef(pl, bin=0.5, main="D. Photosynthesis limitation", col=list("#67001F", "gray30"))
   if (!is.null(png)) dev.off()
 }

### 2016-10-13 JTB_ms/JTB_ms.Rmd
update new C feedback

@@
 ---
-title: A simplified dynamic bioenergetic model for coral-*Symbiodinium* symbioses
+title: A simplified dynamic bioenergetic model for coral bleaching
 author:
@@
   This is the abstract.
+  Some type of positive feedback is necessary.
+  Compare scenarios of recycling CO2, have photodamage increase maintenance that does recycle CO2.
+  Compare recycled N to symbiont instead of host
+  How much do you have to reduce the light to get recovery? i.e., magnitude of hysteresis?
 
@@
 | $\rho_N$  | Nitrogen shared with the symbiont      | molN CmolH^-1^ d^-1^ | 7 |
+| $j_{eC}$  | Excess carbon used to activate host CCMs| molC CmolH^-1^ d^-1^ |  |
 | $j_L$     | Light absorption rate                  | molph CmolS^-1^ d^-1^| 9 |
@@
 | $K_N$ | Half-saturation constant for DIN uptake by host | 0.46e-6 | molN L^-1^ | 
-| $j_{CO_2}^p$ | Passive CO~2~ delivery to symbiont | 4.04e-3 | molC CmolH^-1^ d^-1^ |
-| $j_{CO_2}^a$ | Active CO~2~ delivery to symbiont | 0.32--18.0 | molC CmolH^-1^ d^-1^ |
+| $k_{CO_2}$ | Efficacy of CO~2~ delivery by host CCMs | 5 | molCO~2~ molC^-1^ |
 | $j_{HGm}$ | Maximum specific growth rate of host | 1 | CmolH CmolH^-1^ d^-1^ |
-| $n_{LC}$ | Quantum yield of photosynthesis | 0.1 | molC mol ph^-1^ |
+| $y_{CL}$ | Quantum yield of photosynthesis | 0.1 | molC mol ph^-1^ |
 | $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | 1.34 | m^2^ CmolS^-1^ |
@@
 | $j_{SGm}$ | Maximum specific growth rate of symbiont | 0.25 | CmolS CmolS^-1^ d^-1^ |
-| $b$ | Scaling parameter for strength of bleaching response | 3 | -- |
+| $b$ | Scaling parameter for strength of bleaching response | 5 | -- |
 
@@
 Parallel complementary synthesizing units (SUs) are represented by large circles, and rejection fluxes from these SUs are indicated by dashed lines.
-$c_{ROS}$ is a proportional rate that impacts other model fluxes by inhibition or acceleration.
+$c_{ROS}$ is a proportional rate that impacts other model fluxes by inhibition or acceleration; likewise, $j_{eC}$ accelerates the rate of $j_{CO_2}$.
 Recycled mass fluxes from biomass turnover are not shown for clarity (but see Table 1 for definitions).](../img/Fig1.png)
@@
 
-\begin{equation} r_{NH}=\sigma_{NH}j_{HT}^0 \end{equation}
+\begin{equation} r_{NH}=\sigma_{NH}n_{NH}j_{HT}^0 \end{equation}
 
@@
 
+The carbon rejected from this SU reflects the amount of excess fixed carbon available to the host that is not used in biomass formation:
+
+\begin{equation} j_{eC} = (jX + \rho_C{S \over H} - j_{HG})_+ \end{equation}
+
+This flux, $j_{eC}$, is assumed to be available as a respiratory substrate to support various energetically-demanding processes; of particular importance is the host's active supply of CO~2~ for symbiont photosynthesis [@Wooldridge:2013tj; @Hopkinson:2015ev] (see below).
+
 Due to the inherent inefficiency of the parallel complementary SU formulation, there is always some nitrogen shared with the symbiont even when coral biomass formation is strongly nitrogen-limited.
-Likewise, there is always a non-zero rejection flux of carbon from the coral biomass SU, which is assumed to be lost to the environment.
+Likewise, there is always a non-zero rejection flux of excess carbon from the coral biomass SU.
 
@@
 
-We then specify two pathways for input of inorganic carbon to the photosynthesis SU: 1) passive diffusion of CO~2~ from the external environment ($j_{CO_2}^p$), and 2) active delivery of CO~2~ to the symbiont by the host ($j_{CO_2}^a$).
-The passive flux ensures that some CO~2~ is always available to photosynthesis, and the active flux encompasses potentially diverse mechanisms by which the host may enhance CO~2~ availability for the symbiont, including active transport of bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to CO~2~ to promote diffusion toward the symbiont [@Tansik:2015cm], and acidification of the symbiosome to increase localized CO~2~ concentrations around the symbiont [@Barott:2014gx].
-Since the host physically separates the symbiont from the external environment, both the passive and active flux rates are proportional to host biomass.
-A more rigorous, mechanistic model of inorganic carbon processing would require spatially explicit internal pools accounting for pH and carbon speciation, which is beyond the current scope of this work.
-Instead, the specification of fixed passive and active CO~2~ input rates, $j_{CO_2}^p$ and $j_{CO_2}^a$, offers the user the opportunity to compare different rates of CO~2~ delivery that may characterize different coral species [@Wooldridge:2014di].
+We then specify the input of CO~2~ to photosynthesis as $j_{CO_2}$, which is mediated by the host.
+This encompasses potentially diverse mechanisms by which the host may enhance CO~2~ availability for photosynthesis (i.e., carbon-concentrating mechanisms (CCMs)), including active transport of bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to CO~2~ to promote diffusion toward the symbiont [@Tansik:2015cm], and acidification of the symbiosome to increase localized CO~2~ concentrations around the symbiont [@Barott:2014gx].
+Since these mechanisms require energetic input by the host, the magnitude of CO~2~ delivery is set proportional to $j_{eC}$, the amount of excess carbon from the host biomass SU, assuming that some of this carbon is respired to energize these CCMs.
+This formulation means that the symbiont indirectly ensures its own CO~2~ supply by providing the host with fixed carbon [@Wooldridge:2013tj], which establishes a positive feedback in the model that proves to be essential in simulating the phenomenon of coral bleaching.
+The parameter $k_{CO_2}$ is included to scale the efficacy of host CCMs, i.e., how many moles of CO~2~ can be delivered from the environment to the symbiont per mole of fixed carbon.
+Adjusting this parameter offers the user the opportunity to compare different rates of CO~2~ delivery that may characterize different coral species [@Wooldridge:2014di].
 The input of CO~2~ to the photosynthesis SU is therefore specified as:
 
-\begin{equation} j_{CO_2} = j_{CO_2}^p + j_{CO_2}^a \end{equation}
+\begin{equation} j_{CO_2} = k_{CO_2}j_{eC} \end{equation}
 
@@
 
-\begin{equation} j_{CP} = \bigg({1 \over j_{CPm}} + {1 \over {n_{LC} j_L }} + {1 \over {(j_{CO_2} + r_{CH}){H \over S} + r_{CS}}} - {1 \over {n_{LC} j_L + (j_{CO_2} + r_{CH}){H \over S} + r_{CS}}}\bigg)^{-1} \cdot c_{ROS}^{-1} \end{equation}
+\begin{equation} j_{CP} = \bigg({1 \over j_{CPm}} + {1 \over {y_{CL} j_L }} + {1 \over {(j_{CO_2} + r_{CH}){H \over S} + r_{CS}}} - {1 \over {y_{CL} j_L + (j_{CO_2} + r_{CH}){H \over S} + r_{CS}}}\bigg)^{-1} \cdot c_{ROS}^{-1} \end{equation}
 
@@
 
-\begin{equation} j_{eL} = (j_L - j_{CP} / n_{LC})_+ \end{equation}
+\begin{equation} j_{eL} = (j_L - j_{CP} / y_{CL})_+ \end{equation}
 
@@
 *Symbiodinium* utilize a variety of pathways for non-photochemical quenching [NPQ; @Roth:2014wf], which we collect in a relative NPQ capacity specified as a parameter of the symbiont ($k_{NPQ}$).
-NPQ capacity may also be susceptible to reduction by photoinhibition, so the total NPQ capacity is specified as:
+The total NPQ flux is specified as a single-substrate SU formula:
 
-\begin{equation} j_{NPQ}= k_{NPQ} \cdot c_{ROS}^{-1} \end{equation}
+\begin{equation} j_{NPQ} = \bigg({1 \over k_{NPQ}} + {1 \over j_{eL}}\bigg)^{-1} \end{equation}
 
-If light energy further exceeds the total capacity of both carbon fixation and NPQ, then reactive oxygen species (ROS) are produced.
+If light energy further exceeds the total capacity of both photochemistry and NPQ, then reactive oxygen species (ROS) are produced.
 We represent this as a relative quantity $c_{ROS}$, which takes a value of 1 when all light absorbed can be quenched by photochemistry and NPQ, and increases as the amount of excess excitation energy further increases:
@@
 
-\begin{equation} r_{NS}=\sigma_{NS}j_{ST}^0 \end{equation}
+\begin{equation} r_{NS}=\sigma_{NS}n_{NS}j_{ST}^0 \end{equation}
 
@@
 
-This second component of symbiont biomass loss can represent both photodamage and/or symbiont expulsion (i.e., bleaching), both of which occur in response to high levels or ROS production.
+This second component of symbiont biomass loss can represent both photodamage and/or symbiont expulsion (i.e., bleaching), both of which occur in response to high levels of ROS production.
 The parameter $b$ is included to scale biomass loss due to bleaching in response to ROS.
@@
 Specifically, the rejection fluxes of carbon and nitrogen from the symbiont and host biomass SUs act as reciprocal input fluxes to the other SU.
-In addition, the rejection flux of excitation energy from the photosynthesis SU acts to reduce its own production flux (i.e., photoinhibition), and hence a discretized time-stepping procedure was necessary.
+In addition, the photosynthesis SU receives CO~2~ proportional to the carbon rejection flux from the host biomass SU, and the rejection flux of excitation energy from the photosynthesis SU acts to reduce its own production (i.e., photoinhibition).
 A vector of time values was created for each simulation run, along which dynamic environmental forcing functions (irradiance, DIN, and prey abundance) can be designed.
@@
 
-# Sensitivity analysis -- should come after the steady states bc describes changes in steady states
+# Results
+## Qualitative dynamics
+In a constant environment, the system ultimately reaches a steady state of exponential growth or decline.
+Under some conditions, alternate steady states can be reached depending on initial values of symbiont and host biomass.
+The mechanism that produces these alternate steady states is the positive feedback between carbon-limitation of the host and CO~2~-limitation of photosynthesis.
+If the host becomes carbon-limited, the efficacy of its CCMs are reduced and photosynthesis becomes progressively more CO~2~-limited.
+In turn, less photosynthate is translocated to the host, causing further carbon-limitation and an ultimate decline in biomass; this steady state represents a dysfunctional symbiosis.
+However, if the host is not carbon-limited, it effectively delivers CO~2~ to photosynthesis and ensures its continued receipt of fixed carbon from the symbiont; this system is nitrogen-limited and represents a functional symbiosis with positive exponential growth.
+Either of these stable states may be reached under a constant environment depending on the initial values of symbiont and host biomass: if symbiont biomass is initially very low (i.e., a "bleached" coral), the system cannot escape the carbon-limited state and cannot grow.
+However, if symbiont biomass is initially high (i.e., a healthy coral), then the system maintains positive growth in a nitrogen-limited state.
+For practical purposes, we consider only positive growth states in the "steady state behavior" section below.
+In the "dynamic behavior" section, we explore the conditions that cause the system to switch from one state to another, which in a biological context indicates coral bleaching.
 
-Each default parameter value used in the model (Table 2) is based on carefully selected relevant literature.
-A supplementary document detailing the derivations of each of these default parameter values is provided.
-Nevertheless, it is important to evaluate the sensitivity of the model to changes in these parameter values.
-We measured fractional change in steady state values in response to fractional change in parameter values, relative to their default values (Table 2).
-
-![Sensitivity analysis.
-Plots show the fractional change in steady state values in response to fractional changes in default parameter values (see Table 2 for default values).
-Parameters are grouped by which processes they are involved in.
-Sensitivity in response to changes in each parameter was evaluated in each of four environments (indicated by line type): 1) LLLN=low light (2 mol ph m^-2^ d^-1^) and low DIN (1e-7 mol L^-1^), 2) LLHN=low light (2 mol ph m^-2^ d^-1^) and high DIN (4e-6 mol L^-1^), 3) HLLN=high light (30 mol ph m^-2^ d^-1^) and low DIN (1e-7 mol L^-1^), and 4) HLHN=high light (30 mol ph m^-2^ d^-1^) and high DIN (4e-6 mol L^-1^).](../img/Fig3.png)
-
-# Steady state behavior
+## Steady state behavior
 
@@
 
-# Dynamic behavior
+## Sensitivity analysis -- should come after the steady states bc describes changes in steady states
+
+Each default parameter value used in the model (Table 2) is based on carefully selected relevant literature.
+A supplementary document detailing the derivations of each of these default parameter values is provided.
+Nevertheless, it is important to evaluate the sensitivity of the model to changes in these parameter values.
+We measured fractional change in steady state values in response to fractional change in parameter values, relative to their default values (Table 2).
+
+![Sensitivity analysis.
+Plots show the fractional change in steady state values in response to fractional changes in default parameter values (see Table 2 for default values).
+Parameters are grouped by which processes they are involved in.
+Sensitivity in response to changes in each parameter was evaluated in each of four environments (indicated by line type): 1) LLLN=low light (2 mol ph m^-2^ d^-1^) and low DIN (1e-7 mol L^-1^), 2) LLHN=low light (2 mol ph m^-2^ d^-1^) and high DIN (4e-6 mol L^-1^), 3) HLLN=high light (30 mol ph m^-2^ d^-1^) and low DIN (1e-7 mol L^-1^), and 4) HLHN=high light (30 mol ph m^-2^ d^-1^) and high DIN (4e-6 mol L^-1^).](../img/Fig3.png)
+
+## Dynamic behavior
+Therefore, we are highly interested in the conditions under which the system switches from one state to another, representing the conditions under which corals "bleach" due to progressive, reinforcing carbon-limitation.
+Interestingly, this state can be reached in multiple ways, by the acceleration of photoinhibition and photodamage, or by high external DIN, which causes the symbiont to share less carbon, subsequently causing its own CO2-limitation, and progressive C-limitation of the system.
 ## varying light: consequences of excess light and light-limitation

### 2016-10-13 JTB_ms/JTB_ms.tex
update new C feedback

@@
             pdfauthor={},
-            pdftitle={A simplified dynamic bioenergetic model for coral-Symbiodinium symbioses},
+            pdftitle={A simplified dynamic bioenergetic model for coral bleaching},
             colorlinks=true,
@@
 
-  \title{A simplified dynamic bioenergetic model for coral-\emph{Symbiodinium}
-symbioses}
+  \title{A simplified dynamic bioenergetic model for coral bleaching}
     \author[University of Hawaii]{Ross Cunning\corref{c1}}
@@
   \begin{abstract}
-  This is the abstract.
+  This is the abstract. Some type of positive feedback is necessary.
+  Compare scenarios of recycling CO2, have photodamage increase
+  maintenance that does recycle CO2. Compare recycled N to symbiont
+  instead of host How much do you have to reduce the light to get
+  recovery? i.e., magnitude of hysteresis?
   \end{abstract}
@@
 \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(j_{eC}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Excess carbon used to activate host CCMs
+\strut\end{minipage} &
+\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+molC CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(j_L\)
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{CO_2}^p\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Passive CO\textsubscript{2} delivery to symbiont
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-4.04e-3
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-molC CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{CO_2}^a\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Active CO\textsubscript{2} delivery to symbiont
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-0.32--18.0
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-molC CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\(k_{CO_2}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Efficacy of CO\textsubscript{2} delivery by host CCMs
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+5
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+molCO\textsubscript{2} molC\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(n_{LC}\)
+\(y_{CL}\)
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-3
+5
 \strut\end{minipage} &
@@
 proportional rate that impacts other model fluxes by inhibition or
-acceleration. Recycled mass fluxes from biomass turnover are not shown
-for clarity (but see Table 1 for definitions).}
+acceleration; likewise, \(j_{eC}\) accelerates the rate of \(j_{CO_2}\).
+Recycled mass fluxes from biomass turnover are not shown for clarity
+(but see Table 1 for definitions).}
 \end{figure}
@@
 
+The carbon rejected from this SU reflects the amount of excess fixed
+carbon available to the host that is not used in biomass formation:
+
+\begin{equation} j_{eC} = (jX + \rho_C{S \over H} - j_{HG})_+ \end{equation}
+
+This flux, \(j_{eC}\), is assumed to be available as a respiratory
+substrate to support various energetically-demanding processes; of
+particular importance is the host's active supply of CO\textsubscript{2}
+for symbiont photosynthesis (Wooldridge 2013; Hopkinson, Tansik, and
+Fitt 2015) (see below).
+
 Due to the inherent inefficiency of the parallel complementary SU
@@
 when coral biomass formation is strongly nitrogen-limited. Likewise,
-there is always a non-zero rejection flux of carbon from the coral
-biomass SU, which is assumed to be lost to the environment.
+there is always a non-zero rejection flux of excess carbon from the
+coral biomass SU.
 
@@
 
-We then specify two pathways for input of inorganic carbon to the
-photosynthesis SU: 1) passive diffusion of CO\textsubscript{2} from the
-external environment (\(j_{CO_2}^p\)), and 2) active delivery of
-CO\textsubscript{2} to the symbiont by the host (\(j_{CO_2}^a\)). The
-passive flux ensures that some CO\textsubscript{2} is always available
-to photosynthesis, and the active flux encompasses potentially diverse
-mechanisms by which the host may enhance CO\textsubscript{2}
-availability for the symbiont, including active transport of
+We then specify the input of CO\textsubscript{2} to photosynthesis as
+\(j_{CO_2}\), which is mediated by the host. This encompasses
+potentially diverse mechanisms by which the host may enhance
+CO\textsubscript{2} availability for photosynthesis (i.e.,
+carbon-concentrating mechanisms (CCMs)), including active transport of
 bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to
@@
 increase localized CO\textsubscript{2} concentrations around the
-symbiont (Barott et al. 2014). Since the host physically separates the
-symbiont from the external environment, both the passive and active flux
-rates are proportional to host biomass. A more rigorous, mechanistic
-model of inorganic carbon processing would require spatially explicit
-internal pools accounting for pH and carbon speciation, which is beyond
-the current scope of this work. Instead, the specification of fixed
-passive and active CO\textsubscript{2} input rates, \(j_{CO_2}^p\) and
-\(j_{CO_2}^a\), offers the user the opportunity to compare different
-rates of CO\textsubscript{2} delivery that may characterize different
-coral species (Wooldridge 2014a). The input of CO\textsubscript{2} to
-the photosynthesis SU is therefore specified as:
-
-\begin{equation} j_{CO_2} = j_{CO_2}^p + j_{CO_2}^a \end{equation}
+symbiont (Barott et al. 2014). Since these mechanisms require energetic
+input by the host, the magnitude of CO\textsubscript{2} delivery is set
+proportional to \(j_{eC}\), the amount of excess carbon from the host
+biomass SU, assuming that some of this carbon is respired to energize
+these CCMs. This formulation means that the symbiont indirectly ensures
+its own CO\textsubscript{2} supply by providing the host with fixed
+carbon (Wooldridge 2013), which establishes a positive feedback in the
+model that proves to be essential in simulating the phenomenon of coral
+bleaching. The parameter \(k_{CO_2}\) is included to scale the efficacy
+of host CCMs, i.e., how many moles of CO\textsubscript{2} can be
+delivered from the environment to the symbiont per mole of fixed carbon.
+Adjusting this parameter offers the user the opportunity to compare
+different rates of CO\textsubscript{2} delivery that may characterize
+different coral species (Wooldridge 2014a). The input of
+CO\textsubscript{2} to the photosynthesis SU is therefore specified as:
+
+\begin{equation} j_{CO_2} = k_{CO_2}j_{eC} \end{equation}
 
@@
 quenching (NPQ; Roth 2014), which we collect in a relative NPQ capacity
-specified as a parameter of the symbiont (\(k_{NPQ}\)). NPQ capacity may
-also be susceptible to reduction by photoinhibition, so the total NPQ
-capacity is specified as:
-
-\begin{equation} j_{NPQ}= k_{NPQ} \cdot c_{ROS}^{-1} \end{equation}
+specified as a parameter of the symbiont (\(k_{NPQ}\)). The total NPQ
+flux is specified as a single-substrate SU formula:
+
+\begin{equation} j_{NPQ} = \bigg({1 \over k_{NPQ}} + {1 \over j_{eL}}\bigg)^{-1} \end{equation}
 
@@
 
-\section{Sensitivity analysis}\label{sensitivity-analysis}
-
-Each default parameter value used in the model (Table 2) is based on
-carefully selected relevant literature. A supplementary document
-detailing the derivations of each of these default parameter values is
-provided. Nevertheless, it is important to evaluate the sensitivity of
-the model to changes in these parameter values. We measured fractional
-change in steady state values in response to fractional change in
-parameter values, relative to their default values (Table 2).
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{../img/Fig3.png}
-\caption{Sensitivity analysis. Plots show the fractional change in
-steady state values in response to fractional changes in default
-parameter values (see Table 2 for default values). Parameters are
-grouped by which processes they are involved in. Sensitivity in response
-to changes in each parameter was evaluated in each of four environments
-(indicated by line type): 1) LLLN=low light (2 mol ph
-m\textsuperscript{-2} d\textsuperscript{-1}) and low DIN (1e-7 mol
-L\textsuperscript{-1}), 2) LLHN=low light (2 mol ph
-m\textsuperscript{-2} d\textsuperscript{-1}) and high DIN (4e-6 mol
-L\textsuperscript{-1}), 3) HLLN=high light (30 mol ph
-m\textsuperscript{-2} d\textsuperscript{-1}) and low DIN (1e-7 mol
-L\textsuperscript{-1}), and 4) HLHN=high light (30 mol ph
-m\textsuperscript{-2} d\textsuperscript{-1}) and high DIN (4e-6 mol
-L\textsuperscript{-1}).}
-\end{figure}
-
-\section{Steady state behavior}\label{steady-state-behavior}
-
-To analyze qualitative behavior, we ran the model to steady state across
-gradients of external irradiance and nutrients (Fig. 2), which revealed
-patterns consistent with observed phenomena in corals. Predicted growth
-rates are low at low light and DIN, and begin increasing as both of
-these factors increase (Fig. 2A). Low light is limiting to
-photosynthesis (Fig. 2D), which leads to carbon-limitation of host
-growth (Fig. 2B) and an associated increase in the symbiont to host
-biomass ratio (Fig. 2C). In agreement with this trend are many
-observations of negative correlation between irradiance and symbiont
-density (Stimson 1997; Brown et al. 1999; Fitt et al. 2000; Titlyanov et
-al. 2001). As higher light alleviates light-limitation of
+\section{Results}\label{results}
+
+\subsection{Qualitative dynamics}\label{qualitative-dynamics}
+
+simulations in constant environs show either positive or negative
+exponential growth. some environments show two steady states, one with
+positive and one with negative growth, depending on initial conditions.
+based on numerical analysis, we see ultimate exponential growth or
+decline. the mechanism that produces these alternate steady states is:
+\ldots{}. . for practical purposes, we are interested in the positive
+growth steady states and their properties and how they change across
+environmental gradients (see section X). but we are also interested in
+the possibility for flips to the alternate steady states (see section
+Y).
+
+\subsection{Steady state behavior}\label{steady-state-behavior}
+
+\emph{emphasize that steady state is exponential growth, and fluxes are
+per unit biomass/specific rates} To analyze qualitative behavior, we ran
+the model to steady state across gradients of external irradiance and
+nutrients (Fig. 2), which revealed patterns consistent with observed
+phenomena in corals. Predicted growth rates are low at low light and
+DIN, and begin increasing as both of these factors increase (Fig. 2A).
+Low light is limiting to photosynthesis (Fig. 2D), which leads to
+carbon-limitation of host growth (Fig. 2B) and an associated increase in
+the symbiont to host biomass ratio (Fig. 2C). In agreement with this
+trend are many observations of negative correlation between irradiance
+and symbiont density (Stimson 1997; Brown et al. 1999; Fitt et al. 2000;
+Titlyanov et al. 2001). As higher light alleviates light-limitation of
 photosynthesis, growth becomes less carbon-limited. Similarly, elevating
@@
 
-\section{Dynamic behavior}\label{dynamic-behavior}
+\subsection{Sensitivity analysis -- should come after the steady states
+bc describes changes in steady
+states}\label{sensitivity-analysis-should-come-after-the-steady-states-bc-describes-changes-in-steady-states}
+
+Each default parameter value used in the model (Table 2) is based on
+carefully selected relevant literature. A supplementary document
+detailing the derivations of each of these default parameter values is
+provided. Nevertheless, it is important to evaluate the sensitivity of
+the model to changes in these parameter values. We measured fractional
+change in steady state values in response to fractional change in
+parameter values, relative to their default values (Table 2).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{../img/Fig3.png}
+\caption{Sensitivity analysis. Plots show the fractional change in
+steady state values in response to fractional changes in default
+parameter values (see Table 2 for default values). Parameters are
+grouped by which processes they are involved in. Sensitivity in response
+to changes in each parameter was evaluated in each of four environments
+(indicated by line type): 1) LLLN=low light (2 mol ph
+m\textsuperscript{-2} d\textsuperscript{-1}) and low DIN (1e-7 mol
+L\textsuperscript{-1}), 2) LLHN=low light (2 mol ph
+m\textsuperscript{-2} d\textsuperscript{-1}) and high DIN (4e-6 mol
+L\textsuperscript{-1}), 3) HLLN=high light (30 mol ph
+m\textsuperscript{-2} d\textsuperscript{-1}) and low DIN (1e-7 mol
+L\textsuperscript{-1}), and 4) HLHN=high light (30 mol ph
+m\textsuperscript{-2} d\textsuperscript{-1}) and high DIN (4e-6 mol
+L\textsuperscript{-1}).}
+\end{figure}
+
+\subsection{Dynamic behavior}\label{dynamic-behavior}
 

### 2016-10-13 Rmd/analyze.Rmd
update

--- 2016-09-23 Rmd/analyze.Rmd
@@
 ```{r initialize, include=FALSE}
-# Initialize SU function
-synth <- function(x, y, m) 1/((1/m)+(1/x)+(1/y)-(1/(x+y)))
 # Source default parameters and functions
 sapply(c("R/conversions.R", "R/def_pars.R", "R/init_env.R", "R/plot_results.R",
-         "R/run_coral.R", "R/sym_PI.R"), source, .GlobalEnv)
+         "R/run_coral.R", "R/run_coral_ss.R", "R/sym_PI.R", "R/run_steady_states.R", "R/plot_steady_states.R",
+         "R/explore_NPQ.R"), source, .GlobalEnv)
 ```
 
-```{r lightamp, fig.width=5, fig.height=5, eval=F}
+```{r lightamp, fig.width=5, fig.height=5, eval=T, include=T}
 # Import data from Marcelino et al. 2013
@@
 
-### Run 1
-```{r run1}
+### Steady states values with default parameters - no feeding
+```{r ss, cache=T, fig.width=5, fig.height=4.4}
+# Set parameters
+defpars <- def_pars()  # Get default parameters
+
+# Set values of L and N at which to get steady state values
+at <- expand.grid(L=seq(from=0, to=60, length.out=21),
+                  N=seq(from=0, to=4e-6, length.out=21))
+
+# Run to steady state and plot results
+ssdat <- run_steady_states(pars=replace(defpars, c("initS", "kROS"), c(1, 100)), at=at, food=0, runtime=T)
+plot_steady_states(ssdat)
+```
+With default parameters, steady state values of specific growth are higher than usually observed in corals (skeletal growth), but patterns across environments are consistent with literature.
+
+### Steady states values with default parameters + feeding
+```{r ss_feeding, cache=T, fig.width=5, fig.height=4.4}
+# Set parameters
+defpars <- def_pars()  # Get default parameters
+
+# Set values of L and N at which to get steady state values
+at <- expand.grid(L=seq(from=0, to=60, length.out=21),
+                  N=seq(from=0, to=4e-6, length.out=21))
+
+# Run to steady state and plot results
+ssdat <- run_steady_states(pars=replace(defpars, c("initS", "kROS"), c(1, 100)), at=at, food=5e-6, runtime=T)
+plot_steady_states(ssdat)
+```
+With default parameters, steady state values of specific growth are higher than usually observed in corals (skeletal growth), but patterns across environments are consistent with literature.
+
+### Steady states values with default parameters - + tolerant symbiont
+```{r ss_tolerant, cache=T, fig.width=5, fig.height=4.4}
+# Set parameters
+defpars <- def_pars()  # Get default parameters
+
+# Set values of L and N at which to get steady state values
+at <- expand.grid(L=seq(from=0, to=60, length.out=21),
+                  N=seq(from=0, to=4e-6, length.out=21))
+
+# Run to steady state and plot results
+ssdat <- run_steady_states(pars=replace(defpars, c("initS", "kROS"), c(1, 200)), at=at, food=0, runtime=T)
+plot_steady_states(ssdat)
+```
+With some feeding (more realistic), can maintain growth with very low symbionts (bleached) at very high or low light with high nutrients.
+
+### Vary Light
+```{r varyL, eval=F, include=F}
 # Set run time vector
-time <- seq(1, 365, 0.1)  # if single model run, use time input
+time <- seq(1, 365, 0.5)  # if single model run, use time input
 
 # Initialize environment
-env <- init_env(time=time, L=c(20,32,2), N=c(1e-6,1e-6,0), X=c(0,0,0))
+env <- init_env(time=time, L=c(5,30,2), N=c(1e-7,1e-7,0), X=c(5e-6,5e-6,0))
 
 # Set parameters
-pars <- def_pars()  # Get default parameters
-#pars$rpars$gammaR <- 0.03  # Change any default parameters
-#pars$jCO2a <- 1
-#pars$jCO2a <- 0.1
+defpars <- def_pars()  # Get default parameters
 
 # Run simulation
-run <- run_coral(time=time, env=env, pars=pars)
+run <- run_coral(time=time, env=env, pars=replace(defpars, "initS", 0.2))#
+plot_run(run)
+```
+
+### Vary nutrients
+```{r varyN, eval=F, include=F}
+# Set run time vector
+time <- seq(1, 365, 0.5)  # if single model run, use time input
+
+# Initialize environment
+env <- init_env(time=time, L=c(20,20,2), N=c(1e-7,4e-6,2), X=c(5e-6,5e-6,0))
+
+# Set parameters
+defpars <- def_pars()  # Get default parameters
+
+# Run simulation
+run <- run_coral(time=time, env=env, pars=replace(defpars, "initS", 0.25))#
+plot_run(run)
+```
+
+### Vary nutrients @ very low light
+```{r varyN_lowL, eval=F, include=F}
+# Set run time vector
+time <- seq(1, 365, 0.5)  # if single model run, use time input
+
+# Initialize environment
+env <- init_env(time=time, L=c(5,5,2), N=c(1e-7,4e-6,2), X=c(5e-6,5e-6,0))
+
+# Set parameters
+defpars <- def_pars()  # Get default parameters
+
+# Run simulation
+run <- run_coral(time=time, env=env, pars=replace(defpars, "initS", 1))#
+plot_run(run)
+```
+
+### Steadily increasing light
+```{r incrL, eval=F, include=F}
+# Set run time vector
+time <- seq(1, 365, 0.5)  # if single model run, use time input
+
+# Initialize environment
+env <- init_env(time=time, L=c(30,60,0), N=c(1e-7,1e-7,2), X=c(5e-6,5e-6,0))
+
+# Set parameters
+defpars <- def_pars()  # Get default parameters
+
+# Run simulation
+run <- run_coral(time=time, env=env, pars=replace(defpars, "initS", 0.1))#
+plot_run(run)
+```
+
+### Steadily increasing light, + high N
+```{r incrL_highN, eval=F, include=F}
+# Set run time vector
+time <- seq(1, 365, 0.5)  # if single model run, use time input
+
+# Initialize environment
+env <- init_env(time=time, L=c(15,60,0), N=c(2e-6,2e-6,2), X=c(5e-6,5e-6,0))
+
+# Set parameters
+defpars <- def_pars()  # Get default parameters
+
+# Run simulation
+run <- run_coral(time=time, env=env, pars=replace(defpars, "initS", 0.5))#
+plot_run(run)
+```
+
+### Both on same plot
+```{r incrL_multiline, eval=F, include=F}
+time <- seq(1, 365, 0.5)  # Set time
+defpars <- def_pars()  # Get default parameters
+
+# Run 1
+env1 <- init_env(time=time, L=c(15,60,0), N=c(1e-7,1e-7,2), X=c(5e-6,5e-6,0))
+run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", 0.1))
+# Run 2
+env2 <- init_env(time=time, L=c(15,60,0), N=c(2e-6,2e-6,2), X=c(5e-6,5e-6,0))
+run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", 0.5))
+# Run 3
+env3 <- init_env(time=time, L=c(15,60,0), N=c(4e-6,4e-6,2), X=c(5e-6,5e-6,0))
+run3 <- run_coral(time=time, env=env3, pars=replace(defpars, "initS", 0.5))
+# Run 4
+env4 <- init_env(time=time, L=c(15,60,0), N=c(1e-7,1e-7,2), X=c(0,0,0))
+run4 <- run_coral(time=time, env=env4, pars=replace(defpars, "initS", 0.2))
+
+# Plot
+plot(NA, xlim=range(time), ylim=c(0,0.5))
+with(run1, lines(time, S$S/H$H, lty=1))
+with(run2, lines(time, S$S/H$H, lty=2))
+with(run3, lines(time, S$S/H$H, lty=3))
+with(run4, lines(time, S$S/H$H, lty=4))
+```
+
+```{r incrL_multiline1, include=F, eval=F}
+time <- seq(1, 200, 0.5)  # Set time
+defpars <- def_pars()  # Get default parameters
+defpars <- replace(defpars, "kNPQ", 112)
+
+# Run 1
+env1 <- init_env(time=time, L=c(15,40,2), N=c(1e-7,1e-7,2), X=c(5e-6,5e-6,0))
+run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", 0.12))
+# Run 2 - increase food
+env2 <- init_env(time=time, L=c(15,40,2), N=c(1e-7,1e-7,2), X=c(10e-6,10e-6,0))
+run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", 0.12))
+# Run 3 - increase DIN
+env3 <- init_env(time=time, L=c(15,40,2), N=c(2e-6,2e-6,2), X=c(10e-6,10e-6,0))
+run3 <- run_coral(time=time, env=env3, pars=replace(defpars, "initS", 0.25))
+
+# Plot
+plot(time, env1$L, col=alpha("gold", 0.5), type="l", lwd=2, axes=F)
+par(new=T)
+plot(NA, xlim=range(time), ylim=c(0,0.3), yaxs="i")
+with(run1, lines(time, S$S/H$H, lty=1))
+with(run2, lines(time, S$S/H$H, lty=2))
+with(run3, lines(time, S$S/H$H, lty=3))
+
 ```
@@
 ##### Plot run 1 symbiont PI curve
-```{r run1_PI, fig.height=1}
-#sym_PI(pars)
+```{r run1_PI, cache=T, fig.height=3}
+sym_PI(pars=defpars)
 ```
@@
 ##### Plot run 1 output
-```{r run1_plots}
+```{r abb, fig.height=8, fig.width=3, eval=F, include=F}
+time <- seq(1,365,0.5)
+env <- init_env(time=time, L=c(5,40,2), N=c(1e-7,1e-7,0), X=c(4e-6,4e-6,0))
+run <- run_coral(time=time, env=env, 
+                 pars=replace(defpars, c("kNPQ", "kROS", "k", "b", "jCO2a"), c(30, 10, 1, 5, 0.32)),
+                 NPQ="fixed2", photoinhibition=T)
+plot_run2(run)
+env <- init_env(time=time, L=c(10,40,2), N=c(1e-7,1e-7,0), X=c(4e-6,4e-6,0))
+run <- run_coral_jCO2a(time=time, env=env,
+                 pars=replace(defpars, c("kNPQ", "kROS", "k", "b"), c(80, 80, 1, 10)),
+                 NPQ="fixed2")
+run$H$jCO2a
+run$S$jCP * run$S$S / run$H$H
+plot_run2(run)
+
+env <- init_env(time=time, L=c(40,40,0), N=c(1e-7,1e-7,0), X=c(5e-5,5e-5,0))
+run <- run_coral_jCO2a_manual(time=time, env=env, jCO2a_vals=c(3,1),
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH"), 
+                              c(60,       40,    1,   5,      0,         0)),
+                 NPQ="fixed2")
+plot_run2(run)
+
+##### VARY NITROGEN
+# RUN USING WASTED CARBON AND NO JCO2A
+time <- seq(1,1230,1)
+env <- init_env(time=time, L=c(15,15,2), N=c(1e-7,1e-6,2), X=c(5e-6,5e-6,0))
+run <- run_coral_wC(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
+                              c(40,       30,    1,   5,      0.9,         0.9, 0)),
+                 NPQ="fixed2")
 plot_run(run)
+
+# RUN USING JCO2A (original version of model)
+run <- run_coral(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
+                              c(40,       30,    1,   5,      0.9,         0.9, 0.32)),
+                 NPQ="fixed2")
+plot_run(run)
+
+##### VARY LIGHT
+# RUN USING WASTED CARBON AND NO JCO2A
+time <- seq(1,1230,1)
+env <- init_env(time=time, L=c(10,25,2), N=c(1e-7,1e-7,2), X=c(5e-6,5e-6,0))
+run <- run_coral_wC(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
+                              c(40,       30,    1,   5,      0.9,         0.9, 0)),
+                 NPQ="fixed2")
+plot_run_wC(run)
+
+# RUN USING JCO2A (original version of model)
+run <- run_coral(time=time, env=env,
+                 pars=replace(defpars, 
+                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
+                              c(40,       30,    1,   5,      0.9,         0.9, 0.32)),
+                 NPQ="fixed2")
+plot_run(run)
+```
+
+##### Run 2 -- increase kROS to 100 and b to 5
+```{r, fig.height=7, fig.width=3, eval=F, include=F}
+run <- run_coral(time=time, env=env, pars=replace(defpars, c("kROS", "b"), c(100, 5)))
+plot_run2(run)#
+```
+
+##### Run 2 -- increase kROS to 100 and b to 10
+```{r, fig.height=7, fig.width=3, eval=F, include=F}
+run <- run_coral(time=time, env=env, pars=replace(defpars, c("kROS", "b"), c(100, 10)))
+plot_run2(run)#
+```
+
+##### Run 2 -- fluctuate DIN
+```{r run2_plot, fig.height=7, fig.width=3, eval=F, include=F}
+env2 <- init_env(time=time, L=c(20,20,0), N=c(1e-7,4e-6,2), X=c(0,0,0))
+run2 <- run_coral(time=time, env=env2, pars=pars)
+#plot_run2(run2)
+par(mar=c(2,2,1,1), oma=c(0,0,0,0), mfrow=c(6,1), mgp=c(1,0,0), tck=0.025, lwd=1, xaxs="i", 
+    cex.main=1.2, cex.axis=1, cex.lab=1)
+plot_DIN(run2)
+plot_sh(run2)
+plot_gr(run2)
+plot_Lq(run2)
+plot_pl(run2)
+plot_bl(run2)
+
+```
+
+##### Run 3 -- fluctuate light with higher DIN
+```{r run3_plot, fig.height=3, fig.width=3, eval=F, include=F}
+# Set run time vector
+#time <- seq(1, 365, 0.5)  # if single model run, use time input
+
+# Initialize environment
+env <- init_env(time=time, L=c(2,40,2), N=c(4e-6,4e-6,0), X=c(0,0,0))
+
+# Run simulation
+run3 <- run_coral(time=time, env=env, pars=defpars)
+par(mar=c(2,2,1,1), oma=c(0,0,0,0), mfrow=c(2,1), mgp=c(1,0,0), tck=0.025, lwd=1, xaxs="i", 
+    cex.main=1.2, cex.axis=1, cex.lab=1)
+plot_sh(run)  # run 1 -- lower DIN
+plot_sh(run3)  # run3 -- higher DIN
 

### 2016-10-13 Rmd/analyze.Rmd
update

@@
 
+![](../img/model_new.png)
+
+### Light amplification & self-shading
 ```{r lightamp, fig.width=5, fig.height=5, eval=T, include=T}
@@
 
+### PI curve
+```{r run1_PI, cache=T, fig.height=3}
+sym_PI(pars=defpars)
+```
+
 ### Steady states values with default parameters - no feeding
@@
 # Run to steady state and plot results
-ssdat <- run_steady_states(pars=replace(defpars, c("initS", "kROS"), c(1, 100)), at=at, food=0, runtime=T)
+ssdat <- run_steady_states(pars=replace(defpars, c("initS", "kCO2"), c(1, 10)), at=at, food=0, runtime=T)
 plot_steady_states(ssdat)
@@
 
-### Steady states values with default parameters + feeding
-```{r ss_feeding, cache=T, fig.width=5, fig.height=4.4}
+```{r ss_feeding, cache=T, fig.width=5, fig.height=4.4, eval=F, include=F}
 # Set parameters
@@
 # Run to steady state and plot results
-ssdat <- run_steady_states(pars=replace(defpars, c("initS", "kROS"), c(1, 100)), at=at, food=5e-6, runtime=T)
+ssdat <- run_steady_states(pars=replace(defpars, c("initS", "kCO2"), c(1, 10)), at=at, food=5e-6, runtime=T)
 plot_steady_states(ssdat)
@@
 # Run to steady state and plot results
-ssdat <- run_steady_states(pars=replace(defpars, c("initS", "kROS"), c(1, 200)), at=at, food=0, runtime=T)
+ssdat <- run_steady_states(pars=replace(defpars, c("initS", "kROS", "kCO2"), c(1, 120, 10)), at=at, food=0, runtime=T)
 plot_steady_states(ssdat)
@@
 
-### Vary Light
-```{r varyL, eval=F, include=F}
+### Dynamic behavior: nutrients
+```{r varyN, eval=F, include=F}
 # Set run time vector
@@
 # Initialize environment
-env <- init_env(time=time, L=c(5,30,2), N=c(1e-7,1e-7,0), X=c(5e-6,5e-6,0))
+env <- init_env(time=time, L=c(20,20,2), N=c(1e-7,4e-6,2), X=c(0,0,0))
 
@@
 # Run simulation
-run <- run_coral(time=time, env=env, pars=replace(defpars, "initS", 0.2))#
+run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "kCO2"), c(0.25, 10)))#
 plot_run(run)
@@
 
-### Vary nutrients
-```{r varyN, eval=F, include=F}
-# Set run time vector
-time <- seq(1, 365, 0.5)  # if single model run, use time input
-
-# Initialize environment
-env <- init_env(time=time, L=c(20,20,2), N=c(1e-7,4e-6,2), X=c(5e-6,5e-6,0))
-
-# Set parameters
-defpars <- def_pars()  # Get default parameters
-
-# Run simulation
-run <- run_coral(time=time, env=env, pars=replace(defpars, "initS", 0.25))#
-plot_run(run)
-```
-
-### Vary nutrients @ very low light
 ```{r varyN_lowL, eval=F, include=F}
@@
 
-### Steadily increasing light
-```{r incrL, eval=F, include=F}
+### Dynamic behavior: light
+```{r varyL, eval=T, include=T, fig.width=5}
+# Set run time vector
+time <- seq(1, 365, 0.5)  # if single model run, use time input
+
+# Initialize environment
+env <- init_env(time=time, L=c(5,30,2), N=c(1e-7,1e-7,0), X=c(5e-6,5e-6,0))
+
+# Set parameters
+defpars <- def_pars()  # Get default parameters
+
+# Run simulation
+run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "kCO2"), c(0.2, 10)))#
+plot_run(run)
+```
+
+### Dynamic behavior: steadily increasing light
+```{r incrL, eval=T, include=T, fig.width=5}
 # Set run time vector
@@
 # Run simulation
-run <- run_coral(time=time, env=env, pars=replace(defpars, "initS", 0.1))#
+run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "kCO2"), c(0.1, 10)))#
 plot_run(run)
@@
 
-### Steadily increasing light, + high N
-```{r incrL_highN, eval=F, include=F}
+### Dynamic behavior: steadily increasing light, + high N
+```{r incrL_highN, eval=T, include=T, fig.width=5}
 # Set run time vector
@@
 # Initialize environment
-env <- init_env(time=time, L=c(15,60,0), N=c(2e-6,2e-6,2), X=c(5e-6,5e-6,0))
+env <- init_env(time=time, L=c(30,60,0), N=c(2e-6,2e-6,2), X=c(5e-6,5e-6,0))
 
@@
 # Run simulation
-run <- run_coral(time=time, env=env, pars=replace(defpars, "initS", 0.5))#
+run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "kCO2"), c(0.15, 10)))#
 plot_run(run)
@@
 
-### Both on same plot
-```{r incrL_multiline, eval=F, include=F}
-time <- seq(1, 365, 0.5)  # Set time
+### Dynamic behavior: high light
+```{r varyHL, eval=T, include=T, fig.width=5}
+# Set run time vector
+time <- seq(1, 365, 0.5)  # if single model run, use time input
+
+# Initialize environment
+env <- init_env(time=time, L=c(5,45,2), N=c(1e-7,1e-7,0), X=c(5e-6,5e-6,0))
+
+# Set parameters
 defpars <- def_pars()  # Get default parameters
 
-# Run 1
-env1 <- init_env(time=time, L=c(15,60,0), N=c(1e-7,1e-7,2), X=c(5e-6,5e-6,0))
-run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", 0.1))
-# Run 2
-env2 <- init_env(time=time, L=c(15,60,0), N=c(2e-6,2e-6,2), X=c(5e-6,5e-6,0))
-run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", 0.5))
-# Run 3
-env3 <- init_env(time=time, L=c(15,60,0), N=c(4e-6,4e-6,2), X=c(5e-6,5e-6,0))
-run3 <- run_coral(time=time, env=env3, pars=replace(defpars, "initS", 0.5))
-# Run 4
-env4 <- init_env(time=time, L=c(15,60,0), N=c(1e-7,1e-7,2), X=c(0,0,0))
-run4 <- run_coral(time=time, env=env4, pars=replace(defpars, "initS", 0.2))
-
-# Plot
-plot(NA, xlim=range(time), ylim=c(0,0.5))
-with(run1, lines(time, S$S/H$H, lty=1))
-with(run2, lines(time, S$S/H$H, lty=2))
-with(run3, lines(time, S$S/H$H, lty=3))
-with(run4, lines(time, S$S/H$H, lty=4))
+# Run simulation
+run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "kCO2"), c(0.2, 10)))#
+plot_run(run)
 ```
 
-```{r incrL_multiline1, include=F, eval=F}
+### Dynamic behavior: high light - 3 years
+```{r varyHL3, eval=T, include=T, fig.width=5}
+# Set run time vector
+time <- seq(1, 1095, 0.5)  # if single model run, use time input
+
+# Initialize environment
+env <- init_env(time=time, L=c(5,45,2), N=c(1e-7,1e-7,0), X=c(5e-6,5e-6,0))
+
+# Set parameters
+defpars <- def_pars()  # Get default parameters
+
+# Run simulation
+run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "kCO2"), c(0.2, 10)))#
+plot_run(run)
+```
+
+### Dynamic behavior: high light - 3 years - no feeding
+```{r varyHL3noF, eval=T, include=T, fig.width=5}
+# Set run time vector
+time <- seq(1, 1095, 0.5)  # if single model run, use time input
+
+# Initialize environment
+env <- init_env(time=time, L=c(20,45,2), N=c(1e-7,1e-7,0), X=c(0,0,0))
+
+# Set parameters
+defpars <- def_pars()  # Get default parameters
+
+# Run simulation
+run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "kCO2"), c(0.2, 10)))#
+plot_run(run)
+```
+
+### Dynamic behavior: high light & interacting factors
+```{r incrL_multiline1, include=T, eval=T, fig.width=5}
 time <- seq(1, 200, 0.5)  # Set time
@@
 # Run 1
-env1 <- init_env(time=time, L=c(15,40,2), N=c(1e-7,1e-7,2), X=c(5e-6,5e-6,0))
-run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", 0.12))
+env1 <- init_env(time=time, L=c(15,40,2), N=c(1e-7,1e-7,0), X=c(2e-6,2e-6,0))
+run1 <- run_coral(time=time, env=env1, pars=replace(defpars, c("initS", "kCO2"), c(0.12, 8)))
 # Run 2 - increase food
-env2 <- init_env(time=time, L=c(15,40,2), N=c(1e-7,1e-7,2), X=c(10e-6,10e-6,0))
-run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", 0.12))
+env2 <- init_env(time=time, L=c(15,40,2), N=c(1e-7,1e-7,0), X=c(8e-6,8e-6,0))
+run2 <- run_coral(time=time, env=env2, pars=replace(defpars, c("initS", "kCO2"), c(0.12, 8)))
 # Run 3 - increase DIN
-env3 <- init_env(time=time, L=c(15,40,2), N=c(2e-6,2e-6,2), X=c(10e-6,10e-6,0))
-run3 <- run_coral(time=time, env=env3, pars=replace(defpars, "initS", 0.25))
+env3 <- init_env(time=time, L=c(15,40,2), N=c(1e-6,1e-6,0), X=c(2e-6,2e-6,0))
+run3 <- run_coral(time=time, env=env3, pars=replace(defpars, c("initS", "kCO2"), c(0.16, 8)))
 
 # Plot
-plot(time, env1$L, col=alpha("gold", 0.5), type="l", lwd=2, axes=F)
+par(mfrow=c(1,1), mar=c(4,4,3,4))
+plot(time, env1$L, col=alpha("gold", 0.5), type="l", lwd=2, axes=F, xlab="Time", ylab="S/H biomass ratio")
+axis(side=4); mtext(side=4, text="Light (mol m-2 d-1)", line=2)
 par(new=T)
-plot(NA, xlim=range(time), ylim=c(0,0.3), yaxs="i")
+plot(NA, xlim=range(time), ylim=c(0,0.2), yaxs="i", ylab="", xlab="")
 with(run1, lines(time, S$S/H$H, lty=1))
@@
 with(run3, lines(time, S$S/H$H, lty=3))
+legend("topright", legend=c("default", "+feeding", "+DIN"), lty=c(1,2,3))
 
-```
-
-##### Plot run 1 symbiont PI curve
-```{r run1_PI, cache=T, fig.height=3}
-sym_PI(pars=defpars)
 ```

### 2016-10-15 JTB_supp/JTB_supp.Rmd
Create supplement

---
title: "JTB supplement"
author: "Ross Cunning, Erik B. Muller, Ruth D. Gates, Roger M. Nisbet"
date: "October 14, 2016"
output: html_document
bibliography: library.bib
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(".."))
knitr::opts_chunk$set(message = F, warning = F)
```

```{r initialize, include=FALSE}
# Source default parameters and functions
sapply(c("R/conversions.R", "R/def_pars.R", "R/init_env.R", "R/plot_results.R",
         "R/run_coral.R", "R/run_coral_ss.R", "R/sym_PI.R", "R/run_steady_states.R", 
         "R/plot_steady_states.R", "R/explore_NPQ.R"), source, .GlobalEnv)
# Get default parameters
defpars <- def_pars()
```

This supplement provides justification and derivations from the literature for all parameter values used in the model, as well as additional figures that demonstrate the structure and behavior of the model and its specific components.

## Parameter values

All parameter values used in the model are listed here along with the relevant literature that provides a basis for the value. In most cases, the units of the values reported in the literature needed to be converted to units appropriate for the model, which was accomplished using a set of stoichiometric conversion factors also based on reported literature (see "R/conversions.R" in this paper's data repository). The specific calculations used to convert values from units reported in the literature to relevant units for the model can be found in the R Markdown code used to produce this document.

* $n_{NH}$: N:C molar ratio in host biomass
    + @Horwitz:2015cf measured a C:N ratio of 5.53 in anemones under control conditions, which converts to:
    + Value used: `r defpars$nNH` (N:C ratio, dimensionless)
* $n_{NS}$: N:C molar ratio in symbiont biomass
    + The Redfield ratio (following @Muller:2009io)
    + Value used: `r defpars$nNS` (dimensionless)
* $n_{NX}$: N:C molar ratio in prey biomass
    + Assuming the coral feeds primarily on zooplankton, a general heterotroph has a N:C ratio of 0.2 [@Muller:2009io]
    + Value used: `r defpars$nNX` (N:C ratio, dimensionless)
* $j_{HT}^0$: Specific turnover rate of host biomass
    + @Jacobson:2016bk parameterized a DEB model for coral symbiosis, and reported a specific maintenance rate (analagous to the turnover rate in the current model) of 0.03 CmolH CmolH^-1^ d^-1^.
    + Value used: `r defpars$jHT0` CmolH CmolH^-1^ d^-1^
* $j_{ST}^0$: Specific turnover rate of symbiont biomass
    + Assuming an equivalent maintenance rate as the host:
    + Value used: `r defpars$jST0` CmolS CmolS^-1^ d^-1^
* $\sigma_{NH}, \sigma_{CH}, \sigma_{NS}, \sigma_{CS}$: Proportions of nitrogen and carbon recycled from host and symbiont biomass turnover
    + Assuming highly efficient recycling to avoid loss to the environment:
    + Values used: `r with(defpars, c(sigmaNH, sigmaCH, sigmaNS, sigmaCS))` (dimensionless)
* $j_{Xm}$: Maximum specific feeding rate of host
    + @Wijgerde:2011du measured depletion of carbon from captured prey in *Galaxea fascicularis* (including both intracoelenteric and extracoelenteric feeding), reporting a maximum specific feeding rate of 77.6 µgC cm^-2^ 6h^-1^. This converts to:
    + Value used: `r round(77.6 / µg.g / gC.molC * cm2.m2 / CmolR.m2_Gfas * 4, 2)` molX CmolH^-1^ d^-1^
* $K_X$: Half-saturation constant for prey uptake by host
    + @Purser:2010hp measured feeding rates of the coral *Lophelia pertusa* at different concentrations of *Artemia* nauplii, reporting a half-saturation for feeding of ~345 *Artemia* L^-1^. This converts to:
    + Value used: $`r round(345 * gC.artemia / gC.molC, 7)`$ CmolX L^-1^
* $j_{Nm}$: Maximum specific DIN uptake rate by host
    + @Muscatine:1978fa measured ammonium uptake in 3 coral species and reported an average value of 7.22 µmol NH~4~^+^-N mg chl *a*^-1^ h^-1^ (Table 1, uncorrected), which converts to `r 7.22 * h.d * nmolchlA.cm2 * gchlA.molchlA * mg.g / nmol.mol / CmolR.m2_Spis * cm2.m2 / µmol.mol` mol NH~4~^+^-N  CmolH^-1^ d^-1^.
    + @HoeghGuldberg:1999p3974 measured ammonium and dissolved free amino acid uptake in *Pocillopora damicornis* and reported a maximum rate of 91.8 nmol N cm^-2^ h^-1^, which converts to `r 91.8 * h.d * cm2.m2 / CmolR.m2_Spis / nmol.mol` molN CmolH^-1^ d^-1^.
    + @Godinot:2011jg measured nitrate and ammonium uptake rates in *Stylophora pistillata* under high nutrient concentrations and reported a maximum value of 3.5 µmolN mg chl^-1^ d^-1^ [addition of bars in Fig. 4D,E @ 29°C], which converts to `r 3.5 * h.d * 22 / CmolR.m2_Spis * cm2.m2 / µmol.mol / 1000` molN CmolH^-1^ d^-1^.
    + @Godinot:2011bm measured ammonium uptake kinetics in *Stylophora pistillata* and reported a maximum rate of 5.42 molN mg chl^-1^ h^-1^, which converts to `r 5.42 * h.d * µgchl.cm2_Spis / CmolR.m2_Spis * cm2.m2 / µmol.mol / 1000` molN CmolH^-1^ d^-1^.
    + Averaging these reported values, we arrive at:
    + Value used: `r round(mean(c(7.22 * h.d * nmolchlA.cm2 * gchlA.molchlA * mg.g / nmol.mol / CmolR.m2_Spis * cm2.m2 / µmol.mol, 91.8 * h.d * cm2.m2 / CmolR.m2_Spis / nmol.mol, 3.5 * h.d * 22 / CmolR.m2_Spis * cm2.m2 / µmol.mol / 1000, 5.42 * h.d * µgchl.cm2_Spis / CmolR.m2_Spis * cm2.m2 / µmol.mol / 1000)), 3)` molN CmolH^-1^ d^-1^.
* $K_N$: Half-saturation constant for DIN uptake by host
    + @Muscatine:1978fa measured ammonium uptake in 3 coral species and reported an average half-saturation constant of 1.84 µmolN L^-1^ (Table 1, uncorrected).
    + @Godinot:2011bm measured ammoinum uptake kinetics in *Stylophora pistillata* and reported a half-saturation constant of 0.46 µmol N L^-1^.
    + Averaging these reported values (weighted by number of species measured), we arrive at:
    + Value used: `r round((1.84*3 + 0.46*1)/4, 2)` molN L^-1^
* $k_{CO_2}$: Efficacy of CO~2~ delivery by host CCMs
    + This parameter represents the relative amount of CO~2~ that can be delivered by the host to photosynthesis given the reducing power generated from 1 mol of fixed carbon. We assume that the host uses a portion of its fixed carbon to generate ATP which is then used to activate a variety of carbon concentrating mechanisms (see @Wooldridge:2013tj). We assume these CCMs are efficient and can deliver an order of magnitude more CO~2~ relative to the fixed carbon available.
    + Value used: 10 molCO~2~ molC^-1^
* $j_{HGm}$: Maximum specific growth rate of host 
    + Value used: 1 CmolH CmolH^-1^ d^-1^
    + Reference:
* $y_{CL}$: Quantum yield of photosynthesis
    + The maximum theoretical quantum yield is 0.125 mol C fixed per mol photons. @Brodersen:2014gw presented a balanced light budget based on microenvironmental photobiology of *Montastrea curta*, and reported a quantum yield of 0.102 O~2~ photon^-1^.
    + Value used: 0.1 molC molph^-1^
* $\bar{a}^*$: Effective light-absorbing cross-section of symbiont
    + @Hennige:2009p3439 measured spectra-weighted absoprtion coefficients from 400 to 700nm *ā\**~(400-700)~ for eight *Symbiodinium* types in culture at two different light levels, and reported a (grand mean) value of 0.01095 m^2^ mg chl *a*^-1^ [from Table 6]. This converts to:
    + Value used: `r round(0.01095 * mg.g * gchlA.gC * gC.molC, 2)` m^2^ CmolS^-1^
* $k_{NPQ}$: Capacity of non-photochemical quenching
    + @Gorbunov:2001p4181 report that the capacity of non-photochemical quenching is approximately 4 times the capacity of photochemistry on a sunny day. Therefore, we take the maximum relative capacity of NPQ to be 4 times the maximum rate of photochemical quenching (see below), giving a value of:
    + Value used: `r defpars$jCPm / defpars$yCL * 4` mol ph CmolS^-1^ d^-1^
* $k_{ROS}$: Amount of excitation energy not quenched by photoehcemistry or NPQ that doubles relative production of ROS
    + @McGinty:2012ck report a ~4-fold increase in ROS production in *Symbiodinium* C1 under stressful conditions relative to control. We chose a value of $k_{ROS}$ that similarly produced a ~4-fold increase in ROS production at the highest light level (see Figure S2):
    + Value used: 60 mol ph CmolS^-1^ d^-1^
* $j_{CPm}$: Maximum specific photosynthesis rate of symbiont (molC CmolS^-1^ d^-1^)
    + In theory, this value should be the maximum carboxylation rate of *Symbiodinium* Rubisco (RuBP). @Leggat:2004vx report a maximum carboxylation rate at 29°C of 140 CO~2~ RuBP^-1^ s^-1^ [Max. value, p. 143], which converts to:
    + Value used: `r round(140 * s.h * h.d * RuBP.CmolS, 1)`
* $j_{SGm}$: Maximum specific growth rate of symbiont (CmolS CmolS^-1^ d^-1^)
    + Sugget et al. measured specific growth rates of 5 clades of *Symbiodinium* in culture, and reported values ranging from 0.19 to 0.26 d^-1^ (Table 2). 
    + Value used: 0.25 d^-1^
* $b$: Scaling parameter for strength of bleaching response (dimensionless)
    + This value determines the magnitude of symbiont biomass loss (considered as expulsion of symbionts during bleaching) in response to ROS. In this context, relative ROS concentrations are assumed to trigger symbiont expulsion via unknown signaling mechanisms. Larger values of this parameter cause greater biomass loss in response to ROS signaling (i.e., a more severe bleaching response). Since this represents symbiont cells being expelled from the system, we use a value about an order of magnitude higher than the turnover rate due to maintenance.
    + Value used: 5

## Figure S1. Light amplification and self-shading
```{r lightamp, fig.width=5, fig.height=5, eval=T, include=T, echo=F, fig.cap="The amplification factor (A) is the ratio of internal scalar irradiance experienced by *Symbiodinium* to the external downwelling irradiance, and is greater than 1 due to multiple scattering of incident light by the coral skeleton. The amplification factor decreases as a function of symbiont abundance (converted to symbiont to host biomass ratio) due to self-shading. Data are from Marcelino et al. (2013) using skeletons from three different coral species: pdam=*Pocillopora damicornis*; plob=*Porites lobata*; shys=*Seriatopora hystrix*."}

# Import data from Marcelino et al. 2013
md <- read.csv("data/marcelino2013_fig2a.csv")
md$S.H <- md$density * cm2.m2 / CmolR.m2 * CmolS.cellS  # Convert units of symbiont density to CmolS/CmolR
md$A <- md$E + 1  # Convert "E"=excess light metric of Marcelino to "A"=amplification factor
# Plot data
plot(md$A ~ md$S.H, pch=c(21,23,22)[factor(md$spp)], bg=c("white", "black", "gray")[factor(md$spp)], 
     ylab="A", ylim=c(1,3.2),
     xlab="S:H biomass")
legend("topright", legend=levels(md$spp), pch=c(21,23,22), pt.bg=c("white", "black", "gray"), inset=0.02)
# Fit exponential decay model and plot fitted values
mod <- nls(A ~ x + y * exp(z * S.H), data=md, start=list(x=1, y=2, z=-3))
fit <- predict(mod, list(S.H=seq(0, 0.5, 0.01)))
lines(fit ~ seq(0, 0.5, 0.01))
# Get parameters of fit and add to plot
text(0.13,3.1, labels=paste0(c("A = ", as.character(formula(mod))[3]), collapse=""), adj=0)
coefs <- round(coef(mod), 2)
text(0.13,2.9, labels=paste0(c("x=",coefs[1],"; y=",coefs[2],"; z=",coefs[3]), collapse=""), adj=0)
```

## Figure S2. Photosynthesis-Irradiance curves of modeled symbiont.
```{r run1_PI, cache=T, fig.height=3, fig.width=5, echo=F, fig.cap="The amount of light energy quenched through photochemistry (red line) and non-photochemical quenching (black line), and the relative production of reactive oxygen species (orange line) are plotted as a function of the scalar irradiance experienced by the symbiont (note different y-axes for each line). These simulations represent the behavior of the photosynthesis SU independent of the host and without CO~2~ limitation of photosynthesis (*j~CO2~* set to a very high value). The structure of the photosynthesis SU, with the negative feedback between excess light and photosynthesis rate, produces the classic shape of a P-I curve demonstrating photoinhibition, suggesting the mechanism is robust."}
sym_PI(pars=def_pars())
```

## References

### 2016-10-15 R/def_pars.R
update default parameters, matching supplement

@@ -3,16 +3,16 @@
 def_pars <- function() {
   return(list(
     jHT0=0.03,  # Host specific biomass turnover rate (d^-1)
-    nNH=0.19,  # N:C ratio in host biomass (-)
+    nNH=0.18,  # N:C ratio in host biomass (-)
     nNX=0.2,  # N:C ratio in prey biomass (-)
     sigmaNH=0.9,  # Proportion of host nitrogen turnover recycled (-)
     sigmaCH=0.9,  # Proportion of host carbon turnover recycled (-)
-    jXm=0.1292,  # Maximum specific host feeding rate (molX/CmolH/d)
-    jNm=0.048,  # Maximum specific host DIN uptake rate (molN/CmolH/d)
+    jXm=0.13,  # Maximum specific host feeding rate (molX/CmolH/d)
+    jNm=0.035,  # Maximum specific host DIN uptake rate (molN/CmolH/d)
     jHGm=1,  # Maximum specific host growth rate (CmolH/CmolH/d)
-    kCO2=5,  # Rate of host CCM's (molCO2/molC/d)
-    KN=1e-6,  # Half-saturation constant for host DIN uptake (molN/L)
-    KX=20e-6,  # Half-saturation constant for host feeding (molX/L)
+    kCO2=10,  # Rate of host CCM's (molCO2/molC/d)
+    KN=1.5e-6,  # Half-saturation constant for host DIN uptake (molN/L)
+    KX=1.95e-5,  # Half-saturation constant for host feeding (molX/L)
     initH=1,  # Initial host biomass (CmolH)
     jST0=0.03,  # Symbiont specific biomass turnover rate (d^-1)
     nNS=0.13,  # N:C ratio in symbiont biomass (-)

### 2016-10-15 R/sym_PI.R
plotting tweaks

@@ -31,12 +31,12 @@ sym_PI <- function(pars) {
     # Return ROS and photosynthesis rate for plotting
     par(mfrow=c(1,1), mar=c(3,3,3,7), mgp=c(1.2,0,0), cex=0.7, tck=0.025, xaxs="i")
     plot(jL, jCP/pars$yCL, xlab="Light (mol photons/C-molS/d)", ylab="Photochemical quenching",
-         main="P-I curve", type="l", lwd=3, col="red")
+         type="l", lwd=3, col="red")
     par(new=T)
     plot(jL, jNPQ, type="l", lwd=3, axes=F, xlab="", ylab="")
     axis(side=4); mtext(side=4, text="Non-photochemical quenching", line=1, cex=0.7)
     par(new=T)
-    plot(jL, cROS, type="l", lwd=3, axes=F, col="orange", xlab="", ylab="")
+    plot(jL, cROS, type="l", lwd=3, axes=F, col="orange", xlab="", ylab="", ylim=c(1, max(cROS)*1.5))
     axis(side=4, line=4); mtext(side=4, line=5, text = "ROS production (relative)", cex=0.7)
     legend("bottomright", legend=c("Photo.", "NPQ", "ROS"), lwd=2, col=c("red", "black", "orange"),
            inset=c(0.1, 0.05), bty="n")

### 2016-10-15 Rmd/analyze.Rmd
move amp. & PI curve to supplement

@@
 ![](../img/model_new.png)
-
-### Light amplification & self-shading
-```{r lightamp, fig.width=5, fig.height=5, eval=T, include=T}
-# Import data from Marcelino et al. 2013
-md <- read.csv("data/marcelino2013_fig2a.csv")
-md$S.R <- md$density * cm2.m2 / CmolR.m2 * CmolS.cellS  # Convert units of symbiont density to CmolS/CmolR
-md$A <- md$E + 1  # Convert "E"=excess light metric of Marcelino to "A"=amplification factor
-# Plot data
-plot(md$A ~ md$S.R, pch=c(21,23,22)[factor(md$spp)], bg=c("white", "black", "gray")[factor(md$spp)], 
-     ylab="internal/external light", ylim=c(1,3.2),
-     xlab="Symbiont density (CmolS/CmolR)",
-     main="Light Amplification")
-legend("topright", legend=levels(md$spp), pch=c(21,23,22), pt.bg=c("white", "black", "gray"), inset=0.02)
-# Fit exponential decay model and plot fitted values
-mod <- nls(A ~ x + y * exp(z * S.R), data=md, start=list(x=1, y=2, z=-3))
-fit <- predict(mod, list(S.R=seq(0, 0.5, 0.01)))
-lines(fit ~ seq(0, 0.5, 0.01))
-# Get parameters of fit
-formula(mod); coef(mod)
-```
-
-### PI curve
-```{r run1_PI, cache=T, fig.height=3}
-sym_PI(pars=defpars)
-```
 
@@
 # Run to steady state and plot results
-ssdat <- run_steady_states(pars=replace(defpars, c("initS", "kCO2"), c(1, 10)), at=at, food=5e-6, runtime=T)
+ssdat <- run_steady_states(pars=defpars, at=at, food=5e-6, runtime=T)
 plot_steady_states(ssdat)
@@
 # Initialize environment
-env <- init_env(time=time, L=c(20,20,2), N=c(1e-7,4e-6,2), X=c(0,0,0))
+env <- init_env(time=time, L=c(20,20,2), N=c(1e-8,4e-6,2), X=c(0,0,0))
 
@@
 # Run simulation
-run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "kCO2"), c(0.25, 10)))#
+run <- run_coral(time=time, env=env, pars=defpars)
 plot_run(run)
@@
 # Initialize environment
-env <- init_env(time=time, L=c(5,30,2), N=c(1e-7,1e-7,0), X=c(5e-6,5e-6,0))
+env <- init_env(time=time, L=c(5,30,2), N=c(1e-7,1e-7,0), X=c(0,0,0))
 
@@
 # Run simulation
-run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "kCO2"), c(0.2, 10)))#
+run <- run_coral(time=time, env=env, pars=defpars)#
+plot_run(run)
+```
+
+### Dynamic behavior: food
+```{r varyX, eval=T, include=T, fig.width=5}
+# Set run time vector
+time <- seq(1, 365, 0.5)  # if single model run, use time input
+
+# Initialize environment
+env <- init_env(time=time, L=c(20,20,2), N=c(1e-8,1e-8,0), X=c(0,10e-6,1))
+
+# Set parameters
+defpars <- def_pars()  # Get default parameters
+
+# Run simulation
+run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "kCO2", "nNH"), c(0.2, 10, 0.19)))#
 plot_run(run)

### 2016-10-15 JTB_ms/JTB_ms.Rmd
remove k

@@
 | $K_N$ | Half-saturation constant for DIN uptake by host | 0.46e-6 | molN L^-1^ | 
-| $k_{CO_2}$ | Efficacy of CO~2~ delivery by host CCMs | 5 | molCO~2~ molC^-1^ |
+| $k_{CO_2}$ | Efficacy of CO~2~ delivery by host CCMs | 10 | molCO~2~ molC^-1^ |
 | $j_{HGm}$ | Maximum specific growth rate of host | 1 | CmolH CmolH^-1^ d^-1^ |
@@
 | $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | 1.34 | m^2^ CmolS^-1^ |
-| $k_{NPQ}$ | NPQ capacity of symbiont, relative to C-fixation | 4 | -- |
+| $k_{NPQ}$ | NPQ capacity of symbiont | 112 | mol ph CmolS^-1^ d^-1^ |
 | $k_{ROS}$ | Excess photon energy that doubles ROS production, relative to baseline levels | 40-80 | mol ph CmolS^-1^ d^-1^ |
-| $k$ | Exponent on ROS production rate | 1 | -- |
 | $j_{CPm}$ | Maximum specific photosynthesis rate of symbiont | 2.8 | molC CmolS^-1^ d^-1^ |
@@
 
-\begin{equation} c_{ROS} = 1 + \bigg({{(j_{eL} - j_{NPQ})_+} \over k_{ROS}}\bigg)^k \end{equation}
+\begin{equation} c_{ROS} = 1 + {{(j_{eL} - j_{NPQ})_+} \over k_{ROS}} \end{equation}
 
-where $k_{NPQ}$, $k_{ROS}$, and $k$ are parameters of the symbiont that determine the onset and rate of ROS production.
+where $k_{NPQ}$ and $k_{ROS}$ are parameters of the symbiont that determine the onset and rate of ROS production.
 Importantly, $c_{ROS}$ as specified here is not a function of absolute light absorption, but rather the amount of excess light energy, $j_{eL}$, after accounting for carbon fixation and NPQ.

### 2016-10-18 JTB_ms/JTB_ms.Rmd
text updates

@@
 ---
-title: A simplified dynamic bioenergetic model for coral bleaching
+title: A simplified dynamic bioenergetic model for coral-algal symbiosis and coral bleaching
 author:
@@
 The amount of light absorbed by the symbiont depends on the scalar irradiance at the site of light absorption, which is modified substantially relative to external downwelling irradiance owing to multiple scattering by the coral skeleton and self-shading by surrounding symbionts [@Enriquez:2005p142; @Marcelino:2013hz].
-We used data from @Marcelino:2013hz to empirically derive an "amplification factor" $A$ (the ratio of internal scalar irradiance to external downwelling irradiance) as a function of symbiont density (expressed as symbiont to host biomass ratio), which is specified as:
+We used data from @Marcelino:2013hz to empirically derive an amplification factor, $A$, indicating the ratio of internal scalar irradiance to external downwelling irradiance as a function of symbiont density (S:H biomass), which is specified as:
 
@@
 
-This amplification factor is then multiplied by the external downwelling irradiance $L$ and the effective light-absorbing surface area of symbiont biomass $\bar{a}^*$ to specify the total amount of light absorbed:
+This amplification factor is then multiplied by the external downwelling irradiance $L$ and the effective light-absorbing surface area of symbiont biomass $\bar{a}^*$ to specify the total light absorption:
 
@@
 
-We then specify the input of CO~2~ to photosynthesis as $j_{CO_2}$, which is mediated by the host.
-This encompasses potentially diverse mechanisms by which the host may enhance CO~2~ availability for photosynthesis (i.e., carbon-concentrating mechanisms (CCMs)), including active transport of bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to CO~2~ to promote diffusion toward the symbiont [@Tansik:2015cm], and acidification of the symbiosome to increase localized CO~2~ concentrations around the symbiont [@Barott:2014gx].
-Since these mechanisms require energetic input by the host, the magnitude of CO~2~ delivery is set proportional to $j_{eC}$, the amount of excess carbon from the host biomass SU, assuming that some of this carbon is respired to energize these CCMs.
-This formulation means that the symbiont indirectly ensures its own CO~2~ supply by providing the host with fixed carbon [@Wooldridge:2013tj], which establishes a positive feedback in the model that proves to be essential in simulating the phenomenon of coral bleaching.
-The parameter $k_{CO_2}$ is included to scale the efficacy of host CCMs, i.e., how many moles of CO~2~ can be delivered from the environment to the symbiont per mole of fixed carbon.
-Adjusting this parameter offers the user the opportunity to compare different rates of CO~2~ delivery that may characterize different coral species [@Wooldridge:2014di].
+We then specify the input of CO~2~ to photosynthesis as $j_{CO_2}$, a flux mediated by the host that encompasses potentially diverse carbon-concentrating mechanisms (CCMs), including active transport of bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to CO~2~ to promote diffusion toward the symbiont [@Tansik:2015cm], and acidification of the symbiosome to increase localized CO~2~ concentrations around the symbiont [@Barott:2014gx].
+Since these mechanisms require energetic input by the host, the magnitude of CO~2~ delivery is proportional to the host's excess carbon, $j_{eC}$, assuming that some of this carbon is respired to energize these CCMs.
+This formulation means that the symbiont indirectly ensures its own CO~2~ supply by providing fixed carbon to the host [@Wooldridge:2013tj], which establishes a positive feedback in the model that leads to the existence of multiple stable states, which we discuss below in the context of coral bleaching.
+The parameter $k_{CO_2}$ scales the efficacy of host CCMs, which enables the comparison of different rates of CO~2~ delivery that may characterize different coral species [@Wooldridge:2014di].
 The input of CO~2~ to the photosynthesis SU is therefore specified as:
@@
 where $j_{CPm}$ is the maximum specific rate of photosynthesis, and $c_{ROS}$ is the relative rate of reactive oxygen species production (see below).
-Dividing the photosynthetic rate by $c_{ROS}$ causes a decline in response to photooxidative stress (i.e., photoinhibition).
+Dividing the photosynthetic rate by $c_{ROS}$ causes a decline in response to photooxidative stress at high light levels, and the emergent outcome of this SU formulation demonstrates a classic photoinhibition response (see Fig.
+S1).
 
@@
 
-If light energy further exceeds the total capacity of both photochemistry and NPQ, then reactive oxygen species (ROS) are produced.
-We represent this as a relative quantity $c_{ROS}$, which takes a value of 1 when all light absorbed can be quenched by photochemistry and NPQ, and increases as the amount of excess excitation energy further increases:
+If light energy further exceeds the capacity of both photochemistry and NPQ, then reactive oxygen species (ROS) are produced.
+We represent this as a relative quantity $c_{ROS}$, taking a value of 1 when all light energy is quenched by photochemistry and NPQ, and increasing as the amount of excess excitation energy increases:
 
@@
 
-where $k_{NPQ}$ and $k_{ROS}$ are parameters of the symbiont that determine the onset and rate of ROS production.
-Importantly, $c_{ROS}$ as specified here is not a function of absolute light absorption, but rather the amount of excess light energy, $j_{eL}$, after accounting for carbon fixation and NPQ.
+where $k_{ROS}$ is a parameter of the symbiont that determines the rate of ROS production (specifically, the amount of excess excitation energy that doubles ROS production relative to baseline levels).
+Importantly, $c_{ROS}$ is specified here not as a function of absolute external light, but rather the amount of excess light, $j_{eL}$, after accounting for quenching by carbon fixation and NPQ.
 A direct consequence of this formulation is that CO~2~-limitation of photosynthesis can lead to ROS production, an important mechanism [@Wooldridge:2009p7807] that was not captured by previous representations of photooxidative stress [@Eynaud:2011tv].
-Moreover, this formulation allows functional diversity among symbiont types to be explored by changing the parameters $k_{NPQ}$, $k_{ROS}$, and $k$.
 
@@
 
-This second component of symbiont biomass loss can represent both photodamage and/or symbiont expulsion (i.e., bleaching), both of which occur in response to high levels of ROS production.
+This second component of symbiont biomass loss represents both photodamage and/or symbiont expulsion (i.e., bleaching), both of which occur in response to high levels of ROS production.
 The parameter $b$ is included to scale biomass loss due to bleaching in response to ROS.
-(Note that recycling of symbiont biomass turnover ($rNS$ and $rCS$) only occurs based on the maintenance component of turnover (i.e., $j_{ST}^0$), and not the bleaching-related component, as this loss represents biomass that is damaged or expelled from the host).
+(Note that recycling of symbiont biomass turnover ($rNS$ and $rCS$) only occurs based on the maintenance component of turnover (i.e., $j_{ST}^0$), and not the bleaching-related component, as this loss represents biomass that is damaged or expelled from the host.)
 
@@
 
-A time-stepping Euler method was used to solve the state equations since the production and rejection fluxes of the SUs are imiplicitly defined.
+A time-stepping Euler method was used to solve the state equations since the production and rejection fluxes of the SUs are imiplicitly defined due to feedbacks present in the model.
 Specifically, the rejection fluxes of carbon and nitrogen from the symbiont and host biomass SUs act as reciprocal input fluxes to the other SU.
 In addition, the photosynthesis SU receives CO~2~ proportional to the carbon rejection flux from the host biomass SU, and the rejection flux of excitation energy from the photosynthesis SU acts to reduce its own production (i.e., photoinhibition).
-A vector of time values was created for each simulation run, along which dynamic environmental forcing functions (irradiance, DIN, and prey abundance) can be designed.
+A vector of time steps is used for each simulation, along which dynamic environmental forcing functions (irradiance, DIN, and prey abundance) can be designed.
 These vectors, along with initial values of symbiont and host biomass, then serve as input to the time-stepping function, which solves for the current system state using values of the previous system state where necessary.
@@
 
-where $j_{S1}$ and $j_{S2}$ are the specific input fluxes of the two substrates and $j_{Pm}$ is the maximum specific product formation rate, all with units of Cmol Cmol^-1^ d^-1^.
+where $j_{S1}$ and $j_{S2}$ are the specific input fluxes of the two substrates and $j_{Pm}$ is the maximum specific product formation rate, in units of Cmol Cmol^-1^ d^-1^.
 
@@
 However, if the host is not carbon-limited, it effectively delivers CO~2~ to photosynthesis and ensures its continued receipt of fixed carbon from the symbiont; this system is nitrogen-limited and represents a functional symbiosis with positive exponential growth.
-Either of these stable states may be reached under a constant environment depending on the initial values of symbiont and host biomass: if symbiont biomass is initially very low (i.e., a "bleached" coral), the system cannot escape the carbon-limited state and cannot grow.
+Either of these stable states may be reached under (some sets of) constant environmental conditions, depending on the initial values of symbiont and host biomass: if symbiont biomass is initially very low (i.e., a "bleached" coral), the system cannot escape the carbon-limited state and cannot grow.
 However, if symbiont biomass is initially high (i.e., a healthy coral), then the system maintains positive growth in a nitrogen-limited state.
-For practical purposes, we consider only positive growth states in the "steady state behavior" section below.
-In the "dynamic behavior" section, we explore the conditions that cause the system to switch from one state to another, which in a biological context indicates coral bleaching.
+For practical purposes, we consider only positive growth steady states in the "steady state behavior" section below.
+In the subsequent "dynamic behavior" section, we explore more thoroughly the conditions that cause the system to switch between alternate stable states, and interpret these dynamics in the context of coral bleaching.
 
@@
 
-*emphasize that steady state is exponential growth, and fluxes are per unit biomass/specific rates*
 To analyze qualitative behavior, we ran the model to steady state across gradients of external irradiance and nutrients (Fig.
 2), which revealed patterns consistent with observed phenomena in corals.
-Predicted growth rates are low at low light and DIN, and begin increasing as both of these factors increase (Fig.
+Predicted growth rates are low at low light and DIN (~0.01 d^-1^), and begin increasing as both of these factors increase (Fig.
 2A).
@@
 Increased growth at higher DIN is predicted by the DEB model of @Muller:2009io, and has also been observed experimentally [@MullerParker:1994uo; @Tanaka:2007uv; @Tanaka:2013dj].
-However, DIN elevation beyond a certain point (e.g., 2-3 µM in these simulations) has little effect on growth as carbon becomes limiting (Fig.
+However, DIN elevation beyond a certain point (e.g., ~3-4 µM in these simulations) has little effect on growth as carbon becomes limiting (Fig.
 2B).
@@
 
-The maximum predicted growth rates of ~0.1 d^-1^, occurring between ~10-15 mol photons m^-2^ s^-1^ light and >~3 µM DIN (Fig.
+The maximum predicted growth rates of ~0.1 d^-1^, occurring between ~10-20 mol photons m^-2^ s^-1^ light and ~4 µM DIN (Fig.
 2A), are comparable to the rate of `r Tanaka_gr` d^-1^ measured by @Tanaka:2007uv under similar N-enriched conditions.
-Growth under conditions more typical of reef environments (<0.5 µM DIN) are within the range of ~0.01-0.05 d^-1^; however, these values are still higher than most specific growth rates reported in the literature, which are typically based on skeletal mass, and fall near or below 0.01 d^-1^ [@Osinga:2011kq; @Osinga:2012is] (though values as high as 0.025 d^-1^ have been reported [@Schutter:2010p7758]).
+Growth under conditions more typical of reef environments (<0.5 µM DIN) are within the range of ~0.01-0.03 d^-1^; these values are slightly higher than most specific growth rates reported in the literature, which are typically based on skeletal mass, and fall near or below 0.01 d^-1^ [@Osinga:2011kq; @Osinga:2012is] (though values as high as 0.025 d^-1^ have been reported [@Schutter:2010p7758]).
 These generally low skeletal growth rates relative to predicted biomass growth may be explained by differential substrate-limitation (i.e., CO~2~ may limit calcification, which is not present in the model); indeed, tissue and skeletal growth may not always be well correlated [@Anthony:2002tc].
@@
 
-As irradiance continues to increase above ~15-20 mol photons m^-2^ d^-1^, growth rates decline until positive growth ceases above ~30 µmol photons m^-2^ d^-1^.
-The mechanism underlying this decline is the increase in light energy beyond the capacity of photosynthesis and non-photochemical quenching: excess excitation energy generates reactive oxygen species (ROS) [@Weis:2008p944; @Roth:2014wf], which, in this model, have the phenomenological consequences of reducing the photosynthetic rate (representing photoinhibition) and increasing the symbiont biomass loss rate (representing photodamage and/or symbiont expulsion) [see @Eynaud:2011tv].
+As irradiance continues to increase above ~20 mol photons m^-2^ d^-1^, growth rates decline until positive growth ceases above ~35 µmol photons m^-2^ d^-1^.
+The mechanism underlying this decline is the increase in light energy beyond the capacity of photosynthesis and non-photochemical quenching: excess excitation energy generates reactive oxygen species (ROS) [@Weis:2008p944; @Roth:2014wf], which, in this model, have the phenomenological consequences of reducing the photosynthetic rate (representing photoinhibition) and increasing symbiont biomass loss (representing photodamage and/or symbiont expulsion) [see @Eynaud:2011tv].
 Together, these impacts reduce the symbiont to host biomass ratio (Fig.

### 2016-10-18 JTB_ms/JTB_ms.tex
text updates

@@
 \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-5
+10
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-NPQ capacity of symbiont, relative to C-fixation
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-4
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
---
+NPQ capacity of symbiont
+\strut\end{minipage} &
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+112
+\strut\end{minipage} &
+\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+mol ph CmolS\textsuperscript{-1} d\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 mol ph CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(k\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Exponent on ROS production rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-1
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
---
 \strut\end{minipage}\tabularnewline
@@
 and Iglesias-Prieto 2005; Marcelino et al. 2013). We used data from
-Marcelino et al. (2013) to empirically derive an ``amplification
-factor'' \(A\) (the ratio of internal scalar irradiance to external
-downwelling irradiance) as a function of symbiont density (expressed as
-symbiont to host biomass ratio), which is specified as:
+Marcelino et al. (2013) to empirically derive an amplification factor,
+\(A\), indicating the ratio of internal scalar irradiance to external
+downwelling irradiance as a function of symbiont density (S:H biomass),
+which is specified as:
 
@@
 irradiance \(L\) and the effective light-absorbing surface area of
-symbiont biomass \(\bar{a}^*\) to specify the total amount of light
-absorbed:
+symbiont biomass \(\bar{a}^*\) to specify the total light absorption:
 
@@
 We then specify the input of CO\textsubscript{2} to photosynthesis as
-\(j_{CO_2}\), which is mediated by the host. This encompasses
-potentially diverse mechanisms by which the host may enhance
-CO\textsubscript{2} availability for photosynthesis (i.e.,
-carbon-concentrating mechanisms (CCMs)), including active transport of
-bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to
-CO\textsubscript{2} to promote diffusion toward the symbiont (Tansik,
-Fitt, and Hopkinson 2015), and acidification of the symbiosome to
-increase localized CO\textsubscript{2} concentrations around the
-symbiont (Barott et al. 2014). Since these mechanisms require energetic
-input by the host, the magnitude of CO\textsubscript{2} delivery is set
-proportional to \(j_{eC}\), the amount of excess carbon from the host
-biomass SU, assuming that some of this carbon is respired to energize
-these CCMs. This formulation means that the symbiont indirectly ensures
-its own CO\textsubscript{2} supply by providing the host with fixed
-carbon (Wooldridge 2013), which establishes a positive feedback in the
-model that proves to be essential in simulating the phenomenon of coral
-bleaching. The parameter \(k_{CO_2}\) is included to scale the efficacy
-of host CCMs, i.e., how many moles of CO\textsubscript{2} can be
-delivered from the environment to the symbiont per mole of fixed carbon.
-Adjusting this parameter offers the user the opportunity to compare
+\(j_{CO_2}\), a flux mediated by the host that encompasses potentially
+diverse carbon-concentrating mechanisms (CCMs), including active
+transport of bicarbonate, carbonic anhydrase-catalyzed conversion of
+bicarbonate to CO\textsubscript{2} to promote diffusion toward the
+symbiont (Tansik, Fitt, and Hopkinson 2015), and acidification of the
+symbiosome to increase localized CO\textsubscript{2} concentrations
+around the symbiont (Barott et al. 2014). Since these mechanisms require
+energetic input by the host, the magnitude of CO\textsubscript{2}
+delivery is proportional to the host's excess carbon, \(j_{eC}\),
+assuming that some of this carbon is respired to energize these CCMs.
+This formulation means that the symbiont indirectly ensures its own
+CO\textsubscript{2} supply by providing fixed carbon to the host
+(Wooldridge 2013), which establishes a positive feedback in the model
+that leads to the existence of multiple stable states, which we discuss
+below in the context of coral bleaching. The parameter \(k_{CO_2}\)
+scales the efficacy of host CCMs, which enables the comparison of
 different rates of CO\textsubscript{2} delivery that may characterize
@@
 (see below). Dividing the photosynthetic rate by \(c_{ROS}\) causes a
-decline in response to photooxidative stress (i.e., photoinhibition).
+decline in response to photooxidative stress at high light levels, and
+the emergent outcome of this SU formulation demonstrates a classic
+photoinhibition response (see Fig. S1).
 
@@
 
-If light energy further exceeds the total capacity of both carbon
-fixation and NPQ, then reactive oxygen species (ROS) are produced. We
-represent this as a relative quantity \(c_{ROS}\), which takes a value
-of 1 when all light absorbed can be quenched by photochemistry and NPQ,
-and increases as the amount of excess excitation energy further
-increases:
-
-\begin{equation} c_{ROS} = 1 + \bigg({{(j_{eL} - j_{NPQ})_+} \over k_{ROS}}\bigg)^k \end{equation}
-
-where \(k_{NPQ}\), \(k_{ROS}\), and \(k\) are parameters of the symbiont
-that determine the onset and rate of ROS production. Importantly,
-\(c_{ROS}\) as specified here is not a function of absolute light
-absorption, but rather the amount of excess light energy, \(j_{eL}\),
-after accounting for carbon fixation and NPQ. A direct consequence of
-this formulation is that CO\textsubscript{2}-limitation of
-photosynthesis can lead to ROS production, an important mechanism
+If light energy further exceeds the capacity of both photochemistry and
+NPQ, then reactive oxygen species (ROS) are produced. We represent this
+as a relative quantity \(c_{ROS}\), taking a value of 1 when all light
+energy is quenched by photochemistry and NPQ, and increasing as the
+amount of excess excitation energy increases:
+
+\begin{equation} c_{ROS} = 1 + {{(j_{eL} - j_{NPQ})_+} \over k_{ROS}} \end{equation}
+
+where \(k_{ROS}\) is a parameter of the symbiont that determines the
+rate of ROS production (specifically, the amount of excess excitation
+energy that doubles ROS production relative to baseline levels).
+Importantly, \(c_{ROS}\) is specified here not as a function of absolute
+external light, but rather the amount of excess light, \(j_{eL}\), after
+accounting for quenching by carbon fixation and NPQ. A direct
+consequence of this formulation is that CO\textsubscript{2}-limitation
+of photosynthesis can lead to ROS production, an important mechanism
 (Wooldridge 2009) that was not captured by previous representations of
-photooxidative stress (Eynaud, Nisbet, and Muller 2011). Moreover, this
-formulation allows functional diversity among symbiont types to be
-explored by changing the parameters \(k_{NPQ}\), \(k_{ROS}\), and \(k\).
+photooxidative stress (Eynaud, Nisbet, and Muller 2011).
 
@@
 
-This second component of symbiont biomass loss can represent both
+This second component of symbiont biomass loss represents both
 photodamage and/or symbiont expulsion (i.e., bleaching), both of which
-occur in response to high levels or ROS production. The parameter \(b\)
+occur in response to high levels of ROS production. The parameter \(b\)
 is included to scale biomass loss due to bleaching in response to ROS.
@@
 \(j_{ST}^0\)), and not the bleaching-related component, as this loss
-represents biomass that is damaged or expelled from the host).
+represents biomass that is damaged or expelled from the host.)
 
@@
 A time-stepping Euler method was used to solve the state equations since
-the production and rejection fluxes of the SUs are imiplicitly defined.
-Specifically, the rejection fluxes of carbon and nitrogen from the
-symbiont and host biomass SUs act as reciprocal input fluxes to the
-other SU. In addition, the rejection flux of excitation energy from the
-photosynthesis SU acts to reduce its own production flux (i.e.,
-photoinhibition), and hence a discretized time-stepping procedure was
-necessary. A vector of time values was created for each simulation run,
-along which dynamic environmental forcing functions (irradiance, DIN,
-and prey abundance) can be designed. These vectors, along with initial
-values of symbiont and host biomass, then serve as input to the
-time-stepping function, which solves for the current system state using
-values of the previous system state where necessary. A time step of 0.1
-days was used for all simulations, which were performed using R code
-that is available in the data repository accompanying this article:
-\url{github.com/jrcunning/Rcoral}.
+the production and rejection fluxes of the SUs are imiplicitly defined
+due to feedbacks present in the model. Specifically, the rejection
+fluxes of carbon and nitrogen from the symbiont and host biomass SUs act
+as reciprocal input fluxes to the other SU. In addition, the
+photosynthesis SU receives CO\textsubscript{2} proportional to the
+carbon rejection flux from the host biomass SU, and the rejection flux
+of excitation energy from the photosynthesis SU acts to reduce its own
+production (i.e., photoinhibition). A vector of time steps is used for
+each simulation, along which dynamic environmental forcing functions
+(irradiance, DIN, and prey abundance) can be designed. These vectors,
+along with initial values of symbiont and host biomass, then serve as
+input to the time-stepping function, which solves for the current system
+state using values of the previous system state where necessary. A time
+step of 0.1 days was used for all simulations, which were performed
+using R code that is available in the data repository accompanying this
+article: \url{github.com/jrcunning/Rcoral}.
 
@@
 substrates and \(j_{Pm}\) is the maximum specific product formation
-rate, all with units of Cmol Cmol\textsuperscript{-1}
-d\textsuperscript{-1}.
+rate, in units of Cmol Cmol\textsuperscript{-1} d\textsuperscript{-1}.
 
@@
 
-simulations in constant environs show either positive or negative
-exponential growth. some environments show two steady states, one with
-positive and one with negative growth, depending on initial conditions.
-based on numerical analysis, we see ultimate exponential growth or
-decline. the mechanism that produces these alternate steady states is:
-\ldots{}. . for practical purposes, we are interested in the positive
-growth steady states and their properties and how they change across
-environmental gradients (see section X). but we are also interested in
-the possibility for flips to the alternate steady states (see section
-Y).
+In a constant environment, the system ultimately reaches a steady state
+of exponential growth or decline. Under some conditions, alternate
+steady states can be reached depending on initial values of symbiont and
+host biomass. The mechanism that produces these alternate steady states
+is the positive feedback between carbon-limitation of the host and
+CO\textsubscript{2}-limitation of photosynthesis. If the host becomes
+carbon-limited, the efficacy of its CCMs are reduced and photosynthesis
+becomes progressively more CO\textsubscript{2}-limited. In turn, less
+photosynthate is translocated to the host, causing further
+carbon-limitation and an ultimate decline in biomass; this steady state
+represents a dysfunctional symbiosis. However, if the host is not
+carbon-limited, it effectively delivers CO\textsubscript{2} to
+photosynthesis and ensures its continued receipt of fixed carbon from
+the symbiont; this system is nitrogen-limited and represents a
+functional symbiosis with positive exponential growth. Either of these
+stable states may be reached under (some sets of) constant environmental
+conditions, depending on the initial values of symbiont and host
+biomass: if symbiont biomass is initially very low (i.e., a ``bleached''
+coral), the system cannot escape the carbon-limited state and cannot
+grow. However, if symbiont biomass is initially high (i.e., a healthy
+coral), then the system maintains positive growth in a nitrogen-limited
+state. For practical purposes, we consider only positive growth steady
+states in the ``steady state behavior'' section below. In the subsequent
+``dynamic behavior'' section, we explore more thoroughly the conditions
+that cause the system to switch between alternate stable states, and
+interpret these dynamics in the context of coral bleaching.
 
@@
 
-\emph{emphasize that steady state is exponential growth, and fluxes are
-per unit biomass/specific rates} To analyze qualitative behavior, we ran
-the model to steady state across gradients of external irradiance and
-nutrients (Fig. 2), which revealed patterns consistent with observed
-phenomena in corals. Predicted growth rates are low at low light and
-DIN, and begin increasing as both of these factors increase (Fig. 2A).
-Low light is limiting to photosynthesis (Fig. 2D), which leads to
-carbon-limitation of host growth (Fig. 2B) and an associated increase in
-the symbiont to host biomass ratio (Fig. 2C). In agreement with this
-trend are many observations of negative correlation between irradiance
-and symbiont density (Stimson 1997; Brown et al. 1999; Fitt et al. 2000;
-Titlyanov et al. 2001). As higher light alleviates light-limitation of
+To analyze qualitative behavior, we ran the model to steady state across
+gradients of external irradiance and nutrients (Fig. 2), which revealed
+patterns consistent with observed phenomena in corals. Predicted growth
+rates are low at low light and DIN, and begin increasing as both of
+these factors increase (Fig. 2A). Low light is limiting to
+photosynthesis (Fig. 2D), which leads to carbon-limitation of host
+growth (Fig. 2B) and an associated increase in the symbiont to host
+biomass ratio (Fig. 2C). In agreement with this trend are many
+observations of negative correlation between irradiance and symbiont
+density (Stimson 1997; Brown et al. 1999; Fitt et al. 2000; Titlyanov et
+al. 2001). As higher light alleviates light-limitation of
 photosynthesis, growth becomes less carbon-limited. Similarly, elevating
@@
 
-\subsection{varying light: consequences of excess light and
-light-limitation}\label{varying-light-consequences-of-excess-light-and-light-limitation}
-
-\subsection{varying DIN: role of N-limitation and consequences of
-elevated
-N}\label{varying-din-role-of-n-limitation-and-consequences-of-elevated-n}
-
-\subsection{elevated nutrients cause system to become C-limited (photo:
-CO2 limited; host: fixed-C
-limited)}\label{elevated-nutrients-cause-system-to-become-c-limited-photo-co2-limited-host-fixed-c-limited}
-
-\subsection{C-limitation increases susceptibility to
-bleaching}\label{c-limitation-increases-susceptibility-to-bleaching}
-
-\subsection{consequences of a less efficient but more tolerant
-symbiont}\label{consequences-of-a-less-efficient-but-more-tolerant-symbiont}
-
-\subsection{Rate of symbiont repopulation (Berner et al.
-1993)}\label{rate-of-symbiont-repopulation-berner-et-al.-1993}
-
-\subsection{Bicarbonate addition promotes coral growth (Marubini and
-Thake
-1999)}\label{bicarbonate-addition-promotes-coral-growth-marubini-and-thake-1999}
-
-\subsection{Seasonal change in sym. density is larger when nutrients are
-elevated (Stimson 1997). Such a phenomenon was predicted by Fitt et al.
-2000.}\label{seasonal-change-in-sym.-density-is-larger-when-nutrients-are-elevated-stimson-1997.-such-a-phenomenon-was-predicted-by-fitt-et-al.-2000.}
-
-\subsection{Death rule: negative growth for a certain amount of time
-should be
-defensible.}\label{death-rule-negative-growth-for-a-certain-amount-of-time-should-be-defensible.}
+Therefore, we are highly interested in the conditions under which the
+system switches from one state to another, representing the conditions
+under which corals ``bleach'' due to progressive, reinforcing
+carbon-limitation. Interestingly, this state can be reached in multiple
+ways, by the acceleration of photoinhibition and photodamage, or by high
+external DIN, which causes the symbiont to share less carbon,
+subsequently causing its own CO2-limitation, and progressive
+C-limitation of the system. \#\# varying light: consequences of excess
+light and light-limitation \#\# varying DIN: role of N-limitation and
+consequences of elevated N \#\# elevated nutrients cause system to
+become C-limited (photo: CO2 limited; host: fixed-C limited) \#\#
+C-limitation increases susceptibility to bleaching \#\# consequences of
+a less efficient but more tolerant symbiont \#\# Rate of symbiont
+repopulation (Berner et al. 1993) \#\# Bicarbonate addition promotes
+coral growth (Marubini and Thake 1999) \#\# Seasonal change in sym.
+density is larger when nutrients are elevated (Stimson 1997). Such a
+phenomenon was predicted by Fitt et al. 2000. \#\# Death rule: negative
+growth for a certain amount of time should be defensible.
 

### 2016-10-18 JTB_supp/JTB_supp.Rmd
updates

@@
 ---
-title: "JTB supplement"
+title: "A simplified dynamic bioenergetic model for coral-algal symbiosis and coral bleaching"
 author: "Ross Cunning, Erik B. Muller, Ruth D. Gates, Roger M. Nisbet"
-date: "October 14, 2016"
 output: html_document
@@
 ---
+
+# Supplementary Information
+This supplement provides justification and derivations from the literature for all parameter values used in the model, as well as additional figures that demonstrate the structure and behavior of the model and its specific components.
 
@@
 
-This supplement provides justification and derivations from the literature for all parameter values used in the model, as well as additional figures that demonstrate the structure and behavior of the model and its specific components.
-
-## Parameter values
-
-All parameter values used in the model are listed here along with the relevant literature that provides a basis for the value.
-In most cases, the units of the values reported in the literature needed to be converted to units appropriate for the model, which was accomplished using a set of stoichiometric conversion factors also based on reported literature (see "R/conversions.R" in this paper's data repository).
-The specific calculations used to convert values from units reported in the literature to relevant units for the model can be found in the R Markdown code used to produce this document.
+## Parameter value derivations
+All parameters in the model are listed here along with the relevant literature that provides a basis for the value used.
+In most cases, the units of the values reported in the literature needed to be converted to units appropriate for the model, which was accomplished using a consistent set of stoichiometric conversion factors also based on reported literature (see "R/conversions.R" in this paper's data repository).
+The specific calculations used to convert values from the units reported in the literature to the relevant units for the model can be found in the RMarkdown code used to produce this document.
 
@@
 * $n_{NS}$: N:C molar ratio in symbiont biomass
-    + The Redfield ratio (following @Muller:2009io)
+    + This value should approximate the Redfield ratio (following @Muller:2009io).
     + Value used: `r defpars$nNS` (dimensionless)
 * $n_{NX}$: N:C molar ratio in prey biomass
-    + Assuming the coral feeds primarily on zooplankton, a general heterotroph has a N:C ratio of 0.2 [@Muller:2009io]
+    + Assuming the coral feeds primarily on zooplankton, a general heterotroph has a N:C ratio of 0.2 [@Muller:2009io].
     + Value used: `r defpars$nNX` (N:C ratio, dimensionless)
@@
     + @Gorbunov:2001p4181 report that the capacity of non-photochemical quenching is approximately 4 times the capacity of photochemistry on a sunny day.
-Therefore, we take the maximum relative capacity of NPQ to be 4 times the maximum rate of photochemical quenching (see below), giving a value of:
+Therefore, we take the maximum relative capacity of NPQ to be 4 times the maximum rate of photochemical quenching ($j_{CPm}$; see below), giving a value of:
     + Value used: `r defpars$jCPm / defpars$yCL * 4` mol ph CmolS^-1^ d^-1^
@@
 # Plot data
+par(mfrow=c(1,1), mar=c(3,3,3,3), mgp=c(1.2,0,0), cex=1, tck=0.025)
 plot(md$A ~ md$S.H, pch=c(21,23,22)[factor(md$spp)], bg=c("white", "black", "gray")[factor(md$spp)], 
-     ylab="A", ylim=c(1,3.2),
+     ylab="A (Amplification factor)", ylim=c(1,3.2),
      xlab="S:H biomass")
@@
 ## Figure S2. Photosynthesis-Irradiance curves of modeled symbiont.
-```{r run1_PI, cache=T, fig.height=3, fig.width=5, echo=F, fig.cap="The amount of light energy quenched through photochemistry (red line) and non-photochemical quenching (black line), and the relative production of reactive oxygen species (orange line) are plotted as a function of the scalar irradiance experienced by the symbiont (note different y-axes for each line).
+```{r run1_PI, cache=F, fig.height=4, fig.width=7, echo=F, fig.cap="The amount of light energy quenched through photochemistry (red line) and non-photochemical quenching (black line), and the relative production of reactive oxygen species (orange line) are plotted as a function of the scalar irradiance experienced by the symbiont (note different y-axes for each line).
 These simulations represent the behavior of the photosynthesis SU independent of the host and without CO~2~ limitation of photosynthesis (*j~CO2~* set to a very high value).

### 2016-10-18 R/sym_PI.R
par tweaks

@@ -29,15 +29,15 @@ sym_PI <- function(pars) {
       cROS[t] <- 1 + ((jeL[t] - jNPQ[t]) / pars$kROS)^pars$k
     }
     # Return ROS and photosynthesis rate for plotting
-    par(mfrow=c(1,1), mar=c(3,3,3,7), mgp=c(1.2,0,0), cex=0.7, tck=0.025, xaxs="i")
+    par(mfrow=c(1,1), mar=c(3,3,3,7), mgp=c(1.2,0,0), cex=1, tck=0.025, xaxs="i")
     plot(jL, jCP/pars$yCL, xlab="Light (mol photons/C-molS/d)", ylab="Photochemical quenching",
          type="l", lwd=3, col="red")
     par(new=T)
     plot(jL, jNPQ, type="l", lwd=3, axes=F, xlab="", ylab="")
-    axis(side=4); mtext(side=4, text="Non-photochemical quenching", line=1, cex=0.7)
+    axis(side=4); mtext(side=4, text="Non-photochemical quenching", line=1, cex=1)
     par(new=T)
     plot(jL, cROS, type="l", lwd=3, axes=F, col="orange", xlab="", ylab="", ylim=c(1, max(cROS)*1.5))
-    axis(side=4, line=4); mtext(side=4, line=5, text = "ROS production (relative)", cex=0.7)
+    axis(side=4, line=4); mtext(side=4, line=5, text = "ROS production (relative)", cex=1)
     legend("bottomright", legend=c("Photo.", "NPQ", "ROS"), lwd=2, col=c("red", "black", "orange"),
            inset=c(0.1, 0.05), bty="n")
   })

### 2016-10-18 JTB_ms/JTB_ms.Rmd
update sensitivity analyses

@@
 ![Sensitivity analysis.
-Plots show the fractional change in steady state values in response to fractional changes in default parameter values (see Table 2 for default values).
+Plots show the fractional change in steady state values of growth (solid lines) and S:H biomass (dashed lines) in response to fractional changes in default parameter values (see Table 2 for default values).
 Parameters are grouped by which processes they are involved in.
-Sensitivity in response to changes in each parameter was evaluated in each of four environments (indicated by line type): 1) LLLN=low light (2 mol ph m^-2^ d^-1^) and low DIN (1e-7 mol L^-1^), 2) LLHN=low light (2 mol ph m^-2^ d^-1^) and high DIN (4e-6 mol L^-1^), 3) HLLN=high light (30 mol ph m^-2^ d^-1^) and low DIN (1e-7 mol L^-1^), and 4) HLHN=high light (30 mol ph m^-2^ d^-1^) and high DIN (4e-6 mol L^-1^).](../img/Fig3.png)
+This sensitivity analysis was conducted at conditions typical for coral reef environments: low DIN (1e-7 M) and intermediate light (15 mol m^-2^ d^-2^).
+Sensitivity analyses conducted other environmental conditions are presented in the Supplementary Information.](../img/Fig3.png)
 

### 2016-10-18 JTB_ms/JTB_ms.tex
update sensitivity analyses

@@
             pdfauthor={},
-            pdftitle={A simplified dynamic bioenergetic model for coral bleaching},
+            pdftitle={A simplified dynamic bioenergetic model for coral-algal symbiosis and coral bleaching},
             colorlinks=true,
@@
 
-  \title{A simplified dynamic bioenergetic model for coral bleaching}
+  \title{A simplified dynamic bioenergetic model for coral-algal symbiosis and
+coral bleaching}
     \author[University of Hawaii]{Ross Cunning\corref{c1}}
@@
 patterns consistent with observed phenomena in corals. Predicted growth
-rates are low at low light and DIN, and begin increasing as both of
-these factors increase (Fig. 2A). Low light is limiting to
-photosynthesis (Fig. 2D), which leads to carbon-limitation of host
-growth (Fig. 2B) and an associated increase in the symbiont to host
-biomass ratio (Fig. 2C). In agreement with this trend are many
-observations of negative correlation between irradiance and symbiont
-density (Stimson 1997; Brown et al. 1999; Fitt et al. 2000; Titlyanov et
-al. 2001). As higher light alleviates light-limitation of
-photosynthesis, growth becomes less carbon-limited. Similarly, elevating
-DIN up to \textasciitilde{}2 µM alleviates nitrogen-limitation (Fig.
-2B). Increased growth at higher DIN is predicted by the DEB model of
-Muller et al. (2009), and has also been observed experimentally
-(Muller-Parker et al. 1994; Tanaka et al. 2007; Tanaka et al. 2013).
-However, DIN elevation beyond a certain point (e.g., 2-3 µM in these
-simulations) has little effect on growth as carbon becomes limiting
-(Fig. 2B). Very high nutrient levels may even reduce growth (Shantz,
-Lemoine, and Burkepile 2015), although these impacts are not likely to
-occur within the range of concentrations considered here (\textless{}4
-µM) (Ferrier-Pagès et al. 2000).
+rates are low at low light and DIN (\textasciitilde{}0.01
+d\textsuperscript{-1}), and begin increasing as both of these factors
+increase (Fig. 2A). Low light is limiting to photosynthesis (Fig. 2D),
+which leads to carbon-limitation of host growth (Fig. 2B) and an
+associated increase in the symbiont to host biomass ratio (Fig. 2C). In
+agreement with this trend are many observations of negative correlation
+between irradiance and symbiont density (Stimson 1997; Brown et al.
+1999; Fitt et al. 2000; Titlyanov et al. 2001). As higher light
+alleviates light-limitation of photosynthesis, growth becomes less
+carbon-limited. Similarly, elevating DIN up to \textasciitilde{}2 µM
+alleviates nitrogen-limitation (Fig. 2B). Increased growth at higher DIN
+is predicted by the DEB model of Muller et al. (2009), and has also been
+observed experimentally (Muller-Parker et al. 1994; Tanaka et al. 2007;
+Tanaka et al. 2013). However, DIN elevation beyond a certain point
+(e.g., \textasciitilde{}3-4 µM in these simulations) has little effect
+on growth as carbon becomes limiting (Fig. 2B). Very high nutrient
+levels may even reduce growth (Shantz, Lemoine, and Burkepile 2015),
+although these impacts are not likely to occur within the range of
+concentrations considered here (\textless{}4 µM) (Ferrier-Pagès et al.
+2000).
 
@@
 The maximum predicted growth rates of \textasciitilde{}0.1
-d\textsuperscript{-1}, occurring between \textasciitilde{}10-15 mol
+d\textsuperscript{-1}, occurring between \textasciitilde{}10-20 mol
 photons m\textsuperscript{-2} s\textsuperscript{-1} light and
-\textgreater{}\textasciitilde{}3 µM DIN (Fig. 2A), are comparable to the
-rate of 0.07 d\textsuperscript{-1} measured by Tanaka et al. (2007)
-under similar N-enriched conditions. Growth under conditions more
-typical of reef environments (\textless{}0.5 µM DIN) are within the
-range of \textasciitilde{}0.01-0.05 d\textsuperscript{-1}; however,
-these values are still higher than most specific growth rates reported
-in the literature, which are typically based on skeletal mass, and fall
-near or below 0.01 d\textsuperscript{-1} (Osinga et al. 2011; Osinga et
-al. 2012) (though values as high as 0.025 d\textsuperscript{-1} have
-been reported (Schutter et al. 2010)). These generally low skeletal
-growth rates relative to predicted biomass growth may be explained by
+\textasciitilde{}4 µM DIN (Fig. 2A), are comparable to the rate of 0.07
+d\textsuperscript{-1} measured by Tanaka et al. (2007) under similar
+N-enriched conditions. Growth under conditions more typical of reef
+environments (\textless{}0.5 µM DIN) are within the range of
+\textasciitilde{}0.01-0.03 d\textsuperscript{-1}; these values are
+slightly higher than most specific growth rates reported in the
+literature, which are typically based on skeletal mass, and fall near or
+below 0.01 d\textsuperscript{-1} (Osinga et al. 2011; Osinga et al.
+2012) (though values as high as 0.025 d\textsuperscript{-1} have been
+reported (Schutter et al. 2010)). These generally low skeletal growth
+rates relative to predicted biomass growth may be explained by
 differential substrate-limitation (i.e., CO\textsubscript{2} may limit
@@
 
-As irradiance continues to increase above \textasciitilde{}15-20 mol
+As irradiance continues to increase above \textasciitilde{}20 mol
 photons m\textsuperscript{-2} d\textsuperscript{-1}, growth rates
-decline until positive growth ceases above \textasciitilde{}30 µmol
+decline until positive growth ceases above \textasciitilde{}35 µmol
 photons m\textsuperscript{-2} d\textsuperscript{-1}. The mechanism
@@
 of reducing the photosynthetic rate (representing photoinhibition) and
-increasing the symbiont biomass loss rate (representing photodamage
-and/or symbiont expulsion) (see Eynaud, Nisbet, and Muller 2011).
-Together, these impacts reduce the symbiont to host biomass ratio (Fig.
-2C), as occurs during coral bleaching. This reduction in symbionts
-consequently reduces the flux of fixed carbon to the host, resulting in
-increasing carbon-limitation (Fig. 2B) and eventual cessation of growth
-(Fig. 2A). Importantly, these consequences arise in the present model
-only when light energy exceeds the capacity of photochemistry and NPQ,
-which in turn depends on CO\textsubscript{2} availability. While
-previous models framed photophysiological stress as a fixed response to
-absolute irradiance levels (Eynaud, Nisbet, and Muller 2011), the
-current implementation considers the dynamic balance of multiple energy
-sinks in the causation of stress, which is more consistent with current
+increasing symbiont biomass loss (representing photodamage and/or
+symbiont expulsion) (see Eynaud, Nisbet, and Muller 2011). Together,
+these impacts reduce the symbiont to host biomass ratio (Fig. 2C), as
+occurs during coral bleaching. This reduction in symbionts consequently
+reduces the flux of fixed carbon to the host, resulting in increasing
+carbon-limitation (Fig. 2B) and eventual cessation of growth (Fig. 2A).
+Importantly, these consequences arise in the present model only when
+light energy exceeds the capacity of photochemistry and NPQ, which in
+turn depends on CO\textsubscript{2} availability. While previous models
+framed photophysiological stress as a fixed response to absolute
+irradiance levels (Eynaud, Nisbet, and Muller 2011), the current
+implementation considers the dynamic balance of multiple energy sinks in
+the causation of stress, which is more consistent with current
 understanding of symbiosis dysfunction (Wooldridge 2013), and
@@
 \caption{Sensitivity analysis. Plots show the fractional change in
-steady state values in response to fractional changes in default
-parameter values (see Table 2 for default values). Parameters are
-grouped by which processes they are involved in. Sensitivity in response
-to changes in each parameter was evaluated in each of four environments
-(indicated by line type): 1) LLLN=low light (2 mol ph
-m\textsuperscript{-2} d\textsuperscript{-1}) and low DIN (1e-7 mol
-L\textsuperscript{-1}), 2) LLHN=low light (2 mol ph
-m\textsuperscript{-2} d\textsuperscript{-1}) and high DIN (4e-6 mol
-L\textsuperscript{-1}), 3) HLLN=high light (30 mol ph
-m\textsuperscript{-2} d\textsuperscript{-1}) and low DIN (1e-7 mol
-L\textsuperscript{-1}), and 4) HLHN=high light (30 mol ph
-m\textsuperscript{-2} d\textsuperscript{-1}) and high DIN (4e-6 mol
-L\textsuperscript{-1}).}
+steady state values of growth (solid lines) and S:H biomass (dashed
+lines) in response to fractional changes in default parameter values
+(see Table 2 for default values). Parameters are grouped by which
+processes they are involved in. This sensitivity analysis was conducted
+at conditions typical for coral reef environments: low DIN (1e-7 M) and
+intermediate light (15 mol m\textsuperscript{-2} d\textsuperscript{-2}).
+Sensitivity analyses conducted other environmental conditions are
+presented in the Supplementary Information.}
 \end{figure}

### 2016-10-18 JTB_supp/JTB_supp.Rmd
update sensitivity analyses

@@
          "R/run_coral.R", "R/run_coral_ss.R", "R/sym_PI.R", "R/run_steady_states.R", 
-         "R/plot_steady_states.R", "R/explore_NPQ.R"), source, .GlobalEnv)
+         "R/plot_steady_states.R", "R/sensitivity.R"), source, .GlobalEnv)
 # Get default parameters
@@
 
+## Figure S3. Sensitivity analysis under high DIN (4e-6 M) and intermediate light (15 mol m^-2^ d^-1^).
+```{r sens_ILHN, cache=T, echo=F}
+# Set up time vector and parameters
+defpars <- def_pars()  # Get default parameters
+defpars <- replace(defpars, "initS", 1)  # Start with high S biomass to avoid alternate steady state of negative growth
+
+# Define environments for conducting sensitivity analysis
+envs <- list(
+  ILHN=list(L=15, N=4e-6, X=0)
+)
+
+# Define relative changes in each parameter to evaluate
+levs <- seq(0.5, 1.5, 0.05)  # Set levels of % change in parameter value to evaluate
+
+# Define all simulations to be run for a given parameter
+sims <- expand.grid(env=names(envs), change=levs)
+n <- nrow(sims)
+
+# Set up cluster for parallel processing
+cl <- makeCluster(detectCores())  # Initiate cluster
+registerDoParallel(cl)
+
+# Run sensitivity analysis in each environment for each parameter
+jHT0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHT0", change=change))
+jNm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jNm", change=change))
+jHGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHGm", change=change))
+kCO2.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kCO2", change=change))
+KN.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="KN", change=change))
+jST0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jST0", change=change))
+kNPQ.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kNPQ", change=change))
+kROS.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kROS", change=change))
+astar.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="astar", change=change))
+jCPm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jCPm", change=change))
+jSGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jSGm", change=change))
+b.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="b", change=change))
+
+stopCluster(cl)  # Stop cluster
+
+# Combine output with input
+sens.all <- ls(pattern="*.sens")
+for (i in 1:length(sens.all)) {
+  assign(sens.all[i], cbind(sims, get(sens.all[i])))
+}
+
+# When growth rate becomes negative, set all responses to NA
+for (i in 1:length(sens.all)) {
+  if (any(get(sens.all[i])$grchange < 0, na.rm=T)) {
+    assign(sens.all[i], within(get(sens.all[i]), {
+      shchange <- replace(shchange, grchange < 0, NA)
+      cROSchange <- replace(cROSchange, grchange < 0, NA)
+      grchange <- replace(grchange, grchange < 0, NA)
+    }))
+  }
+}
+
+# Plot Figure
+par(mfrow=c(2,2), mar=c(3,3,1,1), mgp=c(2,0.5,0))
+sens.plot(response=c("grchange", "shchange"), pars=c("jCPm", "kCO2", "astar"), cols=c("black", "red", "gold"))
+sens.plot(response=c("grchange", "shchange"), pars=c("jNm", "KN"), cols=c("black", "blue"))
+sens.plot(response=c("grchange", "shchange"), pars=c("jST0", "jHT0", "jSGm", "jHGm"), cols=c("burlywood4", "burlywood3", "darkolivegreen4", "darkolivegreen3"))
+sens.plot(response=c("grchange", "shchange"), pars=c("kNPQ", "kROS", "b"), cols=c("black", "orange", "purple"))
+```
+
+## Figure S4. Sensitivity analysis under low DIN (1e-7 M) and low light (2 mol m^-2^ d^-1^).
+```{r sens_LLLN, cache=T, echo=F}
+# Set up time vector and parameters
+defpars <- def_pars()  # Get default parameters
+defpars <- replace(defpars, "initS", 1)  # Start with high S biomass to avoid alternate steady state of negative growth
+
+# Define environments for conducting sensitivity analysis
+envs <- list(
+  LLLN=list(L=2, N=1e-7, X=0)
+)
+
+# Define relative changes in each parameter to evaluate
+levs <- seq(0.5, 1.5, 0.05)  # Set levels of % change in parameter value to evaluate
+
+# Define all simulations to be run for a given parameter
+sims <- expand.grid(env=names(envs), change=levs)
+n <- nrow(sims)
+
+# Set up cluster for parallel processing
+cl <- makeCluster(detectCores())  # Initiate cluster
+registerDoParallel(cl)
+
+# Run sensitivity analysis in each environment for each parameter
+jHT0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHT0", change=change))
+jNm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jNm", change=change))
+jHGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHGm", change=change))
+kCO2.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kCO2", change=change))
+KN.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="KN", change=change))
+jST0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jST0", change=change))
+kNPQ.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kNPQ", change=change))
+kROS.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kROS", change=change))
+astar.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="astar", change=change))
+jCPm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jCPm", change=change))
+jSGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jSGm", change=change))
+b.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="b", change=change))
+
+stopCluster(cl)  # Stop cluster
+
+# Combine output with input
+sens.all <- ls(pattern="*.sens")
+for (i in 1:length(sens.all)) {
+  assign(sens.all[i], cbind(sims, get(sens.all[i])))
+}
+
+# When growth rate becomes negative, set all responses to NA
+for (i in 1:length(sens.all)) {
+  if (any(get(sens.all[i])$grchange < 0, na.rm=T)) {
+    assign(sens.all[i], within(get(sens.all[i]), {
+      shchange <- replace(shchange, grchange < 0, NA)
+      cROSchange <- replace(cROSchange, grchange < 0, NA)
+      grchange <- replace(grchange, grchange < 0, NA)
+    }))
+  }
+}
+
+# Plot Figure
+par(mfrow=c(2,2), mar=c(3,3,1,1), mgp=c(2,0.5,0))
+sens.plot(response=c("grchange", "shchange"), pars=c("jCPm", "kCO2", "astar"), cols=c("black", "red", "gold"))
+sens.plot(response=c("grchange", "shchange"), pars=c("jNm", "KN"), cols=c("black", "blue"))
+sens.plot(response=c("grchange", "shchange"), pars=c("jST0", "jHT0", "jSGm", "jHGm"), cols=c("burlywood4", "burlywood3", "darkolivegreen4", "darkolivegreen3"))
+sens.plot(response=c("grchange", "shchange"), pars=c("kNPQ", "kROS", "b"), cols=c("black", "orange", "purple"))
+```
+
+## Figure S5. Sensitivity analysis under high DIN (4e-6 M) and low light (2 mol m^-2^ d^-1^).
+```{r sens_LLHN, cache=T, echo=F}
+# Set up time vector and parameters
+defpars <- def_pars()  # Get default parameters
+defpars <- replace(defpars, "initS", 1)  # Start with high S biomass to avoid alternate steady state of negative growth
+
+# Define environments for conducting sensitivity analysis
+envs <- list(
+  LLHN=list(L=2, N=4e-6, X=0)
+)
+
+# Define relative changes in each parameter to evaluate
+levs <- seq(0.5, 1.5, 0.05)  # Set levels of % change in parameter value to evaluate
+
+# Define all simulations to be run for a given parameter
+sims <- expand.grid(env=names(envs), change=levs)
+n <- nrow(sims)
+
+# Set up cluster for parallel processing
+cl <- makeCluster(detectCores())  # Initiate cluster
+registerDoParallel(cl)
+
+# Run sensitivity analysis in each environment for each parameter
+jHT0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHT0", change=change))
+jNm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jNm", change=change))
+jHGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHGm", change=change))
+kCO2.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kCO2", change=change))
+KN.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="KN", change=change))
+jST0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jST0", change=change))
+kNPQ.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kNPQ", change=change))
+kROS.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kROS", change=change))
+astar.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="astar", change=change))
+jCPm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jCPm", change=change))
+jSGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jSGm", change=change))
+b.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="b", change=change))
+
+stopCluster(cl)  # Stop cluster
+
+# Combine output with input
+sens.all <- ls(pattern="*.sens")
+for (i in 1:length(sens.all)) {
+  assign(sens.all[i], cbind(sims, get(sens.all[i])))
+}
+
+# When growth rate becomes negative, set all responses to NA
+for (i in 1:length(sens.all)) {
+  if (any(get(sens.all[i])$grchange < 0, na.rm=T)) {
+    assign(sens.all[i], within(get(sens.all[i]), {
+      shchange <- replace(shchange, grchange < 0, NA)
+      cROSchange <- replace(cROSchange, grchange < 0, NA)
+      grchange <- replace(grchange, grchange < 0, NA)
+    }))
+  }
+}
+
+# Plot Figure
+par(mfrow=c(2,2), mar=c(3,3,1,1), mgp=c(2,0.5,0))
+sens.plot(response=c("grchange", "shchange"), pars=c("jCPm", "kCO2", "astar"), cols=c("black", "red", "gold"))
+sens.plot(response=c("grchange", "shchange"), pars=c("jNm", "KN"), cols=c("black", "blue"))
+sens.plot(response=c("grchange", "shchange"), pars=c("jST0", "jHT0", "jSGm", "jHGm"), cols=c("burlywood4", "burlywood3", "darkolivegreen4", "darkolivegreen3"))
+sens.plot(response=c("grchange", "shchange"), pars=c("kNPQ", "kROS", "b"), cols=c("black", "orange", "purple"))
+```
+
+## Figure S6. Sensitivity analysis under low DIN (1e-7 M) and high light (25 mol m^-2^ d^-1^).
+```{r sens_HLLN, cache=T, echo=F}
+# Set up time vector and parameters
+defpars <- def_pars()  # Get default parameters
+defpars <- replace(defpars, "initS", 1)  # Start with high S biomass to avoid alternate steady state of negative growth
+
+# Define environments for conducting sensitivity analysis
+envs <- list(
+  HLLN=list(L=25, N=1e-7, X=0)
+)
+
+# Define relative changes in each parameter to evaluate
+levs <- seq(0.5, 1.5, 0.05)  # Set levels of % change in parameter value to evaluate
+
+# Define all simulations to be run for a given parameter
+sims <- expand.grid(env=names(envs), change=levs)
+n <- nrow(sims)
+
+# Set up cluster for parallel processing
+cl <- makeCluster(detectCores())  # Initiate cluster
+registerDoParallel(cl)
+
+# Run sensitivity analysis in each environment for each parameter
+jHT0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHT0", change=change))
+jNm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jNm", change=change))
+jHGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHGm", change=change))
+kCO2.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kCO2", change=change))
+KN.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="KN", change=change))
+jST0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jST0", change=change))
+kNPQ.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kNPQ", change=change))
+kROS.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kROS", change=change))
+astar.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="astar", change=change))
+jCPm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jCPm", change=change))
+jSGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jSGm", change=change))
+b.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="b", change=change))
+
+stopCluster(cl)  # Stop cluster
+
+# Combine output with input
+sens.all <- ls(pattern="*.sens")
+for (i in 1:length(sens.all)) {
+  assign(sens.all[i], cbind(sims, get(sens.all[i])))
+}
+
+# When growth rate becomes negative, set all responses to NA
+for (i in 1:length(sens.all)) {
+  if (any(get(sens.all[i])$grchange < 0, na.rm=T)) {
+    assign(sens.all[i], within(get(sens.all[i]), {
+      shchange <- replace(shchange, grchange < 0, NA)
+      cROSchange <- replace(cROSchange, grchange < 0, NA)
+      grchange <- replace(grchange, grchange < 0, NA)
+    }))
+  }
+}
+
+# Plot Figure
+par(mfrow=c(2,2), mar=c(3,3,1,1), mgp=c(2,0.5,0))
+sens.plot(response=c("grchange", "shchange"), pars=c("jCPm", "kCO2", "astar"), cols=c("black", "red", "gold"))
+sens.plot(response=c("grchange", "shchange"), pars=c("jNm", "KN"), cols=c("black", "blue"))
+sens.plot(response=c("grchange", "shchange"), pars=c("jST0", "jHT0", "jSGm", "jHGm"), cols=c("burlywood4", "burlywood3", "darkolivegreen4", "darkolivegreen3"))
+sens.plot(response=c("grchange", "shchange"), pars=c("kNPQ", "kROS", "b"), cols=c("black", "orange", "purple"))
+```
+
+## Figure S7. Sensitivity analysis under high DIN (4e-6 M) and high light (25 mol m^-2^ d^-1^).
+```{r sens_HLHN, cache=T, echo=F}
+# Set up time vector and parameters
+defpars <- def_pars()  # Get default parameters
+defpars <- replace(defpars, "initS", 1)  # Start with high S biomass to avoid alternate steady state of negative growth
+
+# Define environments for conducting sensitivity analysis
+envs <- list(
+  HLHN=list(L=25, N=4e-6, X=0)
+)
+
+# Define relative changes in each parameter to evaluate
+levs <- seq(0.5, 1.5, 0.05)  # Set levels of % change in parameter value to evaluate
+
+# Define all simulations to be run for a given parameter
+sims <- expand.grid(env=names(envs), change=levs)
+n <- nrow(sims)
+
+# Set up cluster for parallel processing
+cl <- makeCluster(detectCores())  # Initiate cluster
+registerDoParallel(cl)
+
+# Run sensitivity analysis in each environment for each parameter
+jHT0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHT0", change=change))
+jNm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jNm", change=change))
+jHGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHGm", change=change))
+kCO2.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kCO2", change=change))
+KN.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="KN", change=change))
+jST0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jST0", change=change))
+kNPQ.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kNPQ", change=change))
+kROS.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kROS", change=change))
+astar.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="astar", change=change))
+jCPm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jCPm", change=change))
+jSGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jSGm", change=change))
+b.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="b", change=change))
+
+stopCluster(cl)  # Stop cluster
+
+# Combine output with input
+sens.all <- ls(pattern="*.sens")
+for (i in 1:length(sens.all)) {
+  assign(sens.all[i], cbind(sims, get(sens.all[i])))
+}
+
+# When growth rate becomes negative, set all responses to NA
+for (i in 1:length(sens.all)) {
+  if (any(get(sens.all[i])$grchange < 0, na.rm=T)) {
+    assign(sens.all[i], within(get(sens.all[i]), {
+      shchange <- replace(shchange, grchange < 0, NA)
+      cROSchange <- replace(cROSchange, grchange < 0, NA)
+      grchange <- replace(grchange, grchange < 0, NA)
+    }))
+  }
+}
+
+# Plot Figure
+par(mfrow=c(2,2), mar=c(3,3,1,1), mgp=c(2,0.5,0))
+sens.plot(response=c("grchange", "shchange"), pars=c("jCPm", "kCO2", "astar"), cols=c("black", "red", "gold"))
+sens.plot(response=c("grchange", "shchange"), pars=c("jNm", "KN"), cols=c("black", "blue"))
+sens.plot(response=c("grchange", "shchange"), pars=c("jST0", "jHT0", "jSGm", "jHGm"), cols=c("burlywood4", "burlywood3", "darkolivegreen4", "darkolivegreen3"))
+sens.plot(response=c("grchange", "shchange"), pars=c("kNPQ", "kROS", "b"), cols=c("black", "orange", "purple"))
+```
+
 ## References

### 2016-10-18 R/Fig3.R

@@ -1,0 +1,70 @@
+# Load functions
+sapply(c("R/def_pars.R",
+         "R/run_coral_ss.R", 
+         "R/sensitivity.R"), 
+       source, .GlobalEnv)
+
+# Set up time vector and parameters
+defpars <- def_pars()  # Get default parameters
+defpars <- replace(defpars, "initS", 1)  # Start with high S biomass to avoid alternate steady state of negative growth
+
+# Define environments for conducting sensitivity analysis
+envs <- list(
+  #LLLN=list(L=2, N=1e-7, X=0),
+  #LLHN=list(L=2, N=4e-6, X=0),
+  HLLN=list(L=15, N=1e-7, X=0)#,
+  #HLHN=list(L=20, N=4e-6, X=0)
+)
+
+# Define relative changes in each parameter to evaluate
+levs <- seq(0.5, 1.5, 0.05)  # Set levels of % change in parameter value to evaluate
+
+# Define all simulations to be run for a given parameter
+sims <- expand.grid(env=names(envs), change=levs)
+n <- nrow(sims)
+
+# Set up cluster for parallel processing
+cl <- makeCluster(detectCores())  # Initiate cluster
+registerDoParallel(cl)
+
+# Run sensitivity analysis in each environment for each parameter
+jHT0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHT0", change=change))
+jNm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jNm", change=change))
+jHGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHGm", change=change))
+kCO2.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kCO2", change=change))
+KN.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="KN", change=change))
+jST0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jST0", change=change))
[…]

### 2016-10-18 R/sensitivity.R

@@ -8,15 +8,15 @@ source("R/init_env.R")
 # Sensitivity analysis
 
 # Create function to measure sensitivity of responses to specified change in single parameter
-sens <- function(time, env, pars, par, change) {
+sens <- function(env, pars, par, change) {
   # Run 1 
-  run1 <- run_coral(time=time, env=env, pars=pars)
-  run1ss <- lapply(run1[c("H", "S")], function(x) x[length(time), ])
+  run1 <- run_coral_ss(env=env, pars=pars)
+  run1ss <- lapply(run1[c("H", "S")], function(x) x[nrow(run1$H), ])
   run1gr <- run1ss$H$dH.Hdt
   run1sh <- run1ss$S$S/run1ss$H$H
   # Run 2 - with changed parameter
-  run2 <- run_coral(time=time, env=env, pars=replace(pars, par, with(pars, get(par))*change))
-  run2ss <- lapply(run2[c("H", "S")], function(x) x[length(time), ])
+  run2 <- run_coral_ss(env=env, pars=replace(pars, par, with(pars, get(par))*change))
+  run2ss <- lapply(run2[c("H", "S")], function(x) x[nrow(run2$H), ])
   run2gr <- run2ss$H$dH.Hdt
   run2sh <- run2ss$S$S/run2ss$H$H
   # Calculate relative change in steady state response
@@ -27,363 +27,38 @@ sens <- function(time, env, pars, par, change) {
   return(list(grchange=grchange, shchange=shchange, cROSchange=cROSchange))
 }
 
-# Define function to measure sensitivity across a range of changes in parameter value
-vsens <- function(par, levs, env) {
-  gr <- rep(NA, length(levs))
-  sh <- rep(NA, length(levs))
-  cROS <- rep(NA, length(levs))
-  for (i in 1:length(levs)) {
-    sa <- sens(time=time, env=env, pars=defpars, par=par, change=levs[i])
-    gr[i] <- sa[[1]]
-    sh[i] <- sa[[2]]
-    cROS[i] <- sa[[3]]
+
+# PLOT FUNCTION
[…]

### 2016-10-19 R/plot_steady_states.R
tweak Fig2 formatting

@@ -19,7 +19,7 @@ plot_steady_states <- function(ss, png=NULL) {
   
   # Build plotting function
   imagef <- function(r, bin, main, col) {
-    par(tcl=-0.2, cex.main=1, cex.axis=0.75, mar=c(2.3,2.3,2,3.5), mgp=c(0,0.2,0))
+    par(tcl=-0.2, cex.main=1, cex.axis=0.75, mar=c(2,2,3,3), mgp=c(0,0.1,0))
     if (any(r<0, na.rm=T) & any(r>0, na.rm=T)) {
       nb <- rev(seq(0, -1*min(r[which(r!=0)], na.rm=T)+bin, bin))*-1  # negative portion of breaks
       pb <- seq(0, max(r[which(r!=0)], na.rm=T)+bin, bin)
@@ -50,15 +50,15 @@ plot_steady_states <- function(ss, png=NULL) {
     image(t(r), breaks=breaks, col=colors, xaxt="n", yaxt="n")
     #title(main=main, line=0.5, adj=-0.5, outer=F, xpd=T)
     conv <- diff(grconvertX(0:1, 'inches', 'user')) # length of one inch in user coordinates
-    text(x=par("usr")[1]-par("mai")[2]*conv, y=par("usr")[4]+par("mai")[3]/2.5*conv, adj=c(0,0), 
-         labels=main, xpd=NA, font=2)
-    mtext(side=1, text="[DIN] (mol/L)", line=1.2, cex=0.75)
+    text(x=par("usr")[1]-par("mai")[2]*conv/1.1, y=par("usr")[4]+par("mai")[3]/2.5*conv, adj=c(0,0), 
+         labels=main, xpd=NA, font=2, cex=1)
+    mtext(side=1, text="[DIN] (mol/L)", line=1, cex=0.75)
     axis(side=1, at=seq(0,1,0.125), 
          labels=quantile(seq(as.numeric(first(colnames(r))),
                              as.numeric(last(colnames(r))),
                              length.out=100),
                          probs=seq(0,1,0.125)))
-    mtext(side=2, text="Light (mol ph/m2/d)", line=1.2, cex=0.75)
+    mtext(side=2, text="Light (mol ph/m2/d)", line=1, cex=0.75)
     axis(side=2, at=seq(0,1,0.1), 
          labels=quantile(seq(as.numeric(first(rownames(r))),
                              as.numeric(last(rownames(r))),
[…]

### 2016-10-19 JTB_ms/JTB_ms.Rmd
text updates

@@
 
-To analyze qualitative behavior, we ran the model to steady state across gradients of external irradiance and nutrients (Fig.
-2), which revealed patterns consistent with observed phenomena in corals.
-Predicted growth rates are low at low light and DIN (~0.01 d^-1^), and begin increasing as both of these factors increase (Fig.
-2A).
-Low light is limiting to photosynthesis (Fig.
-2D), which leads to carbon-limitation of host growth (Fig.
-2B) and an associated increase in the symbiont to host biomass ratio (Fig.
-2C).
-In agreement with this trend are many observations of negative correlation between irradiance and symbiont density [@Stimson:1997p3837; @Brown:1999p3534; @Fitt:2000p3792; @Titlyanov:2001p3631].
-As higher light alleviates light-limitation of photosynthesis, growth becomes less carbon-limited.
-Similarly, elevating DIN up to ~2 µM alleviates nitrogen-limitation (Fig.
-2B).
-Increased growth at higher DIN is predicted by the DEB model of @Muller:2009io, and has also been observed experimentally [@MullerParker:1994uo; @Tanaka:2007uv; @Tanaka:2013dj].
-However, DIN elevation beyond a certain point (e.g., ~3-4 µM in these simulations) has little effect on growth as carbon becomes limiting (Fig.
-2B).
-Very high nutrient levels may even reduce growth [@Shantz:2015dz], although these impacts are not likely to occur within the range of concentrations considered here (<4 µM) [@FerrierPages:2000p1950].
-
-In addition to increasing growth, DIN also increases the symbiont to host biomass ratio, a phenomenon also observed in reef corals [@Marubini:1996p8460].
-S:H ratios also increase under low light [@Stimson:1997p3837; @Fitt:2000p3792; @Anthony:2003p3632].
-At low DIN and intermediate light, more typical of coral reef environments, symbiont to host biomass ratios are around ~0.1-0.3, which is consistent with values reported in the literature [@Muscatine:1981jy; @Edmunds:2011bv].
-
 ![Steady state values of **A)** specific growth (Cmol Cmol^-1^ d^-1^), **B)** the symbiont to host biomass ratio (CmolS CmolH^-1^), **C)** relative C- or N-limitation of host biomass formation, and **D)** relative CO~2~- or light-limitation of symbiont photosynthesis, across gradients of external irradiance and dissolved inorganic nitrogen.
@@
 
+To analyze qualitative behavior, we ran the model to steady state across gradients of external irradiance and nutrients (Fig.
+2), which revealed patterns consistent with observed phenomena in corals.
+Predicted growth rates are low at low light and DIN (~0.01 d^-1^), and begin increasing as both of these factors increase (Fig.
+2A).
+Low light limits photosynthetic rates, resulting in less fixed carbon shared with the host and an associated increase in the symbiont to host biomass ratio (Fig.
+2B).
+In agreement with this trend are many observations of negative correlation between irradiance and symbiont density [@Stimson:1997p3837; @Brown:1999p3534; @Fitt:2000p3792; @Titlyanov:2001p3631].
+As higher light alleviates light-limitation of photosynthesis, growth becomes less carbon-limited.
+
+Similarly, increasing DIN up to ~2 µM alleviates nitrogen-limitation (Fig.
+2A).
+Increased growth at higher DIN is predicted by the DEB model of @Muller:2009io, and has also been observed experimentally [@MullerParker:1994uo; @Tanaka:2007uv; @Tanaka:2013dj].
+However, DIN elevation beyond a certain point (e.g., ~3-4 µM in these simulations) has little effect on growth as carbon becomes limiting.
+Very high nutrient levels may even reduce growth [@Shantz:2015dz], although these impacts are not likely to occur within the range of concentrations considered here (<4 µM) [@FerrierPages:2000p1950].
+In addition to increasing growth, DIN also increases the symbiont to host biomass ratio (Fig.
+2B), a phenomenon also observed in reef corals [@Marubini:1996p8460].
+At low DIN and intermediate light, more typical of coral reef environments, symbiont to host biomass ratios are around ~0.1-0.3, which is consistent with values reported in the literature [@Muscatine:1981jy; @Edmunds:2011bv; @Hawkins:2016jj].
+
 The maximum predicted growth rates of ~0.1 d^-1^, occurring between ~10-20 mol photons m^-2^ s^-1^ light and ~4 µM DIN (Fig.
 2A), are comparable to the rate of `r Tanaka_gr` d^-1^ measured by @Tanaka:2007uv under similar N-enriched conditions.
-Growth under conditions more typical of reef environments (<0.5 µM DIN) are within the range of ~0.01-0.03 d^-1^; these values are slightly higher than most specific growth rates reported in the literature, which are typically based on skeletal mass, and fall near or below 0.01 d^-1^ [@Osinga:2011kq; @Osinga:2012is] (though values as high as 0.025 d^-1^ have been reported [@Schutter:2010p7758]).
-These generally low skeletal growth rates relative to predicted biomass growth may be explained by differential substrate-limitation (i.e., CO~2~ may limit calcification, which is not present in the model); indeed, tissue and skeletal growth may not always be well correlated [@Anthony:2002tc].
-Furthermore, bioerosion may reduce net skeletal growth, especially in the field [@Bak:1976bv].
+Predicted growth rates under conditions more typical of reef environments (<0.5 µM DIN) are ~0.01-0.03 d^-1^.
+Observed specific growth rates, which are typically based on skeletal mass, fall near or below 0.01 d^-1^ [@Osinga:2011kq; @Osinga:2012is] (though values as high as 0.025 d^-1^ have been reported [@Schutter:2010p7758]).
+That predicted growth rates are slightly higher than most observations is not surprising, as the model does not account for external ecological factors that may reduce growth, such as competition, predation, and bioerosion.
+Furthermore, the model does not explicitly account for skeletal formation, which may not always be well correlated with biomass growth, potentially due to differential substrate-limitation [@Anthony:2002tc].
 Nevertheless, when rates of increase in coral tissue biomass have been directly measured over short periods, values are consistent with model predictions (e.g., `r Tanaka_gr` d^-1^ in @Tanaka:2007uv), and similar rates (`r Armoza_gr` d^-1^) have been measured in *Aiptasia diaphana*, a non-calcifying symbiotic anemone [@ArmozaZvuloni:2014ju].
 
-As irradiance continues to increase above ~20 mol photons m^-2^ d^-1^, growth rates decline until positive growth ceases above ~35 µmol photons m^-2^ d^-1^.
+As irradiance continues to increase above ~20 mol photons m^-2^ d^-1^, growth rates decline until positive growth ceases above ~35 µmol photons m^-2^ d^-1^ (Fig.
+2A).
 The mechanism underlying this decline is the increase in light energy beyond the capacity of photosynthesis and non-photochemical quenching: excess excitation energy generates reactive oxygen species (ROS) [@Weis:2008p944; @Roth:2014wf], which, in this model, have the phenomenological consequences of reducing the photosynthetic rate (representing photoinhibition) and increasing symbiont biomass loss (representing photodamage and/or symbiont expulsion) [see @Eynaud:2011tv].
 Together, these impacts reduce the symbiont to host biomass ratio (Fig.
-2C), as occurs during coral bleaching.
+2B), as occurs during coral bleaching.
 This reduction in symbionts consequently reduces the flux of fixed carbon to the host, resulting in increasing carbon-limitation (Fig.
@@
 2A).
-Importantly, these consequences arise in the present model only when light energy exceeds the capacity of photochemistry and NPQ, which in turn depends on CO~2~ availability.
-While previous models framed photophysiological stress as a fixed response to absolute irradiance levels [@Eynaud:2011tv], the current implementation considers the dynamic balance of multiple energy sinks in the causation of stress, which is more consistent with current understanding of symbiosis dysfunction [@Wooldridge:2013tj], and establishes an important role of the host in providing CO~2~ for symbiont photosynthesis [@Tansik:2015cm; @Hopkinson:2015ev].
 
@@
 In reality, light and temperature interact synergistically [@Coles:1978p1124; @Jones:1998p3572], and in fact, any stressor that disrupts the quenching of light energy may lead to bleaching [@Wooldridge:2010p7809; @Baker:2015kp].
-This is because the proximate cause of coral bleaching is excess excitation energy, but the upstream events that lead to this situation may be diverse: indeed, elevated temperature inhibits Rubisco functioning [@Jones:1998p3572] and repair of the D1 protein in photosystem II [@Warner:1999p4239], which reduces the capacity of photochemical quenching and leads to an excess of light energy.
+This is because the proximate cause of coral bleaching is excess excitation energy, but the upstream events that lead to this situation may be diverse: indeed, elevated temperature may inhibit Rubisco functioning [@Jones:1998p3572] and the repair of the D1 protein in photosystem II [@Warner:1999p4239], which reduces the capacity of photochemical quenching and leads to an excess of light energy.
 In this way, elevated temperature serves to reduce the threshold above which light causes stress to the system [@HoeghGuldberg:1999p4222]; importantly, light is still the proximate stressor.
-Therefore, we omitted temperature from the model, which serves to maintain a desired level of simplicity, while still allowing photophysiological stress and bleaching to be simulated with biological realism in response to light.
-Thus, the specific light levels that cause stress here should be interpreted in this context, and the effects of high light are analagous to the effects of high temperature.
+Therefore, we omitted temperature from the model to maintain a desired level of simplicity, while still allowing photophysiological stress and bleaching to be simulated with biological realism in response to light.
+Thus, while the model is currently accurate from a conceptual and mechanistic standpoint, its value as a predictive tool under scenarios of changing temperature will require further development.
 
 The incorporation of light stress in the model sets an upper limit to the amount of light at which a stable symbiotic interaction can be maintained, but even below this threshold of breakdown, negative effects of high light reduce steady state growth and symbiont:host biomass (Fig.
-2A,C).
+2).
 This gradual decline is consistent with experimental results showing that high light levels decrease growth [@Schutter:2011ie], and field studies documenting optimum growth rates at intermediate depths [@Baker:1975ip; @Huston:1985ec].
-Consequently, the model predicts greater variation in state variables across light gradients than was predicted by the model of @Muller:2009io, which did not include photoinhibition and photodamage.
-It is important to recognize that these are steady state values predicted under a constant environment at the given light and nutrient levels; a dynamic system may temporarily cross this threshold and experience a period of symbiont loss (bleaching) and reduced growth, after which a return to benign conditions may restore symbiont biomass and positive growth.
-To explore this further and illustrate the behavior of the model in more detail, we evaluate a number of dynamic simulations below.
+By incorporating these impacts of light stress, the model predicts greater, and more realistic, variation in state variables across light gradients than was predicted by the model of @Muller:2009io, which did not include photoinhibition and photodamage.
 
-## Sensitivity analysis -- should come after the steady states bc describes changes in steady states
+It is important to recognize that the boundary set by photo-oxidative stress on a stable symbiosis under steady state conditions (Fig.
+2) may be temporarily crossed by a dynamic system, which may experience a period of symbiont loss (bleaching) and reduced growth, after which a return to benign conditions may restore symbiont biomass and positive growth.
+To explore this further and illustrate the behavior of the model in more detail, we evaluate a number of dynamic simulations below (see "Dynamic behavior").
+
+## Sensitivity analysis
 
 Each default parameter value used in the model (Table 2) is based on carefully selected relevant literature.
-A supplementary document detailing the derivations of each of these default parameter values is provided.
+The derivations of each of these default parameter values are described in the Supplementary Information.
 Nevertheless, it is important to evaluate the sensitivity of the model to changes in these parameter values.
-We measured fractional change in steady state values in response to fractional change in parameter values, relative to their default values (Table 2).
+Moreover, these sensitivity plots serve to demonstrate the behavior of the dynamical system.
+We measured fractional change in steady state values in response to fractional changes in parameter values, relative to their default values, under environmental conditions typical of coral reefs.
 
@@
 Sensitivity analyses conducted other environmental conditions are presented in the Supplementary Information.](../img/Fig3.png)
+
+Overall, relative changes in the steady state of the system were less than the equivalent relative change in parameter value, indicating that the system maintains internal stability.
+However, changes in certain parameter values have greater impacts: increasing $j_{Nm}$ or decreasing $K_N$ both dramatically increase host growth, demonstrating the strong nitrogen-limitation that characterizes these symbioses (REF).
+The parameterr $\bar{a}^*$ has a strong impact on S:H biomass ratios, because this parameter determines the amount of light absorbed by symbionts, thus low values of this parameter increase light-limitation of the system.
+Altering the maximum growth and turnover rates of symbiont and host biomass also have expected impacts on steady states.
+Parameters having to do with photooxidative stress and bleaching have little impact under these environmental conditions, but of course have larger impacts under higher light (e.g., Fig.
+S6).
 
@@
 ## C-limitation increases susceptibility to bleaching
+Importantly, these negative impacts of high irradiance only arise when light energy exceeds the capacity of photochemistry and NPQ, which in turn depends on CO~2~ availability.
+While previous models framed photophysiological stress as a fixed response to absolute irradiance levels [@Eynaud:2011tv], the current implementation considers the dynamic balance of multiple energy sinks in the causation of stress, which is more consistent with current understanding of symbiosis dysfunction [@Wooldridge:2013tj], and establishes an important role of the host in providing CO~2~ for symbiont photosynthesis [@Tansik:2015cm; @Hopkinson:2015ev].
 ## consequences of a less efficient but more tolerant symbiont

### 2016-10-19 JTB_ms/JTB_ms.tex
text updates

@@
 
-To analyze qualitative behavior, we ran the model to steady state across
-gradients of external irradiance and nutrients (Fig. 2), which revealed
-patterns consistent with observed phenomena in corals. Predicted growth
-rates are low at low light and DIN (\textasciitilde{}0.01
-d\textsuperscript{-1}), and begin increasing as both of these factors
-increase (Fig. 2A). Low light is limiting to photosynthesis (Fig. 2D),
-which leads to carbon-limitation of host growth (Fig. 2B) and an
-associated increase in the symbiont to host biomass ratio (Fig. 2C). In
-agreement with this trend are many observations of negative correlation
-between irradiance and symbiont density (Stimson 1997; Brown et al.
-1999; Fitt et al. 2000; Titlyanov et al. 2001). As higher light
-alleviates light-limitation of photosynthesis, growth becomes less
-carbon-limited. Similarly, elevating DIN up to \textasciitilde{}2 µM
-alleviates nitrogen-limitation (Fig. 2B). Increased growth at higher DIN
-is predicted by the DEB model of Muller et al. (2009), and has also been
-observed experimentally (Muller-Parker et al. 1994; Tanaka et al. 2007;
-Tanaka et al. 2013). However, DIN elevation beyond a certain point
-(e.g., \textasciitilde{}3-4 µM in these simulations) has little effect
-on growth as carbon becomes limiting (Fig. 2B). Very high nutrient
-levels may even reduce growth (Shantz, Lemoine, and Burkepile 2015),
-although these impacts are not likely to occur within the range of
-concentrations considered here (\textless{}4 µM) (Ferrier-Pagès et al.
-2000).
-
-In addition to increasing growth, DIN also increases the symbiont to
-host biomass ratio, a phenomenon also observed in reef corals (Marubini
-and Davies 1996). S:H ratios also increase under low light (Stimson
-1997; Fitt et al. 2000; Anthony and Hoegh-Guldberg 2003). At low DIN and
-intermediate light, more typical of coral reef environments, symbiont to
-host biomass ratios are around \textasciitilde{}0.1-0.3, which is
-consistent with values reported in the literature (Muscatine, R
-McCloskey, and E Marian 1981; Edmunds et al. 2011).
-
 \begin{figure}[htbp]
@@
 
+To analyze qualitative behavior, we ran the model to steady state across
+gradients of external irradiance and nutrients (Fig. 2), which revealed
+patterns consistent with observed phenomena in corals. Predicted growth
+rates are low at low light and DIN (\textasciitilde{}0.01
+d\textsuperscript{-1}), and begin increasing as both of these factors
+increase (Fig. 2A). Low light limits photosynthetic rates, resulting in
+less fixed carbon shared with the host and an associated increase in the
+symbiont to host biomass ratio (Fig. 2B). In agreement with this trend
+are many observations of negative correlation between irradiance and
+symbiont density (Stimson 1997; Brown et al. 1999; Fitt et al. 2000;
+Titlyanov et al. 2001). As higher light alleviates light-limitation of
+photosynthesis, growth becomes less carbon-limited.
+
+Similarly, increasing DIN up to \textasciitilde{}2 µM alleviates
+nitrogen-limitation (Fig. 2A). Increased growth at higher DIN is
+predicted by the DEB model of Muller et al. (2009), and has also been
+observed experimentally (Muller-Parker et al. 1994; Tanaka et al. 2007;
+Tanaka et al. 2013). However, DIN elevation beyond a certain point
+(e.g., \textasciitilde{}3-4 µM in these simulations) has little effect
+on growth as carbon becomes limiting. Very high nutrient levels may even
+reduce growth (Shantz, Lemoine, and Burkepile 2015), although these
+impacts are not likely to occur within the range of concentrations
+considered here (\textless{}4 µM) (Ferrier-Pagès et al. 2000). In
+addition to increasing growth, DIN also increases the symbiont to host
+biomass ratio (Fig. 2B), a phenomenon also observed in reef corals
+(Marubini and Davies 1996). At low DIN and intermediate light, more
+typical of coral reef environments, symbiont to host biomass ratios are
+around \textasciitilde{}0.1-0.3, which is consistent with values
+reported in the literature (Muscatine, R McCloskey, and E Marian 1981;
+Edmunds et al. 2011; Hawkins et al. 2016).
+
 The maximum predicted growth rates of \textasciitilde{}0.1
@@
 d\textsuperscript{-1} measured by Tanaka et al. (2007) under similar
-N-enriched conditions. Growth under conditions more typical of reef
-environments (\textless{}0.5 µM DIN) are within the range of
-\textasciitilde{}0.01-0.03 d\textsuperscript{-1}; these values are
-slightly higher than most specific growth rates reported in the
-literature, which are typically based on skeletal mass, and fall near or
+N-enriched conditions. Predicted growth rates under conditions more
+typical of reef environments (\textless{}0.5 µM DIN) are
+\textasciitilde{}0.01-0.03 d\textsuperscript{-1}. Observed specific
+growth rates, which are typically based on skeletal mass, fall near or
 below 0.01 d\textsuperscript{-1} (Osinga et al. 2011; Osinga et al.
 2012) (though values as high as 0.025 d\textsuperscript{-1} have been
-reported (Schutter et al. 2010)). These generally low skeletal growth
-rates relative to predicted biomass growth may be explained by
-differential substrate-limitation (i.e., CO\textsubscript{2} may limit
-calcification, which is not present in the model); indeed, tissue and
-skeletal growth may not always be well correlated (Anthony 2002).
-Furthermore, bioerosion may reduce net skeletal growth, especially in
-the field (Bak 1976). Nevertheless, when rates of increase in coral
-tissue biomass have been directly measured over short periods, values
-are consistent with model predictions (e.g., 0.07 d\textsuperscript{-1}
-in Tanaka et al. (2007)), and similar rates (0.04 d\textsuperscript{-1})
-have been measured in \emph{Aiptasia diaphana}, a non-calcifying
-symbiotic anemone (Armoza-Zvuloni et al. 2014).
+reported (Schutter et al. 2010)). That predicted growth rates are
+slightly higher than most observations is not surprising, as the model
+does not account for external ecological factors that may reduce growth,
+such as competition, predation, and bioerosion. Furthermore, the model
+does not explicitly account for skeletal formation, which may not always
+be well correlated with biomass growth, potentially due to differential
+substrate-limitation (Anthony 2002). Nevertheless, when rates of
+increase in coral tissue biomass have been directly measured over short
+periods, values are consistent with model predictions (e.g., 0.07
+d\textsuperscript{-1} in Tanaka et al. (2007)), and similar rates (0.04
+d\textsuperscript{-1}) have been measured in \emph{Aiptasia diaphana}, a
+non-calcifying symbiotic anemone (Armoza-Zvuloni et al. 2014).
 
@@
 decline until positive growth ceases above \textasciitilde{}35 µmol
-photons m\textsuperscript{-2} d\textsuperscript{-1}. The mechanism
-underlying this decline is the increase in light energy beyond the
-capacity of photosynthesis and non-photochemical quenching: excess
+photons m\textsuperscript{-2} d\textsuperscript{-1} (Fig. 2A). The
+mechanism underlying this decline is the increase in light energy beyond
+the capacity of photosynthesis and non-photochemical quenching: excess
 excitation energy generates reactive oxygen species (ROS) (Weis 2008;
@@
 symbiont expulsion) (see Eynaud, Nisbet, and Muller 2011). Together,
-these impacts reduce the symbiont to host biomass ratio (Fig. 2C), as
+these impacts reduce the symbiont to host biomass ratio (Fig. 2B), as
 occurs during coral bleaching. This reduction in symbionts consequently
@@
 carbon-limitation (Fig. 2B) and eventual cessation of growth (Fig. 2A).
-Importantly, these consequences arise in the present model only when
-light energy exceeds the capacity of photochemistry and NPQ, which in
-turn depends on CO\textsubscript{2} availability. While previous models
-framed photophysiological stress as a fixed response to absolute
-irradiance levels (Eynaud, Nisbet, and Muller 2011), the current
-implementation considers the dynamic balance of multiple energy sinks in
-the causation of stress, which is more consistent with current
-understanding of symbiosis dysfunction (Wooldridge 2013), and
-establishes an important role of the host in providing
-CO\textsubscript{2} for symbiont photosynthesis (Tansik, Fitt, and
-Hopkinson 2015; Hopkinson, Tansik, and Fitt 2015).
 
@@
 energy, but the upstream events that lead to this situation may be
-diverse: indeed, elevated temperature inhibits Rubisco functioning
-(Jones et al. 1998) and repair of the D1 protein in photosystem II
+diverse: indeed, elevated temperature may inhibit Rubisco functioning
+(Jones et al. 1998) and the repair of the D1 protein in photosystem II
 (Warner, Fitt, and Schmidt 1999), which reduces the capacity of
@@
 light is still the proximate stressor. Therefore, we omitted temperature
-from the model, which serves to maintain a desired level of simplicity,
-while still allowing photophysiological stress and bleaching to be
-simulated with biological realism in response to light. Thus, the
-specific light levels that cause stress here should be interpreted in
-this context, and the effects of high light are analagous to the effects
-of high temperature.
+from the model to maintain a desired level of simplicity, while still
+allowing photophysiological stress and bleaching to be simulated with
+biological realism in response to light. Thus, while the model is
+currently accurate from a conceptual and mechanistic standpoint, its
+value as a predictive tool under scenarios of changing temperature will
+require further development.
 
@@
 of high light reduce steady state growth and symbiont:host biomass (Fig.
-2A,C). This gradual decline is consistent with experimental results
-showing that high light levels decrease growth (Schutter et al. 2011),
-and field studies documenting optimum growth rates at intermediate
-depths (Baker and Weber 1975; Huston 1985). Consequently, the model
-predicts greater variation in state variables across light gradients
-than was predicted by the model of Muller et al. (2009), which did not
-include photoinhibition and photodamage. It is important to recognize
-that these are steady state values predicted under a constant
-environment at the given light and nutrient levels; a dynamic system may
-temporarily cross this threshold and experience a period of symbiont
-loss (bleaching) and reduced growth, after which a return to benign
-conditions may restore symbiont biomass and positive growth. To explore
-this further and illustrate the behavior of the model in more detail, we
-evaluate a number of dynamic simulations below.
-
-\subsection{Sensitivity analysis -- should come after the steady states
-bc describes changes in steady
-states}\label{sensitivity-analysis-should-come-after-the-steady-states-bc-describes-changes-in-steady-states}
+2). This gradual decline is consistent with experimental results showing
+that high light levels decrease growth (Schutter et al. 2011), and field
+studies documenting optimum growth rates at intermediate depths (Baker
+and Weber 1975; Huston 1985). By incorporating these impacts of light
+stress, the model predicts greater, and more realistic, variation in
+state variables across light gradients than was predicted by the model
+of Muller et al. (2009), which did not include photoinhibition and
+photodamage.
+
+It is important to recognize that the boundary set by photo-oxidative
+stress on a stable symbiosis under steady state conditions (Fig. 2) may
+be temporarily crossed by a dynamic system, which may experience a
+period of symbiont loss (bleaching) and reduced growth, after which a
+return to benign conditions may restore symbiont biomass and positive
+growth. To explore this further and illustrate the behavior of the model
+in more detail, we evaluate a number of dynamic simulations below (see
+``Dynamic behavior'').
+
+\subsection{Sensitivity analysis}\label{sensitivity-analysis}
 
 Each default parameter value used in the model (Table 2) is based on
-carefully selected relevant literature. A supplementary document
-detailing the derivations of each of these default parameter values is
-provided. Nevertheless, it is important to evaluate the sensitivity of
-the model to changes in these parameter values. We measured fractional
-change in steady state values in response to fractional change in
-parameter values, relative to their default values (Table 2).
+carefully selected relevant literature. The derivations of each of these
+default parameter values are described in the Supplementary Information.
+Nevertheless, it is important to evaluate the sensitivity of the model
+to changes in these parameter values. Moreover, these sensitivity plots
+serve to demonstrate the behavior of the dynamical system. We measured
+fractional change in steady state values in response to fractional
+changes in parameter values, relative to their default values, under
+environmental conditions typical of coral reefs.
 
@@
 
+Overall, relative changes in the steady state of the system were less
+than the equivalent relative change in parameter value, indicating that
+the system maintains internal stability. However, changes in certain
+parameter values have greater impacts: increasing \(j_{Nm}\) or
+decreasing \(K_N\) both dramatically increase host growth, demonstrating
+the strong nitrogen-limitation that characterizes these symbioses (REF).
+The parameterr \(\bar{a}^*\) has a strong impact on S:H biomass ratios,
+because this parameter determines the amount of light absorbed by
+symbionts, thus low values of this parameter increase light-limitation
+of the system. Altering the maximum growth and turnover rates of
+symbiont and host biomass also have expected impacts on steady states.
+Parameters having to do with photooxidative stress and bleaching have
+little impact under these environmental conditions, but of course have
+larger impacts under higher light (e.g., Fig. S6).
+
 \subsection{Dynamic behavior}\label{dynamic-behavior}
@@
 become C-limited (photo: CO2 limited; host: fixed-C limited) \#\#
-C-limitation increases susceptibility to bleaching \#\# consequences of
-a less efficient but more tolerant symbiont \#\# Rate of symbiont
-repopulation (Berner et al. 1993) \#\# Bicarbonate addition promotes
-coral growth (Marubini and Thake 1999) \#\# Seasonal change in sym.
-density is larger when nutrients are elevated (Stimson 1997). Such a
-phenomenon was predicted by Fitt et al. 2000. \#\# Death rule: negative
-growth for a certain amount of time should be defensible.
+C-limitation increases susceptibility to bleaching Importantly, these
+negative impacts of high irradiance only arise when light energy exceeds
+the capacity of photochemistry and NPQ, which in turn depends on
+CO\textsubscript{2} availability. While previous models framed
+photophysiological stress as a fixed response to absolute irradiance
+levels (Eynaud, Nisbet, and Muller 2011), the current implementation
+considers the dynamic balance of multiple energy sinks in the causation
+of stress, which is more consistent with current understanding of
+symbiosis dysfunction (Wooldridge 2013), and establishes an important
+role of the host in providing CO\textsubscript{2} for symbiont
+photosynthesis (Tansik, Fitt, and Hopkinson 2015; Hopkinson, Tansik, and
+Fitt 2015). \#\# consequences of a less efficient but more tolerant
+symbiont \#\# Rate of symbiont repopulation (Berner et al. 1993) \#\#
+Bicarbonate addition promotes coral growth (Marubini and Thake 1999)
+\#\# Seasonal change in sym. density is larger when nutrients are
+elevated (Stimson 1997). Such a phenomenon was predicted by Fitt et al.
+2000. \#\# Death rule: negative growth for a certain amount of time
+should be defensible.
 
@@
 \url{https://www.scopus.com/inward/record.uri?partnerID=HzOxMe3b\&scp=0036733455\&origin=inward}.
-
-\hypertarget{ref-Anthony:2003p3632}{}
-Anthony, Kenneth R N, and Ove Hoegh-Guldberg. 2003. ``Variation in coral
-photosynthesis, respiration and growth characteristics in contrasting
-light microhabitats: an analogue to plants in forest gaps and
-understoreys?'' \emph{Functional Ecology} 17 (April): 246--59.
 
@@
 Science} 69 (1): 79--109.
+
+\hypertarget{ref-Hawkins:2016jj}{}
+Hawkins, Thomas D, Julia C G Hagemeyer, Kenneth D Hoadley, Adam G Marsh,
+and Mark E Warner. 2016. ``Partitioning of Respiration in an
+Animal-Algal Symbiosis: Implications for Different Aerobic Capacity
+between Symbiodinium spp.'' \emph{Frontiers in Physiology} 7 (April).
+Frontiers: 125.
+doi:\href{https://doi.org/10.3389/fphys.2016.00128}{10.3389/fphys.2016.00128}.
 

### 2016-10-25 JTB_ms/JTB_ms.Rmd
finish drafting main text

@@
 _Provide well-documented, open-access code._ In order to facilitate the continued development and application of theoretical modeling tools for coral symbioses, we provide open access to the model in the form of detailed, commented code written in the R language [@RALanguageandEn:2014wf].
-With an accessible and modular framework, we envision this as a resource for futher development by the scientific community to include additional complexity and problem-specific components.
+With an accessible and modular framework, we envision this model as a resource for futher development by the scientific community to include additional complexity and problem-specific components.
 To widen the audience for this work, we chose the R language because it is freely available and in common use by biologists and ecologists.
@@
 The two state variables are symbiont biomass and coral biomass; because resources are acquired proportionally to surface area, and surface area is proportional to volume (i.e., corals are "V1-morphs" in DEB terminology [@Kooijman:2010vd]), biomass increases exponentially during growth (indeed, corals grow exponentially [@Bak:1976bv]).
-The specific growth rate and the ratio of symbiont to host biomass are the responses of interest of the system.
-Below we describe the formulation of each flux involved in producing these responses.
+
+Environmental stress is implemented in the form of photo-oxidative stress, which is thought to be the primary trigger of coral bleaching [@Lesser:1997p3587; @Weis:2008p944; @Wooldridge:2009p7807].
+To simulate these events, we model the absorption and quenching of light energy by photochemistry and non-photochemical quenching, and the phenomenological responses that occur when these capacities are overwhelmed (i.e., photoinhibition, photodamage, and symbiont loss).
+While bleaching in response to high light alone has been observed experimentally [@Schutter:2011ie; @Downs:2013kc], mass coral bleaching events occur in response to high temperature [@HoeghGuldberg:1999p4222]; thus, it is important to justify our consideration of light as the primary stressor in the model.
+In reality, light and temperature interact synergistically [@Coles:1978p1124; @Jones:1998p3572], and in fact, any stressor that disrupts the quenching of light energy may lead to bleaching [@Wooldridge:2010p7809; @Baker:2015kp].
+This is because the proximate cause of photo-oxidative stress is excess excitation energy, but the upstream events that lead to this situation may be diverse: indeed, elevated temperature may inhibit Rubisco functioning [@Jones:1998p3572] and the repair of the D1 protein in photosystem II [@Warner:1999p4239], which reduces the capacity of photochemical quenching and leads to an excess of light energy.
+In this way, elevated temperature serves to reduce the threshold above which light stresses the system [@HoeghGuldberg:1999p4222]; importantly, light is still the proximate stressor.
+Therefore, we omitted temperature from the model to maintain a desired level of simplicity, while still allowing photophysiological stress and bleaching to be simulated with biological realism in response to light.
+
+Below we describe the formulation of each flux in the model.
 
@@
 | Symbol  | Description                            | Units               | Eq. no. |
-| --------- | -------------------------------------- | -------------------- | ------- |
+| ------- | -------------------------------------- | ------------------- | ------- |
 | $j_X$     | Prey uptake rate                       | molX CmolH^-1^ d^-1^ | 3 |
@@
 
-| Symbol  | Description                            | Value  | Units              |
-| ------- | -------------------------------------- | ------ | ------------------ |
-| $n_{NH}$ | N:C molar ratio in host biomass | 0.19 | -- |
-| $n_{NS}$ | N:C molar ratio in symbiont biomass | 0.2 | -- |
-| $n_{NX}$ | N:C molar ratio in prey biomass | 0.13 | -- |
-| $j_{HT}^0$ | Specific turnover rate of host biomass | 0.03 | CmolH CmolH^-1^ d^-1^ |
-| $j_{ST}^0$ | Specific turnover rate of symbiont biomass | 0.03 | CmolS CmolS^-1^ d^-1^ |
-| $\sigma_{NH}$ | Proportion host nitrogen turnover recycled | 0.9 | -- |
-| $\sigma_{CH}$ | Proportion host carbon turnover recycled | 0.9 | -- |
-| $\sigma_{NS}$ | Proportion symbiont nitrogen turnover recycled | 0.9 | -- |
-| $\sigma_{CS}$ | Proportion symbiont carbon turnover recycled | 0.9 | -- |
-| $j_{Xm}$ | Maximum specific feeding rate of host | 0.1292 | molX CmolH^-1^ d^-1^ |
-| $K_X$ | Half-saturation constant for prey uptake by host | 20e-6 | molX L^-1^ | 
-| $j_{Nm}$ | Maximum specific DIN uptake rate by host | 0.048 | molN CmolH^-1^ d^-1^ |
-| $K_N$ | Half-saturation constant for DIN uptake by host | 0.46e-6 | molN L^-1^ | 
-| $k_{CO_2}$ | Efficacy of CO~2~ delivery by host CCMs | 10 | molCO~2~ molC^-1^ |
-| $j_{HGm}$ | Maximum specific growth rate of host | 1 | CmolH CmolH^-1^ d^-1^ |
-| $y_{CL}$ | Quantum yield of photosynthesis | 0.1 | molC mol ph^-1^ |
-| $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | 1.34 | m^2^ CmolS^-1^ |
-| $k_{NPQ}$ | NPQ capacity of symbiont | 112 | mol ph CmolS^-1^ d^-1^ |
-| $k_{ROS}$ | Excess photon energy that doubles ROS production, relative to baseline levels | 40-80 | mol ph CmolS^-1^ d^-1^ |
-| $j_{CPm}$ | Maximum specific photosynthesis rate of symbiont | 2.8 | molC CmolS^-1^ d^-1^ |
-| $j_{SGm}$ | Maximum specific growth rate of symbiont | 0.25 | CmolS CmolS^-1^ d^-1^ |
-| $b$ | Scaling parameter for strength of bleaching response | 5 | -- |
+| Symbol  | Description                            | Units               | Value   |
+| ------- | -------------------------------------- | ------------------- | ------- |
+| $n_{NH}$ | N:C molar ratio in host biomass | -- | 0.19 |
+| $n_{NS}$ | N:C molar ratio in symbiont biomass | -- | 0.2 |
+| $n_{NX}$ | N:C molar ratio in prey biomass | -- | 0.13 |
+| $j_{HT}^0$ | Specific turnover rate of host biomass | CmolH CmolH^-1^ d^-1^ | 0.03 |
+| $j_{ST}^0$ | Specific turnover rate of symbiont biomass | CmolS CmolS^-1^ d^-1^ | 0.03 |
+| $\sigma_{NH}$ | Proportion host nitrogen turnover recycled | -- | 0.9 |
+| $\sigma_{CH}$ | Proportion host carbon turnover recycled | -- | 0.9 |
+| $\sigma_{NS}$ | Proportion symbiont nitrogen turnover recycled | -- | 0.9 |
+| $\sigma_{CS}$ | Proportion symbiont carbon turnover recycled | -- | 0.9 |
+| $j_{Xm}$ | Maximum specific feeding rate of host | molX CmolH^-1^ d^-1^ | 0.1292 |
+| $K_X$ | Half-saturation constant for prey uptake by host | molX L^-1^ | 20e-6 |
+| $j_{Nm}$ | Maximum specific DIN uptake rate by host | molN CmolH^-1^ d^-1^ | 0.048 |
+| $K_N$ | Half-saturation constant for DIN uptake by host | molN L^-1^ | 0.46e-6 |
+| $k_{CO_2}$ | Efficacy of CO~2~ delivery by host CCMs | molCO~2~ molC^-1^ | 10 |
+| $j_{HGm}$ | Maximum specific growth rate of host | CmolH CmolH^-1^ d^-1^ | 1 |
+| $y_{CL}$ | Quantum yield of photosynthesis | molC mol ph^-1^ | 0.1 |
+| $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | m^2^ CmolS^-1^ | 1.34 |
+| $k_{NPQ}$ | NPQ capacity of symbiont | mol ph CmolS^-1^ d^-1^ | 112 |
+| $k_{ROS}$ | Excess photon energy that doubles ROS production, relative to baseline levels | mol ph CmolS^-1^ d^-1^ | 40-80 |
+| $j_{CPm}$ | Maximum specific photosynthesis rate of symbiont | molC CmolS^-1^ d^-1^ | 2.8 |
+| $j_{SGm}$ | Maximum specific growth rate of symbiont | CmolS CmolS^-1^ d^-1^ | 0.25 |
+| $b$ | Scaling parameter for strength of bleaching response | -- | 5 |
 
@@
 We then specify the input of CO~2~ to photosynthesis as $j_{CO_2}$, a flux mediated by the host that encompasses potentially diverse carbon-concentrating mechanisms (CCMs), including active transport of bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to CO~2~ to promote diffusion toward the symbiont [@Tansik:2015cm], and acidification of the symbiosome to increase localized CO~2~ concentrations around the symbiont [@Barott:2014gx].
-Since these mechanisms require energetic input by the host, the magnitude of CO~2~ delivery is proportional to the host's excess carbon, $j_{eC}$, assuming that some of this carbon is respired to energize these CCMs.
-This formulation means that the symbiont indirectly ensures its own CO~2~ supply by providing fixed carbon to the host [@Wooldridge:2013tj], which establishes a positive feedback in the model that leads to the existence of multiple stable states, which we discuss below in the context of coral bleaching.
+Since these active CCMs require energetic input by the host, the magnitude of CO~2~ delivery is proportional to the amount of fixed carbon available to the host after satisfying its growth demands, $j_{eC}$, assuming that some of this carbon is respired to energize these CCMs.
+This formulation means that the symbiont indirectly ensures its own CO~2~ supply by providing fixed carbon (=energy) to the host [@Wooldridge:2013tj], which establishes a positive feedback in the model that leads to the existence of multiple stable states, which we discuss below in the context of coral bleaching.
 The parameter $k_{CO_2}$ scales the efficacy of host CCMs, which enables the comparison of different rates of CO~2~ delivery that may characterize different coral species [@Wooldridge:2014di].
@@
 
-# Results
-## Qualitative dynamics
+# Steady state behavior
 In a constant environment, the system ultimately reaches a steady state of exponential growth or decline.
-Under some conditions, alternate steady states can be reached depending on initial values of symbiont and host biomass.
-The mechanism that produces these alternate steady states is the positive feedback between carbon-limitation of the host and CO~2~-limitation of photosynthesis.
-If the host becomes carbon-limited, the efficacy of its CCMs are reduced and photosynthesis becomes progressively more CO~2~-limited.
-In turn, less photosynthate is translocated to the host, causing further carbon-limitation and an ultimate decline in biomass; this steady state represents a dysfunctional symbiosis.
-However, if the host is not carbon-limited, it effectively delivers CO~2~ to photosynthesis and ensures its continued receipt of fixed carbon from the symbiont; this system is nitrogen-limited and represents a functional symbiosis with positive exponential growth.
-Either of these stable states may be reached under (some sets of) constant environmental conditions, depending on the initial values of symbiont and host biomass: if symbiont biomass is initially very low (i.e., a "bleached" coral), the system cannot escape the carbon-limited state and cannot grow.
-However, if symbiont biomass is initially high (i.e., a healthy coral), then the system maintains positive growth in a nitrogen-limited state.
-For practical purposes, we consider only positive growth steady states in the "steady state behavior" section below.
-In the subsequent "dynamic behavior" section, we explore more thoroughly the conditions that cause the system to switch between alternate stable states, and interpret these dynamics in the context of coral bleaching.
+However, under some conditions, either of these outcomes may occur depending on initial values of symbiont and host biomass, indicating the presence of alternate stable states.
+The mechanism that produces these alternate stable states is the positive feedback between carbon-limitation of the host and CO~2~-limitation of photosynthesis: if symbiont biomass is initially very low (i.e., a "bleached" coral), the system cannot escape positively reinforcing carbon-limitation and cannot grow.
+However, if symbiont biomass is initially high (i.e., a "healthy" coral), then the system maintains positive growth in a nitrogen-limited state.
+For practical purposes, this section of the manuscript considers only positive growth steady states under constant environments; the subsequent "dynamic behavior" section explores how changing environments may cause the system to switch between alternate stable states, which we interpret in the context of coral bleaching.
 
-## Steady state behavior
-
-![Steady state values of **A)** specific growth (Cmol Cmol^-1^ d^-1^), **B)** the symbiont to host biomass ratio (CmolS CmolH^-1^), **C)** relative C- or N-limitation of host biomass formation, and **D)** relative CO~2~- or light-limitation of symbiont photosynthesis, across gradients of external irradiance and dissolved inorganic nitrogen.
-Note that typical conditions for reefs are ~1e-7 M DIN and 10-20 mol photons m^-2^ d^-1^.
-Simulations for each combination of light and nutrients (41 points along each axis) were run for 100 days with a time step of 1 day.
-Negative steady state growth rates, and corresponding S:H ratios, were set to zero.](../img/Fig2.png)
-
-To analyze qualitative behavior, we ran the model to steady state across gradients of external irradiance and nutrients (Fig.
+To analyze qualitative patterns in steady state behavior, we ran the model to steady state across gradients of external irradiance and nutrients (Fig.
 2), which revealed patterns consistent with observed phenomena in corals.
@@
 As higher light alleviates light-limitation of photosynthesis, growth becomes less carbon-limited.
+
+![Steady state values of **A)** specific growth (Cmol Cmol^-1^ d^-1^), **B)** the symbiont to host biomass ratio (CmolS CmolH^-1^), **C)** relative C- or N-limitation of host biomass formation, and **D)** relative CO~2~- or light-limitation of symbiont photosynthesis, across gradients of external irradiance and dissolved inorganic nitrogen.
+Note that typical conditions for reefs are ~1e-7 M DIN and 10-20 mol photons m^-2^ d^-1^.
+Simulations for each combination of light and nutrients (41 points along each axis) were run for 100 days with a time step of 1 day.
+Negative steady state growth rates, and corresponding S:H ratios, were set to zero.](../img/Fig2.png)
 
@@
 Predicted growth rates under conditions more typical of reef environments (<0.5 µM DIN) are ~0.01-0.03 d^-1^.
-Observed specific growth rates, which are typically based on skeletal mass, fall near or below 0.01 d^-1^ [@Osinga:2011kq; @Osinga:2012is] (though values as high as 0.025 d^-1^ have been reported [@Schutter:2010p7758]).
-That predicted growth rates are slightly higher than most observations is not surprising, as the model does not account for external ecological factors that may reduce growth, such as competition, predation, and bioerosion.
-Furthermore, the model does not explicitly account for skeletal formation, which may not always be well correlated with biomass growth, potentially due to differential substrate-limitation [@Anthony:2002tc].
-Nevertheless, when rates of increase in coral tissue biomass have been directly measured over short periods, values are consistent with model predictions (e.g., `r Tanaka_gr` d^-1^ in @Tanaka:2007uv), and similar rates (`r Armoza_gr` d^-1^) have been measured in *Aiptasia diaphana*, a non-calcifying symbiotic anemone [@ArmozaZvuloni:2014ju].
+Observed specific growth rates, which are typically based on skeletal mass, fall near or below 0.01 d^-1^ [@Osinga:2011kq; @Osinga:2012is], though values as high as 0.025 d^-1^ have been reported [@Schutter:2010p7758].
+It is not surprising that predicted growth rates are slightly higher than most observations, as the model does not account for external ecological factors that may reduce growth, such as competition, predation, and bioerosion.
+Furthermore, the model does not explicitly account for skeletal formation, which may not always correlate with biomass growth, potentially due to differential substrate-limitation [@Anthony:2002tc].
+Nevertheless, when growth of coral tissue biomass has been directly measured over short periods, values are consistent with model predictions (e.g., `r Tanaka_gr` d^-1^ in @Tanaka:2007uv), and similar rates (`r Armoza_gr` d^-1^) have been measured in *Aiptasia diaphana*, a non-calcifying symbiotic anemone [@ArmozaZvuloni:2014ju].
 
@@
 
-While bleaching in response to high light alone has been observed experimentally [@Schutter:2011ie; @Downs:2013kc], mass coral bleaching events occur in response to high temperature [@HoeghGuldberg:1999p4222]; thus, it is important to justify our consideration of light as the primary stressor in the model.
-In reality, light and temperature interact synergistically [@Coles:1978p1124; @Jones:1998p3572], and in fact, any stressor that disrupts the quenching of light energy may lead to bleaching [@Wooldridge:2010p7809; @Baker:2015kp].
-This is because the proximate cause of coral bleaching is excess excitation energy, but the upstream events that lead to this situation may be diverse: indeed, elevated temperature may inhibit Rubisco functioning [@Jones:1998p3572] and the repair of the D1 protein in photosystem II [@Warner:1999p4239], which reduces the capacity of photochemical quenching and leads to an excess of light energy.
-In this way, elevated temperature serves to reduce the threshold above which light causes stress to the system [@HoeghGuldberg:1999p4222]; importantly, light is still the proximate stressor.
-Therefore, we omitted temperature from the model to maintain a desired level of simplicity, while still allowing photophysiological stress and bleaching to be simulated with biological realism in response to light.
-Thus, while the model is currently accurate from a conceptual and mechanistic standpoint, its value as a predictive tool under scenarios of changing temperature will require further development.
-
 The incorporation of light stress in the model sets an upper limit to the amount of light at which a stable symbiotic interaction can be maintained, but even below this threshold of breakdown, negative effects of high light reduce steady state growth and symbiont:host biomass (Fig.
@@
 
-## Sensitivity analysis
+# Sensitivity analysis
 
-Each default parameter value used in the model (Table 2) is based on carefully selected relevant literature.
-The derivations of each of these default parameter values are described in the Supplementary Information.
-Nevertheless, it is important to evaluate the sensitivity of the model to changes in these parameter values.
-Moreover, these sensitivity plots serve to demonstrate the behavior of the dynamical system.
+The values used for each parameter in the model (Table 2) are derived from relevant literature; these derivations are described in the Supplementary Information.
+Here we evaluate the sensitivity of the model to changes in these parameter values, which also serves to demonstrate the behavior of the dynamical system.
 We measured fractional change in steady state values in response to fractional changes in parameter values, relative to their default values, under environmental conditions typical of coral reefs.
@@
 
-Overall, relative changes in the steady state of the system were less than the equivalent relative change in parameter value, indicating that the system maintains internal stability.
-However, changes in certain parameter values have greater impacts: increasing $j_{Nm}$ or decreasing $K_N$ both dramatically increase host growth, demonstrating the strong nitrogen-limitation that characterizes these symbioses (REF).
-The parameterr $\bar{a}^*$ has a strong impact on S:H biomass ratios, because this parameter determines the amount of light absorbed by symbionts, thus low values of this parameter increase light-limitation of the system.
-Altering the maximum growth and turnover rates of symbiont and host biomass also have expected impacts on steady states.
-Parameters having to do with photooxidative stress and bleaching have little impact under these environmental conditions, but of course have larger impacts under higher light (e.g., Fig.
+Overall, relative changes in the steady state of the system are less than the equivalent relative change in parameter value, indicating that the system maintains internal stability.
+However, changes in certain parameter values have more significant impacts: increasing $j_{Nm}$ or decreasing $K_N$ both dramatically increase host growth (Fig.
+3), demonstrating the strong nitrogen-limitation that characterizes these symbioses (REF).
+The parameter $\bar{a}^*$ has a strong impact on S:H biomass ratios (Fig.
+3) since this parameter determines the amount of light absorbed by symbionts, with lower values increasing light-limitation.
+Altering maximum growth and turnover rates of symbiont and host also have expected impacts on steady states (Fig.
+3).
+Parameters relating to photooxidative stress and bleaching have little impact under these environmental conditions (Fig.
+3), but of course have larger impacts under higher light (e.g., Fig.
 S6).
+Sensitivity analyses conducted under different combinations of external light and nutrients are presented in the Supplementary Information (Figs.
+S3-S7)
 
-## Dynamic behavior
-Therefore, we are highly interested in the conditions under which the system switches from one state to another, representing the conditions under which corals "bleach" due to progressive, reinforcing carbon-limitation.
-Interestingly, this state can be reached in multiple ways, by the acceleration of photoinhibition and photodamage, or by high external DIN, which causes the symbiont to share less carbon, subsequently causing its own CO2-limitation, and progressive C-limitation of the system.
-## varying light: consequences of excess light and light-limitation
-## varying DIN: role of N-limitation and consequences of elevated N
-## elevated nutrients cause system to become C-limited (photo: CO2 limited; host: fixed-C limited)
-## C-limitation increases susceptibility to bleaching
-Importantly, these negative impacts of high irradiance only arise when light energy exceeds the capacity of photochemistry and NPQ, which in turn depends on CO~2~ availability.
-While previous models framed photophysiological stress as a fixed response to absolute irradiance levels [@Eynaud:2011tv], the current implementation considers the dynamic balance of multiple energy sinks in the causation of stress, which is more consistent with current understanding of symbiosis dysfunction [@Wooldridge:2013tj], and establishes an important role of the host in providing CO~2~ for symbiont photosynthesis [@Tansik:2015cm; @Hopkinson:2015ev].
-## consequences of a less efficient but more tolerant symbiont
-## Rate of symbiont repopulation (Berner et al. 1993)
-## Bicarbonate addition promotes coral growth (Marubini and Thake 1999)
-## Seasonal change in sym. density is larger when nutrients are elevated (Stimson 1997). Such a phenomenon was predicted by Fitt et al. 2000.
-## Death rule: negative growth for a certain amount of time should be defensible.
+# Dynamic behavior
+The dynamic behavior of the model demonstrates its power to integrate multiple environmental forcings simultaneously to produced integrated responses of the system.
+The potential applications of this are wide-ranging, and so the model may be highly useful for generating hypotheses about the way the system will respond to complex sets of environmental dynamics.
+Here we present several scenarios to highlight the power of the model and demonstrate that it can also reproduce more complex phenomena that have been observed in corals.
+
+## Seasonal variability
+Symbiont densities and coral growth rates are known to vary seasonally, representing an integrated response to changes in a suite of environmental factors.
+Light in particular is a strong driver of these trends [Stimson:1997p3837; @Brown:1999p3534; @Fagoonee:1999p4136; @Fitt:2000p3792], with high light associated with lower symbiont abundance and reduced tissue biomass.
+The role of light in driving seasonal changes in symbiont density was demonstrated nicely by @Stimson:1997p3837, who also found that experimental nutrient-enrichment amplified the light-driven seasonal oscillation.
+Using the levels of light and nutrients from this study as inputs, the model reproduces this observed interaction among environmental factors (Fig.
+4), and also provides the mechanism: under high light in summer, photoinhibition reduces symbiont biomass and leads to carbon-limitation of host growth regardless of nutrient status; when light is reduced in winter, these responses become constrained instead by nitrogen, and therefore increase more when DIN is enriched (Fig.
+S8).
+
+![Figure 4.
+Light-driven seasonal dynamics of symbiont abundance and coral growth.
+Light input (upper panel) was designed as a sinusoidal curve with a period of one year, with maximum and minimum values of 40 and 22 mol photons m^-2^ d^-1^, based on @Stimson:1997p3837.
+The dynamic behavior of symiont to host biomass ratio (middle panel) and the specific growth rate of host biomass (lower panel) show seasonal oscillations that are greater in magnitude under higher nutrients (1e-5 molN L^-1^; dashed lines) relative to lower nutrients (1e-7 molN L^-1^; solid lines), consistent with the findings of @Stimson:1997p3837.
+Prey density was set at 1e-5 CmolX L^-1^.](../img/Fig4.png)
+
+The prediction of higher growth when light is reduced indicates that growth is not limited by low light in winter, but is actually reduced by excess light in summer, consistent with the experimental findings of @Schutter:2011ie.
+Seasonal summertime reductions in tissue biomass have also been well-documented in the field [@Fitt:2000p3792], along with reductions in net photosynthetic capacity [@MullerParker:1987cr].
+Importantly, while light alone (at these levels) may drive seasonal dynamics in the ways discussed, temperature fluctuations may attenuate or even reverse the effect of light as cooler winters depress metabolism; thus, the relative magnitude of fluctuation in temperature, light, and other factors may produce wide variability in the direction and magnitude of seasonal changes in growth and symbiont abundance, depending on location and microhabitat.
+Nevertheless, the seasonal variability predicted here (Fig.
+4) is consistent with experimental and field observations for corals, and demonstrates the power of this model to predict dynamic behavior that mechanistically integrate the impacts of multiple environmental drivers.
+
+## Coral bleaching
+Coral bleaching is the stress-induced loss of symbiotic algae from coral tissues, which can occur in response to a variety of environmental stressors.
+In most cases, coral bleaching is thought to begin with photo-oxidative stress in symbiont photosynthesis [@Lesser:1997p3587], which triggers a cascade of events leading to symbiont expulsion [@Weis:2008p944].
+As symbionts are expelled, the host receives less fixed carbon, which may then compromise its ability to deliver CO~2~ to photosynthesis [@Wooldridge:2013tj].
+Increasing CO~2~-limitation for remaining symbionts, along with an amplified internal light environment due to reduced self-shading [@Enriquez:2005p142], may further exacerbate photo-oxidative stress and accelerate symbiont expulsion, driving the coral into a bleached state.
+
+While these positive feedbacks have been discussed previously in the literature, this is the first attempt to implement and explore their properties within a dynamical model.
+Interestingly, these feedbacks lead to alternate stable states.
+The 'healthy' state is characterized by nitrogen-limitation of both symbiont and host: under these conditions, the symbiont translocates sufficient carbon to support host growth and CCMs, which ensures that photosynthesis does not become CO~2~-limited.
+However, if carbon translocation is disrupted (and light is sufficiently high), then the system is driven into the 'bleached' state by photo-oxidative stress and positively reinforcing carbon- and CO~2~-limitation.
+In this context, coral bleaching can be understood as a transition from one stable state to another, and bleaching thresholds are sets of environmental conditions that push a healthy-state coral onto a trajectory leading to a bleached stable state.
+
+We are highly interested in the conditions under which the system switches from a healthy to a bleached state, and can use this model as a tool to explore this dynamic behavior.
+Most straightforwardly, this switch occurs when irradiance exceeds the capacity of photochemistry and NPQ, causing the production of ROS that leads to photoinhibition, photodamage, and symbiont expulsion, thereby engaging the positive feedbacks that lead to a bleached state (Fig.
+5).
+However, the positive feedbacks involved in bleaching are not engaged only in response to high external irradiance; in fact, they depend on the relative balance of light energy absorption and quenching, which in turn depends on the availability of CO~2~ for photosynthesis.
+While previous models framed photo-oxidative stress as a fixed response to absolute external irradiance [@Eynaud:2011tv], our implementation considers the dynamic balance of multiple energy sinks in the causation of stress, which is more consistent with current understanding of symbiosis dysfunction [@Wooldridge:2013tj], and establishes a critical role of host CCMs in providing CO~2~ for photosynthesis [@Tansik:2015cm; @Hopkinson:2015ev].
+
+![Figure 5. Bleaching.](../img/Fig5.png)
+
+The importance of host CCM activity establishes significant interactive roles for other factors in influencing coral bleaching responses.
+For example, simulations of high light stress (Fig.
+6) demonstrate that bleaching can be attenuated by heterotrophic feeding, a phenomenon which has been observed experimentally [@Borell:2008p108].
+The mechanism underlying this prediction is that feeding by the host increases host CCM activity, which delays the onset of CO~2~-limitation of photosynthesis and reduces bleaching severity.
+On the other hand, elevated nutrients exacerbate bleaching (Fig.
+6), since higher symbiont densities are more susceptible to CO~2~-limitation [@Wooldridge:2009p7807].
+Several experimental [@Cunning:2013gp; @VegaThurber:2014hk] and correlational studies [@Wooldridge:2009kb] are consistent with this mechanistic link between high nutrients and bleaching.
+
+![Figure 6. Bleaching and interactive factors.](../img/Fig6.png)
+
+Since bleaching represents an alternative stable state, it follows that a bleached coral cannot recover unless the external condition that induced bleaching is alleviated.
+In natural settings, this typically occurs by seasonal declines in temperature and light.
+However, due to hysteresis, a coral cannot recover along the same trajectory it followed during bleaching; indeed, the stressor must be alleviated well *below* the threshold that initially caused bleaching in order for the system to recover (Fig.
+7).
+This is because under the same external conditions, a bleached coral (relative to a healthy one) is characterized by greater light amplification and weaker CCM activity, which serve to maintain the bleached state.
+The magnitude of hysteresis will depend on external nutrients and prey availability, as well as model parameters, yet the fundamental concept demonstrates a core behavior of the system.
+
+![Figure 7. Hysteresis.](../img/Fig7.png)
+
+The hysteresis demonstrated in Fig.
+7 reveals another interesting behavior of the system: during recovery from bleaching, the S:H biomass ratio reaches a higher value that 'overshoots' the level maintained in the healthy stable state.
+In fact, such a phenomenon has been observed in both experimental [@Cunning:2015ja] and field studies [@Kemp:2014ug] documenting recovery from bleaching, and has been interpreted as a potential 'disequilibrium in host-symbiont regulation' [@Kemp:2014ug].
+The model explains this 'overshoot' as an approach to a bifurcation point in the dynamical system: during recovery from the 'bleached' state, host growth is more strongly carbon-limited than symbiont growth (since the symbiont has priority access to carbon), so S:H biomass increases.
+S:H biomass will continue to increase beyond the alternative steady state levels as long as the symbiont remains carbon-limited; not until nitrogen becomes more strongly limiting to symbiont growth (hence forcing the symbiont to share fixed carbon with the host) will the system 'flip' back to a state of high carbon fixation and overall nitrogen-limitation that characterize a 'healthy' symbiosis.
+Of course, the magnitude of this 'overshoot' will depend on external nutrient concentrations and the rate of change in the stressor during recovery, though simulations predict similar magnitudes to what was observed by Kemp et al.
+(supp fig?).
+
+# Conclusions
+The dynamic behavior of the model demonstrates its ability to mechanistically integrate multiple forcing functions, and the degree to which the system state depends not only on external forcing, but also the past history of the system.
+The model represents a unifying mechanistic framework that reproduces many phenomena observed in corals, suggesting it has captured many of the important features of the system in a relatively simple formulation.
+Furthermore, it allows the model to be used to predict responses of the system to multiple dynamic environmental forcing functions in a way that incorporates all positive and negative feedbacks.
+This makes the model potentially highly useful as a tool for generating hypotheses.
+make N, L, and X go up and down.
+For example, the model demonstrates and provides a mechanistic explanation for: an increase in symbiont abundance with nutrients (REF), an increase in symbiont abundance with heterotrophic feeding (refs), decrease in symbiont abundance with light (ref), optimal growth at intermediate light, rapid 'bleaching' above threshold of high light (ref) along with positive feedbacks (wooldridge, enriquez), mitigation of bleaching by heterotrophic feeding, exacerbation of bleaching by elevated nutrients, higher standing stocks of less productive symbionts, greater seasonal change in symbiont density with elevated nutrients, overshoot of symbiont density after recovery from bleaching.
+
+"host control"=>N-limitation.  C-limitation=host has lost control of symbiosis: either bleaches or symbiont "overgrows" host with net negative growth. (when does this occur?)
+
+In this context, nitrogen-limitation represents a scenario in which the host is 'in control' of the system: negative feedbacks stabilize the symbiont to host biomass ratio and the system maintains positive growth.
+Carbon-limitation represents a loss of 'control', wherein positive feedbacks result in a loss of capacity for carbon fixation and positive growth.
+Importantly, these 'control' mechanisms arise entirely passively through the feedbacks present in the system.
+
+(Very low light also causes the symbiont to not share any carbon, resulting in very high S:H ratios and negative growth of the coral, which we interpret as a dysfunctional (and impossible) condition.)
 

### 2016-10-25 JTB_ms/JTB_ms.tex
finish drafting main text

@@
   language (R Core Team 2014). With an accessible and modular framework,
-  we envision this as a resource for futher development by the
+  we envision this model as a resource for futher development by the
   scientific community to include additional complexity and
@@
 exponentially during growth (indeed, corals grow exponentially (Bak
-1976)). The specific growth rate and the ratio of symbiont to host
-biomass are the responses of interest of the system. Below we describe
-the formulation of each flux involved in producing these responses.
+1976)).
+
+Environmental stress is implemented in the form of photo-oxidative
+stress, which is thought to be the primary trigger of coral bleaching
+(Lesser 1997; Weis 2008; Wooldridge 2009). To simulate these events, we
+model the absorption and quenching of light energy by photochemistry and
+non-photochemical quenching, and the phenomenological responses that
+occur when these capacities are overwhelmed (i.e., photoinhibition,
+photodamage, and symbiont loss). While bleaching in response to high
+light alone has been observed experimentally (Schutter et al. 2011;
+Downs et al. 2013), mass coral bleaching events occur in response to
+high temperature (Hoegh-Guldberg 1999); thus, it is important to justify
+our consideration of light as the primary stressor in the model. In
+reality, light and temperature interact synergistically (Coles and
+Jokiel 1978; Jones et al. 1998), and in fact, any stressor that disrupts
+the quenching of light energy may lead to bleaching (Wooldridge 2010;
+Baker and Cunning 2015). This is because the proximate cause of
+photo-oxidative stress is excess excitation energy, but the upstream
+events that lead to this situation may be diverse: indeed, elevated
+temperature may inhibit Rubisco functioning (Jones et al. 1998) and the
+repair of the D1 protein in photosystem II (Warner, Fitt, and Schmidt
+1999), which reduces the capacity of photochemical quenching and leads
+to an excess of light energy. In this way, elevated temperature serves
+to reduce the threshold above which light stresses the system
+(Hoegh-Guldberg 1999); importantly, light is still the proximate
+stressor. Therefore, we omitted temperature from the model to maintain a
+desired level of simplicity, while still allowing photophysiological
+stress and bleaching to be simulated with biological realism in response
+to light.
+
+Below we describe the formulation of each flux in the model.
 
 Table 1. Model fluxes
-
-\begin{longtable}[c]{@{}llll@{}}
-\toprule
-\begin{minipage}[b]{0.12\columnwidth}\raggedright\strut
-Symbol
-\strut\end{minipage} &
-\begin{minipage}[b]{0.48\columnwidth}\raggedright\strut
-Description
-\strut\end{minipage} &
-\begin{minipage}[b]{0.26\columnwidth}\raggedright\strut
-Units
-\strut\end{minipage} &
-\begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
-Eq. no.
-\strut\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(j_X\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Prey uptake rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-molX CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-3
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(j_N\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Nitrogen uptake rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-molN CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-4
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(j_{HG}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Host biomass formation rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-CmolH CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-5
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(r_{NH}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Recycled nitrogen from host turnover
-\strut\end{minipage} &
-\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-molN CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-6
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(\rho_N\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Nitrogen shared with the symbiont
-\strut\end{minipage} &
-\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-molN CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-7
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(j_{eC}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Excess carbon used to activate host CCMs
-\strut\end{minipage} &
-\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-molC CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(j_L\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Light absorption rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-molph CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-9
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(j_{CO_2}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-CO\textsubscript{2} input to photosynthesis
-\strut\end{minipage} &
-\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-molCO\textsubscript{2} CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-10
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(r_{CH}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Recycled CO\textsubscript{2} from host turnover
-\strut\end{minipage} &
-\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-molCO\textsubscript{2} CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-11
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(r_{CS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Recycled CO\textsubscript{2} from symbiont turnover
-\strut\end{minipage} &
-\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-molCO\textsubscript{2} CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-12
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(j_{CP}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Photosynthesis rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-molC CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-13
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(j_{eL}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Light energy in excess of photochemistry
-\strut\end{minipage} &
-\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-molph CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-14
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(j_{NPQ}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Total capacity of NPQ
-\strut\end{minipage} &
-\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-molph CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-15
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(c_{ROS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-ROS production proportional to baseline
-\strut\end{minipage} &
-\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-16
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(r_{NS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Recycled nitrogen from symbiont turnover
-\strut\end{minipage} &
-\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-molN CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-17
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(j_{SG}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Symbiont biomass formation rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-18
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(\rho_C\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Fixed carbon shared with host
-\strut\end{minipage} &
-\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-molC CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-19
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(j_{ST}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Symbiont turnover rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-20
-\strut\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
-
-Table 2. Model parameters
 
@@
 \strut\end{minipage} &
-\begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
-Value
-\strut\end{minipage} &
-\begin{minipage}[b]{0.23\columnwidth}\raggedright\strut
+\begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
 Units
+\strut\end{minipage} &
+\begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
+Eq. no.
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_X\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Prey uptake rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molX CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+3
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_N\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Nitrogen uptake rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molN CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+4
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{HG}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Host biomass formation rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+CmolH CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+5
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(r_{NH}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Recycled nitrogen from host turnover
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molN CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+6
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(\rho_N\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Nitrogen shared with the symbiont
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molN CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+7
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{eC}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Excess carbon used to activate host CCMs
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molC CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_L\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Light absorption rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molph CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+9
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{CO_2}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+CO\textsubscript{2} input to photosynthesis
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molCO\textsubscript{2} CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+10
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(r_{CH}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Recycled CO\textsubscript{2} from host turnover
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molCO\textsubscript{2} CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+11
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(r_{CS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Recycled CO\textsubscript{2} from symbiont turnover
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molCO\textsubscript{2} CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+12
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{CP}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Photosynthesis rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molC CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+13
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{eL}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Light energy in excess of photochemistry
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molph CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+14
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{NPQ}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Total capacity of NPQ
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molph CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+15
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(c_{ROS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+ROS production proportional to baseline
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+--
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+16
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(r_{NS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Recycled nitrogen from symbiont turnover
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molN CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+17
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{SG}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Symbiont biomass formation rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+18
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(\rho_C\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Fixed carbon shared with host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molC CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+19
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{ST}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Symbiont turnover rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+20
+\strut\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
+
+Table 2. Model parameters
+
+\begin{longtable}[c]{@{}llll@{}}
+\toprule
+\begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
+Symbol
+\strut\end{minipage} &
+\begin{minipage}[b]{0.48\columnwidth}\raggedright\strut
+Description
+\strut\end{minipage} &
+\begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
+Units
+\strut\end{minipage} &
+\begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
+Value
+\strut\end{minipage}\tabularnewline
+\midrule
+\endhead
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 \(n_{NH}\)
@@
 \strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+--
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 0.19
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(n_{NS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+N:C molar ratio in symbiont biomass
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 --
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(n_{NS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-N:C molar ratio in symbiont biomass
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 0.2
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(n_{NX}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+N:C molar ratio in prey biomass
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 --
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(n_{NX}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-N:C molar ratio in prey biomass
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 0.13
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{HT}^0\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Specific turnover rate of host biomass
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+CmolH CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+0.03
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{ST}^0\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Specific turnover rate of symbiont biomass
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+0.03
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(\sigma_{NH}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Proportion host nitrogen turnover recycled
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 --
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{HT}^0\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Specific turnover rate of host biomass
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-0.03
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+0.9
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(\sigma_{CH}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Proportion host carbon turnover recycled
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+--
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+0.9
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(\sigma_{NS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Proportion symbiont nitrogen turnover recycled
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+--
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+0.9
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(\sigma_{CS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Proportion symbiont carbon turnover recycled
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+--
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+0.9
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{Xm}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Maximum specific feeding rate of host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molX CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+0.1292
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(K_X\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Half-saturation constant for prey uptake by host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molX L\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+20e-6
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{Nm}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Maximum specific DIN uptake rate by host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molN CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+0.048
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(K_N\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Half-saturation constant for DIN uptake by host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molN L\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+0.46e-6
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(k_{CO_2}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Efficacy of CO\textsubscript{2} delivery by host CCMs
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molCO\textsubscript{2} molC\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+10
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{HGm}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Maximum specific growth rate of host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 CmolH CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{ST}^0\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Specific turnover rate of symbiont biomass
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-0.03
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(\sigma_{NH}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Proportion host nitrogen turnover recycled
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-0.9
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
---
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(\sigma_{CH}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Proportion host carbon turnover recycled
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-0.9
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
---
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(\sigma_{NS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Proportion symbiont nitrogen turnover recycled
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-0.9
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
---
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(\sigma_{CS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Proportion symbiont carbon turnover recycled
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-0.9
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
---
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{Xm}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Maximum specific feeding rate of host
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-0.1292
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-molX CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(K_X\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Half-saturation constant for prey uptake by host
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-20e-6
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-molX L\textsuperscript{-1}
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{Nm}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Maximum specific DIN uptake rate by host
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-0.048
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-molN CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(K_N\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Half-saturation constant for DIN uptake by host
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-0.46e-6
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-molN L\textsuperscript{-1}
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(k_{CO_2}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Efficacy of CO\textsubscript{2} delivery by host CCMs
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-10
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-molCO\textsubscript{2} molC\textsuperscript{-1}
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{HGm}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Maximum specific growth rate of host
-\strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 1
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-CmolH CmolH\textsuperscript{-1} d\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molC mol ph\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 0.1
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-molC mol ph\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+m\textsuperscript{2} CmolS\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 1.34
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-m\textsuperscript{2} CmolS\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+mol ph CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 112
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-mol ph CmolS\textsuperscript{-1} d\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+mol ph CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 40-80
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-mol ph CmolS\textsuperscript{-1} d\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molC CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 2.8
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-molC CmolS\textsuperscript{-1} d\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 0.25
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
 \strut\end{minipage}\tabularnewline
@@
 \strut\end{minipage} &
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+--
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 5
-\strut\end{minipage} &
-\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
---
 \strut\end{minipage}\tabularnewline
@@
 symbiosome to increase localized CO\textsubscript{2} concentrations
-around the symbiont (Barott et al. 2014). Since these mechanisms require
-energetic input by the host, the magnitude of CO\textsubscript{2}
-delivery is proportional to the host's excess carbon, \(j_{eC}\),
-assuming that some of this carbon is respired to energize these CCMs.
-This formulation means that the symbiont indirectly ensures its own
-CO\textsubscript{2} supply by providing fixed carbon to the host
-(Wooldridge 2013), which establishes a positive feedback in the model
-that leads to the existence of multiple stable states, which we discuss
-below in the context of coral bleaching. The parameter \(k_{CO_2}\)
-scales the efficacy of host CCMs, which enables the comparison of
-different rates of CO\textsubscript{2} delivery that may characterize
-different coral species (Wooldridge 2014a). The input of
+around the symbiont (Barott et al. 2014). Since these active CCMs
+require energetic input by the host, the magnitude of
+CO\textsubscript{2} delivery is proportional to the amount of fixed
+carbon available to the host after satisfying its growth demands,
+\(j_{eC}\), assuming that some of this carbon is respired to energize
+these CCMs. This formulation means that the symbiont indirectly ensures
+its own CO\textsubscript{2} supply by providing fixed carbon (=energy)
+to the host (Wooldridge 2013), which establishes a positive feedback in
+the model that leads to the existence of multiple stable states, which
+we discuss below in the context of coral bleaching. The parameter
+\(k_{CO_2}\) scales the efficacy of host CCMs, which enables the
+comparison of different rates of CO\textsubscript{2} delivery that may
+characterize different coral species (Wooldridge 2014a). The input of
 CO\textsubscript{2} to the photosynthesis SU is therefore specified as:
@@
 
-\section{Results}\label{results}
-
-\subsection{Qualitative dynamics}\label{qualitative-dynamics}
+\section{Steady state behavior}\label{steady-state-behavior}
 
 In a constant environment, the system ultimately reaches a steady state
-of exponential growth or decline. Under some conditions, alternate
-steady states can be reached depending on initial values of symbiont and
-host biomass. The mechanism that produces these alternate steady states
-is the positive feedback between carbon-limitation of the host and
-CO\textsubscript{2}-limitation of photosynthesis. If the host becomes
-carbon-limited, the efficacy of its CCMs are reduced and photosynthesis
-becomes progressively more CO\textsubscript{2}-limited. In turn, less
-photosynthate is translocated to the host, causing further
-carbon-limitation and an ultimate decline in biomass; this steady state
-represents a dysfunctional symbiosis. However, if the host is not
-carbon-limited, it effectively delivers CO\textsubscript{2} to
-photosynthesis and ensures its continued receipt of fixed carbon from
-the symbiont; this system is nitrogen-limited and represents a
-functional symbiosis with positive exponential growth. Either of these
-stable states may be reached under (some sets of) constant environmental
-conditions, depending on the initial values of symbiont and host
-biomass: if symbiont biomass is initially very low (i.e., a ``bleached''
-coral), the system cannot escape the carbon-limited state and cannot
-grow. However, if symbiont biomass is initially high (i.e., a healthy
+of exponential growth or decline. However, under some conditions, either
+of these outcomes may occur depending on initial values of symbiont and
+host biomass, indicating the presence of alternate stable states. The
+mechanism that produces these alternate stable states is the positive
+feedback between carbon-limitation of the host and
+CO\textsubscript{2}-limitation of photosynthesis: if symbiont biomass is
+initially very low (i.e., a ``bleached'' coral), the system cannot
+escape positively reinforcing carbon-limitation and cannot grow.
+However, if symbiont biomass is initially high (i.e., a ``healthy''
 coral), then the system maintains positive growth in a nitrogen-limited
-state. For practical purposes, we consider only positive growth steady
-states in the ``steady state behavior'' section below. In the subsequent
-``dynamic behavior'' section, we explore more thoroughly the conditions
-that cause the system to switch between alternate stable states, and
-interpret these dynamics in the context of coral bleaching.
-
-\subsection{Steady state behavior}\label{steady-state-behavior}
+state. For practical purposes, this section of the manuscript considers
+only positive growth steady states under constant environments; the
+subsequent ``dynamic behavior'' section explores how changing
+environments may cause the system to switch between alternate stable
+states, which we interpret in the context of coral bleaching.
+
+To analyze qualitative patterns in steady state behavior, we ran the
+model to steady state across gradients of external irradiance and
+nutrients (Fig. 2), which revealed patterns consistent with observed
+phenomena in corals. Predicted growth rates are low at low light and DIN
+(\textasciitilde{}0.01 d\textsuperscript{-1}), and begin increasing as
+both of these factors increase (Fig. 2A). Low light limits
+photosynthetic rates, resulting in less fixed carbon shared with the
+host and an associated increase in the symbiont to host biomass ratio
+(Fig. 2B). In agreement with this trend are many observations of
+negative correlation between irradiance and symbiont density (Stimson
+1997; Brown et al. 1999; Fitt et al. 2000; Titlyanov et al. 2001). As
+higher light alleviates light-limitation of photosynthesis, growth
+becomes less carbon-limited.
 
@@
 \end{figure}
-
-To analyze qualitative behavior, we ran the model to steady state across
-gradients of external irradiance and nutrients (Fig. 2), which revealed
-patterns consistent with observed phenomena in corals. Predicted growth
-rates are low at low light and DIN (\textasciitilde{}0.01
-d\textsuperscript{-1}), and begin increasing as both of these factors
-increase (Fig. 2A). Low light limits photosynthetic rates, resulting in
-less fixed carbon shared with the host and an associated increase in the
-symbiont to host biomass ratio (Fig. 2B). In agreement with this trend
-are many observations of negative correlation between irradiance and
-symbiont density (Stimson 1997; Brown et al. 1999; Fitt et al. 2000;
-Titlyanov et al. 2001). As higher light alleviates light-limitation of
-photosynthesis, growth becomes less carbon-limited.
 
@@
 below 0.01 d\textsuperscript{-1} (Osinga et al. 2011; Osinga et al.
-2012) (though values as high as 0.025 d\textsuperscript{-1} have been
-reported (Schutter et al. 2010)). That predicted growth rates are
-slightly higher than most observations is not surprising, as the model
+2012), though values as high as 0.025 d\textsuperscript{-1} have been
+reported (Schutter et al. 2010). It is not surprising that predicted
+growth rates are slightly higher than most observations, as the model
 does not account for external ecological factors that may reduce growth,
@@
 does not explicitly account for skeletal formation, which may not always
-be well correlated with biomass growth, potentially due to differential
-substrate-limitation (Anthony 2002). Nevertheless, when rates of
-increase in coral tissue biomass have been directly measured over short
-periods, values are consistent with model predictions (e.g., 0.07
-d\textsuperscript{-1} in Tanaka et al. (2007)), and similar rates (0.04
-d\textsuperscript{-1}) have been measured in \emph{Aiptasia diaphana}, a
-non-calcifying symbiotic anemone (Armoza-Zvuloni et al. 2014).
+correlate with biomass growth, potentially due to differential
+substrate-limitation (Anthony 2002). Nevertheless, when growth of coral
+tissue biomass has been directly measured over short periods, values are
+consistent with model predictions (e.g., 0.07 d\textsuperscript{-1} in
+Tanaka et al. (2007)), and similar rates (0.04 d\textsuperscript{-1})
+have been measured in \emph{Aiptasia diaphana}, a non-calcifying
+symbiotic anemone (Armoza-Zvuloni et al. 2014).
 
@@
 
-While bleaching in response to high light alone has been observed
-experimentally (Schutter et al. 2011; Downs et al. 2013), mass coral
-bleaching events occur in response to high temperature (Hoegh-Guldberg
-1999); thus, it is important to justify our consideration of light as
-the primary stressor in the model. In reality, light and temperature
-interact synergistically (Coles and Jokiel 1978; Jones et al. 1998), and
-in fact, any stressor that disrupts the quenching of light energy may
-lead to bleaching (Wooldridge 2010; Baker and Cunning 2015). This is
-because the proximate cause of coral bleaching is excess excitation
-energy, but the upstream events that lead to this situation may be
-diverse: indeed, elevated temperature may inhibit Rubisco functioning
-(Jones et al. 1998) and the repair of the D1 protein in photosystem II
-(Warner, Fitt, and Schmidt 1999), which reduces the capacity of
-photochemical quenching and leads to an excess of light energy. In this
-way, elevated temperature serves to reduce the threshold above which
-light causes stress to the system (Hoegh-Guldberg 1999); importantly,
-light is still the proximate stressor. Therefore, we omitted temperature
-from the model to maintain a desired level of simplicity, while still
-allowing photophysiological stress and bleaching to be simulated with
-biological realism in response to light. Thus, while the model is
-currently accurate from a conceptual and mechanistic standpoint, its
-value as a predictive tool under scenarios of changing temperature will
-require further development.
-
 The incorporation of light stress in the model sets an upper limit to
@@
 
-\subsection{Sensitivity analysis}\label{sensitivity-analysis}
-
-Each default parameter value used in the model (Table 2) is based on
-carefully selected relevant literature. The derivations of each of these
-default parameter values are described in the Supplementary Information.
-Nevertheless, it is important to evaluate the sensitivity of the model
-to changes in these parameter values. Moreover, these sensitivity plots
-serve to demonstrate the behavior of the dynamical system. We measured
-fractional change in steady state values in response to fractional
-changes in parameter values, relative to their default values, under
-environmental conditions typical of coral reefs.
+\section{Sensitivity analysis}\label{sensitivity-analysis}
+
+The values used for each parameter in the model (Table 2) are derived
+from relevant literature; these derivations are described in the
+Supplementary Information. Here we evaluate the sensitivity of the model
+to changes in these parameter values, which also serves to demonstrate
+the behavior of the dynamical system. We measured fractional change in
+steady state values in response to fractional changes in parameter
+values, relative to their default values, under environmental conditions
+typical of coral reefs.
 
@@
 
-Overall, relative changes in the steady state of the system were less
+Overall, relative changes in the steady state of the system are less
 than the equivalent relative change in parameter value, indicating that
 the system maintains internal stability. However, changes in certain
-parameter values have greater impacts: increasing \(j_{Nm}\) or
-decreasing \(K_N\) both dramatically increase host growth, demonstrating
-the strong nitrogen-limitation that characterizes these symbioses (REF).
-The parameterr \(\bar{a}^*\) has a strong impact on S:H biomass ratios,
-because this parameter determines the amount of light absorbed by
-symbionts, thus low values of this parameter increase light-limitation
-of the system. Altering the maximum growth and turnover rates of
-symbiont and host biomass also have expected impacts on steady states.
-Parameters having to do with photooxidative stress and bleaching have
-little impact under these environmental conditions, but of course have
-larger impacts under higher light (e.g., Fig. S6).
-
-\subsection{Dynamic behavior}\label{dynamic-behavior}
-
-Therefore, we are highly interested in the conditions under which the
-system switches from one state to another, representing the conditions
-under which corals ``bleach'' due to progressive, reinforcing
-carbon-limitation. Interestingly, this state can be reached in multiple
-ways, by the acceleration of photoinhibition and photodamage, or by high
-external DIN, which causes the symbiont to share less carbon,
-subsequently causing its own CO2-limitation, and progressive
-C-limitation of the system. \#\# varying light: consequences of excess
-light and light-limitation \#\# varying DIN: role of N-limitation and
-consequences of elevated N \#\# elevated nutrients cause system to
-become C-limited (photo: CO2 limited; host: fixed-C limited) \#\#
-C-limitation increases susceptibility to bleaching Importantly, these
-negative impacts of high irradiance only arise when light energy exceeds
-the capacity of photochemistry and NPQ, which in turn depends on
-CO\textsubscript{2} availability. While previous models framed
-photophysiological stress as a fixed response to absolute irradiance
-levels (Eynaud, Nisbet, and Muller 2011), the current implementation
-considers the dynamic balance of multiple energy sinks in the causation
-of stress, which is more consistent with current understanding of
-symbiosis dysfunction (Wooldridge 2013), and establishes an important
-role of the host in providing CO\textsubscript{2} for symbiont
-photosynthesis (Tansik, Fitt, and Hopkinson 2015; Hopkinson, Tansik, and
-Fitt 2015). \#\# consequences of a less efficient but more tolerant
-symbiont \#\# Rate of symbiont repopulation (Berner et al. 1993) \#\#
-Bicarbonate addition promotes coral growth (Marubini and Thake 1999)
-\#\# Seasonal change in sym. density is larger when nutrients are
-elevated (Stimson 1997). Such a phenomenon was predicted by Fitt et al.
-2000. \#\# Death rule: negative growth for a certain amount of time
-should be defensible.
+parameter values have more significant impacts: increasing \(j_{Nm}\) or
+decreasing \(K_N\) both dramatically increase host growth (Fig. 3),
+demonstrating the strong nitrogen-limitation that characterizes these
+symbioses (REF). The parameter \(\bar{a}^*\) has a strong impact on S:H
+biomass ratios (Fig. 3) since this parameter determines the amount of
+light absorbed by symbionts, with lower values increasing
+light-limitation. Altering maximum growth and turnover rates of symbiont
+and host also have expected impacts on steady states (Fig. 3).
+Parameters relating to photooxidative stress and bleaching have little
+impact under these environmental conditions (Fig. 3), but of course have
+larger impacts under higher light (e.g., Fig. S6). Sensitivity analyses
+conducted under different combinations of external light and nutrients
+are presented in the Supplementary Information (Figs. S3-S7)
+
+\section{Dynamic behavior}\label{dynamic-behavior}
+
+The dynamic behavior of the model demonstrates its power to integrate
+multiple environmental forcings simultaneously to produced integrated
+responses of the system. The potential applications of this are
+wide-ranging, and so the model may be highly useful for generating
+hypotheses about the way the system will respond to complex sets of
+environmental dynamics. Here we present several scenarios to highlight
+the power of the model and demonstrate that it can also reproduce more
+complex phenomena that have been observed in corals.
+
+\subsection{Seasonal variability}\label{seasonal-variability}
+
+Symbiont densities and coral growth rates are known to vary seasonally,
+representing an integrated response to changes in a suite of
+environmental factors. Light in particular is a strong driver of these
+trends (Stimson:1997p3837; Brown et al. 1999; Fagoonee et al. 1999; Fitt
+et al. 2000), with high light associated with lower symbiont abundance
+and reduced tissue biomass. The role of light in driving seasonal
+changes in symbiont density was demonstrated nicely by Stimson (1997),
+who also found that experimental nutrient-enrichment amplified the
+light-driven seasonal oscillation. Using the levels of light and
+nutrients from this study as inputs, the model reproduces this observed
+interaction among environmental factors (Fig. 4), and also provides the
+mechanism: under high light in summer, photoinhibition reduces symbiont
+biomass and leads to carbon-limitation of host growth regardless of
+nutrient status; when light is reduced in winter, these responses become
+constrained instead by nitrogen, and therefore increase more when DIN is
+enriched (Fig. S8).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{../img/Fig4.png}
+\caption{Figure 4. Light-driven seasonal dynamics of symbiont abundance
+and coral growth. Light input (upper panel) was designed as a sinusoidal
+curve with a period of one year, with maximum and minimum values of 40
+and 22 mol photons m\textsuperscript{-2} d\textsuperscript{-1}, based on
+Stimson (1997). The dynamic behavior of symiont to host biomass ratio
+(middle panel) and the specific growth rate of host biomass (lower
+panel) show seasonal oscillations that are greater in magnitude under
+higher nutrients (1e-5 molN L\textsuperscript{-1}; dashed lines)
+relative to lower nutrients (1e-7 molN L\textsuperscript{-1}; solid
+lines), consistent with the findings of Stimson (1997). Prey density was
+set at 1e-5 CmolX L\textsuperscript{-1}.}
+\end{figure}
+
+The prediction of higher growth when light is reduced indicates that
+growth is not limited by low light in winter, but is actually reduced by
+excess light in summer, consistent with the experimental findings of
+Schutter et al. (2011). Seasonal summertime reductions in tissue biomass
+have also been well-documented in the field (Fitt et al. 2000), along
+with reductions in net photosynthetic capacity (\textbf{???}).
+Importantly, while light alone (at these levels) may drive seasonal
+dynamics in the ways discussed, temperature fluctuations may attenuate
+or even reverse the effect of light as cooler winters depress
+metabolism; thus, the relative magnitude of fluctuation in temperature,
+light, and other factors may produce wide variability in the direction
+and magnitude of seasonal changes in growth and symbiont abundance,
+depending on location and microhabitat. Nevertheless, the seasonal
+variability predicted here (Fig. 4) is consistent with experimental and
+field observations for corals, and demonstrates the power of this model
+to predict dynamic behavior that mechanistically integrate the impacts
+of multiple environmental drivers.
+
+\subsection{Coral bleaching}\label{coral-bleaching}
+
+Coral bleaching is the stress-induced loss of symbiotic algae from coral
+tissues, which can occur in response to a variety of environmental
+stressors. In most cases, coral bleaching is thought to begin with
+photo-oxidative stress in symbiont photosynthesis (Lesser 1997), which
+triggers a cascade of events leading to symbiont expulsion (Weis 2008).
+As symbionts are expelled, the host receives less fixed carbon, which
+may then compromise its ability to deliver CO\textsubscript{2} to
+photosynthesis (Wooldridge 2013). Increasing
+CO\textsubscript{2}-limitation for remaining symbionts, along with an
+amplified internal light environment due to reduced self-shading
+(Enríquez, Méndez, and Iglesias-Prieto 2005), may further exacerbate
+photo-oxidative stress and accelerate symbiont expulsion, driving the
+coral into a bleached state.
+
+While these positive feedbacks have been discussed previously in the
+literature, this is the first attempt to implement and explore their
+properties within a dynamical model. Interestingly, these feedbacks lead
+to alternate stable states. The `healthy' state is characterized by
+nitrogen-limitation of both symbiont and host: under these conditions,
+the symbiont translocates sufficient carbon to support host growth and
+CCMs, which ensures that photosynthesis does not become
+CO\textsubscript{2}-limited. However, if carbon translocation is
+disrupted (and light is sufficiently high), then the system is driven
+into the `bleached' state by photo-oxidative stress and positively
+reinforcing carbon- and CO\textsubscript{2}-limitation. In this context,
+coral bleaching can be understood as a transition from one stable state
+to another, and bleaching thresholds are sets of environmental
+conditions that push a healthy-state coral onto a trajectory leading to
+a bleached stable state.
+
+We are highly interested in the conditions under which the system
+switches from a healthy to a bleached state, and can use this model as a
+tool to explore this dynamic behavior. Most straightforwardly, this
+switch occurs when irradiance exceeds the capacity of photochemistry and
+NPQ, causing the production of ROS that leads to photoinhibition,
+photodamage, and symbiont expulsion, thereby engaging the positive
+feedbacks that lead to a bleached state (Fig. 5). However, the positive
+feedbacks involved in bleaching are not engaged only in response to high
+external irradiance; in fact, they depend on the relative balance of
+light energy absorption and quenching, which in turn depends on the
+availability of CO\textsubscript{2} for photosynthesis. While previous
+models framed photo-oxidative stress as a fixed response to absolute
+external irradiance (Eynaud, Nisbet, and Muller 2011), our
+implementation considers the dynamic balance of multiple energy sinks in
+the causation of stress, which is more consistent with current
+understanding of symbiosis dysfunction (Wooldridge 2013), and
+establishes a critical role of host CCMs in providing
+CO\textsubscript{2} for photosynthesis (Tansik, Fitt, and Hopkinson
+2015; Hopkinson, Tansik, and Fitt 2015).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{../img/Fig5.png}
+\caption{Figure 5. Bleaching.}
+\end{figure}
+
+The importance of host CCM activity establishes significant interactive
+roles for other factors in influencing coral bleaching responses. For
+example, simulations of high light stress (Fig. 6) demonstrate that
+bleaching can be attenuated by heterotrophic feeding, a phenomenon which
+has been observed experimentally (Borell et al. 2008). The mechanism
+underlying this prediction is that feeding by the host increases host
+CCM activity, which delays the onset of CO\textsubscript{2}-limitation
+of photosynthesis and reduces bleaching severity. On the other hand,
+elevated nutrients exacerbate bleaching (Fig. 6), since higher symbiont
+densities are more susceptible to CO\textsubscript{2}-limitation
+(Wooldridge 2009). Several experimental (Cunning and Baker 2013; Vega
+Thurber et al. 2014) and correlational studies (Wooldridge and Done
+2009) are consistent with this mechanistic link between high nutrients
+and bleaching.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{../img/Fig6.png}
+\caption{Figure 6. Bleaching and interactive factors.}
+\end{figure}
+
+Since bleaching represents an alternative stable state, it follows that
+a bleached coral cannot recover unless the external condition that
+induced bleaching is alleviated. In natural settings, this typically
+occurs by seasonal declines in temperature and light. However, due to
+hysteresis, a coral cannot recover along the same trajectory it followed
+during bleaching; indeed, the stressor must be alleviated well
+\emph{below} the threshold that initially caused bleaching in order for
+the system to recover (Fig. 7). This is because under the same external
+conditions, a bleached coral (relative to a healthy one) is
+characterized by greater light amplification and weaker CCM activity,
+which serve to maintain the bleached state. The magnitude of hysteresis
+will depend on external nutrients and prey availability, as well as
+model parameters, yet the fundamental concept demonstrates a core
+behavior of the system.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{../img/Fig7.png}
+\caption{Figure 7. Hysteresis.}
+\end{figure}
+
+The hysteresis demonstrated in Fig. 7 reveals another interesting
+behavior of the system: during recovery from bleaching, the S:H biomass
+ratio reaches a higher value that `overshoots' the level maintained in
+the healthy stable state. In fact, such a phenomenon has been observed
+in both experimental (Cunning, Silverstein, and Baker 2015) and field
+studies (Kemp et al. 2014) documenting recovery from bleaching, and has
+been interpreted as a potential `disequilibrium in host-symbiont
+regulation' (Kemp et al. 2014). The model explains this `overshoot' as
+an approach to a bifurcation point in the dynamical system: during
+recovery from the `bleached' state, host growth is more strongly
+carbon-limited than symbiont growth (since the symbiont has priority
+access to carbon), so S:H biomass increases. S:H biomass will continue
+to increase beyond the alternative steady state levels as long as the
+symbiont remains carbon-limited; not until nitrogen becomes more
+strongly limiting to symbiont growth (hence forcing the symbiont to
+share fixed carbon with the host) will the system `flip' back to a state
+of high carbon fixation and overall nitrogen-limitation that
+characterize a `healthy' symbiosis. Of course, the magnitude of this
+`overshoot' will depend on external nutrient concentrations and the rate
+of change in the stressor during recovery, though simulations predict
+similar magnitudes to what was observed by Kemp et al. (supp fig?).
+
+\section{Conclusions}\label{conclusions}
+
+The dynamic behavior of the model demonstrates its ability to
+mechanistically integrate multiple forcing functions, and the degree to
+which the system state depends not only on external forcing, but also
+the past history of the system. The model represents a unifying
+mechanistic framework that reproduces many phenomena observed in corals,
+suggesting it has captured many of the important features of the system
+in a relatively simple formulation. Furthermore, it allows the model to
+be used to predict responses of the system to multiple dynamic
+environmental forcing functions in a way that incorporates all positive
+and negative feedbacks. This makes the model potentially highly useful
+as a tool for generating hypotheses. make N, L, and X go up and down.
+For example, the model demonstrates and provides a mechanistic
+explanation for: an increase in symbiont abundance with nutrients (REF),
+an increase in symbiont abundance with heterotrophic feeding (refs),
+decrease in symbiont abundance with light (ref), optimal growth at
+intermediate light, rapid `bleaching' above threshold of high light
+(ref) along with positive feedbacks (wooldridge, enriquez), mitigation
+of bleaching by heterotrophic feeding, exacerbation of bleaching by
+elevated nutrients, higher standing stocks of less productive symbionts,
+greater seasonal change in symbiont density with elevated nutrients,
+overshoot of symbiont density after recovery from bleaching.
+
+``host control''=\textgreater{}N-limitation. C-limitation=host has lost
+control of symbiosis: either bleaches or symbiont ``overgrows'' host
+with net negative growth. (when does this occur?)
+
+In this context, nitrogen-limitation represents a scenario in which the
+host is `in control' of the system: negative feedbacks stabilize the
+symbiont to host biomass ratio and the system maintains positive growth.
+Carbon-limitation represents a loss of `control', wherein positive
+feedbacks result in a loss of capacity for carbon fixation and positive
+growth. Importantly, these `control' mechanisms arise entirely passively
+through the feedbacks present in the system.
+
+(Very low light also causes the symbiont to not share any carbon,
+resulting in very high S:H ratios and negative growth of the coral,
+which we interpret as a dysfunctional (and impossible) condition.)
 
@@
 
+\hypertarget{ref-Borell:2008p108}{}
+Borell, Esther M, Ade Yuliantri, Kai Bischof, and Claudio Richter. 2008.
+``The effect of heterotrophy on photosynthesis and tissue composition of
+two scleractinian corals under elevated temperature.'' \emph{Journal of
+Experimental Marine Biology and Ecology} 364: 116--23.
+
 \hypertarget{ref-Brown:1999p3534}{}
@@
 doi:\href{https://doi.org/10.1016/j.gloenvcha.2014.04.002}{10.1016/j.gloenvcha.2014.04.002}.
+
+\hypertarget{ref-Cunning:2013gp}{}
+Cunning, Ross, and Andrew C Baker. 2013. ``Excess algal symbionts
+increase the susceptibility of reef corals to bleaching.'' \emph{Nature
+Climate Change} 3 (March): 259--62.
+doi:\href{https://doi.org/10.1038/nclimate1711}{10.1038/nclimate1711}.
+
+\hypertarget{ref-Cunning:2015ja}{}
+Cunning, Ross, Rachel N Silverstein, and Andrew C Baker. 2015.
+``Investigating the causes and consequences of symbiont shuffling in a
+multi-partner reef coral symbiosis under environmental change.''
+\emph{Proceedings of the Royal Society B}, 1--9.
+doi:\href{https://doi.org/10.1098/rspb.2014.1725\%7B/\&\%7Ddomain=pdf\%7B/\&\%7Ddate_stamp}{10.1098/rspb.2014.1725\{\textbackslash{}\&\}domain=pdf\{\textbackslash{}\&\}date\_stamp}.
 
@@
 
+\hypertarget{ref-Fagoonee:1999p4136}{}
+Fagoonee, I, H B Wilson, M P Hassell, and J R Turner. 1999. ``The
+Dynamics of Zooxanthellae Populations: A Long-Term Study in the Field.''
+\emph{Science} 283 (5403): 843--45.
+doi:\href{https://doi.org/10.1126/science.283.5403.843}{10.1126/science.283.5403.843}.
+
 \hypertarget{ref-FerrierPages:2000p1950}{}
@@
 
+\hypertarget{ref-Kemp:2014ug}{}
+Kemp, Dustin W, X Hernandez-Pech, Roberto Iglesias-Prieto, William K
+Fitt, and Gregory W Schmidt. 2014. ``Community dynamics and physiology
+of Symbiodinium spp. before, during, and after a coral bleaching
+event.'' \emph{Limnology and Oceanography}.
+\url{http://www.aslo.org/lo/pdf/vol_59/issue_3/0788.pdf}.
+
 \hypertarget{ref-Kooijman:2010vd}{}
@@
 Organization}. 3rd ed. Cambridge University Press.
+
+\hypertarget{ref-Lesser:1997p3587}{}
+Lesser, Michael P. 1997. ``Oxidative stress causes coral bleaching
+during exposure to elevated temperatures.'' \emph{Coral Reefs} 16
+(July): 187--92.
 
@@
 
+\hypertarget{ref-VegaThurber:2014hk}{}
+Vega Thurber, Rebecca L, Deron E Burkepile, Corinne Fuchs, Andrew A
+Shantz, Ryan McMinds, and Jesse R Zaneveld. 2014. ``Chronic nutrient
+enrichment increases prevalence and severity of coral disease and
+bleaching.'' \emph{Global Change Biology} 20 (2): 544--54.
+doi:\href{https://doi.org/10.1111/gcb.12450}{10.1111/gcb.12450}.
+
 \hypertarget{ref-Wang:1998p128}{}
@@
 
+\hypertarget{ref-Wooldridge:2009kb}{}
+Wooldridge, Scott A, and Terry J Done. 2009. ``Improved water quality
+can ameliorate effects of climate change on corals.'' \emph{Ecological
+Applications} 19 (6). Ecological Society of America: 1492--9.
+doi:\href{https://doi.org/10.1890/08-0963.1}{10.1890/08-0963.1}.
+
 \hypertarget{ref-Yellowlees:2008p331}{}

### 2016-10-25 JTB_supp/JTB_supp.Rmd
add FigS8 FigS9

@@
 
+## Figure S8. Full system dynamics of seasonal light variation at two nutrient levels.
+This figure accompanies Figure 4 in the main text, and show additional model fluxes that demonstrate the mechanisms underlying the changes in the system state in these simulations.
+
+```{r FigS8, cache=T, fig.width=7, fig.height=7}
+# Set parameters
+defpars <- def_pars()  # Get default parameters
+
+# Set run time vector
+time <- seq(1, 365, 0.5)  # if single model run, use time input
+
+# Initialize environments
+env1 <- init_env(time=time, L=c(20,40,2), N=c(1e-7,1e-7,0), X=c(1e-5,1e-5,0))
+env2 <- init_env(time=time, L=c(20,40,2), N=c(1e-6,1e-6,0), X=c(1e-5,1e-5,0))
+
+# Run simulations
+ss1 <- with(run_coral_ss(env=list(L=30,N=1e-7,X=1e-5), pars=defpars), last(S$S/H$H))
+run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", ss1))
+ss2 <- with(run_coral_ss(env=list(L=30,N=1e-6,X=1e-5), pars=defpars), last(S$S/H$H))
+run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", ss2))
+
+par(mfcol=c(6,2), mar=c(2,2,1,4), mgp=c(1,0,0), tcl=0.2)
+
+with(run1, {
+      # External irradiance
+    plot(time, env$L, type="l", col="gold", ylim=c(0,60), lwd=2, xlab="", ylab="mol ph/m2/d")
+    title("External irradiance and DIN", adj=0, line=0.25)
+    par(new=T)
+    plot(time, env$N, type="l", col="blue", ylim=c(0,1e-5), lwd=2, axes=F, xlab="", ylab="")
+    axis(side=4); mtext(side=4, text="molN/L", line=1, cex=0.75)
+
+    # Symbiont to host biomass ratio
+    totSH <- S$S / H$H
+    totSHf <- totSH[length(totSH)]
+    plot(time, totSH, type="l", col="black", ylim=c(0, 0.25), ylab="CmolS/CmolH", xlab="", lwd=2)
+    text(time[0.95*length(time)], totSHf, labels=as.character(round(totSHf, 3)), pos=3, xpd=T, cex=0.75)
+    title("Symbiont:host biomass", adj=0, line=0.25)
+
+    # Specific growth rates of host and symbiont
+    Hgrf <- H$dH.Hdt[length(H$dH.Hdt)]
+    plot(time, H$dH.Hdt, type="l", ylim=c(-0.01, 0.13), 
+         xlab="", ylab="d-1", lwd=2, cex=1, cex.lab=1)
+    if(any(c(H$dH.Hdt[-(1:10)], S$dS.Sdt[-(1:10)])<0)) abline(h=0, lwd=0.5, lty=3)
+    text(time[0.95*length(time)], Hgrf, labels=as.character(round(Hgrf, 3)), pos=3, xpd=T, cex=0.75)
+    #title("State variable dynamics", adj=0, cex.main=2, outer = T)
+    title("Specific growth rates", adj=0, line=0.25)
+    lines(time, S$dS.Sdt, col="black", lwd=1)
+    legend("topright", legend=c("Host", "Sym"), lwd=c(2,1), col="black", bty="n", y.intersp=1, cex=0.75)
+
+    # Light quenching (carbon fixation, NPQ, and excess (=ROS producing))
+    plot(NA, xlim=range(time), xlab="", ylim=c(0,130), ylab="mol ph/CmolS/d")
+    title("Light quenching", adj=0, line=0.25)
+    lines(time, S$jCP/pars$yCL, col="gold", lwd=1) # amt. quenched by photosynthesis
+    lines(time, S$jCP/pars$yCL + S$jNPQ, col="gold", lty=1, lwd=2) # amt. quenched by photo + NPQ
+    lines(time, S$jL, col="gold", lty=3, lwd=2) # total amt. absorbed
+    legend("topright", legend=c("ROS", "NPQ", "Photo."), lty=c(3,1,1), lwd=c(2,2,1), col="gold", bty="n", cex=0.75)
+
+    # Photosynthesis: substrate-limitation
+    pl <- log(   pmin((H$jCO2 + H$rCH)*H$H/S$S + S$rCS, pars$jCPm)   /   pmin(S$jL * pars$yCL, pars$jCPm)    )
+    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-0.8, 0.8))
+    title("Photosynthesis: substrate-limitation", adj=0, line=0.25)
+    lines(time, pl, col="black", lty=1, lwd=2)
+    abline(h=0, lty=3)
+    text(par("usr")[2], 0, labels="L-lim.\nCO2-lim.", cex=0.75, adj=1)
+
+    # Biomass formation - substrate-limitation
+    sl <- log(  pmin(S$jCP, pars$jSGm)   /   pmin((H$rhoN*H$H/S$S + S$rNS)/pars$nNS, pars$jSGm)  )
+    hl <- log(  pmin(S$rhoC*S$S/H$H + H$jX, pars$jHGm) / pmin((H$jN + pars$nNX*H$jX + H$rNH) / pars$nNH, pars$jHGm)  )
+    maxabs <- max(c(abs(hl), abs(sl)))
+    range <- range(c(hl, sl))
+    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-1.5,1.5))
+    title("Biomass: substrate-limitation", adj=0, line=0.25)
+    lines(time, hl, col="black", lty=1, lwd=2)
+    lines(time, sl, col="black", lty=1, lwd=1)
+    abline(h=0, lty=3)
+    text(par("usr")[2], 0, labels="N-lim.\nC-lim.", cex=0.75, adj=1)
+    legend("topright", legend=c("Host", "Sym"), lty=1, lwd=c(2,1), col="black", bty="n", cex=0.75)
+})
+
+with(run2, {
+      # External irradiance
+    plot(time, env$L, type="l", col="gold", ylim=c(0,60), lwd=2, xlab="", ylab="mol/m2/d")
+    title("External irradiance and DIN", adj=0, line=0.25)
+    par(new=T)
+    plot(time, env$N, type="l", col="blue", ylim=c(0,1e-5), lwd=2, axes=F, xlab="", ylab="")
+    axis(side=4); mtext(side=4, text="molN/L", line=1, cex=0.75)
+
+    # Symbiont to host biomass ratio
+    totSH <- S$S / H$H
+    totSHf <- totSH[length(totSH)]
+    plot(time, totSH, type="l", col="black", ylim=c(0, 0.25), ylab="CmolS/CmolH", xlab="", lwd=2)
+    text(time[0.95*length(time)], totSHf, labels=as.character(round(totSHf, 3)), pos=3, xpd=T, cex=0.75)
+    title("Symbiont:host biomass", adj=0, line=0.25)
+
+    # Specific growth rates of host and symbiont
+    Hgrf <- H$dH.Hdt[length(H$dH.Hdt)]
+    plot(time, H$dH.Hdt, type="l", ylim=c(-0.01,0.13), 
+         xlab="", ylab="d-1", lwd=2, cex=1, cex.lab=1)
+    if(any(c(H$dH.Hdt[-(1:10)], S$dS.Sdt[-(1:10)])<0)) abline(h=0, lwd=0.5, lty=3)
+    text(time[0.95*length(time)], Hgrf, labels=as.character(round(Hgrf, 3)), pos=3, xpd=T, cex=0.75)
+    #title("State variable dynamics", adj=0, cex.main=2, outer = T)
+    title("Specific growth rates", adj=0, line=0.25)
+    lines(time, S$dS.Sdt, col="black", lwd=1)
+    legend("topright", legend=c("Host", "Sym"), lwd=c(2,1), col="black", bty="n", y.intersp=1, cex=0.75)
+
+    # Light quenching (carbon fixation, NPQ, and excess (=ROS producing))
+    plot(NA, xlim=range(time), xlab="", ylim=c(0,130), ylab="mol ph/CmolS/d")
+    title("Light quenching", adj=0, line=0.25)
+    lines(time, S$jCP/pars$yCL, col="gold", lwd=1) # amt. quenched by photosynthesis
+    lines(time, S$jCP/pars$yCL + S$jNPQ, col="gold", lty=1, lwd=2) # amt. quenched by photo + NPQ
+    lines(time, S$jL, col="gold", lty=3, lwd=2) # total amt. absorbed
+    legend("topright", legend=c("ROS", "NPQ", "Photo."), lty=c(3,1,1), lwd=c(2,2,1), col="gold", bty="n", cex=0.75)
+
+    # Photosynthesis: substrate-limitation
+    pl <- log(   pmin((H$jCO2 + H$rCH)*H$H/S$S + S$rCS, pars$jCPm)   /   pmin(S$jL * pars$yCL, pars$jCPm)    )
+    maxabs <- max(abs(pl))
+    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-0.8, 0.8))
+    title("Photosynthesis: substrate-limitation", adj=0, line=0.25)
+    lines(time, pl, col="black", lty=1, lwd=2)
+    abline(h=0, lty=3)
+    text(par("usr")[2], 0, labels="L-lim.\nCO2-lim.", cex=0.75, adj=1)
+
+    # Biomass formation - substrate-limitation
+    sl <- log(  pmin(S$jCP, pars$jSGm)   /   pmin((H$rhoN*H$H/S$S + S$rNS)/pars$nNS, pars$jSGm)  )
+    hl <- log(  pmin(S$rhoC*S$S/H$H + H$jX, pars$jHGm) / pmin((H$jN + pars$nNX*H$jX + H$rNH) / pars$nNH, pars$jHGm)  )
+    maxabs <- max(c(abs(hl), abs(sl)))
+    range <- range(c(hl, sl))
+    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-1.5,1.5))
+    title("Biomass: substrate-limitation", adj=0, line=0.25)
+    lines(time, hl, col="black", lty=1, lwd=2)
+    lines(time, sl, col="black", lty=1, lwd=1)
+    abline(h=0, lty=3)
+    text(par("usr")[2], 0, labels="N-lim.\nC-lim.", cex=0.75, adj=1)
+    legend("topright", legend=c("Host", "Sym"), lty=1, lwd=c(2,1), col="black", bty="n", cex=0.75)
+})
+```
+
+# Figure S9. Overshoot of symbiont density following bleaching.
+```{r FigS9, cache=T}
+# Set run time vector
+time <- seq(-100, 1095, 0.5)  # if single model run, use time input
+
+# Initialize environment
+env <- init_env(time=time, L=c(60,20,4), N=c(2e-7,2e-7,0), X=c(2e-6,2e-6,0))
+
+# Set parameters
+defpars <- def_pars()  # Get default parameters
+
+# Run simulation
+run <- run_coral(time=time, env=env, pars=replace(defpars, "kCO2", 6))#
+plot_run(run)
+```
+
 ## References

### 2016-10-25 R/Fig4.R
update figures with new rCH

@@ -1,27 +1,43 @@
-library(reshape2)
-library(dplyr)
+# Figure 4
+# Seasonal variation in S:H at different DIN levels
+# Values based on Stimson et al. 1997
 
-source("R/def_pars.R")
-source("R/init_env.R")
-source("R/run_coral.R")
-source("R/plot_results.R")
+# Load functions
+sapply(c("R/def_pars.R",
+         "R/init_env.R",
+         "R/run_coral.R"), 
+       source, .GlobalEnv)
+
+# Set parameters
+defpars <- def_pars()  # Get default parameters
 
 # Set run time vector
 time <- seq(1, 365, 0.1)  # if single model run, use time input
 
-# Initialize environment
-env <- init_env(time=time, 
-                L=c(5,30,2),  # Varying light
-                N=c(1e-7,1e-7,0),  # Low external DIN
-                X=c(0,0,0))  # No feeding
+# Initialize environments
+env1 <- init_env(time=time, L=c(20,40,2), N=c(1e-7,1e-7,0), X=c(1e-5,1e-5,0))
+env2 <- init_env(time=time, L=c(20,40,2), N=c(1e-6,1e-6,0), X=c(1e-5,1e-5,0))
 
-# Set parameters
-pars <- def_pars()  # Get default parameters 
+# Run simulations
+ss1 <- with(run_coral_ss(env=list(L=30,N=1e-7,X=1e-5), pars=defpars), last(S$S/H$H))
+run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", ss1))
+ss2 <- with(run_coral_ss(env=list(L=30,N=1e-6,X=1e-5), pars=defpars), last(S$S/H$H))
+run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", ss2))
 
-# Run simulation
-run <- run_coral(time=time, env=env, pars=pars)
-
-# Make figure
-png("img/Fig4.png", height=6, width=3, units="in", res=300)
-plot_run2(run)
+# Create figure
+png("img/Fig4.png", width=3, height=4, units="in", res=300)
+layout(mat=matrix(c(1,2,2,3,3)))
+par(mar=c(0,3,1,1), mgp=c(1.5,0.1,0), tcl=0.25, xaxs="i")
[…]

### 2016-10-25 R/Fig5.R

@@ -1,0 +1,86 @@
+# Figure 5
+# Bleaching in response to increasing light
+
+# Load functions
+sapply(c("R/def_pars.R",
+         "R/init_env.R",
+         "R/run_coral.R"), 
+       source, .GlobalEnv)
+
+# Set run time vector
+time <- seq(1, 80, 0.1)  # if single model run, use time input
+
+# Initialize environment
+env <- init_env(time=time, L=c(27,40,0), N=c(1e-7,1e-7,0), X=c(0e-6,0e-6,0))
+#L <- list(c(rep(20,100), rep(50,300))[-(1:9)])
+#env <- replace(env, "L", L)
+
+# Set parameters
+defpars <- def_pars()  # Get default parameters
+
+# What if sigmaCH is just proportional to S:H biomass?
+
+
+
+
+# Run simulation - this one looks good with minrule NPQ
+#ss <- with(run_coral_ss(env=list(L=20, N=1e-7, X=0), pars=replace(defpars, c("b", "kNPQ", "kCO2"), c(5, 60, 10))), last(S$S/H$H))
+#run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "b", "kNPQ", "kCO2"), c(0.2, 5, 60, 10)))#
+#plot_run(run)
+
+# Run simulation
+ss <- with(run_coral_ss(env=list(L=25, N=1e-7, X=0e-6), pars=defpars), last(S$S/H$H))
+run <- run_coral(time=time, env=env, pars=replace(defpars, "initS", ss))
+
+
+# Plot results
+
+png("img/Fig5.png", width=3, height=3, units="in", res=300)
+  with(run, {
+    # Set up graphical parameters
+    par(mar=c(1.5,1.5,1,1), oma=c(0,0,0,0), mfcol=c(3,2), mgp=c(0.8,0,0), tck=0.025, lwd=1, xaxs="i", 
+        cex.main=0.9, cex.axis=0.6, cex.lab=0.7)
+    
+    # External irradiance
+    plot(time, env$L, type="l", col="gold", ylim=c(25,45), lwd=2, xlab="", ylab="molph/m2/d")
+    title("A. External irradiance", adj=0, line=0.25)
+    
+    # Plot ROS
+    plot(NA, xlim=range(time), xlab="", ylim=c(1, max(2, max(S$cROS))), ylab="Relative")
[…]

### 2016-10-25 R/Fig6.R

@@ -1,0 +1,32 @@
+time <- seq(1, 200, 0.1)  # Set time
+defpars <- def_pars()  # Get default parameters
+pars <- defpars
+
+# Run 1
+env1 <- init_env(time=time, L=c(20,39,2), N=c(1e-7,1e-7,0), X=c(3e-6,3e-6,0))
+ss1 <- with(run_coral_ss(env=list(L=29.6, N=1e-7, X=3e-6), pars=pars), last(S$S/H$H))
+run1 <- run_coral(time=time, env=env1, pars=replace(pars, "initS", ss1))
+# Run 2 - increase food
+env2 <- init_env(time=time, L=c(20,39,2), N=c(1e-7,1e-7,0), X=c(6e-6,6e-6,0))
+ss2 <- with(run_coral_ss(env=list(L=29.6, N=1e-7, X=6e-6), pars=pars), last(S$S/H$H))
+run2 <- run_coral(time=time, env=env2, pars=replace(pars, "initS", ss2))
+# Run 3 - increase DIN
+env3 <- init_env(time=time, L=c(20,39,2), N=c(1e-6,1e-6,0), X=c(3e-6,3e-6,0))
+ss3 <- with(run_coral_ss(env=list(L=29.6, N=1e-6, X=3e-6), pars=pars), last(S$S/H$H))
+run3 <- run_coral(time=time, env=env3, pars=replace(pars, "initS", ss3))
+
+
+# Plot
+png("img/Fig6.png", width=3, height=3, units="in", res=300)
+par(mfrow=c(1,1), mar=c(2,2,1,2), mgp=c(1,0,0), tcl=0.25, xaxs="i", yaxs="i")
+plot(time, env1$L, col=alpha("gold", 0.5), type="l", lwd=2, axes=F, xlab="Days", ylab="S:H biomass ratio", ylim=c(25,40), cex.lab=0.9)
+axis(side=4, cex.axis=0.75); mtext(side=4, text="Light (mol ph/m2/d)", line=1, cex=0.9)
+par(new=T)
+plot(NA, xlim=range(time), ylim=c(0.0,0.16), yaxs="i", ylab="", xlab="", cex.axis=0.75, cex.lab=0.9)
+with(run1, lines(time, S$S/H$H, lty=1))
+with(run2, lines(time, S$S/H$H, lty=2))
+with(run3, lines(time, S$S/H$H, lty=3))
[…]

### 2016-10-25 R/Fig7.R

@@ -1,0 +1,23 @@
+# Increase light for 1000 days then decrease for 1000 days
+# Set run time vector
+time <- seq(1, 8000, 0.1)  # if single model run, use time input
+# Initialize environment
+env1 <- init_env(time=time, L=c(25,50,0), N=c(2e-7,2e-7,0), X=c(2e-6,2e-6,0))
+L <- list(c(seq(20,40, length.out=40000), seq(40,20, length.out=40000))[-(1:9)])
+env1 <- replace(env1, "L", L)
+
+ss <- with(run_coral_ss(env=list(L=20, N=2e-7, X=2e-6), pars=defpars), last(S$S/H$H))
+
+run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", ss))#
+
+png("img/Fig7.png", width=3, height=3, units="in", res=300)
+par(mar=c(2,2,1,1), mgp=c(1,0,0), tcl=0.25, cex.axis=0.5, cex.lab=0.75, xaxs="i")
+with(run1, {
+  plot(env$L, S$S/H$H, type="l", xlab="Light", ylab="S:H biomass", ylim=c(0,0.24))
+  for (i in seq(3750,length(time), 1900)) {
+    arrows(env$L[i-10], S$S[i-10]/H$H[i-10], 
+           env$L[i], S$S[i]/H$H[i], 
+           code=2, length=0.075)
+  }
+})
+dev.off()

### 2016-10-26 R/Fig4.R
update required scripts sourcing

@@ -5,7 +5,8 @@
 # Load functions
 sapply(c("R/def_pars.R",
          "R/init_env.R",
-         "R/run_coral.R"), 
+         "R/run_coral.R",
+         "R/run_coral_ss.R"), 
        source, .GlobalEnv)
 
 # Set parameters

### 2016-10-26 R/Fig5.R

@@ -4,7 +4,8 @@
 # Load functions
 sapply(c("R/def_pars.R",
          "R/init_env.R",
-         "R/run_coral.R"), 
+         "R/run_coral.R",
+         "R/run_coral_ss.R"), 
        source, .GlobalEnv)
 
 # Set run time vector

### 2016-10-26 R/Fig6.R

@@ -1,3 +1,10 @@
+# Load functions
+sapply(c("R/def_pars.R",
+         "R/init_env.R",
+         "R/run_coral.R",
+         "R/run_coral_ss.R"), 
+       source, .GlobalEnv)
+
 time <- seq(1, 200, 0.1)  # Set time
 defpars <- def_pars()  # Get default parameters
 pars <- defpars

### 2016-10-26 R/Fig7.R

@@ -1,4 +1,11 @@
-# Increase light for 1000 days then decrease for 1000 days
+# Load functions
+sapply(c("R/def_pars.R",
+         "R/init_env.R",
+         "R/run_coral.R",
+         "R/run_coral_ss.R"), 
+       source, .GlobalEnv)
+
+# Increase light for 8000 days then decrease for 8000 days
 # Set run time vector
 time <- seq(1, 8000, 0.1)  # if single model run, use time input
 # Initialize environment

### 2016-10-26 R/Fig7.R
0.008 mol/d, format tweaks

@@ -5,23 +5,30 @@ sapply(c("R/def_pars.R",
          "R/run_coral_ss.R"), 
        source, .GlobalEnv)
 
-# Increase light for 8000 days then decrease for 8000 days
+# Load default parameters
+defpars <- def_pars()
+
 # Set run time vector
-time <- seq(1, 8000, 0.1)  # if single model run, use time input
-# Initialize environment
-env1 <- init_env(time=time, L=c(25,50,0), N=c(2e-7,2e-7,0), X=c(2e-6,2e-6,0))
-L <- list(c(seq(20,40, length.out=40000), seq(40,20, length.out=40000))[-(1:9)])
+time <- seq(1, 5000, 0.1)  # if single model run, use time input
+
+# Initialize environment (with null light profile - to be replaced)
+env1 <- init_env(time=time, L=c(0,0,0), N=c(2e-7,2e-7,0), X=c(2e-6,2e-6,0))
+
+# Create light profile: increase from 20-40 for 2500 days (=.008 mol/d) then decrease for 2500 days
+L.incr <- seq(20, 40, len=length(time) %/% 2)
+L.decr <- seq(40, 20, len=length(time) - length(L.incr))
+L <- list(c(L.incr, L.decr))
 env1 <- replace(env1, "L", L)
 
 ss <- with(run_coral_ss(env=list(L=20, N=2e-7, X=2e-6), pars=defpars), last(S$S/H$H))
 
 run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", ss))#
 
-png("img/Fig7.png", width=3, height=3, units="in", res=300)
+png("img/Fig7.png", width=4, height=4, units="in", res=300)
 par(mar=c(2,2,1,1), mgp=c(1,0,0), tcl=0.25, cex.axis=0.5, cex.lab=0.75, xaxs="i")
 with(run1, {
   plot(env$L, S$S/H$H, type="l", xlab="Light", ylab="S:H biomass", ylim=c(0,0.24))
-  for (i in seq(3750,length(time), 1900)) {
+  for (i in seq(1300, length(time), last(time)/3.5)) {
     arrows(env$L[i-10], S$S[i-10]/H$H[i-10], 
            env$L[i], S$S[i]/H$H[i], 
            code=2, length=0.075)

### 2016-10-26 JTB_ms/JTB_ms.Rmd
Fig4-7 captions

@@
 Light-driven seasonal dynamics of symbiont abundance and coral growth.
-Light input (upper panel) was designed as a sinusoidal curve with a period of one year, with maximum and minimum values of 40 and 22 mol photons m^-2^ d^-1^, based on @Stimson:1997p3837.
-The dynamic behavior of symiont to host biomass ratio (middle panel) and the specific growth rate of host biomass (lower panel) show seasonal oscillations that are greater in magnitude under higher nutrients (1e-5 molN L^-1^; dashed lines) relative to lower nutrients (1e-7 molN L^-1^; solid lines), consistent with the findings of @Stimson:1997p3837.
+Light input (upper panel) was designed as a sinusoidal curve with a period of one year, with maximum and minimum values of 40 and 20 mol photons m^-2^ d^-1^, based on @Stimson:1997p3837.
+The dynamic behavior of symiont to host biomass ratio (middle panel) and the specific growth rate of host biomass (lower panel) show seasonal oscillations that are greater in magnitude under higher nutrients (1e-6 molN L^-1^; dashed lines) relative to lower nutrients (1e-7 molN L^-1^; solid lines), consistent with the findings of @Stimson:1997p3837.
 Prey density was set at 1e-5 CmolX L^-1^.](../img/Fig4.png)
@@
 
-![Figure 5. Bleaching.](../img/Fig5.png)
+![Figure 5.
+Coral bleaching as a switch from a nitrogen- to carbon-limited alternative stable state.
+This transition is demonstrated in response to gradually increasing external light (A), which causes an increase in production of ROS (B) that reduces the photosynthetic rate through photoinhibition (C).
+Decreasing photosynthesis moves the system from overall nitrogen-limitation toward carbon-limitation (D); when this threshold is crossed, the system rapidly becomes highly carbon-limited as photosynthesis becomes CO~2~-limited (E) and symbiont densities rapidly decline (F) into a bleached state.
+(All parameters at default values; external DIN=1e-7 molN/L, prey density=0).](../img/Fig5.png)
 
@@
 
-![Figure 6. Bleaching and interactive factors.](../img/Fig6.png)
+![Figure 6.
+Bleaching with interactive factors.
+Simulations of high light stress (sinusoid with maximum=39 mol ph m^-2^ d^-1^) under default environmental conditions (solid line; DIN=1e-7 molN/L; prey=3e-6 CmolX/L), or with elevated feeding (dashed line; prey=6e-6 CmolX/L), or elevated nutrients (dotted line; DIN=1e-6 molN/L).
+Initial symbiont biomass was set to the steady state for each set of starting conditions, respectively.
+(All parameters set to default values).](../img/Fig6.png)
 
@@
 
-![Figure 7. Hysteresis.](../img/Fig7.png)
+![Figure 7.
+Hysteresis in coral bleaching and recovery.
+The symbiont:host biomass ratio is plotted as a function of light from a simulation of linearly increasing, then decreasing, external light.
+The direction of the trajectory is indicated with arrowheads evenly spaced with respect to time.
+The rate of change in light was slow (0.005 mol ph d^-1^) such that dynamics track closely to steady state values.
+(All parameters at default values; DIN=2e-7 molN/L; Prey=2e-6 CmolX/L).](../img/Fig7.png)
 

### 2016-10-26 R/Fig5.R
clean

@@ -19,23 +19,11 @@ env <- init_env(time=time, L=c(27,40,0), N=c(1e-7,1e-7,0), X=c(0e-6,0e-6,0))
 # Set parameters
 defpars <- def_pars()  # Get default parameters
 
-# What if sigmaCH is just proportional to S:H biomass?
-
-
-
-
-# Run simulation - this one looks good with minrule NPQ
-#ss <- with(run_coral_ss(env=list(L=20, N=1e-7, X=0), pars=replace(defpars, c("b", "kNPQ", "kCO2"), c(5, 60, 10))), last(S$S/H$H))
-#run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "b", "kNPQ", "kCO2"), c(0.2, 5, 60, 10)))#
-#plot_run(run)
-
 # Run simulation
 ss <- with(run_coral_ss(env=list(L=25, N=1e-7, X=0e-6), pars=defpars), last(S$S/H$H))
 run <- run_coral(time=time, env=env, pars=replace(defpars, "initS", ss))
 
-
 # Plot results
-
 png("img/Fig5.png", width=3, height=3, units="in", res=300)
   with(run, {
     # Set up graphical parameters

### 2016-10-27 R/run_coral.R
add sigmaCH back into rCH equation

@@ -59,7 +59,7 @@ run_coral <- function(time, env, pars) {
     # CO2 input flux
     S$rCS[t] <- pars$jST0 * pars$sigmaCS  # metabolic CO2 recycled from symbiont biomass turnover
     #H$rCH[t] <- H$jHT[t-1] * pars$sigmaCH  # metabolic CO2 recycled from host biomass turnover
-    H$rCH[t] <- H$jHT[t-1] * S$S[t-1]/H$H[t-1]  #
+    H$rCH[t] <- H$jHT[t-1] * pars$sigmaCH * S$S[t-1]/H$H[t-1]  #
     H$jCO2[t] <- pars$kCO2 * H$jeC[t-1]  # carbon not used in host biomass is used to activate CCM's that deliver CO2 to photosynthesis
     # Production flux (photosynthetic carbon fixation)
     S$jCP[t] <- synth(S$jL[t] * pars$yCL, (H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t], pars$jCPm) / S$cROS[t-1]

### 2016-10-27 R/run_coral_ss.R

@@ -62,7 +62,7 @@ run_coral_ss <- function(env, pars) {
     # CO2 input flux
     S[t,]$rCS <- pars$jST0 * pars$sigmaCS  # metabolic CO2 recycled from symbiont biomass turnover
     #H[t,]$rCH <- H$jHT[t-1] * pars$sigmaCH  # metabolic CO2 recycled from host biomass turnover
-    H[t,]$rCH <- H$jHT[t-1] * S$S[t-1]/H$H[t-1]  #
+    H[t,]$rCH <- H$jHT[t-1] * pars$sigmaCH * S$S[t-1]/H$H[t-1]  #
     H[t,]$jCO2 <- pars$kCO2 * H$jeC[t-1]  # carbon not used in host biomass is used to activate CCM's that deliver CO2 to photosynthesis
     # Production flux (photosynthetic carbon fixation)
     S[t,]$jCP <- synth(S[t,]$jL * pars$yCL, (H[t,]$jCO2 + H[t,]$rCH)*H$H[t-1]/S$S[t-1] + S[t,]$rCS, pars$jCPm) / S$cROS[t-1]

### 2016-10-27 JTB_ms/JTB_ms.Rmd
abstract

@@
 abstract: |
-  This is the abstract.
-  Some type of positive feedback is necessary.
-  Compare scenarios of recycling CO2, have photodamage increase maintenance that does recycle CO2.
-  Compare recycled N to symbiont instead of host
-  How much do you have to reduce the light to get recovery? i.e., magnitude of hysteresis?
+  Coral reef ecosystems owe their ecological success--and vulnerability to climate change--to the symbiotic metabolism of corals and *Symbiodinium* spp.
+The urgency to understand and predict the stability and breakdown of these symbioses (i.e., coral 'bleaching') calls for the development and application of theoretical tools.
+While the biological complexity of the symbiosis is well-suited to dynamical modeling, the theoretical complexity of certain approaches makes them inaccessible to many biologists, thus limiting their potential application and utility.
+Here, we develop a simplified dynamic bioenergetic model of coral-algal symbiosis that demonstrates realistic steady-state behavior across gradients of light, nutrients, and feeding.
+Furthermore, through a mechanistic treatment of photo-oxidative stress, the model displays dynamic responses of bleaching and recovery, and suggests that these phenomena represent transitions between alternative stable states.
+A suite of complex observed behaviors reproduced by this model suggests it has captured many important attributes of the system; meanwhile, its modular framework and open source R code are designed to facilitate expansion and problem-specific applications.
+We see high utility for this model in elucidating the mechanistic underpinnings of complex responses, generating testable hypotheses, and predicting integrated responses to simultaneous changes in multiple environmental inputs.
 
@@
   rticles::elsevier_article:
-    fig_caption: yes
+    fig_caption: no
 ---
@@
 The parameter $k_{CO_2}$ scales the efficacy of host CCMs, which enables the comparison of different rates of CO~2~ delivery that may characterize different coral species [@Wooldridge:2014di].
-The input of CO~2~ to the photosynthesis SU is therefore specified as:
+The active input of CO~2~ to the photosynthesis SU is therefore specified as:
 
@@
 
-Additional metabolic CO~2~ recycled from both the host and symbiont biomass turnover is also made available to the photosynthesis SU, according to:
+An additional fixed proportion of metabolic CO~2~ generated by host biomass turnover is made available to the photosynthesis SU (scaled to the relative amount of symbiont biomass), according to:
 
-\begin{equation} r_{CH}=\sigma_{CH}j_{HT}^0 \end{equation}
+\begin{equation} r_{CH}=\sigma_{CH}j_{HT}^0 \cdot {S \over H} \end{equation}
+
+along with a fixed proportion of symbiont biomass turnover:
+
 \begin{equation} r_{CS}=\sigma_{CS}j_{ST}^0 \end{equation}

### 2016-10-27 JTB_ms/JTB_ms.tex
abstract

@@
   \begin{abstract}
-  This is the abstract. Some type of positive feedback is necessary.
-  Compare scenarios of recycling CO2, have photodamage increase
-  maintenance that does recycle CO2. Compare recycled N to symbiont
-  instead of host How much do you have to reduce the light to get
-  recovery? i.e., magnitude of hysteresis?
+  Coral reef ecosystems owe their ecological success--and vulnerability to
+  climate change--to the symbiotic metabolism of corals and
+  \emph{Symbiodinium} spp. The urgency to understand and predict the
+  stability and breakdown of these symbioses (i.e., coral `bleaching')
+  calls for the development and application of theoretical tools. While
+  the biological complexity of the symbiosis is well-suited to dynamical
+  modeling, the theoretical complexity of certain approaches makes them
+  inaccessible to many biologists, thus limiting their potential
+  application and utility. Here, we develop a simplified dynamic
+  bioenergetic model of coral-algal symbiosis that demonstrates realistic
+  steady-state behavior across gradients of light, nutrients, and feeding.
+  Furthermore, through a mechanistic treatment of photo-oxidative stress,
+  the model displays dynamic responses of bleaching and recovery, and
+  suggests that these phenomena represent transitions between alternative
+  stable states. A suite of complex observed behaviors reproduced by this
+  model suggests it has captured many important attributes of the system;
+  meanwhile, its modular framework and open source R code are designed to
+  facilitate expansion and problem-specific applications. We see high
+  utility for this model in elucidating the mechanistic underpinnings of
+  complex responses, generating testable hypotheses, and predicting
+  integrated responses to simultaneous changes in multiple environmental
+  inputs.
   \end{abstract}
@@
 
-\begin{figure}[htbp]
-\centering
 \includegraphics{../img/Fig1.png}
-\caption{Graphical representation of coral-algal symbiosis model. Light,
-CO\textsubscript{2}, prey, and DIN are acquired from the external
-environment proportional to the biomass of the partner indicated by the
-black box for uptake. Mass fluxes (see Table 1 for definitions) are
-represented by \(j\)'s with subscripts indicating the type of mass, and
-in some cases the process (e.g., \(j_{CP}\) is the flux of carbon
-produced by photosynthesis), and \(\rho\)'s indicate fluxes that are
-shared by one partner with the other. Parallel complementary
-synthesizing units (SUs) are represented by large circles, and rejection
-fluxes from these SUs are indicated by dashed lines. \(c_{ROS}\) is a
-proportional rate that impacts other model fluxes by inhibition or
-acceleration; likewise, \(j_{eC}\) accelerates the rate of \(j_{CO_2}\).
-Recycled mass fluxes from biomass turnover are not shown for clarity
-(but see Table 1 for definitions).}
-\end{figure}
 
@@
 
-\begin{equation} r_{CH}=\sigma_{CH}j_{HT}^0 \end{equation}\begin{equation} r_{CS}=\sigma_{CS}j_{ST}^0 \end{equation}
+\begin{equation} r_{CH}=j_{HT}^0 \cdot {S \over H} \end{equation}\begin{equation} r_{CS}=\sigma_{CS}j_{ST}^0 \end{equation}
 
@@
 
-\begin{figure}[htbp]
-\centering
 \includegraphics{../img/Fig2.png}
-\caption{Steady state values of \textbf{A)} specific growth (Cmol
-Cmol\textsuperscript{-1} d\textsuperscript{-1}), \textbf{B)} the
-symbiont to host biomass ratio (CmolS CmolH\textsuperscript{-1}),
-\textbf{C)} relative C- or N-limitation of host biomass formation, and
-\textbf{D)} relative CO\textsubscript{2}- or light-limitation of
-symbiont photosynthesis, across gradients of external irradiance and
-dissolved inorganic nitrogen. Note that typical conditions for reefs are
-\textasciitilde{}1e-7 M DIN and 10-20 mol photons m\textsuperscript{-2}
-d\textsuperscript{-1}. Simulations for each combination of light and
-nutrients (41 points along each axis) were run for 100 days with a time
-step of 1 day. Negative steady state growth rates, and corresponding S:H
-ratios, were set to zero.}
-\end{figure}
 
@@
 
-\begin{figure}[htbp]
-\centering
 \includegraphics{../img/Fig3.png}
-\caption{Sensitivity analysis. Plots show the fractional change in
-steady state values of growth (solid lines) and S:H biomass (dashed
-lines) in response to fractional changes in default parameter values
-(see Table 2 for default values). Parameters are grouped by which
-processes they are involved in. This sensitivity analysis was conducted
-at conditions typical for coral reef environments: low DIN (1e-7 M) and
-intermediate light (15 mol m\textsuperscript{-2} d\textsuperscript{-2}).
-Sensitivity analyses conducted other environmental conditions are
-presented in the Supplementary Information.}
-\end{figure}
 
@@
 
-\begin{figure}[htbp]
-\centering
 \includegraphics{../img/Fig4.png}
-\caption{Figure 4. Light-driven seasonal dynamics of symbiont abundance
-and coral growth. Light input (upper panel) was designed as a sinusoidal
-curve with a period of one year, with maximum and minimum values of 40
-and 22 mol photons m\textsuperscript{-2} d\textsuperscript{-1}, based on
-Stimson (1997). The dynamic behavior of symiont to host biomass ratio
-(middle panel) and the specific growth rate of host biomass (lower
-panel) show seasonal oscillations that are greater in magnitude under
-higher nutrients (1e-5 molN L\textsuperscript{-1}; dashed lines)
-relative to lower nutrients (1e-7 molN L\textsuperscript{-1}; solid
-lines), consistent with the findings of Stimson (1997). Prey density was
-set at 1e-5 CmolX L\textsuperscript{-1}.}
-\end{figure}
 
@@
 
-\begin{figure}[htbp]
-\centering
 \includegraphics{../img/Fig5.png}
-\caption{Figure 5. Bleaching.}
-\end{figure}
 
@@
 
-\begin{figure}[htbp]
-\centering
 \includegraphics{../img/Fig6.png}
-\caption{Figure 6. Bleaching and interactive factors.}
-\end{figure}
 
@@
 
-\begin{figure}[htbp]
-\centering
 \includegraphics{../img/Fig7.png}
-\caption{Figure 7. Hysteresis.}
-\end{figure}
 

### 2016-10-27 JTB_ms/JTB_ms.tex
remake with new sigmaC*

@@
 comparison of different rates of CO\textsubscript{2} delivery that may
-characterize different coral species (Wooldridge 2014a). The input of
-CO\textsubscript{2} to the photosynthesis SU is therefore specified as:
+characterize different coral species (Wooldridge 2014a). The active
+input of CO\textsubscript{2} to the photosynthesis SU is therefore
+specified as:
 
@@
 
-Additional metabolic CO\textsubscript{2} recycled from both the host and
-symbiont biomass turnover is also made available to the photosynthesis
-SU, according to:
-
-\begin{equation} r_{CH}=j_{HT}^0 \cdot {S \over H} \end{equation}\begin{equation} r_{CS}=\sigma_{CS}j_{ST}^0 \end{equation}
+An additional fixed proportion of metabolic CO\textsubscript{2}
+generated by host biomass turnover is made available to the
+photosynthesis SU (scaled to the relative amount of symbiont biomass),
+according to:
+
+\begin{equation} r_{CH}=\sigma_{CH}j_{HT}^0 \cdot {S \over H} \end{equation}
+
+along with a fixed proportion of symbiont biomass turnover:
+
+\begin{equation} r_{CS}=\sigma_{CS}j_{ST}^0 \end{equation}
 

### 2016-10-27 JTB_supp/JTB_supp.Rmd
remake with new sigmaC*

@@
 knitr::opts_knit$set(root.dir = normalizePath(".."))
-knitr::opts_chunk$set(message = F, warning = F)
+knitr::opts_chunk$set(message = F, warning = F, echo=F)
 ```
@@
 
-par(mfcol=c(6,2), mar=c(2,2,1,4), mgp=c(1,0,0), tcl=0.2)
+par(mfcol=c(6,2), mar=c(2,2,1,4), mgp=c(1,0,0), tcl=0.2, xaxs="i")
 

### 2016-10-28 JTB_ms/JTB_ms.Rmd
update abstract

@@
 While the biological complexity of the symbiosis is well-suited to dynamical modeling, the theoretical complexity of certain approaches makes them inaccessible to many biologists, thus limiting their potential application and utility.
-Here, we develop a simplified dynamic bioenergetic model of coral-algal symbiosis that demonstrates realistic steady-state behavior across gradients of light, nutrients, and feeding.
+Here, we develop a simplified dynamic bioenergetic model of coral-algal symbiosis that demonstrates realistic steady-state patterns in coral growth and symbiont abundance across gradients of light, nutrients, and feeding.
 Furthermore, through a mechanistic treatment of photo-oxidative stress, the model displays dynamic responses of bleaching and recovery, and suggests that these phenomena represent transitions between alternative stable states.
-A suite of complex observed behaviors reproduced by this model suggests it has captured many important attributes of the system; meanwhile, its modular framework and open source R code are designed to facilitate expansion and problem-specific applications.
-We see high utility for this model in elucidating the mechanistic underpinnings of complex responses, generating testable hypotheses, and predicting integrated responses to simultaneous changes in multiple environmental inputs.
+A suite of complex observed behaviors reproduced by the model suggests it captures many important attributes of the system; meanwhile, its modular framework and open source R code are designed to facilitate expansion and problem-specific applications.
+We see high utility for this modeling framework in generating testable hypotheses and predicting integrated, mechanistic responses of corals to environmental change.
 

### 2016-10-28 JTB_supp/JTB_supp.Rmd
update sigmaC*

--- 2016-10-25 JTB_supp/JTB_supp.Rmd
@@
     + Value used: `r defpars$jST0` CmolS CmolS^-1^ d^-1^
-* $\sigma_{NH}, \sigma_{CH}, \sigma_{NS}, \sigma_{CS}$: Proportions of nitrogen and carbon recycled from host and symbiont biomass turnover
+* $\sigma_{NH}, \sigma_{NS}$: Proportions of nitrogen recycled from host and symbiont biomass turnover
     + Assuming highly efficient recycling to avoid loss to the environment:
     + Values used: `r with(defpars, c(sigmaNH, sigmaCH, sigmaNS, sigmaCS))` (dimensionless)
+* $\sigma_{CH},  \sigma_{CS}$: Proportions of CO~2~ recycled to photosynthesis from host and symbiont biomass turnover
+    + Assuming that metabolic CO~2~ rapidly converts to HCO~3~^-^ at intracellular pH ~ 7.2 [@Venn:2009hy], only ~10% actually remains as CO~2~ [@Zeebe:2001wz] which would be available to diffusively supply photosynthesis.
+This passive supply is differentiated from all active mechanisms (CCMs) that facilitate CO~2~ supply to photosynthesis, which are separately treated collectively as the $jCO~2~$ flux.
+    + Values used: 0.10
 * $j_{Xm}$: Maximum specific feeding rate of host

### 2016-11-01 JTB_ms/JTB_ms.Rmd
sigmaCS=0.9, kROS=80

@@
 
+  Latest development--sigmaCH and sigmaCS changed to 0.1 meaning 10% of metabolic CO2 produced is passively available to photosynthesis, to reflect the fact that that ~10% of total internal DIC at pH 7 is in the form of CO2.
+probably actually less than this since internal pH may be closer to 7.5.
+This does not change any of the figures substantively except the hysteresis figure--overshoot no longer occurs with specific parameters run.
+may need to just change light range or rate of change to reproduce overshoot.
+
 bibliography: library.bib
@@
 source("R/reported_metrics.R")
+# Source default parameter values
+source("R/def_pars.R")
+defpars <- def_pars()
 ```
@@
 | ------- | -------------------------------------- | ------------------- | ------- |
-| $n_{NH}$ | N:C molar ratio in host biomass | -- | 0.19 |
-| $n_{NS}$ | N:C molar ratio in symbiont biomass | -- | 0.2 |
-| $n_{NX}$ | N:C molar ratio in prey biomass | -- | 0.13 |
-| $j_{HT}^0$ | Specific turnover rate of host biomass | CmolH CmolH^-1^ d^-1^ | 0.03 |
-| $j_{ST}^0$ | Specific turnover rate of symbiont biomass | CmolS CmolS^-1^ d^-1^ | 0.03 |
-| $\sigma_{NH}$ | Proportion host nitrogen turnover recycled | -- | 0.9 |
-| $\sigma_{CH}$ | Proportion host carbon turnover recycled | -- | 0.9 |
-| $\sigma_{NS}$ | Proportion symbiont nitrogen turnover recycled | -- | 0.9 |
-| $\sigma_{CS}$ | Proportion symbiont carbon turnover recycled | -- | 0.9 |
-| $j_{Xm}$ | Maximum specific feeding rate of host | molX CmolH^-1^ d^-1^ | 0.1292 |
-| $K_X$ | Half-saturation constant for prey uptake by host | molX L^-1^ | 20e-6 |
-| $j_{Nm}$ | Maximum specific DIN uptake rate by host | molN CmolH^-1^ d^-1^ | 0.048 |
-| $K_N$ | Half-saturation constant for DIN uptake by host | molN L^-1^ | 0.46e-6 |
-| $k_{CO_2}$ | Efficacy of CO~2~ delivery by host CCMs | molCO~2~ molC^-1^ | 10 |
-| $j_{HGm}$ | Maximum specific growth rate of host | CmolH CmolH^-1^ d^-1^ | 1 |
-| $y_{CL}$ | Quantum yield of photosynthesis | molC mol ph^-1^ | 0.1 |
-| $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | m^2^ CmolS^-1^ | 1.34 |
-| $k_{NPQ}$ | NPQ capacity of symbiont | mol ph CmolS^-1^ d^-1^ | 112 |
-| $k_{ROS}$ | Excess photon energy that doubles ROS production, relative to baseline levels | mol ph CmolS^-1^ d^-1^ | 40-80 |
-| $j_{CPm}$ | Maximum specific photosynthesis rate of symbiont | molC CmolS^-1^ d^-1^ | 2.8 |
-| $j_{SGm}$ | Maximum specific growth rate of symbiont | CmolS CmolS^-1^ d^-1^ | 0.25 |
-| $b$ | Scaling parameter for strength of bleaching response | -- | 5 |
+| $n_{NH}$ | N:C molar ratio in host biomass | -- | $`r defpars$nNH`$ |
+| $n_{NS}$ | N:C molar ratio in symbiont biomass | -- | $`r defpars$nNS`$ |
+| $n_{NX}$ | N:C molar ratio in prey biomass | -- | $`r defpars$nNX`$ |
+| $j_{HT}^0$ | Specific turnover rate of host biomass | CmolH CmolH^-1^ d^-1^ | $`r defpars$jHT0`$ |
+| $j_{ST}^0$ | Specific turnover rate of symbiont biomass | CmolS CmolS^-1^ d^-1^ | $`r defpars$jST0`$ |
+| $\sigma_{NH}$ | Proportion host N-turnover recycled to host | -- | $`r defpars$sigmaNH`$ |
+| $\sigma_{CH}$ | Proportion host C-turnover recycled to photosynthesis | -- | $`r defpars$sigmaCH`$ |
+| $\sigma_{NS}$ | Proportion symbiont N-turnover recycled to symbiont | -- | $`r defpars$sigmaNS`$ |
+| $\sigma_{CS}$ | Proportion symbiont C-turnover recycled to photosynthesis | -- | $`r defpars$sigmaCS`$ |
+| $j_{Xm}$ | Maximum specific feeding rate of host | molX CmolH^-1^ d^-1^ | $`r defpars$jXm`$ |
+| $K_X$ | Half-saturation constant for prey uptake by host | molX L^-1^ | $`r defpars$KX`$ |
+| $j_{Nm}$ | Maximum specific DIN uptake rate by host | molN CmolH^-1^ d^-1^ | $`r defpars$jNm`$ |
+| $K_N$ | Half-saturation constant for DIN uptake by host | molN L^-1^ | $`r defpars$KN`$ |
+| $k_{CO_2}$ | Efficacy of CO~2~ delivery by host CCMs | molCO~2~ molC^-1^ | $`r defpars$kCO2`$ |
+| $j_{HGm}$ | Maximum specific growth rate of host | CmolH CmolH^-1^ d^-1^ | $`r defpars$jHGm`$ |
+| $y_{CL}$ | Quantum yield of photosynthesis | molC mol ph^-1^ | $`r defpars$yCL`$ |
+| $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | m^2^ CmolS^-1^ | $`r defpars$astar`$ |
+| $k_{NPQ}$ | NPQ capacity of symbiont | mol ph CmolS^-1^ d^-1^ | $`r defpars$kNPQ`$ |
+| $k_{ROS}$ | Excess photon energy that doubles ROS production, relative to baseline levels | mol ph CmolS^-1^ d^-1^ | $`r defpars$kROS`$ |
+| $j_{CPm}$ | Maximum specific photosynthesis rate of symbiont | molC CmolS^-1^ d^-1^ | $`r defpars$jCPm`$ |
+| $j_{SGm}$ | Maximum specific growth rate of symbiont | CmolS CmolS^-1^ d^-1^ | $`r defpars$jSGm`$ |
+| $b$ | Scaling parameter for strength of bleaching response | -- | $`r defpars$b`$ |
 
@@
 
-An additional fixed proportion of metabolic CO~2~ generated by host biomass turnover is made available to the photosynthesis SU (scaled to the relative amount of symbiont biomass), according to:
+An additional fixed proportion of metabolic CO~2~ generated by host biomass turnover is passively made available to the photosynthesis SU, according to:
 
-\begin{equation} r_{CH}=\sigma_{CH}j_{HT}^0 \cdot {S \over H} \end{equation}
+\begin{equation} r_{CH}=\sigma_{CH}j_{HT}^0 \end{equation}
 

### 2016-11-01 JTB_ms/JTB_ms.tex
sigmaCS=0.9, kROS=80

@@
   bioenergetic model of coral-algal symbiosis that demonstrates realistic
-  steady-state behavior across gradients of light, nutrients, and feeding.
-  Furthermore, through a mechanistic treatment of photo-oxidative stress,
-  the model displays dynamic responses of bleaching and recovery, and
-  suggests that these phenomena represent transitions between alternative
-  stable states. A suite of complex observed behaviors reproduced by this
-  model suggests it has captured many important attributes of the system;
-  meanwhile, its modular framework and open source R code are designed to
-  facilitate expansion and problem-specific applications. We see high
-  utility for this model in elucidating the mechanistic underpinnings of
-  complex responses, generating testable hypotheses, and predicting
-  integrated responses to simultaneous changes in multiple environmental
-  inputs.
+  steady-state patterns in coral growth and symbiont abundance across
+  gradients of light, nutrients, and feeding. Furthermore, through a
+  mechanistic treatment of photo-oxidative stress, the model displays
+  dynamic responses of bleaching and recovery, and suggests that these
+  phenomena represent transitions between alternative stable states. A
+  suite of complex observed behaviors reproduced by the model suggests it
+  captures many important attributes of the system; meanwhile, its modular
+  framework and open source R code are designed to facilitate expansion
+  and problem-specific applications. We see high utility for this modeling
+  framework in generating testable hypotheses and predicting integrated,
+  mechanistic responses of corals to environmental change.
+
+  Latest development--sigmaCH and sigmaCS changed to 0.1 meaning 10\% of
+  metabolic CO2 produced is passively available to photosynthesis, to
+  reflect the fact that that \textasciitilde{}10\% of total internal DIC
+  at pH 7 is in the form of CO2. probably actually less than this since
+  internal pH may be closer to 7.5. This does not change any of the
+  figures substantively except the hysteresis figure--overshoot no longer
+  occurs with specific parameters run. may need to just change light range
+  or rate of change to reproduce overshoot.
   \end{abstract}
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.19
+\(0.18\)
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.2
+\(0.13\)
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.13
+\(0.2\)
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.03
+\(0.03\)
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Proportion host nitrogen turnover recycled
+Proportion host N-turnover recycled to host
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.9
+\(0.9\)
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Proportion host carbon turnover recycled
+Proportion host C-turnover recycled to photosynthesis
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.9
+\(0.1\)
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Proportion symbiont nitrogen turnover recycled
+Proportion symbiont N-turnover recycled to symbiont
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Proportion symbiont carbon turnover recycled
+Proportion symbiont C-turnover recycled to photosynthesis
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.1292
+\(0.13\)
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-20e-6
+\(1.95\times 10^{-5}\)
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.048
+\(0.035\)
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.46e-6
+\(1.5\times 10^{-6}\)
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-10
+\(10\)
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-1
+\(1\)
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.1
+\(0.1\)
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-1.34
+\(1.34\)
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-112
+\(112\)
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-40-80
+\(80\)
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-2.8
+\(2.8\)
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.25
+\(0.25\)
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-5
+\(5\)
 \strut\end{minipage}\tabularnewline
@@
 An additional fixed proportion of metabolic CO\textsubscript{2}
-generated by host biomass turnover is made available to the
-photosynthesis SU (scaled to the relative amount of symbiont biomass),
-according to:
-
-\begin{equation} r_{CH}=\sigma_{CH}j_{HT}^0 \cdot {S \over H} \end{equation}
+generated by host biomass turnover is passively made available to the
+photosynthesis SU, according to:
+
+\begin{equation} r_{CH}=\sigma_{CH}j_{HT}^0 \end{equation}
 

### 2016-11-01 JTB_supp/JTB_supp.Rmd
sigmaCS=0.9, kROS=80

--- 2016-10-27 JTB_supp/JTB_supp.Rmd
@@
     + Value used: `r defpars$jST0` CmolS CmolS^-1^ d^-1^
-* $\sigma_{NH}, \sigma_{CH}, \sigma_{NS}, \sigma_{CS}$: Proportions of nitrogen and carbon recycled from host and symbiont biomass turnover
-    + Assuming highly efficient recycling to avoid loss to the environment:
-    + Values used: `r with(defpars, c(sigmaNH, sigmaCH, sigmaNS, sigmaCS))` (dimensionless)
+* $\sigma_{NH}, \sigma_{NS}$: Proportions of nitrogen recycled from host and symbiont biomass turnover
+    + Assume highly efficient recycling by both symbiont and host to avoid loss to the environment:
+    + Values used: `r with(defpars, c(sigmaNH, sigmaNS))` (dimensionless)
+* $\sigma_{CS}$: Proportion of metabolic CO~2~ produced by symbiont recycled passively to photosynthesis
+    + Assume that CO~2~ produced within symbiont cells is recycled with high efficiency, potentially aided by symbiont carbon concentrating mechanisms
+    + Value used: 0.9 (dimensionless)
+* $\sigma_{CH}$: Proportions of metabolic CO~2~ produced by host recycled passively to photosynthesis 
+    + Assume that most metabolic CO~2~ produced within host cells rapidly converts to HCO~3~^-^ at intracellular pH ~ 7.2 [@Venn:2009hy], and that only ~10% actually remains as CO~2~ [@Zeebe:2001wz] that could diffusively supply photosynthesis.
+This passive supply is differentiated from all active host carbon concentrating mechanisms (CCMs) that facilitate CO~2~ supply to photosynthesis, which are separately treated collectively as the $jCO~2~$ flux.
+    + Values used: 0.1 (dimensionless)
 * $j_{Xm}$: Maximum specific feeding rate of host
@@
 # Set run time vector
-time <- seq(1, 365, 0.5)  # if single model run, use time input
+time <- seq(1, 365, 0.1)  # if single model run, use time input
 
 # Initialize environments
-env1 <- init_env(time=time, L=c(20,40,2), N=c(1e-7,1e-7,0), X=c(1e-5,1e-5,0))
-env2 <- init_env(time=time, L=c(20,40,2), N=c(1e-6,1e-6,0), X=c(1e-5,1e-5,0))
+env1 <- init_env(time=time, L=c(22,44,2), N=c(1e-7,1e-7,0), X=c(1e-5,1e-5,0))
+env2 <- init_env(time=time, L=c(22,44,2), N=c(1e-5,1e-5,0), X=c(1e-5,1e-5,0))
 
 # Run simulations
-ss1 <- with(run_coral_ss(env=list(L=30,N=1e-7,X=1e-5), pars=defpars), last(S$S/H$H))
+ss1 <- with(run_coral_ss(env=list(L=33,N=1e-7,X=1e-5), pars=defpars), last(S$S/H$H))
 run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", ss1))
-ss2 <- with(run_coral_ss(env=list(L=30,N=1e-6,X=1e-5), pars=defpars), last(S$S/H$H))
+ss2 <- with(run_coral_ss(env=list(L=33,N=1e-5,X=1e-5), pars=defpars), last(S$S/H$H))
 run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", ss2))
@@
 # Run simulation
-run <- run_coral(time=time, env=env, pars=replace(defpars, "kCO2", 6))#
+run <- run_coral(time=time, env=env, pars=defpars)
 plot_run(run)

### 2016-11-01 R/Fig4.R

@@ -16,20 +16,20 @@ defpars <- def_pars()  # Get default parameters
 time <- seq(1, 365, 0.1)  # if single model run, use time input
 
 # Initialize environments
-env1 <- init_env(time=time, L=c(20,40,2), N=c(1e-7,1e-7,0), X=c(1e-5,1e-5,0))
-env2 <- init_env(time=time, L=c(20,40,2), N=c(1e-6,1e-6,0), X=c(1e-5,1e-5,0))
+env1 <- init_env(time=time, L=c(22,44,2), N=c(1e-7,1e-7,0), X=c(1e-5,1e-5,0))
+env2 <- init_env(time=time, L=c(22,44,2), N=c(1e-5,1e-5,0), X=c(1e-5,1e-5,0))
 
 # Run simulations
-ss1 <- with(run_coral_ss(env=list(L=30,N=1e-7,X=1e-5), pars=defpars), last(S$S/H$H))
+ss1 <- with(run_coral_ss(env=list(L=33,N=1e-7,X=1e-5), pars=defpars), last(S$S/H$H))
 run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", ss1))
-ss2 <- with(run_coral_ss(env=list(L=30,N=1e-6,X=1e-5), pars=defpars), last(S$S/H$H))
+ss2 <- with(run_coral_ss(env=list(L=33,N=1e-5,X=1e-5), pars=defpars), last(S$S/H$H))
 run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", ss2))
 
 # Create figure
 png("img/Fig4.png", width=3, height=4, units="in", res=300)
 layout(mat=matrix(c(1,2,2,3,3)))
 par(mar=c(0,3,1,1), mgp=c(1.5,0.1,0), tcl=0.25, xaxs="i")
-plot(run1$env$L ~ time, type="l", lwd=3, col=alpha("gold", 0.5), ylim=c(20,40), yaxs="i", 
+plot(run1$env$L ~ time, type="l", lwd=3, col=alpha("gold", 0.5), ylim=c(22,44), yaxs="i", 
      xlab="", ylab="Light", xaxt="n", bty="n", xpd=NA)
 par(mar=c(1.5,3,1.5,1))
 plot(NA, xlim=range(time), ylim=range(0,0.25), xlab="", ylab="S:H biomass", bty="n", xaxt="n")

### 2016-11-01 R/Fig5.R

@@ -12,7 +12,7 @@ sapply(c("R/def_pars.R",
 time <- seq(1, 80, 0.1)  # if single model run, use time input
 
 # Initialize environment
-env <- init_env(time=time, L=c(27,40,0), N=c(1e-7,1e-7,0), X=c(0e-6,0e-6,0))
+env <- init_env(time=time, L=c(30,50,0), N=c(1e-7,1e-7,0), X=c(0e-6,0e-6,0))
 #L <- list(c(rep(20,100), rep(50,300))[-(1:9)])
 #env <- replace(env, "L", L)
 
@@ -31,7 +31,7 @@ png("img/Fig5.png", width=3, height=3, units="in", res=300)
         cex.main=0.9, cex.axis=0.6, cex.lab=0.7)
     
     # External irradiance
-    plot(time, env$L, type="l", col="gold", ylim=c(25,45), lwd=2, xlab="", ylab="molph/m2/d")
+    plot(time, env$L, type="l", col="gold", ylim=c(30,50), lwd=2, xlab="", ylab="molph/m2/d")
     title("A. External irradiance", adj=0, line=0.25)
     
     # Plot ROS

### 2016-11-01 R/Fig6.R

@@ -10,23 +10,23 @@ defpars <- def_pars()  # Get default parameters
 pars <- defpars
 
 # Run 1
-env1 <- init_env(time=time, L=c(20,39,2), N=c(1e-7,1e-7,0), X=c(3e-6,3e-6,0))
-ss1 <- with(run_coral_ss(env=list(L=29.6, N=1e-7, X=3e-6), pars=pars), last(S$S/H$H))
+env1 <- init_env(time=time, L=c(25,48,2), N=c(1e-7,1e-7,0), X=c(3e-6,3e-6,0))
+ss1 <- with(run_coral_ss(env=list(L=36, N=1e-7, X=3e-6), pars=pars), last(S$S/H$H))
 run1 <- run_coral(time=time, env=env1, pars=replace(pars, "initS", ss1))
 # Run 2 - increase food
-env2 <- init_env(time=time, L=c(20,39,2), N=c(1e-7,1e-7,0), X=c(6e-6,6e-6,0))
-ss2 <- with(run_coral_ss(env=list(L=29.6, N=1e-7, X=6e-6), pars=pars), last(S$S/H$H))
+env2 <- init_env(time=time, L=c(25,48,2), N=c(1e-7,1e-7,0), X=c(6e-6,6e-6,0))
+ss2 <- with(run_coral_ss(env=list(L=36, N=1e-7, X=6e-6), pars=pars), last(S$S/H$H))
 run2 <- run_coral(time=time, env=env2, pars=replace(pars, "initS", ss2))
 # Run 3 - increase DIN
-env3 <- init_env(time=time, L=c(20,39,2), N=c(1e-6,1e-6,0), X=c(3e-6,3e-6,0))
-ss3 <- with(run_coral_ss(env=list(L=29.6, N=1e-6, X=3e-6), pars=pars), last(S$S/H$H))
+env3 <- init_env(time=time, L=c(25,48,2), N=c(1e-6,1e-6,0), X=c(3e-6,3e-6,0))
+ss3 <- with(run_coral_ss(env=list(L=36, N=1e-6, X=3e-6), pars=pars), last(S$S/H$H))
 run3 <- run_coral(time=time, env=env3, pars=replace(pars, "initS", ss3))
 
 
 # Plot
 png("img/Fig6.png", width=3, height=3, units="in", res=300)
 par(mfrow=c(1,1), mar=c(2,2,1,2), mgp=c(1,0,0), tcl=0.25, xaxs="i", yaxs="i")
[…]

### 2016-11-01 R/Fig7.R

@@ -7,28 +7,29 @@ sapply(c("R/def_pars.R",
 
 # Load default parameters
 defpars <- def_pars()
+defpars <- replace(defpars, "kROS", 80)
 
 # Set run time vector
 time <- seq(1, 5000, 0.1)  # if single model run, use time input
 
 # Initialize environment (with null light profile - to be replaced)
-env1 <- init_env(time=time, L=c(0,0,0), N=c(2e-7,2e-7,0), X=c(2e-6,2e-6,0))
+env1 <- init_env(time=time, L=c(0,0,0), N=c(1e-7,1e-7,0), X=c(2e-6,2e-6,0))
 
 # Create light profile: increase from 20-40 for 2500 days (=.008 mol/d) then decrease for 2500 days
-L.incr <- seq(20, 40, len=length(time) %/% 2)
-L.decr <- seq(40, 20, len=length(time) - length(L.incr))
+L.incr <- seq(25, 50, len=length(time) %/% 2)
+L.decr <- seq(50, 25, len=length(time) - length(L.incr))
 L <- list(c(L.incr, L.decr))
 env1 <- replace(env1, "L", L)
 
-ss <- with(run_coral_ss(env=list(L=20, N=2e-7, X=2e-6), pars=defpars), last(S$S/H$H))
+ss <- with(run_coral_ss(env=list(L=25, N=1e-7, X=2e-6), pars=defpars), last(S$S/H$H))
 
 run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", ss))#
 
 png("img/Fig7.png", width=4, height=4, units="in", res=300)
-par(mar=c(2,2,1,1), mgp=c(1,0,0), tcl=0.25, cex.axis=0.5, cex.lab=0.75, xaxs="i")
+par(mar=c(2,2,1,1), mgp=c(1,0,0), tcl=0.25, cex.axis=0.5, cex.lab=0.75, xaxs="i", yaxs="i", mfrow=c(1,1))
 with(run1, {
-  plot(env$L, S$S/H$H, type="l", xlab="Light", ylab="S:H biomass", ylim=c(0,0.24))
-  for (i in seq(1300, length(time), last(time)/3.5)) {
+  plot(env$L, S$S/H$H, type="l", xlab="External light (molph/m2/d)", ylab="S:H biomass", ylim=c(0,0.2))
+  for (i in seq(1950, length(time), last(time)/2)) {
[…]

### 2016-11-01 R/def_pars.R

@@ -19,11 +19,11 @@ def_pars <- function() {
     yCL=0.1,  # L:C ratio in fixed carbon (=quantum yield) (molC/mol ph)
     kNPQ=112,  # capacity of non-photochemical quenching (mol ph/CmolS/d)
     # calculated as 4x max. photochemical quenching (Gorbunov et al. 2001)
-    kROS=60,  # amount of excess light beyond NPQ capacity (e.g., jeL-jNPQ) that doubles ROS production relative to baseline (mol ph/CmolS/d)
+    kROS=80,  # amount of excess light beyond NPQ capacity (e.g., jeL-jNPQ) that doubles ROS production relative to baseline (mol ph/CmolS/d)
     k=1,  # exponent on ROS production (-)
     astar=1.34,  # Symbiont specific cross-sectional area (m^2/C-molS)
     sigmaNS=0.9,  # Proportion of symbiont nitrogen turnover recylced (-)
-    sigmaCS=0.1,  # Proportion of symbiont carbon turnover recycled (-)
+    sigmaCS=0.9,  # Proportion of symbiont carbon turnover recycled (-)
     jCPm=2.8,  # Maximum specific photosynthate production rate (Cmol/CmolS/d)
     jSGm=0.25,  # Maximum specific symbiont growth rate (CmolS/CmolS/d)
     initS=0.1,  # Initial symbiont biomass (CmolS)

### 2016-11-01 JTB_ms/JTB_ms.Rmd
complete draft

@@
 ---
-title: A simplified dynamic bioenergetic model for coral-algal symbiosis and coral bleaching
+title: A simplified dynamic bioenergetic model for coral-*Symbiodinium* symbioses and coral bleaching as an alternate stable state
 author:
@@
   Coral reef ecosystems owe their ecological success--and vulnerability to climate change--to the symbiotic metabolism of corals and *Symbiodinium* spp.
-The urgency to understand and predict the stability and breakdown of these symbioses (i.e., coral 'bleaching') calls for the development and application of theoretical tools.
-While the biological complexity of the symbiosis is well-suited to dynamical modeling, the theoretical complexity of certain approaches makes them inaccessible to many biologists, thus limiting their potential application and utility.
-Here, we develop a simplified dynamic bioenergetic model of coral-algal symbiosis that demonstrates realistic steady-state patterns in coral growth and symbiont abundance across gradients of light, nutrients, and feeding.
-Furthermore, through a mechanistic treatment of photo-oxidative stress, the model displays dynamic responses of bleaching and recovery, and suggests that these phenomena represent transitions between alternative stable states.
-A suite of complex observed behaviors reproduced by the model suggests it captures many important attributes of the system; meanwhile, its modular framework and open source R code are designed to facilitate expansion and problem-specific applications.
-We see high utility for this modeling framework in generating testable hypotheses and predicting integrated, mechanistic responses of corals to environmental change.
-
-  Latest development--sigmaCH and sigmaCS changed to 0.1 meaning 10% of metabolic CO2 produced is passively available to photosynthesis, to reflect the fact that that ~10% of total internal DIC at pH 7 is in the form of CO2.
-probably actually less than this since internal pH may be closer to 7.5.
-This does not change any of the figures substantively except the hysteresis figure--overshoot no longer occurs with specific parameters run.
-may need to just change light range or rate of change to reproduce overshoot.
-
+The urgency to understand and predict the stability and breakdown of these symbioses (i.e., coral 'bleaching') demands the development and application of theoretical tools; however, the complexity of some approaches may limit their potential application and utility.
+Here, we develop a simplified dynamic bioenergetic model of coral-*Symbiodinium* symbioses that demonstrates realistic steady-state patterns in coral growth and symbiont abundance across gradients of light, nutrients, and feeding.
+Furthermore, through a mechanistic treatment of photo-oxidative stress, the model displays dynamic behaviors of bleaching and recovery, and suggests that these phenomena represent transitions between alternate stable states.
+A suite of complex responses to multiple, interacting environmental factors reproduced by the model suggests it unifyingly captures many important attributes of the system; meanwhile, its modular framework and open source R code are designed to facilitate expansion and problem-specific applications.
+We see significant potential applications for this modeling framework in generating testable hypotheses and predicting integrated, mechanistic responses of corals to environmental change.
 bibliography: library.bib
@@
   rticles::elsevier_article:
-    fig_caption: no
+    fig_caption: yes
 ---
@@
 However, the complex symbiotic metabolism of corals is vulnerable to disruption by numerous anthropogenic environmental perturbations, jeopardizing their future persistence.
-In order to understand and predict coral responses to complex changes in their environment, a mechanistic understanding of how multiple interacting factors drive the individual and emergent physiology of both symbiotic partners is necessary.
+In order to understand and predict responses of corals to complex changes in their environment, a mechanistic understanding of how multiple interacting factors drive the individual and emergent physiology of both symbiotic partners is necessary.
 Such a task is well suited for theoretical modeling frameworks such as Dynamic Energy Budget (DEB) theory [@Kooijman:2010vd], although the complexity of such theory makes these efforts inaccessible to many biologists [@Jager:2013bj].
@@
 1.
-_Develop a detailed module of environmental stress._ Of primary interest to coral biologists and ecologists is symbiosis dysfunction under environmental stress, resulting in coral “bleaching”--the loss of algal symbionts from the association [@Jokiel:1977p7353].
+_Develop a module of photooxidative stress._ Of primary interest to coral biologists and ecologists is symbiosis dysfunction under environmental stress, resulting in coral “bleaching”--the loss of algal symbionts from the association [@Jokiel:1977p7353].
 Photooxidative stress in *Symbiodinium* is considered a primary trigger of bleaching in response to high temperature and/or light [@Weis:2008p944], and prolonged or severe bleaching can result in mortality, though corals sometimes recover their symbionts.
-Bleaching susceptibility, severity, and recovery may by influenced by interacting factors such as heterotrophy and nutrient availability [@Wooldridge:2014hc], and the genetic identity of *Symbiodinium* [@Glynn:2001p7571].
-To simulate these bleaching-related phenomena, we develop a generalized framework linking overreduction of the photosynthetic light reactions to downstream impacts of photoinhibition and photodamage.
+Bleaching susceptibility, severity, and recovery may be influenced by interacting factors such as heterotrophy and nutrient availability [@Wooldridge:2014hc].
+To simulate these events and interactions, we develop a photooxidative stress module linking overreduction of the photosynthetic light reactions to downstream impacts of photoinhibition and photodamage.
 
 2.
-_Reduce theoretical and mathematical complexity._ Following the logic of @Jager:2013bj, we exclude certain features of formal DEB models in order to capture behaviors of interest with the simplest possible formulation.
+_Reduce theoretical and mathematical complexity._ Following the logic of @Jager:2013bj, we exclude certain features of formal DEB theory in order to capture behaviors of interest with the simplest possible formulation.
 Here, we present a model without reserves, maturity, or reproduction [see @Kooijman:2010vd].
-This formulation restricts the model's scope to the bioenergetics of growth and symbiosis dynamics in adult corals (i.e., reproduction, larval stages, and metamorphosis are not considered), but greatly reduces theoretical complexity and parameter numbers, which is advantageous given the relative paucity of data for corals.
+This formulation restricts the model's scope to the bioenergetics of adult corals (i.e., reproduction, larval stages, and metamorphosis are not considered), but greatly reduces theoretical complexity and parameter numbers, which is advantageous given the relative paucity of data for corals.
 However, our primary motivation for reducing complexity was to increase accessibility and applicability for biologists and ecologists without requiring significant expertise in DEB theory.
@@
 With an accessible and modular framework, we envision this model as a resource for futher development by the scientific community to include additional complexity and problem-specific components.
-To widen the audience for this work, we chose the R language because it is freely available and in common use by biologists and ecologists.
+We chose the R language, which is freely available and in common use by biologists and ecologists, to reach the widest possible audience with this work.
 
-With these as our primary motivations, we describe a simplified approach to dynamic bioenergetic modeling of coral-algal symbioses that tracks carbon and nitrogen acquisition and sharing between partners.
-This theoretical framework dynamically integrates the influence of external irradiance, nutrients, and prey availability on coral growth and symbiosis dynamics (i.e., symbiont:host biomass ratios), allowing for the possibility of coral bleaching in the event of photooxidative stress.
-In the following sections, we describe the formulation of this model and justify its structure and parameter values based on relevant literature.
-We then demonstrate the model's behavior and discuss some of its major implications and outcomes, and the wide range of potential applications for this model in the study of cnidarian-algal symbioses.
+With these as our primary motivations, we describe a simplified approach to dynamic bioenergetic modeling of coral-*Symbiodinium* symbioses that dynamically integrates the influence of external irradiance, nutrients, and prey availability on coral growth and symbiosis dynamics (i.e., symbiont:host biomass ratios), allowing for the possibility of coral bleaching in the event of photooxidative stress.
+In the following sections, we describe and provide rationale for the model structure, and demonstrate a range of steady state and dynamic behaviors that are consistent with observed phenomena.
 
 # Model description
-In this dynamic bioenergetic model of coral-algal symbiosis, carbon and nitrogen are acquired by each partner and used to construct biomass.
+In this dynamical system, both partners acquire and use carbon and nitrogen to construct biomass: the symbiont fixes carbon through photosynthesis, and receives nitrogen shared by the host, while the host acquires nitrogen from the environment and receives carbon shared by the symbiont.
 A graphical representation of the model is presented in Fig.
-1, and each model flux and parameter are defined in Tables 1 and 2, respectively.
-We use C-moles as the unit of biomass for consistency with the rigorous mass balance of DEB theory: 1 C-mole is equivalent to the amount of biomass containing 1 mole of Carbon atoms.
+1, and each model flux and parameter is defined in Tables 1 and 2, respectively.
+We use C-moles as the unit of biomass for consistency with the rigorous mass balance of DEB theory: 1 C-mole is equivalent to the amount of biomass containing 1 mole of carbon atoms.
 Host biomass ($H$), symbiont biomass ($S$), and prey biomass ($X$) have fixed, but different, molar N:C ratios (Table 2).
-Carbon and nitrogen are combined to produce biomass by synthesizing units (SU), which are mathematical specifications of the formation of a product from two substrates; we use the parallel complementary formulation of @Kooijman:2010vd to specify these fluxes.
-The two state variables are symbiont biomass and coral biomass; because resources are acquired proportionally to surface area, and surface area is proportional to volume (i.e., corals are "V1-morphs" in DEB terminology [@Kooijman:2010vd]), biomass increases exponentially during growth (indeed, corals grow exponentially [@Bak:1976bv]).
+Biomass is produced from carbon and nitrogen by synthesizing units (SU), which are mathematical specifications of the formation of a product from two substrates; we use the parallel complementary formulation of @Kooijman:2010vd to specify these fluxes.
+The two state variables of this system are symbiont biomass and coral biomass; because resources are acquired proportionally to surface area, and surface area is proportional to volume (i.e., corals are "V1-morphs" in DEB terminology [@Kooijman:2010vd]), biomass increases exponentially during growth (indeed, corals grow exponentially [@Bak:1976bv]).
 
-Environmental stress is implemented in the form of photo-oxidative stress, which is thought to be the primary trigger of coral bleaching [@Lesser:1997p3587; @Weis:2008p944; @Wooldridge:2009p7807].
+Environmental stress is implemented in the form of photooxidative stress, which is thought to be the primary trigger of coral bleaching [@Lesser:1997p3587; @Weis:2008p944; @Wooldridge:2009p7807].
 To simulate these events, we model the absorption and quenching of light energy by photochemistry and non-photochemical quenching, and the phenomenological responses that occur when these capacities are overwhelmed (i.e., photoinhibition, photodamage, and symbiont loss).
-While bleaching in response to high light alone has been observed experimentally [@Schutter:2011ie; @Downs:2013kc], mass coral bleaching events occur in response to high temperature [@HoeghGuldberg:1999p4222]; thus, it is important to justify our consideration of light as the primary stressor in the model.
+While bleaching in response to high light alone has been observed experimentally [@Schutter:2011ie; @Downs:2013kc], mass coral bleaching events occur in response to high temperature [@HoeghGuldberg:1999p4222]; thus, it is important to justify our consideration of light as the primary stressor.
 In reality, light and temperature interact synergistically [@Coles:1978p1124; @Jones:1998p3572], and in fact, any stressor that disrupts the quenching of light energy may lead to bleaching [@Wooldridge:2010p7809; @Baker:2015kp].
@@
 In this way, elevated temperature serves to reduce the threshold above which light stresses the system [@HoeghGuldberg:1999p4222]; importantly, light is still the proximate stressor.
-Therefore, we omitted temperature from the model to maintain a desired level of simplicity, while still allowing photophysiological stress and bleaching to be simulated with biological realism in response to light.
+Therefore, we omitted temperature from the model to maintain a desired level of simplicity, while still allowing photooxidative stress and bleaching to be simulated with biological realism in response to light.
 
@@
 
-Table 1. Model fluxes
+Table 1. Model fluxes (mass-specific)
 
@@
 ## State equations
-The balance equations for symbiont and host biomass are expressed as:
+The balance equations for symbiont ($S$) and host ($H$) biomass are expressed as:
 
@@
 These are expressed per unit of symbiont and host biomass, respectively.
-The specific biomass growth and turnover rates that define these balance equations are produced by combinations of the individual model fluxes (see Table 1 for definitions and units), which are each expressed as mass-specific rates (e.g., per C-mole of symbiont or host biomass).
+The specific biomass growth and turnover rates that define these balance equations are produced by combinations of the individual model fluxes (see Table 1 for definitions and units), which are each expressed as mass-specific rates (e.g., per C-mole of symbiont or host biomass per day).
 When necessary, conversions between symbiont-mass-specific and host-mass-specific rates are accomplished by multiplying or dividing by the symbiont:host biomass ratio.
@@
 
-where $\rho_C$ is the surplus fixed carbon shared by the symbiont (see below), and $r_{NH}$ is the recycled portion of the nitrogen liberated by host biomass turnover:
+where $\rho_C$ is fixed carbon shared by the symbiont (see Eq. XX), and $r_{NH}$ is recycled nitrogen liberated by host biomass turnover:
 
@@
 
-The carbon rejected from this SU reflects the amount of excess fixed carbon available to the host that is not used in biomass formation:
+Due to the inherent inefficiency of the parallel complementary SU formulation, there is always some nitrogen shared with the symbiont even when coral biomass formation is strongly nitrogen-limited.
+Likewise, there is always a non-zero rejection flux of excess carbon from the coral biomass SU.
+Likewise, the carbon rejected from this SU reflects the amount of excess fixed carbon available to the host that is not used in biomass formation:
 
@@
 
-This flux, $j_{eC}$, is assumed to be available as a respiratory substrate to support various energetically-demanding processes; of particular importance is the host's active supply of CO~2~ for symbiont photosynthesis [@Wooldridge:2013tj; @Hopkinson:2015ev] (see below).
+This flux, $j_{eC}$, is assumed to be available to the host as a respiratory substrate to support various energetically-demanding processes; of particular importance is the host's active carbon concentrating mechanisms (CCMs) that supply CO~2~ for symbiont photosynthesis [@Wooldridge:2013tj; @Hopkinson:2015ev].
+We therefore specify $j_{CO_2}$ as the host-mediated delivery of CO~2~ to photosynthesis that encompasses potentially diverse carbon-concentrating mechanisms (CCMs), including active transport of bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to CO~2~ to promote diffusion toward the symbiont [@Tansik:2015cm], and acidification of the symbiosome to increase localized CO~2~ concentrations around the symbiont [@Barott:2014gx].
+Since these active CCMs require energetic input by the host, we define $j_{CO_2}$ as proportional to $j_{eC}$, assuming that some of this carbon is respired to energize the CCMs.
+This formulation means that the symbiont indirectly ensures its own CO~2~ supply by providing fixed carbon (=energy) to the host [@Wooldridge:2013tj], which establishes a positive feedback in the model that leads to the existence of multiple stable states, which we discuss below in the context of coral bleaching.
+The parameter $k_{CO_2}$ scales the efficacy of host CCMs, which enables the comparison of different rates of CO~2~ delivery that may characterize different coral species [@Wooldridge:2014di].
+The active input of CO~2~ to the photosynthesis SU is therefore specified as:
 
-Due to the inherent inefficiency of the parallel complementary SU formulation, there is always some nitrogen shared with the symbiont even when coral biomass formation is strongly nitrogen-limited.
-Likewise, there is always a non-zero rejection flux of excess carbon from the coral biomass SU.
+\begin{equation} j_{CO_2} = k_{CO_2}j_{eC} \end{equation}
 
@@
 
-We then specify the input of CO~2~ to photosynthesis as $j_{CO_2}$, a flux mediated by the host that encompasses potentially diverse carbon-concentrating mechanisms (CCMs), including active transport of bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to CO~2~ to promote diffusion toward the symbiont [@Tansik:2015cm], and acidification of the symbiosome to increase localized CO~2~ concentrations around the symbiont [@Barott:2014gx].
-Since these active CCMs require energetic input by the host, the magnitude of CO~2~ delivery is proportional to the amount of fixed carbon available to the host after satisfying its growth demands, $j_{eC}$, assuming that some of this carbon is respired to energize these CCMs.
-This formulation means that the symbiont indirectly ensures its own CO~2~ supply by providing fixed carbon (=energy) to the host [@Wooldridge:2013tj], which establishes a positive feedback in the model that leads to the existence of multiple stable states, which we discuss below in the context of coral bleaching.
-The parameter $k_{CO_2}$ scales the efficacy of host CCMs, which enables the comparison of different rates of CO~2~ delivery that may characterize different coral species [@Wooldridge:2014di].
-The active input of CO~2~ to the photosynthesis SU is therefore specified as:
-
-\begin{equation} j_{CO_2} = k_{CO_2}j_{eC} \end{equation}
-
-An additional fixed proportion of metabolic CO~2~ generated by host biomass turnover is passively made available to the photosynthesis SU, according to:
+CO~2~ arrives at the photosynthesis SU from multiple sources: in addition to the CO~2~ actively supplied by the host through its CCMs ($j_{CO_2}$; Eq.
+XX), a fixed proportion of metabolic CO~2~ generated by host biomass turnover is passively available to the photosynthesis SU, according to:
 
@@
 
-along with a fixed proportion of symbiont biomass turnover:
+along with a fixed proportion of CO~2~ generated by symbiont biomass turnover:
 
@@
 
-where $j_{CPm}$ is the maximum specific rate of photosynthesis, and $c_{ROS}$ is the relative rate of reactive oxygen species production (see below).
-Dividing the photosynthetic rate by $c_{ROS}$ causes a decline in response to photooxidative stress at high light levels, and the emergent outcome of this SU formulation demonstrates a classic photoinhibition response (see Fig.
-S1).
+where $j_{CPm}$ is the maximum specific rate of photosynthesis, and $c_{ROS}$ is the relative rate of reactive oxygen species production (see Eq.
+XX).
+Dividing the photosynthetic rate by $c_{ROS}$ causes a decline in response to photooxidative stress at high light levels, and the emergent outcome of this SU formulation demonstrates a classic photoinhibition response (Fig.
+S2).
 
-Light energy absorbed in excess of what is used to fix carbon is specified by the SU "rejection flux", according to:
+Light energy absorbed in excess of what is used to fix carbon is specified by the SU rejection flux, according to:
 
@@
 This excess light energy must be quenched by alternative pathways in order to prevent photooxidative damage [@Powles:1984tm].
-*Symbiodinium* utilize a variety of pathways for non-photochemical quenching [NPQ; @Roth:2014wf], which we collect in a relative NPQ capacity specified as a parameter of the symbiont ($k_{NPQ}$).
-The total NPQ flux is specified as a single-substrate SU formula:
+*Symbiodinium* utilize a variety of pathways for non-photochemical quenching [NPQ; @Roth:2014wf], which we collect in a total NPQ capacity specified as a parameter of the symbiont ($k_{NPQ}$).
+The NPQ flux $j_{NPQ}$ is then specified as a single-substrate SU formula with a maximum of $k_{NPQ}$:
 
@@
 If light energy further exceeds the capacity of both photochemistry and NPQ, then reactive oxygen species (ROS) are produced.
-We represent this as a relative quantity $c_{ROS}$, taking a value of 1 when all light energy is quenched by photochemistry and NPQ, and increasing as the amount of excess excitation energy increases:
+We represent this as a relative quantity $c_{ROS}$, which takes a value of 1 when all light energy is quenched by photochemistry and NPQ, and increases as the amount of excess excitation energy increases, specified as:
 
@@
 where $k_{ROS}$ is a parameter of the symbiont that determines the rate of ROS production (specifically, the amount of excess excitation energy that doubles ROS production relative to baseline levels).
-Importantly, $c_{ROS}$ is specified here not as a function of absolute external light, but rather the amount of excess light, $j_{eL}$, after accounting for quenching by carbon fixation and NPQ.
+Importantly, $c_{ROS}$ is specified here not as a function of absolute external light, but rather the amount of excess light after accounting for quenching by carbon fixation and NPQ.
 A direct consequence of this formulation is that CO~2~-limitation of photosynthesis can lead to ROS production, an important mechanism [@Wooldridge:2009p7807] that was not captured by previous representations of photooxidative stress [@Eynaud:2011tv].
 
-Carbon fixed by photosynthesis ($j_{CP}$) is then combined with nitrogen shared by the host ($\rho_N$) and nitrogen recycled from symbiont biomass turnover
+Carbon fixed by photosynthesis ($j_{CP}$; Eq. XX) is then combined with nitrogen shared by the host ($\rho_N$; Eq. XX) and nitrogen recycled from symbiont biomass turnover
 
@@
 The parameter $b$ is included to scale biomass loss due to bleaching in response to ROS.
-(Note that recycling of symbiont biomass turnover ($rNS$ and $rCS$) only occurs based on the maintenance component of turnover (i.e., $j_{ST}^0$), and not the bleaching-related component, as this loss represents biomass that is damaged or expelled from the host.)
+(Note that recycling of symbiont biomass turnover ($rNS$ and $rCS$) only occurs based on the maintenance component of turnover (i.e., $j_{ST}^0$), and not the photodamage/bleaching component, as this loss represents biomass that is damaged or expelled from the host.)
 
@@
 In addition, the photosynthesis SU receives CO~2~ proportional to the carbon rejection flux from the host biomass SU, and the rejection flux of excitation energy from the photosynthesis SU acts to reduce its own production (i.e., photoinhibition).
-A vector of time steps is used for each simulation, along which dynamic environmental forcing functions (irradiance, DIN, and prey abundance) can be designed.
+A vector of time steps is used for each simulation, along which dynamic environmental forcing functions (irradiance, DIN, and prey abundance).
 These vectors, along with initial values of symbiont and host biomass, then serve as input to the time-stepping function, which solves for the current system state using values of the previous system state where necessary.
 A time step of 0.1 days was used for all simulations, which were performed using R code that is available in the data repository accompanying this article: [github.com/jrcunning/Rcoral](github.com/jrcunning/Rcoral).
+For steady state estimations, simulations were run until the change in both growth and S:H biomass was < 5e-7 per time step.
 
@@
 However, under some conditions, either of these outcomes may occur depending on initial values of symbiont and host biomass, indicating the presence of alternate stable states.
-The mechanism that produces these alternate stable states is the positive feedback between carbon-limitation of the host and CO~2~-limitation of photosynthesis: if symbiont biomass is initially very low (i.e., a "bleached" coral), the system cannot escape positively reinforcing carbon-limitation and cannot grow.
-However, if symbiont biomass is initially high (i.e., a "healthy" coral), then the system maintains positive growth in a nitrogen-limited state.
+The mechanism that produces these alternate stable states is the positive feedback between carbon-limitation of the host and CO~2~-limitation of photosynthesis: if symbiont biomass is initially very low (i.e., a "bleached" coral), the system cannot escape this positive feedback and cannot grow.
+However, if symbiont biomass is initially high (i.e., a "healthy" coral), then the system remains in a nitrogen-limited state with positive growth.
 For practical purposes, this section of the manuscript considers only positive growth steady states under constant environments; the subsequent "dynamic behavior" section explores how changing environments may cause the system to switch between alternate stable states, which we interpret in the context of coral bleaching.
 
-To analyze qualitative patterns in steady state behavior, we ran the model to steady state across gradients of external irradiance and nutrients (Fig.
+To analyze steady state behavior, we ran the model to steady state across gradients of external irradiance and nutrients (Fig.
 2), which revealed patterns consistent with observed phenomena in corals.
@@
 In agreement with this trend are many observations of negative correlation between irradiance and symbiont density [@Stimson:1997p3837; @Brown:1999p3534; @Fitt:2000p3792; @Titlyanov:2001p3631].
-As higher light alleviates light-limitation of photosynthesis, growth becomes less carbon-limited.
+As higher light alleviates light-limitation of photosynthesis, host growth becomes less carbon-limited.
 
@@
 
-Similarly, increasing DIN up to ~2 µM alleviates nitrogen-limitation (Fig.
+Similarly, increasing DIN alleviates nitrogen-limitation (Fig.
 2A).
@@
 2B), a phenomenon also observed in reef corals [@Marubini:1996p8460].
-At low DIN and intermediate light, more typical of coral reef environments, symbiont to host biomass ratios are around ~0.1-0.3, which is consistent with values reported in the literature [@Muscatine:1981jy; @Edmunds:2011bv; @Hawkins:2016jj].
+At low DIN and intermediate light, more typical of coral reef environments, symbiont to host biomass ratios are around ~0.06-0.21, which is consistent with values reported in the literature [@Muscatine:1981jy; @Edmunds:2011bv; @Hawkins:2016jj].
 
-The maximum predicted growth rates of ~0.1 d^-1^, occurring between ~10-20 mol photons m^-2^ s^-1^ light and ~4 µM DIN (Fig.
+The maximum predicted growth rates of ~0.1 d^-1^, occurring between ~10-25 mol photons m^-2^ s^-1^ light and ~4 µM DIN (Fig.
 2A), are comparable to the rate of `r Tanaka_gr` d^-1^ measured by @Tanaka:2007uv under similar N-enriched conditions.
@@
 
-As irradiance continues to increase above ~20 mol photons m^-2^ d^-1^, growth rates decline until positive growth ceases above ~35 µmol photons m^-2^ d^-1^ (Fig.
+As irradiance continues to increase above ~25 mol photons m^-2^ d^-1^, growth rates decline until positive growth ceases above ~40 µmol photons m^-2^ d^-1^ (Fig.
 2A).
-The mechanism underlying this decline is the increase in light energy beyond the capacity of photosynthesis and non-photochemical quenching: excess excitation energy generates reactive oxygen species (ROS) [@Weis:2008p944; @Roth:2014wf], which, in this model, have the phenomenological consequences of reducing the photosynthetic rate (representing photoinhibition) and increasing symbiont biomass loss (representing photodamage and/or symbiont expulsion) [see @Eynaud:2011tv].
+The mechanism underlying this decline is the increase in light energy beyond the capacities of photosynthesis and non-photochemical quenching: excess excitation energy generates reactive oxygen species (ROS) [@Weis:2008p944; @Roth:2014wf], which, in this model, have the phenomenological consequences of reducing the photosynthetic rate (representing photoinhibition) and increasing symbiont biomass loss (representing photodamage and/or symbiont expulsion) [see @Eynaud:2011tv].
 Together, these impacts reduce the symbiont to host biomass ratio (Fig.
@@
 
-The incorporation of light stress in the model sets an upper limit to the amount of light at which a stable symbiotic interaction can be maintained, but even below this threshold of breakdown, negative effects of high light reduce steady state growth and symbiont:host biomass (Fig.
+The incorporation of photooxidative stress in the model sets an upper limit to the amount of light at which a stable symbiotic interaction can be maintained, but even below this threshold of breakdown, negative effects of high light reduce steady state growth and symbiont:host biomass (Fig.
 2).
@@
 By incorporating these impacts of light stress, the model predicts greater, and more realistic, variation in state variables across light gradients than was predicted by the model of @Muller:2009io, which did not include photoinhibition and photodamage.
-
-It is important to recognize that the boundary set by photo-oxidative stress on a stable symbiosis under steady state conditions (Fig.
+It is important to recognize that the boundary set by photooxidative stress on a stable symbiosis under steady state conditions (Fig.
 2) may be temporarily crossed by a dynamic system, which may experience a period of symbiont loss (bleaching) and reduced growth, after which a return to benign conditions may restore symbiont biomass and positive growth.
@@
 This sensitivity analysis was conducted at conditions typical for coral reef environments: low DIN (1e-7 M) and intermediate light (15 mol m^-2^ d^-2^).
-Sensitivity analyses conducted other environmental conditions are presented in the Supplementary Information.](../img/Fig3.png)
+Sensitivity analyses conducted other environmental conditions are presented in Figs S3-S7.](../img/Fig3.png)
 
@@
 3) since this parameter determines the amount of light absorbed by symbionts, with lower values increasing light-limitation.
-Altering maximum growth and turnover rates of symbiont and host also have expected impacts on steady states (Fig.
-3).
+Increasing the maximum growth and turnover rates have the expected effects of increasing and decreasing growth, respectively.
 Parameters relating to photooxidative stress and bleaching have little impact under these environmental conditions (Fig.
-3), but of course have larger impacts under higher light (e.g., Fig.
+3), but have larger impacts under higher light (e.g., Fig.
 S6).
-Sensitivity analyses conducted under different combinations of external light and nutrients are presented in the Supplementary Information (Figs.
-S3-S7)
+Sensitivity analyses conducted under different combinations of external light and nutrients are presented in Figs.
+S3-S7.
 
 # Dynamic behavior
-The dynamic behavior of the model demonstrates its power to integrate multiple environmental forcings simultaneously to produced integrated responses of the system.
-The potential applications of this are wide-ranging, and so the model may be highly useful for generating hypotheses about the way the system will respond to complex sets of environmental dynamics.
-Here we present several scenarios to highlight the power of the model and demonstrate that it can also reproduce more complex phenomena that have been observed in corals.
+The dynamic behavior of the model demonstrates its power to integrate multiple environmental forcings simultaneously.
+Here we present several scenarios that demonstrate the model's ability to reproduce complex phenomena that have been observed in corals.
 
@@
 Symbiont densities and coral growth rates are known to vary seasonally, representing an integrated response to changes in a suite of environmental factors.
-Light in particular is a strong driver of these trends [Stimson:1997p3837; @Brown:1999p3534; @Fagoonee:1999p4136; @Fitt:2000p3792], with high light associated with lower symbiont abundance and reduced tissue biomass.
+Light in particular is a strong driver of these trends [@Stimson:1997p3837; @Brown:1999p3534; @Fagoonee:1999p4136; @Fitt:2000p3792], with high light associated with lower symbiont abundance and reduced tissue biomass.
 The role of light in driving seasonal changes in symbiont density was demonstrated nicely by @Stimson:1997p3837, who also found that experimental nutrient-enrichment amplified the light-driven seasonal oscillation.
 Using the levels of light and nutrients from this study as inputs, the model reproduces this observed interaction among environmental factors (Fig.
-4), and also provides the mechanism: under high light in summer, photoinhibition reduces symbiont biomass and leads to carbon-limitation of host growth regardless of nutrient status; when light is reduced in winter, these responses become constrained instead by nitrogen, and therefore increase more when DIN is enriched (Fig.
+4), and also provides the mechanism: under high light in summer, photooxidative stress decreases symbiont biomass and increases carbon-limitation of host growth regardless of nutrient status; when light is reduced in winter, growth and symbiont biomass become constrained instead by nitrogen, and therefore increase more when DIN is enriched (Fig.
 S8).
@@
 Light-driven seasonal dynamics of symbiont abundance and coral growth.
-Light input (upper panel) was designed as a sinusoidal curve with a period of one year, with maximum and minimum values of 40 and 20 mol photons m^-2^ d^-1^, based on @Stimson:1997p3837.
-The dynamic behavior of symiont to host biomass ratio (middle panel) and the specific growth rate of host biomass (lower panel) show seasonal oscillations that are greater in magnitude under higher nutrients (1e-6 molN L^-1^; dashed lines) relative to lower nutrients (1e-7 molN L^-1^; solid lines), consistent with the findings of @Stimson:1997p3837.
+Light input (upper panel) was designed as a sinusoidal curve with a period of one year, with maximum and minimum values of 44 and 22 mol photons m^-2^ d^-1^, based on @Stimson:1997p3837.
+The dynamic behavior of symiont to host biomass ratio (middle panel) and the specific growth rate of host biomass (lower panel) show seasonal oscillations that are greater in magnitude under higher nutrients (1e-5 molN L^-1^; dashed lines) relative to lower nutrients (1e-7 molN L^-1^; solid lines), consistent with the findings of @Stimson:1997p3837.
 Prey density was set at 1e-5 CmolX L^-1^.](../img/Fig4.png)
@@
 Nevertheless, the seasonal variability predicted here (Fig.
-4) is consistent with experimental and field observations for corals, and demonstrates the power of this model to predict dynamic behavior that mechanistically integrate the impacts of multiple environmental drivers.
+4) is consistent with experimental and field observations for corals, and demonstrates the power of this model to predict dynamic behavior that mechanistically integrates multiple environmental drivers.
 
@@
 Coral bleaching is the stress-induced loss of symbiotic algae from coral tissues, which can occur in response to a variety of environmental stressors.
-In most cases, coral bleaching is thought to begin with photo-oxidative stress in symbiont photosynthesis [@Lesser:1997p3587], which triggers a cascade of events leading to symbiont expulsion [@Weis:2008p944].
-As symbionts are expelled, the host receives less fixed carbon, which may then compromise its ability to deliver CO~2~ to photosynthesis [@Wooldridge:2013tj].
-Increasing CO~2~-limitation for remaining symbionts, along with an amplified internal light environment due to reduced self-shading [@Enriquez:2005p142], may further exacerbate photo-oxidative stress and accelerate symbiont expulsion, driving the coral into a bleached state.
+In most cases, coral bleaching is thought to begin with photooxidative stress in symbiont photosynthesis [@Lesser:1997p3587], which triggers a cascade of events leading to symbiont expulsion [@Weis:2008p944].
+As symbionts are expelled, the host receives less fixed carbon, which may then compromise its ability to activate CCMs that deliver CO~2~ to photosynthesis [@Wooldridge:2013tj].
+Increasing CO~2~-limitation for remaining symbionts, along with an amplified internal light environment due to reduced self-shading [@Enriquez:2005p142], may further exacerbate photooxidative stress and accelerate symbiont expulsion, driving the coral into a bleached state.
 
@@
 The 'healthy' state is characterized by nitrogen-limitation of both symbiont and host: under these conditions, the symbiont translocates sufficient carbon to support host growth and CCMs, which ensures that photosynthesis does not become CO~2~-limited.
-However, if carbon translocation is disrupted (and light is sufficiently high), then the system is driven into the 'bleached' state by photo-oxidative stress and positively reinforcing carbon- and CO~2~-limitation.
-In this context, coral bleaching can be understood as a transition from one stable state to another, and bleaching thresholds are sets of environmental conditions that push a healthy-state coral onto a trajectory leading to a bleached stable state.
+However, if carbon translocation is disrupted (and light is sufficiently high), then the system is driven into the 'bleached' state by photooxidative stress and positively reinforcing carbon- and CO~2~-limitation.
+In this context, coral bleaching can be understood as a transition from one stable state to another, and bleaching thresholds are sets of environmental conditions that push a healthy-state coral onto a trajectory leading to a bleached state.
 
 We are highly interested in the conditions under which the system switches from a healthy to a bleached state, and can use this model as a tool to explore this dynamic behavior.
-Most straightforwardly, this switch occurs when irradiance exceeds the capacity of photochemistry and NPQ, causing the production of ROS that leads to photoinhibition, photodamage, and symbiont expulsion, thereby engaging the positive feedbacks that lead to a bleached state (Fig.
-5).
-However, the positive feedbacks involved in bleaching are not engaged only in response to high external irradiance; in fact, they depend on the relative balance of light energy absorption and quenching, which in turn depends on the availability of CO~2~ for photosynthesis.
-While previous models framed photo-oxidative stress as a fixed response to absolute external irradiance [@Eynaud:2011tv], our implementation considers the dynamic balance of multiple energy sinks in the causation of stress, which is more consistent with current understanding of symbiosis dysfunction [@Wooldridge:2013tj], and establishes a critical role of host CCMs in providing CO~2~ for photosynthesis [@Tansik:2015cm; @Hopkinson:2015ev].
+Most straightforwardly, this switch occurs when increasing external irradiance (Fig.
+5A) causes sufficient ROS production (Fig.
+5B) and photoinhibition (Fig.
+5C) such that the positive feedbacks between host carbon-limitation (Fig.
+5D) and CO~2~-limitation of photosynthesis (Fig.
+5E) are rapidly engaged, leading to even greater photooxidative stress and a rapid decline in S:H biomass (Fig.
+5F), characteristic of coral bleaching.
+However, the positive feedbacks involved in bleaching are not engaged only in response to high external irradiance alone; in fact, they depend on the relative balance of light energy absorption and quenching, which in turn depends on the availability of CO~2~ for photosynthesis.
+While previous models framed photooxidative stress as a fixed response to absolute external irradiance [@Eynaud:2011tv], our implementation considers the dynamic balance of multiple energy sinks in the causation of stress, which is more consistent with current understanding of symbiosis dysfunction [@Wooldridge:2013tj], and establishes a critical role of host CCMs in providing CO~2~ for photosynthesis [@Tansik:2015cm; @Hopkinson:2015ev].
 
@@
 Bleaching with interactive factors.
-Simulations of high light stress (sinusoid with maximum=39 mol ph m^-2^ d^-1^) under default environmental conditions (solid line; DIN=1e-7 molN/L; prey=3e-6 CmolX/L), or with elevated feeding (dashed line; prey=6e-6 CmolX/L), or elevated nutrients (dotted line; DIN=1e-6 molN/L).
+Simulations of high light stress (sinusoid with maximum=48 mol ph m^-2^ d^-1^) under default environmental conditions (solid line; DIN=1e-7 molN/L; prey=3e-6 CmolX/L), or with elevated feeding (dashed line; prey=6e-6 CmolX/L), or elevated nutrients (dotted line; DIN=1e-6 molN/L).
 Initial symbiont biomass was set to the steady state for each set of starting conditions, respectively.
@@
 The direction of the trajectory is indicated with arrowheads evenly spaced with respect to time.
-The rate of change in light was slow (0.005 mol ph d^-1^) such that dynamics track closely to steady state values.
-(All parameters at default values; DIN=2e-7 molN/L; Prey=2e-6 CmolX/L).](../img/Fig7.png)
+The rate of change in light was slow (0.01 mol ph d^-1^) such that dynamics track closely to steady state values.
+(All parameters at default values; DIN=1e-7 molN/L; Prey=2e-6 CmolX/L).](../img/Fig7.png)
 
 The hysteresis demonstrated in Fig.
-7 reveals another interesting behavior of the system: during recovery from bleaching, the S:H biomass ratio reaches a higher value that 'overshoots' the level maintained in the healthy stable state.
+7 reveals another interesting behavior of the system: during recovery from bleaching, the S:H biomass ratio reaches a higher value that 'overshoots' the ratio maintained in the healthy stable state.
 In fact, such a phenomenon has been observed in both experimental [@Cunning:2015ja] and field studies [@Kemp:2014ug] documenting recovery from bleaching, and has been interpreted as a potential 'disequilibrium in host-symbiont regulation' [@Kemp:2014ug].
 The model explains this 'overshoot' as an approach to a bifurcation point in the dynamical system: during recovery from the 'bleached' state, host growth is more strongly carbon-limited than symbiont growth (since the symbiont has priority access to carbon), so S:H biomass increases.
-S:H biomass will continue to increase beyond the alternative steady state levels as long as the symbiont remains carbon-limited; not until nitrogen becomes more strongly limiting to symbiont growth (hence forcing the symbiont to share fixed carbon with the host) will the system 'flip' back to a state of high carbon fixation and overall nitrogen-limitation that characterize a 'healthy' symbiosis.
-Of course, the magnitude of this 'overshoot' will depend on external nutrient concentrations and the rate of change in the stressor during recovery, though simulations predict similar magnitudes to what was observed by Kemp et al.
-(supp fig?).
+S:H biomass will continue to increase beyond the alternative steady state levels as long as the symbiont remains carbon-limited; not until nitrogen becomes more strongly limiting to symbiont growth than carbon (hence forcing the symbiont to share carbon with the host) will the system flip back to the 'healthy' state, characterized by high carbon fixation rates and overall nitrogen-limitation.
+Of course, the magnitude of this 'overshoot' depends on external nutrient concentrations and the rate of change in the stressor during recovery, though simulations predict similar magnitudes to what was observed by @Kemp:2014ug and @Cunning:2015ja (Fig.
+S9).
 
 # Conclusions
-The dynamic behavior of the model demonstrates its ability to mechanistically integrate multiple forcing functions, and the degree to which the system state depends not only on external forcing, but also the past history of the system.
-The model represents a unifying mechanistic framework that reproduces many phenomena observed in corals, suggesting it has captured many of the important features of the system in a relatively simple formulation.
-Furthermore, it allows the model to be used to predict responses of the system to multiple dynamic environmental forcing functions in a way that incorporates all positive and negative feedbacks.
-This makes the model potentially highly useful as a tool for generating hypotheses.
-make N, L, and X go up and down.
-For example, the model demonstrates and provides a mechanistic explanation for: an increase in symbiont abundance with nutrients (REF), an increase in symbiont abundance with heterotrophic feeding (refs), decrease in symbiont abundance with light (ref), optimal growth at intermediate light, rapid 'bleaching' above threshold of high light (ref) along with positive feedbacks (wooldridge, enriquez), mitigation of bleaching by heterotrophic feeding, exacerbation of bleaching by elevated nutrients, higher standing stocks of less productive symbionts, greater seasonal change in symbiont density with elevated nutrients, overshoot of symbiont density after recovery from bleaching.
+This simplified dynamic bioenergetic model of coral-*Symbiodinium* symbioses mechanistically reproduces patterns in steady-state coral growth and symbiont abundance commonly observed in corals, including higher symbiont abundance with higher nutrients and feeding, lower symbiont abundance with increasing light, and optimal growth at intermediate light levels.
+Moreover, the model reproduces complex dynamic behaviors including seasonal changes in symbiont density at different nutrient levels, rapid bleaching above a threshold of high light, mitigation of bleaching by heterotrophic feeding, exacerbation of bleaching by elevated nutrients, and an overshoot of symbiont density during recovery from bleaching.
+These examples demonstrate the model's ability to integrate multiple environmental forcing functions to reproduce complex responses; meanwhile, the diversity of these phenomena suggest the model has captured many of the important features of the system in a unifying mechanistic framework.
+This model also provides a new conceptual framework for considering coral bleaching as a transition to an alternate stable state, which has important implications for understanding the performance and maintenance of symbiotic interactions.
+In this context, the 'healthy' stable state represents a scenario in which the host is 'in control' of the system, with nitrogen-limitation stabilizing the symbiont to host biomass ratio and maintaining positive growth.
+Conversely, carbon-limitation represents a loss of host 'control', wherein positive feedbacks result in a loss of capacity for photosynthetic carbon fixation and positive growth.
+Importantly, these 'control' mechanisms arise entirely passively through feedbacks present in the system.
 
-"host control"=>N-limitation.  C-limitation=host has lost control of symbiosis: either bleaches or symbiont "overgrows" host with net negative growth. (when does this occur?)
-
-In this context, nitrogen-limitation represents a scenario in which the host is 'in control' of the system: negative feedbacks stabilize the symbiont to host biomass ratio and the system maintains positive growth.
-Carbon-limitation represents a loss of 'control', wherein positive feedbacks result in a loss of capacity for carbon fixation and positive growth.
-Importantly, these 'control' mechanisms arise entirely passively through the feedbacks present in the system.
-
-(Very low light also causes the symbiont to not share any carbon, resulting in very high S:H ratios and negative growth of the coral, which we interpret as a dysfunctional (and impossible) condition.)
+In addition to the suite of specific examples presented here, the model can be used to explore many different dynamic environmental scenarios, and represents a tool that biologists and ecologists can use to generate hypotheses and make predictions in both experimental and natural settings.
+Moreover, parameter values can be modified to correspond to different genetic or functional types of coral hosts and *Symbiodinium* partners in order to evaluate variability in system responses.
+Thus, the diversity of potential applications for this model is high, and we envision this work as a foundation for continued development, which may include more detailed treatments of specific modules (e.g., DIC processing), and the incorporation of more external forcing capabilities (e.g., external DIC).
+Importantly, this effort is best undertaken by the scientific community at large, including those with theoretical and empirical backgrounds, which is why the code for this model has been developed openly using the R language.
+Ultimately, the continued refinement of these tools is fundamental in elucidating the mechanisms of symbiosis function and dysfunction, and in predicting coral responses to environmental change.
 

### 2016-11-01 JTB_ms/JTB_ms.tex
complete draft

@@
             pdfauthor={},
-            pdftitle={A simplified dynamic bioenergetic model for coral-algal symbiosis and coral bleaching},
+            pdftitle={A simplified dynamic bioenergetic model for coral-Symbiodinium symbioses and coral bleaching as an alternate stable state},
             colorlinks=true,
@@
 
-  \title{A simplified dynamic bioenergetic model for coral-algal symbiosis and
-coral bleaching}
+  \title{A simplified dynamic bioenergetic model for coral-\emph{Symbiodinium}
+symbioses and coral bleaching as an alternate stable state}
     \author[University of Hawaii]{Ross Cunning\corref{c1}}
@@
   stability and breakdown of these symbioses (i.e., coral `bleaching')
-  calls for the development and application of theoretical tools. While
-  the biological complexity of the symbiosis is well-suited to dynamical
-  modeling, the theoretical complexity of certain approaches makes them
-  inaccessible to many biologists, thus limiting their potential
-  application and utility. Here, we develop a simplified dynamic
-  bioenergetic model of coral-algal symbiosis that demonstrates realistic
+  demands the development and application of theoretical tools; however,
+  the complexity of some approaches may limit their potential application
+  and utility. Here, we develop a simplified dynamic bioenergetic model of
+  coral-\emph{Symbiodinium} symbioses that demonstrates realistic
   steady-state patterns in coral growth and symbiont abundance across
@@
   mechanistic treatment of photo-oxidative stress, the model displays
-  dynamic responses of bleaching and recovery, and suggests that these
-  phenomena represent transitions between alternative stable states. A
-  suite of complex observed behaviors reproduced by the model suggests it
-  captures many important attributes of the system; meanwhile, its modular
-  framework and open source R code are designed to facilitate expansion
-  and problem-specific applications. We see high utility for this modeling
-  framework in generating testable hypotheses and predicting integrated,
-  mechanistic responses of corals to environmental change.
-
-  Latest development--sigmaCH and sigmaCS changed to 0.1 meaning 10\% of
-  metabolic CO2 produced is passively available to photosynthesis, to
-  reflect the fact that that \textasciitilde{}10\% of total internal DIC
-  at pH 7 is in the form of CO2. probably actually less than this since
-  internal pH may be closer to 7.5. This does not change any of the
-  figures substantively except the hysteresis figure--overshoot no longer
-  occurs with specific parameters run. may need to just change light range
-  or rate of change to reproduce overshoot.
+  dynamic behaviors of bleaching and recovery, and suggests that these
+  phenomena represent transitions between alternate stable states. A suite
+  of complex responses to multiple, interacting environmental factors
+  reproduced by the model suggests it unifyingly captures many important
+  attributes of the system; meanwhile, its modular framework and open
+  source R code are designed to facilitate expansion and problem-specific
+  applications. We see significant potential applications for this
+  modeling framework in generating testable hypotheses and predicting
+  integrated, mechanistic responses of corals to environmental change.
   \end{abstract}
@@
 jeopardizing their future persistence. In order to understand and
-predict coral responses to complex changes in their environment, a
+predict responses of corals to complex changes in their environment, a
 mechanistic understanding of how multiple interacting factors drive the
@@
 \item
-  \emph{Develop a detailed module of environmental stress.} Of primary
-  interest to coral biologists and ecologists is symbiosis dysfunction
-  under environmental stress, resulting in coral ``bleaching''--the loss
-  of algal symbionts from the association (Jokiel and Coles 1977).
+  \emph{Develop a module of photooxidative stress.} Of primary interest
+  to coral biologists and ecologists is symbiosis dysfunction under
+  environmental stress, resulting in coral ``bleaching''--the loss of
+  algal symbionts from the association (Jokiel and Coles 1977).
   Photooxidative stress in \emph{Symbiodinium} is considered a primary
@@
   mortality, though corals sometimes recover their symbionts. Bleaching
-  susceptibility, severity, and recovery may by influenced by
+  susceptibility, severity, and recovery may be influenced by
   interacting factors such as heterotrophy and nutrient availability
-  (Wooldridge 2014b), and the genetic identity of \emph{Symbiodinium}
-  (Glynn et al. 2001). To simulate these bleaching-related phenomena, we
-  develop a generalized framework linking overreduction of the
+  (Wooldridge 2014b). To simulate these events and interactions, we
+  develop a photooxidative stress module linking overreduction of the
   photosynthetic light reactions to downstream impacts of
@@
   logic of Jager, Martin, and Zimmer (2013), we exclude certain features
-  of formal DEB models in order to capture behaviors of interest with
+  of formal DEB theory in order to capture behaviors of interest with
   the simplest possible formulation. Here, we present a model without
   reserves, maturity, or reproduction (see Kooijman 2010). This
-  formulation restricts the model's scope to the bioenergetics of growth
-  and symbiosis dynamics in adult corals (i.e., reproduction, larval
-  stages, and metamorphosis are not considered), but greatly reduces
-  theoretical complexity and parameter numbers, which is advantageous
-  given the relative paucity of data for corals. However, our primary
-  motivation for reducing complexity was to increase accessibility and
-  applicability for biologists and ecologists without requiring
-  significant expertise in DEB theory.
+  formulation restricts the model's scope to the bioenergetics of adult
+  corals (i.e., reproduction, larval stages, and metamorphosis are not
+  considered), but greatly reduces theoretical complexity and parameter
+  numbers, which is advantageous given the relative paucity of data for
+  corals. However, our primary motivation for reducing complexity was to
+  increase accessibility and applicability for biologists and ecologists
+  without requiring significant expertise in DEB theory.
 \item
@@
   scientific community to include additional complexity and
-  problem-specific components. To widen the audience for this work, we
-  chose the R language because it is freely available and in common use
-  by biologists and ecologists.
+  problem-specific components. We chose the R language, which is freely
+  available and in common use by biologists and ecologists, to reach the
+  widest possible audience with this work.
 \end{enumerate}
@@
 With these as our primary motivations, we describe a simplified approach
-to dynamic bioenergetic modeling of coral-algal symbioses that tracks
-carbon and nitrogen acquisition and sharing between partners. This
-theoretical framework dynamically integrates the influence of external
-irradiance, nutrients, and prey availability on coral growth and
-symbiosis dynamics (i.e., symbiont:host biomass ratios), allowing for
-the possibility of coral bleaching in the event of photooxidative
-stress. In the following sections, we describe the formulation of this
-model and justify its structure and parameter values based on relevant
-literature. We then demonstrate the model's behavior and discuss some of
-its major implications and outcomes, and the wide range of potential
-applications for this model in the study of cnidarian-algal symbioses.
+to dynamic bioenergetic modeling of coral-\emph{Symbiodinium} symbioses
+that dynamically integrates the influence of external irradiance,
+nutrients, and prey availability on coral growth and symbiosis dynamics
+(i.e., symbiont:host biomass ratios), allowing for the possibility of
+coral bleaching in the event of photooxidative stress. In the following
+sections, we describe and provide rationale for the model structure, and
+demonstrate a range of steady state and dynamic behaviors that are
+consistent with observed phenomena.
 
@@
 
-In this dynamic bioenergetic model of coral-algal symbiosis, carbon and
-nitrogen are acquired by each partner and used to construct biomass. A
-graphical representation of the model is presented in Fig. 1, and each
-model flux and parameter are defined in Tables 1 and 2, respectively. We
-use C-moles as the unit of biomass for consistency with the rigorous
-mass balance of DEB theory: 1 C-mole is equivalent to the amount of
-biomass containing 1 mole of Carbon atoms. Host biomass (\(H\)),
-symbiont biomass (\(S\)), and prey biomass (\(X\)) have fixed, but
-different, molar N:C ratios (Table 2). Carbon and nitrogen are combined
-to produce biomass by synthesizing units (SU), which are mathematical
+In this dynamical system, both partners acquire and use carbon and
+nitrogen to construct biomass: the symbiont fixes carbon through
+photosynthesis, and receives nitrogen shared by the host, while the host
+acquires nitrogen from the environment and receives carbon shared by the
+symbiont. A graphical representation of the model is presented in Fig.
+1, and each model flux and parameter is defined in Tables 1 and 2,
+respectively. We use C-moles as the unit of biomass for consistency with
+the rigorous mass balance of DEB theory: 1 C-mole is equivalent to the
+amount of biomass containing 1 mole of carbon atoms. Host biomass
+(\(H\)), symbiont biomass (\(S\)), and prey biomass (\(X\)) have fixed,
+but different, molar N:C ratios (Table 2). Biomass is produced from
+carbon and nitrogen by synthesizing units (SU), which are mathematical
 specifications of the formation of a product from two substrates; we use
 the parallel complementary formulation of Kooijman (2010) to specify
-these fluxes. The two state variables are symbiont biomass and coral
-biomass; because resources are acquired proportionally to surface area,
-and surface area is proportional to volume (i.e., corals are
-``V1-morphs'' in DEB terminology (Kooijman 2010)), biomass increases
-exponentially during growth (indeed, corals grow exponentially (Bak
-1976)).
-
-Environmental stress is implemented in the form of photo-oxidative
+these fluxes. The two state variables of this system are symbiont
+biomass and coral biomass; because resources are acquired proportionally
+to surface area, and surface area is proportional to volume (i.e.,
+corals are ``V1-morphs'' in DEB terminology (Kooijman 2010)), biomass
+increases exponentially during growth (indeed, corals grow exponentially
+(Bak 1976)).
+
+Environmental stress is implemented in the form of photooxidative
 stress, which is thought to be the primary trigger of coral bleaching
@@
 high temperature (Hoegh-Guldberg 1999); thus, it is important to justify
-our consideration of light as the primary stressor in the model. In
-reality, light and temperature interact synergistically (Coles and
-Jokiel 1978; Jones et al. 1998), and in fact, any stressor that disrupts
-the quenching of light energy may lead to bleaching (Wooldridge 2010;
-Baker and Cunning 2015). This is because the proximate cause of
-photo-oxidative stress is excess excitation energy, but the upstream
-events that lead to this situation may be diverse: indeed, elevated
-temperature may inhibit Rubisco functioning (Jones et al. 1998) and the
-repair of the D1 protein in photosystem II (Warner, Fitt, and Schmidt
-1999), which reduces the capacity of photochemical quenching and leads
-to an excess of light energy. In this way, elevated temperature serves
-to reduce the threshold above which light stresses the system
-(Hoegh-Guldberg 1999); importantly, light is still the proximate
-stressor. Therefore, we omitted temperature from the model to maintain a
-desired level of simplicity, while still allowing photophysiological
-stress and bleaching to be simulated with biological realism in response
-to light.
+our consideration of light as the primary stressor. In reality, light
+and temperature interact synergistically (Coles and Jokiel 1978; Jones
+et al. 1998), and in fact, any stressor that disrupts the quenching of
+light energy may lead to bleaching (Wooldridge 2010; Baker and Cunning
+2015). This is because the proximate cause of photo-oxidative stress is
+excess excitation energy, but the upstream events that lead to this
+situation may be diverse: indeed, elevated temperature may inhibit
+Rubisco functioning (Jones et al. 1998) and the repair of the D1 protein
+in photosystem II (Warner, Fitt, and Schmidt 1999), which reduces the
+capacity of photochemical quenching and leads to an excess of light
+energy. In this way, elevated temperature serves to reduce the threshold
+above which light stresses the system (Hoegh-Guldberg 1999);
+importantly, light is still the proximate stressor. Therefore, we
+omitted temperature from the model to maintain a desired level of
+simplicity, while still allowing photooxidative stress and bleaching to
+be simulated with biological realism in response to light.
 
@@
 
+\begin{figure}[htbp]
+\centering
 \includegraphics{../img/Fig1.png}
+\caption{Graphical representation of coral-algal symbiosis model. Light,
+CO\textsubscript{2}, prey, and DIN are acquired from the external
+environment proportional to the biomass of the partner indicated by the
+black box for uptake. Mass fluxes (see Table 1 for definitions) are
+represented by \(j\)'s with subscripts indicating the type of mass, and
+in some cases the process (e.g., \(j_{CP}\) is the flux of carbon
+produced by photosynthesis), and \(\rho\)'s indicate fluxes that are
+shared by one partner with the other. Parallel complementary
+synthesizing units (SUs) are represented by large circles, and rejection
+fluxes from these SUs are indicated by dashed lines. \(c_{ROS}\) is a
+proportional rate that impacts other model fluxes by inhibition or
+acceleration; likewise, \(j_{eC}\) accelerates the rate of \(j_{CO_2}\).
+Recycled mass fluxes from biomass turnover are not shown for clarity
+(but see Table 1 for definitions).}
+\end{figure}
 
@@
 
-The balance equations for symbiont and host biomass are expressed as:
+The balance equations for symbiont (\(S\)) and host (\(H\)) biomass are
+expressed as:
 
@@
 (see Table 1 for definitions and units), which are each expressed as
-mass-specific rates (e.g., per C-mole of symbiont or host biomass). When
-necessary, conversions between symbiont-mass-specific and
+mass-specific rates (e.g., per C-mole of symbiont or host biomass per
+day). When necessary, conversions between symbiont-mass-specific and
 host-mass-specific rates are accomplished by multiplying or dividing by
@@
 
-where \(\rho_C\) is the surplus fixed carbon shared by the symbiont (see
-below), and \(r_{NH}\) is the recycled portion of the nitrogen liberated
-by host biomass turnover:
+where \(\rho_C\) is fixed carbon shared by the symbiont (see Eq. XX),
+and \(r_{NH}\) is recycled nitrogen liberated by host biomass turnover:
 
@@
 
-The carbon rejected from this SU reflects the amount of excess fixed
-carbon available to the host that is not used in biomass formation:
-
-\begin{equation} j_{eC} = (jX + \rho_C{S \over H} - j_{HG})_+ \end{equation}
-
-This flux, \(j_{eC}\), is assumed to be available as a respiratory
-substrate to support various energetically-demanding processes; of
-particular importance is the host's active supply of CO\textsubscript{2}
-for symbiont photosynthesis (Wooldridge 2013; Hopkinson, Tansik, and
-Fitt 2015) (see below).
-
 Due to the inherent inefficiency of the parallel complementary SU
@@
 there is always a non-zero rejection flux of excess carbon from the
-coral biomass SU.
+coral biomass SU. Likewise, the carbon rejected from this SU reflects
+the amount of excess fixed carbon available to the host that is not used
+in biomass formation:
+
+\begin{equation} j_{eC} = (jX + \rho_C{S \over H} - j_{HG})_+ \end{equation}
+
+This flux, \(j_{eC}\), is assumed to be available to the host as a
+respiratory substrate to support various energetically-demanding
+processes; of particular importance is the host's active carbon
+concentrating mechanisms (CCMs) that supply CO\textsubscript{2} for
+symbiont photosynthesis (Wooldridge 2013; Hopkinson, Tansik, and Fitt
+2015). We therefore specify \(j_{CO_2}\) as the host-mediated delivery
+of CO\textsubscript{2} to photosynthesis that encompasses potentially
+diverse carbon-concentrating mechanisms (CCMs), including active
+transport of bicarbonate, carbonic anhydrase-catalyzed conversion of
+bicarbonate to CO\textsubscript{2} to promote diffusion toward the
+symbiont (Tansik, Fitt, and Hopkinson 2015), and acidification of the
+symbiosome to increase localized CO\textsubscript{2} concentrations
+around the symbiont (Barott et al. 2014). Since these active CCMs
+require energetic input by the host, we define \(j_{CO_2}\) as
+proportional to \(j_{eC}\), assuming that some of this carbon is
+respired to energize the CCMs. This formulation means that the symbiont
+indirectly ensures its own CO\textsubscript{2} supply by providing fixed
+carbon (=energy) to the host (Wooldridge 2013), which establishes a
+positive feedback in the model that leads to the existence of multiple
+stable states, which we discuss below in the context of coral bleaching.
+The parameter \(k_{CO_2}\) scales the efficacy of host CCMs, which
+enables the comparison of different rates of CO\textsubscript{2}
+delivery that may characterize different coral species (Wooldridge
+2014a). The active input of CO\textsubscript{2} to the photosynthesis SU
+is therefore specified as:
+
+\begin{equation} j_{CO_2} = k_{CO_2}j_{eC} \end{equation}
 
@@
 
-We then specify the input of CO\textsubscript{2} to photosynthesis as
-\(j_{CO_2}\), a flux mediated by the host that encompasses potentially
-diverse carbon-concentrating mechanisms (CCMs), including active
-transport of bicarbonate, carbonic anhydrase-catalyzed conversion of
-bicarbonate to CO\textsubscript{2} to promote diffusion toward the
-symbiont (Tansik, Fitt, and Hopkinson 2015), and acidification of the
-symbiosome to increase localized CO\textsubscript{2} concentrations
-around the symbiont (Barott et al. 2014). Since these active CCMs
-require energetic input by the host, the magnitude of
-CO\textsubscript{2} delivery is proportional to the amount of fixed
-carbon available to the host after satisfying its growth demands,
-\(j_{eC}\), assuming that some of this carbon is respired to energize
-these CCMs. This formulation means that the symbiont indirectly ensures
-its own CO\textsubscript{2} supply by providing fixed carbon (=energy)
-to the host (Wooldridge 2013), which establishes a positive feedback in
-the model that leads to the existence of multiple stable states, which
-we discuss below in the context of coral bleaching. The parameter
-\(k_{CO_2}\) scales the efficacy of host CCMs, which enables the
-comparison of different rates of CO\textsubscript{2} delivery that may
-characterize different coral species (Wooldridge 2014a). The active
-input of CO\textsubscript{2} to the photosynthesis SU is therefore
-specified as:
-
-\begin{equation} j_{CO_2} = k_{CO_2}j_{eC} \end{equation}
-
-An additional fixed proportion of metabolic CO\textsubscript{2}
-generated by host biomass turnover is passively made available to the
-photosynthesis SU, according to:
+CO\textsubscript{2} arrives at the photosynthesis SU from multiple
+sources: in addition to the CO\textsubscript{2} actively supplied by the
+host through its CCMs (\(j_{CO_2}\); Eq. XX), a fixed proportion of
+metabolic CO\textsubscript{2} generated by host biomass turnover is
+passively available to the photosynthesis SU, according to:
 
@@
 
-along with a fixed proportion of symbiont biomass turnover:
+along with a fixed proportion of CO\textsubscript{2} generated by
+symbiont biomass turnover:
 
@@
 \(c_{ROS}\) is the relative rate of reactive oxygen species production
-(see below). Dividing the photosynthetic rate by \(c_{ROS}\) causes a
+(see Eq. XX). Dividing the photosynthetic rate by \(c_{ROS}\) causes a
 decline in response to photooxidative stress at high light levels, and
 the emergent outcome of this SU formulation demonstrates a classic
-photoinhibition response (see Fig. S1).
+photoinhibition response (Fig. S2).
 
 Light energy absorbed in excess of what is used to fix carbon is
-specified by the SU ``rejection flux'', according to:
+specified by the SU rejection flux, according to:
 
@@
 \emph{Symbiodinium} utilize a variety of pathways for non-photochemical
-quenching (NPQ; Roth 2014), which we collect in a relative NPQ capacity
-specified as a parameter of the symbiont (\(k_{NPQ}\)). The total NPQ
-flux is specified as a single-substrate SU formula:
+quenching (NPQ; Roth 2014), which we collect in a total NPQ capacity
+specified as a parameter of the symbiont (\(k_{NPQ}\)). The NPQ flux
+\(j_{NPQ}\) is then specified as a single-substrate SU formula with a
+maximum of \(k_{NPQ}\):
 
@@
 NPQ, then reactive oxygen species (ROS) are produced. We represent this
-as a relative quantity \(c_{ROS}\), taking a value of 1 when all light
-energy is quenched by photochemistry and NPQ, and increasing as the
-amount of excess excitation energy increases:
+as a relative quantity \(c_{ROS}\), which takes a value of 1 when all
+light energy is quenched by photochemistry and NPQ, and increases as the
+amount of excess excitation energy increases, specified as:
 
@@
 Importantly, \(c_{ROS}\) is specified here not as a function of absolute
-external light, but rather the amount of excess light, \(j_{eL}\), after
-accounting for quenching by carbon fixation and NPQ. A direct
-consequence of this formulation is that CO\textsubscript{2}-limitation
-of photosynthesis can lead to ROS production, an important mechanism
-(Wooldridge 2009) that was not captured by previous representations of
-photooxidative stress (Eynaud, Nisbet, and Muller 2011).
-
-Carbon fixed by photosynthesis (\(j_{CP}\)) is then combined with
-nitrogen shared by the host (\(\rho_N\)) and nitrogen recycled from
-symbiont biomass turnover
+external light, but rather the amount of excess light after accounting
+for quenching by carbon fixation and NPQ. A direct consequence of this
+formulation is that CO\textsubscript{2}-limitation of photosynthesis can
+lead to ROS production, an important mechanism (Wooldridge 2009) that
+was not captured by previous representations of photooxidative stress
+(Eynaud, Nisbet, and Muller 2011).
+
+Carbon fixed by photosynthesis (\(j_{CP}\); Eq. XX) is then combined
+with nitrogen shared by the host (\(\rho_N\); Eq. XX) and nitrogen
+recycled from symbiont biomass turnover
 
@@
 only occurs based on the maintenance component of turnover (i.e.,
-\(j_{ST}^0\)), and not the bleaching-related component, as this loss
+\(j_{ST}^0\)), and not the photodamage/bleaching component, as this loss
 represents biomass that is damaged or expelled from the host.)
@@
 each simulation, along which dynamic environmental forcing functions
-(irradiance, DIN, and prey abundance) can be designed. These vectors,
-along with initial values of symbiont and host biomass, then serve as
-input to the time-stepping function, which solves for the current system
-state using values of the previous system state where necessary. A time
-step of 0.1 days was used for all simulations, which were performed
-using R code that is available in the data repository accompanying this
-article: \url{github.com/jrcunning/Rcoral}.
+(irradiance, DIN, and prey abundance). These vectors, along with initial
+values of symbiont and host biomass, then serve as input to the
+time-stepping function, which solves for the current system state using
+values of the previous system state where necessary. A time step of 0.1
+days was used for all simulations, which were performed using R code
+that is available in the data repository accompanying this article:
+\url{github.com/jrcunning/Rcoral}. For steady state estimations,
+simulations were run until the change in both growth and S:H biomass was
+\textless{} 5e-7 per time step.
 
@@
 initially very low (i.e., a ``bleached'' coral), the system cannot
-escape positively reinforcing carbon-limitation and cannot grow.
-However, if symbiont biomass is initially high (i.e., a ``healthy''
-coral), then the system maintains positive growth in a nitrogen-limited
-state. For practical purposes, this section of the manuscript considers
-only positive growth steady states under constant environments; the
-subsequent ``dynamic behavior'' section explores how changing
-environments may cause the system to switch between alternate stable
-states, which we interpret in the context of coral bleaching.
-
-To analyze qualitative patterns in steady state behavior, we ran the
-model to steady state across gradients of external irradiance and
-nutrients (Fig. 2), which revealed patterns consistent with observed
-phenomena in corals. Predicted growth rates are low at low light and DIN
+escape this positive feedback and cannot grow. However, if symbiont
+biomass is initially high (i.e., a ``healthy'' coral), then the system
+remains in a nitrogen-limited state with positive growth. For practical
+purposes, this section of the manuscript considers only positive growth
+steady states under constant environments; the subsequent ``dynamic
+behavior'' section explores how changing environments may cause the
+system to switch between alternate stable states, which we interpret in
+the context of coral bleaching.
+
+To analyze steady state behavior, we ran the model to steady state
+across gradients of external irradiance and nutrients (Fig. 2), which
+revealed patterns consistent with observed phenomena in corals.
+Predicted growth rates are low at low light and DIN
 (\textasciitilde{}0.01 d\textsuperscript{-1}), and begin increasing as
@@
 1997; Brown et al. 1999; Fitt et al. 2000; Titlyanov et al. 2001). As
-higher light alleviates light-limitation of photosynthesis, growth
+higher light alleviates light-limitation of photosynthesis, host growth
 becomes less carbon-limited.
 
+\begin{figure}[htbp]
+\centering
 \includegraphics{../img/Fig2.png}
-
-Similarly, increasing DIN up to \textasciitilde{}2 µM alleviates
-nitrogen-limitation (Fig. 2A). Increased growth at higher DIN is
-predicted by the DEB model of Muller et al. (2009), and has also been
-observed experimentally (Muller-Parker et al. 1994; Tanaka et al. 2007;
-Tanaka et al. 2013). However, DIN elevation beyond a certain point
-(e.g., \textasciitilde{}3-4 µM in these simulations) has little effect
-on growth as carbon becomes limiting. Very high nutrient levels may even
-reduce growth (Shantz, Lemoine, and Burkepile 2015), although these
-impacts are not likely to occur within the range of concentrations
-considered here (\textless{}4 µM) (Ferrier-Pagès et al. 2000). In
-addition to increasing growth, DIN also increases the symbiont to host
-biomass ratio (Fig. 2B), a phenomenon also observed in reef corals
-(Marubini and Davies 1996). At low DIN and intermediate light, more
-typical of coral reef environments, symbiont to host biomass ratios are
-around \textasciitilde{}0.1-0.3, which is consistent with values
-reported in the literature (Muscatine, R McCloskey, and E Marian 1981;
-Edmunds et al. 2011; Hawkins et al. 2016).
+\caption{Steady state values of \textbf{A)} specific growth (Cmol
+Cmol\textsuperscript{-1} d\textsuperscript{-1}), \textbf{B)} the
+symbiont to host biomass ratio (CmolS CmolH\textsuperscript{-1}),
+\textbf{C)} relative C- or N-limitation of host biomass formation, and
+\textbf{D)} relative CO\textsubscript{2}- or light-limitation of
+symbiont photosynthesis, across gradients of external irradiance and
+dissolved inorganic nitrogen. Note that typical conditions for reefs are
+\textasciitilde{}1e-7 M DIN and 10-20 mol photons m\textsuperscript{-2}
+d\textsuperscript{-1}. Simulations for each combination of light and
+nutrients (41 points along each axis) were run for 100 days with a time
+step of 1 day. Negative steady state growth rates, and corresponding S:H
+ratios, were set to zero.}
+\end{figure}
+
+Similarly, increasing DIN alleviates nitrogen-limitation (Fig. 2A).
+Increased growth at higher DIN is predicted by the DEB model of Muller
+et al. (2009), and has also been observed experimentally (Muller-Parker
+et al. 1994; Tanaka et al. 2007; Tanaka et al. 2013). However, DIN
+elevation beyond a certain point (e.g., \textasciitilde{}3-4 µM in these
+simulations) has little effect on growth as carbon becomes limiting.
+Very high nutrient levels may even reduce growth (Shantz, Lemoine, and
+Burkepile 2015), although these impacts are not likely to occur within
+the range of concentrations considered here (\textless{}4 µM)
+(Ferrier-Pagès et al. 2000). In addition to increasing growth, DIN also
+increases the symbiont to host biomass ratio (Fig. 2B), a phenomenon
+also observed in reef corals (Marubini and Davies 1996). At low DIN and
+intermediate light, more typical of coral reef environments, symbiont to
+host biomass ratios are around \textasciitilde{}0.06-0.21, which is
+consistent with values reported in the literature (Muscatine, R
+McCloskey, and E Marian 1981; Edmunds et al. 2011; Hawkins et al. 2016).
 
 The maximum predicted growth rates of \textasciitilde{}0.1
-d\textsuperscript{-1}, occurring between \textasciitilde{}10-20 mol
+d\textsuperscript{-1}, occurring between \textasciitilde{}10-25 mol
 photons m\textsuperscript{-2} s\textsuperscript{-1} light and
@@
 
-As irradiance continues to increase above \textasciitilde{}20 mol
+As irradiance continues to increase above \textasciitilde{}25 mol
 photons m\textsuperscript{-2} d\textsuperscript{-1}, growth rates
-decline until positive growth ceases above \textasciitilde{}35 µmol
+decline until positive growth ceases above \textasciitilde{}40 µmol
 photons m\textsuperscript{-2} d\textsuperscript{-1} (Fig. 2A). The
 mechanism underlying this decline is the increase in light energy beyond
-the capacity of photosynthesis and non-photochemical quenching: excess
+the capacities of photosynthesis and non-photochemical quenching: excess
 excitation energy generates reactive oxygen species (ROS) (Weis 2008;
@@
 
-The incorporation of light stress in the model sets an upper limit to
-the amount of light at which a stable symbiotic interaction can be
-maintained, but even below this threshold of breakdown, negative effects
-of high light reduce steady state growth and symbiont:host biomass (Fig.
-2). This gradual decline is consistent with experimental results showing
-that high light levels decrease growth (Schutter et al. 2011), and field
-studies documenting optimum growth rates at intermediate depths (Baker
-and Weber 1975; Huston 1985). By incorporating these impacts of light
-stress, the model predicts greater, and more realistic, variation in
-state variables across light gradients than was predicted by the model
-of Muller et al. (2009), which did not include photoinhibition and
-photodamage.
-
-It is important to recognize that the boundary set by photo-oxidative
-stress on a stable symbiosis under steady state conditions (Fig. 2) may
-be temporarily crossed by a dynamic system, which may experience a
-period of symbiont loss (bleaching) and reduced growth, after which a
-return to benign conditions may restore symbiont biomass and positive
-growth. To explore this further and illustrate the behavior of the model
-in more detail, we evaluate a number of dynamic simulations below (see
-``Dynamic behavior'').
+The incorporation of photooxidative stress in the model sets an upper
+limit to the amount of light at which a stable symbiotic interaction can
+be maintained, but even below this threshold of breakdown, negative
+effects of high light reduce steady state growth and symbiont:host
+biomass (Fig. 2). This gradual decline is consistent with experimental
+results showing that high light levels decrease growth (Schutter et al.
+2011), and field studies documenting optimum growth rates at
+intermediate depths (Baker and Weber 1975; Huston 1985). By
+incorporating these impacts of light stress, the model predicts greater,
+and more realistic, variation in state variables across light gradients
+than was predicted by the model of Muller et al. (2009), which did not
+include photoinhibition and photodamage. It is important to recognize
+that the boundary set by photooxidative stress on a stable symbiosis
+under steady state conditions (Fig. 2) may be temporarily crossed by a
+dynamic system, which may experience a period of symbiont loss
+(bleaching) and reduced growth, after which a return to benign
+conditions may restore symbiont biomass and positive growth. To explore
+this further and illustrate the behavior of the model in more detail, we
+evaluate a number of dynamic simulations below (see ``Dynamic
+behavior'').
 
@@
 
+\begin{figure}[htbp]
+\centering
 \includegraphics{../img/Fig3.png}
+\caption{Sensitivity analysis. Plots show the fractional change in
+steady state values of growth (solid lines) and S:H biomass (dashed
+lines) in response to fractional changes in default parameter values
+(see Table 2 for default values). Parameters are grouped by which
+processes they are involved in. This sensitivity analysis was conducted
+at conditions typical for coral reef environments: low DIN (1e-7 M) and
+intermediate light (15 mol m\textsuperscript{-2} d\textsuperscript{-2}).
+Sensitivity analyses conducted other environmental conditions are
+presented in Figs S3-S7.}
+\end{figure}
 
@@
 light absorbed by symbionts, with lower values increasing
-light-limitation. Altering maximum growth and turnover rates of symbiont
-and host also have expected impacts on steady states (Fig. 3).
+light-limitation. Increasing the maximum growth and turnover rates have
+the expected effects of increasing and decreasing growth, respectively.
 Parameters relating to photooxidative stress and bleaching have little
-impact under these environmental conditions (Fig. 3), but of course have
-larger impacts under higher light (e.g., Fig. S6). Sensitivity analyses
+impact under these environmental conditions (Fig. 3), but have larger
+impacts under higher light (e.g., Fig. S6). Sensitivity analyses
 conducted under different combinations of external light and nutrients
-are presented in the Supplementary Information (Figs. S3-S7)
+are presented in Figs. S3-S7.
 
@@
 The dynamic behavior of the model demonstrates its power to integrate
-multiple environmental forcings simultaneously to produced integrated
-responses of the system. The potential applications of this are
-wide-ranging, and so the model may be highly useful for generating
-hypotheses about the way the system will respond to complex sets of
-environmental dynamics. Here we present several scenarios to highlight
-the power of the model and demonstrate that it can also reproduce more
-complex phenomena that have been observed in corals.
+multiple environmental forcings simultaneously. Here we present several
+scenarios that demonstrate the model's ability to reproduce complex
+phenomena that have been observed in corals.
 
@@
 environmental factors. Light in particular is a strong driver of these
-trends (Stimson:1997p3837; Brown et al. 1999; Fagoonee et al. 1999; Fitt
-et al. 2000), with high light associated with lower symbiont abundance
-and reduced tissue biomass. The role of light in driving seasonal
-changes in symbiont density was demonstrated nicely by Stimson (1997),
-who also found that experimental nutrient-enrichment amplified the
-light-driven seasonal oscillation. Using the levels of light and
-nutrients from this study as inputs, the model reproduces this observed
-interaction among environmental factors (Fig. 4), and also provides the
-mechanism: under high light in summer, photoinhibition reduces symbiont
-biomass and leads to carbon-limitation of host growth regardless of
-nutrient status; when light is reduced in winter, these responses become
-constrained instead by nitrogen, and therefore increase more when DIN is
-enriched (Fig. S8).
-
+trends (Stimson 1997; Brown et al. 1999; Fagoonee et al. 1999; Fitt et
+al. 2000), with high light associated with lower symbiont abundance and
+reduced tissue biomass. The role of light in driving seasonal changes in
+symbiont density was demonstrated nicely by Stimson (1997), who also
+found that experimental nutrient-enrichment amplified the light-driven
+seasonal oscillation. Using the levels of light and nutrients from this
+study as inputs, the model reproduces this observed interaction among
+environmental factors (Fig. 4), and also provides the mechanism: under
+high light in summer, photooxidative stress decreases symbiont biomass
+and increases carbon-limitation of host growth regardless of nutrient
+status; when light is reduced in winter, growth and symbiont biomass
+become constrained instead by nitrogen, and therefore increase more when
+DIN is enriched (Fig. S8).
+
+\begin{figure}[htbp]
+\centering
 \includegraphics{../img/Fig4.png}
+\caption{Figure 4. Light-driven seasonal dynamics of symbiont abundance
+and coral growth. Light input (upper panel) was designed as a sinusoidal
+curve with a period of one year, with maximum and minimum values of 44
+and 22 mol photons m\textsuperscript{-2} d\textsuperscript{-1}, based on
+Stimson (1997). The dynamic behavior of symiont to host biomass ratio
+(middle panel) and the specific growth rate of host biomass (lower
+panel) show seasonal oscillations that are greater in magnitude under
+higher nutrients (1e-5 molN L\textsuperscript{-1}; dashed lines)
+relative to lower nutrients (1e-7 molN L\textsuperscript{-1}; solid
+lines), consistent with the findings of Stimson (1997). Prey density was
+set at 1e-5 CmolX L\textsuperscript{-1}.}
+\end{figure}
 
@@
 field observations for corals, and demonstrates the power of this model
-to predict dynamic behavior that mechanistically integrate the impacts
-of multiple environmental drivers.
+to predict dynamic behavior that mechanistically integrates multiple
+environmental drivers.
 
@@
 stressors. In most cases, coral bleaching is thought to begin with
-photo-oxidative stress in symbiont photosynthesis (Lesser 1997), which
+photooxidative stress in symbiont photosynthesis (Lesser 1997), which
 triggers a cascade of events leading to symbiont expulsion (Weis 2008).
 As symbionts are expelled, the host receives less fixed carbon, which
-may then compromise its ability to deliver CO\textsubscript{2} to
-photosynthesis (Wooldridge 2013). Increasing
+may then compromise its ability to activate CCMs that deliver
+CO\textsubscript{2} to photosynthesis (Wooldridge 2013). Increasing
 CO\textsubscript{2}-limitation for remaining symbionts, along with an
@@
 (Enríquez, Méndez, and Iglesias-Prieto 2005), may further exacerbate
-photo-oxidative stress and accelerate symbiont expulsion, driving the
+photooxidative stress and accelerate symbiont expulsion, driving the
 coral into a bleached state.
@@
 disrupted (and light is sufficiently high), then the system is driven
-into the `bleached' state by photo-oxidative stress and positively
+into the `bleached' state by photooxidative stress and positively
 reinforcing carbon- and CO\textsubscript{2}-limitation. In this context,
@@
 conditions that push a healthy-state coral onto a trajectory leading to
-a bleached stable state.
+a bleached state.
 
@@
 tool to explore this dynamic behavior. Most straightforwardly, this
-switch occurs when irradiance exceeds the capacity of photochemistry and
-NPQ, causing the production of ROS that leads to photoinhibition,
-photodamage, and symbiont expulsion, thereby engaging the positive
-feedbacks that lead to a bleached state (Fig. 5). However, the positive
-feedbacks involved in bleaching are not engaged only in response to high
-external irradiance; in fact, they depend on the relative balance of
-light energy absorption and quenching, which in turn depends on the
-availability of CO\textsubscript{2} for photosynthesis. While previous
-models framed photo-oxidative stress as a fixed response to absolute
-external irradiance (Eynaud, Nisbet, and Muller 2011), our
-implementation considers the dynamic balance of multiple energy sinks in
-the causation of stress, which is more consistent with current
-understanding of symbiosis dysfunction (Wooldridge 2013), and
-establishes a critical role of host CCMs in providing
-CO\textsubscript{2} for photosynthesis (Tansik, Fitt, and Hopkinson
-2015; Hopkinson, Tansik, and Fitt 2015).
-
+switch occurs when increasing external irradiance (Fig. 5A) causes
+sufficient ROS production (Fig. 5B) and photoinhibition (Fig. 5C) such
+that the positive feedbacks between host carbon-limitation (Fig. 5D) and
+CO\textsubscript{2}-limitation of photosynthesis (Fig. 5E) are rapidly
+engaged, leading to even greater photooxidative stress and a rapid
+decline in S:H biomass (Fig. 5F), characteristic of coral bleaching.
+However, the positive feedbacks involved in bleaching are not engaged
+only in response to high external irradiance alone; in fact, they depend
+on the relative balance of light energy absorption and quenching, which
+in turn depends on the availability of CO\textsubscript{2} for
+photosynthesis. While previous models framed photooxidative stress as a
+fixed response to absolute external irradiance (Eynaud, Nisbet, and
+Muller 2011), our implementation considers the dynamic balance of
+multiple energy sinks in the causation of stress, which is more
+consistent with current understanding of symbiosis dysfunction
+(Wooldridge 2013), and establishes a critical role of host CCMs in
+providing CO\textsubscript{2} for photosynthesis (Tansik, Fitt, and
+Hopkinson 2015; Hopkinson, Tansik, and Fitt 2015).
+
+\begin{figure}[htbp]
+\centering
 \includegraphics{../img/Fig5.png}
+\caption{Figure 5. Coral bleaching as a switch from a nitrogen- to
+carbon-limited alternative stable state. This transition is demonstrated
+in response to gradually increasing external light (A), which causes an
+increase in production of ROS (B) that reduces the photosynthetic rate
+through photoinhibition (C). Decreasing photosynthesis moves the system
+from overall nitrogen-limitation toward carbon-limitation (D); when this
+threshold is crossed, the system rapidly becomes highly carbon-limited
+as photosynthesis becomes CO\textsubscript{2}-limited (E) and symbiont
+densities rapidly decline (F) into a bleached state. (All parameters at
+default values; external DIN=1e-7 molN/L, prey density=0).}
+\end{figure}
 
@@
 
+\begin{figure}[htbp]
+\centering
 \includegraphics{../img/Fig6.png}
+\caption{Figure 6. Bleaching with interactive factors. Simulations of
+high light stress (sinusoid with maximum=48 mol ph m\textsuperscript{-2}
+d\textsuperscript{-1}) under default environmental conditions (solid
+line; DIN=1e-7 molN/L; prey=3e-6 CmolX/L), or with elevated feeding
+(dashed line; prey=6e-6 CmolX/L), or elevated nutrients (dotted line;
+DIN=1e-6 molN/L). Initial symbiont biomass was set to the steady state
+for each set of starting conditions, respectively. (All parameters set
+to default values).}
+\end{figure}
 
@@
 
+\begin{figure}[htbp]
+\centering
 \includegraphics{../img/Fig7.png}
+\caption{Figure 7. Hysteresis in coral bleaching and recovery. The
+symbiont:host biomass ratio is plotted as a function of light from a
+simulation of linearly increasing, then decreasing, external light. The
+direction of the trajectory is indicated with arrowheads evenly spaced
+with respect to time. The rate of change in light was slow (0.01 mol ph
+d\textsuperscript{-1}) such that dynamics track closely to steady state
+values. (All parameters at default values; DIN=1e-7 molN/L; Prey=2e-6
+CmolX/L).}
+\end{figure}
 
@@
 behavior of the system: during recovery from bleaching, the S:H biomass
-ratio reaches a higher value that `overshoots' the level maintained in
+ratio reaches a higher value that `overshoots' the ratio maintained in
 the healthy stable state. In fact, such a phenomenon has been observed
@@
 symbiont remains carbon-limited; not until nitrogen becomes more
-strongly limiting to symbiont growth (hence forcing the symbiont to
-share fixed carbon with the host) will the system `flip' back to a state
-of high carbon fixation and overall nitrogen-limitation that
-characterize a `healthy' symbiosis. Of course, the magnitude of this
-`overshoot' will depend on external nutrient concentrations and the rate
-of change in the stressor during recovery, though simulations predict
-similar magnitudes to what was observed by Kemp et al. (supp fig?).
+strongly limiting to symbiont growth than carbon (hence forcing the
+symbiont to share carbon with the host) will the system flip back to the
+`healthy' state, characterized by high carbon fixation rates and overall
+nitrogen-limitation. Of course, the magnitude of this `overshoot'
+depends on external nutrient concentrations and the rate of change in
+the stressor during recovery, though simulations predict similar
+magnitudes to what was observed by Kemp et al. (2014) and Cunning,
+Silverstein, and Baker (2015) (Fig. S9).
 
@@
 
-The dynamic behavior of the model demonstrates its ability to
-mechanistically integrate multiple forcing functions, and the degree to
-which the system state depends not only on external forcing, but also
-the past history of the system. The model represents a unifying
-mechanistic framework that reproduces many phenomena observed in corals,
-suggesting it has captured many of the important features of the system
-in a relatively simple formulation. Furthermore, it allows the model to
-be used to predict responses of the system to multiple dynamic
-environmental forcing functions in a way that incorporates all positive
-and negative feedbacks. This makes the model potentially highly useful
-as a tool for generating hypotheses. make N, L, and X go up and down.
-For example, the model demonstrates and provides a mechanistic
-explanation for: an increase in symbiont abundance with nutrients (REF),
-an increase in symbiont abundance with heterotrophic feeding (refs),
-decrease in symbiont abundance with light (ref), optimal growth at
-intermediate light, rapid `bleaching' above threshold of high light
-(ref) along with positive feedbacks (wooldridge, enriquez), mitigation
-of bleaching by heterotrophic feeding, exacerbation of bleaching by
-elevated nutrients, higher standing stocks of less productive symbionts,
-greater seasonal change in symbiont density with elevated nutrients,
-overshoot of symbiont density after recovery from bleaching.
-
-``host control''=\textgreater{}N-limitation. C-limitation=host has lost
-control of symbiosis: either bleaches or symbiont ``overgrows'' host
-with net negative growth. (when does this occur?)
-
-In this context, nitrogen-limitation represents a scenario in which the
-host is `in control' of the system: negative feedbacks stabilize the
-symbiont to host biomass ratio and the system maintains positive growth.
-Carbon-limitation represents a loss of `control', wherein positive
-feedbacks result in a loss of capacity for carbon fixation and positive
-growth. Importantly, these `control' mechanisms arise entirely passively
-through the feedbacks present in the system.
-
-(Very low light also causes the symbiont to not share any carbon,
-resulting in very high S:H ratios and negative growth of the coral,
-which we interpret as a dysfunctional (and impossible) condition.)
+This simplified dynamic bioenergetic model of coral-\emph{Symbiodinium}
+symbioses mechanistically reproduces patterns in steady-state coral
+growth and symbiont abundance commonly observed in corals, including
+higher symbiont abundance with higher nutrients and feeding, lower
+symbiont abundance with increasing light, and optimal growth at
+intermediate light levels. Moreover, the model reproduces complex
+dynamic behaviors including seasonal changes in symbiont density at
+different nutrient levels, rapid bleaching above a threshold of high
+light, mitigation of bleaching by heterotrophic feeding, exacerbation of
+bleaching by elevated nutrients, and an overshoot of symbiont density
+during recovery from bleaching. These examples demonstrate the model's
+ability to integrate multiple environmental forcing functions to
+reproduce complex responses; meanwhile, the diversity of these phenomena
+suggest the model has captured many of the important features of the
+system in a unifying mechanistic framework. This model also provides a
+new conceptual framework for considering coral bleaching as a transition
+to an alternate stable state, which has important implications for
+understanding the performance and maintenance of symbiotic interactions.
+In this context, the `healthy' stable state represents a scenario in
+which the host is `in control' of the system, with nitrogen-limitation
+stabilizing the symbiont to host biomass ratio and maintaining positive
+growth. Conversely, carbon-limitation represents a loss of host
+`control', wherein positive feedbacks result in a loss of capacity for
+photosynthetic carbon fixation and positive growth. Importantly, these
+`control' mechanisms arise entirely passively through feedbacks present
+in the system.
+
+In addition to the suite of specific examples presented here, the model
+can be used to explore many different dynamic environmental scenarios,
+and represents a tool that biologists and ecologists can use to generate
+hypotheses and make predictions in both experimental and natural
+settings. Moreover, parameter values can be modified to correspond to
+different genetic or functional types of coral hosts and
+\emph{Symbiodinium} partners in order to evaluate variability in system
+responses. Thus, the diversity of potential applications for this model
+is high, and we envision this work as a foundation for continued
+development, which may include more detailed treatments of specific
+modules (e.g., DIC processing), and the incorporation of more external
+forcing capabilities (e.g., external DIC). Importantly, this effort is
+best undertaken by the scientific community at large, including those
+with theoretical and empirical backgrounds, which is why the code for
+this model has been developed openly using the R language. Ultimately,
+the continued refinement of these tools is fundamental in elucidating
+the mechanisms of symbiosis function and dysfunction, and in predicting
+coral responses to environmental change.
 
@@
 \emph{Limnology and Oceanography} 45 (3): 677--85.
-
-\hypertarget{ref-Glynn:2001p7571}{}
-Glynn, Peter W, Juan L Maté, Andrew C Baker, and MO Calderón. 2001.
-``Coral bleaching and mortality in Panama and Ecuador during the
-1997-1998 El Niño-Southern Oscillation event: Spatial/temporal patterns
-and comparisons with the 1982-1983 event.'' \emph{Bulletin of Marine
-Science} 69 (1): 79--109.
 

### 2016-11-01 JTB_supp/JTB_supp.Rmd
complete draft

@@
 # Initialize environments
-env1 <- init_env(time=time, L=c(22,44,2), N=c(1e-7,1e-7,0), X=c(1e-5,1e-5,0))
-env2 <- init_env(time=time, L=c(22,44,2), N=c(1e-5,1e-5,0), X=c(1e-5,1e-5,0))
+env1 <- init_env(time=time, L=c(20,44,2), N=c(0.14e-6,0.14e-6,0), X=c(1e-5,1e-5,0))
+env2 <- init_env(time=time, L=c(20,44,2), N=c(10.14e-6,10.14e-6,0), X=c(1e-5,1e-5,0))
 
 # Run simulations
-ss1 <- with(run_coral_ss(env=list(L=33,N=1e-7,X=1e-5), pars=defpars), last(S$S/H$H))
+ss1 <- with(run_coral_ss(env=list(L=33,N=0.14e-6,X=1e-5), pars=defpars), last(S$S/H$H))
 run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", ss1))
-ss2 <- with(run_coral_ss(env=list(L=33,N=1e-5,X=1e-5), pars=defpars), last(S$S/H$H))
+ss2 <- with(run_coral_ss(env=list(L=33,N=10.14e-6,X=1e-5), pars=defpars), last(S$S/H$H))
 run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", ss2))
@@
     par(new=T)
-    plot(time, env$N, type="l", col="blue", ylim=c(0,1e-5), lwd=2, axes=F, xlab="", ylab="")
+    plot(time, env$N, type="l", col="blue", ylim=c(0,10.14e-6), lwd=2, axes=F, xlab="", ylab="")
     axis(side=4); mtext(side=4, text="molN/L", line=1, cex=0.75)
@@
     pl <- log(   pmin((H$jCO2 + H$rCH)*H$H/S$S + S$rCS, pars$jCPm)   /   pmin(S$jL * pars$yCL, pars$jCPm)    )
-    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-0.8, 0.8))
+    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-2, 2))
     title("Photosynthesis: substrate-limitation", adj=0, line=0.25)
@@
     maxabs <- max(abs(pl))
-    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-0.8, 0.8))
+    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-2, 2))
     title("Photosynthesis: substrate-limitation", adj=0, line=0.25)

### 2016-11-01 R/Fig4.R

@@ -16,13 +16,13 @@ defpars <- def_pars()  # Get default parameters
 time <- seq(1, 365, 0.1)  # if single model run, use time input
 
 # Initialize environments
-env1 <- init_env(time=time, L=c(22,44,2), N=c(1e-7,1e-7,0), X=c(1e-5,1e-5,0))
-env2 <- init_env(time=time, L=c(22,44,2), N=c(1e-5,1e-5,0), X=c(1e-5,1e-5,0))
+env1 <- init_env(time=time, L=c(20,44,2), N=c(0.14e-6,0.14e-6,0), X=c(1e-5,1e-5,0))
+env2 <- init_env(time=time, L=c(20,44,2), N=c(10.14e-6,10.14e-6,0), X=c(1e-5,1e-5,0))
 
 # Run simulations
-ss1 <- with(run_coral_ss(env=list(L=33,N=1e-7,X=1e-5), pars=defpars), last(S$S/H$H))
+ss1 <- with(run_coral_ss(env=list(L=33,N=0.14e-6,X=1e-5), pars=defpars), last(S$S/H$H))
 run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", ss1))
-ss2 <- with(run_coral_ss(env=list(L=33,N=1e-5,X=1e-5), pars=defpars), last(S$S/H$H))
+ss2 <- with(run_coral_ss(env=list(L=33,N=10.14e-6,X=1e-5), pars=defpars), last(S$S/H$H))
 run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", ss2))
 
 # Create figure

### 2016-11-01 R/Fig7.R

@@ -7,7 +7,6 @@ sapply(c("R/def_pars.R",
 
 # Load default parameters
 defpars <- def_pars()
-defpars <- replace(defpars, "kROS", 80)
 
 # Set run time vector
 time <- seq(1, 5000, 0.1)  # if single model run, use time input
@@ -15,7 +14,7 @@ time <- seq(1, 5000, 0.1)  # if single model run, use time input
 # Initialize environment (with null light profile - to be replaced)
 env1 <- init_env(time=time, L=c(0,0,0), N=c(1e-7,1e-7,0), X=c(2e-6,2e-6,0))
 
-# Create light profile: increase from 20-40 for 2500 days (=.008 mol/d) then decrease for 2500 days
+# Create light profile: increase from 25-50 for 2500 days (=.01 mol/d) then decrease for 2500 days
 L.incr <- seq(25, 50, len=length(time) %/% 2)
 L.decr <- seq(50, 25, len=length(time) - length(L.incr))
 L <- list(c(L.incr, L.decr))

### 2016-11-01 R/run_coral_ss.R

@@ -1,6 +1,5 @@
 # This script initializes and runs a simulation using specified environmental inputs and model parameters.
 # The arguments to this function are:
-#   - time: a vector of time steps at which the model should be evaluated (units=days) (e.g., seq(0,365,0.1))
 #   - env: an object with structure matching the output of "init_env.R"
 #   - pars: an object with structure matching the output of "def_pars.R"
 

### 2016-11-01 R/plot_steady_states.R
omit zero-growth ss

@@ -5,7 +5,7 @@ plot_steady_states <- function(ss, png=NULL) {
   require(dplyr)
   
   # For runs that did NOT reach a steady state of positive growth, set all results to zero
-  ss[which(ss$gr < 0), c("gr", "sh")] <- NA#, "hl", "sl", "ee", "pl")] <- NA
+  ss[which(ss$gr <= 0), c("gr", "sh")] <- NA#, "hl", "sl", "ee", "pl")] <- NA
   # For runs where steady state value of S:H > 0.5, set S:H to 0.5
   ss[which(ss$sh > 0.5), "sh"] <- 0.5
   

### 2016-11-01 R/sensitivity.R
modify legends

@@ -55,10 +55,11 @@ sens.plot <- function(response, pars, cols, lines="response") {
   }
  
   # Add legends
-  legend("topright", legend=pars, lty=1, col=cols, bty="n")
+  lr <- ifelse(response=="grchange", "Growth", "S:H")
+  legend("topright", legend=pars, lty=1, col=cols, bty="n", cex=0.7)
   if (lines=="response") {
-    legend("topleft", legend=response, lty=1:nlevels(factor(response)), bty="n")
+    legend("topleft", legend=response, lty=1:nlevels(factor(response)), bty="n", cex=0.7)
   } else {
-    legend("topleft", legend=levels(dat[[1]]$env), lty=1:nlevels(dat[[1]]$env), bty="n")
+    legend("topleft", legend=levels(dat[[1]]$env), lty=1:nlevels(dat[[1]]$env), bty="n", cex=0.7)
   }
 }

### 2016-11-02 JTB_ms/JTB_ms.Rmd
update Eq. nos.

@@
 | $\rho_N$  | Nitrogen shared with the symbiont      | molN CmolH^-1^ d^-1^ | 7 |
-| $j_{eC}$  | Excess carbon used to activate host CCMs| molC CmolH^-1^ d^-1^ |  |
-| $j_L$     | Light absorption rate                  | molph CmolS^-1^ d^-1^| 9 |
-| $j_{CO_2}$| CO~2~ input to photosynthesis          | molCO~2~ CmolH^-1^ d^-1^| 10 |
-| $r_{CH}$  | Recycled CO~2~ from host turnover      | molCO~2~ CmolH^-1^ d^-1^| 11 |
-| $r_{CS}$  | Recycled CO~2~ from symbiont turnover  | molCO~2~ CmolS^-1^ d^-1^| 12 |
-| $j_{CP}$  | Photosynthesis rate                    | molC CmolS^-1^ d^-1^ | 13 |
-| $j_{eL}$  | Light energy in excess of photochemistry| molph CmolS^-1^ d^-1^| 14 |
-| $j_{NPQ}$ | Total capacity of NPQ                  | molph CmolS^-1^ d^-1^ | 15 |
-| $c_{ROS}$ | ROS production proportional to baseline | -- | 16 |
-| $r_{NS}$  | Recycled nitrogen from symbiont turnover| molN CmolS^-1^ d^-1^| 17 |
-| $j_{SG}$  | Symbiont biomass formation rate        | CmolS CmolS^-1^ d^-1^| 18 |
-| $\rho_C$  | Fixed carbon shared with host          | molC CmolS^-1^ d^-1^ | 19 |
-| $j_{ST}$  | Symbiont turnover rate                 | CmolS CmolS^-1^ d^-1^| 20 |
+| $j_{eC}$  | Excess carbon used to activate host CCMs| molC CmolH^-1^ d^-1^ | 8 |
+| $j_{CO_2}$| CO~2~ input to photosynthesis          | molCO~2~ CmolH^-1^ d^-1^| 9 |
+| $j_L$     | Light absorption rate                  | molph CmolS^-1^ d^-1^| 11 |
+| $r_{CH}$  | Recycled CO~2~ from host turnover      | molCO~2~ CmolH^-1^ d^-1^| 12 |
+| $r_{CS}$  | Recycled CO~2~ from symbiont turnover  | molCO~2~ CmolS^-1^ d^-1^| 13 |
+| $j_{CP}$  | Photosynthesis rate                    | molC CmolS^-1^ d^-1^ | 14 |
+| $j_{eL}$  | Light energy in excess of photochemistry| molph CmolS^-1^ d^-1^| 15 |
+| $j_{NPQ}$ | Total capacity of NPQ                  | molph CmolS^-1^ d^-1^ | 16 |
+| $c_{ROS}$ | ROS production proportional to baseline | -- | 17 |
+| $r_{NS}$  | Recycled nitrogen from symbiont turnover| molN CmolS^-1^ d^-1^| 18 |
+| $j_{SG}$  | Symbiont biomass formation rate        | CmolS CmolS^-1^ d^-1^| 19 |
+| $\rho_C$  | Fixed carbon shared with host          | molC CmolS^-1^ d^-1^ | 20 |
+| $j_{ST}$  | Symbiont turnover rate                 | CmolS CmolS^-1^ d^-1^| 21 |
 
-Table 2. Model parameters
+Table 2. Model parameters. Derivations of each parameter value along with supporting references are provided in the Supplementary Information.
 
@@
 
-where $\rho_C$ is fixed carbon shared by the symbiont (see Eq. XX), and $r_{NH}$ is recycled nitrogen liberated by host biomass turnover:
+where $\rho_C$ is fixed carbon shared by the symbiont (see Eq. 20), and $r_{NH}$ is recycled nitrogen liberated by host biomass turnover:
 
@@
 CO~2~ arrives at the photosynthesis SU from multiple sources: in addition to the CO~2~ actively supplied by the host through its CCMs ($j_{CO_2}$; Eq.
-XX), a fixed proportion of metabolic CO~2~ generated by host biomass turnover is passively available to the photosynthesis SU, according to:
+9), a fixed proportion of metabolic CO~2~ generated by host biomass turnover is passively available to the photosynthesis SU, according to:
 
@@
 where $j_{CPm}$ is the maximum specific rate of photosynthesis, and $c_{ROS}$ is the relative rate of reactive oxygen species production (see Eq.
-XX).
+17).
 Dividing the photosynthetic rate by $c_{ROS}$ causes a decline in response to photooxidative stress at high light levels, and the emergent outcome of this SU formulation demonstrates a classic photoinhibition response (Fig.
@@
 
-Carbon fixed by photosynthesis ($j_{CP}$; Eq. XX) is then combined with nitrogen shared by the host ($\rho_N$; Eq. XX) and nitrogen recycled from symbiont biomass turnover
+Carbon fixed by photosynthesis ($j_{CP}$; Eq. 14) is then combined with nitrogen shared by the host ($\rho_N$; Eq. 7) and nitrogen recycled from symbiont biomass turnover
 
@@
 However, changes in certain parameter values have more significant impacts: increasing $j_{Nm}$ or decreasing $K_N$ both dramatically increase host growth (Fig.
-3), demonstrating the strong nitrogen-limitation that characterizes these symbioses (REF).
+3), demonstrating the strong nitrogen-limitation that characterizes these symbioses.
 The parameter $\bar{a}^*$ has a strong impact on S:H biomass ratios (Fig.

### 2016-11-02 JTB_ms/JTB_ms.tex
update Eq. nos.

@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+8
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{CO_2}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+CO\textsubscript{2} input to photosynthesis
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molCO\textsubscript{2} CmolH\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+9
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-9
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{CO_2}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-CO\textsubscript{2} input to photosynthesis
+11
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(r_{CH}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+Recycled CO\textsubscript{2} from host turnover
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-10
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(r_{CH}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Recycled CO\textsubscript{2} from host turnover
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molCO\textsubscript{2} CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-11
+12
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-12
+13
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-13
+14
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-14
+15
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-15
+16
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-16
+17
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-17
+18
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-18
+19
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-19
+20
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-20
+21
 \strut\end{minipage}\tabularnewline
@@
 
-Table 2. Model parameters
+Table 2. Model parameters. Derivations of each parameter value along
+with supporting references are provided in the Supplementary
+Information.
 
@@
 
-where \(\rho_C\) is fixed carbon shared by the symbiont (see Eq. XX),
+where \(\rho_C\) is fixed carbon shared by the symbiont (see Eq. 20),
 and \(r_{NH}\) is recycled nitrogen liberated by host biomass turnover:
@@
 sources: in addition to the CO\textsubscript{2} actively supplied by the
-host through its CCMs (\(j_{CO_2}\); Eq. XX), a fixed proportion of
+host through its CCMs (\(j_{CO_2}\); Eq. 9), a fixed proportion of
 metabolic CO\textsubscript{2} generated by host biomass turnover is
@@
 \(c_{ROS}\) is the relative rate of reactive oxygen species production
-(see Eq. XX). Dividing the photosynthetic rate by \(c_{ROS}\) causes a
+(see Eq. 17). Dividing the photosynthetic rate by \(c_{ROS}\) causes a
 decline in response to photooxidative stress at high light levels, and
@@
 
-Carbon fixed by photosynthesis (\(j_{CP}\); Eq. XX) is then combined
-with nitrogen shared by the host (\(\rho_N\); Eq. XX) and nitrogen
+Carbon fixed by photosynthesis (\(j_{CP}\); Eq. 14) is then combined
+with nitrogen shared by the host (\(\rho_N\); Eq. 7) and nitrogen
 recycled from symbiont biomass turnover
@@
 demonstrating the strong nitrogen-limitation that characterizes these
-symbioses (REF). The parameter \(\bar{a}^*\) has a strong impact on S:H
+symbioses. The parameter \(\bar{a}^*\) has a strong impact on S:H
 biomass ratios (Fig. 3) since this parameter determines the amount of
@@
 have also been well-documented in the field (Fitt et al. 2000), along
-with reductions in net photosynthetic capacity (\textbf{???}).
+with reductions in net photosynthetic capacity (Muller-Parker 1987).
 Importantly, while light alone (at these levels) may drive seasonal
@@
 
+\hypertarget{ref-MullerParker:1987cr}{}
+Muller-Parker, Gisèle. 1987. ``Seasonal variation in light-shade
+adaptation of natural populations of the symbiotic sea anemone
+Aiptasiapulchella (Carlgren, 1943) in Hawaii.'' \emph{Journal of
+Experimental Marine Biology and Ecology} 112 (2): 165--83.
+doi:\href{https://doi.org/10.1016/0022-0981(87)90115-8}{10.1016/0022-0981(87)90115-8}.
+
 \hypertarget{ref-MullerParker:1994uo}{}

### 2016-11-23 JTB_supp/JTB_supp.Rmd
update title
Change dt to 0.1 in Fig S9

@@
 ---
-title: "A simplified dynamic bioenergetic model for coral-algal symbiosis and coral bleaching"
+title: A simplified dynamic bioenergetic model for coral-*Symbiodinium* symbioses and coral bleaching as an alternate stable state
 author: "Ross Cunning, Erik B. Muller, Ruth D. Gates, Roger M. Nisbet"
@@
 # Set run time vector
-time <- seq(-100, 1095, 0.5)  # if single model run, use time input
+time <- seq(-100, 1095, 0.1)  # if single model run, use time input
 

### 2016-12-21 JTB_ms/JTB_ms.Rmd
fix error in jSG eq.

@@
 
-\begin{equation} j_{SG} = \bigg({1 \over j_{SGm}} + {1 \over j_{CP}} + {1 \over {(\rho_N{H \over S} + r_{NS}) / n_{NH}}} - {1 \over {j_{CP} + (\rho_N{H \over S} + r_{NS}) / n_{NH}}}\bigg)^{-1} \end{equation}
+\begin{equation} j_{SG} = \bigg({1 \over j_{SGm}} + {1 \over j_{CP}} + {1 \over {(\rho_N{H \over S} + r_{NS}) / n_{NS}}} - {1 \over {j_{CP} + (\rho_N{H \over S} + r_{NS}) / n_{NS}}}\bigg)^{-1} \end{equation}
 

### 2017-01-05 JTB_ms/JTB_ms.Rmd

@@
 
-A time-stepping Euler method was used to solve the state equations since the production and rejection fluxes of the SUs are imiplicitly defined due to feedbacks present in the model.
-Specifically, the rejection fluxes of carbon and nitrogen from the symbiont and host biomass SUs act as reciprocal input fluxes to the other SU.
-In addition, the photosynthesis SU receives CO~2~ proportional to the carbon rejection flux from the host biomass SU, and the rejection flux of excitation energy from the photosynthesis SU acts to reduce its own production (i.e., photoinhibition).
-A vector of time steps is used for each simulation, along which dynamic environmental forcing functions (irradiance, DIN, and prey abundance).
-These vectors, along with initial values of symbiont and host biomass, then serve as input to the time-stepping function, which solves for the current system state using values of the previous system state where necessary.
-A time step of 0.1 days was used for all simulations, which were performed using R code that is available in the data repository accompanying this article: [github.com/jrcunning/Rcoral](github.com/jrcunning/Rcoral).
-For steady state estimations, simulations were run until the change in both growth and S:H biomass was < 5e-7 per time step.
+The model dynamics are specified by the differential equations (1-2) that impose biomass balance for host and symbiont and by a set of coupled non-linear algebraic equations (3-21) that define fluxes.
+Several of these fluxes are defined *implicitly*; for example, the rejection fluxes of carbon and nitrogen from the symbiont and host biomass SUs, respectively, act as reciprocal input fluxes to the other SU.
+Similarly, the photosynthesis SU receives CO~2~ at a rate proportional to the carbon rejection flux from the host biomass SU, and the rejection flux of excitation energy from the photosynthesis SU acts to reduce its own production through photoinhibition.
+Without further assumptions, however, the dynamical system is not always unambiguously defined because for some combinations of parameters and environmental forcing functions the system of algebraic equations has more than one solution with all fluxes non-negative (see results below).
+In such circumstances, the right hand side of the differential equations (1) and (2) is not uniquely defined even when $S$ and $H$ are specified.
+We resolved this problem by *defining the dynamical system* as the limit as a time step $\Delta t \rightarrow 0$ of a discretized system corresponding to Euler integration of the differential equations, with those fluxes that represent flows of elemental matter implemented by assuming that transfer of material between components of the system takes one time step.
+Thus, for example, CO~2~ rejected from the host SU at time $t$ arrives at the photosynthesis SU at time $t + \Delta t$.
 
-To aid in visualizing model results, we calculated values to indicate the degree to which product formation at an SU was limited by availability of either of its two substrates:
+Simulations using the discretized scheme were performed using R code that is available in the data repository accompanying this article: github.com/jrcunning/Rcoral.
+By experimentation, we found that a time step of 0.1 days gave adequate precision for most simulations (including used to generate Figs.
+XXX in this paper).
+For steady state estimations, simulations were run until the change in specific growth rate of host and symbiont and the $S:H$ biomass ratio was less than 1e-5 per time step.
+In regions of state space where very slow transient dynamics could be expected (i.e.
+near bifurcation points), sample steady state calculations were verified using MATHEMATICA code for numerical root finding (function FindRoot) with the code written independently by a coauthor without reference to the R code.
+
+To aid in visualizing model results, we calculated values to indicate the degree to which product formation at an SU was limited by availability of either of its two substrates using the formula
 

### 2017-01-05 JTB_ms/JTB_ms.tex

@@
 
-A time-stepping Euler method was used to solve the state equations since
-the production and rejection fluxes of the SUs are imiplicitly defined
-due to feedbacks present in the model. Specifically, the rejection
-fluxes of carbon and nitrogen from the symbiont and host biomass SUs act
-as reciprocal input fluxes to the other SU. In addition, the
-photosynthesis SU receives CO\textsubscript{2} proportional to the
-carbon rejection flux from the host biomass SU, and the rejection flux
-of excitation energy from the photosynthesis SU acts to reduce its own
-production (i.e., photoinhibition). A vector of time steps is used for
-each simulation, along which dynamic environmental forcing functions
-(irradiance, DIN, and prey abundance). These vectors, along with initial
-values of symbiont and host biomass, then serve as input to the
-time-stepping function, which solves for the current system state using
-values of the previous system state where necessary. A time step of 0.1
-days was used for all simulations, which were performed using R code
+The model dynamics are specified by the differential equations (1-2)
+that impose biomass balance for host and symbiont and by a set of
+coupled non-linear algebraic equations (3-21) that define fluxes.
+Several of these fluxes are defined \emph{implicitly}; for example, the
+rejection fluxes of carbon and nitrogen from the symbiont and host
+biomass SUs, respectively, act as reciprocal input fluxes to the other
+SU. Similarly, the photosynthesis SU receives CO\textsubscript{2} at a
+rate proportional to the carbon rejection flux from the host biomass SU,
+and the rejection flux of excitation energy from the photosynthesis SU
+acts to reduce its own production through photoinhibition. Without
+further assumptions, however, the dynamical system is not always
+unambiguously defined because for some combinations of parameters and
+environmental forcing functions the system of algebraic equations has
+more than one solution with all fluxes non-negative (see results below).
+In such circumstances, the right hand side of the differential equations
+(1) and (2) is not uniquely defined even when \(S\) and \(H\) are
+specified. We resolved this problem by \emph{defining the dynamical
+system} as the limit as a time step \(\Delta t \rightarrow 0\) of a
+discretized system corresponding to Euler integration of the
+differential equations, with those fluxes that represent flows of
+elemental matter implemented by assuming that transfer of material
+between components of the system takes one time step. Thus, for example,
+CO\textsubscript{2} rejected from the host SU at time \(t\) arrives at
+the photosynthesis SU at time \(t + \Delta t\).
+
+Simulations using the discretized scheme were performed using R code
 that is available in the data repository accompanying this article:
-\url{github.com/jrcunning/Rcoral}. For steady state estimations,
-simulations were run until the change in both growth and S:H biomass was
-\textless{} 5e-7 per time step.
+github.com/jrcunning/Rcoral. By experimentation, we found that a time
+step of 0.1 days gave adequate precision for most simulations (including
+used to generate Figs. XXX in this paper). For steady state estimations,
+simulations were run until the change in specific growth rate of host
+and symbiont and the \(S:H\) biomass ratio was less than 1e-5 per time
+step. In regions of state space where very slow transient dynamics could
+be expected (i.e.~near bifurcation points), sample steady state
+calculations were verified using MATHEMATICA code for numerical root
+finding (function FindRoot) with the code written independently by a
+coauthor without reference to the R code.
 
@@
 the degree to which product formation at an SU was limited by
-availability of either of its two substrates:
+availability of either of its two substrates using the formula
 

### 2017-01-05 JTB_ms/JTB_ms.Rmd
change figure order

@@
 In a constant environment, the system ultimately reaches a steady state of exponential growth or decline.
-However, under some conditions, either of these outcomes may occur depending on initial values of symbiont and host biomass, indicating the presence of alternate stable states.
+However, under some conditions, either of these outcomes may occur depending on initial values of symbiont and host biomass, indicating the presence of alternate stable states, which also correspond to either positive or negative growth (Fig.
+2).
 The mechanism that produces these alternate stable states is the positive feedback between carbon-limitation of the host and CO~2~-limitation of photosynthesis: if symbiont biomass is initially very low (i.e., a "bleached" coral), the system cannot escape this positive feedback and cannot grow.
@@
 
+![Alternate stable states](../img/Fig2.png)
+
 To analyze steady state behavior, we ran the model to steady state across gradients of external irradiance and nutrients (Fig.
-2), which revealed patterns consistent with observed phenomena in corals.
+3), which revealed patterns consistent with observed phenomena in corals.
 Predicted growth rates are low at low light and DIN (~0.01 d^-1^), and begin increasing as both of these factors increase (Fig.
-2A).
+3A).
 Low light limits photosynthetic rates, resulting in less fixed carbon shared with the host and an associated increase in the symbiont to host biomass ratio (Fig.
-2B).
+3B).
 In agreement with this trend are many observations of negative correlation between irradiance and symbiont density [@Stimson:1997p3837; @Brown:1999p3534; @Fitt:2000p3792; @Titlyanov:2001p3631].
@@
 Simulations for each combination of light and nutrients (41 points along each axis) were run for 100 days with a time step of 1 day.
-Negative steady state growth rates, and corresponding S:H ratios, were set to zero.](../img/Fig2.png)
+Negative steady state growth rates, and corresponding S:H ratios, were set to zero.](../img/Fig3.png)
 
 Similarly, increasing DIN alleviates nitrogen-limitation (Fig.
-2A).
+3A).
 Increased growth at higher DIN is predicted by the DEB model of @Muller:2009io, and has also been observed experimentally [@MullerParker:1994uo; @Tanaka:2007uv; @Tanaka:2013dj].
@@
 In addition to increasing growth, DIN also increases the symbiont to host biomass ratio (Fig.
-2B), a phenomenon also observed in reef corals [@Marubini:1996p8460].
+3B), a phenomenon also observed in reef corals [@Marubini:1996p8460].
 At low DIN and intermediate light, more typical of coral reef environments, symbiont to host biomass ratios are around ~0.06-0.21, which is consistent with values reported in the literature [@Muscatine:1981jy; @Edmunds:2011bv; @Hawkins:2016jj].
@@
 The maximum predicted growth rates of ~0.1 d^-1^, occurring between ~10-25 mol photons m^-2^ s^-1^ light and ~4 µM DIN (Fig.
-2A), are comparable to the rate of `r Tanaka_gr` d^-1^ measured by @Tanaka:2007uv under similar N-enriched conditions.
+3A), are comparable to the rate of `r Tanaka_gr` d^-1^ measured by @Tanaka:2007uv under similar N-enriched conditions.
 Predicted growth rates under conditions more typical of reef environments (<0.5 µM DIN) are ~0.01-0.03 d^-1^.
@@
 As irradiance continues to increase above ~25 mol photons m^-2^ d^-1^, growth rates decline until positive growth ceases above ~40 µmol photons m^-2^ d^-1^ (Fig.
-2A).
+3A).
 The mechanism underlying this decline is the increase in light energy beyond the capacities of photosynthesis and non-photochemical quenching: excess excitation energy generates reactive oxygen species (ROS) [@Weis:2008p944; @Roth:2014wf], which, in this model, have the phenomenological consequences of reducing the photosynthetic rate (representing photoinhibition) and increasing symbiont biomass loss (representing photodamage and/or symbiont expulsion) [see @Eynaud:2011tv].
 Together, these impacts reduce the symbiont to host biomass ratio (Fig.
-2B), as occurs during coral bleaching.
+3B), as occurs during coral bleaching.
 This reduction in symbionts consequently reduces the flux of fixed carbon to the host, resulting in increasing carbon-limitation (Fig.
-2B) and eventual cessation of growth (Fig.
-2A).
+3B) and eventual cessation of growth (Fig.
+3A).
 
 The incorporation of photooxidative stress in the model sets an upper limit to the amount of light at which a stable symbiotic interaction can be maintained, but even below this threshold of breakdown, negative effects of high light reduce steady state growth and symbiont:host biomass (Fig.
-2).
+3).
 This gradual decline is consistent with experimental results showing that high light levels decrease growth [@Schutter:2011ie], and field studies documenting optimum growth rates at intermediate depths [@Baker:1975ip; @Huston:1985ec].
@@
 It is important to recognize that the boundary set by photooxidative stress on a stable symbiosis under steady state conditions (Fig.
-2) may be temporarily crossed by a dynamic system, which may experience a period of symbiont loss (bleaching) and reduced growth, after which a return to benign conditions may restore symbiont biomass and positive growth.
+3) may be temporarily crossed by a dynamic system, which may experience a period of symbiont loss (bleaching) and reduced growth, after which a return to benign conditions may restore symbiont biomass and positive growth.
 To explore this further and illustrate the behavior of the model in more detail, we evaluate a number of dynamic simulations below (see "Dynamic behavior").
@@
 This sensitivity analysis was conducted at conditions typical for coral reef environments: low DIN (1e-7 M) and intermediate light (15 mol m^-2^ d^-2^).
-Sensitivity analyses conducted other environmental conditions are presented in Figs S3-S7.](../img/Fig3.png)
+Sensitivity analyses conducted other environmental conditions are presented in Figs S3-S7.](../img/Fig4.png)
 
@@
 However, changes in certain parameter values have more significant impacts: increasing $j_{Nm}$ or decreasing $K_N$ both dramatically increase host growth (Fig.
-3), demonstrating the strong nitrogen-limitation that characterizes these symbioses.
+4), demonstrating the strong nitrogen-limitation that characterizes these symbioses.
 The parameter $\bar{a}^*$ has a strong impact on S:H biomass ratios (Fig.
-3) since this parameter determines the amount of light absorbed by symbionts, with lower values increasing light-limitation.
+4) since this parameter determines the amount of light absorbed by symbionts, with lower values increasing light-limitation.
 Increasing the maximum growth and turnover rates have the expected effects of increasing and decreasing growth, respectively.
 Parameters relating to photooxidative stress and bleaching have little impact under these environmental conditions (Fig.
-3), but have larger impacts under higher light (e.g., Fig.
+4), but have larger impacts under higher light (e.g., Fig.
 S6).
@@
 Using the levels of light and nutrients from this study as inputs, the model reproduces this observed interaction among environmental factors (Fig.
-4), and also provides the mechanism: under high light in summer, photooxidative stress decreases symbiont biomass and increases carbon-limitation of host growth regardless of nutrient status; when light is reduced in winter, growth and symbiont biomass become constrained instead by nitrogen, and therefore increase more when DIN is enriched (Fig.
+5), and also provides the mechanism: under high light in summer, photooxidative stress decreases symbiont biomass and increases carbon-limitation of host growth regardless of nutrient status; when light is reduced in winter, growth and symbiont biomass become constrained instead by nitrogen, and therefore increase more when DIN is enriched (Fig.
 S8).
 
-![Figure 4.
-Light-driven seasonal dynamics of symbiont abundance and coral growth.
+![Light-driven seasonal dynamics of symbiont abundance and coral growth.
 Light input (upper panel) was designed as a sinusoidal curve with a period of one year, with maximum and minimum values of 44 and 22 mol photons m^-2^ d^-1^, based on @Stimson:1997p3837.
 The dynamic behavior of symiont to host biomass ratio (middle panel) and the specific growth rate of host biomass (lower panel) show seasonal oscillations that are greater in magnitude under higher nutrients (1e-5 molN L^-1^; dashed lines) relative to lower nutrients (1e-7 molN L^-1^; solid lines), consistent with the findings of @Stimson:1997p3837.
-Prey density was set at 1e-5 CmolX L^-1^.](../img/Fig4.png)
+Prey density was set at 1e-5 CmolX L^-1^.](../img/Fig5.png)
 
@@
 Nevertheless, the seasonal variability predicted here (Fig.
-4) is consistent with experimental and field observations for corals, and demonstrates the power of this model to predict dynamic behavior that mechanistically integrates multiple environmental drivers.
+5) is consistent with experimental and field observations for corals, and demonstrates the power of this model to predict dynamic behavior that mechanistically integrates multiple environmental drivers.
 
@@
 Most straightforwardly, this switch occurs when increasing external irradiance (Fig.
-5A) causes sufficient ROS production (Fig.
-5B) and photoinhibition (Fig.
-5C) such that the positive feedbacks between host carbon-limitation (Fig.
-5D) and CO~2~-limitation of photosynthesis (Fig.
-5E) are rapidly engaged, leading to even greater photooxidative stress and a rapid decline in S:H biomass (Fig.
-5F), characteristic of coral bleaching.
+6A) causes sufficient ROS production (Fig.
+6B) and photoinhibition (Fig.
+6C) such that the positive feedbacks between host carbon-limitation (Fig.
+6D) and CO~2~-limitation of photosynthesis (Fig.
+6E) are rapidly engaged, leading to even greater photooxidative stress and a rapid decline in S:H biomass (Fig.
+6F), characteristic of coral bleaching.
 However, the positive feedbacks involved in bleaching are not engaged only in response to high external irradiance alone; in fact, they depend on the relative balance of light energy absorption and quenching, which in turn depends on the availability of CO~2~ for photosynthesis.
@@
 
-![Figure 5.
-Coral bleaching as a switch from a nitrogen- to carbon-limited alternative stable state.
+![Coral bleaching as a switch from a nitrogen- to carbon-limited alternative stable state.
 This transition is demonstrated in response to gradually increasing external light (A), which causes an increase in production of ROS (B) that reduces the photosynthetic rate through photoinhibition (C).
 Decreasing photosynthesis moves the system from overall nitrogen-limitation toward carbon-limitation (D); when this threshold is crossed, the system rapidly becomes highly carbon-limited as photosynthesis becomes CO~2~-limited (E) and symbiont densities rapidly decline (F) into a bleached state.
-(All parameters at default values; external DIN=1e-7 molN/L, prey density=0).](../img/Fig5.png)
+(All parameters at default values; external DIN=1e-7 molN/L, prey density=0).](../img/Fig6.png)
 
@@
 For example, simulations of high light stress (Fig.
-6) demonstrate that bleaching can be attenuated by heterotrophic feeding, a phenomenon which has been observed experimentally [@Borell:2008p108].
+7) demonstrate that bleaching can be attenuated by heterotrophic feeding, a phenomenon which has been observed experimentally [@Borell:2008p108].
 The mechanism underlying this prediction is that feeding by the host increases host CCM activity, which delays the onset of CO~2~-limitation of photosynthesis and reduces bleaching severity.
 On the other hand, elevated nutrients exacerbate bleaching (Fig.
-6), since higher symbiont densities are more susceptible to CO~2~-limitation [@Wooldridge:2009p7807].
+7), since higher symbiont densities are more susceptible to CO~2~-limitation [@Wooldridge:2009p7807].
 Several experimental [@Cunning:2013gp; @VegaThurber:2014hk] and correlational studies [@Wooldridge:2009kb] are consistent with this mechanistic link between high nutrients and bleaching.
 
-![Figure 6.
-Bleaching with interactive factors.
+![Bleaching with interactive factors.
 Simulations of high light stress (sinusoid with maximum=48 mol ph m^-2^ d^-1^) under default environmental conditions (solid line; DIN=1e-7 molN/L; prey=3e-6 CmolX/L), or with elevated feeding (dashed line; prey=6e-6 CmolX/L), or elevated nutrients (dotted line; DIN=1e-6 molN/L).
 Initial symbiont biomass was set to the steady state for each set of starting conditions, respectively.
-(All parameters set to default values).](../img/Fig6.png)
+(All parameters set to default values).](../img/Fig7.png)
 
@@
 However, due to hysteresis, a coral cannot recover along the same trajectory it followed during bleaching; indeed, the stressor must be alleviated well *below* the threshold that initially caused bleaching in order for the system to recover (Fig.
-7).
+2; Fig.
+X).
 This is because under the same external conditions, a bleached coral (relative to a healthy one) is characterized by greater light amplification and weaker CCM activity, which serve to maintain the bleached state.
@@
 
-![Figure 7.
-Hysteresis in coral bleaching and recovery.
-The symbiont:host biomass ratio is plotted as a function of light from a simulation of linearly increasing, then decreasing, external light.
-The direction of the trajectory is indicated with arrowheads evenly spaced with respect to time.
-The rate of change in light was slow (0.01 mol ph d^-1^) such that dynamics track closely to steady state values.
-(All parameters at default values; DIN=1e-7 molN/L; Prey=2e-6 CmolX/L).](../img/Fig7.png)
+![new overshoot figure.....](../img/Fig8.png)
 
 The hysteresis demonstrated in Fig.
-7 reveals another interesting behavior of the system: during recovery from bleaching, the S:H biomass ratio reaches a higher value that 'overshoots' the ratio maintained in the healthy stable state.
+8 reveals another interesting behavior of the system: during recovery from bleaching, the S:H biomass ratio reaches a higher value that 'overshoots' the ratio maintained in the healthy stable state.
 In fact, such a phenomenon has been observed in both experimental [@Cunning:2015ja] and field studies [@Kemp:2014ug] documenting recovery from bleaching, and has been interpreted as a potential 'disequilibrium in host-symbiont regulation' [@Kemp:2014ug].

### 2017-01-05 R/Fig2.R

@@ -1,16 +1,49 @@
+# Load libraries
+library(parallel)
+library(doParallel)
+library(foreach)
+
 # Load functions
 sapply(c("R/def_pars.R",
-         "R/run_steady_states.R", 
-         "R/plot_steady_states.R"), 
+         "R/run_coral_ss.R"), 
        source, .GlobalEnv)
 
-# Set parameters
-defpars <- def_pars()  # Get default parameters
+# Get default parameter values
+defpars <- def_pars()
 
-# Set values of L and N at which to get steady state values
-at <- expand.grid(L=seq(from=0, to=60, length.out=41),
-                  N=seq(from=0, to=4e-6, length.out=41))
+# Set up cluster for parallel processing
+cl <- makeCluster(detectCores())  # Initiate cluster
+registerDoParallel(cl)
 
-# Run to steady state and plot results
-ssdat <- run_steady_states(pars=replace(defpars, "initS", 1), at=at, food=0, runtime=T)
-plot_steady_states(ssdat, png="img/Fig2.png")
+# Set input values for steady state runs
+input <- expand.grid(L=seq(20,55,1), initS=c(0.0001,0.1))
+
+# Run steady states in parallel
+output <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
+  run <- run_coral_ss(env=list(L=input$L[i], X=2e-6, N=1e-7), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+  list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
+}
+
+stopCluster(cl)  # Stop cluster
+
+# Collect results
+res <- cbind(input, output)
+
+# Create plot
+png("img/Fig2.png", width=5, height=5, units="in", res=300)
+
+plot(NA, xlim=c(20,55), ylim=c(0,0.15), xlab="Light", ylab="Steady state S:H ratio")
+points(c(48,48,53,53), c(0.12,0.13,0.12,0.13), pch=c(19,1,19,1), cex=c(0.4,1,0.4,1), col=c("black","black","red","red"), xpd=T)
+text(38, 0.125, labels="init. S:H", xpd=T, cex=0.7, srt=90)
+text(46, c(0.12,0.13), labels=c("0.1", "0.0001"), xpd=T, pos=2, cex=0.7)
+text(50, 0.15, labels=c("Steady state growth"), xpd=T, cex=0.7)
[…]

### 2017-01-05 R/Fig3.R

@@ -1,76 +1,16 @@
 # Load functions
 sapply(c("R/def_pars.R",
-         "R/run_coral_ss.R", 
-         "R/sensitivity.R"), 
+         "R/run_steady_states.R", 
+         "R/plot_steady_states.R"), 
        source, .GlobalEnv)
 
-# Set up time vector and parameters
+# Set parameters
 defpars <- def_pars()  # Get default parameters
-defpars <- replace(defpars, "initS", 1)  # Start with high S biomass to avoid alternate steady state of negative growth
 
-# Define environment(s) for conducting sensitivity analysis
-envs <- list(
-  HLLN=list(L=15, N=1e-7, X=0)
-)
+# Set values of L and N at which to get steady state values
+at <- expand.grid(L=seq(from=0, to=60, length.out=41),
+                  N=seq(from=0, to=4e-6, length.out=41))
 
-# Define relative changes in each parameter to evaluate
-levs <- seq(0.5, 1.5, 0.05)  # Set levels of % change in parameter value to evaluate
-
-# Define all simulations to be run for a given parameter
-sims <- expand.grid(env=names(envs), change=levs)
-n <- nrow(sims)
-
-# Set up cluster for parallel processing
-cl <- makeCluster(detectCores())  # Initiate cluster
-registerDoParallel(cl)
-
-# Run sensitivity analysis in each environment for each parameter
-jHT0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHT0", change=change))
-jNm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jNm", change=change))
-jHGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHGm", change=change))
-kCO2.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kCO2", change=change))
[…]

### 2017-01-05 R/Fig4.R

@@ -1,44 +1,76 @@
-# Figure 4
-# Seasonal variation in S:H at different DIN levels
-# Values based on Stimson et al. 1997
-
 # Load functions
 sapply(c("R/def_pars.R",
-         "R/init_env.R",
-         "R/run_coral.R",
-         "R/run_coral_ss.R"), 
+         "R/run_coral_ss.R", 
+         "R/sensitivity.R"), 
        source, .GlobalEnv)
 
-# Set parameters
+# Set up time vector and parameters
 defpars <- def_pars()  # Get default parameters
+defpars <- replace(defpars, "initS", 1)  # Start with high S biomass to avoid alternate steady state of negative growth
 
-# Set run time vector
-time <- seq(1, 365, 0.1)  # if single model run, use time input
+# Define environment(s) for conducting sensitivity analysis
+envs <- list(
+  HLLN=list(L=15, N=1e-7, X=0)
+)
 
-# Initialize environments
-env1 <- init_env(time=time, L=c(20,44,2), N=c(0.14e-6,0.14e-6,0), X=c(1e-5,1e-5,0))
-env2 <- init_env(time=time, L=c(20,44,2), N=c(10.14e-6,10.14e-6,0), X=c(1e-5,1e-5,0))
+# Define relative changes in each parameter to evaluate
+levs <- seq(0.5, 1.5, 0.05)  # Set levels of % change in parameter value to evaluate
 
-# Run simulations
-ss1 <- with(run_coral_ss(env=list(L=33,N=0.14e-6,X=1e-5), pars=defpars, dt=0.1), last(S$S/H$H))
-run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", ss1))
-ss2 <- with(run_coral_ss(env=list(L=33,N=10.14e-6,X=1e-5), pars=defpars, dt=0.1), last(S$S/H$H))
-run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", ss2))
+# Define all simulations to be run for a given parameter
+sims <- expand.grid(env=names(envs), change=levs)
+n <- nrow(sims)
 
-# Create figure
-png("img/Fig4.png", width=3, height=4, units="in", res=300)
-layout(mat=matrix(c(1,2,2,3,3)))
[…]

### 2017-01-05 R/Fig5.R

@@ -1,5 +1,6 @@
 # Figure 5
-# Bleaching in response to increasing light
+# Seasonal variation in S:H at different DIN levels
+# Values based on Stimson et al. 1997
 
 # Load functions
 sapply(c("R/def_pars.R",
@@ -8,68 +9,36 @@ sapply(c("R/def_pars.R",
          "R/run_coral_ss.R"), 
        source, .GlobalEnv)
 
-# Set run time vector
-time <- seq(1, 80, 0.1)  # if single model run, use time input
-
-# Initialize environment
-env <- init_env(time=time, L=c(30,50,0), N=c(1e-7,1e-7,0), X=c(0e-6,0e-6,0))
-#L <- list(c(rep(20,100), rep(50,300))[-(1:9)])
-#env <- replace(env, "L", L)
-
 # Set parameters
 defpars <- def_pars()  # Get default parameters
 
-# Run simulation
-ss <- with(run_coral_ss(env=list(L=25, N=1e-7, X=0e-6), pars=defpars, dt=0.1), last(S$S/H$H))
-run <- run_coral(time=time, env=env, pars=replace(defpars, "initS", ss))
+# Set run time vector
+time <- seq(1, 365, 0.1)  # if single model run, use time input
 
-# Plot results
-png("img/Fig5.png", width=3, height=3, units="in", res=300)
-  with(run, {
-    # Set up graphical parameters
-    par(mar=c(1.5,1.5,1,1), oma=c(0,0,0,0), mfcol=c(3,2), mgp=c(0.8,0,0), tck=0.025, lwd=1, xaxs="i", 
-        cex.main=0.9, cex.axis=0.6, cex.lab=0.7)
-    
-    # External irradiance
-    plot(time, env$L, type="l", col="gold", ylim=c(30,50), lwd=2, xlab="", ylab="molph/m2/d")
-    title("A. External irradiance", adj=0, line=0.25)
-    
-    # Plot ROS
-    plot(NA, xlim=range(time), xlab="", ylim=c(1, max(2, max(S$cROS))), ylab="Relative")
-    title("B. ROS production", adj=0.05, line=0)
-    lines(time, S$cROS, col="orange", lwd=2)
-    
-    # Photosynthesis rate
[…]

### 2017-01-05 R/Fig6.R

@@ -1,3 +1,6 @@
+# Figure 5
+# Bleaching in response to increasing light
+
 # Load functions
 sapply(c("R/def_pars.R",
          "R/init_env.R",
@@ -5,35 +8,68 @@ sapply(c("R/def_pars.R",
          "R/run_coral_ss.R"), 
        source, .GlobalEnv)
 
-time <- seq(1, 200, 0.1)  # Set time
+# Set run time vector
+time <- seq(1, 80, 0.1)  # if single model run, use time input
+
+# Initialize environment
+env <- init_env(time=time, L=c(30,50,0), N=c(1e-7,1e-7,0), X=c(0e-6,0e-6,0))
+#L <- list(c(rep(20,100), rep(50,300))[-(1:9)])
+#env <- replace(env, "L", L)
+
+# Set parameters
 defpars <- def_pars()  # Get default parameters
-pars <- defpars
 
-# Run 1
-env1 <- init_env(time=time, L=c(25,48,2), N=c(1e-7,1e-7,0), X=c(3e-6,3e-6,0))
-ss1 <- with(run_coral_ss(env=list(L=36, N=1e-7, X=3e-6), pars=pars, dt=0.1), last(S$S/H$H))
-run1 <- run_coral(time=time, env=env1, pars=replace(pars, "initS", ss1))
-# Run 2 - increase food
-env2 <- init_env(time=time, L=c(25,48,2), N=c(1e-7,1e-7,0), X=c(6e-6,6e-6,0))
-ss2 <- with(run_coral_ss(env=list(L=36, N=1e-7, X=6e-6), pars=pars, dt=0.1), last(S$S/H$H))
-run2 <- run_coral(time=time, env=env2, pars=replace(pars, "initS", ss2))
-# Run 3 - increase DIN
-env3 <- init_env(time=time, L=c(25,48,2), N=c(1e-6,1e-6,0), X=c(3e-6,3e-6,0))
-ss3 <- with(run_coral_ss(env=list(L=36, N=1e-6, X=3e-6), pars=pars, dt=0.1), last(S$S/H$H))
-run3 <- run_coral(time=time, env=env3, pars=replace(pars, "initS", ss3))
+# Run simulation
+ss <- with(run_coral_ss(env=list(L=25, N=1e-7, X=0e-6), pars=defpars, dt=0.1), last(S$S/H$H))
+run <- run_coral(time=time, env=env, pars=replace(defpars, "initS", ss))
 
-
-# Plot
[…]

### 2017-01-05 R/Fig7.R

@@ -5,33 +5,35 @@ sapply(c("R/def_pars.R",
          "R/run_coral_ss.R"), 
        source, .GlobalEnv)
 
-# Load default parameters
-defpars <- def_pars()
+time <- seq(1, 200, 0.1)  # Set time
+defpars <- def_pars()  # Get default parameters
+pars <- defpars
 
-# Set run time vector
-time <- seq(1, 5000, 1)  # THIS NEEDS TO BE CHANGED
+# Run 1
+env1 <- init_env(time=time, L=c(25,48,2), N=c(1e-7,1e-7,0), X=c(3e-6,3e-6,0))
+ss1 <- with(run_coral_ss(env=list(L=36, N=1e-7, X=3e-6), pars=pars, dt=0.1), last(S$S/H$H))
+run1 <- run_coral(time=time, env=env1, pars=replace(pars, "initS", ss1))
+# Run 2 - increase food
+env2 <- init_env(time=time, L=c(25,48,2), N=c(1e-7,1e-7,0), X=c(6e-6,6e-6,0))
+ss2 <- with(run_coral_ss(env=list(L=36, N=1e-7, X=6e-6), pars=pars, dt=0.1), last(S$S/H$H))
+run2 <- run_coral(time=time, env=env2, pars=replace(pars, "initS", ss2))
+# Run 3 - increase DIN
+env3 <- init_env(time=time, L=c(25,48,2), N=c(1e-6,1e-6,0), X=c(3e-6,3e-6,0))
+ss3 <- with(run_coral_ss(env=list(L=36, N=1e-6, X=3e-6), pars=pars, dt=0.1), last(S$S/H$H))
+run3 <- run_coral(time=time, env=env3, pars=replace(pars, "initS", ss3))
 
-# Initialize environment (with null light profile - to be replaced)
-env1 <- init_env(time=time, L=c(0,0,0), N=c(1e-7,1e-7,0), X=c(2e-6,2e-6,0))
 
-# Create light profile: increase from 25-50 for 2500 days (=.01 mol/d) then decrease for 2500 days
-L.incr <- seq(25, 50, len=length(time) %/% 2)
-L.decr <- seq(50, 25, len=length(time) - length(L.incr))
-L <- list(c(L.incr, L.decr))
-env1 <- replace(env1, "L", L)
-
[…]

### 2017-01-05 R/newFig2.R

@@ -1,49 +1,0 @@
-# Load libraries
-library(parallel)
-library(doParallel)
-library(foreach)
-
-# Load functions
-sapply(c("R/def_pars.R",
-         "R/run_coral_ss.R"), 
-       source, .GlobalEnv)
-
-# Get default parameter values
-defpars <- def_pars()
-
-# Set up cluster for parallel processing
-cl <- makeCluster(detectCores())  # Initiate cluster
-registerDoParallel(cl)
-
-# Set input values for steady state runs
-input <- expand.grid(L=seq(20,55,1), initS=c(0.0001,0.1))
-
-# Run steady states in parallel
-output <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
-  run <- run_coral_ss(env=list(L=input$L[i], X=2e-6, N=1e-7), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
-  list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
-}
-
-stopCluster(cl)  # Stop cluster
-
-# Collect results
-res <- cbind(input, output)
-
-# Create plot
-png("img/newFig2.png", width=5, height=5, units="in", res=300)
-
-plot(NA, xlim=c(20,55), ylim=c(0,0.15), xlab="Light", ylab="Steady state S:H ratio")
-points(c(48,48,53,53), c(0.12,0.13,0.12,0.13), pch=c(19,1,19,1), cex=c(0.4,1,0.4,1), col=c("black","black","red","red"), xpd=T)
-text(38, 0.125, labels="init. S:H", xpd=T, cex=0.7, srt=90)
-text(46, c(0.12,0.13), labels=c("0.1", "0.0001"), xpd=T, pos=2, cex=0.7)
-text(50, 0.15, labels=c("Steady state growth"), xpd=T, cex=0.7)
-text(c(48,53), 0.14, labels=c("pos.", "neg."), xpd=T, cex=0.7)
-
-apply(res, 1, FUN=function(x) {
-  with(x, points(L, sh,
-                 col=ifelse(gr>0, "black", "red"), 
-                 pch=ifelse(initS==1e-01, 19, 1), 
-                 cex=ifelse(initS==1e-01, 0.4, 1)))
-})
-
-dev.off()

### 2017-01-05 R/oldFig7.R

@@ -1,0 +1,37 @@
+# Load functions
+sapply(c("R/def_pars.R",
+         "R/init_env.R",
+         "R/run_coral.R",
+         "R/run_coral_ss.R"), 
+       source, .GlobalEnv)
+
+# Load default parameters
+defpars <- def_pars()
+
+# Set run time vector
+time <- seq(1, 5000, 1)  # THIS NEEDS TO BE CHANGED
+
+# Initialize environment (with null light profile - to be replaced)
+env1 <- init_env(time=time, L=c(0,0,0), N=c(1e-7,1e-7,0), X=c(2e-6,2e-6,0))
+
+# Create light profile: increase from 25-50 for 2500 days (=.01 mol/d) then decrease for 2500 days
+L.incr <- seq(25, 50, len=length(time) %/% 2)
+L.decr <- seq(50, 25, len=length(time) - length(L.incr))
+L <- list(c(L.incr, L.decr))
+env1 <- replace(env1, "L", L)
+
+ss <- with(run_coral_ss(env=list(L=25, N=1e-7, X=2e-6), pars=defpars, dt=0.1), last(S$S/H$H))
+
+run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", ss))#
+
+png("img/oldFig7.png", width=4, height=4, units="in", res=300)
+par(mar=c(2,2,1,1), mgp=c(1,0,0), tcl=0.25, cex.axis=0.5, cex.lab=0.75, xaxs="i", yaxs="i", mfrow=c(1,1))
+with(run1, {
+  plot(env$L, S$S/H$H, type="l", xlab="External light (molph/m2/d)", ylab="S:H biomass", ylim=c(0,0.2))
+  for (i in seq(1950, length(time), last(time)/2)) {
+    arrows(env$L[i-10], S$S[i-10]/H$H[i-10], 
+           env$L[i], S$S[i]/H$H[i], 
+           code=2, length=0.075)
+  }
+})
+dev.off()

### 2017-01-08 JTB_ms/JTB_ms.tex
remake

@@
 of these outcomes may occur depending on initial values of symbiont and
-host biomass, indicating the presence of alternate stable states. The
+host biomass, indicating the presence of alternate stable states,
+corresponding to either positive or negative growth (Fig. 2). The
 mechanism that produces these alternate stable states is the positive
@@
 
-To analyze steady state behavior, we ran the model to steady state
-across gradients of external irradiance and nutrients (Fig. 2), which
-revealed patterns consistent with observed phenomena in corals.
-Predicted growth rates are low at low light and DIN
+\begin{figure}[htbp]
+\centering
+\includegraphics{../img/Fig2.png}
+\caption{Alternate stable states}
+\end{figure}
+
+To analyze positive-growth steady state behavior, we ran the model to
+steady state across gradients of external irradiance and nutrients (Fig.
+3), which revealed patterns consistent with observed phenomena in
+corals. Predicted growth rates are low at low light and DIN
 (\textasciitilde{}0.01 d\textsuperscript{-1}), and begin increasing as
-both of these factors increase (Fig. 2A). Low light limits
+both of these factors increase (Fig. 3A). Low light limits
 photosynthetic rates, resulting in less fixed carbon shared with the
 host and an associated increase in the symbiont to host biomass ratio
-(Fig. 2B). In agreement with this trend are many observations of
+(Fig. 3B). In agreement with this trend are many observations of
 negative correlation between irradiance and symbiont density (Stimson
@@
 \centering
-\includegraphics{../img/Fig2.png}
+\includegraphics{../img/Fig3.png}
 \caption{Steady state values of \textbf{A)} specific growth (Cmol
@@
 
-Similarly, increasing DIN alleviates nitrogen-limitation (Fig. 2A).
+Similarly, increasing DIN alleviates nitrogen-limitation (Fig. 3A).
 Increased growth at higher DIN is predicted by the DEB model of Muller
@@
 (Ferrier-Pagès et al. 2000). In addition to increasing growth, DIN also
-increases the symbiont to host biomass ratio (Fig. 2B), a phenomenon
+increases the symbiont to host biomass ratio (Fig. 3B), a phenomenon
 also observed in reef corals (Marubini and Davies 1996). At low DIN and
@@
 photons m\textsuperscript{-2} s\textsuperscript{-1} light and
-\textasciitilde{}4 µM DIN (Fig. 2A), are comparable to the rate of 0.07
+\textasciitilde{}4 µM DIN (Fig. 3A), are comparable to the rate of 0.07
 d\textsuperscript{-1} measured by Tanaka et al. (2007) under similar
@@
 decline until positive growth ceases above \textasciitilde{}40 µmol
-photons m\textsuperscript{-2} d\textsuperscript{-1} (Fig. 2A). The
+photons m\textsuperscript{-2} d\textsuperscript{-1} (Fig. 3A). The
 mechanism underlying this decline is the increase in light energy beyond
@@
 symbiont expulsion) (see Eynaud, Nisbet, and Muller 2011). Together,
-these impacts reduce the symbiont to host biomass ratio (Fig. 2B), as
+these impacts reduce the symbiont to host biomass ratio (Fig. 3B), as
 occurs during coral bleaching. This reduction in symbionts consequently
 reduces the flux of fixed carbon to the host, resulting in increasing
-carbon-limitation (Fig. 2B) and eventual cessation of growth (Fig. 2A).
+carbon-limitation (Fig. 3B) and eventual cessation of growth (Fig. 3A).
 
@@
 effects of high light reduce steady state growth and symbiont:host
-biomass (Fig. 2). This gradual decline is consistent with experimental
+biomass (Fig. 3). This gradual decline is consistent with experimental
 results showing that high light levels decrease growth (Schutter et al.
@@
 that the boundary set by photooxidative stress on a stable symbiosis
-under steady state conditions (Fig. 2) may be temporarily crossed by a
+under steady state conditions (Fig. 3) may be temporarily crossed by a
 dynamic system, which may experience a period of symbiont loss
@@
 \centering
-\includegraphics{../img/Fig3.png}
+\includegraphics{../img/Fig4.png}
 \caption{Sensitivity analysis. Plots show the fractional change in
@@
 parameter values have more significant impacts: increasing \(j_{Nm}\) or
-decreasing \(K_N\) both dramatically increase host growth (Fig. 3),
+decreasing \(K_N\) both dramatically increase host growth (Fig. 4),
 demonstrating the strong nitrogen-limitation that characterizes these
 symbioses. The parameter \(\bar{a}^*\) has a strong impact on S:H
-biomass ratios (Fig. 3) since this parameter determines the amount of
+biomass ratios (Fig. 4) since this parameter determines the amount of
 light absorbed by symbionts, with lower values increasing
@@
 Parameters relating to photooxidative stress and bleaching have little
-impact under these environmental conditions (Fig. 3), but have larger
+impact under these environmental conditions (Fig. 4), but have larger
 impacts under higher light (e.g., Fig. S6). Sensitivity analyses
@@
 study as inputs, the model reproduces this observed interaction among
-environmental factors (Fig. 4), and also provides the mechanism: under
+environmental factors (Fig. 5), and also provides the mechanism: under
 high light in summer, photooxidative stress decreases symbiont biomass
@@
 \centering
-\includegraphics{../img/Fig4.png}
-\caption{Figure 4. Light-driven seasonal dynamics of symbiont abundance
-and coral growth. Light input (upper panel) was designed as a sinusoidal
-curve with a period of one year, with maximum and minimum values of 44
-and 22 mol photons m\textsuperscript{-2} d\textsuperscript{-1}, based on
+\includegraphics{../img/Fig5.png}
+\caption{Light-driven seasonal dynamics of symbiont abundance and coral
+growth. Light input (upper panel) was designed as a sinusoidal curve
+with a period of one year, with maximum and minimum values of 44 and 22
+mol photons m\textsuperscript{-2} d\textsuperscript{-1}, based on
 Stimson (1997). The dynamic behavior of symiont to host biomass ratio
@@
 depending on location and microhabitat. Nevertheless, the seasonal
-variability predicted here (Fig. 4) is consistent with experimental and
+variability predicted here (Fig. 5) is consistent with experimental and
 field observations for corals, and demonstrates the power of this model
@@
 
-\subsection{Coral bleaching}\label{coral-bleaching}
+\subsection{Coral bleaching and
+recovery}\label{coral-bleaching-and-recovery}
 
@@
 properties within a dynamical model. Interestingly, these feedbacks lead
-to alternate stable states. The `healthy' state is characterized by
-nitrogen-limitation of both symbiont and host: under these conditions,
-the symbiont translocates sufficient carbon to support host growth and
-CCMs, which ensures that photosynthesis does not become
+to alternate stable states in the symbiotic system. The `healthy' state
+is characterized by nitrogen-limitation of both symbiont and host: under
+these conditions, the symbiont translocates sufficient carbon to support
+host growth and CCMs, which ensures that photosynthesis does not become
 CO\textsubscript{2}-limited. However, if carbon translocation is
@@
 tool to explore this dynamic behavior. Most straightforwardly, this
-switch occurs when increasing external irradiance (Fig. 5A) causes
-sufficient ROS production (Fig. 5B) and photoinhibition (Fig. 5C) such
-that the positive feedbacks between host carbon-limitation (Fig. 5D) and
-CO\textsubscript{2}-limitation of photosynthesis (Fig. 5E) are rapidly
+switch occurs when increasing external irradiance (Fig. 6A) causes
+sufficient ROS production (Fig. 6B) and photoinhibition (Fig. 6C) such
+that the positive feedbacks between host carbon-limitation (Fig. 6D) and
+CO\textsubscript{2}-limitation of photosynthesis (Fig. 6E) are rapidly
 engaged, leading to even greater photooxidative stress and a rapid
-decline in S:H biomass (Fig. 5F), characteristic of coral bleaching.
+decline in S:H biomass (Fig. 6F), characteristic of coral bleaching.
 However, the positive feedbacks involved in bleaching are not engaged
@@
 \centering
-\includegraphics{../img/Fig5.png}
-\caption{Figure 5. Coral bleaching as a switch from a nitrogen- to
-carbon-limited alternative stable state. This transition is demonstrated
-in response to gradually increasing external light (A), which causes an
-increase in production of ROS (B) that reduces the photosynthetic rate
-through photoinhibition (C). Decreasing photosynthesis moves the system
-from overall nitrogen-limitation toward carbon-limitation (D); when this
+\includegraphics{../img/Fig6.png}
+\caption{Coral bleaching as a switch from a nitrogen- to carbon-limited
+alternative stable state. This transition is demonstrated in response to
+gradually increasing external light (A), which causes an increase in
+production of ROS (B) that reduces the photosynthetic rate through
+photoinhibition (C). Decreasing photosynthesis moves the system from
+overall nitrogen-limitation toward carbon-limitation (D); when this
 threshold is crossed, the system rapidly becomes highly carbon-limited
@@
 roles for other factors in influencing coral bleaching responses. For
-example, simulations of high light stress (Fig. 6) demonstrate that
+example, simulations of high light stress (Fig. 7) demonstrate that
 bleaching can be attenuated by heterotrophic feeding, a phenomenon which
@@
 of photosynthesis and reduces bleaching severity. On the other hand,
-elevated nutrients exacerbate bleaching (Fig. 6), since higher symbiont
+elevated nutrients exacerbate bleaching (Fig. 7), since higher symbiont
 densities are more susceptible to CO\textsubscript{2}-limitation
@@
 \centering
-\includegraphics{../img/Fig6.png}
-\caption{Figure 6. Bleaching with interactive factors. Simulations of
-high light stress (sinusoid with maximum=48 mol ph m\textsuperscript{-2}
+\includegraphics{../img/Fig7.png}
+\caption{Bleaching with interactive factors. Simulations of high light
+stress (sinusoid with maximum=48 mol ph m\textsuperscript{-2}
 d\textsuperscript{-1}) under default environmental conditions (solid
@@
 
-Since bleaching represents an alternative stable state, it follows that
-a bleached coral cannot recover unless the external condition that
-induced bleaching is alleviated. In natural settings, this typically
-occurs by seasonal declines in temperature and light. However, due to
-hysteresis, a coral cannot recover along the same trajectory it followed
-during bleaching; indeed, the stressor must be alleviated well
+Since bleaching is induced by an external stressor, recovery from
+bleaching cannot occur unless the stressor is alleviated. In natural
+settings, this typically occurs by seasonal declines in temperature and
+light. However, alternate stable states lead to significant hysteresis,
+indicating that a coral cannot recover along the same trajectory it
+followed during bleaching; indeed, the stressor must be alleviated well
 \emph{below} the threshold that initially caused bleaching in order for
-the system to recover (Fig. 7). This is because under the same external
-conditions, a bleached coral (relative to a healthy one) is
+the system to recover (Fig. 2; Fig. X). This is because under the same
+external conditions, a bleached coral (relative to a healthy one) is
 characterized by greater light amplification and weaker CCM activity,
@@
 will depend on external nutrients and prey availability, as well as
-model parameters, yet the fundamental concept demonstrates a core
-behavior of the system.
+model parameters, yet the core behavior of hysteresis illustrates the
+alternate stable states that exist in the symbiotic system.
 
@@
 \centering
-\includegraphics{../img/Fig7.png}
-\caption{Figure 7. Hysteresis in coral bleaching and recovery. The
-symbiont:host biomass ratio is plotted as a function of light from a
-simulation of linearly increasing, then decreasing, external light. The
-direction of the trajectory is indicated with arrowheads evenly spaced
-with respect to time. The rate of change in light was slow (0.01 mol ph
-d\textsuperscript{-1}) such that dynamics track closely to steady state
-values. (All parameters at default values; DIN=1e-7 molN/L; Prey=2e-6
-CmolX/L).}
+\includegraphics{../img/Fig8.png}
+\caption{Recovery from bleaching under varying N:C availability. Higher
+N:C ratios in a heterotrophic food source (effectively representing
+higher external DIN and/or lower feeding rates) cause a larger overshoot
+in the S:H biomass ratio, and prolong the duration of time until the
+system `recovers' by re-establishing nitrogen-limitation (represented by
+black dots on each trajectory). All simulations run with default
+parameters (except varying nNX), and DIN=1e-7 molN/L, prey=1e-6 CmolX/L,
+and initial symbiont biomass=0.0001.}
 \end{figure}
 
-The hysteresis demonstrated in Fig. 7 reveals another interesting
-behavior of the system: during recovery from bleaching, the S:H biomass
-ratio reaches a higher value that `overshoots' the ratio maintained in
-the healthy stable state. In fact, such a phenomenon has been observed
-in both experimental (Cunning, Silverstein, and Baker 2015) and field
-studies (Kemp et al. 2014) documenting recovery from bleaching, and has
-been interpreted as a potential `disequilibrium in host-symbiont
-regulation' (Kemp et al. 2014). The model explains this `overshoot' as
-an approach to a bifurcation point in the dynamical system: during
-recovery from the `bleached' state, host growth is more strongly
-carbon-limited than symbiont growth (since the symbiont has priority
-access to carbon), so S:H biomass increases. S:H biomass will continue
-to increase beyond the alternative steady state levels as long as the
-symbiont remains carbon-limited; not until nitrogen becomes more
-strongly limiting to symbiont growth than carbon (hence forcing the
-symbiont to share carbon with the host) will the system flip back to the
-`healthy' state, characterized by high carbon fixation rates and overall
-nitrogen-limitation. Of course, the magnitude of this `overshoot'
-depends on external nutrient concentrations and the rate of change in
-the stressor during recovery, though simulations predict similar
-magnitudes to what was observed by Kemp et al. (2014) and Cunning,
-Silverstein, and Baker (2015) (Fig. S9).
+The dynamics of recovery from bleaching reveal another interesting
+behavior of the system: under some conditions, the S:H biomass ratio may
+increase to a higher value that `overshoots' the ratio maintained in the
+`healthy' state under the same conditions (Fig. 8; Fig. S9). In fact,
+unusually high symbiont densities have been observed after recovery from
+bleaching in both experimental (Cunning, Silverstein, and Baker 2015)
+and field studies (Kemp et al. 2014), and have been interpreted as a
+potential `disequilibrium in host-symbiont regulation' (Kemp et al.
+2014). The model explains this `overshoot' as an approach to a
+bifurcation point in the dynamical system \textless{}\textgreater{}. As
+the symbiont begins repopulating the host, CO\textsubscript{2} is
+limiting and nitrogen is replete, allowing nearly all fixed carbon to be
+used for symbiont growth. As a result, host growth remains strongly
+carbon-limited, CCM activity remains weak, and photosynthesis remains
+CO\textsubscript{2}-limited. As long as this carbon-limitation persists,
+S:H biomass will continue to increase beyond the `healthy' steady state
+level, until the symbiont population eventually reaches a high enough
+density such that nitrogen limits its growth more strongly than carbon.
+At this point, which represents the peak of the `overshoot', the
+symbiont begins to share carbon with the host, which then supplies more
+CO\textsubscript{2} for photosynthesis, etc., causing the system to
+rapidly escape the positive feedbacks that were maintaining
+carbon-limitation (see Fig. S9 for illustration of all fluxes during
+overshoot). The re-establishment of nitrogen-limitation resumes high
+rates of carbon fixation and coral growth, bringing the S:H biomass
+ratio back down to a stable equilibrium.
+
+The occurrence and the magnitude of this overshoot depend on the
+relative inputs of carbon and nitrogen to the system -- any factor that
+enhances carbon-limitation (e.g.~high DIN and/or low feeding) therefore
+magnifies the overshoot and prolongs the dysfunctional, carbon-limited
+state of the symbiosis (Fig. 8). On the other hand, factors that favor
+nitrogen-limitation, such as low external DIN, high feeding rates, will
+prevent an overshoot from occurring at all, since nitrogen-limitation is
+re-established in the system before S:H biomass ratios reach high
+levels. While many scenarios are possible under different environmental
+conditions, we illustrate the general effects of varying N and C
+availability on recovery from bleaching with a series of simulations
+that vary the N:C ratio of host's heterotrophic food source (Fig. 8):
+lower N:C ratios (effectively representing lower DIN and/or higher
+heterotrophy) favor nitrogen-limitation and more rapid recovery, while
+higher N:C ratios (effectively representing higher DIN and/or lower
+heterotrophy) favor carbon-limitation and prolonged recovery with a
+larger overshoot in S:H biomass.
+
+Importantly, the re-establishment of nitrogen-limitation that
+characterizes the `healthy' symbiosis is the most important diagnostic
+for recovery, as this is when high carbon fixation and growth rates are
+resumed. A high S:H biomass ratio alone does not indicate a `healthy'
+state has been reached, because overall C-limitation could still persist
+and thus the host isn't really growing. This could explain why corals
+that have apparently recovered their symbiont populations after
+bleaching still do not grow or suffer reduced reproduction until much
+later (xxx). Together, these dynamics indicate that acquisition of
+carbon from a source other than the symbiont is extremely important for
+the host in order to recover from bleaching. The system can `recover'
+(in terms of resuming positive growth and N-limitation) more rapidly
+when the coral obtains more carbon. Indeed, feeding has been shown to
+promote more rapid recovery from bleaching (xxx). Moreover, if the host
+could obtain an independent source of carbon, such as direct uptake of
+DOC (xxx), this could be even more important in ensuring rapid recovery
+from a bleached state.
 

### 2017-01-08 JTB_ms/JTB_ms.Rmd
Finish figure 8 and overshoot discussion

@@
 In a constant environment, the system ultimately reaches a steady state of exponential growth or decline.
-However, under some conditions, either of these outcomes may occur depending on initial values of symbiont and host biomass, indicating the presence of alternate stable states, which also correspond to either positive or negative growth (Fig.
+However, under some conditions, either of these outcomes may occur depending on initial values of symbiont and host biomass, indicating the presence of alternate stable states, corresponding to either positive or negative growth (Fig.
 2).
@@
 
-To analyze steady state behavior, we ran the model to steady state across gradients of external irradiance and nutrients (Fig.
+To analyze positive-growth steady state behavior, we ran the model to steady state across gradients of external irradiance and nutrients (Fig.
 3), which revealed patterns consistent with observed phenomena in corals.
@@
 
-## Coral bleaching
+## Coral bleaching and recovery
 Coral bleaching is the stress-induced loss of symbiotic algae from coral tissues, which can occur in response to a variety of environmental stressors.
@@
 While these positive feedbacks have been discussed previously in the literature, this is the first attempt to implement and explore their properties within a dynamical model.
-Interestingly, these feedbacks lead to alternate stable states.
+Interestingly, these feedbacks lead to alternate stable states in the symbiotic system.
 The 'healthy' state is characterized by nitrogen-limitation of both symbiont and host: under these conditions, the symbiont translocates sufficient carbon to support host growth and CCMs, which ensures that photosynthesis does not become CO~2~-limited.
@@
 
-Since bleaching represents an alternative stable state, it follows that a bleached coral cannot recover unless the external condition that induced bleaching is alleviated.
+Since bleaching is induced by an external stressor, recovery from bleaching cannot occur unless the stressor is alleviated.
 In natural settings, this typically occurs by seasonal declines in temperature and light.
-However, due to hysteresis, a coral cannot recover along the same trajectory it followed during bleaching; indeed, the stressor must be alleviated well *below* the threshold that initially caused bleaching in order for the system to recover (Fig.
+However, alternate stable states lead to significant hysteresis, indicating that a coral cannot recover along the same trajectory it followed during bleaching; indeed, the stressor must be alleviated well *below* the threshold that initially caused bleaching in order for the system to recover (Fig.
 2; Fig.
@@
 This is because under the same external conditions, a bleached coral (relative to a healthy one) is characterized by greater light amplification and weaker CCM activity, which serve to maintain the bleached state.
-The magnitude of hysteresis will depend on external nutrients and prey availability, as well as model parameters, yet the fundamental concept demonstrates a core behavior of the system.
+The magnitude of hysteresis will depend on external nutrients and prey availability, as well as model parameters, yet the core behavior of hysteresis illustrates the alternate stable states that exist in the symbiotic system.
 
-![new overshoot figure.....](../img/Fig8.png)
+![Recovery from bleaching under varying N:C availability.
+Higher N:C ratios in a heterotrophic food source (effectively representing higher external DIN and/or lower feeding rates) cause a larger overshoot in the S:H biomass ratio, and prolong the duration of time until the system 'recovers' by re-establishing nitrogen-limitation (represented by black dots on each trajectory).
+All simulations run with default parameters (except varying nNX), and DIN=1e-7 molN/L, prey=1e-6 CmolX/L, and initial symbiont biomass=0.0001.](../img/Fig8.png)
 
-The hysteresis demonstrated in Fig.
-8 reveals another interesting behavior of the system: during recovery from bleaching, the S:H biomass ratio reaches a higher value that 'overshoots' the ratio maintained in the healthy stable state.
-In fact, such a phenomenon has been observed in both experimental [@Cunning:2015ja] and field studies [@Kemp:2014ug] documenting recovery from bleaching, and has been interpreted as a potential 'disequilibrium in host-symbiont regulation' [@Kemp:2014ug].
-The model explains this 'overshoot' as an approach to a bifurcation point in the dynamical system: during recovery from the 'bleached' state, host growth is more strongly carbon-limited than symbiont growth (since the symbiont has priority access to carbon), so S:H biomass increases.
-S:H biomass will continue to increase beyond the alternative steady state levels as long as the symbiont remains carbon-limited; not until nitrogen becomes more strongly limiting to symbiont growth than carbon (hence forcing the symbiont to share carbon with the host) will the system flip back to the 'healthy' state, characterized by high carbon fixation rates and overall nitrogen-limitation.
-Of course, the magnitude of this 'overshoot' depends on external nutrient concentrations and the rate of change in the stressor during recovery, though simulations predict similar magnitudes to what was observed by @Kemp:2014ug and @Cunning:2015ja (Fig.
+The dynamics of recovery from bleaching reveal another interesting behavior of the system: under some conditions, the S:H biomass ratio may increase to a higher value that 'overshoots' the ratio maintained in the 'healthy' state under the same conditions (Fig.
+8; Fig.
 S9).
+In fact, unusually high symbiont densities have been observed after recovery from bleaching in both experimental [@Cunning:2015ja] and field studies [@Kemp:2014ug], and have been interpreted as a potential 'disequilibrium in host-symbiont regulation' [@Kemp:2014ug].
+The model explains this 'overshoot' as an approach to a bifurcation point in the dynamical system <<NOT RIGHT>>.
+As the symbiont begins repopulating the host, CO~2~ is limiting and nitrogen is replete, allowing nearly all fixed carbon to be used for symbiont growth.
+As a result, host growth remains strongly carbon-limited, CCM activity remains weak, and photosynthesis remains CO~2~-limited.
+As long as this carbon-limitation persists, S:H biomass will continue to increase beyond the 'healthy' steady state level, until the symbiont population eventually reaches a high enough density such that nitrogen limits its growth more strongly than carbon.
+At this point, which represents the peak of the 'overshoot', the symbiont begins to share carbon with the host, which then supplies more CO~2~ for photosynthesis, etc., causing the system to rapidly escape the positive feedbacks that were maintaining carbon-limitation (see Fig.
+S9 for illustration of all fluxes during overshoot).
+The re-establishment of nitrogen-limitation resumes high rates of carbon fixation and coral growth, bringing the S:H biomass ratio back down to a stable equilibrium.
+
+The occurrence and the magnitude of this overshoot depend on the relative inputs of carbon and nitrogen to the system -- any factor that enhances carbon-limitation (e.g.
+high DIN and/or low feeding) therefore magnifies the overshoot and prolongs the dysfunctional, carbon-limited state of the symbiosis (Fig.
+8).
+On the other hand, factors that favor nitrogen-limitation, such as low external DIN, high feeding rates, will prevent an overshoot from occurring at all, since nitrogen-limitation is re-established in the system before S:H biomass ratios reach high levels.
+While many scenarios are possible under different environmental conditions, we illustrate the general effects of varying N and C availability on recovery from bleaching with a series of simulations that vary the N:C ratio of host's heterotrophic food source (Fig.
+8): lower N:C ratios (effectively representing lower DIN and/or higher heterotrophy) favor nitrogen-limitation and more rapid recovery, while higher N:C ratios (effectively representing higher DIN and/or lower heterotrophy) favor carbon-limitation and prolonged recovery with a larger overshoot in S:H biomass.
+
+Importantly, the re-establishment of nitrogen-limitation that characterizes the 'healthy' symbiosis is the most important diagnostic for recovery, as this is when high carbon fixation and growth rates are resumed.
+A high S:H biomass ratio alone does not indicate a 'healthy' state has been reached, because overall C-limitation could still persist and thus the host isn't really growing.
+This could explain why corals that have apparently recovered their symbiont populations after bleaching still do not grow or suffer reduced reproduction until much later (xxx).
+Together, these dynamics indicate that acquisition of carbon from a source other than the symbiont is extremely important for the host in order to recover from bleaching.
+The system can 'recover' (in terms of resuming positive growth and N-limitation) more rapidly when the coral obtains more carbon.
+Indeed, feeding has been shown to promote more rapid recovery from bleaching (xxx).
+Moreover, if the host could obtain an independent source of carbon, such as direct uptake of DOC (xxx), this could be even more important in ensuring rapid recovery from a bleached state.
 

### 2017-01-08 R/Fig8.R

@@ -1,0 +1,32 @@
+# Load functions
+sapply(c("R/def_pars.R",
+         "R/run_coral_ss.R"), 
+       source, .GlobalEnv)
+
+# Get default parameter values
+defpars <- def_pars()
+
+# Simulations with different prey N:C ratios
+time <- seq(1,300,0.1)
+run0.4 <- run_coral_ss(env=list(L=10, N=1e-7, X=1e-6), pars=replace(defpars, c("initS", "nNX"), c(0.0001, 0.4)), dt=0.1)
+run0.2 <- run_coral_ss(env=list(L=10, N=1e-7, X=1e-6), pars=replace(defpars, "initS", 0.0001), dt=0.1)
+run0.1 <- run_coral_ss(env=list(L=10, N=1e-7, X=1e-6), pars=replace(defpars, c("initS", "nNX"), c(0.0001, 0.10)), dt=0.1)
+run0.05 <- run_coral_ss(env=list(L=10, N=1e-7, X=1e-6), pars=replace(defpars, c("initS", "nNX"), c(0.0001, 0.05)), dt=0.1)
+
+# Plotting function
+plotrecov <- function(run, lty) with(run, {
+  lines(seq(1, nrow(S)), S$S/H$H, lty=lty)
+  points(min(which(H$dH.Hdt>0)), (S$S/H$H)[min(which(H$dH.Hdt>0))], pch=19, cex=0.75)
+})
+
+# Plot
+png("img/Fig8.png", width=5, height=5, units="in", res=300)
+par(mfrow=c(1,1))
+plot(NA, xlim=c(0,3000), ylim=c(0,1), xlab="Days", ylab="S:H ratio", xaxt="n")
+axis(side=1, at=seq(0,3000,1000), labels=seq(0,3000,1000)*0.1)
+plotrecov(run0.4, lty=4)
+plotrecov(run0.2, lty=3)
+plotrecov(run0.1, lty=2)
+plotrecov(run0.05, lty=1)
+legend("topleft", lty=c(4,3,2,1), legend=c("N:C=0.4", "N:C=0.2", "N:C=0.1", "N:C=0.05"), bty="n")
+dev.off()

### 2017-01-08 JTB_ms/JTB_ms.Rmd
add Fig8 to Makefile

@@
 ![Bleaching with interactive factors.
-Simulations of high light stress (sinusoid with maximum=48 mol ph m^-2^ d^-1^) under default environmental conditions (solid line; DIN=1e-7 molN/L; prey=3e-6 CmolX/L), or with elevated feeding (dashed line; prey=6e-6 CmolX/L), or elevated nutrients (dotted line; DIN=1e-6 molN/L).
+Simulations of high light stress (sinusoid with maximum=48 mol ph m^-2^ d^-1^) under default environmental conditions (solid line; DIN=1e-7 molN L^-1^; prey=3e-6 CmolX L^-1^), or with elevated feeding (dashed line; prey=6e-6 CmolX L^-1^), or elevated nutrients (dotted line; DIN=1e-6 molN L^-1^).
 Initial symbiont biomass was set to the steady state for each set of starting conditions, respectively.
@@
 Higher N:C ratios in a heterotrophic food source (effectively representing higher external DIN and/or lower feeding rates) cause a larger overshoot in the S:H biomass ratio, and prolong the duration of time until the system 'recovers' by re-establishing nitrogen-limitation (represented by black dots on each trajectory).
-All simulations run with default parameters (except varying nNX), and DIN=1e-7 molN/L, prey=1e-6 CmolX/L, and initial symbiont biomass=0.0001.](../img/Fig8.png)
+All simulations run with default parameters (except varying nNX), L=10 mol photons m^2^ d^-1^, DIN=1e-7 molN L^-1^, prey=1e-6 CmolX L^-1^, and initial S:H biomass=0.0001.](../img/Fig8.png)
 

### 2017-01-08 JTB_ms/JTB_ms.tex
add Fig8 to Makefile

@@
 d\textsuperscript{-1}) under default environmental conditions (solid
-line; DIN=1e-7 molN/L; prey=3e-6 CmolX/L), or with elevated feeding
-(dashed line; prey=6e-6 CmolX/L), or elevated nutrients (dotted line;
-DIN=1e-6 molN/L). Initial symbiont biomass was set to the steady state
-for each set of starting conditions, respectively. (All parameters set
-to default values).}
+line; DIN=1e-7 molN L\textsuperscript{-1}; prey=3e-6 CmolX
+L\textsuperscript{-1}), or with elevated feeding (dashed line; prey=6e-6
+CmolX L\textsuperscript{-1}), or elevated nutrients (dotted line;
+DIN=1e-6 molN L\textsuperscript{-1}). Initial symbiont biomass was set
+to the steady state for each set of starting conditions, respectively.
+(All parameters set to default values).}
 \end{figure}
@@
 black dots on each trajectory). All simulations run with default
-parameters (except varying nNX), and DIN=1e-7 molN/L, prey=1e-6 CmolX/L,
-and initial symbiont biomass=0.0001.}
+parameters (except varying nNX), L=10 mol photons m\textsuperscript{2}
+d\textsuperscript{-1}, DIN=1e-7 molN L\textsuperscript{-1}, prey=1e-6
+CmolX L\textsuperscript{-1}, and initial S:H biomass=0.0001.}
 \end{figure}

### 2017-01-08 R/Fig8.R

@@ -20,7 +20,7 @@ plotrecov <- function(run, lty) with(run, {
 })
 
 # Plot
-png("img/Fig8.png", width=5, height=5, units="in", res=300)
+png("img/Fig8.png", width=3, height=3, units="in", res=300)
 par(mfrow=c(1,1))
 plot(NA, xlim=c(0,3000), ylim=c(0,1), xlab="Days", ylab="S:H ratio", xaxt="n")
 axis(side=1, at=seq(0,3000,1000), labels=seq(0,3000,1000)*0.1)

### 2017-01-08 JTB_ms/JTB_ms.Rmd
Fig8 formatting

@@
 
-![Recovery from bleaching under varying N:C availability.
-Higher N:C ratios in a heterotrophic food source (effectively representing higher external DIN and/or lower feeding rates) cause a larger overshoot in the S:H biomass ratio, and prolong the duration of time until the system 'recovers' by re-establishing nitrogen-limitation (represented by black dots on each trajectory).
-All simulations run with default parameters (except varying nNX), L=10 mol photons m^2^ d^-1^, DIN=1e-7 molN L^-1^, prey=1e-6 CmolX L^-1^, and initial S:H biomass=0.0001.](../img/Fig8.png)
-
 The dynamics of recovery from bleaching reveal another interesting behavior of the system: under some conditions, the S:H biomass ratio may increase to a higher value that 'overshoots' the ratio maintained in the 'healthy' state under the same conditions (Fig.
@@
 
+![Recovery from bleaching under varying N:C availability.
+Higher N:C ratios in a heterotrophic food source (effectively representing higher external DIN and/or lower feeding rates) cause a larger overshoot in the S:H biomass ratio, and prolong the duration of time until the system 'recovers' by re-establishing nitrogen-limitation (represented by black dots on each trajectory).
+All simulations run with default parameters (except varying nNX), L=10 mol photons m^2^ d^-1^, DIN=1e-7 molN L^-1^, prey=1e-6 CmolX L^-1^, and initial S:H biomass=0.0001.](../img/Fig8.png)
+
 The occurrence and the magnitude of this overshoot depend on the relative inputs of carbon and nitrogen to the system -- any factor that enhances carbon-limitation (e.g.
@@
 8).
-On the other hand, factors that favor nitrogen-limitation, such as low external DIN, high feeding rates, will prevent an overshoot from occurring at all, since nitrogen-limitation is re-established in the system before S:H biomass ratios reach high levels.
+On the other hand, factors that favor nitrogen-limitation, such as low external DIN and/or high feeding rates, will accelerate the re-establishment of nitrogen-limitation and prevent an overshoot from occurring at all.
 While many scenarios are possible under different environmental conditions, we illustrate the general effects of varying N and C availability on recovery from bleaching with a series of simulations that vary the N:C ratio of host's heterotrophic food source (Fig.
@@
 
-Importantly, the re-establishment of nitrogen-limitation that characterizes the 'healthy' symbiosis is the most important diagnostic for recovery, as this is when high carbon fixation and growth rates are resumed.
-A high S:H biomass ratio alone does not indicate a 'healthy' state has been reached, because overall C-limitation could still persist and thus the host isn't really growing.
-This could explain why corals that have apparently recovered their symbiont populations after bleaching still do not grow or suffer reduced reproduction until much later (xxx).
-Together, these dynamics indicate that acquisition of carbon from a source other than the symbiont is extremely important for the host in order to recover from bleaching.
-The system can 'recover' (in terms of resuming positive growth and N-limitation) more rapidly when the coral obtains more carbon.
+Importantly, the re-establishment of nitrogen-limitation is the most important diagnostic of recovery to a 'healthy' state, as this is when high carbon fixation and growth rates are resumed.
+A high S:H biomass ratio alone does not indicate that a 'healthy' state has been reached, since overall C-limitation may still persist (e.g., high N:C ratios in Fig.
+8).
+This could explain why corals that have apparently recovered their symbiont populations after bleaching still have reduced growth and reproduction rates until much later (xxx, Hughes).
+Therefore, acquisition of carbon from a source other than the symbiont may be extremely important for the host in order to recover from bleaching.
 Indeed, feeding has been shown to promote more rapid recovery from bleaching (xxx).
-Moreover, if the host could obtain an independent source of carbon, such as direct uptake of DOC (xxx), this could be even more important in ensuring rapid recovery from a bleached state.
+Moreover, if the host obtains carbon from another independent source, such as direct uptake of DOC (xxx), this could be even more important in ensuring rapid recovery from a bleached state.
 

### 2017-01-08 JTB_ms/JTB_ms.tex
Fig8 formatting

@@
 state of the symbiosis (Fig. 8). On the other hand, factors that favor
-nitrogen-limitation, such as low external DIN, high feeding rates, will
-prevent an overshoot from occurring at all, since nitrogen-limitation is
-re-established in the system before S:H biomass ratios reach high
-levels. While many scenarios are possible under different environmental
-conditions, we illustrate the general effects of varying N and C
-availability on recovery from bleaching with a series of simulations
-that vary the N:C ratio of host's heterotrophic food source (Fig. 8):
-lower N:C ratios (effectively representing lower DIN and/or higher
-heterotrophy) favor nitrogen-limitation and more rapid recovery, while
-higher N:C ratios (effectively representing higher DIN and/or lower
-heterotrophy) favor carbon-limitation and prolonged recovery with a
-larger overshoot in S:H biomass.
-
-Importantly, the re-establishment of nitrogen-limitation that
-characterizes the `healthy' symbiosis is the most important diagnostic
-for recovery, as this is when high carbon fixation and growth rates are
-resumed. A high S:H biomass ratio alone does not indicate a `healthy'
-state has been reached, because overall C-limitation could still persist
-and thus the host isn't really growing. This could explain why corals
-that have apparently recovered their symbiont populations after
-bleaching still do not grow or suffer reduced reproduction until much
-later (xxx). Together, these dynamics indicate that acquisition of
-carbon from a source other than the symbiont is extremely important for
-the host in order to recover from bleaching. The system can `recover'
-(in terms of resuming positive growth and N-limitation) more rapidly
-when the coral obtains more carbon. Indeed, feeding has been shown to
-promote more rapid recovery from bleaching (xxx). Moreover, if the host
-could obtain an independent source of carbon, such as direct uptake of
-DOC (xxx), this could be even more important in ensuring rapid recovery
-from a bleached state.
+nitrogen-limitation, such as low external DIN and/or high feeding rates,
+will accelerate the re-establishment of nitrogen-limitation and prevent
+an overshoot from occurring at all. While many scenarios are possible
+under different environmental conditions, we illustrate the general
+effects of varying N and C availability on recovery from bleaching with
+a series of simulations that vary the N:C ratio of host's heterotrophic
+food source (Fig. 8): lower N:C ratios (effectively representing lower
+DIN and/or higher heterotrophy) favor nitrogen-limitation and more rapid
+recovery, while higher N:C ratios (effectively representing higher DIN
+and/or lower heterotrophy) favor carbon-limitation and prolonged
+recovery with a larger overshoot in S:H biomass.
+
+Importantly, the re-establishment of nitrogen-limitation is the most
+important diagnostic of recovery to a `healthy' state, as this is when
+high carbon fixation and growth rates are resumed. A high S:H biomass
+ratio alone does not indicate that a `healthy' state has been reached,
+since overall C-limitation may still persist (e.g., high N:C ratios in
+Fig. 8). This could explain why corals that have apparently recovered
+their symbiont populations after bleaching still have reduced growth and
+reproduction rates until much later (xxx, Hughes). Therefore,
+acquisition of carbon from a source other than the symbiont may be
+extremely important for the host in order to recover from bleaching.
+Indeed, feeding has been shown to promote more rapid recovery from
+bleaching (xxx). Moreover, if the host obtains carbon from another
+independent source, such as direct uptake of DOC (xxx), this could be
+even more important in ensuring rapid recovery from a bleached state.
 

### 2017-01-08 R/Fig8.R

@@ -21,12 +21,12 @@ plotrecov <- function(run, lty) with(run, {
 
 # Plot
 png("img/Fig8.png", width=3, height=3, units="in", res=300)
-par(mfrow=c(1,1))
-plot(NA, xlim=c(0,3000), ylim=c(0,1), xlab="Days", ylab="S:H ratio", xaxt="n")
-axis(side=1, at=seq(0,3000,1000), labels=seq(0,3000,1000)*0.1)
+par(mfrow=c(1,1), mar=c(2,2,1,2), mgp=c(1,0,0), tcl=0.25, xaxs="i", yaxs="i")
+plot(NA, xlim=c(0,3000), ylim=c(0,1), xlab="Days", ylab="S:H ratio", xaxt="n", cex.lab=0.9, cex.axis=0.75)
+axis(side=1, at=seq(0,3000,1000), labels=seq(0,3000,1000)*0.1, cex.axis=0.75)
 plotrecov(run0.4, lty=4)
 plotrecov(run0.2, lty=3)
 plotrecov(run0.1, lty=2)
 plotrecov(run0.05, lty=1)
-legend("topleft", lty=c(4,3,2,1), legend=c("N:C=0.4", "N:C=0.2", "N:C=0.1", "N:C=0.05"), bty="n")
+legend("topleft", lty=c(4,3,2,1), legend=c("N:C=0.4", "N:C=0.2", "N:C=0.1", "N:C=0.05"), bty="n", cex=0.5)
 dev.off()

### 2017-01-10 JTB_supp/JTB_supp.Rmd
replace Fig S9 with hysteresis under N/X regimes

@@
 
-# Figure S9. Overshoot of symbiont density following bleaching.
+# Figure S9. Hysteresis under different nutrient and feeding regimes.
 ```{r FigS9, cache=T}
-# Set run time vector
-time <- seq(-100, 1095, 0.1)  # if single model run, use time input
+# Set up cluster for parallel processing
+cl <- makeCluster(detectCores())  # Initiate cluster
+registerDoParallel(cl)
 
-# Initialize environment
-env <- init_env(time=time, L=c(60,20,4), N=c(2e-7,2e-7,0), X=c(2e-6,2e-6,0))
+# Set input values for steady state runs
+input <- expand.grid(L=seq(20,55,1), initS=c(0.0001,0.1))
 
-# Set parameters
-defpars <- def_pars()  # Get default parameters
+# Run steady states in parallel
+output1 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
+  run <- run_coral_ss(env=list(L=input$L[i], X=2e-6, N=1e-7), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+  list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
+}
 
-# Run simulation
-run <- run_coral(time=time, env=env, pars=defpars)
-plot_run(run)
+output2 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
+  run <- run_coral_ss(env=list(L=input$L[i], X=1e-6, N=1e-7), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+  list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
+}
+
+output3 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
+  run <- run_coral_ss(env=list(L=input$L[i], X=2e-6, N=1e-6), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+  list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
+}
+
+output4 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
+  run <- run_coral_ss(env=list(L=input$L[i], X=4e-6, N=1e-6), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+  list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
+}
+
+stopCluster(cl)  # Stop cluster
+
+# Collect results
+res1 <- cbind(input, output1)
+res2 <- cbind(input, output2)
+res3 <- cbind(input, output3)
+res4 <- cbind(input, output4)
+
+# Create plot
+par(mfrow=c(2,2))
+
+plot(NA, xlim=c(20,55), ylim=c(0,0.15), xlab="Light", ylab="Steady state S:H ratio")
+points(c(48,48,53,53), c(0.12,0.13,0.12,0.13), pch=c(19,1,19,1), cex=c(0.4,1,0.4,1), col=c("black","black","red","red"), xpd=T)
+text(38, 0.125, labels="init. S:H", xpd=T, cex=0.7, srt=90)
+text(46, c(0.12,0.13), labels=c("0.1", "0.0001"), xpd=T, pos=2, cex=0.7)
+text(50, 0.15, labels=c("Steady state growth"), xpd=T, cex=0.7)
+text(c(48,53), 0.14, labels=c("pos.", "neg."), xpd=T, cex=0.7)
+apply(res1, 1, FUN=function(x) {
+  with(x, points(L, sh,
+                 col=ifelse(gr>0, "black", "red"), 
+                 pch=ifelse(initS==1e-01, 19, 1), 
+                 cex=ifelse(initS==1e-01, 0.4, 1)))
+})
+
+plot(NA, xlim=c(20,55), ylim=c(0,0.15), xlab="Light", ylab="Steady state S:H ratio")
+apply(res2, 1, FUN=function(x) {
+  with(x, points(L, sh,
+                 col=ifelse(gr>0, "black", "red"), 
+                 pch=ifelse(initS==1e-01, 19, 1), 
+                 cex=ifelse(initS==1e-01, 0.4, 1)))
+})
+
+plot(NA, xlim=c(20,55), ylim=c(0,0.15), xlab="Light", ylab="Steady state S:H ratio")
+apply(res3, 1, FUN=function(x) {
+  with(x, points(L, sh,
+                 col=ifelse(gr>0, "black", "red"), 
+                 pch=ifelse(initS==1e-01, 19, 1), 
+                 cex=ifelse(initS==1e-01, 0.4, 1)))
+})
+
+plot(NA, xlim=c(20,55), ylim=c(0,0.15), xlab="Light", ylab="Steady state S:H ratio")
+apply(res4, 1, FUN=function(x) {
+  with(x, points(L, sh,
+                 col=ifelse(gr>0, "black", "red"), 
+                 pch=ifelse(initS==1e-01, 19, 1), 
+                 cex=ifelse(initS==1e-01, 0.4, 1)))
+})
+
+dev.off()
 ```

### 2017-01-11 JTB_supp/JTB_supp.Rmd
Fig S9 formatting

@@
 # Figure S9. Hysteresis under different nutrient and feeding regimes.
-```{r FigS9, cache=T}
+```{r FigS9, cache=T, fig.width=7, fig.height=7, echo=F, message=F}
 # Set up cluster for parallel processing
@@
 # Create plot
-par(mfrow=c(2,2))
+par(mfrow=c(2,2), mar=c(3,3,1,1))
 
-plot(NA, xlim=c(20,55), ylim=c(0,0.15), xlab="Light", ylab="Steady state S:H ratio")
-points(c(48,48,53,53), c(0.12,0.13,0.12,0.13), pch=c(19,1,19,1), cex=c(0.4,1,0.4,1), col=c("black","black","red","red"), xpd=T)
-text(38, 0.125, labels="init. S:H", xpd=T, cex=0.7, srt=90)
-text(46, c(0.12,0.13), labels=c("0.1", "0.0001"), xpd=T, pos=2, cex=0.7)
-text(50, 0.15, labels=c("Steady state growth"), xpd=T, cex=0.7)
-text(c(48,53), 0.14, labels=c("pos.", "neg."), xpd=T, cex=0.7)
+plot(NA, xlim=c(20,55), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
+points(c(48,48,53,53), c(0.22,0.23,0.22,0.23), pch=c(19,1,19,1), cex=c(0.4,1,0.4,1), col=c("black","black","red","red"), xpd=T)
+text(38, 0.225, labels="init. S:H", xpd=T, cex=0.7, srt=90)
+text(46, c(0.22,0.23), labels=c("0.1", "0.0001"), xpd=T, pos=2, cex=0.7)
+text(50, 0.25, labels=c("Steady state growth"), xpd=T, cex=0.7)
+text(c(48,53), 0.24, labels=c("pos.", "neg."), xpd=T, cex=0.7)
 apply(res1, 1, FUN=function(x) {
@@
 
-plot(NA, xlim=c(20,55), ylim=c(0,0.15), xlab="Light", ylab="Steady state S:H ratio")
+plot(NA, xlim=c(20,55), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
 apply(res2, 1, FUN=function(x) {
@@
 
-plot(NA, xlim=c(20,55), ylim=c(0,0.15), xlab="Light", ylab="Steady state S:H ratio")
+plot(NA, xlim=c(20,55), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
 apply(res3, 1, FUN=function(x) {
@@
 
-plot(NA, xlim=c(20,55), ylim=c(0,0.15), xlab="Light", ylab="Steady state S:H ratio")
+plot(NA, xlim=c(20,55), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
 apply(res4, 1, FUN=function(x) {
@@
 
-dev.off()
 ```

### 2017-01-11 JTB_ms/JTB_ms.tex

@@
 purposes, this section of the manuscript considers only positive growth
-steady states under constant environments; the subsequent ``dynamic
-behavior'' section explores how changing environments may cause the
-system to switch between alternate stable states, which we interpret in
-the context of coral bleaching.
+steady states under constant environments; subsequently, we explore how
+environmental forcing may cause the system to switch between alternate
+stable states, which we interpret in the context of coral bleaching (see
+``Coral Bleaching and Recovery'', below).
 
@@
 \includegraphics{../img/Fig3.png}
-\caption{Steady state values of \textbf{A)} specific growth (Cmol
-Cmol\textsuperscript{-1} d\textsuperscript{-1}), \textbf{B)} the
-symbiont to host biomass ratio (CmolS CmolH\textsuperscript{-1}),
-\textbf{C)} relative C- or N-limitation of host biomass formation, and
-\textbf{D)} relative CO\textsubscript{2}- or light-limitation of
-symbiont photosynthesis, across gradients of external irradiance and
-dissolved inorganic nitrogen. Note that typical conditions for reefs are
-\textasciitilde{}1e-7 M DIN and 10-20 mol photons m\textsuperscript{-2}
-d\textsuperscript{-1}. Simulations for each combination of light and
-nutrients (41 points along each axis) were run for 100 days with a time
-step of 1 day. Negative steady state growth rates, and corresponding S:H
-ratios, were set to zero.}
+\caption{Steady state values of (A) specific growth (Cmol
+Cmol\textsuperscript{-1} d\textsuperscript{-1}) and (B) the symbiont to
+host biomass ratio (CmolS CmolH\textsuperscript{-1}) across gradients of
+external irradiance and dissolved inorganic nitrogen. Note that typical
+conditions for reefs are \textasciitilde{}1e-7 M DIN and 10-20 mol
+photons m\textsuperscript{-2} d\textsuperscript{-1}. Simulations for
+each combination of light and nutrients (41 points along each axis) were
+run for 100 days with a time step of 1 day. All parameters were set to
+default values and prey density set to zero. Negative steady state
+growth rates, and corresponding S:H ratios, were set to zero.}
 \end{figure}
@@
 mol photons m\textsuperscript{-2} d\textsuperscript{-1}, based on
-Stimson (1997). The dynamic behavior of symiont to host biomass ratio
+Stimson (1997). The dynamic behavior of symbiont to host biomass ratio
 (middle panel) and the specific growth rate of host biomass (lower
@@
 densities are more susceptible to CO\textsubscript{2}-limitation
-(Wooldridge 2009). Several experimental (Cunning and Baker 2013; Vega
-Thurber et al. 2014) and correlational studies (Wooldridge and Done
-2009) are consistent with this mechanistic link between high nutrients
-and bleaching.
+(Wooldridge 2009). Several experimental (Wiedenmann et al. 2013; Cunning
+and Baker 2013; Vega Thurber et al. 2014) and correlational studies
+(Wooldridge and Done 2009) are consistent with this mechanistic link
+between high nutrients and bleaching.
 
@@
 
-Since bleaching is induced by an external stressor, recovery from
-bleaching cannot occur unless the stressor is alleviated. In natural
-settings, this typically occurs by seasonal declines in temperature and
-light. However, alternate stable states lead to significant hysteresis,
-indicating that a coral cannot recover along the same trajectory it
-followed during bleaching; indeed, the stressor must be alleviated well
+Since bleaching can be understood as a transition from a
+nitrogen-limited state with high S:H biomass to a carbon-limited state
+with low S:H biomass, induced by an external stressor, recovery can be
+understood as a switch back to the nitrogen-limited state once the
+external stressor is alleviated. In natural settings, this typically
+occurs through seasonal declines in temperature and light. However,
+hysteresis associated with the system's alternate stable states
+indicates that the symbiosis cannot recover along the same trajectory it
+followed during bleaching; indeed, the stressor must be alleviated
 \emph{below} the threshold that initially caused bleaching in order for
-the system to recover (Fig. 2; Fig. X). This is because under the same
-external conditions, a bleached coral (relative to a healthy one) is
-characterized by greater light amplification and weaker CCM activity,
-which serve to maintain the bleached state. The magnitude of hysteresis
-will depend on external nutrients and prey availability, as well as
-model parameters, yet the core behavior of hysteresis illustrates the
-alternate stable states that exist in the symbiotic system.
+the system to recover (Fig. 2). This is because under the same external
+conditions, a bleached coral with low S:H biomass (relative to a healthy
+coral with high S:H biomass) is characterized by greater light
+amplification and weaker CCM activity, which enhance photooxidative
+stress and serve to maintain the carbon-limited state. In order for the
+system to recover, the stressor must be reduced enough such that
+photooxidative stress ceases and translocation of carbon from symbiont
+to host is resumed. Once this occurs, the host can energize its CCMs,
+which further enhances carbon fixation and accelerates the system back
+toward a nitrogen-limited state, indicative of recovery. The conditions
+under which recovery can occur -- which determine the magnitude of
+hysteresis (Fig. SX) -- depend on the relative abundance of nitrogen and
+carbon in the environment. Higher food levels, representing a
+non-autotrophic carbon source for the host, make it easier for the host
+to overcome carbon-limitation, thus providing a potential mechanism
+underlying observations that feeding aids recovery from bleaching
+{[}xxx{]}. Conversely, high external DIN impedes the re-establishment of
+nitrogen-limitation, making recovery from bleaching more difficult
+(i.e., narrowing the sets of conditions under which recovery is
+possible).
+
+Dynamic simulations of recovery reveal another interesting behavior of
+the system: under some conditions, an `overshoot' occurs in which S:H
+biomass temporarily increases beyond the ratio maintained in the
+`healthy' state, before returning to stabilize at this value (Fig. 8;
+Fig. S9). In fact, unusually high symbiont densities have been observed
+following bleaching in both experimental (Cunning, Silverstein, and
+Baker 2015) and field studies (Kemp et al. 2014), and have been
+interpreted as a potential `disequilibrium in host-symbiont regulation'
+(Kemp et al. 2014). The model reveals the dynamics of this `overshoot'
+as follows: as symbionts repopulate the host, photosynthesis becomes
+increasingly CO\textsubscript{2}-limited due to weak CCM activity of the
+carbon-limited host. Thus, although symbiont growth is not yet
+carbon-limited, a growing symbiont population has less and less excess
+carbon (per symbiont) to share, and is thus moving toward
+carbon-limitation. Meanwhile, because S:H biomass is increasing, the
+host receives more and more carbon per unit host biomass, and is thus
+moving away from carbon-limitation. If the host overcomes
+carbon-limitation \emph{before} the symbiont reaches it, then the system
+rapidly transitions to the nitrogen-limited state and the S:H ratio
+stabilizes without an overshoot (Fig. 8A). However, if the symbiont
+becomes carbon-limited first (Fig. 8B, 8C), then carbon translocation
+per symbiont declines further and photosynthesis becomes more
+CO\textsubscript{2}-limited, which maintains carbon-limitation of the
+host. In this situation, continued growth of less and less productive
+symbionts causes S:H biomass to reach a much higher level before the
+host finally overcomes carbon-limitation. At this point, representing
+the peak of the overshoot, the transition to nitrogen-limitation finally
+takes place and the S:H biomass ratio declines and stabilizes.
+
+Whether this `overshoot' occurs or not represents a bifurcation point in
+the dynamical system that is defined by the relative availability of
+carbon and nitrogen to the host -- any factor that enhances
+carbon-limitation of host growth (e.g.~high DIN and/or low feeding)
+therefore magnifies the overshoot and prolongs the dysfunctional,
+carbon-limited state of the symbiosis (Fig. 8). On the other hand,
+factors that favor nitrogen-limitation, such as low external DIN and/or
+high feeding rates, will accelerate the re-establishment of
+nitrogen-limitation and prevent an overshoot from occurring at all.
+While many scenarios are possible under different environmental
+conditions, we illustrate the general effects of varying N and C
+availability on recovery from bleaching with a series of simulations
+that vary the N:C ratio of host's heterotrophic food source (Fig. 8A-C):
+lower N:C ratios (effectively representing lower DIN and/or higher
+heterotrophy) favor nitrogen-limitation and more rapid recovery, while
+higher N:C ratios (effectively representing higher DIN and/or lower
+heterotrophy) favor carbon-limitation and prolonged recovery with a
+larger overshoot in S:H biomass.
 
@@
 \includegraphics{../img/Fig8.png}
-\caption{Recovery from bleaching under varying N:C availability. Higher
-N:C ratios in a heterotrophic food source (effectively representing
-higher external DIN and/or lower feeding rates) cause a larger overshoot
-in the S:H biomass ratio, and prolong the duration of time until the
-system `recovers' by re-establishing nitrogen-limitation (represented by
-black dots on each trajectory). All simulations run with default
-parameters (except varying nNX), L=10 mol photons m\textsuperscript{2}
-d\textsuperscript{-1}, DIN=1e-7 molN L\textsuperscript{-1}, prey=1e-6
-CmolX L\textsuperscript{-1}, and initial S:H biomass=0.0001.}
+\caption{Recovery from bleaching with varying N:C availability (A:
+N:C=0.09; B: N:C=0.10; C: N:C=0.30). Higher N:C ratios in a
+heterotrophic food source (effectively representing higher external DIN
+and/or lower feeding rates) cause a larger overshoot in the S:H biomass
+ratio, and prolong the duration of time until the system `recovers' by
+re-establishing nitrogen-limitation. All simulations run with default
+parameters (except varying nNX), with L=10 mol photons
+m\textsuperscript{2} d\textsuperscript{-1}, DIN=1e-7 molN
+L\textsuperscript{-1}, prey=1e-6 CmolX L\textsuperscript{-1}, and
+initial S:H biomass=0.001.}
 \end{figure}
 
-The dynamics of recovery from bleaching reveal another interesting
-behavior of the system: under some conditions, the S:H biomass ratio may
-increase to a higher value that `overshoots' the ratio maintained in the
-`healthy' state under the same conditions (Fig. 8; Fig. S9). In fact,
-unusually high symbiont densities have been observed after recovery from
-bleaching in both experimental (Cunning, Silverstein, and Baker 2015)
-and field studies (Kemp et al. 2014), and have been interpreted as a
-potential `disequilibrium in host-symbiont regulation' (Kemp et al.
-2014). The model explains this `overshoot' as an approach to a
-bifurcation point in the dynamical system \textless{}\textgreater{}. As
-the symbiont begins repopulating the host, CO\textsubscript{2} is
-limiting and nitrogen is replete, allowing nearly all fixed carbon to be
-used for symbiont growth. As a result, host growth remains strongly
-carbon-limited, CCM activity remains weak, and photosynthesis remains
-CO\textsubscript{2}-limited. As long as this carbon-limitation persists,
-S:H biomass will continue to increase beyond the `healthy' steady state
-level, until the symbiont population eventually reaches a high enough
-density such that nitrogen limits its growth more strongly than carbon.
-At this point, which represents the peak of the `overshoot', the
-symbiont begins to share carbon with the host, which then supplies more
-CO\textsubscript{2} for photosynthesis, etc., causing the system to
-rapidly escape the positive feedbacks that were maintaining
-carbon-limitation (see Fig. S9 for illustration of all fluxes during
-overshoot). The re-establishment of nitrogen-limitation resumes high
-rates of carbon fixation and coral growth, bringing the S:H biomass
-ratio back down to a stable equilibrium.
-
-The occurrence and the magnitude of this overshoot depend on the
-relative inputs of carbon and nitrogen to the system -- any factor that
-enhances carbon-limitation (e.g.~high DIN and/or low feeding) therefore
-magnifies the overshoot and prolongs the dysfunctional, carbon-limited
-state of the symbiosis (Fig. 8). On the other hand, factors that favor
-nitrogen-limitation, such as low external DIN and/or high feeding rates,
-will accelerate the re-establishment of nitrogen-limitation and prevent
-an overshoot from occurring at all. While many scenarios are possible
-under different environmental conditions, we illustrate the general
-effects of varying N and C availability on recovery from bleaching with
-a series of simulations that vary the N:C ratio of host's heterotrophic
-food source (Fig. 8): lower N:C ratios (effectively representing lower
-DIN and/or higher heterotrophy) favor nitrogen-limitation and more rapid
-recovery, while higher N:C ratios (effectively representing higher DIN
-and/or lower heterotrophy) favor carbon-limitation and prolonged
-recovery with a larger overshoot in S:H biomass.
-
-Importantly, the re-establishment of nitrogen-limitation is the most
-important diagnostic of recovery to a `healthy' state, as this is when
-high carbon fixation and growth rates are resumed. A high S:H biomass
-ratio alone does not indicate that a `healthy' state has been reached,
-since overall C-limitation may still persist (e.g., high N:C ratios in
-Fig. 8). This could explain why corals that have apparently recovered
-their symbiont populations after bleaching still have reduced growth and
-reproduction rates until much later (xxx, Hughes). Therefore,
-acquisition of carbon from a source other than the symbiont may be
-extremely important for the host in order to recover from bleaching.
-Indeed, feeding has been shown to promote more rapid recovery from
-bleaching (xxx). Moreover, if the host obtains carbon from another
-independent source, such as direct uptake of DOC (xxx), this could be
-even more important in ensuring rapid recovery from a bleached state.
+These dynamics reveal that the re-establishment of nitrogen-limitation
+is the most important diagnostic of recovery to a `healthy' state, as
+this is when positive growth rates are resumed. A high S:H biomass ratio
+alone does not necessarily indicate that a `healthy' state has been
+reached, since the C-limited state may still persist (e.g., Fig. 8B,
+8C). This could explain why corals that have recovered their symbiont
+populations after bleaching may still exhibit energetic deficits and
+physiological impacts for much longer periods of time (Levitan et al.
+2014, Hughes and Grottoli (2013)). Therefore, host acquisition of carbon
+from a source other than the symbiont may be extremely important for the
+system to recover from bleaching. Indeed, host feeding has been shown to
+promote a more rapid return to pre-bleaching levels of key physiological
+parameters in recovering corals (Connolly, Lopez-Yglesias, and Anthony
+2012). The model predicts that acquisition of carbon from other sources,
+such as direct uptake of DOC (Levas et al. 2015), may also promote more
+rapid recovery from a bleached state.
 
@@
 
+\hypertarget{ref-Connolly:2012ep}{}
+Connolly, Sean R, M A Lopez-Yglesias, and Kenneth R N Anthony. 2012.
+``Food availability promotes rapid recovery from thermal stress in a
+scleractinian coral.'' \emph{Coral Reefs} 31 (4). Springer-Verlag:
+951--60.
+doi:\href{https://doi.org/10.1007/s00338-012-0925-9}{10.1007/s00338-012-0925-9}.
+
 \hypertarget{ref-Costanza:2014ex}{}
@@
 
+\hypertarget{ref-Hughes:2013hd}{}
+Hughes, Adam D, and Andréa G Grottoli. 2013. ``Heterotrophic
+Compensation: A Possible Mechanism for Resilience of Coral Reefs to
+Global Warming or a Sign of Prolonged Stress?'' \emph{PLoS ONE} 8 (11).
+Public Library of Science: e81172.
+doi:\href{https://doi.org/10.1371/journal.pone.0081172}{10.1371/journal.pone.0081172}.
+
 \hypertarget{ref-Huston:1985ec}{}
@@
 (July): 187--92.
+
+\hypertarget{ref-Levas:2015ic}{}
+Levas, Stephen, Andréa G Grottoli, Verena Schoepf, Matthew D
+Aschaffenburg, Justin Baumann, James E Bauer, and Mark E Warner. 2015.
+``Can heterotrophic uptake of dissolved organic carbon and zooplankton
+mitigate carbon budget deficits in annually bleached corals?''
+\emph{Coral Reefs} 35 (2). Springer Berlin Heidelberg: 495--506.
+doi:\href{https://doi.org/10.1007/s00338-015-1390-z}{10.1007/s00338-015-1390-z}.
+
+\hypertarget{ref-Levitan:2014et}{}
+Levitan, D R, W Boudreau, J Jara, and Nancy Knowlton. 2014. ``Long-term
+reduced spawning in Orbicella coral species due to temperature stress.''
+\emph{Marine Ecology Progress Series} 515: 1--10.
+doi:\href{https://doi.org/10.3354/meps11063}{10.3354/meps11063}.
 
@@
 
+\hypertarget{ref-Wiedenmann:2013bn}{}
+Wiedenmann, Jörg, C D'Angelo, Edward G Smith, Alan N Hunt, François-Eric
+Legiret, Anthony D Postle, and Eric P Achterberg. 2013. ``Nutrient
+enrichment can increase the susceptibility of reef corals to
+bleaching.'' \emph{Nature Climate Change} 3. Nature Publishing Group:
+160--64.
+doi:\href{https://doi.org/10.1038/nclimate1661}{10.1038/nclimate1661}.
+
 \hypertarget{ref-Wooldridge:2009p7807}{}

### 2017-01-11 JTB_ms/JTB_ms.Rmd
complete draft

@@
 Likewise, there is always a non-zero rejection flux of excess carbon from the coral biomass SU.
-Likewise, the carbon rejected from this SU reflects the amount of excess fixed carbon available to the host that is not used in biomass formation:
+The carbon rejected from this SU reflects the amount of excess fixed carbon available to the host that is not used in biomass formation:
 
@@
 By experimentation, we found that a time step of 0.1 days gave adequate precision for most simulations (including used to generate Figs.
-XXX in this paper).
-For steady state estimations, simulations were run until the change in specific growth rate of host and symbiont and the $S:H$ biomass ratio was less than 1e-5 per time step.
+2-8 in this paper).
+For steady state estimations, simulations were run until the change in specific growth rate of host and symbiont and the S:H biomass ratio was less than 1e-5 per time step.
 In regions of state space where very slow transient dynamics could be expected (i.e.
@@
 However, if symbiont biomass is initially high (i.e., a "healthy" coral), then the system remains in a nitrogen-limited state with positive growth.
-For practical purposes, this section of the manuscript considers only positive growth steady states under constant environments; the subsequent "dynamic behavior" section explores how changing environments may cause the system to switch between alternate stable states, which we interpret in the context of coral bleaching.
+For practical purposes, this section of the manuscript considers only positive growth steady states under constant environments; subsequently, we explore how environmental forcing may cause the system to switch between alternate stable states, which we interpret in the context of coral bleaching (see "Coral Bleaching and Recovery", below).
 
@@
 
-![Steady state values of **A)** specific growth (Cmol Cmol^-1^ d^-1^), **B)** the symbiont to host biomass ratio (CmolS CmolH^-1^), **C)** relative C- or N-limitation of host biomass formation, and **D)** relative CO~2~- or light-limitation of symbiont photosynthesis, across gradients of external irradiance and dissolved inorganic nitrogen.
+![Steady state values of (A) specific growth (Cmol Cmol^-1^ d^-1^) and (B) the symbiont to host biomass ratio (CmolS CmolH^-1^) across gradients of external irradiance and dissolved inorganic nitrogen.
 Note that typical conditions for reefs are ~1e-7 M DIN and 10-20 mol photons m^-2^ d^-1^.
 Simulations for each combination of light and nutrients (41 points along each axis) were run for 100 days with a time step of 1 day.
+All parameters were set to default values and prey density set to zero.
 Negative steady state growth rates, and corresponding S:H ratios, were set to zero.](../img/Fig3.png)
@@
 Light input (upper panel) was designed as a sinusoidal curve with a period of one year, with maximum and minimum values of 44 and 22 mol photons m^-2^ d^-1^, based on @Stimson:1997p3837.
-The dynamic behavior of symiont to host biomass ratio (middle panel) and the specific growth rate of host biomass (lower panel) show seasonal oscillations that are greater in magnitude under higher nutrients (1e-5 molN L^-1^; dashed lines) relative to lower nutrients (1e-7 molN L^-1^; solid lines), consistent with the findings of @Stimson:1997p3837.
+The dynamic behavior of symbiont to host biomass ratio (middle panel) and the specific growth rate of host biomass (lower panel) show seasonal oscillations that are greater in magnitude under higher nutrients (1e-5 molN L^-1^; dashed lines) relative to lower nutrients (1e-7 molN L^-1^; solid lines), consistent with the findings of @Stimson:1997p3837.
 Prey density was set at 1e-5 CmolX L^-1^.](../img/Fig5.png)
@@
 Decreasing photosynthesis moves the system from overall nitrogen-limitation toward carbon-limitation (D); when this threshold is crossed, the system rapidly becomes highly carbon-limited as photosynthesis becomes CO~2~-limited (E) and symbiont densities rapidly decline (F) into a bleached state.
-(All parameters at default values; external DIN=1e-7 molN/L, prey density=0).](../img/Fig6.png)
+Substrate limitation coefficients were calculated using Eq.
+22.
+All parameters were set at default values with external DIN=1e-7 molN/L and prey density=0).](../img/Fig6.png)
 
@@
 7), since higher symbiont densities are more susceptible to CO~2~-limitation [@Wooldridge:2009p7807].
-Several experimental [@Cunning:2013gp; @VegaThurber:2014hk] and correlational studies [@Wooldridge:2009kb] are consistent with this mechanistic link between high nutrients and bleaching.
+Several experimental [@Wiedenmann:2013bn; @Cunning:2013gp; @VegaThurber:2014hk] and correlational studies [@Wooldridge:2009kb] are consistent with this mechanistic link between high nutrients and bleaching.
 
@@
 
-Since bleaching is induced by an external stressor, recovery from bleaching cannot occur unless the stressor is alleviated.
-In natural settings, this typically occurs by seasonal declines in temperature and light.
-However, alternate stable states lead to significant hysteresis, indicating that a coral cannot recover along the same trajectory it followed during bleaching; indeed, the stressor must be alleviated well *below* the threshold that initially caused bleaching in order for the system to recover (Fig.
-2; Fig.
-X).
-This is because under the same external conditions, a bleached coral (relative to a healthy one) is characterized by greater light amplification and weaker CCM activity, which serve to maintain the bleached state.
-The magnitude of hysteresis will depend on external nutrients and prey availability, as well as model parameters, yet the core behavior of hysteresis illustrates the alternate stable states that exist in the symbiotic system.
+Since bleaching can be understood as a transition from a nitrogen-limited state with high S:H biomass to a carbon-limited state with low S:H biomass, induced by an external stressor, recovery can be understood as a switch back to the nitrogen-limited state once the external stressor is alleviated.
+In natural settings, this typically occurs through seasonal declines in temperature and light.
+However, hysteresis associated with the system's alternate stable states indicates that the symbiosis cannot recover along the same trajectory it followed during bleaching; indeed, the stressor must be alleviated *below* the threshold that initially caused bleaching in order for the system to recover (Fig.
+2).
+This is because under the same external conditions, a bleached coral with low S:H biomass (relative to a healthy coral with high S:H biomass) is characterized by greater light amplification and weaker CCM activity, which enhance photooxidative stress and serve to maintain the carbon-limited state.
+In order for the system to recover, the stressor must be reduced enough such that photooxidative stress ceases and translocation of carbon from symbiont to host is resumed.
+Once this occurs, the host can energize its CCMs, which further enhances carbon fixation and accelerates the system back toward a nitrogen-limited state, indicative of recovery.
+The conditions under which recovery can occur -- which determine the magnitude of hysteresis (Fig.
+S9) -- depend on the relative abundance of nitrogen and carbon in the environment.
+Higher food levels, representing a non-autotrophic carbon source for the host, make it easier for the host to overcome carbon-limitation, thus providing a potential mechanism underlying observations that feeding aids recovery from bleaching [@Grottoli:2006p3692; @Connolly:2012ep].
+Conversely, high external DIN impedes the re-establishment of nitrogen-limitation, making recovery from bleaching more difficult (i.e., narrowing the sets of conditions under which recovery is possible).
 
-The dynamics of recovery from bleaching reveal another interesting behavior of the system: under some conditions, the S:H biomass ratio may increase to a higher value that 'overshoots' the ratio maintained in the 'healthy' state under the same conditions (Fig.
-8; Fig.
-S9).
-In fact, unusually high symbiont densities have been observed after recovery from bleaching in both experimental [@Cunning:2015ja] and field studies [@Kemp:2014ug], and have been interpreted as a potential 'disequilibrium in host-symbiont regulation' [@Kemp:2014ug].
-The model explains this 'overshoot' as an approach to a bifurcation point in the dynamical system <<NOT RIGHT>>.
-As the symbiont begins repopulating the host, CO~2~ is limiting and nitrogen is replete, allowing nearly all fixed carbon to be used for symbiont growth.
-As a result, host growth remains strongly carbon-limited, CCM activity remains weak, and photosynthesis remains CO~2~-limited.
-As long as this carbon-limitation persists, S:H biomass will continue to increase beyond the 'healthy' steady state level, until the symbiont population eventually reaches a high enough density such that nitrogen limits its growth more strongly than carbon.
-At this point, which represents the peak of the 'overshoot', the symbiont begins to share carbon with the host, which then supplies more CO~2~ for photosynthesis, etc., causing the system to rapidly escape the positive feedbacks that were maintaining carbon-limitation (see Fig.
-S9 for illustration of all fluxes during overshoot).
-The re-establishment of nitrogen-limitation resumes high rates of carbon fixation and coral growth, bringing the S:H biomass ratio back down to a stable equilibrium.
+Dynamic simulations of recovery reveal another interesting behavior of the system: under some conditions, an 'overshoot' occurs in which S:H biomass temporarily increases beyond the ratio maintained in the 'healthy' state, before returning to stabilize at this value (Fig.
+8).
+In fact, unusually high symbiont densities have been observed following bleaching in both experimental [@Cunning:2015ja] and field studies [@Kemp:2014ug], and have been interpreted as a potential 'disequilibrium in host-symbiont regulation' [@Kemp:2014ug].
+The model reveals the dynamics of this 'overshoot' as follows: as symbionts repopulate the host, photosynthesis becomes increasingly CO~2~-limited due to weak CCM activity of the carbon-limited host.
+Thus, although symbiont growth is not yet carbon-limited, a growing symbiont population has less and less excess carbon (per symbiont) to share, and is thus moving toward carbon-limitation.
+Meanwhile, because S:H biomass is increasing, the host receives more and more carbon per unit host biomass, and is thus moving away from carbon-limitation.
+If the host overcomes carbon-limitation *before* the symbiont reaches it, then the system rapidly transitions to the nitrogen-limited state and the S:H ratio stabilizes without an overshoot (Fig.
+8A).
+However, if the symbiont becomes carbon-limited first (Fig.
+8B, 8C), then carbon translocation per symbiont declines further and photosynthesis becomes more CO~2~-limited, which maintains carbon-limitation of the host.
+In this situation, continued growth of less and less productive symbionts drives S:H biomass to a much higher level before the host finally overcomes carbon-limitation.
+At this point, representing the peak of the overshoot, the transition to nitrogen-limitation finally takes place and the S:H biomass ratio declines and stabilizes as positive growth is resumed.
 
-![Recovery from bleaching under varying N:C availability.
-Higher N:C ratios in a heterotrophic food source (effectively representing higher external DIN and/or lower feeding rates) cause a larger overshoot in the S:H biomass ratio, and prolong the duration of time until the system 'recovers' by re-establishing nitrogen-limitation (represented by black dots on each trajectory).
-All simulations run with default parameters (except varying nNX), L=10 mol photons m^2^ d^-1^, DIN=1e-7 molN L^-1^, prey=1e-6 CmolX L^-1^, and initial S:H biomass=0.0001.](../img/Fig8.png)
-
-The occurrence and the magnitude of this overshoot depend on the relative inputs of carbon and nitrogen to the system -- any factor that enhances carbon-limitation (e.g.
+Whether this 'overshoot' occurs or not represents a bifurcation point in the dynamical system that is defined by the relative availability of carbon and nitrogen to the host -- any factor that enhances carbon-limitation of host growth (e.g.
 high DIN and/or low feeding) therefore magnifies the overshoot and prolongs the dysfunctional, carbon-limited state of the symbiosis (Fig.
@@
 While many scenarios are possible under different environmental conditions, we illustrate the general effects of varying N and C availability on recovery from bleaching with a series of simulations that vary the N:C ratio of host's heterotrophic food source (Fig.
-8): lower N:C ratios (effectively representing lower DIN and/or higher heterotrophy) favor nitrogen-limitation and more rapid recovery, while higher N:C ratios (effectively representing higher DIN and/or lower heterotrophy) favor carbon-limitation and prolonged recovery with a larger overshoot in S:H biomass.
+8A-C): lower N:C ratios (effectively representing lower DIN and/or higher heterotrophy) favor nitrogen-limitation and more rapid recovery, while higher N:C ratios (effectively representing higher DIN and/or lower heterotrophy) favor carbon-limitation and prolonged recovery with a larger overshoot in S:H biomass.
 
-Importantly, the re-establishment of nitrogen-limitation is the most important diagnostic of recovery to a 'healthy' state, as this is when high carbon fixation and growth rates are resumed.
-A high S:H biomass ratio alone does not indicate that a 'healthy' state has been reached, since overall C-limitation may still persist (e.g., high N:C ratios in Fig.
-8).
-This could explain why corals that have apparently recovered their symbiont populations after bleaching still have reduced growth and reproduction rates until much later (xxx, Hughes).
-Therefore, acquisition of carbon from a source other than the symbiont may be extremely important for the host in order to recover from bleaching.
-Indeed, feeding has been shown to promote more rapid recovery from bleaching (xxx).
-Moreover, if the host obtains carbon from another independent source, such as direct uptake of DOC (xxx), this could be even more important in ensuring rapid recovery from a bleached state.
+![Recovery from bleaching with varying N:C availability (A: N:C=0.09; B: N:C=0.10; C: N:C=0.30).
+Higher N:C ratios in a heterotrophic food source (effectively representing higher external DIN and/or lower feeding rates) cause a larger overshoot in the S:H biomass ratio, and prolong the duration of time until the system 'recovers' by re-establishing nitrogen-limitation.
+Substrate limitation coefficients were calculated using Eq.
+22.
+All simulations were run with default parameters (except varying $n_{NX}$), with L=10 mol photons m^2^ d^-1^, DIN=1e-7 molN L^-1^, prey=1e-6 CmolX L^-1^, and initial S:H biomass=0.001.](../img/Fig8.png)
+
+These dynamics reveal that the re-establishment of nitrogen-limitation is the most important diagnostic of recovery to a 'healthy' state, as this is when positive growth rates are resumed.
+A high S:H biomass ratio alone does not necessarily indicate that a 'healthy' state has been reached, since the C-limited state may still persist (e.g., Fig.
+8B, 8C).
+This could explain why corals that have recovered their symbiont populations after bleaching may still exhibit energetic deficits and physiological impacts for much longer periods of time [@Levitan:2014et, @Hughes:2013hd].
+Therefore, host acquisition of carbon from a source other than the symbiont may be extremely important for the system to recover from bleaching.
+Indeed, host feeding has been shown to promote a more rapid return to pre-bleaching levels of key physiological parameters in recovering corals [@Connolly:2012ep].
+The model predicts that acquisition of carbon from other sources, such as direct uptake of DOC [@Levas:2015ic], may also promote more rapid recovery from a bleached state.
 
@@
 This model also provides a new conceptual framework for considering coral bleaching as a transition to an alternate stable state, which has important implications for understanding the performance and maintenance of symbiotic interactions.
-In this context, the 'healthy' stable state represents a scenario in which the host is 'in control' of the system, with nitrogen-limitation stabilizing the symbiont to host biomass ratio and maintaining positive growth.
-Conversely, carbon-limitation represents a loss of host 'control', wherein positive feedbacks result in a loss of capacity for photosynthetic carbon fixation and positive growth.
+In this context, the 'healthy' stable state represents a scenario in which nitrogen-limitation stabilizes the symbiont to host biomass ratio and maintains positive growth.
+Conversely, carbon-limitation represents a dysfunctional state wherein positive feedbacks result in a loss of photosynthetic capacity and negative growth.
 Importantly, these 'control' mechanisms arise entirely passively through feedbacks present in the system.
@@
 Moreover, parameter values can be modified to correspond to different genetic or functional types of coral hosts and *Symbiodinium* partners in order to evaluate variability in system responses.
-Thus, the diversity of potential applications for this model is high, and we envision this work as a foundation for continued development, which may include more detailed treatments of specific modules (e.g., DIC processing), and the incorporation of more external forcing capabilities (e.g., external DIC).
-Importantly, this effort is best undertaken by the scientific community at large, including those with theoretical and empirical backgrounds, which is why the code for this model has been developed openly using the R language.
+Thus, the diversity of potential applications for this model is high, and we envision this work as a foundation for continued development, which may include more detailed treatments of specific modules (e.g., DIC processing), and the incorporation of more external forcing capabilities (e.g., external DIC, temperature).
+Importantly, this effort may be undertaken by the scientific community at large, including those with theoretical and empirical backgrounds, which is why the code for this model has been developed openly using the R language.
 Ultimately, the continued refinement of these tools is fundamental in elucidating the mechanisms of symbiosis function and dysfunction, and in predicting coral responses to environmental change.

### 2017-01-11 JTB_supp/JTB_supp.Rmd
Fig S9 light range 10-55
fix xlim in Fig S9

@@
 # Set input values for steady state runs
-input <- expand.grid(L=seq(20,55,1), initS=c(0.0001,0.1))
+input <- expand.grid(L=seq(10,55,1), initS=c(0.0001,0.1))
 
@@
 
-plot(NA, xlim=c(20,55), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
+plot(NA, xlim=range(input$L), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
 points(c(48,48,53,53), c(0.22,0.23,0.22,0.23), pch=c(19,1,19,1), cex=c(0.4,1,0.4,1), col=c("black","black","red","red"), xpd=T)
@@
 
-plot(NA, xlim=c(20,55), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
+plot(NA, xlim=range(input$L), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
 apply(res2, 1, FUN=function(x) {
@@
 
-plot(NA, xlim=c(20,55), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
+plot(NA, xlim=range(input$L), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
 apply(res3, 1, FUN=function(x) {
@@
 
-plot(NA, xlim=c(20,55), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
+plot(NA, xlim=range(input$L), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
 apply(res4, 1, FUN=function(x) {

### 2017-01-11 JTB_supp/JTB_supp.Rmd
separate FigS9 calc and plot

@@
 # Figure S9. Hysteresis under different nutrient and feeding regimes.
-```{r FigS9, cache=T, fig.width=7, fig.height=7, echo=F, message=F}
+```{r runFigS9, cache=T, include=F}
 # Set up cluster for parallel processing
@@
 res4 <- cbind(input, output4)
+```
 
+```{r plotFigS9, cache=T, fig.width=7, fig.height=7, echo=F, message=F}
 # Create plot

### 2017-01-11 JTB_supp/JTB_supp.Rmd
title and caption Fig S9

@@
 plot(NA, xlim=range(input$L), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
+title("DIN=1e-7 mol/L; X=2e-6 mol/L")
 points(c(48,48,53,53), c(0.22,0.23,0.22,0.23), pch=c(19,1,19,1), cex=c(0.4,1,0.4,1), col=c("black","black","red","red"), xpd=T)
@@
 plot(NA, xlim=range(input$L), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
+title("DIN=1e-7 mol/L; X=1e-6 mol/L")
 apply(res2, 1, FUN=function(x) {
@@
 plot(NA, xlim=range(input$L), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
+title("DIN=1e-6 mol/L; X=2e-6 mol/L")
 apply(res3, 1, FUN=function(x) {
@@
 plot(NA, xlim=range(input$L), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
+title("DIN=1e-6 mol/L; X=4e-6 mol/L")
 apply(res4, 1, FUN=function(x) {
@@
 
+These figures show steady state S:H biomass ratios in constant environments with varying DIN and food (X) levels, with initial S:H biomass values of either 0.1 or 0.0001.
+Points are colored according to whether the host exhibits positive (black) or negative (red) growth at steady state.
+
 ## References

### 2017-01-11 JTB_ms/JTB_ms.tex
finish draft

@@
 there is always a non-zero rejection flux of excess carbon from the
-coral biomass SU. Likewise, the carbon rejected from this SU reflects
-the amount of excess fixed carbon available to the host that is not used
-in biomass formation:
+coral biomass SU. The carbon rejected from this SU reflects the amount
+of excess fixed carbon available to the host that is not used in biomass
+formation:
 
@@
 step of 0.1 days gave adequate precision for most simulations (including
-used to generate Figs. XXX in this paper). For steady state estimations,
+used to generate Figs. 2-8 in this paper). For steady state estimations,
 simulations were run until the change in specific growth rate of host
-and symbiont and the \(S:H\) biomass ratio was less than 1e-5 per time
-step. In regions of state space where very slow transient dynamics could
-be expected (i.e.~near bifurcation points), sample steady state
+and symbiont and the S:H biomass ratio was less than 1e-5 per time step.
+In regions of state space where very slow transient dynamics could be
+expected (i.e.~near bifurcation points), sample steady state
 calculations were verified using MATHEMATICA code for numerical root
@@
 as photosynthesis becomes CO\textsubscript{2}-limited (E) and symbiont
-densities rapidly decline (F) into a bleached state. (All parameters at
-default values; external DIN=1e-7 molN/L, prey density=0).}
+densities rapidly decline (F) into a bleached state. Substrate
+limitation coefficients were calculated using Eq. 22. All parameters
+were set at default values with external DIN=1e-7 molN/L and prey
+density=0).}
 \end{figure}
@@
 under which recovery can occur -- which determine the magnitude of
-hysteresis (Fig. SX) -- depend on the relative abundance of nitrogen and
+hysteresis (Fig. S9) -- depend on the relative abundance of nitrogen and
 carbon in the environment. Higher food levels, representing a
@@
 underlying observations that feeding aids recovery from bleaching
-{[}xxx{]}. Conversely, high external DIN impedes the re-establishment of
-nitrogen-limitation, making recovery from bleaching more difficult
-(i.e., narrowing the sets of conditions under which recovery is
-possible).
+(Grottoli, Rodrigues, and Palardy 2006; Connolly, Lopez-Yglesias, and
+Anthony 2012). Conversely, high external DIN impedes the
+re-establishment of nitrogen-limitation, making recovery from bleaching
+more difficult (i.e., narrowing the sets of conditions under which
+recovery is possible).
 
@@
 biomass temporarily increases beyond the ratio maintained in the
-`healthy' state, before returning to stabilize at this value (Fig. 8;
-Fig. S9). In fact, unusually high symbiont densities have been observed
-following bleaching in both experimental (Cunning, Silverstein, and
-Baker 2015) and field studies (Kemp et al. 2014), and have been
-interpreted as a potential `disequilibrium in host-symbiont regulation'
-(Kemp et al. 2014). The model reveals the dynamics of this `overshoot'
-as follows: as symbionts repopulate the host, photosynthesis becomes
-increasingly CO\textsubscript{2}-limited due to weak CCM activity of the
+`healthy' state, before returning to stabilize at this value (Fig. 8).
+In fact, unusually high symbiont densities have been observed following
+bleaching in both experimental (Cunning, Silverstein, and Baker 2015)
+and field studies (Kemp et al. 2014), and have been interpreted as a
+potential `disequilibrium in host-symbiont regulation' (Kemp et al.
+2014). The model reveals the dynamics of this `overshoot' as follows: as
+symbionts repopulate the host, photosynthesis becomes increasingly
+CO\textsubscript{2}-limited due to weak CCM activity of the
 carbon-limited host. Thus, although symbiont growth is not yet
@@
 host. In this situation, continued growth of less and less productive
-symbionts causes S:H biomass to reach a much higher level before the
-host finally overcomes carbon-limitation. At this point, representing
-the peak of the overshoot, the transition to nitrogen-limitation finally
-takes place and the S:H biomass ratio declines and stabilizes.
+symbionts drives S:H biomass to a much higher level before the host
+finally overcomes carbon-limitation. At this point, representing the
+peak of the overshoot, the transition to nitrogen-limitation finally
+takes place and the S:H biomass ratio declines and stabilizes as
+positive growth is resumed.
 
@@
 ratio, and prolong the duration of time until the system `recovers' by
-re-establishing nitrogen-limitation. All simulations run with default
-parameters (except varying nNX), with L=10 mol photons
+re-establishing nitrogen-limitation. Substrate limitation coefficients
+were calculated using Eq. 22. All simulations were run with default
+parameters (except varying \(n_{NX}\)), with L=10 mol photons
 m\textsuperscript{2} d\textsuperscript{-1}, DIN=1e-7 molN
@@
 In this context, the `healthy' stable state represents a scenario in
-which the host is `in control' of the system, with nitrogen-limitation
-stabilizing the symbiont to host biomass ratio and maintaining positive
-growth. Conversely, carbon-limitation represents a loss of host
-`control', wherein positive feedbacks result in a loss of capacity for
-photosynthetic carbon fixation and positive growth. Importantly, these
+which nitrogen-limitation stabilizes the symbiont to host biomass ratio
+and maintains positive growth. Conversely, carbon-limitation represents
+a dysfunctional state wherein positive feedbacks result in a loss of
+photosynthetic capacity and negative growth. Importantly, these
 `control' mechanisms arise entirely passively through feedbacks present
@@
 modules (e.g., DIC processing), and the incorporation of more external
-forcing capabilities (e.g., external DIC). Importantly, this effort is
-best undertaken by the scientific community at large, including those
-with theoretical and empirical backgrounds, which is why the code for
-this model has been developed openly using the R language. Ultimately,
-the continued refinement of these tools is fundamental in elucidating
-the mechanisms of symbiosis function and dysfunction, and in predicting
-coral responses to environmental change.
+forcing capabilities (e.g., external DIC, temperature). Importantly,
+this effort may be undertaken by the scientific community at large,
+including those with theoretical and empirical backgrounds, which is why
+the code for this model has been developed openly using the R language.
+Ultimately, the continued refinement of these tools is fundamental in
+elucidating the mechanisms of symbiosis function and dysfunction, and in
+predicting coral responses to environmental change.
 
@@
 \emph{Limnology and Oceanography} 45 (3): 677--85.
+
+\hypertarget{ref-Grottoli:2006p3692}{}
+Grottoli, Andréa G, Lisa J Rodrigues, and James E Palardy. 2006.
+``Heterotrophic plasticity and resilience in bleached corals.''
+\emph{Nature} 440 (7088): 1186--9.
+doi:\href{https://doi.org/10.1038/nature04565}{10.1038/nature04565}.
 

### 2017-01-11 JTB_supp/JTB_supp.Rmd
add large overshoot points to top FigS9

@@
                  cex=ifelse(initS==1e-01, 0.4, 1)))
+  with(x[x$sh>0.3, ], {
+    points(L, rep(0.3, length(L)), pch="^", col="red")
+    points(L, rep(0.29, length(L)), pch=1, col="red", cex=1)
+  })
 })

### 2017-01-11 JTB_supp/JTB_supp.Rmd
fix plotting error

@@
                  cex=ifelse(initS==1e-01, 0.4, 1)))
-  with(x[x$sh>0.3, ], {
-    points(L, rep(0.3, length(L)), pch="^", col="red")
-    points(L, rep(0.29, length(L)), pch=1, col="red", cex=1)
+  with(x, if (sh > 0.3) {
+    points(L, 0.3, pch="^", col="red")
+    points(L, 0.29, pch=1, col="red", cex=1)
   })

### 2017-01-11 JTB_supp/JTB_supp.Rmd
rerun Fig S9 from 1-60

@@
 # Set input values for steady state runs
-input <- expand.grid(L=seq(10,55,1), initS=c(0.0001,0.1))
+input <- expand.grid(L=seq(1,60,1), initS=c(0.0001,0.1))
 
@@
 # Create plot
-par(mfrow=c(2,2), mar=c(3,3,1,1))
+par(mfrow=c(2,2), mar=c(3,3,1,1), mgp=c(1.5,0.5,0), tcl=-0.025)
 

### 2017-01-11 JTB_supp/JTB_supp.Rmd
change Fig S9 ylim to 0-0.5

@@
 
-plot(NA, xlim=range(input$L), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
+plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab="Light", ylab="Steady state S:H ratio")
 title("DIN=1e-7 mol/L; X=2e-6 mol/L")
@@
                  cex=ifelse(initS==1e-01, 0.4, 1)))
-  with(x, if (sh > 0.3) {
-    points(L, 0.3, pch="^", col="red")
-    points(L, 0.29, pch=1, col="red", cex=1)
+  with(x, if (sh > 0.5) {
+    points(L, 0.5, pch="^", 
+           col=ifelse(gr>0, "black", "red"))
+    points(L, 0.49, 
+           col=ifelse(gr>0, "black", "red"),
+           pch=ifelse(initS==1e-01, 19, 1), 
+           cex=ifelse(initS==1e-01, 0.4, 1))
   })
@@
 
-plot(NA, xlim=range(input$L), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
+plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab="Light", ylab="Steady state S:H ratio")
 title("DIN=1e-7 mol/L; X=1e-6 mol/L")
@@
 
-plot(NA, xlim=range(input$L), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
+plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab="Light", ylab="Steady state S:H ratio")
 title("DIN=1e-6 mol/L; X=2e-6 mol/L")
@@
 
-plot(NA, xlim=range(input$L), ylim=c(0,0.3), xlab="Light", ylab="Steady state S:H ratio")
+plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab="Light", ylab="Steady state S:H ratio")
 title("DIN=1e-6 mol/L; X=4e-6 mol/L")

### 2017-01-11 JTB_supp/JTB_supp.Rmd
reposition Fig S9 legend

@@
 title("DIN=1e-7 mol/L; X=2e-6 mol/L")
-points(c(48,48,53,53), c(0.22,0.23,0.22,0.23), pch=c(19,1,19,1), cex=c(0.4,1,0.4,1), col=c("black","black","red","red"), xpd=T)
-text(38, 0.225, labels="init. S:H", xpd=T, cex=0.7, srt=90)
-text(46, c(0.22,0.23), labels=c("0.1", "0.0001"), xpd=T, pos=2, cex=0.7)
-text(50, 0.25, labels=c("Steady state growth"), xpd=T, cex=0.7)
-text(c(48,53), 0.24, labels=c("pos.", "neg."), xpd=T, cex=0.7)
+points(c(48,48,53,53), c(0.42,0.45,0.42,0.45), pch=c(19,1,19,1), cex=c(0.4,1,0.4,1), col=c("black","black","red","red"), xpd=T)
+text(38, 0.435, labels="init. S:H", xpd=T, cex=0.7, srt=90)
+text(46, c(0.42,0.45), labels=c("0.1", "0.0001"), xpd=T, pos=2, cex=0.7)
+text(50, 0.49, labels=c("Steady state growth"), xpd=T, cex=0.7)
+text(c(48,53), 0.47, labels=c("pos.", "neg."), xpd=T, cex=0.7)
 apply(res1, 1, FUN=function(x) {

### 2017-02-07 R/sym_PI.R
add y-axis units

@@ -30,11 +30,14 @@ sym_PI <- function(pars) {
     }
     # Return ROS and photosynthesis rate for plotting
     par(mfrow=c(1,1), mar=c(3,3,3,7), mgp=c(1.2,0,0), cex=1, tck=0.025, xaxs="i")
-    plot(jL, jCP/pars$yCL, xlab="Light (mol photons/C-molS/d)", ylab="Photochemical quenching",
+    plot(jL, jCP/pars$yCL, xlab="Light (mol photons/C-molS/d)", ylab="",
          type="l", lwd=3, col="red")
+    mtext(side=2, text="Photochemical quenching", cex=1, line=1.8)
+    mtext(side=2, text="(mol photons/CmolS/d)", cex=0.8, line=1)
     par(new=T)
     plot(jL, jNPQ, type="l", lwd=3, axes=F, xlab="", ylab="")
     axis(side=4); mtext(side=4, text="Non-photochemical quenching", line=1, cex=1)
+                  mtext(side=4, text="(mol photons/CmolS/d)", line=1.8, cex=0.8)
     par(new=T)
     plot(jL, cROS, type="l", lwd=3, axes=F, col="orange", xlab="", ylab="", ylim=c(1, max(cROS)*1.5))
     axis(side=4, line=4); mtext(side=4, line=5, text = "ROS production (relative)", cex=1)

### 2017-02-07 R/conversions.R
delete unused

@@ -25,5 +25,4 @@ nmolchlA.cm2 <- 9.66  # Edmunds et al. 2011
 CmolR.m2_Gfas <- 2  # Schutter et al. 2010 [based on reported 6 mg afdw cm-2 (Fig. 9) and 41.4% Carbon by weight (Results section)]
 gC.artemia <- 0.678e-6  # Wijgerde et al. 2011 JEB
 pgC.cellS <- 234.5 # Average of two clade A phylotypes in Brading et al. 2013
-#pgC.cellS <- 100 # Average of two types from Cassiopeia in Verde & McCloskey 1998
 CmolS.cellS <- pgC.cellS / pg.g / gC.molC

### 2017-02-07 JTB_ms/JTB_ms.Rmd
address EM + RN comments

@@
 ---
-title: A simplified dynamic bioenergetic model for coral-*Symbiodinium* symbioses and coral bleaching as an alternate stable state
+title: A dynamic bioenergetic model for coral-*Symbiodinium* symbioses and coral bleaching as an alternate stable state
 author:
@@
   - name: Erik B. Muller
-    affiliation: University of California
+    affiliation: UCSB, MSI
   - name: Ruth D. Gates
-    affiliation: University of Hawaii
+    affiliation: HIMB
   - name: Roger M. Nisbet
-    affiliation: University of California
+    affiliation: UCSB
 address:
-  - code: University of Hawaii
-    address: Hawaii Institute of Marine Biology, Kaneohe, HI 96744, USA
-  - code: University of California
-    address: Department of Ecology, Evolution, and Marine Biology, Santa Barbara, CA 93106, USA
+  - code: HIMB
+    address: Hawaii Institute of Marine Biology, School of Ocean and Earth Science and Technology, University of Hawaii at Mānoa, USA
+  - code: UCSB
+    address: Department of Ecology, Evolution, and Marine Biology, University of California, Santa Barbara, USA
+  - code: MSI
+    address: Marine Science Institute, University of California, Santa Barbara, USA
 abstract: |
   Coral reef ecosystems owe their ecological success--and vulnerability to climate change--to the symbiotic metabolism of corals and *Symbiodinium* spp.
-The urgency to understand and predict the stability and breakdown of these symbioses (i.e., coral 'bleaching') demands the development and application of theoretical tools; however, the complexity of some approaches may limit their potential application and utility.
-Here, we develop a simplified dynamic bioenergetic model of coral-*Symbiodinium* symbioses that demonstrates realistic steady-state patterns in coral growth and symbiont abundance across gradients of light, nutrients, and feeding.
-Furthermore, through a mechanistic treatment of photo-oxidative stress, the model displays dynamic behaviors of bleaching and recovery, and suggests that these phenomena represent transitions between alternate stable states.
-A suite of complex responses to multiple, interacting environmental factors reproduced by the model suggests it unifyingly captures many important attributes of the system; meanwhile, its modular framework and open source R code are designed to facilitate expansion and problem-specific applications.
-We see significant potential applications for this modeling framework in generating testable hypotheses and predicting integrated, mechanistic responses of corals to environmental change.
+The urgency to understand and predict the stability and breakdown of these symbioses (i.e., coral 'bleaching') demands the development and application of theoretical tools.
+Here, we develop a dynamic bioenergetic model of coral-*Symbiodinium* symbioses that demonstrates realistic steady-state patterns in coral growth and symbiont abundance across gradients of light, nutrients, and feeding.
+Furthermore, by including a mechanistic treatment of photo-oxidative stress, the model displays dynamics of bleaching and recovery that can be explained as transitions between alternate stable states.
+These dynamics reveal that 'healthy' and 'bleached' states correspond broadly to nitrogen- and carbon-limitation of the system, with transitions between them occurring as integrated responses to multiple environmental factors.
+Indeed, a suite of complex emergent behaviors reproduced by the model (e.g., bleaching is exacerbated by nutrients and attenuated by feeding) suggests it unifyingly captures many important attributes of the system; meanwhile, its modular framework and open source R code are designed to facilitate further problem-specific development.
+We see significant potential for this modeling framework to generate testable hypotheses and predict integrated, mechanistic responses of corals to environmental change, with important implications for understanding the performance and maintenance of symbiotic systems.
 bibliography: library.bib
@@
 
+MAKE INTRO INCLUDE SOMETHING ABOUT ALTERNATIVE STABLE STATES
+
 The nutritional exchange between corals and *Symbiodinium* directly underlies the capacity of corals to build coral reef ecosystems, worth trillions of US Dollars annually [@Costanza:2014ex].
@@
 Previous application of DEB theory to this syntrophic system [@Muller:2009io] demonstrated a stable symbiotic relationship and qualitatively realistic growth and biomass ratios across gradients of ambient irradiance, nutrients, and food.
-This model assumed that 1) *Symbiodinium* has priority access to fixed carbon through photosynthesis, 2) the coral animal has priority access to dissolved nitrogen through contact with seawater, and 3) each partner shares with the other only what it cannot use for its own growth.
+This model (as well as the present work) assumes that 1) *Symbiodinium* has priority access to fixed carbon through photosynthesis, 2) the coral animal has priority access to nitrogen through contact with seawater and feeding on prey, and 3) each partner shares with the other only what it cannot use for its own growth.
 In its simplest form, this principle of sharing the surplus is sufficient to describe the dynamics of diverse syntrophic organs and organisms (e.g., trees, duckweeds, corals), suggesting the mechanism is mathematically and evolutionarily robust (Nisbet et al., in prep.).
@@
 1.
-_Develop a module of photooxidative stress._ Of primary interest to coral biologists and ecologists is symbiosis dysfunction under environmental stress, resulting in coral “bleaching”--the loss of algal symbionts from the association [@Jokiel:1977p7353].
-Photooxidative stress in *Symbiodinium* is considered a primary trigger of bleaching in response to high temperature and/or light [@Weis:2008p944], and prolonged or severe bleaching can result in mortality, though corals sometimes recover their symbionts.
-Bleaching susceptibility, severity, and recovery may be influenced by interacting factors such as heterotrophy and nutrient availability [@Wooldridge:2014hc].
-To simulate these events and interactions, we develop a photooxidative stress module linking overreduction of the photosynthetic light reactions to downstream impacts of photoinhibition and photodamage.
+_Develop a new module of photo-oxidative stress._ Of primary interest to coral biologists and ecologists is symbiosis dysfunction under environmental stress, resulting in coral “bleaching”--the loss of algal symbionts from the association [@Jokiel:1977p7353].
+Bleaching is thought to be triggered by photo-oxidative stress in *Symbiodinium* [@Weis:2008p944], which has been modeled previously as a response to absolute external irradiance [@Eynaud:2011tv].
+However, this response may also depend on self-shading by other symbionts, CO~2~ availability at the site of photosynthesis, and other non-photochemical quenching mechanisms.
+We incorporate these features into a new photo-oxidative stress module linking overreduction of the photosynthetic light reactions to downstream impacts of photoinhibition and photodamage.
+Importantly, this formulation introduces a link between CO~2~-limitation of photosynthesis and bleaching, and potential synergistic roles for heterotrophy and nutrient availability in influencing bleaching responses [@Wooldridge:2014hc]
 
@@
 Here, we present a model without reserves, maturity, or reproduction [see @Kooijman:2010vd].
-This formulation restricts the model's scope to the bioenergetics of adult corals (i.e., reproduction, larval stages, and metamorphosis are not considered), but greatly reduces theoretical complexity and parameter numbers, which is advantageous given the relative paucity of data for corals.
-However, our primary motivation for reducing complexity was to increase accessibility and applicability for biologists and ecologists without requiring significant expertise in DEB theory.
+This formulation precludes modeling the full life cycle of corals as reproduction, larval stages, and metamorphosis are not considered, but greatly reduces theoretical complexity and parameter numbers, which is advantageous given the relative paucity of data for corals.
+Moreover, dynamics of the symbiosis (i.e., changes in symbiont to host biomass, including bleaching and recovery) and coral biomass growth are efficiently captured with this simpler formulation, which also increases accessibility for biologists and ecologists without requiring significant expertise in DEB theory.
 
@@
 _Provide well-documented, open-access code._ In order to facilitate the continued development and application of theoretical modeling tools for coral symbioses, we provide open access to the model in the form of detailed, commented code written in the R language [@RALanguageandEn:2014wf].
-With an accessible and modular framework, we envision this model as a resource for futher development by the scientific community to include additional complexity and problem-specific components.
+With an accessible and modular framework, we envision this model as a resource for further development by the scientific community to include additional complexity and problem-specific components.
 We chose the R language, which is freely available and in common use by biologists and ecologists, to reach the widest possible audience with this work.
@@
 1, and each model flux and parameter is defined in Tables 1 and 2, respectively.
-We use C-moles as the unit of biomass for consistency with the rigorous mass balance of DEB theory: 1 C-mole is equivalent to the amount of biomass containing 1 mole of carbon atoms.
-Host biomass ($H$), symbiont biomass ($S$), and prey biomass ($X$) have fixed, but different, molar N:C ratios (Table 2).
-Biomass is produced from carbon and nitrogen by synthesizing units (SU), which are mathematical specifications of the formation of a product from two substrates; we use the parallel complementary formulation of @Kooijman:2010vd to specify these fluxes.
-The two state variables of this system are symbiont biomass and coral biomass; because resources are acquired proportionally to surface area, and surface area is proportional to volume (i.e., corals are "V1-morphs" in DEB terminology [@Kooijman:2010vd]), biomass increases exponentially during growth (indeed, corals grow exponentially [@Bak:1976bv]).
+We use C-moles (C-mol) as the unit of biomass for consistency with the rigorous mass balance of DEB theory: 1 C-mol is equivalent to the amount of biomass containing 1 mole of carbon atoms.
+Host biomass (expressed as C-mol H), symbiont biomass (expressed as C-mol S), and prey biomass (expressed as C-mol X) have fixed, but different, molar N:C ratios (Table 2).
+Biomass is produced from carbon and nitrogen by synthesizing units (SU), which are mathematical specifications of the formation of a product from two substrates; we use the "parallel complementary" formulation of @Kooijman:2010vd (page 105, Fig.
+3.7) to specify these fluxes.
+The two state variables of this system are symbiont biomass and coral biomass; because resources are acquired proportionally to surface area, and surface area is proportional to volume, biomass increases exponentially during growth (indeed, corals grow exponentially [@Bak:1976bv]).
 
-Environmental stress is implemented in the form of photooxidative stress, which is thought to be the primary trigger of coral bleaching [@Lesser:1997p3587; @Weis:2008p944; @Wooldridge:2009p7807].
-To simulate these events, we model the absorption and quenching of light energy by photochemistry and non-photochemical quenching, and the phenomenological responses that occur when these capacities are overwhelmed (i.e., photoinhibition, photodamage, and symbiont loss).
-While bleaching in response to high light alone has been observed experimentally [@Schutter:2011ie; @Downs:2013kc], mass coral bleaching events occur in response to high temperature [@HoeghGuldberg:1999p4222]; thus, it is important to justify our consideration of light as the primary stressor.
+Environmental stress is implemented in the form of photooxidative stress, which is thought to be a primary trigger of coral bleaching [@Lesser:1997p3587; @Weis:2008p944; @Wooldridge:2009p7807].
+To simulate bleaching, we model the absorption and quenching of light energy by photochemistry and non-photochemical quenching, and the responses that occur (i.e., photoinhibition, photodamage, and symbiont loss) when these quenching capacities are overwhelmed.
+While bleaching in response to high light alone has been observed experimentally [@Schutter:2011ie; @Downs:2013kc], mass coral bleaching events occur concurrently with high temperature [@HoeghGuldberg:1999p4222]; thus, it is important to justify our consideration of light as the primary stressor.
 In reality, light and temperature interact synergistically [@Coles:1978p1124; @Jones:1998p3572], and in fact, any stressor that disrupts the quenching of light energy may lead to bleaching [@Wooldridge:2010p7809; @Baker:2015kp].
@@
 
-Table 1. Model fluxes (mass-specific)
+Table 1. Model fluxes (mass-specific). Units are explained in the text.
 
@@
 | ------- | -------------------------------------- | ------------------- | ------- |
-| $j_X$     | Prey uptake rate                       | molX CmolH^-1^ d^-1^ | 3 |
-| $j_N$     | Nitrogen uptake rate                   | molN CmolH^-1^ d^-1^ | 4 |
-| $j_{HG}$  | Host biomass formation rate            | CmolH CmolH^-1^ d^-1^| 5 |
-| $r_{NH}$  | Recycled nitrogen from host turnover   | molN CmolH^-1^ d^-1^ | 6 |
-| $\rho_N$  | Nitrogen shared with the symbiont      | molN CmolH^-1^ d^-1^ | 7 |
-| $j_{eC}$  | Excess carbon used to activate host CCMs| molC CmolH^-1^ d^-1^ | 8 |
-| $j_{CO_2}$| CO~2~ input to photosynthesis          | molCO~2~ CmolH^-1^ d^-1^| 9 |
-| $j_L$     | Light absorption rate                  | molph CmolS^-1^ d^-1^| 11 |
-| $r_{CH}$  | Recycled CO~2~ from host turnover      | molCO~2~ CmolH^-1^ d^-1^| 12 |
-| $r_{CS}$  | Recycled CO~2~ from symbiont turnover  | molCO~2~ CmolS^-1^ d^-1^| 13 |
-| $j_{CP}$  | Photosynthesis rate                    | molC CmolS^-1^ d^-1^ | 14 |
-| $j_{eL}$  | Light energy in excess of photochemistry| molph CmolS^-1^ d^-1^| 15 |
-| $j_{NPQ}$ | Total capacity of NPQ                  | molph CmolS^-1^ d^-1^ | 16 |
+| $j_X$     | Prey uptake rate                       | mol X C-mol H^-1^ d^-1^ | 3 |
+| $j_N$     | Nitrogen uptake rate                   | mol N Cmol H^-1^ d^-1^ | 4 |
+| $j_{HG}$  | Host biomass formation rate            | C-mol H C-mol H^-1^ d^-1^| 5 |
+| $j_{HT}$  | Host biomass turnover rate             | C-mol H C-mol H^-1^ d^-1^| 21 |
+| $r_{NH}$  | Recycled nitrogen from host turnover   | mol N C-mol H^-1^ d^-1^ | 6 |
+| $\rho_N$  | Nitrogen shared with the symbiont      | mol N C-mol H^-1^ d^-1^ | 7 |
+| $j_{eC}$  | Excess carbon used to activate host CCMs| mol C C-mol H^-1^ d^-1^ | 8 |
+| $j_{CO_2}$| CO~2~ input to photosynthesis          | mol CO~2~ C-mol H^-1^ d^-1^| 9 |
+| $j_L$     | Light absorption rate                  | mol q C-mol S^-1^ d^-1^| 11 |
+| $r_{CH}$  | Recycled CO~2~ from host turnover      | mol CO~2~ C-mol H^-1^ d^-1^| 12 |
+| $r_{CS}$  | Recycled CO~2~ from symbiont turnover  | mol CO~2~ C-mol S^-1^ d^-1^| 13 |
+| $j_{CP}$  | Photosynthesis rate                    | mol C C-mol S^-1^ d^-1^ | 14 |
+| $j_{eL}$  | Light energy in excess of photochemistry| mol q C-mol S^-1^ d^-1^| 15 |
+| $j_{NPQ}$ | Total capacity of NPQ                  | mol q C-mol S^-1^ d^-1^ | 16 |
 | $c_{ROS}$ | ROS production proportional to baseline | -- | 17 |
-| $r_{NS}$  | Recycled nitrogen from symbiont turnover| molN CmolS^-1^ d^-1^| 18 |
-| $j_{SG}$  | Symbiont biomass formation rate        | CmolS CmolS^-1^ d^-1^| 19 |
-| $\rho_C$  | Fixed carbon shared with host          | molC CmolS^-1^ d^-1^ | 20 |
-| $j_{ST}$  | Symbiont turnover rate                 | CmolS CmolS^-1^ d^-1^| 21 |
+| $r_{NS}$  | Recycled nitrogen from symbiont turnover| mol N C-mol S^-1^ d^-1^| 18 |
+| $j_{SG}$  | Symbiont biomass formation rate        | C-mol S C-mol S^-1^ d^-1^| 19 |
+| $\rho_C$  | Fixed carbon shared with host          | mol C C-mol S^-1^ d^-1^ | 20 |
+| $j_{ST}$  | Symbiont biomass turnover rate         | C-mol S C-mol S^-1^ d^-1^| 21 |
 
-Table 2. Model parameters. Derivations of each parameter value along with supporting references are provided in the Supplementary Information.
+Table 2. Model parameters (units explained in the text). Derivations of each parameter value along with supporting references are provided in the Supplementary Information.
 
@@
 | $n_{NX}$ | N:C molar ratio in prey biomass | -- | $`r defpars$nNX`$ |
-| $j_{HT}^0$ | Specific turnover rate of host biomass | CmolH CmolH^-1^ d^-1^ | $`r defpars$jHT0`$ |
-| $j_{ST}^0$ | Specific turnover rate of symbiont biomass | CmolS CmolS^-1^ d^-1^ | $`r defpars$jST0`$ |
+| $j_{HT}^0$ | Specific maintenance rate of host biomass | C-mol H C-mol H^-1^ d^-1^ | $`r defpars$jHT0`$ |
+| $j_{ST}^0$ | Specific maintenance rate of symbiont biomass | C-mol S C-mol S^-1^ d^-1^ | $`r defpars$jST0`$ |
 | $\sigma_{NH}$ | Proportion host N-turnover recycled to host | -- | $`r defpars$sigmaNH`$ |
@@
 | $\sigma_{CS}$ | Proportion symbiont C-turnover recycled to photosynthesis | -- | $`r defpars$sigmaCS`$ |
-| $j_{Xm}$ | Maximum specific feeding rate of host | molX CmolH^-1^ d^-1^ | $`r defpars$jXm`$ |
+| $j_{Xm}$ | Maximum specific feeding rate of host | molX C-mol H^-1^ d^-1^ | $`r defpars$jXm`$ |
 | $K_X$ | Half-saturation constant for prey uptake by host | molX L^-1^ | $`r defpars$KX`$ |
-| $j_{Nm}$ | Maximum specific DIN uptake rate by host | molN CmolH^-1^ d^-1^ | $`r defpars$jNm`$ |
-| $K_N$ | Half-saturation constant for DIN uptake by host | molN L^-1^ | $`r defpars$KN`$ |
-| $k_{CO_2}$ | Efficacy of CO~2~ delivery by host CCMs | molCO~2~ molC^-1^ | $`r defpars$kCO2`$ |
-| $j_{HGm}$ | Maximum specific growth rate of host | CmolH CmolH^-1^ d^-1^ | $`r defpars$jHGm`$ |
-| $y_{CL}$ | Quantum yield of photosynthesis | molC mol ph^-1^ | $`r defpars$yCL`$ |
-| $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | m^2^ CmolS^-1^ | $`r defpars$astar`$ |
-| $k_{NPQ}$ | NPQ capacity of symbiont | mol ph CmolS^-1^ d^-1^ | $`r defpars$kNPQ`$ |
-| $k_{ROS}$ | Excess photon energy that doubles ROS production, relative to baseline levels | mol ph CmolS^-1^ d^-1^ | $`r defpars$kROS`$ |
-| $j_{CPm}$ | Maximum specific photosynthesis rate of symbiont | molC CmolS^-1^ d^-1^ | $`r defpars$jCPm`$ |
-| $j_{SGm}$ | Maximum specific growth rate of symbiont | CmolS CmolS^-1^ d^-1^ | $`r defpars$jSGm`$ |
+| $j_{Nm}$ | Maximum specific DIN uptake rate by host | mol N C-mol H^-1^ d^-1^ | $`r defpars$jNm`$ |
+| $K_N$ | Half-saturation constant for DIN uptake by host | mol N L^-1^ | $`r defpars$KN`$ |
+| $k_{CO_2}$ | Efficacy of CO~2~ delivery by host CCMs | mol CO~2~ mol C^-1^ | $`r defpars$kCO2`$ |
+| $j_{HGm}$ | Maximum specific growth rate of host | C-mol H C-mol H^-1^ d^-1^ | $`r defpars$jHGm`$ |
+| $y_{CL}$ | Quantum yield of photosynthesis | mol C mol q^-1^ | $`r defpars$yCL`$ |
+| $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | m^2^ C-mol S^-1^ | $`r defpars$astar`$ |
+| $k_{NPQ}$ | NPQ capacity of symbiont | mol ph C-mol S^-1^ d^-1^ | $`r defpars$kNPQ`$ |
+| $k_{ROS}$ | Excess photon energy that doubles ROS production, relative to baseline levels | mol ph C-mol S^-1^ d^-1^ | $`r defpars$kROS`$ |
+| $j_{CPm}$ | Maximum specific photosynthesis rate of symbiont | mol C C-mol S^-1^ d^-1^ | $`r defpars$jCPm`$ |
+| $j_{SGm}$ | Maximum specific growth rate of symbiont | C-mol S C-mol S^-1^ d^-1^ | $`r defpars$jSGm`$ |
 | $b$ | Scaling parameter for strength of bleaching response | -- | $`r defpars$b`$ |
@@
 ## State equations
-The balance equations for symbiont ($S$) and host ($H$) biomass are expressed as:
+The balance equations for symbiont ($S$) and host ($H$) biomass are expressed as "specific" rates, i.e. rates per unit of symbiont and host biomass, respectively:
 
@@
 
-\begin{equation} {dH \over Hdt} = j_{HG} - j_{HT}^0 \end{equation}
+\begin{equation} {dH \over Hdt} = j_{HG} - j_{HT} \end{equation}
 
-These are expressed per unit of symbiont and host biomass, respectively.
 The specific biomass growth and turnover rates that define these balance equations are produced by combinations of the individual model fluxes (see Table 1 for definitions and units), which are each expressed as mass-specific rates (e.g., per C-mole of symbiont or host biomass per day).
@@
 Additionally, the coral animal acquires nitrogen dissolved in the surrounding seawater, which is assumed to represent ammonium, the primary form utilized by corals [@Yellowlees:2008p331].
-This gives the host (rather than the symbiont) priority in nitrogen utilization; this capacity is supported by experimental evidence [@Wang:1998p128] and is consistent with the physical arrangment of the partners, where the host is in direct contact with the external environment.
+This gives the host (rather than the symbiont) priority in nitrogen utilization; this capacity is supported by experimental evidence [@Wang:1998p128] and is consistent with the spatial arrangment of the partners, where the host is in direct contact with the external environment.
 The uptake of nitrogen from the environment is thus specified by Michaelis-Menten kinetics using a maximum uptake rate $j_{Nm}$ and half-saturation constant $K_N$:
@@
 
-Coral biomass formation is then specified by a parallel complementary SU that combines carbon and nitrogen, according to:
+Coral biomass formation is then specified by a parallel complementary SU (formula in Fig. 3.7 of @Kooijman:2010vd]) that combines carbon and nitrogen, according to:
 
@@
 
-where $\rho_C$ is fixed carbon shared by the symbiont (see Eq. 20), and $r_{NH}$ is recycled nitrogen liberated by host biomass turnover:
+where $\rho_C$ is fixed carbon shared by the symbiont (see Eq.
+20), and $r_{NH}$ is recycled nitrogen liberated by host biomass turnover (see Eq.
+XX).
+Host biomass turnover is equal to the specific maintenance rate of host biomass, which is set as a parameter $j_{HT}^0$,
 
-\begin{equation} r_{NH}=\sigma_{NH}n_{NH}j_{HT}^0 \end{equation}
+\begin{equation} j_{HT}=j_{HT}^0 \end{equation}
+
+and the specific flux of nitrogen that is recycled to the host biomass SU is calcluated as:
+
+\begin{equation} r_{NH}=\sigma_{NH}n_{NH}j_{HT} \end{equation}
 
@@
 Since these active CCMs require energetic input by the host, we define $j_{CO_2}$ as proportional to $j_{eC}$, assuming that some of this carbon is respired to energize the CCMs.
-This formulation means that the symbiont indirectly ensures its own CO~2~ supply by providing fixed carbon (=energy) to the host [@Wooldridge:2013tj], which establishes a positive feedback in the model that leads to the existence of multiple stable states, which we discuss below in the context of coral bleaching.
+This formulation means that the symbiont indirectly ensures its own CO~2~ supply by providing fixed carbon (=energy) to the host [@Wooldridge:2013tj].
 The parameter $k_{CO_2}$ scales the efficacy of host CCMs, which enables the comparison of different rates of CO~2~ delivery that may characterize different coral species [@Wooldridge:2014di].
@@
 
-This amplification factor is then multiplied by the external downwelling irradiance $L$ and the effective light-absorbing surface area of symbiont biomass $\bar{a}^*$ to specify the total light absorption:
+This amplification factor is then multiplied by the external downwelling irradiance $L$ and a parameter representing the effective light-absorbing surface area of symbiont biomass $\bar{a}^*$ to specify the total light absorption:
 
@@
 CO~2~ arrives at the photosynthesis SU from multiple sources: in addition to the CO~2~ actively supplied by the host through its CCMs ($j_{CO_2}$; Eq.
-9), a fixed proportion of metabolic CO~2~ generated by host biomass turnover is passively available to the photosynthesis SU, according to:
+9), we assume a fixed proportion $\sigma_{CH}$ of metabolic CO~2~ generated by host biomass turnover is passively available to the photosynthesis SU, according to:
 
-\begin{equation} r_{CH}=\sigma_{CH}j_{HT}^0 \end{equation}
+\begin{equation} r_{CH}=\sigma_{CH}j_{HT} \end{equation}
 
-along with a fixed proportion of CO~2~ generated by symbiont biomass turnover:
+along with a fixed proportion of CO~2~ generated by symbiont biomass turnover[^2]:
 
@@
 where $k_{ROS}$ is a parameter of the symbiont that determines the rate of ROS production (specifically, the amount of excess excitation energy that doubles ROS production relative to baseline levels).
-Importantly, $c_{ROS}$ is specified here not as a function of absolute external light, but rather the amount of excess light after accounting for quenching by carbon fixation and NPQ.
+Importantly, $c_{ROS}$ is specified here not as a function of absolute external light, but rather the amount of excess light energy after accounting for quenching by carbon fixation and NPQ.
 A direct consequence of this formulation is that CO~2~-limitation of photosynthesis can lead to ROS production, an important mechanism [@Wooldridge:2009p7807] that was not captured by previous representations of photooxidative stress [@Eynaud:2011tv].
+With this single SU, both the light and dark reactions of photosynthesis are represented, allowing for sink-limitation (i.e., CO~2~-limitation) to cause overreduction of the electron transport chain and ROS production.
 
-Carbon fixed by photosynthesis ($j_{CP}$; Eq. 14) is then combined with nitrogen shared by the host ($\rho_N$; Eq. 7) and nitrogen recycled from symbiont biomass turnover
+Carbon fixed by photosynthesis ($j_{CP}$; Eq. 14) is then combined with nitrogen shared by the host ($\rho_N$; Eq. 7) and nitrogen recycled from symbiont biomass turnover[^2]
 
@@
 
-The rejection flux of nitrogen from the symbiont biomass SU is lost to the environment.
+Nitrogen rejected by the symbiont biomass SU, which has already been rejected by the host biomass SU, cannot be used by either partner and is thus lost to the environment.
 
@@
 The parameter $b$ is included to scale biomass loss due to bleaching in response to ROS.
-(Note that recycling of symbiont biomass turnover ($rNS$ and $rCS$) only occurs based on the maintenance component of turnover (i.e., $j_{ST}^0$), and not the photodamage/bleaching component, as this loss represents biomass that is damaged or expelled from the host.)
+
+[^2]: Note that recycling of symbiont biomass turnover ($rNS$ and $rCS$) only occurs based on the maintenance component of turnover (i.e., $j_{ST}^0$), and not the photodamage/bleaching component, as this loss represents biomass that is expelled from the host.
 
@@
 In a constant environment, the system ultimately reaches a steady state of exponential growth or decline.
-However, under some conditions, either of these outcomes may occur depending on initial values of symbiont and host biomass, indicating the presence of alternate stable states, corresponding to either positive or negative growth (Fig.
+However, under some conditions, either of these outcomes may occur depending on initial values of symbiont and host biomass, indicating the presence of alternate stable states (Fig.
 2).
-The mechanism that produces these alternate stable states is the positive feedback between carbon-limitation of the host and CO~2~-limitation of photosynthesis: if symbiont biomass is initially very low (i.e., a "bleached" coral), the system cannot escape this positive feedback and cannot grow.
+The mechanism that produces these alternate stable states is the positive feedback between carbon-limitation of the host and CO~2~-limitation of photosynthesis: if symbiont biomass is initially very low (i.e., a "bleached" coral), the system cannot escape this positive feedback and cannot grow (unless feeding is sufficiently high).
 However, if symbiont biomass is initially high (i.e., a "healthy" coral), then the system remains in a nitrogen-limited state with positive growth.
@@
 
-![Steady state values of (A) specific growth (Cmol Cmol^-1^ d^-1^) and (B) the symbiont to host biomass ratio (CmolS CmolH^-1^) across gradients of external irradiance and dissolved inorganic nitrogen.
+![Steady state values of (A) specific growth (Cmol Cmol^-1^ d^-1^) and (B) the symbiont to host biomass ratio (C-mol S C-mol H^-1^) across gradients of external irradiance and dissolved inorganic nitrogen.
 Note that typical conditions for reefs are ~1e-7 M DIN and 10-20 mol photons m^-2^ d^-1^.
@@
 However, DIN elevation beyond a certain point (e.g., ~3-4 µM in these simulations) has little effect on growth as carbon becomes limiting.
-Very high nutrient levels may even reduce growth [@Shantz:2015dz], although these impacts are not likely to occur within the range of concentrations considered here (<4 µM) [@FerrierPages:2000p1950].
+Although very high nutrient levels may reduce growth in nature [@Shantz:2015dz], these impacts are not likely to occur within the range of concentrations considered here (<4 µM) [@FerrierPages:2000p1950].
 In addition to increasing growth, DIN also increases the symbiont to host biomass ratio (Fig.
@@
 The maximum predicted growth rates of ~0.1 d^-1^, occurring between ~10-25 mol photons m^-2^ s^-1^ light and ~4 µM DIN (Fig.
-3A), are comparable to the rate of `r Tanaka_gr` d^-1^ measured by @Tanaka:2007uv under similar N-enriched conditions.
-Predicted growth rates under conditions more typical of reef environments (<0.5 µM DIN) are ~0.01-0.03 d^-1^.
-Observed specific growth rates, which are typically based on skeletal mass, fall near or below 0.01 d^-1^ [@Osinga:2011kq; @Osinga:2012is], though values as high as 0.025 d^-1^ have been reported [@Schutter:2010p7758].
-It is not surprising that predicted growth rates are slightly higher than most observations, as the model does not account for external ecological factors that may reduce growth, such as competition, predation, and bioerosion.
-Furthermore, the model does not explicitly account for skeletal formation, which may not always correlate with biomass growth, potentially due to differential substrate-limitation [@Anthony:2002tc].
-Nevertheless, when growth of coral tissue biomass has been directly measured over short periods, values are consistent with model predictions (e.g., `r Tanaka_gr` d^-1^ in @Tanaka:2007uv), and similar rates (`r Armoza_gr` d^-1^) have been measured in *Aiptasia diaphana*, a non-calcifying symbiotic anemone [@ArmozaZvuloni:2014ju].
+3A), are comparable to the rate of `r Tanaka_gr` d^-1^ measured by @Tanaka:2007uv in *Acropora pulchra* under similar N-enriched conditions.
+Under conditions more typical of reef environments (<0.5 µM DIN), predicted growth rates are ~0.01-0.03 d^-1^.
+Observed specific growth rates in several coral species fall near or below the lower end of this range (~0.01 d^-1^) [@Osinga:2011kq; @Osinga:2012is], though values as high as 0.025 d^-1^ have been reported in *Galaxea fascicularis* [@Schutter:2010p7758], and `r Armoza_gr` d^-1^ in *Aiptasia diaphana*, a non-calcifying symbiotic anemone [@ArmozaZvuloni:2014ju].
+It is not surprising that the model predicts slightly higher growth than is typically observed, since it does not account for ecological factors that may limit growth (e.g., competition, predation, bioerosion).
+Furthermore, while most measurements are made on skeletal growth, the model predicts biomass growth, which may not always be strongly correlated [@Anthony:2002tc].
 
@@
 This gradual decline is consistent with experimental results showing that high light levels decrease growth [@Schutter:2011ie], and field studies documenting optimum growth rates at intermediate depths [@Baker:1975ip; @Huston:1985ec].
-By incorporating these impacts of light stress, the model predicts greater, and more realistic, variation in state variables across light gradients than was predicted by the model of @Muller:2009io, which did not include photoinhibition and photodamage.
-It is important to recognize that the boundary set by photooxidative stress on a stable symbiosis under steady state conditions (Fig.
+By incorporating these impacts of light stress, the model predicts greater, and more realistic, variation in state variables across light gradients than was predicted by the models of @Muller:2009io, which did not include photoinhibition or photodamage, or @Eynaud:2011tv, which included representations of photoinhibition or photodamage separately.
+It is important to recognize that the upper light limit set by photooxidative stress on a stable symbiosis under steady state conditions (Fig.
 3) may be temporarily crossed by a dynamic system, which may experience a period of symbiont loss (bleaching) and reduced growth, after which a return to benign conditions may restore symbiont biomass and positive growth.
@@
 Here we evaluate the sensitivity of the model to changes in these parameter values, which also serves to demonstrate the behavior of the dynamical system.
-We measured fractional change in steady state values in response to fractional changes in parameter values, relative to their default values, under environmental conditions typical of coral reefs.
+We calculated fractional change in steady state values in response to fractional changes in parameter values, relative to their default values, under environmental conditions typical of coral reefs.
 
@@
 
-Overall, relative changes in the steady state of the system are less than the equivalent relative change in parameter value, indicating that the system maintains internal stability.
+Overall, relative changes in the steady state of the system are less than the equivalent relative change in parameter value.
 However, changes in certain parameter values have more significant impacts: increasing $j_{Nm}$ or decreasing $K_N$ both dramatically increase host growth (Fig.
@@
 Using the levels of light and nutrients from this study as inputs, the model reproduces this observed interaction among environmental factors (Fig.
-5), and also provides the mechanism: under high light in summer, photooxidative stress decreases symbiont biomass and increases carbon-limitation of host growth regardless of nutrient status; when light is reduced in winter, growth and symbiont biomass become constrained instead by nitrogen, and therefore increase more when DIN is enriched (Fig.
+5), and also provides the mechanism: as increasing light in summer decreases carbon fixation due to photooxidative stress, symbionts grow slower than the host, leading to decreasing S:H ratios.
+Under nutrient enrichment, this effect is more pronounced, as CO~2~-limitation (due to higher symbiont standing stocks) causes mild bleaching that results in a similar summertime minimum S:H as the ambient nutrient case ('physiological bleaching' *sensu* @Fitt:2001p832).
+Decreasing light into winter then alleviates the photooxidative stress constraints on carbon fixation, such that nitrogen constrains the S:H ratio, explaining why S:H increases more when DIN is enriched (Fig.
+S8).
+The model of @Muller would also predict higher S:H ratios in winter if light became limiting to photosynthesis, but this is distinct from the mechanism here, in which stress in the summer drives a reduction in S:H.
+These can be distinguished as follows:
+
+winter increase in S:H as light limits carbon fixation - decreases growth in winter
+summer decrease in S:H as photo-oxidative stress limits carbon fixation - decreases growth in summer
+
+under high light in summer, photooxidative stress decreases symbiont biomass and increases carbon-limitation of host growth regardless of nutrient status; when light is reduced in winter, growth and symbiont biomass become constrained instead by nitrogen, and therefore increase more when DIN is enriched (Fig.
 S8).
@@
 Light input (upper panel) was designed as a sinusoidal curve with a period of one year, with maximum and minimum values of 44 and 22 mol photons m^-2^ d^-1^, based on @Stimson:1997p3837.
-The dynamic behavior of symbiont to host biomass ratio (middle panel) and the specific growth rate of host biomass (lower panel) show seasonal oscillations that are greater in magnitude under higher nutrients (1e-5 molN L^-1^; dashed lines) relative to lower nutrients (1e-7 molN L^-1^; solid lines), consistent with the findings of @Stimson:1997p3837.
+The dynamic behavior of symbiont to host biomass ratio (middle panel) and the specific growth rate of host biomass (lower panel) show seasonal oscillations that are greater in magnitude under higher nutrients (1e-5 mol N L^-1^; dashed lines) relative to lower nutrients (1e-7 mol N L^-1^; solid lines), consistent with the findings of @Stimson:1997p3837.
 Prey density was set at 1e-5 CmolX L^-1^.](../img/Fig5.png)
 
-The prediction of higher growth when light is reduced indicates that growth is not limited by low light in winter, but is actually reduced by excess light in summer, consistent with the experimental findings of @Schutter:2011ie.
+The prediction of higher growth when light is reduced indicates that growth is not limited by low light in winter, but is actually reduced by excess light in summer[^2], consistent with the experimental findings of @Schutter:2011ie.
 Seasonal summertime reductions in tissue biomass have also been well-documented in the field [@Fitt:2000p3792], along with reductions in net photosynthetic capacity [@MullerParker:1987cr].
-Importantly, while light alone (at these levels) may drive seasonal dynamics in the ways discussed, temperature fluctuations may attenuate or even reverse the effect of light as cooler winters depress metabolism; thus, the relative magnitude of fluctuation in temperature, light, and other factors may produce wide variability in the direction and magnitude of seasonal changes in growth and symbiont abundance, depending on location and microhabitat.
+Importantly, while light alone may drive seasonal dynamics in the ways discussed[^2], temperature fluctuations may attenuate or even reverse the effect of light as cooler winters depress metabolism; thus, the relative magnitude of fluctuation in temperature, light, and other factors may produce wide variability in the direction and magnitude of seasonal changes in growth and symbiont abundance, depending on location and microhabitat.
 Nevertheless, the seasonal variability predicted here (Fig.
 5) is consistent with experimental and field observations for corals, and demonstrates the power of this model to predict dynamic behavior that mechanistically integrates multiple environmental drivers.
+
+[^2]: at the light levels indicated in @Stimson:1997p3837.
+Note that if light levels were reduced throughout the year (e.g.
+for a coral at greater depth) such that light did not cause photo-oxidative stress in summer but became limiting to photosynthesis in winter, the S:H ratio would still increase in winter, but growth would decrease; the latter scenario is predicted both by the present model as well as that of @Muller:2009io, since photo-oxidative stress does not become relevant.
 
@@
 6B) and photoinhibition (Fig.
-6C) such that the positive feedbacks between host carbon-limitation (Fig.
+6C) that the positive feedbacks between host carbon-limitation (Fig.
 6D) and CO~2~-limitation of photosynthesis (Fig.
@@
 22.
-All parameters were set at default values with external DIN=1e-7 molN/L and prey density=0).](../img/Fig6.png)
+All parameters were set at default values with external DIN=1e-7 mol N/L and prey density=0).](../img/Fig6.png)
 
@@
 ![Bleaching with interactive factors.
-Simulations of high light stress (sinusoid with maximum=48 mol ph m^-2^ d^-1^) under default environmental conditions (solid line; DIN=1e-7 molN L^-1^; prey=3e-6 CmolX L^-1^), or with elevated feeding (dashed line; prey=6e-6 CmolX L^-1^), or elevated nutrients (dotted line; DIN=1e-6 molN L^-1^).
+Simulations of high light stress (sinusoid with maximum=48 mol ph m^-2^ d^-1^) under default environmental conditions (solid line; DIN=1e-7 mol N L^-1^; prey=3e-6 CmolX L^-1^), or with elevated feeding (dashed line; prey=6e-6 CmolX L^-1^), or elevated nutrients (dotted line; DIN=1e-6 mol N L^-1^).
 Initial symbiont biomass was set to the steady state for each set of starting conditions, respectively.
@@
 
-Whether this 'overshoot' occurs or not represents a bifurcation point in the dynamical system that is defined by the relative availability of carbon and nitrogen to the host -- any factor that enhances carbon-limitation of host growth (e.g.
+Whether this 'overshoot' occurs or not is determined by the relative availability of carbon and nitrogen to the host -- any factor that enhances carbon-limitation of host growth (e.g.
 high DIN and/or low feeding) therefore magnifies the overshoot and prolongs the dysfunctional, carbon-limited state of the symbiosis (Fig.
@@
 22.
-All simulations were run with default parameters (except varying $n_{NX}$), with L=10 mol photons m^2^ d^-1^, DIN=1e-7 molN L^-1^, prey=1e-6 CmolX L^-1^, and initial S:H biomass=0.001.](../img/Fig8.png)
+All simulations were run with default parameters (except varying $n_{NX}$), with L=10 mol photons m^2^ d^-1^, DIN=1e-7 mol N L^-1^, prey=1e-6 CmolX L^-1^, and initial S:H biomass=0.001.](../img/Fig8.png)
 
 These dynamics reveal that the re-establishment of nitrogen-limitation is the most important diagnostic of recovery to a 'healthy' state, as this is when positive growth rates are resumed.
-A high S:H biomass ratio alone does not necessarily indicate that a 'healthy' state has been reached, since the C-limited state may still persist (e.g., Fig.
+A high S:H biomass ratio alone does not necessarily indicate that a 'healthy' state has been reached, since the C-limited state may still persist (e.g.
+Fig.
 8B, 8C).
-This could explain why corals that have recovered their symbiont populations after bleaching may still exhibit energetic deficits and physiological impacts for much longer periods of time [@Levitan:2014et, @Hughes:2013hd].
+This could explain why corals that have recovered their symbiont populations after bleaching may still exhibit energetic deficits and physiological impacts, possibly for months to years [@Levitan:2014et, @Hughes:2013hd]).
 Therefore, host acquisition of carbon from a source other than the symbiont may be extremely important for the system to recover from bleaching.
@@
 # Conclusions
-This simplified dynamic bioenergetic model of coral-*Symbiodinium* symbioses mechanistically reproduces patterns in steady-state coral growth and symbiont abundance commonly observed in corals, including higher symbiont abundance with higher nutrients and feeding, lower symbiont abundance with increasing light, and optimal growth at intermediate light levels.
+This dynamic bioenergetic model of coral-*Symbiodinium* symbioses mechanistically reproduces patterns in steady-state coral growth and symbiont abundance commonly observed in corals, including higher symbiont abundance with higher nutrients and feeding, lower symbiont abundance with increasing light, and optimal growth at intermediate light levels.
 Moreover, the model reproduces complex dynamic behaviors including seasonal changes in symbiont density at different nutrient levels, rapid bleaching above a threshold of high light, mitigation of bleaching by heterotrophic feeding, exacerbation of bleaching by elevated nutrients, and an overshoot of symbiont density during recovery from bleaching.

### 2017-02-07 JTB_supp/JTB_supp.Rmd
address EM + RN comments

@@
 ---
-title: A simplified dynamic bioenergetic model for coral-*Symbiodinium* symbioses and coral bleaching as an alternate stable state
+title: "Supplementary Information for:"
+subtitle: A simplified dynamic bioenergetic model for coral-*Symbiodinium* symbioses and coral bleaching as an alternate stable state
 author: "Ross Cunning, Erik B. Muller, Ruth D. Gates, Roger M. Nisbet"
-output: html_document
+output:
+  html_document:
+    toc: true
+    toc_float: true
+    toc_depth: 1
 bibliography: library.bib
@@
 
-# Supplementary Information
+-----
+
 This supplement provides justification and derivations from the literature for all parameter values used in the model, as well as additional figures that demonstrate the structure and behavior of the model and its specific components.
@@
 ```{r setup, include=FALSE}
+require(pander)
+require(stringr)
 knitr::opts_knit$set(root.dir = normalizePath(".."))
-knitr::opts_chunk$set(message = F, warning = F, echo=F)
+knitr::opts_chunk$set(message = F, warning = F, echo=T, results='asis')
+knitr::opts_chunk$set(results = "asis", render=pander)
 ```
@@
 
+-----
+
+# Model parameters
+All parameters in the model are listed here along with the relevant literature that provides a basis for the value used.
+In most cases, the units of the values reported in the literature were converted to units appropriate for the model, which was accomplished using a consistent set of stoichiometric conversion factors also based on reported literature (see [Conversion factors used in parameter derivation]).
+Calculations executed in R are shown for each parameter (when relevant) as well as the final value used.
+
 ## Parameter value derivations
-All parameters in the model are listed here along with the relevant literature that provides a basis for the value used.
-In most cases, the units of the values reported in the literature needed to be converted to units appropriate for the model, which was accomplished using a consistent set of stoichiometric conversion factors also based on reported literature (see "R/conversions.R" in this paper's data repository).
-The specific calculations used to convert values from the units reported in the literature to the relevant units for the model can be found in the RMarkdown code used to produce this document.
 
-* $n_{NH}$: N:C molar ratio in host biomass
-    + @Horwitz:2015cf measured a C:N ratio of 5.53 in anemones under control conditions, which converts to:
-    + Value used: `r defpars$nNH` (N:C ratio, dimensionless)
-* $n_{NS}$: N:C molar ratio in symbiont biomass
-    + This value should approximate the Redfield ratio (following @Muller:2009io).
-    + Value used: `r defpars$nNS` (dimensionless)
-* $n_{NX}$: N:C molar ratio in prey biomass
-    + Assuming the coral feeds primarily on zooplankton, a general heterotroph has a N:C ratio of 0.2 [@Muller:2009io].
-    + Value used: `r defpars$nNX` (N:C ratio, dimensionless)
-* $j_{HT}^0$: Specific turnover rate of host biomass
-    + @Jacobson:2016bk parameterized a DEB model for coral symbiosis, and reported a specific maintenance rate (analagous to the turnover rate in the current model) of 0.03 CmolH CmolH^-1^ d^-1^.
-    + Value used: `r defpars$jHT0` CmolH CmolH^-1^ d^-1^
-* $j_{ST}^0$: Specific turnover rate of symbiont biomass
-    + Assuming an equivalent maintenance rate as the host:
-    + Value used: `r defpars$jST0` CmolS CmolS^-1^ d^-1^
-* $\sigma_{NH}, \sigma_{NS}$: Proportions of nitrogen recycled from host and symbiont biomass turnover
-    + Assume highly efficient recycling by both symbiont and host to avoid loss to the environment:
-    + Values used: `r with(defpars, c(sigmaNH, sigmaNS))` (dimensionless)
-* $\sigma_{CS}$: Proportion of metabolic CO~2~ produced by symbiont recycled passively to photosynthesis
-    + Assume that CO~2~ produced within symbiont cells is recycled with high efficiency, potentially aided by symbiont carbon concentrating mechanisms
-    + Value used: 0.9 (dimensionless)
-* $\sigma_{CH}$: Proportions of metabolic CO~2~ produced by host recycled passively to photosynthesis 
-    + Assume that most metabolic CO~2~ produced within host cells rapidly converts to HCO~3~^-^ at intracellular pH ~ 7.2 [@Venn:2009hy], and that only ~10% actually remains as CO~2~ [@Zeebe:2001wz] that could diffusively supply photosynthesis.
+#### $n_{NH}$: N:C molar ratio in host biomass {-}
+@Horwitz:2015cf measured a molar C:N ratio of 5.53 in anemones under control conditions, which converts to:
+```{r nNH}
+1 / 5.53
+```
+N:C ratio (dimensionless)
+<br><br>
+
+#### $n_{NS}$: N:C molar ratio in symbiont biomass {-}
+This value should approximate the Redfield ratio (following @Muller:2009io).  
+```{r nNS, echo=F}
+0.13
+```
+N:C ratio (dimensionless)
+<br><br>
+
+#### $n_{NX}$: N:C molar ratio in prey biomass {-}
+Assuming the coral feeds primarily on zooplankton, a general heterotroph has a molar N:C ratio of 0.2 [@Muller:2009io].  
+```{r nNX, echo=F}
+0.20
+```
+N:C ratio (dimensionless)
+<br><br>
+
+#### $j_{HT}^0$: Specific turnover rate of host biomass {-}
+@Jacobson:2016bk parameterized a DEB model for coral symbiosis, and reported a specific maintenance rate of 0.03 CmolH CmolH^-1^ d^-1^.
+Although this value is per unit structural biomass (in the DEB sense) rather than per unit total biomass (as needed here), use of this value generates plausible values of specific growth (see main text).
+```{r jHT, echo=F}
+0.03
+```
+CmolH CmolH^-1^ d^-1^
+<br><br>
+
+#### $j_{ST}^0$: Specific turnover rate of symbiont biomass {-}
+Assuming an equivalent maintenance rate as the host:
+```{r jST, echo=F}
+0.03
+```
+CmolS CmolS^-1^ d^-1^
+<br><br>
+
+#### $\sigma_{NH}, \sigma_{NS}$: Proportions of nitrogen recycled from host and symbiont biomass turnover {-}
+Assume highly efficient recycling by both symbiont and host to avoid loss to the environment:  
+```{r sigmaNHsigmaNS, echo=F}
+c(0.03, 0.03)
+```
+(dimensionless)
+<br><br>
+
+#### $\sigma_{CS}$: Proportion of metabolic CO~2~ produced by symbiont recycled passively to photosynthesis {-}
+Assume that CO~2~ produced within symbiont cells is recycled with high efficiency, potentially aided by symbiont carbon concentrating mechanisms.  
+```{r sigmaCS, echo=F}
+0.9
+```
+(dimensionless)
+<br><br>
+
+#### $\sigma_{CH}$: Proportion of metabolic CO~2~ produced by host recycled passively to photosynthesis {-}
+Assume that most metabolic CO~2~ produced within host cells rapidly converts to HCO~3~^-^ at intracellular pH ~ 7.2 [@Venn:2009hy], and that only ~10% actually remains as CO~2~ [@Zeebe:2001wz] that could diffusively supply photosynthesis.
 This passive supply is differentiated from all active host carbon concentrating mechanisms (CCMs) that facilitate CO~2~ supply to photosynthesis, which are separately treated collectively as the $jCO~2~$ flux.
-    + Values used: 0.1 (dimensionless)
-* $j_{Xm}$: Maximum specific feeding rate of host
-    + @Wijgerde:2011du measured depletion of carbon from captured prey in *Galaxea fascicularis* (including both intracoelenteric and extracoelenteric feeding), reporting a maximum specific feeding rate of 77.6 µgC cm^-2^ 6h^-1^.
-This converts to:
-    + Value used: `r round(77.6 / µg.g / gC.molC * cm2.m2 / CmolR.m2_Gfas * 4, 2)` molX CmolH^-1^ d^-1^
-* $K_X$: Half-saturation constant for prey uptake by host
-    + @Purser:2010hp measured feeding rates of the coral *Lophelia pertusa* at different concentrations of *Artemia* nauplii, reporting a half-saturation for feeding of ~345 *Artemia* L^-1^.
-This converts to:
-    + Value used: $`r round(345 * gC.artemia / gC.molC, 7)`$ CmolX L^-1^
-* $j_{Nm}$: Maximum specific DIN uptake rate by host
-    + @Muscatine:1978fa measured ammonium uptake in 3 coral species and reported an average value of 7.22 µmol NH~4~^+^-N mg chl *a*^-1^ h^-1^ (Table 1, uncorrected), which converts to `r 7.22 * h.d * nmolchlA.cm2 * gchlA.molchlA * mg.g / nmol.mol / CmolR.m2_Spis * cm2.m2 / µmol.mol` mol NH~4~^+^-N  CmolH^-1^ d^-1^.
-    + @HoeghGuldberg:1999p3974 measured ammonium and dissolved free amino acid uptake in *Pocillopora damicornis* and reported a maximum rate of 91.8 nmol N cm^-2^ h^-1^, which converts to `r 91.8 * h.d * cm2.m2 / CmolR.m2_Spis / nmol.mol` molN CmolH^-1^ d^-1^.
-    + @Godinot:2011jg measured nitrate and ammonium uptake rates in *Stylophora pistillata* under high nutrient concentrations and reported a maximum value of 3.5 µmolN mg chl^-1^ d^-1^ [addition of bars in Fig.
-4D,E @ 29°C], which converts to `r 3.5 * h.d * 22 / CmolR.m2_Spis * cm2.m2 / µmol.mol / 1000` molN CmolH^-1^ d^-1^.
-    + @Godinot:2011bm measured ammonium uptake kinetics in *Stylophora pistillata* and reported a maximum rate of 5.42 molN mg chl^-1^ h^-1^, which converts to `r 5.42 * h.d * µgchl.cm2_Spis / CmolR.m2_Spis * cm2.m2 / µmol.mol / 1000` molN CmolH^-1^ d^-1^.
-    + Averaging these reported values, we arrive at:
-    + Value used: `r round(mean(c(7.22 * h.d * nmolchlA.cm2 * gchlA.molchlA * mg.g / nmol.mol / CmolR.m2_Spis * cm2.m2 / µmol.mol, 91.8 * h.d * cm2.m2 / CmolR.m2_Spis / nmol.mol, 3.5 * h.d * 22 / CmolR.m2_Spis * cm2.m2 / µmol.mol / 1000, 5.42 * h.d * µgchl.cm2_Spis / CmolR.m2_Spis * cm2.m2 / µmol.mol / 1000)), 3)` molN CmolH^-1^ d^-1^.
-* $K_N$: Half-saturation constant for DIN uptake by host
-    + @Muscatine:1978fa measured ammonium uptake in 3 coral species and reported an average half-saturation constant of 1.84 µmolN L^-1^ (Table 1, uncorrected).
-    + @Godinot:2011bm measured ammoinum uptake kinetics in *Stylophora pistillata* and reported a half-saturation constant of 0.46 µmol N L^-1^.
-    + Averaging these reported values (weighted by number of species measured), we arrive at:
-    + Value used: `r round((1.84*3 + 0.46*1)/4, 2)` molN L^-1^
-* $k_{CO_2}$: Efficacy of CO~2~ delivery by host CCMs
-    + This parameter represents the relative amount of CO~2~ that can be delivered by the host to photosynthesis given the reducing power generated from 1 mol of fixed carbon.
+```{r sigmaCH, echo=F}
+0.10
+```
+(dimensionless)
+<br><br>
+
+#### $j_{Xm}$: Maximum specific feeding rate of host {-}
+@Wijgerde:2011du measured depletion of carbon from captured prey in *Galaxea fascicularis* (including both intracoelenteric and extracoelenteric feeding), reporting a maximum specific feeding rate of 77.6 µgC cm^-2^ 6h^-1^.
+```{r jXm}
+77.6 / µg.g / gC.molC * cm2.m2 / CmolR.m2_Gfas * 24/6
+```
+molX CmolH^-1^ d^-1^
+<br><br>
+
+#### $K_X$: Half-saturation constant for prey uptake by host {-}
+@Purser:2010hp measured feeding rates of the coral *Lophelia pertusa* at different concentrations of *Artemia* nauplii, reporting a half-saturation for feeding of ~345 *Artemia* L^-1^.
+```{r KX}
+345 * gC.artemia / gC.molC
+```
+CmolX L^-1^
+<br><br>
+
+#### $j_{Nm}$: Maximum specific DIN uptake rate by host {-}
+@Muscatine:1978fa measured ammonium uptake in 3 coral species and reported an average value of 7.22 µmol NH~4~^+^-N mg chl *a*^-1^ h^-1^ (Table 1, uncorrected), which converts to:
+```{r jNm1}
+7.22 * h.d * nmolchlA.cm2 * gchlA.molchlA * mg.g / nmol.mol / CmolR.m2_Spis * cm2.m2 / µmol.mol
+```
+mol NH~4~^+^-N  CmolH^-1^ d^-1^  
+<br>
+@HoeghGuldberg:1999p3974 measured ammonium and dissolved free amino acid uptake in *Pocillopora damicornis* and reported a maximum rate of 91.8 nmol N cm^-2^ h^-1^, which converts to:
+```{r jNm2}
+91.8 * h.d * cm2.m2 / CmolR.m2_Spis / nmol.mol
+``` 
+molN CmolH^-1^ d^-1^  
+<br>
+@Godinot:2011jg measured nitrate and ammonium uptake rates in *Stylophora pistillata* under high nutrient concentrations and reported a maximum value of 3.5 µmolN mg chl^-1^ d^-1^ [addition of bars in Fig.
+4D,E @ 29°C], which converts to:
+```{r jNm3}
+3.5 * h.d * 22 / CmolR.m2_Spis * cm2.m2 / µmol.mol / 1000
+```
+molN CmolH^-1^ d^-1^   
+<br>
+@Godinot:2011bm measured ammonium uptake kinetics in *Stylophora pistillata* and reported a maximum rate of 5.42 molN mg chl^-1^ h^-1^, which converts to:
+```{r jNm4}
+5.42 * h.d * µgchl.cm2_Spis / CmolR.m2_Spis * cm2.m2 / µmol.mol / 1000
+```
+molN CmolH^-1^ d^-1^   
+<br>
+Averaging these reported values, we arrive at:  
+```{r jNm_avg, echo=F}
+round(mean(c(7.22 * h.d * nmolchlA.cm2 * gchlA.molchlA * mg.g / nmol.mol / CmolR.m2_Spis * cm2.m2 / µmol.mol, 91.8 * h.d * cm2.m2 / CmolR.m2_Spis / nmol.mol, 3.5 * h.d * 22 / CmolR.m2_Spis * cm2.m2 / µmol.mol / 1000, 5.42 * h.d * µgchl.cm2_Spis / CmolR.m2_Spis * cm2.m2 / µmol.mol / 1000)), 3)
+```
+molN CmolH^-1^ d^-1^
+<br><br>
+
+#### $K_N$: Half-saturation constant for DIN uptake by host {-}
+@Muscatine:1978fa measured ammonium uptake in 3 coral species and reported an average half-saturation constant of 1.84 µmolN L^-1^ (Table 1, uncorrected).  
+@Godinot:2011bm measured ammoinum uptake kinetics in *Stylophora pistillata* and reported a half-saturation constant of 0.46 µmol N L^-1^.  
+Averaging these reported values (weighted by number of species measured), we arrive at:  
+```{r KN}
+(1.84*3 + 0.46*1)/4
+``` 
+molN L^-1^
+<br><br>
+
+#### $k_{CO_2}$: Efficacy of CO~2~ delivery by host CCMs {-}
+This parameter represents the relative amount of CO~2~ that can be delivered by the host to photosynthesis given the reducing power generated from 1 mol of fixed carbon.
 We assume that the host uses a portion of its fixed carbon to generate ATP which is then used to activate a variety of carbon concentrating mechanisms (see @Wooldridge:2013tj).
 We assume these CCMs are efficient and can deliver an order of magnitude more CO~2~ relative to the fixed carbon available.
-    + Value used: 10 molCO~2~ molC^-1^
-* $j_{HGm}$: Maximum specific growth rate of host 
-    + Value used: 1 CmolH CmolH^-1^ d^-1^
-    + Reference:
-* $y_{CL}$: Quantum yield of photosynthesis
-    + The maximum theoretical quantum yield is 0.125 mol C fixed per mol photons.
+```{r kCO2, echo=F}
+10
+```
+molCO~2~ molC^-1^
+<br><br>
+
+#### $j_{HGm}$: Maximum specific growth rate of host {-}
+We set this to a high value such that constraints on growth are primarily determined by substrate availabilities rather than the maximum value.  
+```{r jHGm, echo=F}
+1
+```
+CmolH CmolH^-1^ d^-1^
+<br><br>
+
+#### $y_{CL}$: Quantum yield of photosynthesis {-}
+The maximum theoretical quantum yield is 0.125 mol C fixed per mol photons.
 @Brodersen:2014gw presented a balanced light budget based on microenvironmental photobiology of *Montastrea curta*, and reported a quantum yield of 0.102 O~2~ photon^-1^.
-    + Value used: 0.1 molC molph^-1^
-* $\bar{a}^*$: Effective light-absorbing cross-section of symbiont
-    + @Hennige:2009p3439 measured spectra-weighted absoprtion coefficients from 400 to 700nm *ā\**~(400-700)~ for eight *Symbiodinium* types in culture at two different light levels, and reported a (grand mean) value of 0.01095 m^2^ mg chl *a*^-1^ [from Table 6].
+```{r yCL, echo=F}
+0.1
+```
+molC molph^-1^
+<br><br>
+
+#### $\bar{a}^*$: Effective light-absorbing cross-section of symbiont {-}
+@Hennige:2009p3439 measured spectra-weighted absoprtion coefficients from 400 to 700nm *ā\**~(400-700)~ for eight *Symbiodinium* types in culture at two different light levels, and reported a (grand mean) value of 0.01095 m^2^ mg chl *a*^-1^ [from Table 6].
 This converts to:
-    + Value used: `r round(0.01095 * mg.g * gchlA.gC * gC.molC, 2)` m^2^ CmolS^-1^
-* $k_{NPQ}$: Capacity of non-photochemical quenching
-    + @Gorbunov:2001p4181 report that the capacity of non-photochemical quenching is approximately 4 times the capacity of photochemistry on a sunny day.
+```{r astar}
+0.01095 * mg.g * gchlA.gC * gC.molC
+```
+m^2^ CmolS^-1^
+<br><br>
+
+#### $k_{NPQ}$: Capacity of non-photochemical quenching {-}
+@Gorbunov:2001p4181 report that the capacity of non-photochemical quenching is approximately 4 times the capacity of photochemistry on a sunny day.
 Therefore, we take the maximum relative capacity of NPQ to be 4 times the maximum rate of photochemical quenching ($j_{CPm}$; see below), giving a value of:
-    + Value used: `r defpars$jCPm / defpars$yCL * 4` mol ph CmolS^-1^ d^-1^
-* $k_{ROS}$: Amount of excitation energy not quenched by photoehcemistry or NPQ that doubles relative production of ROS
-    + @McGinty:2012ck report a ~4-fold increase in ROS production in *Symbiodinium* C1 under stressful conditions relative to control.
+```{r kNPQ}
+defpars$jCPm / defpars$yCL * 4
+```
+mol ph CmolS^-1^ d^-1^
+<br><br>
+
+#### $k_{ROS}$: Amount of excitation energy not quenched by photoehcemistry or NPQ that doubles relative production of ROS {-}
+@McGinty:2012ck report a ~4-fold increase in ROS production in *Symbiodinium* C1 under stressful conditions relative to control.
 We chose a value of $k_{ROS}$ that similarly produced a ~4-fold increase in ROS production at the highest light level (see Figure S2):
-    + Value used: 60 mol ph CmolS^-1^ d^-1^
-* $j_{CPm}$: Maximum specific photosynthesis rate of symbiont (molC CmolS^-1^ d^-1^)
-    + In theory, this value should be the maximum carboxylation rate of *Symbiodinium* Rubisco (RuBP).
+```{r kROS, echo=F}
+80
+```
+mol ph CmolS^-1^ d^-1^
+<br><br>
+
+#### $j_{CPm}$: Maximum specific photosynthesis rate of symbiont {-}
+In theory, this value should be the maximum carboxylation rate of *Symbiodinium* Rubisco (RuBP).
 @Leggat:2004vx report a maximum carboxylation rate at 29°C of 140 CO~2~ RuBP^-1^ s^-1^ [Max.
@@
 143], which converts to:
-    + Value used: `r round(140 * s.h * h.d * RuBP.CmolS, 1)`
-* $j_{SGm}$: Maximum specific growth rate of symbiont (CmolS CmolS^-1^ d^-1^)
-    + Sugget et al. measured specific growth rates of 5 clades of *Symbiodinium* in culture, and reported values ranging from 0.19 to 0.26 d^-1^ (Table 2). 
-    + Value used: 0.25 d^-1^
-* $b$: Scaling parameter for strength of bleaching response (dimensionless)
-    + This value determines the magnitude of symbiont biomass loss (considered as expulsion of symbionts during bleaching) in response to ROS.
+```{r jCPm}
+140 * s.h * h.d * RuBP.CmolS
+```
+molC CmolS^-1^ d^-1^
+<br><br>
+
+#### $j_{SGm}$: Maximum specific growth rate of symbiont {-}
+@Suggett:2015dv measured specific growth rates of 5 clades of *Symbiodinium* in culture, and reported values ranging from 0.19 to 0.26 d^-1^ (Table 2).  
+```{r jSGm, echo=F}
+0.25
+```
+CmolS CmolS^-1^ d^-1^
+<br><br>
+
+#### $b$: Scaling parameter for strength of bleaching response {-}
+This value determines the magnitude of symbiont biomass loss (considered as expulsion of symbionts during bleaching) in response to ROS.
 In this context, relative ROS concentrations are assumed to trigger symbiont expulsion via unknown signaling mechanisms.
@@
 Since this represents symbiont cells being expelled from the system, we use a value about an order of magnitude higher than the turnover rate due to maintenance.
-    + Value used: 5
+```{r b, echo=F}
+5
+```
+(dimensionless)
+<br><br>
 
-## Figure S1. Light amplification and self-shading
-```{r lightamp, fig.width=5, fig.height=5, eval=T, include=T, echo=F, fig.cap="The amplification factor (A) is the ratio of internal scalar irradiance experienced by *Symbiodinium* to the external downwelling irradiance, and is greater than 1 due to multiple scattering of incident light by the coral skeleton.
+## Conversion factors used in parameter derivation
+```{r, code = readLines('../R/conversions.R')}
+```
+
+------
+
+# Figure S1 
+```{r lightamp, fig.width=5, fig.height=5, eval=T, include=T, echo=F, fig.cap="**Fig.
+S1.
+Light amplification and self-shading.** The amplification factor (A) is the ratio of internal scalar irradiance experienced by *Symbiodinium* to the external downwelling irradiance, and is greater than 1 due to multiple scattering of incident light by the coral skeleton.
 The amplification factor decreases as a function of symbiont abundance (converted to symbiont to host biomass ratio) due to self-shading.
@@
 # Get parameters of fit and add to plot
-text(0.13,3.1, labels=paste0(c("A = ", as.character(formula(mod))[3]), collapse=""), adj=0)
-coefs <- round(coef(mod), 2)
-text(0.13,2.9, labels=paste0(c("x=",coefs[1],"; y=",coefs[2],"; z=",coefs[3]), collapse=""), adj=0)
+fiteq <- str_replace_all(format(formula(mod)), c("~"="= ", " x"=coefs[[1]], "y"=coefs[[2]], "z"=coefs[[3]]))
+text(0.05,2.8, fiteq, adj=0)
+
 ```
 
-## Figure S2. Photosynthesis-Irradiance curves of modeled symbiont.
-```{r run1_PI, cache=F, fig.height=4, fig.width=7, echo=F, fig.cap="The amount of light energy quenched through photochemistry (red line) and non-photochemical quenching (black line), and the relative production of reactive oxygen species (orange line) are plotted as a function of the scalar irradiance experienced by the symbiont (note different y-axes for each line).
-These simulations represent the behavior of the photosynthesis SU independent of the host and without CO~2~ limitation of photosynthesis (*j~CO2~* set to a very high value).
+-----
+
+# Figure S2 
+```{r run1_PI, cache=F, fig.height=4, fig.width=7, echo=F, fig.cap="**Fig.
+S2.
+Photosynthesis-Irradiance curves of modeled symbiont.** The rates of light energy quenching through photochemistry (red line) and non-photochemical quenching (black line), and the relative production of reactive oxygen species (orange line) are plotted as a function of the scalar irradiance experienced by the symbiont (note different y-axes for each line).
+These curves represent the behavior of the photosynthesis SU independent of the host and without CO~2~ limitation of photosynthesis (*j~CO2~* set to a very high value).
 The structure of the photosynthesis SU, with the negative feedback between excess light and photosynthesis rate, produces the classic shape of a P-I curve demonstrating photoinhibition, suggesting the mechanism is robust."}
@@
 
-## Figure S3. Sensitivity analysis under high DIN (4e-6 M) and intermediate light (15 mol m^-2^ d^-1^).
-```{r sens_ILHN, cache=T, echo=F}
+-----
+
+# Figure S3 
+```{r sens_ILHN, cache=T, echo=F, eval=F, fig.cap="**Fig. S3. Sensitivity analysis under high DIN (4e-6 M) and intermediate light (15 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
 # Set up time vector and parameters
@@
 
-## Figure S4. Sensitivity analysis under low DIN (1e-7 M) and low light (2 mol m^-2^ d^-1^).
-```{r sens_LLLN, cache=T, echo=F}
+-----
+
+# Figure S4 
+```{r sens_LLLN, cache=T, echo=F, eval=F, fig.cap="**Sensitivity analysis under low DIN (1e-7 M) and low light (2 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
 # Set up time vector and parameters
@@
 
-## Figure S5. Sensitivity analysis under high DIN (4e-6 M) and low light (2 mol m^-2^ d^-1^).
-```{r sens_LLHN, cache=T, echo=F}
+-----
+
+# Figure S5
+```{r sens_LLHN, cache=T, echo=F, eval=F, fig.cap="**Fig. S5. Sensitivity analysis under high DIN (4e-6 M) and low light (2 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
 # Set up time vector and parameters
@@
 
-## Figure S6. Sensitivity analysis under low DIN (1e-7 M) and high light (25 mol m^-2^ d^-1^).
-```{r sens_HLLN, cache=T, echo=F}
+-----
+
+# Figure S6
+```{r sens_HLLN, cache=T, echo=F, eval=F, fig.cap="**Fig. S6. Sensitivity analysis under low DIN (1e-7 M) and high light (25 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
 # Set up time vector and parameters
@@
 
-## Figure S7. Sensitivity analysis under high DIN (4e-6 M) and high light (25 mol m^-2^ d^-1^).
-```{r sens_HLHN, cache=T, echo=F}
+-----
+
+# Figure S7
+```{r sens_HLHN, cache=T, echo=F, eval=F, fig.cap="Fig. S7. Sensitivity analysis under high DIN (4e-6 M) and high light (25 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
 # Set up time vector and parameters
@@
 
-## Figure S8. Full system dynamics of seasonal light variation at two nutrient levels.
-This figure accompanies Figure 4 in the main text, and show additional model fluxes that demonstrate the mechanisms underlying the changes in the system state in these simulations.
+-----
 
-```{r FigS8, cache=T, fig.width=7, fig.height=7}
+# Figure S8
+```{r FigS8, cache=T, fig.width=7, fig.height=7, echo=F, eval=F, fig.cap="**Fig.
+S8.
+Full system dynamics of seasonal light variation at two nutrient levels.** This figure accompanies Figure 5 in the main text, and shows additional model fluxes that demonstrate the mechanisms underlying the changes in the system state in these simulations."}
 # Set parameters
@@
 # Run simulations
-ss1 <- with(run_coral_ss(env=list(L=33,N=0.14e-6,X=1e-5), pars=defpars), last(S$S/H$H))
+ss1 <- with(run_coral_ss(env=list(L=33,N=0.14e-6,X=1e-5), pars=defpars, dt=0.1), last(S$S/H$H))
 run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", ss1))
-ss2 <- with(run_coral_ss(env=list(L=33,N=10.14e-6,X=1e-5), pars=defpars), last(S$S/H$H))
+ss2 <- with(run_coral_ss(env=list(L=33,N=10.14e-6,X=1e-5), pars=defpars, dt=0.1), last(S$S/H$H))
 run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", ss2))
@@
 
-# Figure S9. Hysteresis under different nutrient and feeding regimes.
-```{r runFigS9, cache=T, include=F}
+-----
+
+# Figure S9
+```{r runFigS9, cache=T, include=F, eval=F}
 # Set up cluster for parallel processing
@@
 
-```{r plotFigS9, cache=T, fig.width=7, fig.height=7, echo=F, message=F}
+```{r plotFigS9, cache=T, fig.width=7, fig.height=7, echo=F, message=F, eval=F, fig.cap="**Fig.
+S9.
+Hysteresis under different nutrient and feeding regimes.** This figure accompanies Fig.
+2 in the main text.
+These figures show steady state S:H biomass ratios in constant environments with varying DIN and food (X) levels, with initial S:H biomass values of either 0.1 or 0.0001.
+Points are colored according to whether the host exhibits positive (black) or negative (red) growth at steady state."}
 # Create plot
@@
 
-These figures show steady state S:H biomass ratios in constant environments with varying DIN and food (X) levels, with initial S:H biomass values of either 0.1 or 0.0001.
-Points are colored according to whether the host exhibits positive (black) or negative (red) growth at steady state.
+-----
 
-## References
+# References

### 2017-02-07 JTB_supp/JTB_supp.Rmd
turn on eval to run on remote server

@@
 # Figure S3 
-```{r sens_ILHN, cache=T, echo=F, eval=F, fig.cap="**Fig. S3. Sensitivity analysis under high DIN (4e-6 M) and intermediate light (15 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
+```{r sens_ILHN, cache=T, echo=F, eval=T, fig.cap="**Fig. S3. Sensitivity analysis under high DIN (4e-6 M) and intermediate light (15 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
 # Set up time vector and parameters
@@
 # Figure S4 
-```{r sens_LLLN, cache=T, echo=F, eval=F, fig.cap="**Sensitivity analysis under low DIN (1e-7 M) and low light (2 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
+```{r sens_LLLN, cache=T, echo=F, eval=T, fig.cap="**Sensitivity analysis under low DIN (1e-7 M) and low light (2 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
 # Set up time vector and parameters
@@
 # Figure S5
-```{r sens_LLHN, cache=T, echo=F, eval=F, fig.cap="**Fig. S5. Sensitivity analysis under high DIN (4e-6 M) and low light (2 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
+```{r sens_LLHN, cache=T, echo=F, eval=T, fig.cap="**Fig. S5. Sensitivity analysis under high DIN (4e-6 M) and low light (2 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
 # Set up time vector and parameters
@@
 # Figure S6
-```{r sens_HLLN, cache=T, echo=F, eval=F, fig.cap="**Fig. S6. Sensitivity analysis under low DIN (1e-7 M) and high light (25 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
+```{r sens_HLLN, cache=T, echo=F, eval=T, fig.cap="**Fig. S6. Sensitivity analysis under low DIN (1e-7 M) and high light (25 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
 # Set up time vector and parameters
@@
 # Figure S7
-```{r sens_HLHN, cache=T, echo=F, eval=F, fig.cap="Fig. S7. Sensitivity analysis under high DIN (4e-6 M) and high light (25 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
+```{r sens_HLHN, cache=T, echo=F, eval=T, fig.cap="Fig. S7. Sensitivity analysis under high DIN (4e-6 M) and high light (25 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
 # Set up time vector and parameters
@@
 # Figure S8
-```{r FigS8, cache=T, fig.width=7, fig.height=7, echo=F, eval=F, fig.cap="**Fig.
+```{r FigS8, cache=T, fig.width=7, fig.height=7, echo=F, eval=T, fig.cap="**Fig.
 S8.
@@
 # Figure S9
-```{r runFigS9, cache=T, include=F, eval=F}
+```{r runFigS9, cache=T, include=F, eval=T}
 # Set up cluster for parallel processing
@@
 
-```{r plotFigS9, cache=T, fig.width=7, fig.height=7, echo=F, message=F, eval=F, fig.cap="**Fig.
+```{r plotFigS9, cache=T, fig.width=7, fig.height=7, echo=F, message=F, eval=T, fig.cap="**Fig.
 S9.

### 2017-02-07 JTB_ms/JTB_ms.Rmd
fix footnote duplication

@@
 S8).
-The model of @Muller would also predict higher S:H ratios in winter if light became limiting to photosynthesis, but this is distinct from the mechanism here, in which stress in the summer drives a reduction in S:H.
-These can be distinguished as follows:
-
-winter increase in S:H as light limits carbon fixation - decreases growth in winter
-summer decrease in S:H as photo-oxidative stress limits carbon fixation - decreases growth in summer
-
-under high light in summer, photooxidative stress decreases symbiont biomass and increases carbon-limitation of host growth regardless of nutrient status; when light is reduced in winter, growth and symbiont biomass become constrained instead by nitrogen, and therefore increase more when DIN is enriched (Fig.
-S8).
 
@@
 
-The prediction of higher growth when light is reduced indicates that growth is not limited by low light in winter, but is actually reduced by excess light in summer[^2], consistent with the experimental findings of @Schutter:2011ie.
+The prediction of higher growth when light is reduced indicates that growth is not limited by low light in winter, but is actually reduced by excess light in summer[^3], consistent with the experimental findings of @Schutter:2011ie.
 Seasonal summertime reductions in tissue biomass have also been well-documented in the field [@Fitt:2000p3792], along with reductions in net photosynthetic capacity [@MullerParker:1987cr].
-Importantly, while light alone may drive seasonal dynamics in the ways discussed[^2], temperature fluctuations may attenuate or even reverse the effect of light as cooler winters depress metabolism; thus, the relative magnitude of fluctuation in temperature, light, and other factors may produce wide variability in the direction and magnitude of seasonal changes in growth and symbiont abundance, depending on location and microhabitat.
+Importantly, while light alone may drive seasonal dynamics in the ways discussed[^3], temperature fluctuations may attenuate or even reverse the effect of light as cooler winters depress metabolism; thus, the relative magnitude of fluctuation in temperature, light, and other factors may produce wide variability in the direction and magnitude of seasonal changes in growth and symbiont abundance, depending on location and microhabitat.
 Nevertheless, the seasonal variability predicted here (Fig.
@@
 
-[^2]: at the light levels indicated in @Stimson:1997p3837.
+[^3]: at the light levels indicated in @Stimson:1997p3837.
 Note that if light levels were reduced throughout the year (e.g.

### 2017-02-07 JTB_ms/JTB_ms.Rmd
remove kahako (causing latex error)

@@
   - code: HIMB
-    address: Hawaii Institute of Marine Biology, School of Ocean and Earth Science and Technology, University of Hawaii at Mānoa, USA
+    address: Hawaii Institute of Marine Biology, School of Ocean and Earth Science and Technology, University of Hawaii at Manoa, USA
   - code: UCSB

### 2017-02-07 JTB_ms/JTB_ms.tex
remove kahako (causing latex error)

@@
             pdfauthor={},
-            pdftitle={A simplified dynamic bioenergetic model for coral-Symbiodinium symbioses and coral bleaching as an alternate stable state},
+            pdftitle={A dynamic bioenergetic model for coral-Symbiodinium symbioses and coral bleaching as an alternate stable state},
             colorlinks=true,
@@
 
-  \title{A simplified dynamic bioenergetic model for coral-\emph{Symbiodinium}
-symbioses and coral bleaching as an alternate stable state}
+  \title{A dynamic bioenergetic model for coral-\emph{Symbiodinium} symbioses and
+coral bleaching as an alternate stable state}
     \author[University of Hawaii]{Ross Cunning\corref{c1}}
@@
    \cortext[c1]{Corresponding Author}
-    \author[University of California]{Erik B. Muller}
-
-    \author[University of Hawaii]{Ruth D. Gates}
-
-    \author[University of California]{Roger M. Nisbet}
-
-      \address[University of Hawaii]{Hawaii Institute of Marine Biology, Kaneohe, HI 96744, USA}
-    \address[University of California]{Department of Ecology, Evolution, and Marine Biology, Santa Barbara, CA
-93106, USA}
+    \author[UCSB, MSI]{Erik B. Muller}
+
+    \author[HIMB]{Ruth D. Gates}
+
+    \author[UCSB]{Roger M. Nisbet}
+
+      \address[HIMB]{Hawaii Institute of Marine Biology, School of Ocean and Earth Science
+and Technology, University of Hawaii at Manoa, USA}
+    \address[UCSB]{Department of Ecology, Evolution, and Marine Biology, University of
+California, Santa Barbara, USA}
+    \address[MSI]{Marine Science Institute, University of California, Santa Barbara, USA}
 
@@
   stability and breakdown of these symbioses (i.e., coral `bleaching')
-  demands the development and application of theoretical tools; however,
-  the complexity of some approaches may limit their potential application
-  and utility. Here, we develop a simplified dynamic bioenergetic model of
-  coral-\emph{Symbiodinium} symbioses that demonstrates realistic
-  steady-state patterns in coral growth and symbiont abundance across
-  gradients of light, nutrients, and feeding. Furthermore, through a
-  mechanistic treatment of photo-oxidative stress, the model displays
-  dynamic behaviors of bleaching and recovery, and suggests that these
-  phenomena represent transitions between alternate stable states. A suite
-  of complex responses to multiple, interacting environmental factors
-  reproduced by the model suggests it unifyingly captures many important
-  attributes of the system; meanwhile, its modular framework and open
-  source R code are designed to facilitate expansion and problem-specific
-  applications. We see significant potential applications for this
-  modeling framework in generating testable hypotheses and predicting
-  integrated, mechanistic responses of corals to environmental change.
+  demands the development and application of theoretical tools. Here, we
+  develop a dynamic bioenergetic model of coral-\emph{Symbiodinium}
+  symbioses that demonstrates realistic steady-state patterns in coral
+  growth and symbiont abundance across gradients of light, nutrients, and
+  feeding. Furthermore, by including a mechanistic treatment of
+  photo-oxidative stress, the model displays dynamics of bleaching and
+  recovery that can be explained as transitions between alternate stable
+  states. These dynamics reveal that `healthy' and `bleached' states
+  correspond broadly to nitrogen- and carbon-limitation of the system,
+  with transitions between them occurring as integrated responses to
+  multiple environmental factors. Indeed, a suite of complex emergent
+  behaviors reproduced by the model (e.g., bleaching is exacerbated by
+  nutrients and attenuated by feeding) suggests it unifyingly captures
+  many important attributes of the system; meanwhile, its modular
+  framework and open source R code are designed to facilitate further
+  problem-specific development. We see significant potential for this
+  modeling framework to generate testable hypotheses and predict
+  integrated, mechanistic responses of corals to environmental change,
+  with important implications for understanding the performance and
+  maintenance of symbiotic systems.
   \end{abstract}
@@
 \section{Introduction}\label{introduction}
+
+MAKE INTRO INCLUDE SOMETHING ABOUT ALTERNATIVE STABLE STATES
 
@@
 across gradients of ambient irradiance, nutrients, and food. This model
-assumed that 1) \emph{Symbiodinium} has priority access to fixed carbon
-through photosynthesis, 2) the coral animal has priority access to
-dissolved nitrogen through contact with seawater, and 3) each partner
-shares with the other only what it cannot use for its own growth. In its
-simplest form, this principle of sharing the surplus is sufficient to
-describe the dynamics of diverse syntrophic organs and organisms (e.g.,
-trees, duckweeds, corals), suggesting the mechanism is mathematically
-and evolutionarily robust (Nisbet et al., in prep.).
+(as well as the present work) assumes that 1) \emph{Symbiodinium} has
+priority access to fixed carbon through photosynthesis, 2) the coral
+animal has priority access to nitrogen through contact with seawater and
+feeding on prey, and 3) each partner shares with the other only what it
+cannot use for its own growth. In its simplest form, this principle of
+sharing the surplus is sufficient to describe the dynamics of diverse
+syntrophic organs and organisms (e.g., trees, duckweeds, corals),
+suggesting the mechanism is mathematically and evolutionarily robust
+(Nisbet et al., in prep.).
 
@@
 \item
-  \emph{Develop a module of photooxidative stress.} Of primary interest
-  to coral biologists and ecologists is symbiosis dysfunction under
-  environmental stress, resulting in coral ``bleaching''--the loss of
-  algal symbionts from the association (Jokiel and Coles 1977).
-  Photooxidative stress in \emph{Symbiodinium} is considered a primary
-  trigger of bleaching in response to high temperature and/or light
-  (Weis 2008), and prolonged or severe bleaching can result in
-  mortality, though corals sometimes recover their symbionts. Bleaching
-  susceptibility, severity, and recovery may be influenced by
-  interacting factors such as heterotrophy and nutrient availability
-  (Wooldridge 2014b). To simulate these events and interactions, we
-  develop a photooxidative stress module linking overreduction of the
-  photosynthetic light reactions to downstream impacts of
-  photoinhibition and photodamage.
+  \emph{Develop a new module of photo-oxidative stress.} Of primary
+  interest to coral biologists and ecologists is symbiosis dysfunction
+  under environmental stress, resulting in coral ``bleaching''--the loss
+  of algal symbionts from the association (Jokiel and Coles 1977).
+  Bleaching is thought to be triggered by photo-oxidative stress in
+  \emph{Symbiodinium} (Weis 2008), which has been modeled previously as
+  a response to absolute external irradiance (Eynaud, Nisbet, and Muller
+  2011). However, this response may also depend on self-shading by other
+  symbionts, CO\textsubscript{2} availability at the site of
+  photosynthesis, and other non-photochemical quenching mechanisms. We
+  incorporate these features into a new photo-oxidative stress module
+  linking overreduction of the photosynthetic light reactions to
+  downstream impacts of photoinhibition and photodamage. Importantly,
+  this formulation introduces a link between
+  CO\textsubscript{2}-limitation of photosynthesis and bleaching, and
+  potential synergistic roles for heterotrophy and nutrient availability
+  in influencing bleaching responses (Wooldridge 2014b)
 \item
@@
   reserves, maturity, or reproduction (see Kooijman 2010). This
-  formulation restricts the model's scope to the bioenergetics of adult
-  corals (i.e., reproduction, larval stages, and metamorphosis are not
-  considered), but greatly reduces theoretical complexity and parameter
-  numbers, which is advantageous given the relative paucity of data for
-  corals. However, our primary motivation for reducing complexity was to
-  increase accessibility and applicability for biologists and ecologists
-  without requiring significant expertise in DEB theory.
+  formulation precludes modeling the full life cycle of corals as
+  reproduction, larval stages, and metamorphosis are not considered, but
+  greatly reduces theoretical complexity and parameter numbers, which is
+  advantageous given the relative paucity of data for corals. Moreover,
+  dynamics of the symbiosis (i.e., changes in symbiont to host biomass,
+  including bleaching and recovery) and coral biomass growth are
+  efficiently captured with this simpler formulation, which also
+  increases accessibility for biologists and ecologists without
+  requiring significant expertise in DEB theory.
 \item
@@
   language (R Core Team 2014). With an accessible and modular framework,
-  we envision this model as a resource for futher development by the
+  we envision this model as a resource for further development by the
   scientific community to include additional complexity and
@@
 1, and each model flux and parameter is defined in Tables 1 and 2,
-respectively. We use C-moles as the unit of biomass for consistency with
-the rigorous mass balance of DEB theory: 1 C-mole is equivalent to the
-amount of biomass containing 1 mole of carbon atoms. Host biomass
-(\(H\)), symbiont biomass (\(S\)), and prey biomass (\(X\)) have fixed,
-but different, molar N:C ratios (Table 2). Biomass is produced from
-carbon and nitrogen by synthesizing units (SU), which are mathematical
+respectively. We use C-moles (C-mol) as the unit of biomass for
+consistency with the rigorous mass balance of DEB theory: 1 C-mol is
+equivalent to the amount of biomass containing 1 mole of carbon atoms.
+Host biomass (expressed as C-mol H), symbiont biomass (expressed as
+C-mol S), and prey biomass (expressed as C-mol X) have fixed, but
+different, molar N:C ratios (Table 2). Biomass is produced from carbon
+and nitrogen by synthesizing units (SU), which are mathematical
 specifications of the formation of a product from two substrates; we use
-the parallel complementary formulation of Kooijman (2010) to specify
-these fluxes. The two state variables of this system are symbiont
-biomass and coral biomass; because resources are acquired proportionally
-to surface area, and surface area is proportional to volume (i.e.,
-corals are ``V1-morphs'' in DEB terminology (Kooijman 2010)), biomass
-increases exponentially during growth (indeed, corals grow exponentially
-(Bak 1976)).
+the ``parallel complementary'' formulation of Kooijman (2010) (page 105,
+Fig. 3.7) to specify these fluxes. The two state variables of this
+system are symbiont biomass and coral biomass; because resources are
+acquired proportionally to surface area, and surface area is
+proportional to volume, biomass increases exponentially during growth
+(indeed, corals grow exponentially (Bak 1976)).
 
 Environmental stress is implemented in the form of photooxidative
-stress, which is thought to be the primary trigger of coral bleaching
-(Lesser 1997; Weis 2008; Wooldridge 2009). To simulate these events, we
+stress, which is thought to be a primary trigger of coral bleaching
+(Lesser 1997; Weis 2008; Wooldridge 2009). To simulate bleaching, we
 model the absorption and quenching of light energy by photochemistry and
-non-photochemical quenching, and the phenomenological responses that
-occur when these capacities are overwhelmed (i.e., photoinhibition,
-photodamage, and symbiont loss). While bleaching in response to high
-light alone has been observed experimentally (Schutter et al. 2011;
-Downs et al. 2013), mass coral bleaching events occur in response to
-high temperature (Hoegh-Guldberg 1999); thus, it is important to justify
-our consideration of light as the primary stressor. In reality, light
-and temperature interact synergistically (Coles and Jokiel 1978; Jones
-et al. 1998), and in fact, any stressor that disrupts the quenching of
+non-photochemical quenching, and the responses that occur (i.e.,
+photoinhibition, photodamage, and symbiont loss) when these quenching
+capacities are overwhelmed. While bleaching in response to high light
+alone has been observed experimentally (Schutter et al. 2011; Downs et
+al. 2013), mass coral bleaching events occur concurrently with high
+temperature (Hoegh-Guldberg 1999); thus, it is important to justify our
+consideration of light as the primary stressor. In reality, light and
+temperature interact synergistically (Coles and Jokiel 1978; Jones et
+al. 1998), and in fact, any stressor that disrupts the quenching of
 light energy may lead to bleaching (Wooldridge 2010; Baker and Cunning
@@
 
-Table 1. Model fluxes (mass-specific)
-
-\begin{longtable}[c]{@{}llll@{}}
+Table 1. Model fluxes (mass-specific). Units are explained in the text.
+
+\begin{longtable}[]{@{}llll@{}}
 \toprule
-\begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
-Symbol
-\strut\end{minipage} &
-\begin{minipage}[b]{0.48\columnwidth}\raggedright\strut
-Description
-\strut\end{minipage} &
-\begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
-Units
-\strut\end{minipage} &
-\begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
-Eq. no.
-\strut\end{minipage}\tabularnewline
+Symbol & Description & Units & Eq. no.\tabularnewline
 \midrule
 \endhead
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_X\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Prey uptake rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molX CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-3
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_N\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Nitrogen uptake rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molN CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-4
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{HG}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Host biomass formation rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-CmolH CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-5
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(r_{NH}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Recycled nitrogen from host turnover
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molN CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-6
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(\rho_N\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Nitrogen shared with the symbiont
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molN CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-7
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{eC}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Excess carbon used to activate host CCMs
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molC CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-8
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{CO_2}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-CO\textsubscript{2} input to photosynthesis
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molCO\textsubscript{2} CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-9
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_L\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Light absorption rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molph CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-11
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(r_{CH}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Recycled CO\textsubscript{2} from host turnover
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molCO\textsubscript{2} CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-12
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(r_{CS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Recycled CO\textsubscript{2} from symbiont turnover
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molCO\textsubscript{2} CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-13
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{CP}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Photosynthesis rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molC CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-14
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{eL}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Light energy in excess of photochemistry
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molph CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-15
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{NPQ}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Total capacity of NPQ
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molph CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-16
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(c_{ROS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-ROS production proportional to baseline
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-17
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(r_{NS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Recycled nitrogen from symbiont turnover
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molN CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-18
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{SG}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Symbiont biomass formation rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-19
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(\rho_C\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Fixed carbon shared with host
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molC CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-20
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{ST}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Symbiont turnover rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-21
-\strut\end{minipage}\tabularnewline
+\(j_X\) & Prey uptake rate & mol X C-mol H\textsuperscript{-1}
+d\textsuperscript{-1} & 3\tabularnewline
+\(j_N\) & Nitrogen uptake rate & mol N Cmol H\textsuperscript{-1}
+d\textsuperscript{-1} & 4\tabularnewline
+\(j_{HG}\) & Host biomass formation rate & C-mol H C-mol
+H\textsuperscript{-1} d\textsuperscript{-1} & 5\tabularnewline
+\(j_{HT}\) & Host biomass turnover rate & C-mol H C-mol
+H\textsuperscript{-1} d\textsuperscript{-1} & 21\tabularnewline
+\(r_{NH}\) & Recycled nitrogen from host turnover & mol N C-mol
+H\textsuperscript{-1} d\textsuperscript{-1} & 6\tabularnewline
+\(\rho_N\) & Nitrogen shared with the symbiont & mol N C-mol
+H\textsuperscript{-1} d\textsuperscript{-1} & 7\tabularnewline
+\(j_{eC}\) & Excess carbon used to activate host CCMs & mol C C-mol
+H\textsuperscript{-1} d\textsuperscript{-1} & 8\tabularnewline
+\(j_{CO_2}\) & CO\textsubscript{2} input to photosynthesis & mol
+CO\textsubscript{2} C-mol H\textsuperscript{-1} d\textsuperscript{-1} &
+9\tabularnewline
+\(j_L\) & Light absorption rate & mol q C-mol S\textsuperscript{-1}
+d\textsuperscript{-1} & 11\tabularnewline
+\(r_{CH}\) & Recycled CO\textsubscript{2} from host turnover & mol
+CO\textsubscript{2} C-mol H\textsuperscript{-1} d\textsuperscript{-1} &
+12\tabularnewline
+\(r_{CS}\) & Recycled CO\textsubscript{2} from symbiont turnover & mol
+CO\textsubscript{2} C-mol S\textsuperscript{-1} d\textsuperscript{-1} &
+13\tabularnewline
+\(j_{CP}\) & Photosynthesis rate & mol C C-mol S\textsuperscript{-1}
+d\textsuperscript{-1} & 14\tabularnewline
+\(j_{eL}\) & Light energy in excess of photochemistry & mol q C-mol
+S\textsuperscript{-1} d\textsuperscript{-1} & 15\tabularnewline
+\(j_{NPQ}\) & Total capacity of NPQ & mol q C-mol S\textsuperscript{-1}
+d\textsuperscript{-1} & 16\tabularnewline
+\(c_{ROS}\) & ROS production proportional to baseline & -- &
+17\tabularnewline
+\(r_{NS}\) & Recycled nitrogen from symbiont turnover & mol N C-mol
+S\textsuperscript{-1} d\textsuperscript{-1} & 18\tabularnewline
+\(j_{SG}\) & Symbiont biomass formation rate & C-mol S C-mol
+S\textsuperscript{-1} d\textsuperscript{-1} & 19\tabularnewline
+\(\rho_C\) & Fixed carbon shared with host & mol C C-mol
+S\textsuperscript{-1} d\textsuperscript{-1} & 20\tabularnewline
+\(j_{ST}\) & Symbiont biomass turnover rate & C-mol S C-mol
+S\textsuperscript{-1} d\textsuperscript{-1} & 21\tabularnewline
 \bottomrule
@@
 
-Table 2. Model parameters. Derivations of each parameter value along
-with supporting references are provided in the Supplementary
-Information.
-
-\begin{longtable}[c]{@{}llll@{}}
+Table 2. Model parameters (units explained in the text). Derivations of
+each parameter value along with supporting references are provided in
+the Supplementary Information.
+
+\begin{longtable}[]{@{}llll@{}}
 \toprule
-\begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
-Symbol
-\strut\end{minipage} &
-\begin{minipage}[b]{0.48\columnwidth}\raggedright\strut
-Description
-\strut\end{minipage} &
-\begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
-Units
-\strut\end{minipage} &
-\begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
-Value
-\strut\end{minipage}\tabularnewline
+\begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
+Symbol\strut
+\end{minipage} & \begin{minipage}[b]{0.46\columnwidth}\raggedright\strut
+Description\strut
+\end{minipage} & \begin{minipage}[b]{0.24\columnwidth}\raggedright\strut
+Units\strut
+\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
+Value\strut
+\end{minipage}\tabularnewline
 \midrule
 \endhead
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(n_{NH}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-N:C molar ratio in host biomass
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(0.18\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(n_{NS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-N:C molar ratio in symbiont biomass
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(0.13\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(n_{NX}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-N:C molar ratio in prey biomass
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(0.2\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{HT}^0\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Specific turnover rate of host biomass
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-CmolH CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(0.03\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{ST}^0\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Specific turnover rate of symbiont biomass
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(0.03\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(\sigma_{NH}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Proportion host N-turnover recycled to host
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(0.9\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(\sigma_{CH}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Proportion host C-turnover recycled to photosynthesis
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(0.1\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(\sigma_{NS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Proportion symbiont N-turnover recycled to symbiont
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(0.9\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(\sigma_{CS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Proportion symbiont C-turnover recycled to photosynthesis
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(0.9\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{Xm}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Maximum specific feeding rate of host
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molX CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(0.13\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(K_X\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Half-saturation constant for prey uptake by host
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molX L\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(1.95\times 10^{-5}\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{Nm}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Maximum specific DIN uptake rate by host
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molN CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(0.035\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(K_N\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Half-saturation constant for DIN uptake by host
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molN L\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(1.5\times 10^{-6}\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(k_{CO_2}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Efficacy of CO\textsubscript{2} delivery by host CCMs
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molCO\textsubscript{2} molC\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(10\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{HGm}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Maximum specific growth rate of host
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-CmolH CmolH\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(1\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(y_{CL}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Quantum yield of photosynthesis
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molC mol ph\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(0.1\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(\bar{a}^*\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Effective light-absorbing cross-section of symbiont
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-m\textsuperscript{2} CmolS\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(1.34\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(k_{NPQ}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-NPQ capacity of symbiont
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-mol ph CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(112\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(k_{ROS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(n_{NH}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+N:C molar ratio in host biomass\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+--\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(0.18\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(n_{NS}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+N:C molar ratio in symbiont biomass\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+--\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(0.13\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(n_{NX}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+N:C molar ratio in prey biomass\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+--\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(0.2\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{HT}^0\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+Specific maintenance rate of host biomass\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(0.03\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{ST}^0\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+Specific maintenance rate of symbiont biomass\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(0.03\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(\sigma_{NH}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+Proportion host N-turnover recycled to host\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+--\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(0.9\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(\sigma_{CH}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+Proportion host C-turnover recycled to photosynthesis\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+--\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(0.1\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(\sigma_{NS}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+Proportion symbiont N-turnover recycled to symbiont\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+--\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(0.9\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(\sigma_{CS}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+Proportion symbiont C-turnover recycled to photosynthesis\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+--\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(0.9\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{Xm}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+Maximum specific feeding rate of host\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+molX C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(0.13\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(K_X\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+Half-saturation constant for prey uptake by host\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+molX L\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(1.95\times 10^{-5}\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{Nm}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+Maximum specific DIN uptake rate by host\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+mol N C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(0.035\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(K_N\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+Half-saturation constant for DIN uptake by host\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+mol N L\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(1.5\times 10^{-6}\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(k_{CO_2}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+Efficacy of CO\textsubscript{2} delivery by host CCMs\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+mol CO\textsubscript{2} mol C\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(10\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{HGm}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+Maximum specific growth rate of host\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(1\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(y_{CL}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+Quantum yield of photosynthesis\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+mol C mol q\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(0.1\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(\bar{a}^*\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+Effective light-absorbing cross-section of symbiont\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+m\textsuperscript{2} C-mol S\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(1.34\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(k_{NPQ}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+NPQ capacity of symbiont\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+mol ph C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(112\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(k_{ROS}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
 Excess photon energy that doubles ROS production, relative to baseline
-levels
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-mol ph CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(80\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{CPm}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Maximum specific photosynthesis rate of symbiont
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molC CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(2.8\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{SGm}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Maximum specific growth rate of symbiont
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(0.25\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(b\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.48\columnwidth}\raggedright\strut
-Scaling parameter for strength of bleaching response
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(5\)
-\strut\end{minipage}\tabularnewline
+levels\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+mol ph C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(80\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{CPm}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+Maximum specific photosynthesis rate of symbiont\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+mol C C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(2.8\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{SGm}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+Maximum specific growth rate of symbiont\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(0.25\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(b\)\strut
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+Scaling parameter for strength of bleaching response\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+--\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(5\)\strut
+\end{minipage}\tabularnewline
 \bottomrule
@@
 The balance equations for symbiont (\(S\)) and host (\(H\)) biomass are
-expressed as:
+expressed as ``specific'' rates, i.e.~rates per unit of symbiont and
+host biomass, respectively:
 
@@
 
-\begin{equation} {dH \over Hdt} = j_{HG} - j_{HT}^0 \end{equation}
-
-These are expressed per unit of symbiont and host biomass, respectively.
+\begin{equation} {dH \over Hdt} = j_{HG} - j_{HT} \end{equation}
+
 The specific biomass growth and turnover rates that define these balance
@@
 utilization; this capacity is supported by experimental evidence (Wang
-and Douglas 1998) and is consistent with the physical arrangment of the
+and Douglas 1998) and is consistent with the spatial arrangment of the
 partners, where the host is in direct contact with the external
@@
 Coral biomass formation is then specified by a parallel complementary SU
-that combines carbon and nitrogen, according to:
+(formula in Fig. 3.7 of Kooijman (2010){]}) that combines carbon and
+nitrogen, according to:
 
@@
 where \(\rho_C\) is fixed carbon shared by the symbiont (see Eq. 20),
-and \(r_{NH}\) is recycled nitrogen liberated by host biomass turnover:
-
-\begin{equation} r_{NH}=\sigma_{NH}n_{NH}j_{HT}^0 \end{equation}
+and \(r_{NH}\) is recycled nitrogen liberated by host biomass turnover
+(see Eq. XX). Host biomass turnover is equal to the specific maintenance
+rate of host biomass, which is set as a parameter \(j_{HT}^0\),
+
+\begin{equation} j_{HT}=j_{HT}^0 \end{equation}
+
+and the specific flux of nitrogen that is recycled to the host biomass
+SU is calcluated as:
+
+\begin{equation} r_{NH}=\sigma_{NH}n_{NH}j_{HT} \end{equation}
 
@@
 indirectly ensures its own CO\textsubscript{2} supply by providing fixed
-carbon (=energy) to the host (Wooldridge 2013), which establishes a
-positive feedback in the model that leads to the existence of multiple
-stable states, which we discuss below in the context of coral bleaching.
-The parameter \(k_{CO_2}\) scales the efficacy of host CCMs, which
-enables the comparison of different rates of CO\textsubscript{2}
-delivery that may characterize different coral species (Wooldridge
-2014a). The active input of CO\textsubscript{2} to the photosynthesis SU
-is therefore specified as:
+carbon (=energy) to the host (Wooldridge 2013). The parameter
+\(k_{CO_2}\) scales the efficacy of host CCMs, which enables the
+comparison of different rates of CO\textsubscript{2} delivery that may
+characterize different coral species (Wooldridge 2014a). The active
+input of CO\textsubscript{2} to the photosynthesis SU is therefore
+specified as:
 
@@
 This amplification factor is then multiplied by the external downwelling
-irradiance \(L\) and the effective light-absorbing surface area of
-symbiont biomass \(\bar{a}^*\) to specify the total light absorption:
+irradiance \(L\) and a parameter representing the effective
+light-absorbing surface area of symbiont biomass \(\bar{a}^*\) to
+specify the total light absorption:
 
@@
 sources: in addition to the CO\textsubscript{2} actively supplied by the
-host through its CCMs (\(j_{CO_2}\); Eq. 9), a fixed proportion of
-metabolic CO\textsubscript{2} generated by host biomass turnover is
-passively available to the photosynthesis SU, according to:
-
-\begin{equation} r_{CH}=\sigma_{CH}j_{HT}^0 \end{equation}
+host through its CCMs (\(j_{CO_2}\); Eq. 9), we assume a fixed
+proportion \(\sigma_{CH}\) of metabolic CO\textsubscript{2} generated by
+host biomass turnover is passively available to the photosynthesis SU,
+according to:
+
+\begin{equation} r_{CH}=\sigma_{CH}j_{HT} \end{equation}
 
 along with a fixed proportion of CO\textsubscript{2} generated by
-symbiont biomass turnover:
+symbiont biomass turnover\footnote{Note that recycling of symbiont
+  biomass turnover (\(rNS\) and \(rCS\)) only occurs based on the
+  maintenance component of turnover (i.e., \(j_{ST}^0\)), and not the
+  photodamage/bleaching component, as this loss represents biomass that
+  is expelled from the host.}:
 
@@
 Importantly, \(c_{ROS}\) is specified here not as a function of absolute
-external light, but rather the amount of excess light after accounting
-for quenching by carbon fixation and NPQ. A direct consequence of this
-formulation is that CO\textsubscript{2}-limitation of photosynthesis can
-lead to ROS production, an important mechanism (Wooldridge 2009) that
-was not captured by previous representations of photooxidative stress
-(Eynaud, Nisbet, and Muller 2011).
+external light, but rather the amount of excess light energy after
+accounting for quenching by carbon fixation and NPQ. A direct
+consequence of this formulation is that CO\textsubscript{2}-limitation
+of photosynthesis can lead to ROS production, an important mechanism
+(Wooldridge 2009) that was not captured by previous representations of
+photooxidative stress (Eynaud, Nisbet, and Muller 2011). With this
+single SU, both the light and dark reactions of photosynthesis are
+represented, allowing for sink-limitation (i.e.,
+CO\textsubscript{2}-limitation) to cause overreduction of the electron
+transport chain and ROS production.
 
@@
 with nitrogen shared by the host (\(\rho_N\); Eq. 7) and nitrogen
-recycled from symbiont biomass turnover
+recycled from symbiont biomass turnover\footnote{Note that recycling of
+  symbiont biomass turnover (\(rNS\) and \(rCS\)) only occurs based on
+  the maintenance component of turnover (i.e., \(j_{ST}^0\)), and not
+  the photodamage/bleaching component, as this loss represents biomass
+  that is expelled from the host.}
 
@@
 
-The rejection flux of nitrogen from the symbiont biomass SU is lost to
-the environment.
+Nitrogen rejected by the symbiont biomass SU, which has already been
+rejected by the host biomass SU, cannot be used by either partner and is
+thus lost to the environment.
 
@@
 is included to scale biomass loss due to bleaching in response to ROS.
-(Note that recycling of symbiont biomass turnover (\(rNS\) and \(rCS\))
-only occurs based on the maintenance component of turnover (i.e.,
-\(j_{ST}^0\)), and not the photodamage/bleaching component, as this loss
-represents biomass that is damaged or expelled from the host.)
 
@@
 of these outcomes may occur depending on initial values of symbiont and
-host biomass, indicating the presence of alternate stable states,
-corresponding to either positive or negative growth (Fig. 2). The
-mechanism that produces these alternate stable states is the positive
-feedback between carbon-limitation of the host and
+host biomass, indicating the presence of alternate stable states (Fig.
+2). The mechanism that produces these alternate stable states is the
+positive feedback between carbon-limitation of the host and
 CO\textsubscript{2}-limitation of photosynthesis: if symbiont biomass is
 initially very low (i.e., a ``bleached'' coral), the system cannot
-escape this positive feedback and cannot grow. However, if symbiont
-biomass is initially high (i.e., a ``healthy'' coral), then the system
-remains in a nitrogen-limited state with positive growth. For practical
-purposes, this section of the manuscript considers only positive growth
-steady states under constant environments; subsequently, we explore how
+escape this positive feedback and cannot grow (unless feeding is
+sufficiently high). However, if symbiont biomass is initially high
+(i.e., a ``healthy'' coral), then the system remains in a
+nitrogen-limited state with positive growth. For practical purposes,
+this section of the manuscript considers only positive growth steady
+states under constant environments; subsequently, we explore how
 environmental forcing may cause the system to switch between alternate
@@
 Cmol\textsuperscript{-1} d\textsuperscript{-1}) and (B) the symbiont to
-host biomass ratio (CmolS CmolH\textsuperscript{-1}) across gradients of
-external irradiance and dissolved inorganic nitrogen. Note that typical
-conditions for reefs are \textasciitilde{}1e-7 M DIN and 10-20 mol
-photons m\textsuperscript{-2} d\textsuperscript{-1}. Simulations for
-each combination of light and nutrients (41 points along each axis) were
-run for 100 days with a time step of 1 day. All parameters were set to
-default values and prey density set to zero. Negative steady state
-growth rates, and corresponding S:H ratios, were set to zero.}
+host biomass ratio (C-mol S C-mol H\textsuperscript{-1}) across
+gradients of external irradiance and dissolved inorganic nitrogen. Note
+that typical conditions for reefs are \textasciitilde{}1e-7 M DIN and
+10-20 mol photons m\textsuperscript{-2} d\textsuperscript{-1}.
+Simulations for each combination of light and nutrients (41 points along
+each axis) were run for 100 days with a time step of 1 day. All
+parameters were set to default values and prey density set to zero.
+Negative steady state growth rates, and corresponding S:H ratios, were
+set to zero.}
 \end{figure}
@@
 simulations) has little effect on growth as carbon becomes limiting.
-Very high nutrient levels may even reduce growth (Shantz, Lemoine, and
-Burkepile 2015), although these impacts are not likely to occur within
-the range of concentrations considered here (\textless{}4 µM)
+Although very high nutrient levels may reduce growth in nature (Shantz,
+Lemoine, and Burkepile 2015), these impacts are not likely to occur
+within the range of concentrations considered here (\textless{}4 µM)
 (Ferrier-Pagès et al. 2000). In addition to increasing growth, DIN also
@@
 \textasciitilde{}4 µM DIN (Fig. 3A), are comparable to the rate of 0.07
-d\textsuperscript{-1} measured by Tanaka et al. (2007) under similar
-N-enriched conditions. Predicted growth rates under conditions more
-typical of reef environments (\textless{}0.5 µM DIN) are
-\textasciitilde{}0.01-0.03 d\textsuperscript{-1}. Observed specific
-growth rates, which are typically based on skeletal mass, fall near or
-below 0.01 d\textsuperscript{-1} (Osinga et al. 2011; Osinga et al.
-2012), though values as high as 0.025 d\textsuperscript{-1} have been
-reported (Schutter et al. 2010). It is not surprising that predicted
-growth rates are slightly higher than most observations, as the model
-does not account for external ecological factors that may reduce growth,
-such as competition, predation, and bioerosion. Furthermore, the model
-does not explicitly account for skeletal formation, which may not always
-correlate with biomass growth, potentially due to differential
-substrate-limitation (Anthony 2002). Nevertheless, when growth of coral
-tissue biomass has been directly measured over short periods, values are
-consistent with model predictions (e.g., 0.07 d\textsuperscript{-1} in
-Tanaka et al. (2007)), and similar rates (0.04 d\textsuperscript{-1})
-have been measured in \emph{Aiptasia diaphana}, a non-calcifying
-symbiotic anemone (Armoza-Zvuloni et al. 2014).
+d\textsuperscript{-1} measured by Tanaka et al. (2007) in \emph{Acropora
+pulchra} under similar N-enriched conditions. Under conditions more
+typical of reef environments (\textless{}0.5 µM DIN), predicted growth
+rates are \textasciitilde{}0.01-0.03 d\textsuperscript{-1}. Observed
+specific growth rates in several coral species fall near or below the
+lower end of this range (\textasciitilde{}0.01 d\textsuperscript{-1})
+(Osinga et al. 2011; Osinga et al. 2012), though values as high as 0.025
+d\textsuperscript{-1} have been reported in \emph{Galaxea fascicularis}
+(Schutter et al. 2010), and 0.04 d\textsuperscript{-1} in \emph{Aiptasia
+diaphana}, a non-calcifying symbiotic anemone (Armoza-Zvuloni et al.
+2014). It is not surprising that the model predicts slightly higher
+growth than is typically observed, since it does not account for
+ecological factors that may limit growth (e.g., competition, predation,
+bioerosion). Furthermore, while most measurements are made on skeletal
+growth, the model predicts biomass growth, which may not always be
+strongly correlated (Anthony 2002).
 
@@
 and more realistic, variation in state variables across light gradients
-than was predicted by the model of Muller et al. (2009), which did not
-include photoinhibition and photodamage. It is important to recognize
-that the boundary set by photooxidative stress on a stable symbiosis
-under steady state conditions (Fig. 3) may be temporarily crossed by a
-dynamic system, which may experience a period of symbiont loss
-(bleaching) and reduced growth, after which a return to benign
-conditions may restore symbiont biomass and positive growth. To explore
-this further and illustrate the behavior of the model in more detail, we
-evaluate a number of dynamic simulations below (see ``Dynamic
-behavior'').
+than was predicted by the models of Muller et al. (2009), which did not
+include photoinhibition or photodamage, or Eynaud, Nisbet, and Muller
+(2011), which included representations of photoinhibition or photodamage
+separately. It is important to recognize that the upper light limit set
+by photooxidative stress on a stable symbiosis under steady state
+conditions (Fig. 3) may be temporarily crossed by a dynamic system,
+which may experience a period of symbiont loss (bleaching) and reduced
+growth, after which a return to benign conditions may restore symbiont
+biomass and positive growth. To explore this further and illustrate the
+behavior of the model in more detail, we evaluate a number of dynamic
+simulations below (see ``Dynamic behavior'').
 
@@
 to changes in these parameter values, which also serves to demonstrate
-the behavior of the dynamical system. We measured fractional change in
+the behavior of the dynamical system. We calculated fractional change in
 steady state values in response to fractional changes in parameter
@@
 Overall, relative changes in the steady state of the system are less
-than the equivalent relative change in parameter value, indicating that
-the system maintains internal stability. However, changes in certain
-parameter values have more significant impacts: increasing \(j_{Nm}\) or
-decreasing \(K_N\) both dramatically increase host growth (Fig. 4),
-demonstrating the strong nitrogen-limitation that characterizes these
-symbioses. The parameter \(\bar{a}^*\) has a strong impact on S:H
-biomass ratios (Fig. 4) since this parameter determines the amount of
-light absorbed by symbionts, with lower values increasing
+than the equivalent relative change in parameter value. However, changes
+in certain parameter values have more significant impacts: increasing
+\(j_{Nm}\) or decreasing \(K_N\) both dramatically increase host growth
+(Fig. 4), demonstrating the strong nitrogen-limitation that
+characterizes these symbioses. The parameter \(\bar{a}^*\) has a strong
+impact on S:H biomass ratios (Fig. 4) since this parameter determines
+the amount of light absorbed by symbionts, with lower values increasing
 light-limitation. Increasing the maximum growth and turnover rates have
@@
 study as inputs, the model reproduces this observed interaction among
-environmental factors (Fig. 5), and also provides the mechanism: under
-high light in summer, photooxidative stress decreases symbiont biomass
-and increases carbon-limitation of host growth regardless of nutrient
-status; when light is reduced in winter, growth and symbiont biomass
-become constrained instead by nitrogen, and therefore increase more when
-DIN is enriched (Fig. S8).
+environmental factors (Fig. 5), and also provides the mechanism: as
+increasing light in summer decreases carbon fixation due to
+photooxidative stress, symbionts grow slower than the host, leading to
+decreasing S:H ratios. Under nutrient enrichment, this effect is more
+pronounced, as CO\textsubscript{2}-limitation (due to higher symbiont
+standing stocks) causes mild bleaching that results in a similar
+summertime minimum S:H as the ambient nutrient case (`physiological
+bleaching' \emph{sensu} Fitt et al. (2001)). Decreasing light into
+winter then alleviates the photooxidative stress constraints on carbon
+fixation, such that nitrogen constrains the S:H ratio, explaining why
+S:H increases more when DIN is enriched (Fig. S8).
 
@@
 panel) show seasonal oscillations that are greater in magnitude under
-higher nutrients (1e-5 molN L\textsuperscript{-1}; dashed lines)
-relative to lower nutrients (1e-7 molN L\textsuperscript{-1}; solid
+higher nutrients (1e-5 mol N L\textsuperscript{-1}; dashed lines)
+relative to lower nutrients (1e-7 mol N L\textsuperscript{-1}; solid
 lines), consistent with the findings of Stimson (1997). Prey density was
@@
 growth is not limited by low light in winter, but is actually reduced by
-excess light in summer, consistent with the experimental findings of
-Schutter et al. (2011). Seasonal summertime reductions in tissue biomass
-have also been well-documented in the field (Fitt et al. 2000), along
-with reductions in net photosynthetic capacity (Muller-Parker 1987).
-Importantly, while light alone (at these levels) may drive seasonal
-dynamics in the ways discussed, temperature fluctuations may attenuate
-or even reverse the effect of light as cooler winters depress
+excess light in summer\footnote{at the light levels indicated in Stimson
+  (1997). Note that if light levels were reduced throughout the year
+  (e.g.~for a coral at greater depth) such that light did not cause
+  photo-oxidative stress in summer but became limiting to photosynthesis
+  in winter, the S:H ratio would still increase in winter, but growth
+  would decrease; the latter scenario is predicted both by the present
+  model as well as that of Muller et al. (2009), since photo-oxidative
+  stress does not become relevant.}, consistent with the experimental
+findings of Schutter et al. (2011). Seasonal summertime reductions in
+tissue biomass have also been well-documented in the field (Fitt et al.
+2000), along with reductions in net photosynthetic capacity
+(Muller-Parker 1987). Importantly, while light alone may drive seasonal
+dynamics in the ways discussed\footnote{at the light levels indicated in
+  Stimson (1997). Note that if light levels were reduced throughout the
+  year (e.g.~for a coral at greater depth) such that light did not cause
+  photo-oxidative stress in summer but became limiting to photosynthesis
+  in winter, the S:H ratio would still increase in winter, but growth
+  would decrease; the latter scenario is predicted both by the present
+  model as well as that of Muller et al. (2009), since photo-oxidative
+  stress does not become relevant.}, temperature fluctuations may
+attenuate or even reverse the effect of light as cooler winters depress
 metabolism; thus, the relative magnitude of fluctuation in temperature,
@@
 switch occurs when increasing external irradiance (Fig. 6A) causes
-sufficient ROS production (Fig. 6B) and photoinhibition (Fig. 6C) such
-that the positive feedbacks between host carbon-limitation (Fig. 6D) and
+sufficient ROS production (Fig. 6B) and photoinhibition (Fig. 6C) that
+the positive feedbacks between host carbon-limitation (Fig. 6D) and
 CO\textsubscript{2}-limitation of photosynthesis (Fig. 6E) are rapidly
@@
 limitation coefficients were calculated using Eq. 22. All parameters
-were set at default values with external DIN=1e-7 molN/L and prey
+were set at default values with external DIN=1e-7 mol N/L and prey
 density=0).}
@@
 d\textsuperscript{-1}) under default environmental conditions (solid
-line; DIN=1e-7 molN L\textsuperscript{-1}; prey=3e-6 CmolX
+line; DIN=1e-7 mol N L\textsuperscript{-1}; prey=3e-6 CmolX
 L\textsuperscript{-1}), or with elevated feeding (dashed line; prey=6e-6
 CmolX L\textsuperscript{-1}), or elevated nutrients (dotted line;
-DIN=1e-6 molN L\textsuperscript{-1}). Initial symbiont biomass was set
+DIN=1e-6 mol N L\textsuperscript{-1}). Initial symbiont biomass was set
 to the steady state for each set of starting conditions, respectively.
@@
 
-Whether this `overshoot' occurs or not represents a bifurcation point in
-the dynamical system that is defined by the relative availability of
-carbon and nitrogen to the host -- any factor that enhances
-carbon-limitation of host growth (e.g.~high DIN and/or low feeding)
-therefore magnifies the overshoot and prolongs the dysfunctional,
-carbon-limited state of the symbiosis (Fig. 8). On the other hand,
-factors that favor nitrogen-limitation, such as low external DIN and/or
-high feeding rates, will accelerate the re-establishment of
+Whether this `overshoot' occurs or not is determined by the relative
+availability of carbon and nitrogen to the host -- any factor that
+enhances carbon-limitation of host growth (e.g.~high DIN and/or low
+feeding) therefore magnifies the overshoot and prolongs the
+dysfunctional, carbon-limited state of the symbiosis (Fig. 8). On the
+other hand, factors that favor nitrogen-limitation, such as low external
+DIN and/or high feeding rates, will accelerate the re-establishment of
 nitrogen-limitation and prevent an overshoot from occurring at all.
@@
 parameters (except varying \(n_{NX}\)), with L=10 mol photons
-m\textsuperscript{2} d\textsuperscript{-1}, DIN=1e-7 molN
+m\textsuperscript{2} d\textsuperscript{-1}, DIN=1e-7 mol N
 L\textsuperscript{-1}, prey=1e-6 CmolX L\textsuperscript{-1}, and
@@
 alone does not necessarily indicate that a `healthy' state has been
-reached, since the C-limited state may still persist (e.g., Fig. 8B,
-8C). This could explain why corals that have recovered their symbiont
+reached, since the C-limited state may still persist (e.g.~Fig. 8B, 8C).
+This could explain why corals that have recovered their symbiont
 populations after bleaching may still exhibit energetic deficits and
-physiological impacts for much longer periods of time (Levitan et al.
-2014, Hughes and Grottoli (2013)). Therefore, host acquisition of carbon
-from a source other than the symbiont may be extremely important for the
-system to recover from bleaching. Indeed, host feeding has been shown to
-promote a more rapid return to pre-bleaching levels of key physiological
-parameters in recovering corals (Connolly, Lopez-Yglesias, and Anthony
-2012). The model predicts that acquisition of carbon from other sources,
-such as direct uptake of DOC (Levas et al. 2015), may also promote more
-rapid recovery from a bleached state.
+physiological impacts, possibly for months to years (Levitan et al.
+2014, Hughes and Grottoli (2013))). Therefore, host acquisition of
+carbon from a source other than the symbiont may be extremely important
+for the system to recover from bleaching. Indeed, host feeding has been
+shown to promote a more rapid return to pre-bleaching levels of key
+physiological parameters in recovering corals (Connolly, Lopez-Yglesias,
+and Anthony 2012). The model predicts that acquisition of carbon from
+other sources, such as direct uptake of DOC (Levas et al. 2015), may
+also promote more rapid recovery from a bleached state.
 
@@
 
-This simplified dynamic bioenergetic model of coral-\emph{Symbiodinium}
-symbioses mechanistically reproduces patterns in steady-state coral
-growth and symbiont abundance commonly observed in corals, including
-higher symbiont abundance with higher nutrients and feeding, lower
-symbiont abundance with increasing light, and optimal growth at
-intermediate light levels. Moreover, the model reproduces complex
-dynamic behaviors including seasonal changes in symbiont density at
-different nutrient levels, rapid bleaching above a threshold of high
-light, mitigation of bleaching by heterotrophic feeding, exacerbation of
-bleaching by elevated nutrients, and an overshoot of symbiont density
-during recovery from bleaching. These examples demonstrate the model's
-ability to integrate multiple environmental forcing functions to
-reproduce complex responses; meanwhile, the diversity of these phenomena
-suggest the model has captured many of the important features of the
-system in a unifying mechanistic framework. This model also provides a
-new conceptual framework for considering coral bleaching as a transition
-to an alternate stable state, which has important implications for
+This dynamic bioenergetic model of coral-\emph{Symbiodinium} symbioses
+mechanistically reproduces patterns in steady-state coral growth and
+symbiont abundance commonly observed in corals, including higher
+symbiont abundance with higher nutrients and feeding, lower symbiont
+abundance with increasing light, and optimal growth at intermediate
+light levels. Moreover, the model reproduces complex dynamic behaviors
+including seasonal changes in symbiont density at different nutrient
+levels, rapid bleaching above a threshold of high light, mitigation of
+bleaching by heterotrophic feeding, exacerbation of bleaching by
+elevated nutrients, and an overshoot of symbiont density during recovery
+from bleaching. These examples demonstrate the model's ability to
+integrate multiple environmental forcing functions to reproduce complex
+responses; meanwhile, the diversity of these phenomena suggest the model
+has captured many of the important features of the system in a unifying
+mechanistic framework. This model also provides a new conceptual
+framework for considering coral bleaching as a transition to an
+alternate stable state, which has important implications for
 understanding the performance and maintenance of symbiotic interactions.
@@
 
+\hypertarget{ref-Fitt:2001p832}{}
+Fitt, William K, Barbara E Brown, Mark E Warner, and R P Dunne. 2001.
+``Coral bleaching: interpretation of thermal tolerance limits and
+thermal thresholds in tropical corals.'' \emph{Coral Reefs} 20: 51--65.
+
 \hypertarget{ref-Fitt:2000p3792}{}

### 2017-02-07 JTB_supp/JTB_supp.Rmd
fix accidental deletion of coefs in Fig S1 code

@@
 # Get parameters of fit and add to plot
+coefs <- round(coef(mod), 2)
 fiteq <- str_replace_all(format(formula(mod)), c("~"="= ", " x"=coefs[[1]], "y"=coefs[[2]], "z"=coefs[[3]]))

### 2017-02-07 JTB_ms/JTB_ms.Rmd
add Fig. 2 legend

@@
 
-![Alternate stable states](../img/Fig2.png)
+![Alternate stable states in S:H biomass and growth across a light gradient.
+Alternate stable states occur between ~33-44 mol photons m^-2^ d^-1^ under these conditions (DIN=1e-7 molN L^-1^; X=2e-6 molX L^-1^), depending on whether initial S:H is high (0.1, closed circles), representing a healthy coral, or low (0.0001, open circles), representing a bleached coral.](../img/Fig2.png)
 

### 2017-02-07 JTB_supp/JTB_supp.Rmd
combine FigS9 run and plot

@@
 # Figure S9
-```{r runFigS9, cache=T, include=F, eval=T}
+```{r FigS9, cache=T, fig.width=7, fig.height=7, echo=F, message=F, eval=T, fig.cap="**Fig.
+S9.
+Hysteresis under different nutrient and feeding regimes.** This figure accompanies Fig.
+2 in the main text.
+These figures show steady state S:H biomass ratios in constant environments with varying DIN and food (X) levels, with initial S:H biomass values of either 0.1 or 0.0001.
+Points are colored according to whether the host exhibits positive (black) or negative (red) growth at steady state."}
+
 # Set up cluster for parallel processing
@@
 res4 <- cbind(input, output4)
-```
 
-```{r plotFigS9, cache=T, fig.width=7, fig.height=7, echo=F, message=F, eval=T, fig.cap="**Fig.
-S9.
-Hysteresis under different nutrient and feeding regimes.** This figure accompanies Fig.
-2 in the main text.
-These figures show steady state S:H biomass ratios in constant environments with varying DIN and food (X) levels, with initial S:H biomass values of either 0.1 or 0.0001.
-Points are colored according to whether the host exhibits positive (black) or negative (red) growth at steady state."}
 # Create plot

### 2017-02-08 JTB_supp/test_figs9.Rmd
test figs9

--- 2017-02-07 JTB_supp/JTB_supp.Rmd
@@
 ---
-title: "Supplementary Information for:"
-subtitle: A simplified dynamic bioenergetic model for coral-*Symbiodinium* symbioses and coral bleaching as an alternate stable state
+title: "TestFigS9"
 author: "Ross Cunning, Erik B. Muller, Ruth D. Gates, Roger M. Nisbet"
@@
 ---
-
------
-
-This supplement provides justification and derivations from the literature for all parameter values used in the model, as well as additional figures that demonstrate the structure and behavior of the model and its specific components.
 
@@
 
------
-
-# Model parameters
-All parameters in the model are listed here along with the relevant literature that provides a basis for the value used.
-In most cases, the units of the values reported in the literature were converted to units appropriate for the model, which was accomplished using a consistent set of stoichiometric conversion factors also based on reported literature (see [Conversion factors used in parameter derivation]).
-Calculations executed in R are shown for each parameter (when relevant) as well as the final value used.
-
-## Parameter value derivations
-
-#### $n_{NH}$: N:C molar ratio in host biomass {-}
-@Horwitz:2015cf measured a molar C:N ratio of 5.53 in anemones under control conditions, which converts to:
-```{r nNH}
-1 / 5.53
-```
-N:C ratio (dimensionless)
-<br><br>
-
-#### $n_{NS}$: N:C molar ratio in symbiont biomass {-}
-This value should approximate the Redfield ratio (following @Muller:2009io).  
-```{r nNS, echo=F}
-0.13
-```
-N:C ratio (dimensionless)
-<br><br>
-
-#### $n_{NX}$: N:C molar ratio in prey biomass {-}
-Assuming the coral feeds primarily on zooplankton, a general heterotroph has a molar N:C ratio of 0.2 [@Muller:2009io].  
-```{r nNX, echo=F}
-0.20
-```
-N:C ratio (dimensionless)
-<br><br>
-
-#### $j_{HT}^0$: Specific turnover rate of host biomass {-}
-@Jacobson:2016bk parameterized a DEB model for coral symbiosis, and reported a specific maintenance rate of 0.03 CmolH CmolH^-1^ d^-1^.
-Although this value is per unit structural biomass (in the DEB sense) rather than per unit total biomass (as needed here), use of this value generates plausible values of specific growth (see main text).
-```{r jHT, echo=F}
-0.03
-```
-CmolH CmolH^-1^ d^-1^
-<br><br>
-
-#### $j_{ST}^0$: Specific turnover rate of symbiont biomass {-}
-Assuming an equivalent maintenance rate as the host:
-```{r jST, echo=F}
-0.03
-```
-CmolS CmolS^-1^ d^-1^
-<br><br>
-
-#### $\sigma_{NH}, \sigma_{NS}$: Proportions of nitrogen recycled from host and symbiont biomass turnover {-}
-Assume highly efficient recycling by both symbiont and host to avoid loss to the environment:  
-```{r sigmaNHsigmaNS, echo=F}
-c(0.03, 0.03)
-```
-(dimensionless)
-<br><br>
-
-#### $\sigma_{CS}$: Proportion of metabolic CO~2~ produced by symbiont recycled passively to photosynthesis {-}
-Assume that CO~2~ produced within symbiont cells is recycled with high efficiency, potentially aided by symbiont carbon concentrating mechanisms.  
-```{r sigmaCS, echo=F}
-0.9
-```
-(dimensionless)
-<br><br>
-
-#### $\sigma_{CH}$: Proportion of metabolic CO~2~ produced by host recycled passively to photosynthesis {-}
-Assume that most metabolic CO~2~ produced within host cells rapidly converts to HCO~3~^-^ at intracellular pH ~ 7.2 [@Venn:2009hy], and that only ~10% actually remains as CO~2~ [@Zeebe:2001wz] that could diffusively supply photosynthesis.
-This passive supply is differentiated from all active host carbon concentrating mechanisms (CCMs) that facilitate CO~2~ supply to photosynthesis, which are separately treated collectively as the $jCO~2~$ flux.
-```{r sigmaCH, echo=F}
-0.10
-```
-(dimensionless)
-<br><br>
-
-#### $j_{Xm}$: Maximum specific feeding rate of host {-}
-@Wijgerde:2011du measured depletion of carbon from captured prey in *Galaxea fascicularis* (including both intracoelenteric and extracoelenteric feeding), reporting a maximum specific feeding rate of 77.6 µgC cm^-2^ 6h^-1^.
-```{r jXm}
-77.6 / µg.g / gC.molC * cm2.m2 / CmolR.m2_Gfas * 24/6
-```
-molX CmolH^-1^ d^-1^
-<br><br>
-
-#### $K_X$: Half-saturation constant for prey uptake by host {-}
-@Purser:2010hp measured feeding rates of the coral *Lophelia pertusa* at different concentrations of *Artemia* nauplii, reporting a half-saturation for feeding of ~345 *Artemia* L^-1^.
-```{r KX}
-345 * gC.artemia / gC.molC
-```
-CmolX L^-1^
-<br><br>
-
-#### $j_{Nm}$: Maximum specific DIN uptake rate by host {-}
-@Muscatine:1978fa measured ammonium uptake in 3 coral species and reported an average value of 7.22 µmol NH~4~^+^-N mg chl *a*^-1^ h^-1^ (Table 1, uncorrected), which converts to:
-```{r jNm1}
-7.22 * h.d * nmolchlA.cm2 * gchlA.molchlA * mg.g / nmol.mol / CmolR.m2_Spis * cm2.m2 / µmol.mol
-```
-mol NH~4~^+^-N  CmolH^-1^ d^-1^  
-<br>
-@HoeghGuldberg:1999p3974 measured ammonium and dissolved free amino acid uptake in *Pocillopora damicornis* and reported a maximum rate of 91.8 nmol N cm^-2^ h^-1^, which converts to:
-```{r jNm2}
-91.8 * h.d * cm2.m2 / CmolR.m2_Spis / nmol.mol
-``` 
-molN CmolH^-1^ d^-1^  
-<br>
-@Godinot:2011jg measured nitrate and ammonium uptake rates in *Stylophora pistillata* under high nutrient concentrations and reported a maximum value of 3.5 µmolN mg chl^-1^ d^-1^ [addition of bars in Fig.
-4D,E @ 29°C], which converts to:
-```{r jNm3}
-3.5 * h.d * 22 / CmolR.m2_Spis * cm2.m2 / µmol.mol / 1000
-```
-molN CmolH^-1^ d^-1^   
-<br>
-@Godinot:2011bm measured ammonium uptake kinetics in *Stylophora pistillata* and reported a maximum rate of 5.42 molN mg chl^-1^ h^-1^, which converts to:
-```{r jNm4}
-5.42 * h.d * µgchl.cm2_Spis / CmolR.m2_Spis * cm2.m2 / µmol.mol / 1000
-```
-molN CmolH^-1^ d^-1^   
-<br>
-Averaging these reported values, we arrive at:  
-```{r jNm_avg, echo=F}
-round(mean(c(7.22 * h.d * nmolchlA.cm2 * gchlA.molchlA * mg.g / nmol.mol / CmolR.m2_Spis * cm2.m2 / µmol.mol, 91.8 * h.d * cm2.m2 / CmolR.m2_Spis / nmol.mol, 3.5 * h.d * 22 / CmolR.m2_Spis * cm2.m2 / µmol.mol / 1000, 5.42 * h.d * µgchl.cm2_Spis / CmolR.m2_Spis * cm2.m2 / µmol.mol / 1000)), 3)
-```
-molN CmolH^-1^ d^-1^
-<br><br>
-
-#### $K_N$: Half-saturation constant for DIN uptake by host {-}
-@Muscatine:1978fa measured ammonium uptake in 3 coral species and reported an average half-saturation constant of 1.84 µmolN L^-1^ (Table 1, uncorrected).  
-@Godinot:2011bm measured ammoinum uptake kinetics in *Stylophora pistillata* and reported a half-saturation constant of 0.46 µmol N L^-1^.  
-Averaging these reported values (weighted by number of species measured), we arrive at:  
-```{r KN}
-(1.84*3 + 0.46*1)/4
-``` 
-molN L^-1^
-<br><br>
-
-#### $k_{CO_2}$: Efficacy of CO~2~ delivery by host CCMs {-}
-This parameter represents the relative amount of CO~2~ that can be delivered by the host to photosynthesis given the reducing power generated from 1 mol of fixed carbon.
-We assume that the host uses a portion of its fixed carbon to generate ATP which is then used to activate a variety of carbon concentrating mechanisms (see @Wooldridge:2013tj).
-We assume these CCMs are efficient and can deliver an order of magnitude more CO~2~ relative to the fixed carbon available.
-```{r kCO2, echo=F}
-10
-```
-molCO~2~ molC^-1^
-<br><br>
-
-#### $j_{HGm}$: Maximum specific growth rate of host {-}
-We set this to a high value such that constraints on growth are primarily determined by substrate availabilities rather than the maximum value.  
-```{r jHGm, echo=F}
-1
-```
-CmolH CmolH^-1^ d^-1^
-<br><br>
-
-#### $y_{CL}$: Quantum yield of photosynthesis {-}
-The maximum theoretical quantum yield is 0.125 mol C fixed per mol photons.
-@Brodersen:2014gw presented a balanced light budget based on microenvironmental photobiology of *Montastrea curta*, and reported a quantum yield of 0.102 O~2~ photon^-1^.
-```{r yCL, echo=F}
-0.1
-```
-molC molph^-1^
-<br><br>
-
-#### $\bar{a}^*$: Effective light-absorbing cross-section of symbiont {-}
-@Hennige:2009p3439 measured spectra-weighted absoprtion coefficients from 400 to 700nm *ā\**~(400-700)~ for eight *Symbiodinium* types in culture at two different light levels, and reported a (grand mean) value of 0.01095 m^2^ mg chl *a*^-1^ [from Table 6].
-This converts to:
-```{r astar}
-0.01095 * mg.g * gchlA.gC * gC.molC
-```
-m^2^ CmolS^-1^
-<br><br>
-
-#### $k_{NPQ}$: Capacity of non-photochemical quenching {-}
-@Gorbunov:2001p4181 report that the capacity of non-photochemical quenching is approximately 4 times the capacity of photochemistry on a sunny day.
-Therefore, we take the maximum relative capacity of NPQ to be 4 times the maximum rate of photochemical quenching ($j_{CPm}$; see below), giving a value of:
-```{r kNPQ}
-defpars$jCPm / defpars$yCL * 4
-```
-mol ph CmolS^-1^ d^-1^
-<br><br>
-
-#### $k_{ROS}$: Amount of excitation energy not quenched by photoehcemistry or NPQ that doubles relative production of ROS {-}
-@McGinty:2012ck report a ~4-fold increase in ROS production in *Symbiodinium* C1 under stressful conditions relative to control.
-We chose a value of $k_{ROS}$ that similarly produced a ~4-fold increase in ROS production at the highest light level (see Figure S2):
-```{r kROS, echo=F}
-80
-```
-mol ph CmolS^-1^ d^-1^
-<br><br>
-
-#### $j_{CPm}$: Maximum specific photosynthesis rate of symbiont {-}
-In theory, this value should be the maximum carboxylation rate of *Symbiodinium* Rubisco (RuBP).
-@Leggat:2004vx report a maximum carboxylation rate at 29°C of 140 CO~2~ RuBP^-1^ s^-1^ [Max.
-value, p.
-143], which converts to:
-```{r jCPm}
-140 * s.h * h.d * RuBP.CmolS
-```
-molC CmolS^-1^ d^-1^
-<br><br>
-
-#### $j_{SGm}$: Maximum specific growth rate of symbiont {-}
-@Suggett:2015dv measured specific growth rates of 5 clades of *Symbiodinium* in culture, and reported values ranging from 0.19 to 0.26 d^-1^ (Table 2).  
-```{r jSGm, echo=F}
-0.25
-```
-CmolS CmolS^-1^ d^-1^
-<br><br>
-
-#### $b$: Scaling parameter for strength of bleaching response {-}
-This value determines the magnitude of symbiont biomass loss (considered as expulsion of symbionts during bleaching) in response to ROS.
-In this context, relative ROS concentrations are assumed to trigger symbiont expulsion via unknown signaling mechanisms.
-Larger values of this parameter cause greater biomass loss in response to ROS signaling (i.e., a more severe bleaching response).
-Since this represents symbiont cells being expelled from the system, we use a value about an order of magnitude higher than the turnover rate due to maintenance.
-```{r b, echo=F}
-5
-```
-(dimensionless)
-<br><br>
-
-## Conversion factors used in parameter derivation
-```{r, code = readLines('../R/conversions.R')}
-```
-
-------
-
-# Figure S1 
-```{r lightamp, fig.width=5, fig.height=5, eval=T, include=T, echo=F, fig.cap="**Fig.
-S1.
-Light amplification and self-shading.** The amplification factor (A) is the ratio of internal scalar irradiance experienced by *Symbiodinium* to the external downwelling irradiance, and is greater than 1 due to multiple scattering of incident light by the coral skeleton.
-The amplification factor decreases as a function of symbiont abundance (converted to symbiont to host biomass ratio) due to self-shading.
-Data are from Marcelino et al.
-(2013) using skeletons from three different coral species: pdam=*Pocillopora damicornis*; plob=*Porites lobata*; shys=*Seriatopora hystrix*."}
-
-# Import data from Marcelino et al. 2013
-md <- read.csv("data/marcelino2013_fig2a.csv")
-md$S.H <- md$density * cm2.m2 / CmolR.m2 * CmolS.cellS  # Convert units of symbiont density to CmolS/CmolR
-md$A <- md$E + 1  # Convert "E"=excess light metric of Marcelino to "A"=amplification factor
-# Plot data
-par(mfrow=c(1,1), mar=c(3,3,3,3), mgp=c(1.2,0,0), cex=1, tck=0.025)
-plot(md$A ~ md$S.H, pch=c(21,23,22)[factor(md$spp)], bg=c("white", "black", "gray")[factor(md$spp)], 
-     ylab="A (Amplification factor)", ylim=c(1,3.2),
-     xlab="S:H biomass")
-legend("topright", legend=levels(md$spp), pch=c(21,23,22), pt.bg=c("white", "black", "gray"), inset=0.02)
-# Fit exponential decay model and plot fitted values
-mod <- nls(A ~ x + y * exp(z * S.H), data=md, start=list(x=1, y=2, z=-3))
-fit <- predict(mod, list(S.H=seq(0, 0.5, 0.01)))
-lines(fit ~ seq(0, 0.5, 0.01))
-# Get parameters of fit and add to plot
-coefs <- round(coef(mod), 2)
-fiteq <- str_replace_all(format(formula(mod)), c("~"="= ", " x"=coefs[[1]], "y"=coefs[[2]], "z"=coefs[[3]]))
-text(0.05,2.8, fiteq, adj=0)
-
-```
-
------
-
-# Figure S2 
-```{r run1_PI, cache=F, fig.height=4, fig.width=7, echo=F, fig.cap="**Fig.
-S2.
-Photosynthesis-Irradiance curves of modeled symbiont.** The rates of light energy quenching through photochemistry (red line) and non-photochemical quenching (black line), and the relative production of reactive oxygen species (orange line) are plotted as a function of the scalar irradiance experienced by the symbiont (note different y-axes for each line).
-These curves represent the behavior of the photosynthesis SU independent of the host and without CO~2~ limitation of photosynthesis (*j~CO2~* set to a very high value).
-The structure of the photosynthesis SU, with the negative feedback between excess light and photosynthesis rate, produces the classic shape of a P-I curve demonstrating photoinhibition, suggesting the mechanism is robust."}
-sym_PI(pars=def_pars())
-```
-
------
-
-# Figure S3 
-```{r sens_ILHN, cache=T, echo=F, eval=T, fig.cap="**Fig. S3. Sensitivity analysis under high DIN (4e-6 M) and intermediate light (15 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
-# Set up time vector and parameters
-defpars <- def_pars()  # Get default parameters
-defpars <- replace(defpars, "initS", 1)  # Start with high S biomass to avoid alternate steady state of negative growth
-
-# Define environments for conducting sensitivity analysis
-envs <- list(
-  ILHN=list(L=15, N=4e-6, X=0)
-)
-
-# Define relative changes in each parameter to evaluate
-levs <- seq(0.5, 1.5, 0.05)  # Set levels of % change in parameter value to evaluate
-
-# Define all simulations to be run for a given parameter
-sims <- expand.grid(env=names(envs), change=levs)
-n <- nrow(sims)
-
-# Set up cluster for parallel processing
-cl <- makeCluster(detectCores())  # Initiate cluster
-registerDoParallel(cl)
-
-# Run sensitivity analysis in each environment for each parameter
-jHT0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHT0", change=change))
-jNm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jNm", change=change))
-jHGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHGm", change=change))
-kCO2.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kCO2", change=change))
-KN.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="KN", change=change))
-jST0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jST0", change=change))
-kNPQ.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kNPQ", change=change))
-kROS.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kROS", change=change))
-astar.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="astar", change=change))
-jCPm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jCPm", change=change))
-jSGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jSGm", change=change))
-b.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="b", change=change))
-
-stopCluster(cl)  # Stop cluster
-
-# Combine output with input
-sens.all <- ls(pattern="*.sens")
-for (i in 1:length(sens.all)) {
-  assign(sens.all[i], cbind(sims, get(sens.all[i])))
-}
-
-# When growth rate becomes negative, set all responses to NA
-for (i in 1:length(sens.all)) {
-  if (any(get(sens.all[i])$grchange < 0, na.rm=T)) {
-    assign(sens.all[i], within(get(sens.all[i]), {
-      shchange <- replace(shchange, grchange < 0, NA)
-      cROSchange <- replace(cROSchange, grchange < 0, NA)
-      grchange <- replace(grchange, grchange < 0, NA)
-    }))
-  }
-}
-
-# Plot Figure
-par(mfrow=c(2,2), mar=c(3,3,1,1), mgp=c(2,0.5,0))
-sens.plot(response=c("grchange", "shchange"), pars=c("jCPm", "kCO2", "astar"), cols=c("black", "red", "gold"))
-sens.plot(response=c("grchange", "shchange"), pars=c("jNm", "KN"), cols=c("black", "blue"))
-sens.plot(response=c("grchange", "shchange"), pars=c("jST0", "jHT0", "jSGm", "jHGm"), cols=c("burlywood4", "burlywood3", "darkolivegreen4", "darkolivegreen3"))
-sens.plot(response=c("grchange", "shchange"), pars=c("kNPQ", "kROS", "b"), cols=c("black", "orange", "purple"))
-```
-
------
-
-# Figure S4 
-```{r sens_LLLN, cache=T, echo=F, eval=T, fig.cap="**Sensitivity analysis under low DIN (1e-7 M) and low light (2 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
-# Set up time vector and parameters
-defpars <- def_pars()  # Get default parameters
-defpars <- replace(defpars, "initS", 1)  # Start with high S biomass to avoid alternate steady state of negative growth
-
-# Define environments for conducting sensitivity analysis
-envs <- list(
-  LLLN=list(L=2, N=1e-7, X=0)
-)
-
-# Define relative changes in each parameter to evaluate
-levs <- seq(0.5, 1.5, 0.05)  # Set levels of % change in parameter value to evaluate
-
-# Define all simulations to be run for a given parameter
-sims <- expand.grid(env=names(envs), change=levs)
-n <- nrow(sims)
-
-# Set up cluster for parallel processing
-cl <- makeCluster(detectCores())  # Initiate cluster
-registerDoParallel(cl)
-
-# Run sensitivity analysis in each environment for each parameter
-jHT0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHT0", change=change))
-jNm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jNm", change=change))
-jHGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHGm", change=change))
-kCO2.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kCO2", change=change))
-KN.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="KN", change=change))
-jST0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jST0", change=change))
-kNPQ.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kNPQ", change=change))
-kROS.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kROS", change=change))
-astar.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="astar", change=change))
-jCPm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jCPm", change=change))
-jSGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jSGm", change=change))
-b.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="b", change=change))
-
-stopCluster(cl)  # Stop cluster
-
-# Combine output with input
-sens.all <- ls(pattern="*.sens")
-for (i in 1:length(sens.all)) {
-  assign(sens.all[i], cbind(sims, get(sens.all[i])))
-}
-
-# When growth rate becomes negative, set all responses to NA
-for (i in 1:length(sens.all)) {
-  if (any(get(sens.all[i])$grchange < 0, na.rm=T)) {
-    assign(sens.all[i], within(get(sens.all[i]), {
-      shchange <- replace(shchange, grchange < 0, NA)
-      cROSchange <- replace(cROSchange, grchange < 0, NA)
-      grchange <- replace(grchange, grchange < 0, NA)
-    }))
-  }
-}
-
-# Plot Figure
-par(mfrow=c(2,2), mar=c(3,3,1,1), mgp=c(2,0.5,0))
-sens.plot(response=c("grchange", "shchange"), pars=c("jCPm", "kCO2", "astar"), cols=c("black", "red", "gold"))
-sens.plot(response=c("grchange", "shchange"), pars=c("jNm", "KN"), cols=c("black", "blue"))
-sens.plot(response=c("grchange", "shchange"), pars=c("jST0", "jHT0", "jSGm", "jHGm"), cols=c("burlywood4", "burlywood3", "darkolivegreen4", "darkolivegreen3"))
-sens.plot(response=c("grchange", "shchange"), pars=c("kNPQ", "kROS", "b"), cols=c("black", "orange", "purple"))
-```
-
------
-
-# Figure S5
-```{r sens_LLHN, cache=T, echo=F, eval=T, fig.cap="**Fig. S5. Sensitivity analysis under high DIN (4e-6 M) and low light (2 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
-# Set up time vector and parameters
-defpars <- def_pars()  # Get default parameters
-defpars <- replace(defpars, "initS", 1)  # Start with high S biomass to avoid alternate steady state of negative growth
-
-# Define environments for conducting sensitivity analysis
-envs <- list(
-  LLHN=list(L=2, N=4e-6, X=0)
-)
-
-# Define relative changes in each parameter to evaluate
-levs <- seq(0.5, 1.5, 0.05)  # Set levels of % change in parameter value to evaluate
-
-# Define all simulations to be run for a given parameter
-sims <- expand.grid(env=names(envs), change=levs)
-n <- nrow(sims)
-
-# Set up cluster for parallel processing
-cl <- makeCluster(detectCores())  # Initiate cluster
-registerDoParallel(cl)
-
-# Run sensitivity analysis in each environment for each parameter
-jHT0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHT0", change=change))
-jNm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jNm", change=change))
-jHGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHGm", change=change))
-kCO2.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kCO2", change=change))
-KN.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="KN", change=change))
-jST0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jST0", change=change))
-kNPQ.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kNPQ", change=change))
-kROS.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kROS", change=change))
-astar.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="astar", change=change))
-jCPm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jCPm", change=change))
-jSGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jSGm", change=change))
-b.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="b", change=change))
-
-stopCluster(cl)  # Stop cluster
-
-# Combine output with input
-sens.all <- ls(pattern="*.sens")
-for (i in 1:length(sens.all)) {
-  assign(sens.all[i], cbind(sims, get(sens.all[i])))
-}
-
-# When growth rate becomes negative, set all responses to NA
-for (i in 1:length(sens.all)) {
-  if (any(get(sens.all[i])$grchange < 0, na.rm=T)) {
-    assign(sens.all[i], within(get(sens.all[i]), {
-      shchange <- replace(shchange, grchange < 0, NA)
-      cROSchange <- replace(cROSchange, grchange < 0, NA)
-      grchange <- replace(grchange, grchange < 0, NA)
-    }))
-  }
-}
-
-# Plot Figure
-par(mfrow=c(2,2), mar=c(3,3,1,1), mgp=c(2,0.5,0))
-sens.plot(response=c("grchange", "shchange"), pars=c("jCPm", "kCO2", "astar"), cols=c("black", "red", "gold"))
-sens.plot(response=c("grchange", "shchange"), pars=c("jNm", "KN"), cols=c("black", "blue"))
-sens.plot(response=c("grchange", "shchange"), pars=c("jST0", "jHT0", "jSGm", "jHGm"), cols=c("burlywood4", "burlywood3", "darkolivegreen4", "darkolivegreen3"))
-sens.plot(response=c("grchange", "shchange"), pars=c("kNPQ", "kROS", "b"), cols=c("black", "orange", "purple"))
-```
-
------
-
-# Figure S6
-```{r sens_HLLN, cache=T, echo=F, eval=T, fig.cap="**Fig. S6. Sensitivity analysis under low DIN (1e-7 M) and high light (25 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
-# Set up time vector and parameters
-defpars <- def_pars()  # Get default parameters
-defpars <- replace(defpars, "initS", 1)  # Start with high S biomass to avoid alternate steady state of negative growth
-
-# Define environments for conducting sensitivity analysis
-envs <- list(
-  HLLN=list(L=25, N=1e-7, X=0)
-)
-
-# Define relative changes in each parameter to evaluate
-levs <- seq(0.5, 1.5, 0.05)  # Set levels of % change in parameter value to evaluate
-
-# Define all simulations to be run for a given parameter
-sims <- expand.grid(env=names(envs), change=levs)
-n <- nrow(sims)
-
-# Set up cluster for parallel processing
-cl <- makeCluster(detectCores())  # Initiate cluster
-registerDoParallel(cl)
-
-# Run sensitivity analysis in each environment for each parameter
-jHT0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHT0", change=change))
-jNm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jNm", change=change))
-jHGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHGm", change=change))
-kCO2.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kCO2", change=change))
-KN.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="KN", change=change))
-jST0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jST0", change=change))
-kNPQ.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kNPQ", change=change))
-kROS.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kROS", change=change))
-astar.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="astar", change=change))
-jCPm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jCPm", change=change))
-jSGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jSGm", change=change))
-b.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="b", change=change))
-
-stopCluster(cl)  # Stop cluster
-
-# Combine output with input
-sens.all <- ls(pattern="*.sens")
-for (i in 1:length(sens.all)) {
-  assign(sens.all[i], cbind(sims, get(sens.all[i])))
-}
-
-# When growth rate becomes negative, set all responses to NA
-for (i in 1:length(sens.all)) {
-  if (any(get(sens.all[i])$grchange < 0, na.rm=T)) {
-    assign(sens.all[i], within(get(sens.all[i]), {
-      shchange <- replace(shchange, grchange < 0, NA)
-      cROSchange <- replace(cROSchange, grchange < 0, NA)
-      grchange <- replace(grchange, grchange < 0, NA)
-    }))
-  }
-}
-
-# Plot Figure
-par(mfrow=c(2,2), mar=c(3,3,1,1), mgp=c(2,0.5,0))
-sens.plot(response=c("grchange", "shchange"), pars=c("jCPm", "kCO2", "astar"), cols=c("black", "red", "gold"))
-sens.plot(response=c("grchange", "shchange"), pars=c("jNm", "KN"), cols=c("black", "blue"))
-sens.plot(response=c("grchange", "shchange"), pars=c("jST0", "jHT0", "jSGm", "jHGm"), cols=c("burlywood4", "burlywood3", "darkolivegreen4", "darkolivegreen3"))
-sens.plot(response=c("grchange", "shchange"), pars=c("kNPQ", "kROS", "b"), cols=c("black", "orange", "purple"))
-```
-
------
-
-# Figure S7
-```{r sens_HLHN, cache=T, echo=F, eval=T, fig.cap="Fig. S7. Sensitivity analysis under high DIN (4e-6 M) and high light (25 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
-# Set up time vector and parameters
-defpars <- def_pars()  # Get default parameters
-defpars <- replace(defpars, "initS", 1)  # Start with high S biomass to avoid alternate steady state of negative growth
-
-# Define environments for conducting sensitivity analysis
-envs <- list(
-  HLHN=list(L=25, N=4e-6, X=0)
-)
-
-# Define relative changes in each parameter to evaluate
-levs <- seq(0.5, 1.5, 0.05)  # Set levels of % change in parameter value to evaluate
-
-# Define all simulations to be run for a given parameter
-sims <- expand.grid(env=names(envs), change=levs)
-n <- nrow(sims)
-
-# Set up cluster for parallel processing
-cl <- makeCluster(detectCores())  # Initiate cluster
-registerDoParallel(cl)
-
-# Run sensitivity analysis in each environment for each parameter
-jHT0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHT0", change=change))
-jNm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jNm", change=change))
-jHGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jHGm", change=change))
-kCO2.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kCO2", change=change))
-KN.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="KN", change=change))
-jST0.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jST0", change=change))
-kNPQ.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kNPQ", change=change))
-kROS.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="kROS", change=change))
-astar.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="astar", change=change))
-jCPm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jCPm", change=change))
-jSGm.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="jSGm", change=change))
-b.sens <- foreach(i=1:n, .combine=rbind) %dopar% with(sims[i,], sens(env=envs[[env]], pars=defpars, par="b", change=change))
-
-stopCluster(cl)  # Stop cluster
-
-# Combine output with input
-sens.all <- ls(pattern="*.sens")
-for (i in 1:length(sens.all)) {
-  assign(sens.all[i], cbind(sims, get(sens.all[i])))
-}
-
-# When growth rate becomes negative, set all responses to NA
-for (i in 1:length(sens.all)) {
-  if (any(get(sens.all[i])$grchange < 0, na.rm=T)) {
-    assign(sens.all[i], within(get(sens.all[i]), {
-      shchange <- replace(shchange, grchange < 0, NA)
-      cROSchange <- replace(cROSchange, grchange < 0, NA)
-      grchange <- replace(grchange, grchange < 0, NA)
-    }))
-  }
-}
-
-# Plot Figure
-par(mfrow=c(2,2), mar=c(3,3,1,1), mgp=c(2,0.5,0))
-sens.plot(response=c("grchange", "shchange"), pars=c("jCPm", "kCO2", "astar"), cols=c("black", "red", "gold"))
-sens.plot(response=c("grchange", "shchange"), pars=c("jNm", "KN"), cols=c("black", "blue"))
-sens.plot(response=c("grchange", "shchange"), pars=c("jST0", "jHT0", "jSGm", "jHGm"), cols=c("burlywood4", "burlywood3", "darkolivegreen4", "darkolivegreen3"))
-sens.plot(response=c("grchange", "shchange"), pars=c("kNPQ", "kROS", "b"), cols=c("black", "orange", "purple"))
-```
-
------
-
-# Figure S8
-```{r FigS8, cache=T, fig.width=7, fig.height=7, echo=F, eval=T, fig.cap="**Fig.
-S8.
-Full system dynamics of seasonal light variation at two nutrient levels.** This figure accompanies Figure 5 in the main text, and shows additional model fluxes that demonstrate the mechanisms underlying the changes in the system state in these simulations."}
-# Set parameters
-defpars <- def_pars()  # Get default parameters
-
-# Set run time vector
-time <- seq(1, 365, 0.1)  # if single model run, use time input
-
-# Initialize environments
-env1 <- init_env(time=time, L=c(20,44,2), N=c(0.14e-6,0.14e-6,0), X=c(1e-5,1e-5,0))
-env2 <- init_env(time=time, L=c(20,44,2), N=c(10.14e-6,10.14e-6,0), X=c(1e-5,1e-5,0))
-
-# Run simulations
-ss1 <- with(run_coral_ss(env=list(L=33,N=0.14e-6,X=1e-5), pars=defpars, dt=0.1), last(S$S/H$H))
-run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", ss1))
-ss2 <- with(run_coral_ss(env=list(L=33,N=10.14e-6,X=1e-5), pars=defpars, dt=0.1), last(S$S/H$H))
-run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", ss2))
-
-par(mfcol=c(6,2), mar=c(2,2,1,4), mgp=c(1,0,0), tcl=0.2, xaxs="i")
-
-with(run1, {
-      # External irradiance
-    plot(time, env$L, type="l", col="gold", ylim=c(0,60), lwd=2, xlab="", ylab="mol ph/m2/d")
-    title("External irradiance and DIN", adj=0, line=0.25)
-    par(new=T)
-    plot(time, env$N, type="l", col="blue", ylim=c(0,10.14e-6), lwd=2, axes=F, xlab="", ylab="")
-    axis(side=4); mtext(side=4, text="molN/L", line=1, cex=0.75)
-
-    # Symbiont to host biomass ratio
-    totSH <- S$S / H$H
-    totSHf <- totSH[length(totSH)]
-    plot(time, totSH, type="l", col="black", ylim=c(0, 0.25), ylab="CmolS/CmolH", xlab="", lwd=2)
-    text(time[0.95*length(time)], totSHf, labels=as.character(round(totSHf, 3)), pos=3, xpd=T, cex=0.75)
-    title("Symbiont:host biomass", adj=0, line=0.25)
-
-    # Specific growth rates of host and symbiont
-    Hgrf <- H$dH.Hdt[length(H$dH.Hdt)]
-    plot(time, H$dH.Hdt, type="l", ylim=c(-0.01, 0.13), 
-         xlab="", ylab="d-1", lwd=2, cex=1, cex.lab=1)
-    if(any(c(H$dH.Hdt[-(1:10)], S$dS.Sdt[-(1:10)])<0)) abline(h=0, lwd=0.5, lty=3)
-    text(time[0.95*length(time)], Hgrf, labels=as.character(round(Hgrf, 3)), pos=3, xpd=T, cex=0.75)
-    #title("State variable dynamics", adj=0, cex.main=2, outer = T)
-    title("Specific growth rates", adj=0, line=0.25)
-    lines(time, S$dS.Sdt, col="black", lwd=1)
-    legend("topright", legend=c("Host", "Sym"), lwd=c(2,1), col="black", bty="n", y.intersp=1, cex=0.75)
-
-    # Light quenching (carbon fixation, NPQ, and excess (=ROS producing))
-    plot(NA, xlim=range(time), xlab="", ylim=c(0,130), ylab="mol ph/CmolS/d")
-    title("Light quenching", adj=0, line=0.25)
-    lines(time, S$jCP/pars$yCL, col="gold", lwd=1) # amt. quenched by photosynthesis
-    lines(time, S$jCP/pars$yCL + S$jNPQ, col="gold", lty=1, lwd=2) # amt. quenched by photo + NPQ
-    lines(time, S$jL, col="gold", lty=3, lwd=2) # total amt. absorbed
-    legend("topright", legend=c("ROS", "NPQ", "Photo."), lty=c(3,1,1), lwd=c(2,2,1), col="gold", bty="n", cex=0.75)
-
-    # Photosynthesis: substrate-limitation
-    pl <- log(   pmin((H$jCO2 + H$rCH)*H$H/S$S + S$rCS, pars$jCPm)   /   pmin(S$jL * pars$yCL, pars$jCPm)    )
-    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-2, 2))
-    title("Photosynthesis: substrate-limitation", adj=0, line=0.25)
-    lines(time, pl, col="black", lty=1, lwd=2)
-    abline(h=0, lty=3)
-    text(par("usr")[2], 0, labels="L-lim.\nCO2-lim.", cex=0.75, adj=1)
-
-    # Biomass formation - substrate-limitation
-    sl <- log(  pmin(S$jCP, pars$jSGm)   /   pmin((H$rhoN*H$H/S$S + S$rNS)/pars$nNS, pars$jSGm)  )
-    hl <- log(  pmin(S$rhoC*S$S/H$H + H$jX, pars$jHGm) / pmin((H$jN + pars$nNX*H$jX + H$rNH) / pars$nNH, pars$jHGm)  )
-    maxabs <- max(c(abs(hl), abs(sl)))
-    range <- range(c(hl, sl))
-    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-1.5,1.5))
-    title("Biomass: substrate-limitation", adj=0, line=0.25)
-    lines(time, hl, col="black", lty=1, lwd=2)
-    lines(time, sl, col="black", lty=1, lwd=1)
-    abline(h=0, lty=3)
-    text(par("usr")[2], 0, labels="N-lim.\nC-lim.", cex=0.75, adj=1)
-    legend("topright", legend=c("Host", "Sym"), lty=1, lwd=c(2,1), col="black", bty="n", cex=0.75)
-})
-
-with(run2, {
-      # External irradiance
-    plot(time, env$L, type="l", col="gold", ylim=c(0,60), lwd=2, xlab="", ylab="mol/m2/d")
-    title("External irradiance and DIN", adj=0, line=0.25)
-    par(new=T)
-    plot(time, env$N, type="l", col="blue", ylim=c(0,10.14e-6), lwd=2, axes=F, xlab="", ylab="")
-    axis(side=4); mtext(side=4, text="molN/L", line=1, cex=0.75)
-
-    # Symbiont to host biomass ratio
-    totSH <- S$S / H$H
-    totSHf <- totSH[length(totSH)]
-    plot(time, totSH, type="l", col="black", ylim=c(0, 0.25), ylab="CmolS/CmolH", xlab="", lwd=2)
-    text(time[0.95*length(time)], totSHf, labels=as.character(round(totSHf, 3)), pos=3, xpd=T, cex=0.75)
-    title("Symbiont:host biomass", adj=0, line=0.25)
-
-    # Specific growth rates of host and symbiont
-    Hgrf <- H$dH.Hdt[length(H$dH.Hdt)]
-    plot(time, H$dH.Hdt, type="l", ylim=c(-0.01,0.13), 
-         xlab="", ylab="d-1", lwd=2, cex=1, cex.lab=1)
-    if(any(c(H$dH.Hdt[-(1:10)], S$dS.Sdt[-(1:10)])<0)) abline(h=0, lwd=0.5, lty=3)
-    text(time[0.95*length(time)], Hgrf, labels=as.character(round(Hgrf, 3)), pos=3, xpd=T, cex=0.75)
-    #title("State variable dynamics", adj=0, cex.main=2, outer = T)
-    title("Specific growth rates", adj=0, line=0.25)
-    lines(time, S$dS.Sdt, col="black", lwd=1)
-    legend("topright", legend=c("Host", "Sym"), lwd=c(2,1), col="black", bty="n", y.intersp=1, cex=0.75)
-
-    # Light quenching (carbon fixation, NPQ, and excess (=ROS producing))
-    plot(NA, xlim=range(time), xlab="", ylim=c(0,130), ylab="mol ph/CmolS/d")
-    title("Light quenching", adj=0, line=0.25)
-    lines(time, S$jCP/pars$yCL, col="gold", lwd=1) # amt. quenched by photosynthesis
-    lines(time, S$jCP/pars$yCL + S$jNPQ, col="gold", lty=1, lwd=2) # amt. quenched by photo + NPQ
-    lines(time, S$jL, col="gold", lty=3, lwd=2) # total amt. absorbed
-    legend("topright", legend=c("ROS", "NPQ", "Photo."), lty=c(3,1,1), lwd=c(2,2,1), col="gold", bty="n", cex=0.75)
-
-    # Photosynthesis: substrate-limitation
-    pl <- log(   pmin((H$jCO2 + H$rCH)*H$H/S$S + S$rCS, pars$jCPm)   /   pmin(S$jL * pars$yCL, pars$jCPm)    )
-    maxabs <- max(abs(pl))
-    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-2, 2))
-    title("Photosynthesis: substrate-limitation", adj=0, line=0.25)
-    lines(time, pl, col="black", lty=1, lwd=2)
-    abline(h=0, lty=3)
-    text(par("usr")[2], 0, labels="L-lim.\nCO2-lim.", cex=0.75, adj=1)
-
-    # Biomass formation - substrate-limitation
-    sl <- log(  pmin(S$jCP, pars$jSGm)   /   pmin((H$rhoN*H$H/S$S + S$rNS)/pars$nNS, pars$jSGm)  )
-    hl <- log(  pmin(S$rhoC*S$S/H$H + H$jX, pars$jHGm) / pmin((H$jN + pars$nNX*H$jX + H$rNH) / pars$nNH, pars$jHGm)  )
-    maxabs <- max(c(abs(hl), abs(sl)))
-    range <- range(c(hl, sl))
-    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-1.5,1.5))
-    title("Biomass: substrate-limitation", adj=0, line=0.25)
-    lines(time, hl, col="black", lty=1, lwd=2)
-    lines(time, sl, col="black", lty=1, lwd=1)
-    abline(h=0, lty=3)
-    text(par("usr")[2], 0, labels="N-lim.\nC-lim.", cex=0.75, adj=1)
-    legend("topright", legend=c("Host", "Sym"), lty=1, lwd=c(2,1), col="black", bty="n", cex=0.75)
-})
-```
-
------
-
 # Figure S9
@@
 # Set input values for steady state runs
-input <- expand.grid(L=seq(1,60,1), initS=c(0.0001,0.1))
+input <- expand.grid(L=seq(59,60,1), initS=c(0.0001,0.1))
 
@@
 ```
-
------
-
-# References

### 2017-02-08 JTB_supp/test_figs9.Rmd
try removing chunk opts?

@@
 # Figure S9
-```{r FigS9, cache=T, fig.width=7, fig.height=7, echo=F, message=F, eval=T, fig.cap="**Fig.
-S9.
-Hysteresis under different nutrient and feeding regimes.** This figure accompanies Fig.
-2 in the main text.
-These figures show steady state S:H biomass ratios in constant environments with varying DIN and food (X) levels, with initial S:H biomass values of either 0.1 or 0.0001.
-Points are colored according to whether the host exhibits positive (black) or negative (red) growth at steady state."}
+```{r FigS9}
 

### 2017-02-08 JTB_supp/JTB_supp.Rmd
suppress unwanted NULL output from apply functions in Fig S9 that were causing knit error

@@
 # Figure S9
-```{r FigS9, cache=T, fig.width=7, fig.height=7, echo=F, message=F, eval=T, fig.cap="**Fig.
+```{r FigS9, cache=T, fig.width=7, fig.height=7, echo=F, message=F, eval=T, render=NULL, fig.cap="**Fig.
 S9.
@@
 text(c(48,53), 0.47, labels=c("pos.", "neg."), xpd=T, cex=0.7)
-apply(res1, 1, FUN=function(x) {
+p1 <- apply(res1, 1, FUN=function(x) {
   with(x, points(L, sh,
@@
 title("DIN=1e-7 mol/L; X=1e-6 mol/L")
-apply(res2, 1, FUN=function(x) {
+p2 <- apply(res2, 1, FUN=function(x) {
   with(x, points(L, sh,
@@
 title("DIN=1e-6 mol/L; X=2e-6 mol/L")
-apply(res3, 1, FUN=function(x) {
+p3 <- apply(res3, 1, FUN=function(x) {
   with(x, points(L, sh,
@@
 title("DIN=1e-6 mol/L; X=4e-6 mol/L")
-apply(res4, 1, FUN=function(x) {
+p4 <- apply(res4, 1, FUN=function(x) {
   with(x, points(L, sh,

### 2017-02-08 JTB_ms/JTB_ms.Rmd
more updates for revision

@@
 
-MAKE INTRO INCLUDE SOMETHING ABOUT ALTERNATIVE STABLE STATES
-
 The nutritional exchange between corals and *Symbiodinium* directly underlies the capacity of corals to build coral reef ecosystems, worth trillions of US Dollars annually [@Costanza:2014ex].
@@
 Bleaching is thought to be triggered by photo-oxidative stress in *Symbiodinium* [@Weis:2008p944], which has been modeled previously as a response to absolute external irradiance [@Eynaud:2011tv].
-However, this response may also depend on self-shading by other symbionts, CO~2~ availability at the site of photosynthesis, and other non-photochemical quenching mechanisms.
+However, this response may also depend on self-shading by other symbionts [@Enriquez:2005p142], CO~2~ availability at the site of photosynthesis [@Wooldridge:2009p7807], and other non-photochemical quenching mechanisms [@Roth:2014wf].
 We incorporate these features into a new photo-oxidative stress module linking overreduction of the photosynthetic light reactions to downstream impacts of photoinhibition and photodamage.
@@
 
-With these as our primary motivations, we describe a simplified approach to dynamic bioenergetic modeling of coral-*Symbiodinium* symbioses that dynamically integrates the influence of external irradiance, nutrients, and prey availability on coral growth and symbiosis dynamics (i.e., symbiont:host biomass ratios), allowing for the possibility of coral bleaching in the event of photooxidative stress.
-In the following sections, we describe and provide rationale for the model structure, and demonstrate a range of steady state and dynamic behaviors that are consistent with observed phenomena.
+With these as our primary motivations, we describe a simplified approach to bioenergetic modeling of coral-*Symbiodinium* symbioses that dynamically integrates the influences of external irradiance, nutrients, and prey availability on coral growth and symbiosis dynamics (i.e., changes in symbiont:host biomass ratios), allowing for the possibility of coral bleaching in response to photooxidative stress.
+An emergent finding of this work is that coral bleaching can be interpreted as an alternate stable state of the symbiotic system, which provides a new framework for understanding the mechanisms that drive a coral into a bleached state, as well as those that facilitate recovery.
+In the following sections, we describe and provide rationale for the model structure, demonstrate a range of steady state and dynamic behaviors that are consistent with observed phenomena, and discuss new insights from this work in understanding responses of coral-*Symbiodinium* symbioses to environmental change.
 
 # Model description
-In this dynamical system, both partners acquire and use carbon and nitrogen to construct biomass: the symbiont fixes carbon through photosynthesis, and receives nitrogen shared by the host, while the host acquires nitrogen from the environment and receives carbon shared by the symbiont.
+In this dynamical system, both the coral host and algal symbiont acquire and use carbon and nitrogen to construct biomass.
+The symbiont fixes carbon through photosynthesis and receives nitrogen shared by the host, while the host acquires nitrogen from the environment and receives carbon shared by the symbiont.
 A graphical representation of the model is presented in Fig.
@@
 To simulate bleaching, we model the absorption and quenching of light energy by photochemistry and non-photochemical quenching, and the responses that occur (i.e., photoinhibition, photodamage, and symbiont loss) when these quenching capacities are overwhelmed.
-While bleaching in response to high light alone has been observed experimentally [@Schutter:2011ie; @Downs:2013kc], mass coral bleaching events occur concurrently with high temperature [@HoeghGuldberg:1999p4222]; thus, it is important to justify our consideration of light as the primary stressor.
+While bleaching in response to high light alone has been observed experimentally [@Schutter:2011ie; @Downs:2013kc], mass coral bleaching events occur concurrently with high temperature [@HoeghGuldberg:1999p4222].
+Thus, it is important to justify our consideration of light as the primary stressor.
 In reality, light and temperature interact synergistically [@Coles:1978p1124; @Jones:1998p3572], and in fact, any stressor that disrupts the quenching of light energy may lead to bleaching [@Wooldridge:2010p7809; @Baker:2015kp].
@@
 
-Below we describe the formulation of each flux in the model.
-
 Table 1. Model fluxes (mass-specific). Units are explained in the text.
@@
 | Symbol  | Description                          | Units                 | Eq. no. |
-| ------- | -------------------------------------- | ------------------- | ------- |
+| ------- | ------------------------------------ | --------------------- | ------- |
 | $j_X$     | Prey uptake rate                       | mol X C-mol H^-1^ d^-1^ | 3 |
@@
 | $j_{HG}$  | Host biomass formation rate            | C-mol H C-mol H^-1^ d^-1^| 5 |
-| $j_{HT}$  | Host biomass turnover rate             | C-mol H C-mol H^-1^ d^-1^| 21 |
-| $r_{NH}$  | Recycled nitrogen from host turnover   | mol N C-mol H^-1^ d^-1^ | 6 |
-| $\rho_N$  | Nitrogen shared with the symbiont      | mol N C-mol H^-1^ d^-1^ | 7 |
-| $j_{eC}$  | Excess carbon used to activate host CCMs| mol C C-mol H^-1^ d^-1^ | 8 |
-| $j_{CO_2}$| CO~2~ input to photosynthesis          | mol CO~2~ C-mol H^-1^ d^-1^| 9 |
-| $j_L$     | Light absorption rate                  | mol q C-mol S^-1^ d^-1^| 11 |
-| $r_{CH}$  | Recycled CO~2~ from host turnover      | mol CO~2~ C-mol H^-1^ d^-1^| 12 |
-| $r_{CS}$  | Recycled CO~2~ from symbiont turnover  | mol CO~2~ C-mol S^-1^ d^-1^| 13 |
-| $j_{CP}$  | Photosynthesis rate                    | mol C C-mol S^-1^ d^-1^ | 14 |
-| $j_{eL}$  | Light energy in excess of photochemistry| mol q C-mol S^-1^ d^-1^| 15 |
-| $j_{NPQ}$ | Total capacity of NPQ                  | mol q C-mol S^-1^ d^-1^ | 16 |
-| $c_{ROS}$ | ROS production proportional to baseline | -- | 17 |
-| $r_{NS}$  | Recycled nitrogen from symbiont turnover| mol N C-mol S^-1^ d^-1^| 18 |
-| $j_{SG}$  | Symbiont biomass formation rate        | C-mol S C-mol S^-1^ d^-1^| 19 |
-| $\rho_C$  | Fixed carbon shared with host          | mol C C-mol S^-1^ d^-1^ | 20 |
-| $j_{ST}$  | Symbiont biomass turnover rate         | C-mol S C-mol S^-1^ d^-1^| 21 |
+| $j_{HT}$  | Host biomass turnover rate             | C-mol H C-mol H^-1^ d^-1^| 6 |
+| $r_{NH}$  | Recycled nitrogen from host turnover   | mol N C-mol H^-1^ d^-1^ | 7 |
+| $\rho_N$  | Nitrogen shared with the symbiont      | mol N C-mol H^-1^ d^-1^ | 8 |
+| $j_{eC}$  | Excess carbon used to activate host CCMs| mol C C-mol H^-1^ d^-1^ | 9 |
+| $j_{CO_2}$| CO~2~ input to photosynthesis          | mol CO~2~ C-mol H^-1^ d^-1^| 10 |
+| $j_L$     | Light absorption rate                  | mol photons C-mol S^-1^ d^-1^| 12 |
+| $r_{CH}$  | Recycled CO~2~ from host turnover      | mol CO~2~ C-mol H^-1^ d^-1^| 13 |
+| $r_{CS}$  | Recycled CO~2~ from symbiont turnover  | mol CO~2~ C-mol S^-1^ d^-1^| 14 |
+| $j_{CP}$  | Photosynthesis rate                    | mol C C-mol S^-1^ d^-1^ | 15 |
+| $j_{eL}$  | Light energy in excess of photochemistry| mol photons C-mol S^-1^ d^-1^| 16 |
+| $j_{NPQ}$ | Total capacity of NPQ                  | mol photons C-mol S^-1^ d^-1^ | 17 |
+| $c_{ROS}$ | ROS production proportional to baseline | -- | 18 |
+| $r_{NS}$  | Recycled nitrogen from symbiont turnover| mol N C-mol S^-1^ d^-1^| 19 |
+| $j_{SG}$  | Symbiont biomass formation rate        | C-mol S C-mol S^-1^ d^-1^| 20 |
+| $\rho_C$  | Fixed carbon shared with host          | mol C C-mol S^-1^ d^-1^ | 21 |
+| $j_{ST}$  | Symbiont biomass turnover rate         | C-mol S C-mol S^-1^ d^-1^| 22 |
 
-Table 2. Model parameters (units explained in the text). Derivations of each parameter value along with supporting references are provided in the Supplementary Information.
+Table 2. Model parameters (units explained in the text). Justification and/or derivation of each parameter value along with supporting references are provided in the Supplementary Information.
 
 | Symbol  | Description                          | Units               | Value     |
-| ------- | -------------------------------------- | ------------------- | ------- |
+| ------- | ------------------------------------ | ------------------- | --------- |
 | $n_{NH}$ | N:C molar ratio in host biomass | -- | $`r defpars$nNH`$ |
@@
 | $j_{HGm}$ | Maximum specific growth rate of host | C-mol H C-mol H^-1^ d^-1^ | $`r defpars$jHGm`$ |
-| $y_{CL}$ | Quantum yield of photosynthesis | mol C mol q^-1^ | $`r defpars$yCL`$ |
+| $y_{CL}$ | Quantum yield of photosynthesis | mol C mol photons^-1^ | $`r defpars$yCL`$ |
 | $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | m^2^ C-mol S^-1^ | $`r defpars$astar`$ |
-| $k_{NPQ}$ | NPQ capacity of symbiont | mol ph C-mol S^-1^ d^-1^ | $`r defpars$kNPQ`$ |
-| $k_{ROS}$ | Excess photon energy that doubles ROS production, relative to baseline levels | mol ph C-mol S^-1^ d^-1^ | $`r defpars$kROS`$ |
+| $k_{NPQ}$ | NPQ capacity of symbiont | mol photons C-mol S^-1^ d^-1^ | $`r defpars$kNPQ`$ |
+| $k_{ROS}$ | Excess photon energy that doubles ROS production, relative to baseline levels | mol photons C-mol S^-1^ d^-1^ | $`r defpars$kROS`$ |
 | $j_{CPm}$ | Maximum specific photosynthesis rate of symbiont | mol C C-mol S^-1^ d^-1^ | $`r defpars$jCPm`$ |
 | $j_{SGm}$ | Maximum specific growth rate of symbiont | C-mol S C-mol S^-1^ d^-1^ | $`r defpars$jSGm`$ |
-| $b$ | Scaling parameter for strength of bleaching response | -- | $`r defpars$b`$ |
+| $b$ | Scaling parameter for bleaching response | -- | $`r defpars$b`$ |
 
@@
 
-Additionally, the coral animal acquires nitrogen dissolved in the surrounding seawater, which is assumed to represent ammonium, the primary form utilized by corals [@Yellowlees:2008p331].
+Additionally, the coral animal acquires dissolved inorganic nitrogen (DIN) from the surrounding seawater, which is assumed to represent ammonium, the primary form utilized by corals [@Yellowlees:2008p331].
 This gives the host (rather than the symbiont) priority in nitrogen utilization; this capacity is supported by experimental evidence [@Wang:1998p128] and is consistent with the spatial arrangment of the partners, where the host is in direct contact with the external environment.
@@
 where $\rho_C$ is fixed carbon shared by the symbiont (see Eq.
-20), and $r_{NH}$ is recycled nitrogen liberated by host biomass turnover (see Eq.
-XX).
-Host biomass turnover is equal to the specific maintenance rate of host biomass, which is set as a parameter $j_{HT}^0$,
+21), and $r_{NH}$ is recycled nitrogen liberated by host biomass turnover (see Eq.
+7).
+Host biomass turnover is equal to the specific maintenance rate of host biomass,
 
@@
 CO~2~ arrives at the photosynthesis SU from multiple sources: in addition to the CO~2~ actively supplied by the host through its CCMs ($j_{CO_2}$; Eq.
-9), we assume a fixed proportion $\sigma_{CH}$ of metabolic CO~2~ generated by host biomass turnover is passively available to the photosynthesis SU, according to:
+10), we assume a fixed proportion $\sigma_{CH}$ of metabolic CO~2~ generated by host biomass turnover is passively available to the photosynthesis SU, according to:
 
@@
 where $j_{CPm}$ is the maximum specific rate of photosynthesis, and $c_{ROS}$ is the relative rate of reactive oxygen species production (see Eq.
-17).
+18).
 Dividing the photosynthetic rate by $c_{ROS}$ causes a decline in response to photooxidative stress at high light levels, and the emergent outcome of this SU formulation demonstrates a classic photoinhibition response (Fig.
@@
 
-Carbon fixed by photosynthesis ($j_{CP}$; Eq. 14) is then combined with nitrogen shared by the host ($\rho_N$; Eq. 7) and nitrogen recycled from symbiont biomass turnover[^2]
+Carbon fixed by photosynthesis ($j_{CP}$; Eq. 15) is then combined with nitrogen shared by the host ($\rho_N$; Eq. 8) and nitrogen recycled from symbiont biomass turnover[^3]
 
@@
 [^2]: Note that recycling of symbiont biomass turnover ($rNS$ and $rCS$) only occurs based on the maintenance component of turnover (i.e., $j_{ST}^0$), and not the photodamage/bleaching component, as this loss represents biomass that is expelled from the host.
+[^3]: See footnote 2.
 
@@
 
-The model dynamics are specified by the differential equations (1-2) that impose biomass balance for host and symbiont and by a set of coupled non-linear algebraic equations (3-21) that define fluxes.
+The model dynamics are specified by the differential equations (1-2) that impose biomass balance for host and symbiont and by a set of coupled non-linear algebraic equations (3-22) that define fluxes.
 Several of these fluxes are defined *implicitly*; for example, the rejection fluxes of carbon and nitrogen from the symbiont and host biomass SUs, respectively, act as reciprocal input fluxes to the other SU.
@@
 2).
-The mechanism that produces these alternate stable states is the positive feedback between carbon-limitation of the host and CO~2~-limitation of photosynthesis: if symbiont biomass is initially very low (i.e., a "bleached" coral), the system cannot escape this positive feedback and cannot grow (unless feeding is sufficiently high).
+The mechanism that produces these alternate stable states is the positive feedback between carbon-limitation of the host and CO~2~-limitation of photosynthesis: if symbiont biomass is initially very low (i.e., a "bleached" coral), very little carbon is fixed, and the system cannot escape this positive feedback and cannot grow (unless feeding is sufficiently high).
 However, if symbiont biomass is initially high (i.e., a "healthy" coral), then the system remains in a nitrogen-limited state with positive growth.
@@
 Observed specific growth rates in several coral species fall near or below the lower end of this range (~0.01 d^-1^) [@Osinga:2011kq; @Osinga:2012is], though values as high as 0.025 d^-1^ have been reported in *Galaxea fascicularis* [@Schutter:2010p7758], and `r Armoza_gr` d^-1^ in *Aiptasia diaphana*, a non-calcifying symbiotic anemone [@ArmozaZvuloni:2014ju].
-It is not surprising that the model predicts slightly higher growth than is typically observed, since it does not account for ecological factors that may limit growth (e.g., competition, predation, bioerosion).
+However, it is not surprising that observed growth rates are often lower than model predictions, since the model does not account for ecological factors that may limit growth (e.g., competition, predation, bioerosion).
 Furthermore, while most measurements are made on skeletal growth, the model predicts biomass growth, which may not always be strongly correlated [@Anthony:2002tc].
 
-As irradiance continues to increase above ~25 mol photons m^-2^ d^-1^, growth rates decline until positive growth ceases above ~40 µmol photons m^-2^ d^-1^ (Fig.
+At irradiance levels above ~25 mol photons m^-2^ d^-1^, steady state growth rates decline until positive growth ceases above ~40 µmol photons m^-2^ d^-1^ (Fig.
 3A).
@@
 
-The prediction of higher growth when light is reduced indicates that growth is not limited by low light in winter, but is actually reduced by excess light in summer[^3], consistent with the experimental findings of @Schutter:2011ie.
+The prediction of higher growth when light is reduced indicates that growth is not limited by low light in winter, but is actually reduced by excess light in summer[^4], consistent with the experimental findings of @Schutter:2011ie.
 Seasonal summertime reductions in tissue biomass have also been well-documented in the field [@Fitt:2000p3792], along with reductions in net photosynthetic capacity [@MullerParker:1987cr].
-Importantly, while light alone may drive seasonal dynamics in the ways discussed[^3], temperature fluctuations may attenuate or even reverse the effect of light as cooler winters depress metabolism; thus, the relative magnitude of fluctuation in temperature, light, and other factors may produce wide variability in the direction and magnitude of seasonal changes in growth and symbiont abundance, depending on location and microhabitat.
+Importantly, while light alone may drive seasonal dynamics in the ways discussed[^5], temperature fluctuations may attenuate or even reverse the effect of light as cooler winters depress metabolism; thus, the relative magnitude of fluctuation in temperature, light, and other factors may produce wide variability in the direction and magnitude of seasonal changes in growth and symbiont abundance, depending on location and microhabitat.
 Nevertheless, the seasonal variability predicted here (Fig.
@@
 
-[^3]: at the light levels indicated in @Stimson:1997p3837.
+[^4]: at the light levels indicated in @Stimson:1997p3837.
 Note that if light levels were reduced throughout the year (e.g.
 for a coral at greater depth) such that light did not cause photo-oxidative stress in summer but became limiting to photosynthesis in winter, the S:H ratio would still increase in winter, but growth would decrease; the latter scenario is predicted both by the present model as well as that of @Muller:2009io, since photo-oxidative stress does not become relevant.
+[^5]: See footnote 4.
 
@@
 Substrate limitation coefficients were calculated using Eq.
-22.
+23.
 All parameters were set at default values with external DIN=1e-7 mol N/L and prey density=0).](../img/Fig6.png)
@@
 Substrate limitation coefficients were calculated using Eq.
-22.
+23.
 All simulations were run with default parameters (except varying $n_{NX}$), with L=10 mol photons m^2^ d^-1^, DIN=1e-7 mol N L^-1^, prey=1e-6 CmolX L^-1^, and initial S:H biomass=0.001.](../img/Fig8.png)

### 2017-02-08 JTB_ms/JTB_ms.tex
more updates for revision

@@
 
-MAKE INTRO INCLUDE SOMETHING ABOUT ALTERNATIVE STABLE STATES
-
 The nutritional exchange between corals and \emph{Symbiodinium} directly
@@
   2011). However, this response may also depend on self-shading by other
-  symbionts, CO\textsubscript{2} availability at the site of
-  photosynthesis, and other non-photochemical quenching mechanisms. We
-  incorporate these features into a new photo-oxidative stress module
-  linking overreduction of the photosynthetic light reactions to
-  downstream impacts of photoinhibition and photodamage. Importantly,
-  this formulation introduces a link between
+  symbionts (Enríquez, Méndez, and Iglesias-Prieto 2005),
+  CO\textsubscript{2} availability at the site of photosynthesis
+  (Wooldridge 2009), and other non-photochemical quenching mechanisms
+  (Roth 2014). We incorporate these features into a new photo-oxidative
+  stress module linking overreduction of the photosynthetic light
+  reactions to downstream impacts of photoinhibition and photodamage.
+  Importantly, this formulation introduces a link between
   CO\textsubscript{2}-limitation of photosynthesis and bleaching, and
@@
 With these as our primary motivations, we describe a simplified approach
-to dynamic bioenergetic modeling of coral-\emph{Symbiodinium} symbioses
-that dynamically integrates the influence of external irradiance,
-nutrients, and prey availability on coral growth and symbiosis dynamics
-(i.e., symbiont:host biomass ratios), allowing for the possibility of
-coral bleaching in the event of photooxidative stress. In the following
-sections, we describe and provide rationale for the model structure, and
-demonstrate a range of steady state and dynamic behaviors that are
-consistent with observed phenomena.
+to bioenergetic modeling of coral-\emph{Symbiodinium} symbioses that
+dynamically integrates the influences of external irradiance, nutrients,
+and prey availability on coral growth and symbiosis dynamics (i.e.,
+changes in symbiont:host biomass ratios), allowing for the possibility
+of coral bleaching in response to photooxidative stress. An emergent
+finding of this work is that coral bleaching can be interpreted as an
+alternate stable state of the symbiotic system, which provides a new
+framework for understanding the mechanisms that drive a coral into a
+bleached state, as well as those that facilitate recovery. In the
+following sections, we describe and provide rationale for the model
+structure, demonstrate a range of steady state and dynamic behaviors
+that are consistent with observed phenomena, and discuss new insights
+from this work in understanding responses of coral-\emph{Symbiodinium}
+symbioses to environmental change.
 
@@
 
-In this dynamical system, both partners acquire and use carbon and
-nitrogen to construct biomass: the symbiont fixes carbon through
-photosynthesis, and receives nitrogen shared by the host, while the host
-acquires nitrogen from the environment and receives carbon shared by the
-symbiont. A graphical representation of the model is presented in Fig.
-1, and each model flux and parameter is defined in Tables 1 and 2,
-respectively. We use C-moles (C-mol) as the unit of biomass for
-consistency with the rigorous mass balance of DEB theory: 1 C-mol is
-equivalent to the amount of biomass containing 1 mole of carbon atoms.
-Host biomass (expressed as C-mol H), symbiont biomass (expressed as
-C-mol S), and prey biomass (expressed as C-mol X) have fixed, but
+In this dynamical system, both the coral host and algal symbiont acquire
+and use carbon and nitrogen to construct biomass. The symbiont fixes
+carbon through photosynthesis and receives nitrogen shared by the host,
+while the host acquires nitrogen from the environment and receives
+carbon shared by the symbiont. A graphical representation of the model
+is presented in Fig. 1, and each model flux and parameter is defined in
+Tables 1 and 2, respectively. We use C-moles (C-mol) as the unit of
+biomass for consistency with the rigorous mass balance of DEB theory: 1
+C-mol is equivalent to the amount of biomass containing 1 mole of carbon
+atoms. Host biomass (expressed as C-mol H), symbiont biomass (expressed
+as C-mol S), and prey biomass (expressed as C-mol X) have fixed, but
 different, molar N:C ratios (Table 2). Biomass is produced from carbon
@@
 al. 2013), mass coral bleaching events occur concurrently with high
-temperature (Hoegh-Guldberg 1999); thus, it is important to justify our
+temperature (Hoegh-Guldberg 1999). Thus, it is important to justify our
 consideration of light as the primary stressor. In reality, light and
@@
 
-Below we describe the formulation of each flux in the model.
-
 Table 1. Model fluxes (mass-specific). Units are explained in the text.
-
-\begin{longtable}[]{@{}llll@{}}
-\toprule
-Symbol & Description & Units & Eq. no.\tabularnewline
-\midrule
-\endhead
-\(j_X\) & Prey uptake rate & mol X C-mol H\textsuperscript{-1}
-d\textsuperscript{-1} & 3\tabularnewline
-\(j_N\) & Nitrogen uptake rate & mol N Cmol H\textsuperscript{-1}
-d\textsuperscript{-1} & 4\tabularnewline
-\(j_{HG}\) & Host biomass formation rate & C-mol H C-mol
-H\textsuperscript{-1} d\textsuperscript{-1} & 5\tabularnewline
-\(j_{HT}\) & Host biomass turnover rate & C-mol H C-mol
-H\textsuperscript{-1} d\textsuperscript{-1} & 21\tabularnewline
-\(r_{NH}\) & Recycled nitrogen from host turnover & mol N C-mol
-H\textsuperscript{-1} d\textsuperscript{-1} & 6\tabularnewline
-\(\rho_N\) & Nitrogen shared with the symbiont & mol N C-mol
-H\textsuperscript{-1} d\textsuperscript{-1} & 7\tabularnewline
-\(j_{eC}\) & Excess carbon used to activate host CCMs & mol C C-mol
-H\textsuperscript{-1} d\textsuperscript{-1} & 8\tabularnewline
-\(j_{CO_2}\) & CO\textsubscript{2} input to photosynthesis & mol
-CO\textsubscript{2} C-mol H\textsuperscript{-1} d\textsuperscript{-1} &
-9\tabularnewline
-\(j_L\) & Light absorption rate & mol q C-mol S\textsuperscript{-1}
-d\textsuperscript{-1} & 11\tabularnewline
-\(r_{CH}\) & Recycled CO\textsubscript{2} from host turnover & mol
-CO\textsubscript{2} C-mol H\textsuperscript{-1} d\textsuperscript{-1} &
-12\tabularnewline
-\(r_{CS}\) & Recycled CO\textsubscript{2} from symbiont turnover & mol
-CO\textsubscript{2} C-mol S\textsuperscript{-1} d\textsuperscript{-1} &
-13\tabularnewline
-\(j_{CP}\) & Photosynthesis rate & mol C C-mol S\textsuperscript{-1}
-d\textsuperscript{-1} & 14\tabularnewline
-\(j_{eL}\) & Light energy in excess of photochemistry & mol q C-mol
-S\textsuperscript{-1} d\textsuperscript{-1} & 15\tabularnewline
-\(j_{NPQ}\) & Total capacity of NPQ & mol q C-mol S\textsuperscript{-1}
-d\textsuperscript{-1} & 16\tabularnewline
-\(c_{ROS}\) & ROS production proportional to baseline & -- &
-17\tabularnewline
-\(r_{NS}\) & Recycled nitrogen from symbiont turnover & mol N C-mol
-S\textsuperscript{-1} d\textsuperscript{-1} & 18\tabularnewline
-\(j_{SG}\) & Symbiont biomass formation rate & C-mol S C-mol
-S\textsuperscript{-1} d\textsuperscript{-1} & 19\tabularnewline
-\(\rho_C\) & Fixed carbon shared with host & mol C C-mol
-S\textsuperscript{-1} d\textsuperscript{-1} & 20\tabularnewline
-\(j_{ST}\) & Symbiont biomass turnover rate & C-mol S C-mol
-S\textsuperscript{-1} d\textsuperscript{-1} & 21\tabularnewline
-\bottomrule
-\end{longtable}
-
-Table 2. Model parameters (units explained in the text). Derivations of
-each parameter value along with supporting references are provided in
-the Supplementary Information.
 
@@
 Symbol\strut
-\end{minipage} & \begin{minipage}[b]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[b]{0.44\columnwidth}\raggedright\strut
+Description\strut
+\end{minipage} & \begin{minipage}[b]{0.26\columnwidth}\raggedright\strut
+Units\strut
+\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
+Eq. no.\strut
+\end{minipage}\tabularnewline
+\midrule
+\endhead
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_X\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Prey uptake rate\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol X C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+3\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_N\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Nitrogen uptake rate\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol N Cmol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+4\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{HG}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Host biomass formation rate\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+5\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{HT}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Host biomass turnover rate\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+6\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(r_{NH}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Recycled nitrogen from host turnover\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol N C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+7\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(\rho_N\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Nitrogen shared with the symbiont\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol N C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+8\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{eC}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Excess carbon used to activate host CCMs\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol C C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+9\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{CO_2}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+CO\textsubscript{2} input to photosynthesis\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol CO\textsubscript{2} C-mol H\textsuperscript{-1}
+d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+10\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_L\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Light absorption rate\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+12\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(r_{CH}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Recycled CO\textsubscript{2} from host turnover\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol CO\textsubscript{2} C-mol H\textsuperscript{-1}
+d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+13\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(r_{CS}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Recycled CO\textsubscript{2} from symbiont turnover\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol CO\textsubscript{2} C-mol S\textsuperscript{-1}
+d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+14\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{CP}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Photosynthesis rate\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol C C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+15\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{eL}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Light energy in excess of photochemistry\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+16\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{NPQ}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Total capacity of NPQ\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+17\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(c_{ROS}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+ROS production proportional to baseline\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+--\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+18\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(r_{NS}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Recycled nitrogen from symbiont turnover\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol N C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+19\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{SG}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Symbiont biomass formation rate\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+20\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(\rho_C\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Fixed carbon shared with host\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol C C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+21\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{ST}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Symbiont biomass turnover rate\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+22\strut
+\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
+
+Table 2. Model parameters (units explained in the text). Justification
+and/or derivation of each parameter value along with supporting
+references are provided in the Supplementary Information.
+
+\begin{longtable}[]{@{}llll@{}}
+\toprule
+\begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
+Symbol\strut
+\end{minipage} & \begin{minipage}[b]{0.44\columnwidth}\raggedright\strut
 Description\strut
@@
 Units\strut
-\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\raggedright\strut
 Value\strut
@@
 \(n_{NH}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 N:C molar ratio in host biomass\strut
@@
 --\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(0.18\)\strut
@@
 \(n_{NS}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 N:C molar ratio in symbiont biomass\strut
@@
 --\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(0.13\)\strut
@@
 \(n_{NX}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 N:C molar ratio in prey biomass\strut
@@
 --\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(0.2\)\strut
@@
 \(j_{HT}^0\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Specific maintenance rate of host biomass\strut
@@
 C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(0.03\)\strut
@@
 \(j_{ST}^0\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Specific maintenance rate of symbiont biomass\strut
@@
 C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(0.03\)\strut
@@
 \(\sigma_{NH}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Proportion host N-turnover recycled to host\strut
@@
 --\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(0.9\)\strut
@@
 \(\sigma_{CH}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Proportion host C-turnover recycled to photosynthesis\strut
@@
 --\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(0.1\)\strut
@@
 \(\sigma_{NS}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Proportion symbiont N-turnover recycled to symbiont\strut
@@
 \(\sigma_{CS}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Proportion symbiont C-turnover recycled to photosynthesis\strut
@@
 \(j_{Xm}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Maximum specific feeding rate of host\strut
@@
 molX C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(0.13\)\strut
@@
 \(K_X\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Half-saturation constant for prey uptake by host\strut
@@
 molX L\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(1.95\times 10^{-5}\)\strut
@@
 \(j_{Nm}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Maximum specific DIN uptake rate by host\strut
@@
 mol N C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(0.035\)\strut
@@
 \(K_N\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Half-saturation constant for DIN uptake by host\strut
@@
 mol N L\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(1.5\times 10^{-6}\)\strut
@@
 \(k_{CO_2}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Efficacy of CO\textsubscript{2} delivery by host CCMs\strut
@@
 mol CO\textsubscript{2} mol C\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(10\)\strut
@@
 \(j_{HGm}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Maximum specific growth rate of host\strut
@@
 C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(1\)\strut
@@
 \(y_{CL}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Quantum yield of photosynthesis\strut
 \end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-mol C mol q\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+mol C mol photons\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(0.1\)\strut
@@
 \(\bar{a}^*\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Effective light-absorbing cross-section of symbiont\strut
@@
 m\textsuperscript{2} C-mol S\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(1.34\)\strut
@@
 \(k_{NPQ}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 NPQ capacity of symbiont\strut
 \end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-mol ph C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(112\)\strut
@@
 \(k_{ROS}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Excess photon energy that doubles ROS production, relative to baseline
@@
 \end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-mol ph C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(80\)\strut
@@
 \(j_{CPm}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Maximum specific photosynthesis rate of symbiont\strut
@@
 mol C C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(2.8\)\strut
@@
 \(j_{SGm}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Maximum specific growth rate of symbiont\strut
@@
 C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(0.25\)\strut
@@
 \(b\)\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright\strut
-Scaling parameter for strength of bleaching response\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Scaling parameter for bleaching response\strut
 \end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
 --\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
 \(5\)\strut
@@
 
-Additionally, the coral animal acquires nitrogen dissolved in the
-surrounding seawater, which is assumed to represent ammonium, the
-primary form utilized by corals (Yellowlees, Rees, and Leggat 2008).
-This gives the host (rather than the symbiont) priority in nitrogen
-utilization; this capacity is supported by experimental evidence (Wang
-and Douglas 1998) and is consistent with the spatial arrangment of the
-partners, where the host is in direct contact with the external
-environment. The uptake of nitrogen from the environment is thus
-specified by Michaelis-Menten kinetics using a maximum uptake rate
+Additionally, the coral animal acquires dissolved inorganic nitrogen
+(DIN) from the surrounding seawater, which is assumed to represent
+ammonium, the primary form utilized by corals (Yellowlees, Rees, and
+Leggat 2008). This gives the host (rather than the symbiont) priority in
+nitrogen utilization; this capacity is supported by experimental
+evidence (Wang and Douglas 1998) and is consistent with the spatial
+arrangment of the partners, where the host is in direct contact with the
+external environment. The uptake of nitrogen from the environment is
+thus specified by Michaelis-Menten kinetics using a maximum uptake rate
 \(j_{Nm}\) and half-saturation constant \(K_N\):
@@
 
-where \(\rho_C\) is fixed carbon shared by the symbiont (see Eq. 20),
+where \(\rho_C\) is fixed carbon shared by the symbiont (see Eq. 21),
 and \(r_{NH}\) is recycled nitrogen liberated by host biomass turnover
-(see Eq. XX). Host biomass turnover is equal to the specific maintenance
-rate of host biomass, which is set as a parameter \(j_{HT}^0\),
+(see Eq. 7). Host biomass turnover is equal to the specific maintenance
+rate of host biomass,
 
@@
 sources: in addition to the CO\textsubscript{2} actively supplied by the
-host through its CCMs (\(j_{CO_2}\); Eq. 9), we assume a fixed
+host through its CCMs (\(j_{CO_2}\); Eq. 10), we assume a fixed
 proportion \(\sigma_{CH}\) of metabolic CO\textsubscript{2} generated by
@@
 \(c_{ROS}\) is the relative rate of reactive oxygen species production
-(see Eq. 17). Dividing the photosynthetic rate by \(c_{ROS}\) causes a
+(see Eq. 18). Dividing the photosynthetic rate by \(c_{ROS}\) causes a
 decline in response to photooxidative stress at high light levels, and
@@
 
-Carbon fixed by photosynthesis (\(j_{CP}\); Eq. 14) is then combined
-with nitrogen shared by the host (\(\rho_N\); Eq. 7) and nitrogen
-recycled from symbiont biomass turnover\footnote{Note that recycling of
-  symbiont biomass turnover (\(rNS\) and \(rCS\)) only occurs based on
-  the maintenance component of turnover (i.e., \(j_{ST}^0\)), and not
-  the photodamage/bleaching component, as this loss represents biomass
-  that is expelled from the host.}
+Carbon fixed by photosynthesis (\(j_{CP}\); Eq. 15) is then combined
+with nitrogen shared by the host (\(\rho_N\); Eq. 8) and nitrogen
+recycled from symbiont biomass turnover\footnote{See footnote 2.}
 
@@
 that impose biomass balance for host and symbiont and by a set of
-coupled non-linear algebraic equations (3-21) that define fluxes.
+coupled non-linear algebraic equations (3-22) that define fluxes.
 Several of these fluxes are defined \emph{implicitly}; for example, the
@@
 CO\textsubscript{2}-limitation of photosynthesis: if symbiont biomass is
-initially very low (i.e., a ``bleached'' coral), the system cannot
-escape this positive feedback and cannot grow (unless feeding is
-sufficiently high). However, if symbiont biomass is initially high
-(i.e., a ``healthy'' coral), then the system remains in a
-nitrogen-limited state with positive growth. For practical purposes,
-this section of the manuscript considers only positive growth steady
-states under constant environments; subsequently, we explore how
+initially very low (i.e., a ``bleached'' coral), very little carbon is
+fixed, and the system cannot escape this positive feedback and cannot
+grow (unless feeding is sufficiently high). However, if symbiont biomass
+is initially high (i.e., a ``healthy'' coral), then the system remains
+in a nitrogen-limited state with positive growth. For practical
+purposes, this section of the manuscript considers only positive growth
+steady states under constant environments; subsequently, we explore how
 environmental forcing may cause the system to switch between alternate
@@
 \includegraphics{../img/Fig2.png}
-\caption{Alternate stable states}
+\caption{Alternate stable states in S:H biomass and growth across a
+light gradient. Alternate stable states occur between
+\textasciitilde{}33-44 mol photons m\textsuperscript{-2}
+d\textsuperscript{-1} under these conditions (DIN=1e-7 molN
+L\textsuperscript{-1}; X=2e-6 molX L\textsuperscript{-1}), depending on
+whether initial S:H is high (0.1, closed circles), representing a
+healthy coral, or low (0.0001, open circles), representing a bleached
+coral.}
 \end{figure}
@@
 diaphana}, a non-calcifying symbiotic anemone (Armoza-Zvuloni et al.
-2014). It is not surprising that the model predicts slightly higher
-growth than is typically observed, since it does not account for
+2014). However, it is not surprising that observed growth rates are
+often lower than model predictions, since the model does not account for
 ecological factors that may limit growth (e.g., competition, predation,
@@
 
-As irradiance continues to increase above \textasciitilde{}25 mol
-photons m\textsuperscript{-2} d\textsuperscript{-1}, growth rates
+At irradiance levels above \textasciitilde{}25 mol photons
+m\textsuperscript{-2} d\textsuperscript{-1}, steady state growth rates
 decline until positive growth ceases above \textasciitilde{}40 µmol
@@
 (Muller-Parker 1987). Importantly, while light alone may drive seasonal
-dynamics in the ways discussed\footnote{at the light levels indicated in
-  Stimson (1997). Note that if light levels were reduced throughout the
-  year (e.g.~for a coral at greater depth) such that light did not cause
-  photo-oxidative stress in summer but became limiting to photosynthesis
-  in winter, the S:H ratio would still increase in winter, but growth
-  would decrease; the latter scenario is predicted both by the present
-  model as well as that of Muller et al. (2009), since photo-oxidative
-  stress does not become relevant.}, temperature fluctuations may
-attenuate or even reverse the effect of light as cooler winters depress
-metabolism; thus, the relative magnitude of fluctuation in temperature,
-light, and other factors may produce wide variability in the direction
-and magnitude of seasonal changes in growth and symbiont abundance,
-depending on location and microhabitat. Nevertheless, the seasonal
-variability predicted here (Fig. 5) is consistent with experimental and
-field observations for corals, and demonstrates the power of this model
-to predict dynamic behavior that mechanistically integrates multiple
-environmental drivers.
+dynamics in the ways discussed\footnote{See footnote 4.}, temperature
+fluctuations may attenuate or even reverse the effect of light as cooler
+winters depress metabolism; thus, the relative magnitude of fluctuation
+in temperature, light, and other factors may produce wide variability in
+the direction and magnitude of seasonal changes in growth and symbiont
+abundance, depending on location and microhabitat. Nevertheless, the
+seasonal variability predicted here (Fig. 5) is consistent with
+experimental and field observations for corals, and demonstrates the
+power of this model to predict dynamic behavior that mechanistically
+integrates multiple environmental drivers.
 
@@
 densities rapidly decline (F) into a bleached state. Substrate
-limitation coefficients were calculated using Eq. 22. All parameters
+limitation coefficients were calculated using Eq. 23. All parameters
 were set at default values with external DIN=1e-7 mol N/L and prey
@@
 re-establishing nitrogen-limitation. Substrate limitation coefficients
-were calculated using Eq. 22. All simulations were run with default
+were calculated using Eq. 23. All simulations were run with default
 parameters (except varying \(n_{NX}\)), with L=10 mol photons

### 2017-03-21 JTB_ms/JTB_ms.Rmd
update fig caps, yC effects

@@
 Furthermore, by including a mechanistic treatment of photo-oxidative stress, the model displays dynamics of bleaching and recovery that can be explained as transitions between alternate stable states.
-These dynamics reveal that 'healthy' and 'bleached' states correspond broadly to nitrogen- and carbon-limitation of the system, with transitions between them occurring as integrated responses to multiple environmental factors.
-Indeed, a suite of complex emergent behaviors reproduced by the model (e.g., bleaching is exacerbated by nutrients and attenuated by feeding) suggests it unifyingly captures many important attributes of the system; meanwhile, its modular framework and open source R code are designed to facilitate further problem-specific development.
+These dynamics reveal that 'healthy' and 'bleached' states correspond broadly to nitrogen- and carbon-limitation in the system, with transitions between them occurring as integrated responses to multiple environmental factors.
+Indeed, a suite of complex emergent behaviors reproduced by the model (e.g., bleaching is exacerbated by nutrients and attenuated by feeding) suggests it captures many important attributes of the system; meanwhile, its modular framework and open source R code are designed to facilitate further problem-specific development.
 We see significant potential for this modeling framework to generate testable hypotheses and predict integrated, mechanistic responses of corals to environmental change, with important implications for understanding the performance and maintenance of symbiotic systems.
@@
 | $j_L$     | Light absorption rate                  | mol photons C-mol S^-1^ d^-1^| 12 |
-| $r_{CH}$  | Recycled CO~2~ from host turnover      | mol CO~2~ C-mol H^-1^ d^-1^| 13 |
-| $r_{CS}$  | Recycled CO~2~ from symbiont turnover  | mol CO~2~ C-mol S^-1^ d^-1^| 14 |
+| $r_{CH}$  | Recycled CO~2~ from host               | mol CO~2~ C-mol H^-1^ d^-1^| 13 |
+| $r_{CS}$  | Recycled CO~2~ from symbiont           | mol CO~2~ C-mol S^-1^ d^-1^| 14 |
 | $j_{CP}$  | Photosynthesis rate                    | mol C C-mol S^-1^ d^-1^ | 15 |
@@
 | $\sigma_{CS}$ | Proportion symbiont C-turnover recycled to photosynthesis | -- | $`r defpars$sigmaCS`$ |
-| $j_{Xm}$ | Maximum specific feeding rate of host | molX C-mol H^-1^ d^-1^ | $`r defpars$jXm`$ |
+| $j_{Xm}$ | Maximum specific assimilation rate from host feeding | molX C-mol H^-1^ d^-1^ | $`r defpars$jXm`$ |
 | $K_X$ | Half-saturation constant for prey uptake by host | molX L^-1^ | $`r defpars$KX`$ |
@@
 | $y_{CL}$ | Quantum yield of photosynthesis | mol C mol photons^-1^ | $`r defpars$yCL`$ |
+| $y_{C}$ | Yield of biomass formation from carbon | C-mol mol C^-1^ | $`r defpars$yC`$ |
 | $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | m^2^ C-mol S^-1^ | $`r defpars$astar`$ |
@@
 $c_{ROS}$ is a proportional rate that impacts other model fluxes by inhibition or acceleration; likewise, $j_{eC}$ accelerates the rate of $j_{CO_2}$.
-Recycled mass fluxes from biomass turnover are not shown for clarity (but see Table 1 for definitions).](../img/Fig1.png)
+Recycling fluxes are not shown for clarity (but see Table 1 for definitions).](../img/Fig1.png)
 
@@
 The coral animal acquires both carbon and nitrogen from feeding on prey from the environment.
-Prey acquisition is specified by Michaelis-Menten kinetics (i.e., a Holling type II function) with a maximum feeding rate $j_{Xm}$ and half-saturation constant $K_X$:
+Assimilation from feeding is specified by Michaelis-Menten kinetics (i.e., a Holling type II function) with a maximum rate of $j_{Xm}$ and half-saturation constant $K_X$:
 
@@
 
-\begin{equation} j_{HG} = \bigg({1 \over j_{HGm}} + {1 \over {\rho_C{S \over H} + j_X}} + {1 \over {(j_N + n_{NX}j_X + r_{NH}) / n_{NH}}} - {1 \over {{\rho_C{S \over H} + j_X} + (j_N + n_{NX}j_X + r_{NH}) / n_{NH}}}\bigg)^{-1} \end{equation}
+\begin{equation} j_{HG} = \bigg({1 \over j_{HGm}} + {1 \over {y_{C}\rho_C{S \over H} + j_X}} + {1 \over {(j_N + n_{NX}j_X + r_{NH}) / n_{NH}}} - {1 \over {y_{C}{\rho_C{S \over H} + j_X} + (j_N + n_{NX}j_X + r_{NH}) / n_{NH}}}\bigg)^{-1} \end{equation}
 
@@
 7).
+The parameter $y_C$ specifies the yield of biomass from organic carbon, which we take to be 0.8 for redox balance (see @Muller:2009io).
 Host biomass turnover is equal to the specific maintenance rate of host biomass,
@@
 
-\begin{equation} \rho_N = (j_N + n_{NX}j_X + r_{NH} - n_{NH}j_{HG})_+ \end{equation}
+\begin{equation} \rho_N = (j_N + n_{NX}j_X + r_{NH} - n_{NH}j_{HG}y_{C}^{-1})_+ \end{equation}
 
@@
 
-\begin{equation} j_{eC} = (jX + \rho_C{S \over H} - j_{HG})_+ \end{equation}
+\begin{equation} j_{eC} = (j_X + \rho_C{S \over H} - j_{HG}y_{C}^{-1})_+ \end{equation}
 
@@
 CO~2~ arrives at the photosynthesis SU from multiple sources: in addition to the CO~2~ actively supplied by the host through its CCMs ($j_{CO_2}$; Eq.
-10), we assume a fixed proportion $\sigma_{CH}$ of metabolic CO~2~ generated by host biomass turnover is passively available to the photosynthesis SU, according to:
+10), we assume a fixed proportion $\sigma_{CH}$ of metabolic CO~2~ generated by the host from both biomass turnover and formation is passively available to the photosynthesis SU, according to:
 
-\begin{equation} r_{CH}=\sigma_{CH}j_{HT} \end{equation}
+\begin{equation} r_{CH}=\sigma_{CH}(j_{HT} + (1-y_C)j_{HG}/y_C) \end{equation}
 
-along with a fixed proportion of CO~2~ generated by symbiont biomass turnover[^2]:
+along with a fixed proportion of CO~2~ generated by symbiont biomass turnover[^2] and formation:
 
-\begin{equation} r_{CS}=\sigma_{CS}j_{ST}^0 \end{equation}
+\begin{equation} r_{CS}=\sigma_{CS}(j_{ST}^0 + (1-y_C)j_{SG}/y_C) \end{equation}
 
@@
 Importantly, $c_{ROS}$ is specified here not as a function of absolute external light, but rather the amount of excess light energy after accounting for quenching by carbon fixation and NPQ.
-A direct consequence of this formulation is that CO~2~-limitation of photosynthesis can lead to ROS production, an important mechanism [@Wooldridge:2009p7807] that was not captured by previous representations of photooxidative stress [@Eynaud:2011tv].
+A direct consequence of this formulation is that CO~2~-limitation of photosynthesis can lead to ROS production, an important mechanism [@Butow:1998gz; @Wooldridge:2009p7807] that was not captured by previous representations of photooxidative stress [@Eynaud:2011tv].
 With this single SU, both the light and dark reactions of photosynthesis are represented, allowing for sink-limitation (i.e., CO~2~-limitation) to cause overreduction of the electron transport chain and ROS production.
@@
 
-\begin{equation} j_{SG} = \bigg({1 \over j_{SGm}} + {1 \over j_{CP}} + {1 \over {(\rho_N{H \over S} + r_{NS}) / n_{NS}}} - {1 \over {j_{CP} + (\rho_N{H \over S} + r_{NS}) / n_{NS}}}\bigg)^{-1} \end{equation}
+\begin{equation} j_{SG} = \bigg({1 \over j_{SGm}} + {1 \over y_{C}j_{CP}} + {1 \over {(\rho_N{H \over S} + r_{NS}) / n_{NS}}} - {1 \over {y_{C}j_{CP} + (\rho_N{H \over S} + r_{NS}) / n_{NS}}}\bigg)^{-1} \end{equation}
 
@@
 
-\begin{equation} \rho_C = (j_{CP} - j_{SG})_+ \end{equation}
+\begin{equation} \rho_C = (j_{CP} - j_{SG}y_{C}^{-1})_+ \end{equation}
 
@@
 where $j_{S1}$ and $j_{S2}$ are the specific input fluxes of the two substrates and $j_{Pm}$ is the maximum specific product formation rate, in units of Cmol Cmol^-1^ d^-1^.
+When both substrate input fluxes are higher than what can be used at the maximum production rate, this limitation coefficient is zero, implying that neither substrate is limiting production.
 
@@
 ![Alternate stable states in S:H biomass and growth across a light gradient.
-Alternate stable states occur between ~33-44 mol photons m^-2^ d^-1^ under these conditions (DIN=1e-7 molN L^-1^; X=2e-6 molX L^-1^), depending on whether initial S:H is high (0.1, closed circles), representing a healthy coral, or low (0.0001, open circles), representing a bleached coral.](../img/Fig2.png)
+Alternate stable states occur between ~25-42 mol photons m^-2^ d^-1^ under these conditions (DIN=1e-7 molN L^-1^; X=1e-7 CmolX L^-1^), depending on whether initial S:H is high (1, closed circles), representing a healthy coral, or low (0.0001, open circles), representing a bleached coral.
+Arrow above point at L=25 indicates a S:H ratio beyond the axis range; this 'overshoot' phenomenon, in which initially  bleached corals may achieve high S:H ratios while remaining in a carbon-limited state is discussed in the *Coral Bleaching and Recovery* section.](../img/Fig2.png)
 
@@
 
-![Steady state values of (A) specific growth (Cmol Cmol^-1^ d^-1^) and (B) the symbiont to host biomass ratio (C-mol S C-mol H^-1^) across gradients of external irradiance and dissolved inorganic nitrogen.
+![Steady state values of (A) specific growth (C-mol H C-mol H^-1^ d^-1^) and (B) the symbiont to host biomass ratio (C-mol S C-mol H^-1^) across gradients of external irradiance and dissolved inorganic nitrogen.
 Note that typical conditions for reefs are ~1e-7 M DIN and 10-20 mol photons m^-2^ d^-1^.
-Simulations for each combination of light and nutrients (41 points along each axis) were run for 100 days with a time step of 1 day.
-All parameters were set to default values and prey density set to zero.
+Simulations for each combination of light and nutrients (41 points along each axis) were run to steady state with all parameters at default values and prey density set to zero.
 Negative steady state growth rates, and corresponding S:H ratios, were set to zero.](../img/Fig3.png)
@@
 Plots show the fractional change in steady state values of growth (solid lines) and S:H biomass (dashed lines) in response to fractional changes in default parameter values (see Table 2 for default values).
-Parameters are grouped by which processes they are involved in.
-This sensitivity analysis was conducted at conditions typical for coral reef environments: low DIN (1e-7 M) and intermediate light (15 mol m^-2^ d^-2^).
+Parameters are grouped by the processes in which they are involved.
+This sensitivity analysis was conducted at conditions typical for coral reef environments: low DIN (1e-7 M) and intermediate light (15 mol photons m^-2^ d^-2^), with prey density set to zero.
 Sensitivity analyses conducted other environmental conditions are presented in Figs S3-S7.](../img/Fig4.png)
@@
 ![Light-driven seasonal dynamics of symbiont abundance and coral growth.
-Light input (upper panel) was designed as a sinusoidal curve with a period of one year, with maximum and minimum values of 44 and 22 mol photons m^-2^ d^-1^, based on @Stimson:1997p3837.
-The dynamic behavior of symbiont to host biomass ratio (middle panel) and the specific growth rate of host biomass (lower panel) show seasonal oscillations that are greater in magnitude under higher nutrients (1e-5 mol N L^-1^; dashed lines) relative to lower nutrients (1e-7 mol N L^-1^; solid lines), consistent with the findings of @Stimson:1997p3837.
-Prey density was set at 1e-5 CmolX L^-1^.](../img/Fig5.png)
+Light input (upper panel) was designed as a sinusoidal curve with a period of one year, with maximum and minimum values of 44 and 20 mol photons m^-2^ d^-1^, corresponding to those measured by @Stimson:1997p3837.
+The dynamic behavior of symbiont to host biomass ratio (middle panel) and the specific growth rate of host biomass (lower panel) show seasonal oscillations that are greater in magnitude under high nutrients (15.14 $\mu$M N; dashed lines) relative to low nutrients (0.14 $\mu$M N; solid lines), consistent with the findings of @Stimson:1997p3837.
+Prey density was set at 1e-6 CmolX L^-1^.](../img/Fig5.png)
 
@@
 23.
-All parameters were set at default values with external DIN=1e-7 mol N/L and prey density=0).](../img/Fig6.png)
+All parameters were set at default values with external DIN=1e-7 mol N/L and prey density set to zero).](../img/Fig6.png)
 
@@
 ![Bleaching with interactive factors.
-Simulations of high light stress (sinusoid with maximum=48 mol ph m^-2^ d^-1^) under default environmental conditions (solid line; DIN=1e-7 mol N L^-1^; prey=3e-6 CmolX L^-1^), or with elevated feeding (dashed line; prey=6e-6 CmolX L^-1^), or elevated nutrients (dotted line; DIN=1e-6 mol N L^-1^).
-Initial symbiont biomass was set to the steady state for each set of starting conditions, respectively.
-(All parameters set to default values).](../img/Fig7.png)
+Simulations of high light stress (sinusoid with maximum=48 mol ph m^-2^ d^-1^) under default environmental conditions (solid line; DIN=1e-7 mol N L^-1^; prey=2e-7 C-molX L^-1^), or with elevated feeding (dashed line; prey=1e-6 C-molX L^-1^), or elevated nutrients (dotted line; DIN=4e-6 mol N L^-1^).
+Initial symbiont biomass was set to the steady state for each set of starting conditions, with all other parameters at default values.](../img/Fig7.png)
 
@@
 
-![Recovery from bleaching with varying N:C availability (A: N:C=0.09; B: N:C=0.10; C: N:C=0.30).
+![Recovery from bleaching with varying N:C availability (A: N:C=0.100; B: N:C=0.175; C: N:C=0.220).
 Higher N:C ratios in a heterotrophic food source (effectively representing higher external DIN and/or lower feeding rates) cause a larger overshoot in the S:H biomass ratio, and prolong the duration of time until the system 'recovers' by re-establishing nitrogen-limitation.
@@
 23.
-All simulations were run with default parameters (except varying $n_{NX}$), with L=10 mol photons m^2^ d^-1^, DIN=1e-7 mol N L^-1^, prey=1e-6 CmolX L^-1^, and initial S:H biomass=0.001.](../img/Fig8.png)
+All simulations were run with default parameters (except varying $n_{NX}$), with L=20 mol photons m^2^ d^-1^, DIN=1e-7 mol N L^-1^, prey=1e-7 CmolX L^-1^, and initial S:H biomass=0.001.](../img/Fig8.png)
 

### 2017-03-21 JTB_supp/JTB_supp.Rmd

@@
 #### $j_{HT}^0$: Specific turnover rate of host biomass {-}
-@Jacobson:2016bk parameterized a DEB model for coral symbiosis, and reported a specific maintenance rate of 0.03 CmolH CmolH^-1^ d^-1^.
-Although this value is per unit structural biomass (in the DEB sense) rather than per unit total biomass (as needed here), use of this value generates plausible values of specific growth (see main text).
-```{r jHT, echo=F}
+@Jacobson:2016bk parameterized a DEB model for coral symbiosis, and reported a specific maintenance rate of 0.03 mol C~E~ mol C~V~^-1^ d^-1^, where subscripts E and V refer to reserve and structural biomass.
+We assume (simplistically) the same rate per unit of total biomass:
+```{r jHT, echo=T}
 0.03
@@
 #### $j_{ST}^0$: Specific turnover rate of symbiont biomass {-}
-Assuming an equivalent maintenance rate as the host:
-```{r jST, echo=F}
+Values of *Symbiodinium* and other dinoflagellate respiration rates may provide a reasonable approximation of the specific turnover rate of symbiont biomass, yet values calculated from various studies range widely from 0.01 to nearly 0.2.
+Without data for a more confident estimation of this parameter, we choose a value at the low end of this range, since 0.2 seems unrealistic and many rates may be depressed in symbiotic dinoflagellates relative to their free-living counterparts.
+```{r jST, echo=T}
 0.03
@@
 ```{r sigmaNHsigmaNS, echo=F}
-c(0.03, 0.03)
+c(0.9, 0.9)
 ```
@@
 ```{r sigmaCH, echo=F}
-0.10
+0.1
 ```
@@
 #### $j_{Xm}$: Maximum specific feeding rate of host {-}
-@Wijgerde:2011du measured depletion of carbon from captured prey in *Galaxea fascicularis* (including both intracoelenteric and extracoelenteric feeding), reporting a maximum specific feeding rate of 77.6 µgC cm^-2^ 6h^-1^.
+@Wijgerde:2011du measured depletion of carbon from captured prey in *Galaxea fascicularis* (including both intracoelenteric and extracoelenteric feeding), reporting a maximum specific carbon assimilation rate of 77.6 µgC cm^-2^ 6h^-1^.
 ```{r jXm}
@@
 
+@Osinga:2011di measured a maximum feeding rate of 
+half sat = 294119 nauplii L^-1^
+294119 artemia/L * 6.78e-7 gC/artemia = 0.19941268 gC/L / 12.011g/mol = 0.01656815
+
 #### $K_X$: Half-saturation constant for prey uptake by host {-}
-@Purser:2010hp measured feeding rates of the coral *Lophelia pertusa* at different concentrations of *Artemia* nauplii, reporting a half-saturation for feeding of ~345 *Artemia* L^-1^.
+For consistency, and due to a lack of available data, we use the same value as @Muller:2009io.
 ```{r KX}
-345 * gC.artemia / gC.molC
+1e-6
 ```
@@
 #### $j_{SGm}$: Maximum specific growth rate of symbiont {-}
-@Suggett:2015dv measured specific growth rates of 5 clades of *Symbiodinium* in culture, and reported values ranging from 0.19 to 0.26 d^-1^ (Table 2).  
+We use the maximum specific growth rate of *Symbiodinium* that is reported by @Chang:1983p8066:
 ```{r jSGm, echo=F}
@@
 Larger values of this parameter cause greater biomass loss in response to ROS signaling (i.e., a more severe bleaching response).
-Since this represents symbiont cells being expelled from the system, we use a value about an order of magnitude higher than the turnover rate due to maintenance.
+Since this represents symbionts being expelled from the system, we use a value about two orders of magnitude higher than the turnover rate due to maintenance.
 ```{r b, echo=F}
@@
 # Initialize environments
-env1 <- init_env(time=time, L=c(20,44,2), N=c(0.14e-6,0.14e-6,0), X=c(1e-5,1e-5,0))
-env2 <- init_env(time=time, L=c(20,44,2), N=c(10.14e-6,10.14e-6,0), X=c(1e-5,1e-5,0))
+env1 <- init_env(time=time, L=c(20,44,2), N=c(0.14e-6,0.14e-6,0), X=c(1e-6,1e-6,0))
+env2 <- init_env(time=time, L=c(20,44,2), N=c(15.14e-6,15.14e-6,0), X=c(1e-6,1e-6,0))
 
 # Run simulations
-ss1 <- with(run_coral_ss(env=list(L=33,N=0.14e-6,X=1e-5), pars=defpars, dt=0.1), last(S$S/H$H))
+ss1 <- with(run_coral_ss(env=list(L=32,N=0.14e-6,X=1e-6), pars=defpars, dt=0.1), last(S$S/H$H))
 run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", ss1))
-ss2 <- with(run_coral_ss(env=list(L=33,N=10.14e-6,X=1e-5), pars=defpars, dt=0.1), last(S$S/H$H))
+ss2 <- with(run_coral_ss(env=list(L=32,N=15.14e-6,X=1e-6), pars=defpars, dt=0.1), last(S$S/H$H))
 run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", ss2))
@@
 
-with(run1, {
-      # External irradiance
-    plot(time, env$L, type="l", col="gold", ylim=c(0,60), lwd=2, xlab="", ylab="mol ph/m2/d")
-    title("External irradiance and DIN", adj=0, line=0.25)
-    par(new=T)
-    plot(time, env$N, type="l", col="blue", ylim=c(0,10.14e-6), lwd=2, axes=F, xlab="", ylab="")
-    axis(side=4); mtext(side=4, text="molN/L", line=1, cex=0.75)
-
-    # Symbiont to host biomass ratio
-    totSH <- S$S / H$H
-    totSHf <- totSH[length(totSH)]
-    plot(time, totSH, type="l", col="black", ylim=c(0, 0.25), ylab="CmolS/CmolH", xlab="", lwd=2)
-    text(time[0.95*length(time)], totSHf, labels=as.character(round(totSHf, 3)), pos=3, xpd=T, cex=0.75)
-    title("Symbiont:host biomass", adj=0, line=0.25)
-
-    # Specific growth rates of host and symbiont
-    Hgrf <- H$dH.Hdt[length(H$dH.Hdt)]
-    plot(time, H$dH.Hdt, type="l", ylim=c(-0.01, 0.13), 
-         xlab="", ylab="d-1", lwd=2, cex=1, cex.lab=1)
-    if(any(c(H$dH.Hdt[-(1:10)], S$dS.Sdt[-(1:10)])<0)) abline(h=0, lwd=0.5, lty=3)
-    text(time[0.95*length(time)], Hgrf, labels=as.character(round(Hgrf, 3)), pos=3, xpd=T, cex=0.75)
-    #title("State variable dynamics", adj=0, cex.main=2, outer = T)
-    title("Specific growth rates", adj=0, line=0.25)
-    lines(time, S$dS.Sdt, col="black", lwd=1)
-    legend("topright", legend=c("Host", "Sym"), lwd=c(2,1), col="black", bty="n", y.intersp=1, cex=0.75)
-
-    # Light quenching (carbon fixation, NPQ, and excess (=ROS producing))
-    plot(NA, xlim=range(time), xlab="", ylim=c(0,130), ylab="mol ph/CmolS/d")
-    title("Light quenching", adj=0, line=0.25)
-    lines(time, S$jCP/pars$yCL, col="gold", lwd=1) # amt. quenched by photosynthesis
-    lines(time, S$jCP/pars$yCL + S$jNPQ, col="gold", lty=1, lwd=2) # amt. quenched by photo + NPQ
-    lines(time, S$jL, col="gold", lty=3, lwd=2) # total amt. absorbed
-    legend("topright", legend=c("ROS", "NPQ", "Photo."), lty=c(3,1,1), lwd=c(2,2,1), col="gold", bty="n", cex=0.75)
-
-    # Photosynthesis: substrate-limitation
-    pl <- log(   pmin((H$jCO2 + H$rCH)*H$H/S$S + S$rCS, pars$jCPm)   /   pmin(S$jL * pars$yCL, pars$jCPm)    )
-    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-2, 2))
-    title("Photosynthesis: substrate-limitation", adj=0, line=0.25)
-    lines(time, pl, col="black", lty=1, lwd=2)
-    abline(h=0, lty=3)
-    text(par("usr")[2], 0, labels="L-lim.\nCO2-lim.", cex=0.75, adj=1)
-
-    # Biomass formation - substrate-limitation
-    sl <- log(  pmin(S$jCP, pars$jSGm)   /   pmin((H$rhoN*H$H/S$S + S$rNS)/pars$nNS, pars$jSGm)  )
-    hl <- log(  pmin(S$rhoC*S$S/H$H + H$jX, pars$jHGm) / pmin((H$jN + pars$nNX*H$jX + H$rNH) / pars$nNH, pars$jHGm)  )
-    maxabs <- max(c(abs(hl), abs(sl)))
-    range <- range(c(hl, sl))
-    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-1.5,1.5))
-    title("Biomass: substrate-limitation", adj=0, line=0.25)
-    lines(time, hl, col="black", lty=1, lwd=2)
-    lines(time, sl, col="black", lty=1, lwd=1)
-    abline(h=0, lty=3)
-    text(par("usr")[2], 0, labels="N-lim.\nC-lim.", cex=0.75, adj=1)
-    legend("topright", legend=c("Host", "Sym"), lty=1, lwd=c(2,1), col="black", bty="n", cex=0.75)
-})
-
-with(run2, {
-      # External irradiance
-    plot(time, env$L, type="l", col="gold", ylim=c(0,60), lwd=2, xlab="", ylab="mol/m2/d")
-    title("External irradiance and DIN", adj=0, line=0.25)
-    par(new=T)
-    plot(time, env$N, type="l", col="blue", ylim=c(0,10.14e-6), lwd=2, axes=F, xlab="", ylab="")
-    axis(side=4); mtext(side=4, text="molN/L", line=1, cex=0.75)
-
-    # Symbiont to host biomass ratio
-    totSH <- S$S / H$H
-    totSHf <- totSH[length(totSH)]
-    plot(time, totSH, type="l", col="black", ylim=c(0, 0.25), ylab="CmolS/CmolH", xlab="", lwd=2)
-    text(time[0.95*length(time)], totSHf, labels=as.character(round(totSHf, 3)), pos=3, xpd=T, cex=0.75)
-    title("Symbiont:host biomass", adj=0, line=0.25)
-
-    # Specific growth rates of host and symbiont
-    Hgrf <- H$dH.Hdt[length(H$dH.Hdt)]
-    plot(time, H$dH.Hdt, type="l", ylim=c(-0.01,0.13), 
-         xlab="", ylab="d-1", lwd=2, cex=1, cex.lab=1)
-    if(any(c(H$dH.Hdt[-(1:10)], S$dS.Sdt[-(1:10)])<0)) abline(h=0, lwd=0.5, lty=3)
-    text(time[0.95*length(time)], Hgrf, labels=as.character(round(Hgrf, 3)), pos=3, xpd=T, cex=0.75)
-    #title("State variable dynamics", adj=0, cex.main=2, outer = T)
-    title("Specific growth rates", adj=0, line=0.25)
-    lines(time, S$dS.Sdt, col="black", lwd=1)
-    legend("topright", legend=c("Host", "Sym"), lwd=c(2,1), col="black", bty="n", y.intersp=1, cex=0.75)
-
-    # Light quenching (carbon fixation, NPQ, and excess (=ROS producing))
-    plot(NA, xlim=range(time), xlab="", ylim=c(0,130), ylab="mol ph/CmolS/d")
-    title("Light quenching", adj=0, line=0.25)
-    lines(time, S$jCP/pars$yCL, col="gold", lwd=1) # amt. quenched by photosynthesis
-    lines(time, S$jCP/pars$yCL + S$jNPQ, col="gold", lty=1, lwd=2) # amt. quenched by photo + NPQ
-    lines(time, S$jL, col="gold", lty=3, lwd=2) # total amt. absorbed
-    legend("topright", legend=c("ROS", "NPQ", "Photo."), lty=c(3,1,1), lwd=c(2,2,1), col="gold", bty="n", cex=0.75)
-
-    # Photosynthesis: substrate-limitation
-    pl <- log(   pmin((H$jCO2 + H$rCH)*H$H/S$S + S$rCS, pars$jCPm)   /   pmin(S$jL * pars$yCL, pars$jCPm)    )
-    maxabs <- max(abs(pl))
-    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-2, 2))
-    title("Photosynthesis: substrate-limitation", adj=0, line=0.25)
-    lines(time, pl, col="black", lty=1, lwd=2)
-    abline(h=0, lty=3)
-    text(par("usr")[2], 0, labels="L-lim.\nCO2-lim.", cex=0.75, adj=1)
-
-    # Biomass formation - substrate-limitation
-    sl <- log(  pmin(S$jCP, pars$jSGm)   /   pmin((H$rhoN*H$H/S$S + S$rNS)/pars$nNS, pars$jSGm)  )
-    hl <- log(  pmin(S$rhoC*S$S/H$H + H$jX, pars$jHGm) / pmin((H$jN + pars$nNX*H$jX + H$rNH) / pars$nNH, pars$jHGm)  )
-    maxabs <- max(c(abs(hl), abs(sl)))
-    range <- range(c(hl, sl))
-    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-1.5,1.5))
-    title("Biomass: substrate-limitation", adj=0, line=0.25)
-    lines(time, hl, col="black", lty=1, lwd=2)
-    lines(time, sl, col="black", lty=1, lwd=1)
-    abline(h=0, lty=3)
-    text(par("usr")[2], 0, labels="N-lim.\nC-lim.", cex=0.75, adj=1)
-    legend("topright", legend=c("Host", "Sym"), lty=1, lwd=c(2,1), col="black", bty="n", cex=0.75)
-})
+plot_run_v2(run1)
+plot_run_v2(run2)
 ```
@@
 2 in the main text.
-These figures show steady state S:H biomass ratios in constant environments with varying DIN and food (X) levels, with initial S:H biomass values of either 0.1 or 0.0001.
+These figures show steady state S:H biomass ratios in constant environments with varying DIN and food (X) levels, with initial S:H biomass values of either 1 or 0.0001.
 Points are colored according to whether the host exhibits positive (black) or negative (red) growth at steady state."}
@@
 # Set input values for steady state runs
-input <- expand.grid(L=seq(1,60,1), initS=c(0.0001,0.1))
+input <- expand.grid(L=seq(1,60,1), initS=c(0.0001,1))
 
@@
 output1 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
-  run <- run_coral_ss(env=list(L=input$L[i], X=2e-6, N=1e-7), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+  run <- run_coral_ss(env=list(L=input$L[i], X=5e-7, N=1e-7), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
   list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
@@
 output2 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
-  run <- run_coral_ss(env=list(L=input$L[i], X=1e-6, N=1e-7), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+  run <- run_coral_ss(env=list(L=input$L[i], X=1e-7, N=1e-7), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
   list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
@@
 output3 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
-  run <- run_coral_ss(env=list(L=input$L[i], X=2e-6, N=1e-6), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+  run <- run_coral_ss(env=list(L=input$L[i], X=5e-7, N=2e-6), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
   list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
@@
 output4 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
-  run <- run_coral_ss(env=list(L=input$L[i], X=4e-6, N=1e-6), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+  run <- run_coral_ss(env=list(L=input$L[i], X=1e-6, N=2e-6), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
   list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
@@
                  col=ifelse(gr>0, "black", "red"), 
-                 pch=ifelse(initS==1e-01, 19, 1), 
-                 cex=ifelse(initS==1e-01, 0.4, 1)))
+                 pch=ifelse(initS==1, 19, 1), 
+                 cex=ifelse(initS==1, 0.4, 1)))
   with(x, if (sh > 0.5) {
@@
            col=ifelse(gr>0, "black", "red"),
-           pch=ifelse(initS==1e-01, 19, 1), 
-           cex=ifelse(initS==1e-01, 0.4, 1))
+           pch=ifelse(initS==1, 19, 1), 
+           cex=ifelse(initS==1, 0.4, 1))
   })

### 2017-03-21 JTB_supp/JTB_supp.Rmd
adj wording for feeding

@@
 
-#### $j_{Xm}$: Maximum specific feeding rate of host {-}
+#### $j_{Xm}$: Maximum specific rate of assimilation from host feeding {-}
 @Wijgerde:2011du measured depletion of carbon from captured prey in *Galaxea fascicularis* (including both intracoelenteric and extracoelenteric feeding), reporting a maximum specific carbon assimilation rate of 77.6 µgC cm^-2^ 6h^-1^.
@@
 
-#### $K_X$: Half-saturation constant for prey uptake by host {-}
+#### $K_X$: Half-saturation constant of assimilation from host feeding {-}
 For consistency, and due to a lack of available data, we use the same value as @Muller:2009io.

### 2017-03-21 R/Fig3.R
write to img/Fig3.png

@@ -13,4 +13,4 @@ at <- expand.grid(L=seq(from=0, to=50, length.out=41),
 
 # Run to steady state and plot results
 ssdat <- run_steady_states(pars=replace(defpars, "initS", 1), at=at, food=0, runtime=T)
-plot_steady_states(ssdat, png=NULL)#"img/Fig3.png")
+plot_steady_states(ssdat, png="img/Fig3.png")

### 2017-03-21 JTB_ms/JTB_ms.Rmd
text revisions

@@
 This formulation precludes modeling the full life cycle of corals as reproduction, larval stages, and metamorphosis are not considered, but greatly reduces theoretical complexity and parameter numbers, which is advantageous given the relative paucity of data for corals.
-Moreover, dynamics of the symbiosis (i.e., changes in symbiont to host biomass, including bleaching and recovery) and coral biomass growth are efficiently captured with this simpler formulation, which also increases accessibility for biologists and ecologists without requiring significant expertise in DEB theory.
+Moreover, dynamics of the symbiosis (i.e., changes in symbiont to host biomass, including bleaching and recovery) and coral biomass growth are efficiently captured with this simpler formulation, which also increases accessibility for biologists and ecologists without requiring expertise in DEB theory.
 
 3.
-_Provide well-documented, open-access code._ In order to facilitate the continued development and application of theoretical modeling tools for coral symbioses, we provide open access to the model in the form of detailed, commented code written in the R language [@RALanguageandEn:2014wf].
+_Provide well-documented, open source code._ In order to facilitate the continued development and application of theoretical modeling tools for coral symbioses, we provide open access to the model in the form of detailed, commented code written in the R language [@RALanguageandEn:2014wf].
 With an accessible and modular framework, we envision this model as a resource for further development by the scientific community to include additional complexity and problem-specific components.
-We chose the R language, which is freely available and in common use by biologists and ecologists, to reach the widest possible audience with this work.
+We chose R, an open source programming language in common use by biologists and ecologists, to reach the widest possible audience with this work.
 
@@
 In reality, light and temperature interact synergistically [@Coles:1978p1124; @Jones:1998p3572], and in fact, any stressor that disrupts the quenching of light energy may lead to bleaching [@Wooldridge:2010p7809; @Baker:2015kp].
-This is because the proximate cause of photo-oxidative stress is excess excitation energy, but the upstream events that lead to this situation may be diverse: indeed, elevated temperature may inhibit Rubisco functioning [@Jones:1998p3572] and the repair of the D1 protein in photosystem II [@Warner:1999p4239], which reduces the capacity of photochemical quenching and leads to an excess of light energy.
+This is because the proximate cause of photo-oxidative stress is excess excitation energy, but the upstream events that lead to this situation may be diverse.
+Indeed, elevated temperature may inhibit Rubisco functioning [@Jones:1998p3572] and the repair of the D1 protein in photosystem II [@Warner:1999p4239], which reduces the capacity of photochemical quenching and leads to an excess of light energy.
 In this way, elevated temperature serves to reduce the threshold above which light stresses the system [@HoeghGuldberg:1999p4222]; importantly, light is still the proximate stressor.
@@
 7).
-The parameter $y_C$ specifies the yield of biomass from organic carbon, which we take to be 0.8 for redox balance (see @Muller:2009io).
+The parameter $y_C$ specifies the yield of biomass from organic carbon, which we take to be 0.8 to satisfy redox balance (see @Muller:2009io).
+
 Host biomass turnover is equal to the specific maintenance rate of host biomass,
@@
 
-This flux, $j_{eC}$, is assumed to be available to the host as a respiratory substrate to support various energetically-demanding processes; of particular importance is the host's active carbon concentrating mechanisms (CCMs) that supply CO~2~ for symbiont photosynthesis [@Wooldridge:2013tj; @Hopkinson:2015ev].
+This flux, $j_{eC}$, is assumed to be available to the host as a respiratory substrate to support energetically-demanding processes; of particular importance is the host's active carbon concentrating mechanisms (CCMs) that supply CO~2~ for symbiont photosynthesis [@Wooldridge:2013tj; @Hopkinson:2015ev].
 We therefore specify $j_{CO_2}$ as the host-mediated delivery of CO~2~ to photosynthesis that encompasses potentially diverse carbon-concentrating mechanisms (CCMs), including active transport of bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to CO~2~ to promote diffusion toward the symbiont [@Tansik:2015cm], and acidification of the symbiosome to increase localized CO~2~ concentrations around the symbiont [@Barott:2014gx].
@@
 Simulations for each combination of light and nutrients (41 points along each axis) were run to steady state with all parameters at default values and prey density set to zero.
-Negative steady state growth rates, and corresponding S:H ratios, were set to zero.](../img/Fig3.png)
+Negative steady state growth rates and corresponding S:H ratios were set to zero, and a ceiling of 0.5 was imposed on S:H ratios to aid in visualization.](../img/Fig3.png)
 
@@
 
-The values used for each parameter in the model (Table 2) are derived from relevant literature; these derivations are described in the Supplementary Information.
+The values used for each parameter in the model (Table 2) are derived from relevant literature (see Supplementary Information).
 Here we evaluate the sensitivity of the model to changes in these parameter values, which also serves to demonstrate the behavior of the dynamical system.
@@
 Increasing the maximum growth and turnover rates have the expected effects of increasing and decreasing growth, respectively.
-Parameters relating to photooxidative stress and bleaching have little impact under these environmental conditions (Fig.
+Parameters relating to photooxidative stress and bleaching have little impact under low nutrients and intermediate light (Fig.
 4), but have larger impacts under higher light (e.g., Fig.
@@
 Using the levels of light and nutrients from this study as inputs, the model reproduces this observed interaction among environmental factors (Fig.
-5), and also provides the mechanism: as increasing light in summer decreases carbon fixation due to photooxidative stress, symbionts grow slower than the host, leading to decreasing S:H ratios.
-Under nutrient enrichment, this effect is more pronounced, as CO~2~-limitation (due to higher symbiont standing stocks) causes mild bleaching that results in a similar summertime minimum S:H as the ambient nutrient case ('physiological bleaching' *sensu* @Fitt:2001p832).
-Decreasing light into winter then alleviates the photooxidative stress constraints on carbon fixation, such that nitrogen constrains the S:H ratio, explaining why S:H increases more when DIN is enriched (Fig.
+5), and also provides the mechanism: increasing light in summer decreases symbiont growth rates due to photooxidative stress, leading to decreasing S:H ratios.
+Under nutrient enrichment, this effect is more pronounced, as CO~2~-limitation of photosynthesis (due to higher symbiont standing stocks) causes mild bleaching that results in a similar summertime minimum S:H as the ambient nutrient case ('physiological bleaching' *sensu* @Fitt:2001p832).
+Decreasing light into winter then alleviates the photooxidative stress constraints on carbon fixation such that nitrogen-limitation constrains the S:H ratio, explaining why S:H increases more when DIN is enriched (Fig.
 S8).
@@
 Nevertheless, the seasonal variability predicted here (Fig.
-5) is consistent with experimental and field observations for corals, and demonstrates the power of this model to predict dynamic behavior that mechanistically integrates multiple environmental drivers.
+5) is consistent with experimental and field observations for corals, and demonstrates the model's ability to predict dynamic behavior that mechanistically integrates multiple environmental drivers.
 
@@
 These dynamics reveal that the re-establishment of nitrogen-limitation is the most important diagnostic of recovery to a 'healthy' state, as this is when positive growth rates are resumed.
-A high S:H biomass ratio alone does not necessarily indicate that a 'healthy' state has been reached, since the C-limited state may still persist (e.g.
+A high S:H biomass ratio alone does not necessarily indicate that a 'healthy' state has been reached, since the carbon-limited state may still persist (e.g.
 Fig.
@@
 This could explain why corals that have recovered their symbiont populations after bleaching may still exhibit energetic deficits and physiological impacts, possibly for months to years [@Levitan:2014et, @Hughes:2013hd]).
-Therefore, host acquisition of carbon from a source other than the symbiont may be extremely important for the system to recover from bleaching.
+These findings suggest that host acquisition of carbon from a source other than the symbiont may be extremely important for the system to recover from bleaching.
 Indeed, host feeding has been shown to promote a more rapid return to pre-bleaching levels of key physiological parameters in recovering corals [@Connolly:2012ep].
-The model predicts that acquisition of carbon from other sources, such as direct uptake of DOC [@Levas:2015ic], may also promote more rapid recovery from a bleached state.
+Additional carbon sources for the host, such as direct uptake of dissolved organic carbon [@Levas:2015ic], may also promote more rapid recovery from a bleached state.
 
@@
 Thus, the diversity of potential applications for this model is high, and we envision this work as a foundation for continued development, which may include more detailed treatments of specific modules (e.g., DIC processing), and the incorporation of more external forcing capabilities (e.g., external DIC, temperature).
-Importantly, this effort may be undertaken by the scientific community at large, including those with theoretical and empirical backgrounds, which is why the code for this model has been developed openly using the R language.
+Importantly, open source R code allows this effort to benefit from contributions from the wider scientific community, including those with empirical and theoretical backgrounds.
 Ultimately, the continued refinement of these tools is fundamental in elucidating the mechanisms of symbiosis function and dysfunction, and in predicting coral responses to environmental change.

### 2017-03-21 JTB_supp/JTB_supp.Rmd
change light levels for sensitivity

@@
 # Figure S3 
-```{r sens_ILHN, cache=T, echo=F, eval=T, fig.cap="**Fig. S3. Sensitivity analysis under high DIN (4e-6 M) and intermediate light (15 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
+```{r sens_ILHN, cache=T, echo=F, eval=T, fig.cap="**Fig. S3. Sensitivity analysis under high DIN (4e-6 M) and intermediate light (20 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
 # Set up time vector and parameters
@@
 envs <- list(
-  ILHN=list(L=15, N=4e-6, X=0)
+  ILHN=list(L=20, N=4e-6, X=0)
 )
@@
 # Figure S4 
-```{r sens_LLLN, cache=T, echo=F, eval=T, fig.cap="**Sensitivity analysis under low DIN (1e-7 M) and low light (2 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
+```{r sens_LLLN, cache=T, echo=F, eval=T, fig.cap="**Sensitivity analysis under low DIN (1e-7 M) and low light (5 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
 # Set up time vector and parameters
@@
 envs <- list(
-  LLLN=list(L=2, N=1e-7, X=0)
+  LLLN=list(L=5, N=1e-7, X=0)
 )
@@
 # Figure S5
-```{r sens_LLHN, cache=T, echo=F, eval=T, fig.cap="**Fig. S5. Sensitivity analysis under high DIN (4e-6 M) and low light (2 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
+```{r sens_LLHN, cache=T, echo=F, eval=T, fig.cap="**Fig. S5. Sensitivity analysis under high DIN (4e-6 M) and low light (5 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
 # Set up time vector and parameters
@@
 envs <- list(
-  LLHN=list(L=2, N=4e-6, X=0)
+  LLHN=list(L=5, N=4e-6, X=0)
 )
@@
 # Figure S6
-```{r sens_HLLN, cache=T, echo=F, eval=T, fig.cap="**Fig. S6. Sensitivity analysis under low DIN (1e-7 M) and high light (25 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
+```{r sens_HLLN, cache=T, echo=F, eval=T, fig.cap="**Fig. S6. Sensitivity analysis under low DIN (1e-7 M) and high light (30 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
 # Set up time vector and parameters
@@
 envs <- list(
-  HLLN=list(L=25, N=1e-7, X=0)
+  HLLN=list(L=30, N=1e-7, X=0)
 )
@@
 # Figure S7
-```{r sens_HLHN, cache=T, echo=F, eval=T, fig.cap="Fig. S7. Sensitivity analysis under high DIN (4e-6 M) and high light (25 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
+```{r sens_HLHN, cache=T, echo=F, eval=T, fig.cap="Fig. S7. Sensitivity analysis under high DIN (4e-6 M) and high light (30 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
 # Set up time vector and parameters
@@
 envs <- list(
-  HLHN=list(L=25, N=4e-6, X=0)
+  HLHN=list(L=30, N=4e-6, X=0)
 )

### 2017-03-22 JTB_ms/JTB_ms.tex
make with changes

@@
   states. These dynamics reveal that `healthy' and `bleached' states
-  correspond broadly to nitrogen- and carbon-limitation of the system,
+  correspond broadly to nitrogen- and carbon-limitation in the system,
   with transitions between them occurring as integrated responses to
@@
   behaviors reproduced by the model (e.g., bleaching is exacerbated by
-  nutrients and attenuated by feeding) suggests it unifyingly captures
-  many important attributes of the system; meanwhile, its modular
-  framework and open source R code are designed to facilitate further
-  problem-specific development. We see significant potential for this
-  modeling framework to generate testable hypotheses and predict
-  integrated, mechanistic responses of corals to environmental change,
-  with important implications for understanding the performance and
-  maintenance of symbiotic systems.
+  nutrients and attenuated by feeding) suggests it captures many important
+  attributes of the system; meanwhile, its modular framework and open
+  source R code are designed to facilitate further problem-specific
+  development. We see significant potential for this modeling framework to
+  generate testable hypotheses and predict integrated, mechanistic
+  responses of corals to environmental change, with important implications
+  for understanding the performance and maintenance of symbiotic systems.
   \end{abstract}
@@
   increases accessibility for biologists and ecologists without
-  requiring significant expertise in DEB theory.
+  requiring expertise in DEB theory.
 \item
-  \emph{Provide well-documented, open-access code.} In order to
+  \emph{Provide well-documented, open source code.} In order to
   facilitate the continued development and application of theoretical
@@
   scientific community to include additional complexity and
-  problem-specific components. We chose the R language, which is freely
-  available and in common use by biologists and ecologists, to reach the
+  problem-specific components. We chose R, an open source programming
+  language in common use by biologists and ecologists, to reach the
   widest possible audience with this work.
@@
 excess excitation energy, but the upstream events that lead to this
-situation may be diverse: indeed, elevated temperature may inhibit
+situation may be diverse. Indeed, elevated temperature may inhibit
 Rubisco functioning (Jones et al. 1998) and the repair of the D1 protein
@@
 
-\begin{longtable}[]{@{}llll@{}}
+\begin{longtable}[c]{@{}llll@{}}
 \toprule
-\begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
-Symbol\strut
-\end{minipage} & \begin{minipage}[b]{0.44\columnwidth}\raggedright\strut
-Description\strut
-\end{minipage} & \begin{minipage}[b]{0.26\columnwidth}\raggedright\strut
-Units\strut
-\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
-Eq. no.\strut
-\end{minipage}\tabularnewline
+\begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
+Symbol
+\strut\end{minipage} &
+\begin{minipage}[b]{0.45\columnwidth}\raggedright\strut
+Description
+\strut\end{minipage} &
+\begin{minipage}[b]{0.27\columnwidth}\raggedright\strut
+Units
+\strut\end{minipage} &
+\begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
+Eq. no.
+\strut\end{minipage}\tabularnewline
 \midrule
 \endhead
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_X\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Prey uptake rate\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-mol X C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-3\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_N\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Nitrogen uptake rate\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-mol N Cmol H\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-4\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_{HG}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Host biomass formation rate\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-5\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_{HT}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Host biomass turnover rate\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-6\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(r_{NH}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Recycled nitrogen from host turnover\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-mol N C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-7\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(\rho_N\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Nitrogen shared with the symbiont\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-mol N C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-8\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_{eC}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Excess carbon used to activate host CCMs\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-mol C C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-9\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_{CO_2}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-CO\textsubscript{2} input to photosynthesis\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_X\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Prey uptake rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+mol X C-mol H\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+3
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_N\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Nitrogen uptake rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+mol N Cmol H\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+4
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{HG}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Host biomass formation rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+5
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{HT}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Host biomass turnover rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+6
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(r_{NH}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Recycled nitrogen from host turnover
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+mol N C-mol H\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+7
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(\rho_N\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Nitrogen shared with the symbiont
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+mol N C-mol H\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+8
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{eC}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Excess carbon used to activate host CCMs
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+mol C C-mol H\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+9
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{CO_2}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+CO\textsubscript{2} input to photosynthesis
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
 mol CO\textsubscript{2} C-mol H\textsuperscript{-1}
-d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-10\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_L\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Light absorption rate\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-12\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(r_{CH}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Recycled CO\textsubscript{2} from host turnover\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+10
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_L\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Light absorption rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+12
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(r_{CH}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Recycled CO\textsubscript{2} from host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
 mol CO\textsubscript{2} C-mol H\textsuperscript{-1}
-d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-13\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(r_{CS}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Recycled CO\textsubscript{2} from symbiont turnover\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+13
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(r_{CS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Recycled CO\textsubscript{2} from symbiont
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
 mol CO\textsubscript{2} C-mol S\textsuperscript{-1}
-d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-14\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_{CP}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Photosynthesis rate\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-mol C C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-15\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_{eL}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Light energy in excess of photochemistry\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-16\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_{NPQ}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Total capacity of NPQ\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-17\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(c_{ROS}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-ROS production proportional to baseline\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
---\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-18\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(r_{NS}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Recycled nitrogen from symbiont turnover\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-mol N C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-19\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_{SG}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Symbiont biomass formation rate\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-20\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(\rho_C\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Fixed carbon shared with host\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-mol C C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-21\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_{ST}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Symbiont biomass turnover rate\strut
-\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
-C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-22\strut
-\end{minipage}\tabularnewline
+d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+14
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{CP}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Photosynthesis rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+mol C C-mol S\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+15
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{eL}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Light energy in excess of photochemistry
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+16
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{NPQ}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Total capacity of NPQ
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+17
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(c_{ROS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+ROS production proportional to baseline
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+--
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+18
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(r_{NS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Recycled nitrogen from symbiont turnover
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+mol N C-mol S\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+19
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{SG}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Symbiont biomass formation rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+20
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(\rho_C\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Fixed carbon shared with host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+mol C C-mol S\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+21
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{ST}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Symbiont biomass turnover rate
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+22
+\strut\end{minipage}\tabularnewline
 \bottomrule
@@
 
-\begin{longtable}[]{@{}llll@{}}
+\begin{longtable}[c]{@{}llll@{}}
 \toprule
-\begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
-Symbol\strut
-\end{minipage} & \begin{minipage}[b]{0.44\columnwidth}\raggedright\strut
-Description\strut
-\end{minipage} & \begin{minipage}[b]{0.24\columnwidth}\raggedright\strut
-Units\strut
-\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\raggedright\strut
-Value\strut
-\end{minipage}\tabularnewline
+\begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
+Symbol
+\strut\end{minipage} &
+\begin{minipage}[b]{0.45\columnwidth}\raggedright\strut
+Description
+\strut\end{minipage} &
+\begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
+Units
+\strut\end{minipage} &
+\begin{minipage}[b]{0.12\columnwidth}\raggedright\strut
+Value
+\strut\end{minipage}\tabularnewline
 \midrule
 \endhead
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(n_{NH}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-N:C molar ratio in host biomass\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
---\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.18\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(n_{NS}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-N:C molar ratio in symbiont biomass\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
---\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.13\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(n_{NX}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-N:C molar ratio in prey biomass\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
---\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.2\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_{HT}^0\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Specific maintenance rate of host biomass\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.03\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_{ST}^0\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Specific maintenance rate of symbiont biomass\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.03\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(\sigma_{NH}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Proportion host N-turnover recycled to host\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
---\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.9\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(\sigma_{CH}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Proportion host C-turnover recycled to photosynthesis\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
---\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.1\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(\sigma_{NS}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Proportion symbiont N-turnover recycled to symbiont\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
---\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.9\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(\sigma_{CS}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Proportion symbiont C-turnover recycled to photosynthesis\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
---\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.9\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_{Xm}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Maximum specific feeding rate of host\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-molX C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.13\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(K_X\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Half-saturation constant for prey uptake by host\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-molX L\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(1.95\times 10^{-5}\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_{Nm}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Maximum specific DIN uptake rate by host\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-mol N C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.035\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(K_N\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Half-saturation constant for DIN uptake by host\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-mol N L\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(1.5\times 10^{-6}\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(k_{CO_2}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Efficacy of CO\textsubscript{2} delivery by host CCMs\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-mol CO\textsubscript{2} mol C\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(10\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_{HGm}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Maximum specific growth rate of host\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(1\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(y_{CL}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Quantum yield of photosynthesis\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-mol C mol photons\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.1\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(\bar{a}^*\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Effective light-absorbing cross-section of symbiont\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-m\textsuperscript{2} C-mol S\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(1.34\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(k_{NPQ}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-NPQ capacity of symbiont\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(112\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(k_{ROS}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(n_{NH}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+N:C molar ratio in host biomass
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+--
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.18\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(n_{NS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+N:C molar ratio in symbiont biomass
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+--
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.13\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(n_{NX}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+N:C molar ratio in prey biomass
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+--
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.2\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{HT}^0\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Specific maintenance rate of host biomass
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.03\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{ST}^0\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Specific maintenance rate of symbiont biomass
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.03\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(\sigma_{NH}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Proportion host N-turnover recycled to host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+--
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.9\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(\sigma_{CH}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Proportion host C-turnover recycled to photosynthesis
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+--
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.1\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(\sigma_{NS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Proportion symbiont N-turnover recycled to symbiont
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+--
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.9\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(\sigma_{CS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Proportion symbiont C-turnover recycled to photosynthesis
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+--
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.9\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{Xm}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Maximum specific assimilation rate from host feeding
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molX C-mol H\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.13\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(K_X\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Half-saturation constant for prey uptake by host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+molX L\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(10^{-6}\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{Nm}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Maximum specific DIN uptake rate by host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+mol N C-mol H\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.035\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(K_N\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Half-saturation constant for DIN uptake by host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+mol N L\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(1.5\times 10^{-6}\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(k_{CO_2}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Efficacy of CO\textsubscript{2} delivery by host CCMs
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+mol CO\textsubscript{2} mol C\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(10\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{HGm}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Maximum specific growth rate of host
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(1\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(y_{CL}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Quantum yield of photosynthesis
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+mol C mol photons\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.1\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(y_{C}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Yield of biomass formation from carbon
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+C-mol mol C\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.8\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(\bar{a}^*\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Effective light-absorbing cross-section of symbiont
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+m\textsuperscript{2} C-mol S\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(1.34\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(k_{NPQ}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+NPQ capacity of symbiont
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(112\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(k_{ROS}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
 Excess photon energy that doubles ROS production, relative to baseline
-levels\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(80\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_{CPm}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Maximum specific photosynthesis rate of symbiont\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-mol C C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(2.8\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(j_{SGm}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Maximum specific growth rate of symbiont\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.25\)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
-\(b\)\strut
-\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
-Scaling parameter for bleaching response\strut
-\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
---\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(5\)\strut
-\end{minipage}\tabularnewline
+levels
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(80\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{CPm}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Maximum specific photosynthesis rate of symbiont
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+mol C C-mol S\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(2.8\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(j_{SGm}\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Maximum specific growth rate of symbiont
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.25\)
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\(b\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Scaling parameter for bleaching response
+\strut\end{minipage} &
+\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
+--
+\strut\end{minipage} &
+\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(5\)
+\strut\end{minipage}\tabularnewline
 \bottomrule
@@
 acceleration; likewise, \(j_{eC}\) accelerates the rate of \(j_{CO_2}\).
-Recycled mass fluxes from biomass turnover are not shown for clarity
-(but see Table 1 for definitions).}
+Recycling fluxes are not shown for clarity (but see Table 1 for
+definitions).}
 \end{figure}
@@
 The coral animal acquires both carbon and nitrogen from feeding on prey
-from the environment. Prey acquisition is specified by Michaelis-Menten
-kinetics (i.e., a Holling type II function) with a maximum feeding rate
-\(j_{Xm}\) and half-saturation constant \(K_X\):
+from the environment. Assimilation from feeding is specified by
+Michaelis-Menten kinetics (i.e., a Holling type II function) with a
+maximum rate of \(j_{Xm}\) and half-saturation constant \(K_X\):
 
@@
 and \(r_{NH}\) is recycled nitrogen liberated by host biomass turnover
-(see Eq. 7). Host biomass turnover is equal to the specific maintenance
-rate of host biomass,
+(see Eq. 7). The parameter \(y_C\) specifies the yield of biomass from
+organic carbon, which we take to be 0.8 to satisfy redox balance (see
+Muller et al. (2009)).
+
+Host biomass turnover is equal to the specific maintenance rate of host
+biomass,
 
@@
 
-\begin{equation} j_{eC} = (jX + \rho_C{S \over H} - j_{HG})_+ \end{equation}
+\begin{equation} j_{eC} = (j_X + \rho_C{S \over H} - j_{HG}y_{C}^{-1})_+ \end{equation}
 
 This flux, \(j_{eC}\), is assumed to be available to the host as a
-respiratory substrate to support various energetically-demanding
-processes; of particular importance is the host's active carbon
-concentrating mechanisms (CCMs) that supply CO\textsubscript{2} for
-symbiont photosynthesis (Wooldridge 2013; Hopkinson, Tansik, and Fitt
-2015). We therefore specify \(j_{CO_2}\) as the host-mediated delivery
-of CO\textsubscript{2} to photosynthesis that encompasses potentially
+respiratory substrate to support energetically-demanding processes; of
+particular importance is the host's active carbon concentrating
+mechanisms (CCMs) that supply CO\textsubscript{2} for symbiont
+photosynthesis (Wooldridge 2013; Hopkinson, Tansik, and Fitt 2015). We
+therefore specify \(j_{CO_2}\) as the host-mediated delivery of
+CO\textsubscript{2} to photosynthesis that encompasses potentially
 diverse carbon-concentrating mechanisms (CCMs), including active
@@
 proportion \(\sigma_{CH}\) of metabolic CO\textsubscript{2} generated by
-host biomass turnover is passively available to the photosynthesis SU,
-according to:
-
-\begin{equation} r_{CH}=\sigma_{CH}j_{HT} \end{equation}
+the host from both biomass turnover and formation is passively available
+to the photosynthesis SU, according to:
+
+\begin{equation} r_{CH}=\sigma_{CH}(j_{HT} + (1-y_C)j_{HG}/y_C) \end{equation}
 
@@
   photodamage/bleaching component, as this loss represents biomass that
-  is expelled from the host.}:
-
-\begin{equation} r_{CS}=\sigma_{CS}j_{ST}^0 \end{equation}
+  is expelled from the host.} and formation:
+
+\begin{equation} r_{CS}=\sigma_{CS}(j_{ST}^0 + (1-y_C)j_{SG}/y_C) \end{equation}
 
@@
 of photosynthesis can lead to ROS production, an important mechanism
-(Wooldridge 2009) that was not captured by previous representations of
-photooxidative stress (Eynaud, Nisbet, and Muller 2011). With this
-single SU, both the light and dark reactions of photosynthesis are
-represented, allowing for sink-limitation (i.e.,
+(Butow et al. 1998; Wooldridge 2009) that was not captured by previous
+representations of photooxidative stress (Eynaud, Nisbet, and Muller
+2011). With this single SU, both the light and dark reactions of
+photosynthesis are represented, allowing for sink-limitation (i.e.,
 CO\textsubscript{2}-limitation) to cause overreduction of the electron
@@
 rate, in units of Cmol Cmol\textsuperscript{-1} d\textsuperscript{-1}.
+When both substrate input fluxes are higher than what can be used at the
+maximum production rate, this limitation coefficient is zero, implying
+that neither substrate is limiting production.
 
@@
 light gradient. Alternate stable states occur between
-\textasciitilde{}33-44 mol photons m\textsuperscript{-2}
+\textasciitilde{}25-42 mol photons m\textsuperscript{-2}
 d\textsuperscript{-1} under these conditions (DIN=1e-7 molN
-L\textsuperscript{-1}; X=2e-6 molX L\textsuperscript{-1}), depending on
-whether initial S:H is high (0.1, closed circles), representing a
-healthy coral, or low (0.0001, open circles), representing a bleached
-coral.}
+L\textsuperscript{-1}; X=1e-7 CmolX L\textsuperscript{-1}), depending on
+whether initial S:H is high (1, closed circles), representing a healthy
+coral, or low (0.0001, open circles), representing a bleached coral.
+Arrow above point at L=25 indicates a S:H ratio beyond the axis range;
+this `overshoot' phenomenon, in which initially bleached corals may
+achieve high S:H ratios while remaining in a carbon-limited state is
+discussed in the \emph{Coral Bleaching and Recovery} section.}
 \end{figure}
@@
 \includegraphics{../img/Fig3.png}
-\caption{Steady state values of (A) specific growth (Cmol
-Cmol\textsuperscript{-1} d\textsuperscript{-1}) and (B) the symbiont to
+\caption{Steady state values of (A) specific growth (C-mol H C-mol
+H\textsuperscript{-1} d\textsuperscript{-1}) and (B) the symbiont to
 host biomass ratio (C-mol S C-mol H\textsuperscript{-1}) across
@@
 Simulations for each combination of light and nutrients (41 points along
-each axis) were run for 100 days with a time step of 1 day. All
-parameters were set to default values and prey density set to zero.
-Negative steady state growth rates, and corresponding S:H ratios, were
-set to zero.}
+each axis) were run to steady state with all parameters at default
+values and prey density set to zero. Negative steady state growth rates
+and corresponding S:H ratios were set to zero, and a ceiling of 0.5 was
+imposed on S:H ratios to aid in visualization.}
 \end{figure}
@@
 The values used for each parameter in the model (Table 2) are derived
-from relevant literature; these derivations are described in the
-Supplementary Information. Here we evaluate the sensitivity of the model
-to changes in these parameter values, which also serves to demonstrate
-the behavior of the dynamical system. We calculated fractional change in
-steady state values in response to fractional changes in parameter
-values, relative to their default values, under environmental conditions
-typical of coral reefs.
+from relevant literature (see Supplementary Information). Here we
+evaluate the sensitivity of the model to changes in these parameter
+values, which also serves to demonstrate the behavior of the dynamical
+system. We calculated fractional change in steady state values in
+response to fractional changes in parameter values, relative to their
+default values, under environmental conditions typical of coral reefs.
 
@@
 lines) in response to fractional changes in default parameter values
-(see Table 2 for default values). Parameters are grouped by which
-processes they are involved in. This sensitivity analysis was conducted
-at conditions typical for coral reef environments: low DIN (1e-7 M) and
-intermediate light (15 mol m\textsuperscript{-2} d\textsuperscript{-2}).
-Sensitivity analyses conducted other environmental conditions are
-presented in Figs S3-S7.}
+(see Table 2 for default values). Parameters are grouped by the
+processes in which they are involved. This sensitivity analysis was
+conducted at conditions typical for coral reef environments: low DIN
+(1e-7 M) and intermediate light (15 mol photons m\textsuperscript{-2}
+d\textsuperscript{-2}), with prey density set to zero. Sensitivity
+analyses conducted other environmental conditions are presented in Figs
+S3-S7.}
 \end{figure}
@@
 Parameters relating to photooxidative stress and bleaching have little
-impact under these environmental conditions (Fig. 4), but have larger
-impacts under higher light (e.g., Fig. S6). Sensitivity analyses
+impact under low nutrients and intermediate light (Fig. 4), but have
+larger impacts under higher light (e.g., Fig. S6). Sensitivity analyses
 conducted under different combinations of external light and nutrients
@@
 study as inputs, the model reproduces this observed interaction among
-environmental factors (Fig. 5), and also provides the mechanism: as
-increasing light in summer decreases carbon fixation due to
-photooxidative stress, symbionts grow slower than the host, leading to
-decreasing S:H ratios. Under nutrient enrichment, this effect is more
-pronounced, as CO\textsubscript{2}-limitation (due to higher symbiont
+environmental factors (Fig. 5), and also provides the mechanism:
+increasing light in summer decreases symbiont growth rates due to
+photooxidative stress, leading to decreasing S:H ratios. Under nutrient
+enrichment, this effect is more pronounced, as
+CO\textsubscript{2}-limitation of photosynthesis (due to higher symbiont
 standing stocks) causes mild bleaching that results in a similar
@@
 winter then alleviates the photooxidative stress constraints on carbon
-fixation, such that nitrogen constrains the S:H ratio, explaining why
-S:H increases more when DIN is enriched (Fig. S8).
+fixation such that nitrogen-limitation constrains the S:H ratio,
+explaining why S:H increases more when DIN is enriched (Fig. S8).
 
@@
 growth. Light input (upper panel) was designed as a sinusoidal curve
-with a period of one year, with maximum and minimum values of 44 and 22
-mol photons m\textsuperscript{-2} d\textsuperscript{-1}, based on
-Stimson (1997). The dynamic behavior of symbiont to host biomass ratio
-(middle panel) and the specific growth rate of host biomass (lower
-panel) show seasonal oscillations that are greater in magnitude under
-higher nutrients (1e-5 mol N L\textsuperscript{-1}; dashed lines)
-relative to lower nutrients (1e-7 mol N L\textsuperscript{-1}; solid
-lines), consistent with the findings of Stimson (1997). Prey density was
-set at 1e-5 CmolX L\textsuperscript{-1}.}
+with a period of one year, with maximum and minimum values of 44 and 20
+mol photons m\textsuperscript{-2} d\textsuperscript{-1}, corresponding
+to those measured by Stimson (1997). The dynamic behavior of symbiont to
+host biomass ratio (middle panel) and the specific growth rate of host
+biomass (lower panel) show seasonal oscillations that are greater in
+magnitude under high nutrients (15.14 \(\mu\)M N; dashed lines) relative
+to low nutrients (0.14 \(\mu\)M N; solid lines), consistent with the
+findings of Stimson (1997). Prey density was set at 1e-6 CmolX
+L\textsuperscript{-1}.}
 \end{figure}
@@
 experimental and field observations for corals, and demonstrates the
-power of this model to predict dynamic behavior that mechanistically
+model's ability to predict dynamic behavior that mechanistically
 integrates multiple environmental drivers.
@@
 were set at default values with external DIN=1e-7 mol N/L and prey
-density=0).}
+density set to zero).}
 \end{figure}
@@
 d\textsuperscript{-1}) under default environmental conditions (solid
-line; DIN=1e-7 mol N L\textsuperscript{-1}; prey=3e-6 CmolX
-L\textsuperscript{-1}), or with elevated feeding (dashed line; prey=6e-6
-CmolX L\textsuperscript{-1}), or elevated nutrients (dotted line;
-DIN=1e-6 mol N L\textsuperscript{-1}). Initial symbiont biomass was set
-to the steady state for each set of starting conditions, respectively.
-(All parameters set to default values).}
+line; DIN=1e-7 mol N L\textsuperscript{-1}; prey=2e-7 C-molX
+L\textsuperscript{-1}), or with elevated feeding (dashed line; prey=1e-6
+C-molX L\textsuperscript{-1}), or elevated nutrients (dotted line;
+DIN=4e-6 mol N L\textsuperscript{-1}). Initial symbiont biomass was set
+to the steady state for each set of starting conditions, with all other
+parameters at default values.}
 \end{figure}
@@
 \caption{Recovery from bleaching with varying N:C availability (A:
-N:C=0.09; B: N:C=0.10; C: N:C=0.30). Higher N:C ratios in a
+N:C=0.100; B: N:C=0.175; C: N:C=0.220). Higher N:C ratios in a
 heterotrophic food source (effectively representing higher external DIN
@@
 were calculated using Eq. 23. All simulations were run with default
-parameters (except varying \(n_{NX}\)), with L=10 mol photons
+parameters (except varying \(n_{NX}\)), with L=20 mol photons
 m\textsuperscript{2} d\textsuperscript{-1}, DIN=1e-7 mol N
-L\textsuperscript{-1}, prey=1e-6 CmolX L\textsuperscript{-1}, and
+L\textsuperscript{-1}, prey=1e-7 CmolX L\textsuperscript{-1}, and
 initial S:H biomass=0.001.}
@@
 alone does not necessarily indicate that a `healthy' state has been
-reached, since the C-limited state may still persist (e.g.~Fig. 8B, 8C).
-This could explain why corals that have recovered their symbiont
+reached, since the carbon-limited state may still persist (e.g.~Fig. 8B,
+8C). This could explain why corals that have recovered their symbiont
 populations after bleaching may still exhibit energetic deficits and
 physiological impacts, possibly for months to years (Levitan et al.
-2014, Hughes and Grottoli (2013))). Therefore, host acquisition of
-carbon from a source other than the symbiont may be extremely important
-for the system to recover from bleaching. Indeed, host feeding has been
-shown to promote a more rapid return to pre-bleaching levels of key
-physiological parameters in recovering corals (Connolly, Lopez-Yglesias,
-and Anthony 2012). The model predicts that acquisition of carbon from
-other sources, such as direct uptake of DOC (Levas et al. 2015), may
-also promote more rapid recovery from a bleached state.
+2014, Hughes and Grottoli (2013))). These findings suggest that host
+acquisition of carbon from a source other than the symbiont may be
+extremely important for the system to recover from bleaching. Indeed,
+host feeding has been shown to promote a more rapid return to
+pre-bleaching levels of key physiological parameters in recovering
+corals (Connolly, Lopez-Yglesias, and Anthony 2012). Additional carbon
+sources for the host, such as direct uptake of dissolved organic carbon
+(Levas et al. 2015), may also promote more rapid recovery from a
+bleached state.
 
@@
 forcing capabilities (e.g., external DIC, temperature). Importantly,
-this effort may be undertaken by the scientific community at large,
-including those with theoretical and empirical backgrounds, which is why
-the code for this model has been developed openly using the R language.
-Ultimately, the continued refinement of these tools is fundamental in
-elucidating the mechanisms of symbiosis function and dysfunction, and in
-predicting coral responses to environmental change.
+open source R code allows this effort to benefit from contributions from
+the wider scientific community, including those with empirical and
+theoretical backgrounds. Ultimately, the continued refinement of these
+tools is fundamental in elucidating the mechanisms of symbiosis function
+and dysfunction, and in predicting coral responses to environmental
+change.
 
@@
 53--69.
+
+\hypertarget{ref-Butow:1998gz}{}
+Butow, B, D Wynne, A Sukenik, O Hadas, and E Tel-Or. 1998. ``The
+synergistic effect of carbon concentration and high temperature on lipid
+peroxidation in Peridinium gatunense.'' \emph{Journal of Plankton
+Research} 20 (2). Oxford University Press: 355--69.
+doi:\href{https://doi.org/10.1093/plankt/20.2.355}{10.1093/plankt/20.2.355}.
 

### 2017-03-22 JTB_ms/JTB_ms.Rmd
minor text edits

@@
 | ------- | ------------------------------------ | --------------------- | ------- |
-| $j_X$     | Prey uptake rate                       | mol X C-mol H^-1^ d^-1^ | 3 |
+| $j_X$     | Prey assimilation (feeding) rate                       | C-mol X C-mol H^-1^ d^-1^ | 3 |
 | $j_N$     | Nitrogen uptake rate                   | mol N Cmol H^-1^ d^-1^ | 4 |
@@
 | $n_{NX}$ | N:C molar ratio in prey biomass | -- | $`r defpars$nNX`$ |
-| $j_{HT}^0$ | Specific maintenance rate of host biomass | C-mol H C-mol H^-1^ d^-1^ | $`r defpars$jHT0`$ |
-| $j_{ST}^0$ | Specific maintenance rate of symbiont biomass | C-mol S C-mol S^-1^ d^-1^ | $`r defpars$jST0`$ |
-| $\sigma_{NH}$ | Proportion host N-turnover recycled to host | -- | $`r defpars$sigmaNH`$ |
-| $\sigma_{CH}$ | Proportion host C-turnover recycled to photosynthesis | -- | $`r defpars$sigmaCH`$ |
-| $\sigma_{NS}$ | Proportion symbiont N-turnover recycled to symbiont | -- | $`r defpars$sigmaNS`$ |
-| $\sigma_{CS}$ | Proportion symbiont C-turnover recycled to photosynthesis | -- | $`r defpars$sigmaCS`$ |
-| $j_{Xm}$ | Maximum specific assimilation rate from host feeding | molX C-mol H^-1^ d^-1^ | $`r defpars$jXm`$ |
-| $K_X$ | Half-saturation constant for prey uptake by host | molX L^-1^ | $`r defpars$KX`$ |
-| $j_{Nm}$ | Maximum specific DIN uptake rate by host | mol N C-mol H^-1^ d^-1^ | $`r defpars$jNm`$ |
-| $K_N$ | Half-saturation constant for DIN uptake by host | mol N L^-1^ | $`r defpars$KN`$ |
-| $k_{CO_2}$ | Efficacy of CO~2~ delivery by host CCMs | mol CO~2~ mol C^-1^ | $`r defpars$kCO2`$ |
+| $j_{HT}^0$ | Maintenance rate of host biomass | C-mol H C-mol H^-1^ d^-1^ | $`r defpars$jHT0`$ |
+| $j_{ST}^0$ | Maintenance rate of symbiont biomass | C-mol S C-mol S^-1^ d^-1^ | $`r defpars$jST0`$ |
+| $\sigma_{NH}$ | Proportion N turnover recycled in host | -- | $`r defpars$sigmaNH`$ |
+| $\sigma_{CH}$ | Proportion host metabolic CO~2~ recycled to photosynthesis | -- | $`r defpars$sigmaCH`$ |
+| $\sigma_{NS}$ | Proportion N turnover recycled in symbiont | -- | $`r defpars$sigmaNS`$ |
+| $\sigma_{CS}$ | Proportion symbiont metabolic CO~2~ recycled to photosynthesis | -- | $`r defpars$sigmaCS`$ |
+| $j_{Xm}$ | Maximum prey assimilation rate from host feeding | molX C-mol H^-1^ d^-1^ | $`r defpars$jXm`$ |
+| $K_X$ | Half-saturation constant for prey assimilation | molX L^-1^ | $`r defpars$KX`$ |
+| $j_{Nm}$ | Maximum host DIN uptake rate | mol N C-mol H^-1^ d^-1^ | $`r defpars$jNm`$ |
+| $K_N$ | Half-saturation constant for host DIN uptake | mol N L^-1^ | $`r defpars$KN`$ |
+| $k_{CO_2}$ | Efficacy of CO~2~ delivery to photosynthesis by host CCMs | mol CO~2~ mol C^-1^ | $`r defpars$kCO2`$ |
 | $j_{HGm}$ | Maximum specific growth rate of host | C-mol H C-mol H^-1^ d^-1^ | $`r defpars$jHGm`$ |
@@
 
-\begin{equation} j_{HG} = \bigg({1 \over j_{HGm}} + {1 \over {y_{C}\rho_C{S \over H} + j_X}} + {1 \over {(j_N + n_{NX}j_X + r_{NH}) / n_{NH}}} - {1 \over {y_{C}{\rho_C{S \over H} + j_X} + (j_N + n_{NX}j_X + r_{NH}) / n_{NH}}}\bigg)^{-1} \end{equation}
+\begin{equation} j_{HG} = \bigg({1 \over j_{HGm}} + {1 \over {y_{C}\rho_C{S \over H} + j_X}} + {1 \over {(j_N + n_{NX}j_X + r_{NH})n_{NH}^{-1}}} - {1 \over {y_{C}{\rho_C{S \over H} + j_X} + (j_N + n_{NX}j_X + r_{NH})n_{NH}^{-1}}}\bigg)^{-1} \end{equation}
 
@@
 This flux, $j_{eC}$, is assumed to be available to the host as a respiratory substrate to support energetically-demanding processes; of particular importance is the host's active carbon concentrating mechanisms (CCMs) that supply CO~2~ for symbiont photosynthesis [@Wooldridge:2013tj; @Hopkinson:2015ev].
-We therefore specify $j_{CO_2}$ as the host-mediated delivery of CO~2~ to photosynthesis that encompasses potentially diverse carbon-concentrating mechanisms (CCMs), including active transport of bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to CO~2~ to promote diffusion toward the symbiont [@Tansik:2015cm], and acidification of the symbiosome to increase localized CO~2~ concentrations around the symbiont [@Barott:2014gx].
+We therefore specify $j_{CO_2}$ as the host-mediated delivery of CO~2~ to photosynthesis that encompasses potentially diverse CCMs, including active transport of bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to CO~2~ to promote diffusion toward the symbiont [@Tansik:2015cm], and acidification of the symbiosome to increase localized CO~2~ concentrations around the symbiont [@Barott:2014gx].
 Since these active CCMs require energetic input by the host, we define $j_{CO_2}$ as proportional to $j_{eC}$, assuming that some of this carbon is respired to energize the CCMs.
@@
 
-\begin{equation} r_{CH}=\sigma_{CH}(j_{HT} + (1-y_C)j_{HG}/y_C) \end{equation}
+\begin{equation} r_{CH}=\sigma_{CH}(j_{HT} + (1-y_C)j_{HG}y_C^{-1}) \end{equation}
 
@@
 
-\begin{equation} r_{CS}=\sigma_{CS}(j_{ST}^0 + (1-y_C)j_{SG}/y_C) \end{equation}
+\begin{equation} r_{CS}=\sigma_{CS}(j_{ST}^0 + (1-y_C)j_{SG}y_C^{-1}) \end{equation}
 
@@
 
-\begin{equation} j_{eL} = (j_L - j_{CP} / y_{CL})_+ \end{equation}
+\begin{equation} j_{eL} = (j_L - j_{CP}y_{CL}^{-1})_+ \end{equation}
 
@@
 
-\begin{equation} j_{NPQ} = \bigg({1 \over k_{NPQ}} + {1 \over j_{eL}}\bigg)^{-1} \end{equation}
+\begin{equation} j_{NPQ} = (k_{NPQ}^{-1} + j_{eL}^{-1})^{-1} \end{equation}
 
@@
 
-\begin{equation} j_{SG} = \bigg({1 \over j_{SGm}} + {1 \over y_{C}j_{CP}} + {1 \over {(\rho_N{H \over S} + r_{NS}) / n_{NS}}} - {1 \over {y_{C}j_{CP} + (\rho_N{H \over S} + r_{NS}) / n_{NS}}}\bigg)^{-1} \end{equation}
+\begin{equation} j_{SG} = \bigg({1 \over j_{SGm}} + {1 \over y_{C}j_{CP}} + {1 \over {(\rho_N{H \over S} + r_{NS})n_{NS}^{-1}}} - {1 \over {y_{C}j_{CP} + (\rho_N{H \over S} + r_{NS})n_{NS}^{-1}}}\bigg)^{-1} \end{equation}
 

### 2017-03-22 JTB_supp/JTB_supp.Rmd
add yC description

@@
 molC molph^-1^
+<br><br>
+
+#### $y_{C}$: Yield of biomass formation from photosynthate {-}
+This coefficient specifies the efficiency of conversion of photosynthate into biomass at the biomass synthesizing units, which combine photosynthate (CH~2~O), ammonia (NH~3~) and O~2~ to form biomass, water, and CO~2~.
+In order to balance this macrochemical equation, there must be a carbon loss, which is accounted for using $y_{C}$ < 1.
+We take a value of 0.8 following @Muller:2009io.
+```{r yC, echo=F}
+0.8
+```
+C-mol mol C^-1^
+
+Following this macrochemical equation, the amount of CO~2~ produced during biomass formation can be specified as $(1-y_C)j_P/y_C$, where j_P is the production flux of the SU (see equations 13 and 14 in the main text).
 <br><br>

### 2017-03-22 R/run_coral_ss.R
fix ss oscillation filter

@@ -126,8 +126,8 @@ run_coral_ss <- function(env, pars, dt) {
       shss <- ifelse(abs(S$S[t]/H$H[t] - S$S[t-10/dt]/H$H[t-10/dt]) < 0.00001, T, F)
     }
     #Test if system is oscillating but with negative growth (e.g., L=1.2, N=2.6e-6, X=0)
-    if (t > 1000) {
-     if (any(H$dH.Hdt[(t-500):t] < 0)) {
+    if (t > 10000) {
+     if (any(H$dH.Hdt[(t-1000):t] < 0)) {
        grss <- T; shss <- T
        H$dH.Hdt[t] <- 0
      }

### 2017-03-22 JTB_ms/JTB_ms.tex
remake w fixed ss oscill filter

@@
 \begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Prey uptake rate
+Prey assimilation (feeding) rate
 \strut\end{minipage} &
 \begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-mol X C-mol H\textsuperscript{-1} d\textsuperscript{-1}
+C-mol X C-mol H\textsuperscript{-1} d\textsuperscript{-1}
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Specific maintenance rate of host biomass
+Maintenance rate of host biomass
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Specific maintenance rate of symbiont biomass
+Maintenance rate of symbiont biomass
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Proportion host N-turnover recycled to host
+Proportion N turnover recycled in host
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Proportion host C-turnover recycled to photosynthesis
+Proportion host metabolic CO\textsubscript{2} recycled to photosynthesis
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Proportion symbiont N-turnover recycled to symbiont
+Proportion N turnover recycled in symbiont
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Proportion symbiont C-turnover recycled to photosynthesis
+Proportion symbiont metabolic CO\textsubscript{2} recycled to
+photosynthesis
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Maximum specific assimilation rate from host feeding
+Maximum prey assimilation rate from host feeding
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Half-saturation constant for prey uptake by host
+Half-saturation constant for prey assimilation
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Maximum specific DIN uptake rate by host
+Maximum host DIN uptake rate
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Half-saturation constant for DIN uptake by host
+Half-saturation constant for host DIN uptake
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Efficacy of CO\textsubscript{2} delivery by host CCMs
+Efficacy of CO\textsubscript{2} delivery to photosynthesis by host CCMs
 \strut\end{minipage} &
@@
 CO\textsubscript{2} to photosynthesis that encompasses potentially
-diverse carbon-concentrating mechanisms (CCMs), including active
-transport of bicarbonate, carbonic anhydrase-catalyzed conversion of
-bicarbonate to CO\textsubscript{2} to promote diffusion toward the
-symbiont (Tansik, Fitt, and Hopkinson 2015), and acidification of the
-symbiosome to increase localized CO\textsubscript{2} concentrations
-around the symbiont (Barott et al. 2014). Since these active CCMs
-require energetic input by the host, we define \(j_{CO_2}\) as
-proportional to \(j_{eC}\), assuming that some of this carbon is
-respired to energize the CCMs. This formulation means that the symbiont
-indirectly ensures its own CO\textsubscript{2} supply by providing fixed
-carbon (=energy) to the host (Wooldridge 2013). The parameter
-\(k_{CO_2}\) scales the efficacy of host CCMs, which enables the
-comparison of different rates of CO\textsubscript{2} delivery that may
-characterize different coral species (Wooldridge 2014a). The active
+diverse CCMs, including active transport of bicarbonate, carbonic
+anhydrase-catalyzed conversion of bicarbonate to CO\textsubscript{2} to
+promote diffusion toward the symbiont (Tansik, Fitt, and Hopkinson
+2015), and acidification of the symbiosome to increase localized
+CO\textsubscript{2} concentrations around the symbiont (Barott et al.
+2014). Since these active CCMs require energetic input by the host, we
+define \(j_{CO_2}\) as proportional to \(j_{eC}\), assuming that some of
+this carbon is respired to energize the CCMs. This formulation means
+that the symbiont indirectly ensures its own CO\textsubscript{2} supply
+by providing fixed carbon (=energy) to the host (Wooldridge 2013). The
+parameter \(k_{CO_2}\) scales the efficacy of host CCMs, which enables
+the comparison of different rates of CO\textsubscript{2} delivery that
+may characterize different coral species (Wooldridge 2014a). The active
 input of CO\textsubscript{2} to the photosynthesis SU is therefore

### 2017-03-22 R/run_coral_ss.R
fix ss oscill filter

@@ -125,9 +125,10 @@ run_coral_ss <- function(env, pars, dt) {
       grss <- ifelse(abs(H$dH.Hdt[t] - H$dH.Hdt[t-10/dt]) < 0.00001, T, F)
       shss <- ifelse(abs(S$S[t]/H$H[t] - S$S[t-10/dt]/H$H[t-10/dt]) < 0.00001, T, F)
     }
-    #Test if system is oscillating but with negative growth (e.g., L=1.2, N=2.6e-6, X=0)
-    if (t > 10000) {
-     if (any(H$dH.Hdt[(t-1000):t] < 0)) {
+    
+    # If system is oscillating after 1000 days, classify as steady state with no growth
+    if (t > 1000/dt) {
+     if (sum(diff(sign(diff(S$S/H$H)))!=0)>=3) {  # if slope of S/H has changed three times or more...then oscillating
        grss <- T; shss <- T
        H$dH.Hdt[t] <- 0
      }

### 2017-03-22 R/Fig2.R
add y-axis units

@@ -32,7 +32,8 @@ res <- cbind(input, output)
 # Create plot
 png("img/Fig2.png", width=5, height=5, units="in", res=300)
 
-plot(NA, xlim=c(20,50), ylim=c(0,0.20), xlab=expression("Light"~(mol~photons~m^{-2}~d^{-1})), ylab="Steady state S:H ratio")
+plot(NA, xlim=c(20,50), ylim=c(0,0.20), xlab=expression("Light"~(mol~photons~m^{-2}~d^{-1})), 
+     ylab=expression("Steady state S:H"~(C-mol~S~C-mol~H^{-1})))
 points(c(43,43,48,48), c(0.17,0.18,0.17,0.18), pch=c(19,1,19,1), cex=c(0.4,1,0.4,1), col=c("black","black","red","red"), xpd=T)
 text(41, 0.20, labels="init. S:H", xpd=T, cex=0.7, pos=2)
 text(41, c(0.17,0.18), labels=c("1", "0.0001"), xpd=T, pos=2, cex=0.7)

### 2017-03-22 JTB_ms/JTB_ms.Rmd
formatting

@@
 ![Alternate stable states in S:H biomass and growth across a light gradient.
-Alternate stable states occur between ~25-42 mol photons m^-2^ d^-1^ under these conditions (DIN=1e-7 molN L^-1^; X=1e-7 CmolX L^-1^), depending on whether initial S:H is high (1, closed circles), representing a healthy coral, or low (0.0001, open circles), representing a bleached coral.
+Alternate stable states occur between ~25-42 mol photons m^-2^ d^-1^ under these conditions (DIN=1e-7 mol N L^-1^; X=1e-7 C-mol X L^-1^), depending on whether initial S:H is high (1, closed circles), representing a healthy coral, or low (0.0001, open circles), representing a bleached coral.
 Arrow above point at L=25 indicates a S:H ratio beyond the axis range; this 'overshoot' phenomenon, in which initially  bleached corals may achieve high S:H ratios while remaining in a carbon-limited state is discussed in the *Coral Bleaching and Recovery* section.](../img/Fig2.png)

### 2017-03-22 JTB_ms/JTB_ms.tex
remake after fix ss oscill filter

@@
 \textasciitilde{}25-42 mol photons m\textsuperscript{-2}
-d\textsuperscript{-1} under these conditions (DIN=1e-7 molN
-L\textsuperscript{-1}; X=1e-7 CmolX L\textsuperscript{-1}), depending on
-whether initial S:H is high (1, closed circles), representing a healthy
-coral, or low (0.0001, open circles), representing a bleached coral.
-Arrow above point at L=25 indicates a S:H ratio beyond the axis range;
-this `overshoot' phenomenon, in which initially bleached corals may
-achieve high S:H ratios while remaining in a carbon-limited state is
+d\textsuperscript{-1} under these conditions (DIN=1e-7 mol N
+L\textsuperscript{-1}; X=1e-7 C-mol X L\textsuperscript{-1}), depending
+on whether initial S:H is high (1, closed circles), representing a
+healthy coral, or low (0.0001, open circles), representing a bleached
+coral. Arrow above point at L=25 indicates a S:H ratio beyond the axis
+range; this `overshoot' phenomenon, in which initially bleached corals
+may achieve high S:H ratios while remaining in a carbon-limited state is
 discussed in the \emph{Coral Bleaching and Recovery} section.}

### 2017-03-23 JTB_ms/JTB_ms.Rmd
format units, change Fig S9 conditions

@@
 | $\sigma_{CS}$ | Proportion symbiont metabolic CO~2~ recycled to photosynthesis | -- | $`r defpars$sigmaCS`$ |
-| $j_{Xm}$ | Maximum prey assimilation rate from host feeding | molX C-mol H^-1^ d^-1^ | $`r defpars$jXm`$ |
-| $K_X$ | Half-saturation constant for prey assimilation | molX L^-1^ | $`r defpars$KX`$ |
+| $j_{Xm}$ | Maximum prey assimilation rate from host feeding | C-mol X C-mol H^-1^ d^-1^ | $`r defpars$jXm`$ |
+| $K_X$ | Half-saturation constant for prey assimilation | C-mol X L^-1^ | $`r defpars$KX`$ |
 | $j_{Nm}$ | Maximum host DIN uptake rate | mol N C-mol H^-1^ d^-1^ | $`r defpars$jNm`$ |

### 2017-03-23 JTB_supp/JTB_supp.Rmd
format units, change Fig S9 conditions

@@
 We assume (simplistically) the same rate per unit of total biomass:
-```{r jHT, echo=T}
+```{r jHT, echo=F}
 0.03
 ```
-CmolH CmolH^-1^ d^-1^
+C-mol H C-mol H^-1^ d^-1^
 <br><br>
@@
 Without data for a more confident estimation of this parameter, we choose a value at the low end of this range, since 0.2 seems unrealistic and many rates may be depressed in symbiotic dinoflagellates relative to their free-living counterparts.
-```{r jST, echo=T}
+```{r jST, echo=F}
 0.03
 ```
-CmolS CmolS^-1^ d^-1^  
+C-mol S C-mol S^-1^ d^-1^  
 <br><br>
@@
 ```
-molX CmolH^-1^ d^-1^
+C-mol X C-mol H^-1^ d^-1^
 <br><br>
-
-@Osinga:2011di measured a maximum feeding rate of 
-half sat = 294119 nauplii L^-1^
-294119 artemia/L * 6.78e-7 gC/artemia = 0.19941268 gC/L / 12.011g/mol = 0.01656815
 
@@
 ```
-CmolX L^-1^
+C-mol X L^-1^
 <br><br>
@@
 ```
-mol NH~4~^+^-N  CmolH^-1^ d^-1^  
+mol N C-mol H^-1^ d^-1^  
 <br>
@@
 ``` 
-molN CmolH^-1^ d^-1^  
+mol N C-mol H^-1^ d^-1^  
 <br>
@@
 ```
-molN CmolH^-1^ d^-1^   
+mol N C-mol H^-1^ d^-1^   
 <br>
@@
 ```
-molN CmolH^-1^ d^-1^
+mol N C-mol H^-1^ d^-1^
 <br><br>
@@
 ``` 
-molN L^-1^
+mol N L^-1^
 <br><br>
@@
 ```
-molCO~2~ molC^-1^
+mol CO~2~ mol C^-1^
 <br><br>
@@
 ```
-CmolH CmolH^-1^ d^-1^
+C-mol H C-mol H^-1^ d^-1^
 <br><br>
@@
 ```
-molC molph^-1^
+mol C mol photons^-1^
 <br><br>
@@
 ```
-m^2^ CmolS^-1^
+m^2^ C-mol S^-1^
 <br><br>
@@
 ```
-mol ph CmolS^-1^ d^-1^
+mol photons C-mol S^-1^ d^-1^
 <br><br>
@@
 ```
-molC CmolS^-1^ d^-1^
+mol C C-mol S^-1^ d^-1^
 <br><br>
@@
 #### $j_{SGm}$: Maximum specific growth rate of symbiont {-}
-We use the maximum specific growth rate of *Symbiodinium* that is reported by @Chang:1983p8066:
+We use the maximum specific growth rate of *Symbiodinium* that is reported by @Chang:1983p8066  
 ```{r jSGm, echo=F}
@@
 ```
-CmolS CmolS^-1^ d^-1^
+C-mol S C-mol S^-1^ d^-1^
 <br><br>
@@
 # Figure S7
-```{r sens_HLHN, cache=T, echo=F, eval=T, fig.cap="Fig. S7. Sensitivity analysis under high DIN (4e-6 M) and high light (30 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
+```{r sens_HLHN, cache=T, echo=F, eval=T, fig.cap="**Fig. S7. Sensitivity analysis under high DIN (4e-6 M) and high light (30 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
 # Set up time vector and parameters
@@
 output1 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
-  run <- run_coral_ss(env=list(L=input$L[i], X=5e-7, N=1e-7), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+  run <- run_coral_ss(env=list(L=input$L[i], X=2e-7, N=1e-7), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
   list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
@@
 output2 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
-  run <- run_coral_ss(env=list(L=input$L[i], X=1e-7, N=1e-7), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+  run <- run_coral_ss(env=list(L=input$L[i], X=0, N=1e-7), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
   list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
@@
 output3 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
-  run <- run_coral_ss(env=list(L=input$L[i], X=5e-7, N=2e-6), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+  run <- run_coral_ss(env=list(L=input$L[i], X=1e-7, N=2e-6), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
   list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
@@
 plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab="Light", ylab="Steady state S:H ratio")
-title("DIN=1e-7 mol/L; X=2e-6 mol/L")
+title("DIN=1e-7 mol/L; X=2e-7 mol/L")
 points(c(48,48,53,53), c(0.42,0.45,0.42,0.45), pch=c(19,1,19,1), cex=c(0.4,1,0.4,1), col=c("black","black","red","red"), xpd=T)
@@
 plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab="Light", ylab="Steady state S:H ratio")
-title("DIN=1e-7 mol/L; X=1e-6 mol/L")
+title("DIN=2e-6 mol/L; X=1e-7 mol/L")
 p2 <- apply(res2, 1, FUN=function(x) {
@@
 plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab="Light", ylab="Steady state S:H ratio")
-title("DIN=1e-6 mol/L; X=2e-6 mol/L")
+title("DIN=2e-6 mol/L; X=1e-6 mol/L")
 p3 <- apply(res3, 1, FUN=function(x) {

### 2017-03-23 JTB_supp/JTB_supp.Rmd
adj fig s9 conditions

@@
 output3 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
-  run <- run_coral_ss(env=list(L=input$L[i], X=1e-7, N=2e-6), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+  run <- run_coral_ss(env=list(L=input$L[i], X=2e-7, N=2e-6), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
   list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
@@
 output4 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
-  run <- run_coral_ss(env=list(L=input$L[i], X=1e-6, N=2e-6), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+  run <- run_coral_ss(env=list(L=input$L[i], X=0, N=2e-6), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
   list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
@@
 plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab="Light", ylab="Steady state S:H ratio")
-title("DIN=2e-6 mol/L; X=1e-7 mol/L")
+title("DIN=1e-7 mol/L; X=0 mol/L")
 p2 <- apply(res2, 1, FUN=function(x) {
@@
 plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab="Light", ylab="Steady state S:H ratio")
-title("DIN=2e-6 mol/L; X=1e-6 mol/L")
+title("DIN=2e-6 mol/L; X=2e-7 mol/L")
 p3 <- apply(res3, 1, FUN=function(x) {
@@
 plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab="Light", ylab="Steady state S:H ratio")
-title("DIN=1e-6 mol/L; X=4e-6 mol/L")
+title("DIN=2e-6 mol/L; X=0 mol/L")
 p4 <- apply(res4, 1, FUN=function(x) {

### 2017-03-23 JTB_supp/JTB_supp.Rmd
FigS9 panel 4 > X=4e-7

@@
 output4 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
-  run <- run_coral_ss(env=list(L=input$L[i], X=0, N=2e-6), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+  run <- run_coral_ss(env=list(L=input$L[i], X=4e-7, N=2e-6), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
   list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
@@
 plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab="Light", ylab="Steady state S:H ratio")
-title("DIN=2e-6 mol/L; X=0 mol/L")
+title("DIN=2e-6 mol/L; X=4e-7 mol/L")
 p4 <- apply(res4, 1, FUN=function(x) {

### 2017-03-23 JTB_supp/JTB_supp.Rmd
FigS8 > side-by-side. code-folding: hide

@@
     toc_depth: 1
+    code_folding: hide
 bibliography: library.bib
@@
 
-plot_run_v2(run1)
-plot_run_v2(run2)
+with(run1, {
+      # External irradiance
+    plot(time, env$L, type="l", col="gold", ylim=c(0,60), lwd=2, xlab="", ylab="mol ph/m2/d")
+    title("External irradiance and DIN", adj=0, line=0.25)
+    par(new=T)
+    plot(time, env$N, type="l", col="blue", ylim=c(0,10.14e-6), lwd=2, axes=F, xlab="", ylab="")
+    axis(side=4); mtext(side=4, text="molN/L", line=1, cex=0.75)
+
+    # Symbiont to host biomass ratio
+    totSH <- S$S / H$H
+    totSHf <- totSH[length(totSH)]
+    plot(time, totSH, type="l", col="black", ylim=c(0, 0.25), ylab="CmolS/CmolH", xlab="", lwd=2)
+    text(time[0.95*length(time)], totSHf, labels=as.character(round(totSHf, 3)), pos=3, xpd=T, cex=0.75)
+    title("Symbiont:host biomass", adj=0, line=0.25)
+
+    # Specific growth rates of host and symbiont
+    Hgrf <- H$dH.Hdt[length(H$dH.Hdt)]
+    plot(time, H$dH.Hdt, type="l", ylim=c(-0.01, 0.13), 
+         xlab="", ylab="d-1", lwd=2, cex=1, cex.lab=1)
+    if(any(c(H$dH.Hdt[-(1:10)], S$dS.Sdt[-(1:10)])<0)) abline(h=0, lwd=0.5, lty=3)
+    text(time[0.95*length(time)], Hgrf, labels=as.character(round(Hgrf, 3)), pos=3, xpd=T, cex=0.75)
+    #title("State variable dynamics", adj=0, cex.main=2, outer = T)
+    title("Specific growth rates", adj=0, line=0.25)
+    lines(time, S$dS.Sdt, col="black", lwd=1)
+    legend("topright", legend=c("Host", "Sym"), lwd=c(2,1), col="black", bty="n", y.intersp=1, cex=0.75)
+
+    # Light quenching (carbon fixation, NPQ, and excess (=ROS producing))
+    plot(NA, xlim=range(time), xlab="", ylim=c(0,130), ylab="mol ph/CmolS/d")
+    title("Light quenching", adj=0, line=0.25)
+    lines(time, S$jCP/pars$yCL, col="gold", lwd=1) # amt. quenched by photosynthesis
+    lines(time, S$jCP/pars$yCL + S$jNPQ, col="gold", lty=1, lwd=2) # amt. quenched by photo + NPQ
+    lines(time, S$jL, col="gold", lty=3, lwd=2) # total amt. absorbed
+    legend("topright", legend=c("ROS", "NPQ", "Photo."), lty=c(3,1,1), lwd=c(2,2,1), col="gold", bty="n", cex=0.75)
+
+    # Photosynthesis: substrate-limitation
+    pl <- log(   pmin((H$jCO2 + H$rCH)*H$H/S$S + S$rCS, pars$jCPm)   /   pmin(S$jL * pars$yCL, pars$jCPm)    )
+    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-2, 2))
+    title("Photosynthesis: substrate-limitation", adj=0, line=0.25)
+    lines(time, pl, col="black", lty=1, lwd=2)
+    abline(h=0, lty=3)
+    text(par("usr")[2], 0, labels="L-lim.\nCO2-lim.", cex=0.75, adj=1)
+
+    # Biomass formation - substrate-limitation
+    sl <- log(  pmin(pars$yC*S$jCP, pars$jSGm)   /   pmin((H$rhoN*H$H/S$S + S$rNS)/pars$nNS, pars$jSGm)  )
+    hl <- log(  pmin(pars$yC*S$rhoC*S$S/H$H + H$jX, pars$jHGm) / pmin((H$jN + pars$nNX*H$jX + H$rNH) / pars$nNH, pars$jHGm)  )
+    maxabs <- max(c(abs(hl), abs(sl)))
+    range <- range(c(hl, sl))
+    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-1.5,1.5))
+    title("Biomass: substrate-limitation", adj=0, line=0.25)
+    lines(time, hl, col="black", lty=1, lwd=2)
+    lines(time, sl, col="black", lty=1, lwd=1)
+    abline(h=0, lty=3)
+    text(par("usr")[2], 0, labels="N-lim.\nC-lim.", cex=0.75, adj=1)
+    legend("topright", legend=c("Host", "Sym"), lty=1, lwd=c(2,1), col="black", bty="n", cex=0.75)
+})
+
+with(run2, {
+      # External irradiance
+    plot(time, env$L, type="l", col="gold", ylim=c(0,60), lwd=2, xlab="", ylab="mol/m2/d")
+    title("External irradiance and DIN", adj=0, line=0.25)
+    par(new=T)
+    plot(time, env$N, type="l", col="blue", ylim=c(0,10.14e-6), lwd=2, axes=F, xlab="", ylab="")
+    axis(side=4); mtext(side=4, text="molN/L", line=1, cex=0.75)
+
+    # Symbiont to host biomass ratio
+    totSH <- S$S / H$H
+    totSHf <- totSH[length(totSH)]
+    plot(time, totSH, type="l", col="black", ylim=c(0, 0.25), ylab="CmolS/CmolH", xlab="", lwd=2)
+    text(time[0.95*length(time)], totSHf, labels=as.character(round(totSHf, 3)), pos=3, xpd=T, cex=0.75)
+    title("Symbiont:host biomass", adj=0, line=0.25)
+
+    # Specific growth rates of host and symbiont
+    Hgrf <- H$dH.Hdt[length(H$dH.Hdt)]
+    plot(time, H$dH.Hdt, type="l", ylim=c(-0.01,0.13), 
+         xlab="", ylab="d-1", lwd=2, cex=1, cex.lab=1)
+    if(any(c(H$dH.Hdt[-(1:10)], S$dS.Sdt[-(1:10)])<0)) abline(h=0, lwd=0.5, lty=3)
+    text(time[0.95*length(time)], Hgrf, labels=as.character(round(Hgrf, 3)), pos=3, xpd=T, cex=0.75)
+    #title("State variable dynamics", adj=0, cex.main=2, outer = T)
+    title("Specific growth rates", adj=0, line=0.25)
+    lines(time, S$dS.Sdt, col="black", lwd=1)
+    legend("topright", legend=c("Host", "Sym"), lwd=c(2,1), col="black", bty="n", y.intersp=1, cex=0.75)
+
+    # Light quenching (carbon fixation, NPQ, and excess (=ROS producing))
+    plot(NA, xlim=range(time), xlab="", ylim=c(0,130), ylab="mol ph/CmolS/d")
+    title("Light quenching", adj=0, line=0.25)
+    lines(time, S$jCP/pars$yCL, col="gold", lwd=1) # amt. quenched by photosynthesis
+    lines(time, S$jCP/pars$yCL + S$jNPQ, col="gold", lty=1, lwd=2) # amt. quenched by photo + NPQ
+    lines(time, S$jL, col="gold", lty=3, lwd=2) # total amt. absorbed
+    legend("topright", legend=c("ROS", "NPQ", "Photo."), lty=c(3,1,1), lwd=c(2,2,1), col="gold", bty="n", cex=0.75)
+
+    # Photosynthesis: substrate-limitation
+    pl <- log(   pmin((H$jCO2 + H$rCH)*H$H/S$S + S$rCS, pars$jCPm)   /   pmin(S$jL * pars$yCL, pars$jCPm)    )
+    maxabs <- max(abs(pl))
+    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-2, 2))
+    title("Photosynthesis: substrate-limitation", adj=0, line=0.25)
+    lines(time, pl, col="black", lty=1, lwd=2)
+    abline(h=0, lty=3)
+    text(par("usr")[2], 0, labels="L-lim.\nCO2-lim.", cex=0.75, adj=1)
+
+    # Biomass formation - substrate-limitation
+    sl <- log(  pmin(pars$yC*S$jCP, pars$jSGm)   /   pmin((H$rhoN*H$H/S$S + S$rNS)/pars$nNS, pars$jSGm)  )
+    hl <- log(  pmin(pars$yC*S$rhoC*S$S/H$H + H$jX, pars$jHGm) / pmin((H$jN + pars$nNX*H$jX + H$rNH) / pars$nNH, pars$jHGm)  )
+    maxabs <- max(c(abs(hl), abs(sl)))
+    range <- range(c(hl, sl))
+    plot(NA, xlim=range(time), xlab="", ylab="", ylim=c(-1.5,1.5))
+    title("Biomass: substrate-limitation", adj=0, line=0.25)
+    lines(time, hl, col="black", lty=1, lwd=2)
+    lines(time, sl, col="black", lty=1, lwd=1)
+    abline(h=0, lty=3)
+    text(par("usr")[2], 0, labels="N-lim.\nC-lim.", cex=0.75, adj=1)
+    legend("topright", legend=c("Host", "Sym"), lty=1, lwd=c(2,1), col="black", bty="n", cex=0.75)
+})
 ```

### 2017-03-26 ms/ms.Rmd
update filenames

--- 2017-03-23 JTB_ms/JTB_ms.Rmd
@@
     email: ross.cunning@gmail.com
-    affiliation: University of Hawaii
+    affiliation: HIMB
     footnote: Corresponding Author
   - name: Erik B. Muller
-    affiliation: UCSB, MSI
+    affiliation: UCSB,MSI
   - name: Ruth D. Gates
@@
 
+# Acknowledgements
+This work was supported by a NSF Postdoctoral Research Fellowship in Biology to RC (#1400787).
+
 References {#references .unnumbered}

### 2017-03-26 ms/ms.Rmd
update links

@@
 # Source default parameter values
-source("R/def_pars.R")
-defpars <- def_pars()
+defpars <- coRal::def_pars()
 ```
@@
 3.
-_Provide well-documented, open source code._ In order to facilitate the continued development and application of theoretical modeling tools for coral symbioses, we provide open access to the model in the form of detailed, commented code written in the R language [@RALanguageandEn:2014wf].
-With an accessible and modular framework, we envision this model as a resource for further development by the scientific community to include additional complexity and problem-specific components.
-We chose R, an open source programming language in common use by biologists and ecologists, to reach the widest possible audience with this work.
+_Provide well-documented, open source code._ In order to facilitate the continued development and application of theoretical modeling tools for coral symbioses, we provide access to the model in the form of an R package called coRal [github.com/jrcunning/coRal](github.com/jrcunning/coRal), which users may install to run and visualize model simulations.
+With an accessible and modular framework, we envision this code base as a resource for further development by the scientific community to include additional complexity and problem-specific components.
+We chose R [@RALanguageandEn:2014wf], an open source programming language in common use by biologists and ecologists, to reach the widest possible audience with this work.
 
@@
 # Model description
+
 In this dynamical system, both the coral host and algal symbiont acquire and use carbon and nitrogen to construct biomass.
@@
 
+To aid in visualizing model results, we calculated values to indicate the degree to which product formation at an SU was limited by availability of either of its two substrates using the formula
+
+\begin{equation} \log \bigg({{\min(j_{S1}, j_{Pm})} \over {\min(j_{S2}, j_{Pm})}}\bigg) \end{equation}
+
+where $j_{S1}$ and $j_{S2}$ are the specific input fluxes of the two substrates and $j_{Pm}$ is the maximum specific product formation rate, in units of Cmol Cmol^-1^ d^-1^.
+When both substrate input fluxes are higher than what can be used at the maximum production rate, this limitation coefficient is zero, implying that neither substrate is limiting production.
+
 ## Numerical analysis
@@
 
-Simulations using the discretized scheme were performed using R code that is available in the data repository accompanying this article: github.com/jrcunning/Rcoral.
+Simulations using the discretized scheme were performed using R code developed in the coRal R package [github.com/jrcunning/coRal](github.com/jrcunning/coRal).
 By experimentation, we found that a time step of 0.1 days gave adequate precision for most simulations (including used to generate Figs.
 2-8 in this paper).
-For steady state estimations, simulations were run until the change in specific growth rate of host and symbiont and the S:H biomass ratio was less than 1e-5 per time step.
+For steady state estimations, simulations were run until the changes in specific growth rate of the host and the S:H biomass ratio were less than 1e-5 per time step.
 In regions of state space where very slow transient dynamics could be expected (i.e.
 near bifurcation points), sample steady state calculations were verified using MATHEMATICA code for numerical root finding (function FindRoot) with the code written independently by a coauthor without reference to the R code.
-
-To aid in visualizing model results, we calculated values to indicate the degree to which product formation at an SU was limited by availability of either of its two substrates using the formula
-
-\begin{equation} \log \bigg({{\min(j_{S1}, j_{Pm})} \over {\min(j_{S2}, j_{Pm})}}\bigg) \end{equation}
-
-where $j_{S1}$ and $j_{S2}$ are the specific input fluxes of the two substrates and $j_{Pm}$ is the maximum specific product formation rate, in units of Cmol Cmol^-1^ d^-1^.
-When both substrate input fluxes are higher than what can be used at the maximum production rate, this limitation coefficient is zero, implying that neither substrate is limiting production.
+All of the R code for the simulations and figures presented in this paper can be found in the accompanying data repository at [github.com/jrcunning/coRal-analysis](github.com/jrcunning/coRal-analysis).
 
 # Steady state behavior
+
 In a constant environment, the system ultimately reaches a steady state of exponential growth or decline.
@@
 # Dynamic behavior
+
 The dynamic behavior of the model demonstrates its power to integrate multiple environmental forcings simultaneously.
@@
 ## Seasonal variability
+
 Symbiont densities and coral growth rates are known to vary seasonally, representing an integrated response to changes in a suite of environmental factors.
@@
 ## Coral bleaching and recovery
+
 Coral bleaching is the stress-induced loss of symbiotic algae from coral tissues, which can occur in response to a variety of environmental stressors.
@@
 # Conclusions
+
 This dynamic bioenergetic model of coral-*Symbiodinium* symbioses mechanistically reproduces patterns in steady-state coral growth and symbiont abundance commonly observed in corals, including higher symbiont abundance with higher nutrients and feeding, lower symbiont abundance with increasing light, and optimal growth at intermediate light levels.
@@
 # Acknowledgements
+
 This work was supported by a NSF Postdoctoral Research Fellowship in Biology to RC (#1400787).

### 2017-03-26 ms/ms.tex
update filenames
update links

--- 2017-03-22 JTB_ms/JTB_ms.tex
@@
 coral bleaching as an alternate stable state}
-    \author[University of Hawaii]{Ross Cunning\corref{c1}}
+    \author[HIMB]{Ross Cunning\corref{c1}}
    \ead{ross.cunning@gmail.com} 
    \cortext[c1]{Corresponding Author}
-    \author[UCSB, MSI]{Erik B. Muller}
+    \author[UCSB,MSI]{Erik B. Muller}
 
@@
   facilitate the continued development and application of theoretical
-  modeling tools for coral symbioses, we provide open access to the
-  model in the form of detailed, commented code written in the R
-  language (R Core Team 2014). With an accessible and modular framework,
-  we envision this model as a resource for further development by the
-  scientific community to include additional complexity and
-  problem-specific components. We chose R, an open source programming
-  language in common use by biologists and ecologists, to reach the
-  widest possible audience with this work.
+  modeling tools for coral symbioses, we provide access to the model in
+  the form of an R package called coRal
+  \url{github.com/jrcunning/coRal}, which users may install to run and
+  visualize model simulations. With an accessible and modular framework,
+  we envision this code base as a resource for further development by
+  the scientific community to include additional complexity and
+  problem-specific components. We chose R (R Core Team 2014), an open
+  source programming language in common use by biologists and
+  ecologists, to reach the widest possible audience with this work.
 \end{enumerate}
@@
 
-\begin{longtable}[c]{@{}llll@{}}
+\begin{longtable}[]{@{}llll@{}}
 \toprule
-\begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
-Symbol
-\strut\end{minipage} &
-\begin{minipage}[b]{0.45\columnwidth}\raggedright\strut
-Description
-\strut\end{minipage} &
-\begin{minipage}[b]{0.27\columnwidth}\raggedright\strut
-Units
-\strut\end{minipage} &
-\begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
-Eq. no.
-\strut\end{minipage}\tabularnewline
+\begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
+Symbol\strut
+\end{minipage} & \begin{minipage}[b]{0.44\columnwidth}\raggedright\strut
+Description\strut
+\end{minipage} & \begin{minipage}[b]{0.26\columnwidth}\raggedright\strut
+Units\strut
+\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
+Eq. no.\strut
+\end{minipage}\tabularnewline
 \midrule
 \endhead
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_X\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Prey assimilation (feeding) rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-C-mol X C-mol H\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-3
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_N\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Nitrogen uptake rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-mol N Cmol H\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-4
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{HG}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Host biomass formation rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-5
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{HT}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Host biomass turnover rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-6
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(r_{NH}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Recycled nitrogen from host turnover
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-mol N C-mol H\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-7
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(\rho_N\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Nitrogen shared with the symbiont
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-mol N C-mol H\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-8
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{eC}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Excess carbon used to activate host CCMs
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-mol C C-mol H\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-9
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{CO_2}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-CO\textsubscript{2} input to photosynthesis
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_X\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Prey assimilation (feeding) rate\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+C-mol X C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+3\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_N\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Nitrogen uptake rate\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol N Cmol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+4\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{HG}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Host biomass formation rate\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+5\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{HT}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Host biomass turnover rate\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+6\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(r_{NH}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Recycled nitrogen from host turnover\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol N C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+7\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(\rho_N\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Nitrogen shared with the symbiont\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol N C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+8\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{eC}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Excess carbon used to activate host CCMs\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol C C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+9\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{CO_2}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+CO\textsubscript{2} input to photosynthesis\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
 mol CO\textsubscript{2} C-mol H\textsuperscript{-1}
-d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-10
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_L\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Light absorption rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-12
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(r_{CH}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Recycled CO\textsubscript{2} from host
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+10\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_L\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Light absorption rate\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+12\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(r_{CH}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Recycled CO\textsubscript{2} from host\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
 mol CO\textsubscript{2} C-mol H\textsuperscript{-1}
-d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-13
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(r_{CS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Recycled CO\textsubscript{2} from symbiont
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+13\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(r_{CS}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Recycled CO\textsubscript{2} from symbiont\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
 mol CO\textsubscript{2} C-mol S\textsuperscript{-1}
-d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-14
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{CP}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Photosynthesis rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-mol C C-mol S\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-15
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{eL}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Light energy in excess of photochemistry
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-16
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{NPQ}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Total capacity of NPQ
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-17
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(c_{ROS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-ROS production proportional to baseline
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-18
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(r_{NS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Recycled nitrogen from symbiont turnover
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-mol N C-mol S\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-19
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{SG}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Symbiont biomass formation rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-20
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(\rho_C\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Fixed carbon shared with host
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-mol C C-mol S\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-21
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{ST}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Symbiont biomass turnover rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-22
-\strut\end{minipage}\tabularnewline
+d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+14\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{CP}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Photosynthesis rate\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol C C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+15\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{eL}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Light energy in excess of photochemistry\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+16\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{NPQ}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Total capacity of NPQ\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+17\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(c_{ROS}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+ROS production proportional to baseline\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+--\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+18\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(r_{NS}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Recycled nitrogen from symbiont turnover\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol N C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+19\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{SG}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Symbiont biomass formation rate\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+20\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(\rho_C\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Fixed carbon shared with host\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+mol C C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+21\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{ST}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Symbiont biomass turnover rate\strut
+\end{minipage} & \begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
+C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+22\strut
+\end{minipage}\tabularnewline
 \bottomrule
@@
 
-\begin{longtable}[c]{@{}llll@{}}
+\begin{longtable}[]{@{}llll@{}}
 \toprule
-\begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
-Symbol
-\strut\end{minipage} &
-\begin{minipage}[b]{0.45\columnwidth}\raggedright\strut
-Description
-\strut\end{minipage} &
-\begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
-Units
-\strut\end{minipage} &
-\begin{minipage}[b]{0.12\columnwidth}\raggedright\strut
-Value
-\strut\end{minipage}\tabularnewline
+\begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
+Symbol\strut
+\end{minipage} & \begin{minipage}[b]{0.44\columnwidth}\raggedright\strut
+Description\strut
+\end{minipage} & \begin{minipage}[b]{0.24\columnwidth}\raggedright\strut
+Units\strut
+\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\raggedright\strut
+Value\strut
+\end{minipage}\tabularnewline
 \midrule
 \endhead
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(n_{NH}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-N:C molar ratio in host biomass
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.18\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(n_{NS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-N:C molar ratio in symbiont biomass
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.13\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(n_{NX}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-N:C molar ratio in prey biomass
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.2\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{HT}^0\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Maintenance rate of host biomass
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.03\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{ST}^0\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Maintenance rate of symbiont biomass
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.03\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(\sigma_{NH}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Proportion N turnover recycled in host
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.9\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(\sigma_{CH}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Proportion host metabolic CO\textsubscript{2} recycled to photosynthesis
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.1\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(\sigma_{NS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Proportion N turnover recycled in symbiont
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.9\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(\sigma_{CS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(n_{NH}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+N:C molar ratio in host biomass\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+--\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.18\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(n_{NS}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+N:C molar ratio in symbiont biomass\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+--\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.13\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(n_{NX}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+N:C molar ratio in prey biomass\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+--\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.2\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{HT}^0\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Maintenance rate of host biomass\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.03\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{ST}^0\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Maintenance rate of symbiont biomass\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.03\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(\sigma_{NH}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Proportion N turnover recycled in host\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+--\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.9\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(\sigma_{CH}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Proportion host metabolic CO\textsubscript{2} recycled to
+photosynthesis\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+--\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.1\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(\sigma_{NS}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Proportion N turnover recycled in symbiont\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+--\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.9\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(\sigma_{CS}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Proportion symbiont metabolic CO\textsubscript{2} recycled to
-photosynthesis
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.9\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{Xm}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Maximum prey assimilation rate from host feeding
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molX C-mol H\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.13\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(K_X\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Half-saturation constant for prey assimilation
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molX L\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(10^{-6}\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{Nm}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Maximum host DIN uptake rate
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-mol N C-mol H\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.035\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(K_N\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Half-saturation constant for host DIN uptake
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-mol N L\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(1.5\times 10^{-6}\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(k_{CO_2}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Efficacy of CO\textsubscript{2} delivery to photosynthesis by host CCMs
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-mol CO\textsubscript{2} mol C\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(10\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{HGm}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Maximum specific growth rate of host
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(1\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(y_{CL}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Quantum yield of photosynthesis
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-mol C mol photons\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.1\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(y_{C}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Yield of biomass formation from carbon
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-C-mol mol C\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.8\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(\bar{a}^*\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Effective light-absorbing cross-section of symbiont
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-m\textsuperscript{2} C-mol S\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(1.34\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(k_{NPQ}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-NPQ capacity of symbiont
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(112\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(k_{ROS}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+photosynthesis\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+--\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.9\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{Xm}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Maximum prey assimilation rate from host feeding\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+C-mol X C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.13\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(K_X\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Half-saturation constant for prey assimilation\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+C-mol X L\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(10^{-6}\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{Nm}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Maximum host DIN uptake rate\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+mol N C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.035\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(K_N\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Half-saturation constant for host DIN uptake\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+mol N L\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(1.5\times 10^{-6}\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(k_{CO_2}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Efficacy of CO\textsubscript{2} delivery to photosynthesis by host
+CCMs\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+mol CO\textsubscript{2} mol C\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(10\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{HGm}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Maximum specific growth rate of host\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+C-mol H C-mol H\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(1\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(y_{CL}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Quantum yield of photosynthesis\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+mol C mol photons\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.1\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(y_{C}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Yield of biomass formation from carbon\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+C-mol mol C\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.8\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(\bar{a}^*\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Effective light-absorbing cross-section of symbiont\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+m\textsuperscript{2} C-mol S\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(1.34\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(k_{NPQ}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+NPQ capacity of symbiont\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(112\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(k_{ROS}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
 Excess photon energy that doubles ROS production, relative to baseline
-levels
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(80\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{CPm}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Maximum specific photosynthesis rate of symbiont
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-mol C C-mol S\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(2.8\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(j_{SGm}\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Maximum specific growth rate of symbiont
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(0.25\)
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-\(b\)
-\strut\end{minipage} &
-\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
-Scaling parameter for bleaching response
-\strut\end{minipage} &
-\begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
---
-\strut\end{minipage} &
-\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-\(5\)
-\strut\end{minipage}\tabularnewline
+levels\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+mol photons C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(80\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{CPm}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Maximum specific photosynthesis rate of symbiont\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+mol C C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(2.8\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(j_{SGm}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Maximum specific growth rate of symbiont\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+C-mol S C-mol S\textsuperscript{-1} d\textsuperscript{-1}\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(0.25\)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\(b\)\strut
+\end{minipage} & \begin{minipage}[t]{0.44\columnwidth}\raggedright\strut
+Scaling parameter for bleaching response\strut
+\end{minipage} & \begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
+--\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\(5\)\strut
+\end{minipage}\tabularnewline
 \bottomrule
@@
 is included to scale biomass loss due to bleaching in response to ROS.
+
+To aid in visualizing model results, we calculated values to indicate
+the degree to which product formation at an SU was limited by
+availability of either of its two substrates using the formula
+
+\begin{equation} \log \bigg({{\min(j_{S1}, j_{Pm})} \over {\min(j_{S2}, j_{Pm})}}\bigg) \end{equation}
+
+where \(j_{S1}\) and \(j_{S2}\) are the specific input fluxes of the two
+substrates and \(j_{Pm}\) is the maximum specific product formation
+rate, in units of Cmol Cmol\textsuperscript{-1} d\textsuperscript{-1}.
+When both substrate input fluxes are higher than what can be used at the
+maximum production rate, this limitation coefficient is zero, implying
+that neither substrate is limiting production.
 
@@
 Simulations using the discretized scheme were performed using R code
-that is available in the data repository accompanying this article:
-github.com/jrcunning/Rcoral. By experimentation, we found that a time
-step of 0.1 days gave adequate precision for most simulations (including
-used to generate Figs. 2-8 in this paper). For steady state estimations,
-simulations were run until the change in specific growth rate of host
-and symbiont and the S:H biomass ratio was less than 1e-5 per time step.
-In regions of state space where very slow transient dynamics could be
-expected (i.e.~near bifurcation points), sample steady state
-calculations were verified using MATHEMATICA code for numerical root
-finding (function FindRoot) with the code written independently by a
-coauthor without reference to the R code.
-
-To aid in visualizing model results, we calculated values to indicate
-the degree to which product formation at an SU was limited by
-availability of either of its two substrates using the formula
-
-\begin{equation} \log \bigg({{\min(j_{S1}, j_{Pm})} \over {\min(j_{S2}, j_{Pm})}}\bigg) \end{equation}
-
-where \(j_{S1}\) and \(j_{S2}\) are the specific input fluxes of the two
-substrates and \(j_{Pm}\) is the maximum specific product formation
-rate, in units of Cmol Cmol\textsuperscript{-1} d\textsuperscript{-1}.
-When both substrate input fluxes are higher than what can be used at the
-maximum production rate, this limitation coefficient is zero, implying
-that neither substrate is limiting production.
+developed in the coRal R package \url{github.com/jrcunning/coRal}. By
+experimentation, we found that a time step of 0.1 days gave adequate
+precision for most simulations (including used to generate Figs. 2-8 in
+this paper). For steady state estimations, simulations were run until
+the changes in specific growth rate of the host and the S:H biomass
+ratio were less than 1e-5 per time step. In regions of state space where
+very slow transient dynamics could be expected (i.e.~near bifurcation
+points), sample steady state calculations were verified using
+MATHEMATICA code for numerical root finding (function FindRoot) with the
+code written independently by a coauthor without reference to the R
+code. All of the R code for the simulations and figures presented in
+this paper can be found in the accompanying data repository at
+\url{github.com/jrcunning/coRal-analysis}.
 
@@
 
+\section{Acknowledgements}\label{acknowledgements}
+
+This work was supported by a NSF Postdoctoral Research Fellowship in
+Biology to RC (\#1400787).
+
 \section*{References}\label{references}

### 2017-07-11 ms/ms.Rmd
revisions

@@
 We use C-moles (C-mol) as the unit of biomass for consistency with the rigorous mass balance of DEB theory: 1 C-mol is equivalent to the amount of biomass containing 1 mole of carbon atoms.
-Host biomass (expressed as C-mol H), symbiont biomass (expressed as C-mol S), and prey biomass (expressed as C-mol X) have fixed, but different, molar N:C ratios (Table 2).
-Biomass is produced from carbon and nitrogen by synthesizing units (SU), which are mathematical specifications of the formation of a product from two substrates; we use the "parallel complementary" formulation of @Kooijman:2010vd (page 105, Fig.
-3.7) to specify these fluxes.
-The two state variables of this system are symbiont biomass and coral biomass; because resources are acquired proportionally to surface area, and surface area is proportional to volume, biomass increases exponentially during growth (indeed, corals grow exponentially [@Bak:1976bv]).
+Host biomass (expressed as C-mol H), symbiont biomass (expressed as C-mol S), and prey biomass (expressed as C-mol X) have fixed, but different, molar N:C ratios (Table 2), consistent with the assumption of strong homeostasis in DEB theory [@Kooijman:2010vd].
+Biomass is produced from carbon and nitrogen by parallel complementary synthesizing units (SUs) -- mathematical specifications of the formation of a product by a metabolic network processing two complementary substrates in parallel (@Kooijman:2010vd, p.
+105, Fig.
+3.7).
+The two state variables in the model are symbiont biomass and coral biomass; because resources are acquired proportionally to surface area, and surface area is proportional to volume (corals are V1-morphs in a DEB context (@Kooijman:2010vd, p.
+122)), biomass increases exponentially during growth (indeed, corals grow exponentially [@Bak:1976bv]).
 
@@
 
-Table 1. Model fluxes (mass-specific). Units are explained in the text.
+Table 1. Model fluxes (mass-specific). Units are explained in the text. (Abbreviations: CCMs=carbon concentrating mechanisms; NPQ=non-photochemical quenching; ROS=reactive oxygen species)
 
@@
 
-Table 2. Model parameters (units explained in the text). Justification and/or derivation of each parameter value along with supporting references are provided in the Supplementary Information.
+Table 2.
+Model parameters (units explained in the text).
+Justification and/or derivation of each parameter value along with supporting references are provided in the Supplementary Information.
+(Abbreviations: CCMs=carbon concentrating mechanisms; NPQ=non-photochemical quenching; ROS=reactive oxygen species; DIN=dissolved inorganic nitrogen)
 
@@
 
-Coral biomass formation is then specified by a parallel complementary SU (formula in Fig. 3.7 of @Kooijman:2010vd]) that combines carbon and nitrogen, according to:
+Coral biomass formation is then specified by a parallel complementary SU (formula in Fig.
+3.7 of @Kooijman:2010vd]).
+The general form of this equation is $(m^{-1} + x^{-1} + y^{-1} - (x + y)^{-1})^{-1}$, where $m$ is the maximum production flux, and $x$ and $y$ are the input fluxes of the two substrates.
+Coral biomass is thus produced from carbon and nitrogen according to:
 
@@
 The amount of light absorbed by the symbiont depends on the scalar irradiance at the site of light absorption, which is modified substantially relative to external downwelling irradiance owing to multiple scattering by the coral skeleton and self-shading by surrounding symbionts [@Enriquez:2005p142; @Marcelino:2013hz].
-We used data from @Marcelino:2013hz to empirically derive an amplification factor, $A$, indicating the ratio of internal scalar irradiance to external downwelling irradiance as a function of symbiont density (S:H biomass), which is specified as:
+We used skeletal light amplification measurements from @Marcelino:2013hz to empirically derive an amplification factor, $A$ (Figure S1), indicating the ratio of internal scalar irradiance to external downwelling irradiance as a function of symbiont density (S:H biomass), which is specified as:
 
@@
 
-[^2]: Note that recycling of symbiont biomass turnover ($rNS$ and $rCS$) only occurs based on the maintenance component of turnover (i.e., $j_{ST}^0$), and not the photodamage/bleaching component, as this loss represents biomass that is expelled from the host.
+[^2]: Note that recycling of symbiont biomass turnover ($r_{NS}$ and $r_{CS}$) only occurs based on the maintenance component of turnover (i.e., $j_{ST}^0$), and not the photodamage/bleaching component, as this loss represents biomass that is expelled from the host.
 [^3]: See footnote 2.
@@
 ![Sensitivity analysis.
-Plots show the fractional change in steady state values of growth (solid lines) and S:H biomass (dashed lines) in response to fractional changes in default parameter values (see Table 2 for default values).
+Plots show the fractional change in steady state values of growth rate (solid lines) and S:H biomass (dashed lines) in response to fractional changes in default parameter values (see Table 2 for default values).
 Parameters are grouped by the processes in which they are involved.
@@
 Seasonal summertime reductions in tissue biomass have also been well-documented in the field [@Fitt:2000p3792], along with reductions in net photosynthetic capacity [@MullerParker:1987cr].
-Importantly, while light alone may drive seasonal dynamics in the ways discussed[^5], temperature fluctuations may attenuate or even reverse the effect of light as cooler winters depress metabolism; thus, the relative magnitude of fluctuation in temperature, light, and other factors may produce wide variability in the direction and magnitude of seasonal changes in growth and symbiont abundance, depending on location and microhabitat.
-Nevertheless, the seasonal variability predicted here (Fig.
+Importantly, while light alone may drive seasonal dynamics in the ways discussed[^5], temperature fluctuations may attenuate or even reverse the effect of light as cooler winters depress metabolism.
+Seasonal changes in nutrients associated with fertilization use and runoff during growing seasons may also impact these dynamics.
+Thus, the relative magnitude of fluctuation in temperature, light, and nutrients may produce wide variability in the direction and magnitude of seasonal changes in growth and symbiont abundance, depending on location and microhabitat.
+Nevertheless, the seasonal variability predicted here in response to light (Fig.
 5) is consistent with experimental and field observations for corals, and demonstrates the model's ability to predict dynamic behavior that mechanistically integrates multiple environmental drivers.

### 2017-07-11 ms/ms.tex
revisions

@@
 as C-mol S), and prey biomass (expressed as C-mol X) have fixed, but
-different, molar N:C ratios (Table 2). Biomass is produced from carbon
-and nitrogen by synthesizing units (SU), which are mathematical
-specifications of the formation of a product from two substrates; we use
-the ``parallel complementary'' formulation of Kooijman (2010) (page 105,
-Fig. 3.7) to specify these fluxes. The two state variables of this
-system are symbiont biomass and coral biomass; because resources are
+different, molar N:C ratios (Table 2), consistent with the assumption of
+strong homeostasis in DEB theory (Kooijman 2010). Biomass is produced
+from carbon and nitrogen by parallel complementary synthesizing units
+(SUs) -- mathematical specifications of the formation of a product by a
+metabolic network processing two complementary substrates in parallel
+(Kooijman (2010), p.~105, Fig. 3.7). The two state variables in the
+model are symbiont biomass and coral biomass; because resources are
 acquired proportionally to surface area, and surface area is
-proportional to volume, biomass increases exponentially during growth
-(indeed, corals grow exponentially (Bak 1976)).
+proportional to volume (corals are V1-morphs in a DEB context (Kooijman
+(2010), p.~122)), biomass increases exponentially during growth (indeed,
+corals grow exponentially (Bak 1976)).
 
@@
 Table 1. Model fluxes (mass-specific). Units are explained in the text.
+(Abbreviations: CCMs=carbon concentrating mechanisms;
+NPQ=non-photochemical quenching; ROS=reactive oxygen species)
 
@@
 references are provided in the Supplementary Information.
+(Abbreviations: CCMs=carbon concentrating mechanisms;
+NPQ=non-photochemical quenching; ROS=reactive oxygen species;
+DIN=dissolved inorganic nitrogen)
 
@@
 Coral biomass formation is then specified by a parallel complementary SU
-(formula in Fig. 3.7 of Kooijman (2010){]}) that combines carbon and
-nitrogen, according to:
+(formula in Fig. 3.7 of Kooijman (2010){]}). The general form of this
+equation is \((m^{-1} + x^{-1} + y^{-1} - (x + y)^{-1})^{-1}\), where
+\(m\) is the maximum production flux, and \(x\) and \(y\) are the input
+fluxes of the two substrates. Coral biomass is thus produced from carbon
+and nitrogen according to:
 
@@
 skeleton and self-shading by surrounding symbionts (Enríquez, Méndez,
-and Iglesias-Prieto 2005; Marcelino et al. 2013). We used data from
-Marcelino et al. (2013) to empirically derive an amplification factor,
-\(A\), indicating the ratio of internal scalar irradiance to external
-downwelling irradiance as a function of symbiont density (S:H biomass),
-which is specified as:
+and Iglesias-Prieto 2005; Marcelino et al. 2013). We used skeletal light
+amplification measurements from Marcelino et al. (2013) to empirically
+derive an amplification factor, \(A\) (Figure S1), indicating the ratio
+of internal scalar irradiance to external downwelling irradiance as a
+function of symbiont density (S:H biomass), which is specified as:
 
@@
 symbiont biomass turnover\footnote{Note that recycling of symbiont
-  biomass turnover (\(rNS\) and \(rCS\)) only occurs based on the
+  biomass turnover (\(r_{NS}\) and \(r_{CS}\)) only occurs based on the
   maintenance component of turnover (i.e., \(j_{ST}^0\)), and not the
@@
 \caption{Sensitivity analysis. Plots show the fractional change in
-steady state values of growth (solid lines) and S:H biomass (dashed
+steady state values of growth rate (solid lines) and S:H biomass (dashed
 lines) in response to fractional changes in default parameter values
@@
 fluctuations may attenuate or even reverse the effect of light as cooler
-winters depress metabolism; thus, the relative magnitude of fluctuation
-in temperature, light, and other factors may produce wide variability in
-the direction and magnitude of seasonal changes in growth and symbiont
+winters depress metabolism. Seasonal changes in nutrients associated
+with fertilization use and runoff during growing seasons may also impact
+these dynamics. Thus, the relative magnitude of fluctuation in
+temperature, light, and nutrients may produce wide variability in the
+direction and magnitude of seasonal changes in growth and symbiont
 abundance, depending on location and microhabitat. Nevertheless, the
-seasonal variability predicted here (Fig. 5) is consistent with
-experimental and field observations for corals, and demonstrates the
-model's ability to predict dynamic behavior that mechanistically
-integrates multiple environmental drivers.
+seasonal variability predicted here in response to light (Fig. 5) is
+consistent with experimental and field observations for corals, and
+demonstrates the model's ability to predict dynamic behavior that
+mechanistically integrates multiple environmental drivers.
 

### 2017-07-11 ms/ms.Rmd
Fig8 -> Fig9, FigS9 -> Fig8

@@
 The conditions under which recovery can occur -- which determine the magnitude of hysteresis (Fig.
-S9) -- depend on the relative abundance of nitrogen and carbon in the environment.
-Higher food levels, representing a non-autotrophic carbon source for the host, make it easier for the host to overcome carbon-limitation, thus providing a potential mechanism underlying observations that feeding aids recovery from bleaching [@Grottoli:2006p3692; @Connolly:2012ep].
-Conversely, high external DIN impedes the re-establishment of nitrogen-limitation, making recovery from bleaching more difficult (i.e., narrowing the sets of conditions under which recovery is possible).
+8) -- depend on the relative abundance of nitrogen and carbon in the environment.
+Higher food levels, representing a non-autotrophic carbon source for the host, make it easier for the host to overcome carbon-limitation (Fig.
+8A, D), thus providing a potential mechanism underlying observations that feeding aids recovery from bleaching [@Grottoli:2006p3692; @Connolly:2012ep].
+Conversely, high external DIN impedes the re-establishment of nitrogen-limitation, making recovery from bleaching more difficult (i.e., narrowing - or eliminating - conditions under which recovery is possible, Fig.
+8C).
+
+![Hysteresis under different nutrient and feeding regimes.
+Steady state S:H biomass ratios in constant environments with (A) low DIN (1e-7 M) and low food (2e-7 M), (B) low DIN (1e-7 M) and no food, (C) high DIN (2e-6 M) and low food (2e-7 M), and (D) high DIN (2e-6 M) and high food (4e-7 M).
+In each panel, steady states are shown starting from both high initial S:H biomass (1, i.e.
+healthy corals - closed circles) and low initial initial S:H biomass (0.0001, i.e.
+bleached corals - open circles).
+Points are colored according to whether the host exhibits positive (black) or negative (red) growth at steady state.](../img/Fig8.png)
 
 Dynamic simulations of recovery reveal another interesting behavior of the system: under some conditions, an 'overshoot' occurs in which S:H biomass temporarily increases beyond the ratio maintained in the 'healthy' state, before returning to stabilize at this value (Fig.
-8).
+9).
 In fact, unusually high symbiont densities have been observed following bleaching in both experimental [@Cunning:2015ja] and field studies [@Kemp:2014ug], and have been interpreted as a potential 'disequilibrium in host-symbiont regulation' [@Kemp:2014ug].
@@
 If the host overcomes carbon-limitation *before* the symbiont reaches it, then the system rapidly transitions to the nitrogen-limited state and the S:H ratio stabilizes without an overshoot (Fig.
-8A).
+9A).
 However, if the symbiont becomes carbon-limited first (Fig.
-8B, 8C), then carbon translocation per symbiont declines further and photosynthesis becomes more CO~2~-limited, which maintains carbon-limitation of the host.
+9B, 9C), then carbon translocation per symbiont declines further and photosynthesis becomes more CO~2~-limited, which maintains carbon-limitation of the host.
 In this situation, continued growth of less and less productive symbionts drives S:H biomass to a much higher level before the host finally overcomes carbon-limitation.
@@
 high DIN and/or low feeding) therefore magnifies the overshoot and prolongs the dysfunctional, carbon-limited state of the symbiosis (Fig.
-8).
+9).
 On the other hand, factors that favor nitrogen-limitation, such as low external DIN and/or high feeding rates, will accelerate the re-establishment of nitrogen-limitation and prevent an overshoot from occurring at all.
 While many scenarios are possible under different environmental conditions, we illustrate the general effects of varying N and C availability on recovery from bleaching with a series of simulations that vary the N:C ratio of host's heterotrophic food source (Fig.
-8A-C): lower N:C ratios (effectively representing lower DIN and/or higher heterotrophy) favor nitrogen-limitation and more rapid recovery, while higher N:C ratios (effectively representing higher DIN and/or lower heterotrophy) favor carbon-limitation and prolonged recovery with a larger overshoot in S:H biomass.
+9A-C): lower N:C ratios (effectively representing lower DIN and/or higher heterotrophy) favor nitrogen-limitation and more rapid recovery, while higher N:C ratios (effectively representing higher DIN and/or lower heterotrophy) favor carbon-limitation and prolonged recovery with a larger overshoot in S:H biomass.
 
@@
 23.
-All simulations were run with default parameters (except varying $n_{NX}$), with L=20 mol photons m^2^ d^-1^, DIN=1e-7 mol N L^-1^, prey=1e-7 CmolX L^-1^, and initial S:H biomass=0.001.](../img/Fig8.png)
+All simulations were run with default parameters (except varying $n_{NX}$), with L=20 mol photons m^2^ d^-1^, DIN=1e-7 mol N L^-1^, prey=1e-7 CmolX L^-1^, and initial S:H biomass=0.001.](../img/Fig9.png)
 

### 2017-07-11 supp/supp.Rmd
update filenames
Fig8 -> Fig9, FigS9 -> Fig8

--- 2017-03-23 JTB_supp/JTB_supp.Rmd
@@
     toc_depth: 1
-    code_folding: hide
 bibliography: library.bib
@@
 # Figure S5
-```{r sens_LLHN, cache=T, echo=F, eval=T, fig.cap="**Fig. S5. Sensitivity analysis under high DIN (4e-6 M) and low light (5 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
+```{r sens_LLHN, cache=T, echo=F, eval=T, fig.cap="**Fig.
+S5.
+Sensitivity analysis under high DIN (4e-6 M) and low light (5 mol m^-2^ d^-1^).** See Fig.
+4 in main text.
+Open circles represent points at which the parameter value causes the system to move into the negative growth stable state."}
 # Set up time vector and parameters
@@
 # Figure S6
-```{r sens_HLLN, cache=T, echo=F, eval=T, fig.cap="**Fig. S6. Sensitivity analysis under low DIN (1e-7 M) and high light (30 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
+```{r sens_HLLN, cache=T, echo=F, eval=T, fig.cap="**Fig.
+S6.
+Sensitivity analysis under low DIN (1e-7 M) and high light (30 mol m^-2^ d^-1^).** See Fig.
+4 in main text.
+Open circles represent points at which the parameter value causes the system to move into the negative growth stable state."}
 # Set up time vector and parameters
@@
 # Figure S7
-```{r sens_HLHN, cache=T, echo=F, eval=T, fig.cap="**Fig. S7. Sensitivity analysis under high DIN (4e-6 M) and high light (30 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
+```{r sens_HLHN, cache=T, echo=F, eval=T, fig.cap="**Fig.
+S7.
+Sensitivity analysis under high DIN (4e-6 M) and high light (30 mol m^-2^ d^-1^).** See Fig.
+4 in main text.
+Open circles represent points at which the parameter value causes the system to move into the negative growth stable state."}
 # Set up time vector and parameters
@@
 
-# Figure S9
-```{r FigS9, cache=T, fig.width=7, fig.height=7, echo=F, message=F, eval=T, render=NULL, fig.cap="**Fig.
-S9.
-Hysteresis under different nutrient and feeding regimes.** This figure accompanies Fig.
-2 in the main text.
-These figures show steady state S:H biomass ratios in constant environments with varying DIN and food (X) levels, with initial S:H biomass values of either 1 or 0.0001.
-Points are colored according to whether the host exhibits positive (black) or negative (red) growth at steady state."}
-
-# Set up cluster for parallel processing
-cl <- makeCluster(detectCores())  # Initiate cluster
-registerDoParallel(cl)
-
-# Set input values for steady state runs
-input <- expand.grid(L=seq(1,60,1), initS=c(0.0001,1))
-
-# Run steady states in parallel
-output1 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
-  run <- run_coral_ss(env=list(L=input$L[i], X=2e-7, N=1e-7), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
-  list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
-}
-
-output2 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
-  run <- run_coral_ss(env=list(L=input$L[i], X=0, N=1e-7), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
-  list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
-}
-
-output3 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
-  run <- run_coral_ss(env=list(L=input$L[i], X=2e-7, N=2e-6), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
-  list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
-}
-
-output4 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
-  run <- run_coral_ss(env=list(L=input$L[i], X=4e-7, N=2e-6), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
-  list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
-}
-
-stopCluster(cl)  # Stop cluster
-
-# Collect results
-res1 <- cbind(input, output1)
-res2 <- cbind(input, output2)
-res3 <- cbind(input, output3)
-res4 <- cbind(input, output4)
-
-# Create plot
-par(mfrow=c(2,2), mar=c(3,3,1,1), mgp=c(1.5,0.5,0), tcl=-0.025)
-
-plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab="Light", ylab="Steady state S:H ratio")
-title("DIN=1e-7 mol/L; X=2e-7 mol/L")
-points(c(48,48,53,53), c(0.42,0.45,0.42,0.45), pch=c(19,1,19,1), cex=c(0.4,1,0.4,1), col=c("black","black","red","red"), xpd=T)
-text(38, 0.435, labels="init. S:H", xpd=T, cex=0.7, srt=90)
-text(46, c(0.42,0.45), labels=c("0.1", "0.0001"), xpd=T, pos=2, cex=0.7)
-text(50, 0.49, labels=c("Steady state growth"), xpd=T, cex=0.7)
-text(c(48,53), 0.47, labels=c("pos.", "neg."), xpd=T, cex=0.7)
-p1 <- apply(res1, 1, FUN=function(x) {
-  with(x, points(L, sh,
-                 col=ifelse(gr>0, "black", "red"), 
-                 pch=ifelse(initS==1, 19, 1), 
-                 cex=ifelse(initS==1, 0.4, 1)))
-  with(x, if (sh > 0.5) {
-    points(L, 0.5, pch="^", 
-           col=ifelse(gr>0, "black", "red"))
-    points(L, 0.49, 
-           col=ifelse(gr>0, "black", "red"),
-           pch=ifelse(initS==1, 19, 1), 
-           cex=ifelse(initS==1, 0.4, 1))
-  })
-})
-
-plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab="Light", ylab="Steady state S:H ratio")
-title("DIN=1e-7 mol/L; X=0 mol/L")
-p2 <- apply(res2, 1, FUN=function(x) {
-  with(x, points(L, sh,
-                 col=ifelse(gr>0, "black", "red"), 
-                 pch=ifelse(initS==1, 19, 1), 
-                 cex=ifelse(initS==1, 0.4, 1)))
-  with(x, if (sh > 0.5) {
-    points(L, 0.5, pch="^", 
-           col=ifelse(gr>0, "black", "red"))
-    points(L, 0.49, 
-           col=ifelse(gr>0, "black", "red"),
-           pch=ifelse(initS==1, 19, 1), 
-           cex=ifelse(initS==1, 0.4, 1))
-  })
-})
-
-plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab="Light", ylab="Steady state S:H ratio")
-title("DIN=2e-6 mol/L; X=2e-7 mol/L")
-p3 <- apply(res3, 1, FUN=function(x) {
-  with(x, points(L, sh,
-                 col=ifelse(gr>0, "black", "red"), 
-                 pch=ifelse(initS==1, 19, 1), 
-                 cex=ifelse(initS==1, 0.4, 1)))
-  with(x, if (sh > 0.5) {
-    points(L, 0.5, pch="^", 
-           col=ifelse(gr>0, "black", "red"))
-    points(L, 0.49, 
-           col=ifelse(gr>0, "black", "red"),
-           pch=ifelse(initS==1, 19, 1), 
-           cex=ifelse(initS==1, 0.4, 1))
-  })
-})
-
-plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab="Light", ylab="Steady state S:H ratio")
-title("DIN=2e-6 mol/L; X=4e-7 mol/L")
-p4 <- apply(res4, 1, FUN=function(x) {
-  with(x, points(L, sh,
-                 col=ifelse(gr>0, "black", "red"), 
-                 pch=ifelse(initS==1, 19, 1), 
-                 cex=ifelse(initS==1, 0.4, 1)))
-  with(x, if (sh > 0.5) {
-    points(L, 0.5, pch="^", 
-           col=ifelse(gr>0, "black", "red"))
-    points(L, 0.49, 
-           col=ifelse(gr>0, "black", "red"),
-           pch=ifelse(initS==1, 19, 1), 
-           cex=ifelse(initS==1, 0.4, 1))
-  })
-})
-
-```
-
------
-
 # References

### 2017-07-11 R/Fig8.R

@@ -1,87 +1,121 @@
-# Load functions
-library(coRal)
+# Figure 8
 
-# Set parameters and time
-defpars <- coRal::def_pars()  # Get default parameters
-time <- seq(0, 250, 0.1)
+# Run simulations if .RData file does not exist
+if (!file.exists("R/Fig8.RData")) {
+  # Set up cluster for parallel processing
+  cl <- makeCluster(detectCores())  # Initiate cluster
+  registerDoParallel(cl)
+  
+  # Set input values for steady state runs
+  input <- expand.grid(L=seq(1,60,1), initS=c(0.0001,1))
+  
+  # Run steady states in parallel
+  output1 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
+    run <- run_coral_ss(env=list(L=input$L[i], X=2e-7, N=1e-7), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+    list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
+  }
+  
+  output2 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
+    run <- run_coral_ss(env=list(L=input$L[i], X=0, N=1e-7), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+    list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
+  }
+  
+  output3 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
+    run <- run_coral_ss(env=list(L=input$L[i], X=2e-7, N=2e-6), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+    list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
+  }
+  
+  output4 <- foreach(i=1:nrow(input), .combine=rbind) %dopar% {
+    run <- run_coral_ss(env=list(L=input$L[i], X=4e-7, N=2e-6), pars=replace(defpars, "initS", input$initS[i]), dt=0.1)
+    list(gr=last(run$H$dH.Hdt), sh=last(run$S$S/run$H$H))
+  }
+  
+  stopCluster(cl)  # Stop cluster
+  
[…]

### 2017-07-11 R/Fig9.R

@@ -1,0 +1,87 @@
+# Load functions
+library(coRal)
+
+# Set parameters and time
+defpars <- coRal::def_pars()  # Get default parameters
+time <- seq(0, 250, 0.1)
+
+# Set up limitation calculation
+limcoef <- function(x,y,m) log(pmin(x, m)/pmin(y, m))
+
+# Run no overshoot
+env1 <- init_env(time=time, L=c(20,20,0), N=c(1e-7,1e-7,0), X=c(1e-7,1e-7,0))
+run1 <- run_coral(time=time, env=env1, pars=replace(defpars, c("initS", "nNX"), c(0.001, 0.1)))
+run1 <- within(run1, {
+  pl <- limcoef((H$jCO2 + H$rCH)*H$H/S$S + S$rCS, S$jL * pars$yCL, pars$jCPm)
+  sl <- limcoef(pars$yC*S$jCP, (H$rhoN*H$H/S$S + S$rNS)/pars$nNS, pars$jSGm)
+  hl <- limcoef(pars$yC*S$rhoC*S$S/H$H + H$jX, (H$jN + pars$nNX*H$jX + H$rNH) / pars$nNH, pars$jHGm)
+})
+# Run overshoot
+env2 <- init_env(time=time, L=c(20,20,0), N=c(1e-7,1e-7,0), X=c(1e-7,1e-7,0))
+run2 <- run_coral(time=time, env=env2, pars=replace(defpars, c("initS", "nNX"), c(0.001, 0.175)))
+run2 <- within(run2, {
+  pl <- limcoef((H$jCO2 + H$rCH)*H$H/S$S + S$rCS, S$jL * pars$yCL, pars$jCPm)
+  sl <- limcoef(pars$yC*S$jCP, (H$rhoN*H$H/S$S + S$rNS)/pars$nNS, pars$jSGm)
+  hl <- limcoef(pars$yC*S$rhoC*S$S/H$H + H$jX, (H$jN + pars$nNX*H$jX + H$rNH) / pars$nNH, pars$jHGm)
+})
+# Run larger overshoot
+env3 <- init_env(time=time, L=c(20,20,0), N=c(1e-7,1e-7,0), X=c(1e-7,1e-7,0))
+run3 <- run_coral(time=time, env=env3, pars=replace(defpars, c("initS", "nNX"), c(0.001, 0.22)))
+run3 <- within(run3, {
[…]

### 2017-07-11 R/Fig8.R
format panel and axis labels

@@ -45,8 +45,9 @@ if (!file.exists("R/Fig8.RData")) {
 png("img/Fig8.png", width=5, height=5, units="in", res=300)
 par(mfrow=c(2,2), mar=c(3,3,1,1), mgp=c(1.5,0.5,0), tcl=-0.025)
 
-plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab="Light", ylab="Steady state S:H ratio")
-title("DIN=1e-7 mol/L; X=2e-7 mol/L")
+plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab=expression("Light" ~ (mol~photons~m^{2}~d^{-1})), ylab="Steady state S:H ratio")
+text(x=par("usr")[1]-par("mai")[2]*conv/1.1, y=par("usr")[4]+par("mai")[3]/2.5*conv, adj=c(0,0), 
+     labels="A.", xpd=NA, font=2, cex=1)
 points(c(48,48,53,53), c(0.42,0.45,0.42,0.45), pch=c(19,1,19,1), cex=c(0.4,1,0.4,1), col=c("black","black","red","red"), xpd=T)
 text(38, 0.435, labels="init. S:H", xpd=T, cex=0.7, srt=90)
 text(46, c(0.42,0.45), labels=c("0.1", "0.0001"), xpd=T, pos=2, cex=0.7)
@@ -67,8 +68,9 @@ p1 <- apply(res1, 1, FUN=function(x) {
   })
 })
 
-plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab="Light", ylab="Steady state S:H ratio")
-title("DIN=1e-7 mol/L; X=0 mol/L")
+plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab=expression("Light" ~ (mol~photons~m^{2}~d^{-1})), ylab="Steady state S:H ratio")
+text(x=par("usr")[1]-par("mai")[2]*conv/1.1, y=par("usr")[4]+par("mai")[3]/2.5*conv, adj=c(0,0), 
+     labels="B.", xpd=NA, font=2, cex=1)
 p2 <- apply(res2, 1, FUN=function(x) {
   with(x, points(L, sh,
                  col=ifelse(gr>0, "black", "red"), 
@@ -84,8 +86,9 @@ p2 <- apply(res2, 1, FUN=function(x) {
   })
 })
 
[…]

### 2017-07-11 R/Fig8.R
formatting tweaks

@@ -53,16 +53,16 @@ if (!file.exists("R/Fig8.RData")) {
 
 # Create plot
 png("img/Fig8.png", width=5, height=5, units="in", res=300)
-par(mfrow=c(2,2), tcl=-0.2, cex.main=1, cex.axis=0.6, mar=c(2,2,2,1), mgp=c(1,0.1,0))
-conv <- diff(grconvertX(0:1, 'inches', 'user')) # length of one inch in user coordinates
+par(mfrow=c(2,2), tcl=-0.2, cex.main=1, cex.axis=0.6, mar=c(2.5,2.5,1.5,0.5), mgp=c(1,0.1,0), cex.lab=0.8)
 plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab=expression("Light" ~ (mol~photons~m^{2}~d^{-1})), ylab="Steady state S:H ratio")
-text(x=par("usr")[1]-par("mai")[2]*conv/1.1, y=par("usr")[4]+par("mai")[3]/2.5*conv, adj=c(0,0), 
+#conv <- diff(grconvertX(0:1, 'inches', 'user')) # length of one inch in user coordinates
+text(x=par("usr")[1]-8, y=par("usr")[4]+0.025, adj=c(0,0), 
      labels="A.", xpd=NA, font=2, cex=1)
-points(c(48,48,53,53), c(0.42,0.45,0.42,0.45), pch=c(19,1,19,1), cex=c(0.4,1,0.4,1), col=c("black","black","red","red"), xpd=T)
-text(38, 0.435, labels="init. S:H", xpd=T, cex=0.7, srt=90)
-text(46, c(0.42,0.45), labels=c("0.1", "0.0001"), xpd=T, pos=2, cex=0.7)
-text(50, 0.49, labels=c("Steady state growth"), xpd=T, cex=0.7)
-text(c(48,53), 0.47, labels=c("pos.", "neg."), xpd=T, cex=0.7)
+points(c(45,45,52,52), c(0.39,0.43,0.39,0.43), pch=c(19,1,19,1), cex=c(0.4,1,0.4,1), col=c("black","black","red","red"), xpd=T)
+text(30, 0.415, labels="init. S:H", xpd=T, cex=0.7, srt=90)
+text(45, c(0.39,0.43), labels=c("0.1", "0.0001"), xpd=T, pos=2, cex=0.7)
[…]

### 2017-07-11 ms/ms.Rmd
fix unit error in text
more text revisions

@@
 
-The maximum predicted growth rates of ~0.1 d^-1^, occurring between ~10-25 mol photons m^-2^ s^-1^ light and ~4 µM DIN (Fig.
+The maximum predicted growth rates of ~0.1 d^-1^, occurring between ~10-25 mol photons m^-2^ d^-1^ light and ~4 µM DIN (Fig.
 3A), are comparable to the rate of `r Tanaka_gr` d^-1^ measured by @Tanaka:2007uv in *Acropora pulchra* under similar N-enriched conditions.
@@
 In this context, the 'healthy' stable state represents a scenario in which nitrogen-limitation stabilizes the symbiont to host biomass ratio and maintains positive growth.
-Conversely, carbon-limitation represents a dysfunctional state wherein positive feedbacks result in a loss of photosynthetic capacity and negative growth.
-Importantly, these 'control' mechanisms arise entirely passively through feedbacks present in the system.
+Conversely, carbon-limitation represents a dysfunctional state wherein positive feedbacks result in coral bleaching and negative growth, with hysteresis maintaining this alternate state until a favorable environment is re-established.
+Interestingly, alternate stable states have received much attention in coral reef community ecology with respect to coral- and macroalgal-dominance on reefs [@Mumby:2007cr].
+Here, we find that the same principles may also mediate bleaching and recovery in individual corals.
 
-In addition to the suite of specific examples presented here, the model can be used to explore many different dynamic environmental scenarios, and represents a tool that biologists and ecologists can use to generate hypotheses and make predictions in both experimental and natural settings.
+The model developed here can be used to explore many different dynamic environmental scenarios, and represents a tool that biologists and ecologists can use to generate hypotheses and make predictions in both experimental and natural settings.
 Moreover, parameter values can be modified to correspond to different genetic or functional types of coral hosts and *Symbiodinium* partners in order to evaluate variability in system responses.

### 2017-07-11 ms/ms.tex
remake with revisions

--- 2017-03-22 JTB_ms/JTB_ms.tex
@@
 \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molX C-mol H\textsuperscript{-1} d\textsuperscript{-1}
+C-mol X C-mol H\textsuperscript{-1} d\textsuperscript{-1}
 \strut\end{minipage} &
@@
 \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-molX L\textsuperscript{-1}
+C-mol X L\textsuperscript{-1}
 \strut\end{minipage} &
@@
 d\textsuperscript{-1}, occurring between \textasciitilde{}10-25 mol
-photons m\textsuperscript{-2} s\textsuperscript{-1} light and
+photons m\textsuperscript{-2} d\textsuperscript{-1} light and
 \textasciitilde{}4 µM DIN (Fig. 3A), are comparable to the rate of 0.07
@@
 under which recovery can occur -- which determine the magnitude of
-hysteresis (Fig. S9) -- depend on the relative abundance of nitrogen and
+hysteresis (Fig. 8) -- depend on the relative abundance of nitrogen and
 carbon in the environment. Higher food levels, representing a
 non-autotrophic carbon source for the host, make it easier for the host
-to overcome carbon-limitation, thus providing a potential mechanism
-underlying observations that feeding aids recovery from bleaching
-(Grottoli, Rodrigues, and Palardy 2006; Connolly, Lopez-Yglesias, and
-Anthony 2012). Conversely, high external DIN impedes the
-re-establishment of nitrogen-limitation, making recovery from bleaching
-more difficult (i.e., narrowing the sets of conditions under which
-recovery is possible).
+to overcome carbon-limitation (Fig. 8A, D), thus providing a potential
+mechanism underlying observations that feeding aids recovery from
+bleaching (Grottoli, Rodrigues, and Palardy 2006; Connolly,
+Lopez-Yglesias, and Anthony 2012). Conversely, high external DIN impedes
+the re-establishment of nitrogen-limitation, making recovery from
+bleaching more difficult (i.e., narrowing - or eliminating - conditions
+under which recovery is possible, Fig. 8C).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{../img/Fig8.png}
+\caption{Hysteresis under different nutrient and feeding regimes. Steady
+state S:H biomass ratios in constant environments with (A) low DIN (1e-7
+M) and low food (2e-7 M), (B) low DIN (1e-7 M) and no food, (C) high DIN
+(2e-6 M) and low food (2e-7 M), and (D) high DIN (2e-6 M) and high food
+(4e-7 M). In each panel, steady states are shown starting from both high
+initial S:H biomass (1, i.e.~healthy corals - closed circles) and low
+initial initial S:H biomass (0.0001, i.e.~bleached corals - open
+circles). Points are colored according to whether the host exhibits
+positive (black) or negative (red) growth at steady state.}
+\end{figure}
 
@@
 biomass temporarily increases beyond the ratio maintained in the
-`healthy' state, before returning to stabilize at this value (Fig. 8).
+`healthy' state, before returning to stabilize at this value (Fig. 9).
 In fact, unusually high symbiont densities have been observed following
@@
 rapidly transitions to the nitrogen-limited state and the S:H ratio
-stabilizes without an overshoot (Fig. 8A). However, if the symbiont
-becomes carbon-limited first (Fig. 8B, 8C), then carbon translocation
+stabilizes without an overshoot (Fig. 9A). However, if the symbiont
+becomes carbon-limited first (Fig. 9B, 9C), then carbon translocation
 per symbiont declines further and photosynthesis becomes more
@@
 feeding) therefore magnifies the overshoot and prolongs the
-dysfunctional, carbon-limited state of the symbiosis (Fig. 8). On the
+dysfunctional, carbon-limited state of the symbiosis (Fig. 9). On the
 other hand, factors that favor nitrogen-limitation, such as low external
@@
 availability on recovery from bleaching with a series of simulations
-that vary the N:C ratio of host's heterotrophic food source (Fig. 8A-C):
+that vary the N:C ratio of host's heterotrophic food source (Fig. 9A-C):
 lower N:C ratios (effectively representing lower DIN and/or higher
@@
 \centering
-\includegraphics{../img/Fig8.png}
+\includegraphics{../img/Fig9.png}
 \caption{Recovery from bleaching with varying N:C availability (A:
@@
 and maintains positive growth. Conversely, carbon-limitation represents
-a dysfunctional state wherein positive feedbacks result in a loss of
-photosynthetic capacity and negative growth. Importantly, these
-`control' mechanisms arise entirely passively through feedbacks present
-in the system.
-
-In addition to the suite of specific examples presented here, the model
-can be used to explore many different dynamic environmental scenarios,
-and represents a tool that biologists and ecologists can use to generate
-hypotheses and make predictions in both experimental and natural
-settings. Moreover, parameter values can be modified to correspond to
-different genetic or functional types of coral hosts and
-\emph{Symbiodinium} partners in order to evaluate variability in system
-responses. Thus, the diversity of potential applications for this model
-is high, and we envision this work as a foundation for continued
-development, which may include more detailed treatments of specific
-modules (e.g., DIC processing), and the incorporation of more external
-forcing capabilities (e.g., external DIC, temperature). Importantly,
-open source R code allows this effort to benefit from contributions from
-the wider scientific community, including those with empirical and
-theoretical backgrounds. Ultimately, the continued refinement of these
-tools is fundamental in elucidating the mechanisms of symbiosis function
-and dysfunction, and in predicting coral responses to environmental
-change.
+a dysfunctional state wherein positive feedbacks result in coral
+bleaching and negative growth, with hysteresis maintaining this
+alternate state until a favorable environment is re-established.
+Interestingly, alternate stable states have received much attention in
+coral reef community ecology with respect to coral- and
+macroalgal-dominance on reefs (Mumby, Hastings, and Edwards 2007). Here,
+we find that the same principles may also mediate bleaching and recovery
+in individual corals.
+
+The model developed here can be used to explore many different dynamic
+environmental scenarios, and represents a tool that biologists and
+ecologists can use to generate hypotheses and make predictions in both
+experimental and natural settings. Moreover, parameter values can be
+modified to correspond to different genetic or functional types of coral
+hosts and \emph{Symbiodinium} partners in order to evaluate variability
+in system responses. Thus, the diversity of potential applications for
+this model is high, and we envision this work as a foundation for
+continued development, which may include more detailed treatments of
+specific modules (e.g., DIC processing), and the incorporation of more
+external forcing capabilities (e.g., external DIC, temperature).
+Importantly, open source R code allows this effort to benefit from
+contributions from the wider scientific community, including those with
+empirical and theoretical backgrounds. Ultimately, the continued
+refinement of these tools is fundamental in elucidating the mechanisms
+of symbiosis function and dysfunction, and in predicting coral responses
+to environmental change.
+
+\section{Acknowledgements}\label{acknowledgements}
+
+This work was supported by a NSF Postdoctoral Research Fellowship in
+Biology to RC (\#1400787).
 
@@
 multi-partner reef coral symbiosis under environmental change.''
-\emph{Proceedings of the Royal Society B}, 1--9.
+\emph{Proceedings of the Royal Society B} 282: 20141725.
 doi:\href{https://doi.org/10.1098/rspb.2014.1725\%7B/\&\%7Ddomain=pdf\%7B/\&\%7Ddate_stamp}{10.1098/rspb.2014.1725\{\textbackslash{}\&\}domain=pdf\{\textbackslash{}\&\}date\_stamp}.
@@
 (3). \url{http://scholarspace.manoa.hawaii.edu/handle/10125/2236}.
+
+\hypertarget{ref-Mumby:2007cr}{}
+Mumby, Peter J, Alan Hastings, and Helen J. Edwards. 2007. ``Thresholds
+and the resilience of Caribbean coral reefs.'' \emph{Nature} 450 (7166).
+Nature Publishing Group: 98--101.
+doi:\href{https://doi.org/10.1038/nature06252}{10.1038/nature06252}.
 

### 2017-07-11 R/Fig2.R
tweak legend

@@ -34,10 +34,10 @@ png("img/Fig2.png", width=5, height=5, units="in", res=300)
 plot(NA, xlim=c(20,50), ylim=c(0,0.20), xlab=expression("Light"~(mol~photons~m^{-2}~d^{-1})), 
      ylab=expression("Steady state S:H"~(C-mol~S~C-mol~H^{-1})))
 points(c(43,43,48,48), c(0.17,0.18,0.17,0.18), pch=c(19,1,19,1), cex=c(0.4,1,0.4,1), col=c("black","black","red","red"), xpd=T)
-text(41, 0.20, labels="init. S:H", xpd=T, cex=0.7, pos=2)
+text(35, 0.175, labels="init. S:H", xpd=T, cex=0.7, srt=90)
 text(41, c(0.17,0.18), labels=c("1", "0.0001"), xpd=T, pos=2, cex=0.7)
-text(46, 0.20, labels=c("Steady state growth"), xpd=T, cex=0.7)
-text(c(43,48), 0.19, labels=c(">0", "<0"), xpd=T, cex=0.7)
+text(45, 0.20, labels=c("Steady state growth"), xpd=T, cex=0.7)
+text(c(43,48), 0.19, labels=c("pos.", "neg."), xpd=T, cex=0.7)
 
 apply(res, 1, FUN=function(x) {
   with(x, points(L, sh,

### 2017-07-11 R/Fig8.R
add y-axis units

@@ -53,8 +53,10 @@ if (!file.exists("R/Fig8.RData")) {
 
 # Create plot
 png("img/Fig8.png", width=5, height=5, units="in", res=300)
-par(mfrow=c(2,2), tcl=-0.2, cex.main=1, cex.axis=0.6, mar=c(2.5,2.5,1.5,0.5), mgp=c(1,0.1,0), cex.lab=0.8)
-plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab=expression("Light" ~ (mol~photons~m^{2}~d^{-1})), ylab="Steady state S:H ratio")
+par(mfrow=c(2,2), tcl=-0.2, cex.main=1, cex.axis=0.6, mar=c(2,2.5,1.5,0.5), mgp=c(1,0.1,0), cex.lab=0.8)
+plot(NA, xlim=range(input$L), ylim=c(0,0.5), 
+     xlab=expression("Light" ~ (mol~photons~m^{2}~d^{-1})), 
+     ylab=expression("Steady state S:H"~(C-mol~S~C-mol~H^{-1})))
 #conv <- diff(grconvertX(0:1, 'inches', 'user')) # length of one inch in user coordinates
 text(x=par("usr")[1]-8, y=par("usr")[4]+0.025, adj=c(0,0), 
      labels="A.", xpd=NA, font=2, cex=1)
@@ -78,7 +80,9 @@ p1 <- apply(res1, 1, FUN=function(x) {
   })
 })
 
-plot(NA, xlim=range(input$L), ylim=c(0,0.5), xlab=expression("Light" ~ (mol~photons~m^{2}~d^{-1})), ylab="Steady state S:H ratio")
+plot(NA, xlim=range(input$L), ylim=c(0,0.5), 
+     xlab=expression("Light" ~ (mol~photons~m^{2}~d^{-1})), 
+     ylab=expression("Steady state S:H"~(C-mol~S~C-mol~H^{-1})))
 text(x=par("usr")[1]-8, y=par("usr")[4]+0.025, adj=c(0,0), 
      labels="B.", xpd=NA, font=2, cex=1)
 p2 <- apply(res2, 1, FUN=function(x) {
@@ -96,7 +100,9 @@ p2 <- apply(res2, 1, FUN=function(x) {
   })
 })
 
[…]

### 2017-07-11 ms/ms.Rmd
fig legends

@@
 Alternate stable states occur between ~25-42 mol photons m^-2^ d^-1^ under these conditions (DIN=1e-7 mol N L^-1^; X=1e-7 C-mol X L^-1^), depending on whether initial S:H is high (1, closed circles), representing a healthy coral, or low (0.0001, open circles), representing a bleached coral.
-Arrow above point at L=25 indicates a S:H ratio beyond the axis range; this 'overshoot' phenomenon, in which initially  bleached corals may achieve high S:H ratios while remaining in a carbon-limited state is discussed in the *Coral Bleaching and Recovery* section.](../img/Fig2.png)
+Arrow above point at L=25 indicates a S:H ratio beyond the y-axis range; this 'overshoot' phenomenon, in which initially  bleached corals may achieve high S:H ratios while remaining in a carbon-limited state is discussed in the *Coral Bleaching and Recovery* section.](../img/Fig2.png)
 
@@
 bleached corals - open circles).
-Points are colored according to whether the host exhibits positive (black) or negative (red) growth at steady state.](../img/Fig8.png)
+Points are colored according to whether the host exhibits positive (black) or negative (red) growth at steady state.
+Arrows above points indicate values beyond the y-axis range.](../img/Fig8.png)
 

### 2017-07-11 supp/supp.Rmd
remove old scripts

@@
 # Source default parameters and functions
-sapply(c("R/conversions.R", "R/def_pars.R", "R/init_env.R", "R/plot_results.R",
-         "R/run_coral.R", "R/run_coral_ss.R", "R/sym_PI.R", "R/run_steady_states.R", 
-         "R/plot_steady_states.R", "R/sensitivity.R"), source, .GlobalEnv)
+sapply(c("R/run_steady_states.R", "R/plot_steady_states.R", "R/reported_metrics.R",
+         "R/sensitivity.R", "R/conversions.R"), source, .GlobalEnv)
 # Get default parameters

### 2017-07-11 ms/ms.tex

@@
 healthy coral, or low (0.0001, open circles), representing a bleached
-coral. Arrow above point at L=25 indicates a S:H ratio beyond the axis
+coral. Arrow above point at L=25 indicates a S:H ratio beyond the y-axis
 range; this `overshoot' phenomenon, in which initially bleached corals
@@
 circles). Points are colored according to whether the host exhibits
-positive (black) or negative (red) growth at steady state.}
+positive (black) or negative (red) growth at steady state. Arrows above
+points indicate values beyond the y-axis range.}
 \end{figure}

### 2017-07-11 R/sensitivity.R
use coRal library

@@ -1,9 +1,7 @@
 library(parallel)
 library(doParallel)
 library(foreach)
-source("R/def_pars.R")
-source("R/run_coral_ss.R")
-source("R/init_env.R")
+library(coRal)
 
 # Sensitivity analysis
 

### 2017-07-11 supp/supp.Rmd
load coRal library

@@
 require(stringr)
+require(coRal)
 knitr::opts_knit$set(root.dir = normalizePath(".."))
@@
 # Get default parameters
-defpars <- def_pars()
+defpars <- coRal::def_pars()
 ```
@@
 # Run simulations
-ss1 <- with(run_coral_ss(env=list(L=32,N=0.14e-6,X=1e-6), pars=defpars, dt=0.1), last(S$S/H$H))
-run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", ss1))
-ss2 <- with(run_coral_ss(env=list(L=32,N=15.14e-6,X=1e-6), pars=defpars, dt=0.1), last(S$S/H$H))
-run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", ss2))
+ss1 <- with(coRal::run_coral_ss(env=list(L=32,N=0.14e-6,X=1e-6), pars=defpars, dt=0.1), dplyr::last(S$S/H$H))
+run1 <- coRal::run_coral(time=time, env=env1, pars=replace(defpars, "initS", ss1))
+ss2 <- with(coRal::run_coral_ss(env=list(L=32,N=15.14e-6,X=1e-6), pars=defpars, dt=0.1), dplyr::last(S$S/H$H))
+run2 <- coRal::run_coral(time=time, env=env2, pars=replace(defpars, "initS", ss2))
 

### 2017-07-11 R/sensitivity.R
use coRal::

@@ -8,12 +8,12 @@ library(coRal)
 # Create function to measure sensitivity of responses to specified change in single parameter
 sens <- function(env, pars, par, change) {
   # Run 1 
-  run1 <- run_coral_ss(env=env, pars=pars, dt=0.1)
+  run1 <- coRal::run_coral_ss(env=env, pars=pars, dt=0.1)
   run1ss <- lapply(run1[c("H", "S")], function(x) x[nrow(run1$H), ])
   run1gr <- run1ss$H$dH.Hdt
   run1sh <- run1ss$S$S/run1ss$H$H
   # Run 2 - with changed parameter
-  run2 <- run_coral_ss(env=env, pars=replace(pars, par, with(pars, get(par))*change), dt=0.1)
+  run2 <- coRal::run_coral_ss(env=env, pars=replace(pars, par, with(pars, get(par))*change), dt=0.1)
   run2ss <- lapply(run2[c("H", "S")], function(x) x[nrow(run2$H), ])
   run2gr <- run2ss$H$dH.Hdt
   run2sh <- run2ss$S$S/run2ss$H$H

### 2017-07-11 supp/supp.Rmd
do not return sym_PI to stdout

@@
 The structure of the photosynthesis SU, with the negative feedback between excess light and photosynthesis rate, produces the classic shape of a P-I curve demonstrating photoinhibition, suggesting the mechanism is robust."}
-sym_PI(pars=def_pars())
+sympi <- sym_PI(pars=def_pars())
 ```

### 2017-07-12 supp/supp.Rmd
add CO2 flux/y_C description

@@
 knitr::opts_chunk$set(message = F, warning = F, echo=T, results='asis')
-knitr::opts_chunk$set(results = "asis", render=pander)
+knitr::opts_chunk$set(results = "asis", render=pander, cache=T)
 ```
@@
 This coefficient specifies the efficiency of conversion of photosynthate into biomass at the biomass synthesizing units, which combine photosynthate (CH~2~O), ammonia (NH~3~) and O~2~ to form biomass, water, and CO~2~.
+Because of thermodynamic constraints, there is an energetic overhead on converting carbon into biomass.
 In order to balance this macrochemical equation, there must be a carbon loss, which is accounted for using $y_{C}$ < 1.
@@
 
-Following this macrochemical equation, the amount of CO~2~ produced during biomass formation can be specified as $(1-y_C)j_P/y_C$, where j_P is the production flux of the SU (see equations 13 and 14 in the main text).
+If $y_C$ represents the yield (in C-mol) of biomass per C-mol photosynthate, we assume that the energetic cost of this conversion, met from respiration, will lead to simultaneous production of $(1 – y_C)$ C-mol of CO~2~.
+Thus the ratio of the overhead contribution to CO~2~ flux to the growth flux is $(1 – y_C)/y_C$, and hence the CO~2~ flux ($j_{CO_2}$) can be specified in terms of the growth flux ($j_G$) as $j_{CO_2} = {(1-y_C) j_G \over y_C}$ (see equations 13 and 14 in the main text).
 <br><br>

### 2017-07-12 supp/supp.Rmd
formatting tweaks

@@
 #### $\sigma_{CS}$: Proportion of metabolic CO~2~ produced by symbiont recycled passively to photosynthesis {-}
-Assume that CO~2~ produced within symbiont cells is recycled with high efficiency, potentially aided by symbiont carbon concentrating mechanisms.  
+Assume that CO~2~ produced within symbiont cells is recycled with high efficiency, potentially aided by symbiont carbon concentrating mechanisms:  
 ```{r sigmaCS, echo=F}
@@
 Assume that most metabolic CO~2~ produced within host cells rapidly converts to HCO~3~^-^ at intracellular pH ~ 7.2 [@Venn:2009hy], and that only ~10% actually remains as CO~2~ [@Zeebe:2001wz] that could diffusively supply photosynthesis.
-This passive supply is differentiated from all active host carbon concentrating mechanisms (CCMs) that facilitate CO~2~ supply to photosynthesis, which are separately treated collectively as the $jCO~2~$ flux.
+This passive supply is differentiated from all active host carbon concentrating mechanisms (CCMs) that facilitate CO~2~ supply to photosynthesis, which are separately treated collectively as the $j_{CO_2}$ flux.
 ```{r sigmaCH, echo=F}
@@
 #### $j_{Xm}$: Maximum specific rate of assimilation from host feeding {-}
-@Wijgerde:2011du measured depletion of carbon from captured prey in *Galaxea fascicularis* (including both intracoelenteric and extracoelenteric feeding), reporting a maximum specific carbon assimilation rate of 77.6 µgC cm^-2^ 6h^-1^.
+@Wijgerde:2011du measured depletion of carbon from captured prey in *Galaxea fascicularis* (including both intracoelenteric and extracoelenteric feeding), reporting a maximum specific carbon assimilation rate of 77.6 µgC cm^-2^ 6h^-1^, which converts to:
 ```{r jXm}
@@
 #### $K_X$: Half-saturation constant of assimilation from host feeding {-}
-For consistency, and due to a lack of available data, we use the same value as @Muller:2009io.
+For consistency, and due to a lack of available data, we use the same value as @Muller:2009io:  
 ```{r KX}
@@
 We assume that the host uses a portion of its fixed carbon to generate ATP which is then used to activate a variety of carbon concentrating mechanisms (see @Wooldridge:2013tj).
-We assume these CCMs are efficient and can deliver an order of magnitude more CO~2~ relative to the fixed carbon available.
+We assume these CCMs are efficient and can deliver an order of magnitude more CO~2~ relative to the fixed carbon available:
 ```{r kCO2, echo=F}
@@
 #### $j_{HGm}$: Maximum specific growth rate of host {-}
-We set this to a high value such that constraints on growth are primarily determined by substrate availabilities rather than the maximum value.  
+We set this to a high value such that constraints on growth are primarily determined by substrate availabilities rather than the maximum value:  
 ```{r jHGm, echo=F}
@@
 The maximum theoretical quantum yield is 0.125 mol C fixed per mol photons.
-@Brodersen:2014gw presented a balanced light budget based on microenvironmental photobiology of *Montastrea curta*, and reported a quantum yield of 0.102 O~2~ photon^-1^.
+@Brodersen:2014gw presented a balanced light budget based on microenvironmental photobiology of *Montastrea curta*, and reported a quantum yield of 0.102 O~2~ photon^-1^:
 ```{r yCL, echo=F}
@@
 In order to balance this macrochemical equation, there must be a carbon loss, which is accounted for using $y_{C}$ < 1.
-We take a value of 0.8 following @Muller:2009io.
+We take a value of 0.8 following @Muller:2009io:
+
 ```{r yC, echo=F}
@@
 #### $j_{SGm}$: Maximum specific growth rate of symbiont {-}
-We use the maximum specific growth rate of *Symbiodinium* that is reported by @Chang:1983p8066  
+We use the maximum specific growth rate of *Symbiodinium* that is reported by @Chang:1983p8066:  
 ```{r jSGm, echo=F}
@@
 Larger values of this parameter cause greater biomass loss in response to ROS signaling (i.e., a more severe bleaching response).
-Since this represents symbionts being expelled from the system, we use a value about two orders of magnitude higher than the turnover rate due to maintenance.
+Since this represents symbionts being expelled from the system, we use a value about two orders of magnitude higher than the turnover rate due to maintenance:
 ```{r b, echo=F}

### 2017-07-12 supp/supp.Rmd
add code_folding

@@
   html_document:
+    code_folding: hide
     toc: true
@@
 #### $K_X$: Half-saturation constant of assimilation from host feeding {-}
-For consistency, and due to a lack of available data, we use the same value as @Muller:2009io:  
+For consistency, and due to a lack of available data, we use the same value as @Muller:2009io :  
 ```{r KX}
@@
 In order to balance this macrochemical equation, there must be a carbon loss, which is accounted for using $y_{C}$ < 1.
-We take a value of 0.8 following @Muller:2009io:
+We take a value of 0.8 following @Muller:2009io :
 
@@
 #### $j_{SGm}$: Maximum specific growth rate of symbiont {-}
-We use the maximum specific growth rate of *Symbiodinium* that is reported by @Chang:1983p8066:  
+We use the maximum specific growth rate of *Symbiodinium* that is reported by @Chang:1983p8066 :  
 ```{r jSGm, echo=F}

### 2017-07-12 supp/supp.Rmd
echo=T for chunks

@@
 # Figure S2 
-```{r run1_PI, cache=F, fig.height=4, fig.width=7, echo=F, fig.cap="**Fig.
+```{r run1_PI, cache=F, fig.height=4, fig.width=7, fig.cap="**Fig.
 S2.
@@
 # Figure S3 
-```{r sens_ILHN, cache=T, echo=F, eval=T, fig.cap="**Fig. S3. Sensitivity analysis under high DIN (4e-6 M) and intermediate light (20 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
+```{r sens_ILHN, cache=T, eval=T, fig.cap="**Fig. S3. Sensitivity analysis under high DIN (4e-6 M) and intermediate light (20 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
 # Set up time vector and parameters
@@
 # Figure S4 
-```{r sens_LLLN, cache=T, echo=F, eval=T, fig.cap="**Sensitivity analysis under low DIN (1e-7 M) and low light (5 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
+```{r sens_LLLN, cache=T, eval=T, fig.cap="**Sensitivity analysis under low DIN (1e-7 M) and low light (5 mol m^-2^ d^-1^).** See Fig. 4 in main text."}
 # Set up time vector and parameters
@@
 # Figure S5
-```{r sens_LLHN, cache=T, echo=F, eval=T, fig.cap="**Fig.
+```{r sens_LLHN, cache=T, eval=T, fig.cap="**Fig.
 S5.
@@
 # Figure S6
-```{r sens_HLLN, cache=T, echo=F, eval=T, fig.cap="**Fig.
+```{r sens_HLLN, cache=T, eval=T, fig.cap="**Fig.
 S6.
@@
 # Figure S7
-```{r sens_HLHN, cache=T, echo=F, eval=T, fig.cap="**Fig.
+```{r sens_HLHN, cache=T, eval=T, fig.cap="**Fig.
 S7.
@@
 # Figure S8
-```{r FigS8, cache=T, fig.width=7, fig.height=7, echo=F, eval=T, fig.cap="**Fig.
+```{r FigS8, cache=T, fig.width=7, fig.height=7, eval=T, fig.cap="**Fig.
 S8.

### 2017-07-12 supp/supp.Rmd
fix title, add link to Rmd

@@
 title: "Supplementary Information for:"
-subtitle: A simplified dynamic bioenergetic model for coral-*Symbiodinium* symbioses and coral bleaching as an alternate stable state
+subtitle: A dynamic bioenergetic model for coral-*Symbiodinium* symbioses and coral bleaching as an alternate stable state
 author: "Ross Cunning, Erik B. Muller, Ruth D. Gates, Roger M. Nisbet"
@@
 This supplement provides justification and derivations from the literature for all parameter values used in the model, as well as additional figures that demonstrate the structure and behavior of the model and its specific components.
+R code for each chunk of output in this document can be shown and hidden using the right-aligned "Code" buttons, or for the entire document using the button at the top.
+This HTML output was knit from an RMarkdown document that can be found [here](https://github.com/jrcunning/coRal-analysis/blob/master/supp/supp.Rmd).
 

### 2017-07-13 ms/ms.Rmd
add line numbers

@@
     fig_caption: yes
+header-includes:
+    - \usepackage{lineno}
+    - \linenumbers
 ---

### 2017-07-13 ms/ms.tex
add line numbers

--- 2017-07-11 ms/ms.tex
@@
 \setcounter{secnumdepth}{0}
+\usepackage{lineno}
+\linenumbers
 % Pandoc toggle for numbering sections (defaults to be off)
@@
 % Pandoc header
+\usepackage{lineno}
+\linenumbers
 
@@
 under which recovery can occur -- which determine the magnitude of
-hysteresis (Fig. S9) -- depend on the relative abundance of nitrogen and
+hysteresis (Fig. 8) -- depend on the relative abundance of nitrogen and
 carbon in the environment. Higher food levels, representing a
 non-autotrophic carbon source for the host, make it easier for the host
-to overcome carbon-limitation, thus providing a potential mechanism
-underlying observations that feeding aids recovery from bleaching
-(Grottoli, Rodrigues, and Palardy 2006; Connolly, Lopez-Yglesias, and
-Anthony 2012). Conversely, high external DIN impedes the
-re-establishment of nitrogen-limitation, making recovery from bleaching
-more difficult (i.e., narrowing the sets of conditions under which
-recovery is possible).
+to overcome carbon-limitation (Fig. 8A, D), thus providing a potential
+mechanism underlying observations that feeding aids recovery from
+bleaching (Grottoli, Rodrigues, and Palardy 2006; Connolly,
+Lopez-Yglesias, and Anthony 2012). Conversely, high external DIN impedes
+the re-establishment of nitrogen-limitation, making recovery from
+bleaching more difficult (i.e., narrowing - or eliminating - conditions
+under which recovery is possible, Fig. 8C).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{../img/Fig8.png}
+\caption{Hysteresis under different nutrient and feeding regimes. Steady
+state S:H biomass ratios in constant environments with (A) low DIN (1e-7
+M) and low food (2e-7 M), (B) low DIN (1e-7 M) and no food, (C) high DIN
+(2e-6 M) and low food (2e-7 M), and (D) high DIN (2e-6 M) and high food
+(4e-7 M). In each panel, steady states are shown starting from both high
+initial S:H biomass (1, i.e.~healthy corals - closed circles) and low
+initial initial S:H biomass (0.0001, i.e.~bleached corals - open
+circles). Points are colored according to whether the host exhibits
+positive (black) or negative (red) growth at steady state. Arrows above
+points indicate values beyond the y-axis range.}
+\end{figure}
 
@@
 and maintains positive growth. Conversely, carbon-limitation represents
-a dysfunctional state wherein positive feedbacks result in a loss of
-photosynthetic capacity and negative growth. Importantly, these
-`control' mechanisms arise entirely passively through feedbacks present
-in the system.
-
-In addition to the suite of specific examples presented here, the model
-can be used to explore many different dynamic environmental scenarios,
-and represents a tool that biologists and ecologists can use to generate
-hypotheses and make predictions in both experimental and natural
-settings. Moreover, parameter values can be modified to correspond to
-different genetic or functional types of coral hosts and
-\emph{Symbiodinium} partners in order to evaluate variability in system
-responses. Thus, the diversity of potential applications for this model
-is high, and we envision this work as a foundation for continued
-development, which may include more detailed treatments of specific
-modules (e.g., DIC processing), and the incorporation of more external
-forcing capabilities (e.g., external DIC, temperature). Importantly,
-open source R code allows this effort to benefit from contributions from
-the wider scientific community, including those with empirical and
-theoretical backgrounds. Ultimately, the continued refinement of these
-tools is fundamental in elucidating the mechanisms of symbiosis function
-and dysfunction, and in predicting coral responses to environmental
-change.
+a dysfunctional state wherein positive feedbacks result in coral
+bleaching and negative growth, with hysteresis maintaining this
+alternate state until a favorable environment is re-established.
+Interestingly, alternate stable states have received much attention in
+coral reef community ecology with respect to coral- and
+macroalgal-dominance on reefs (Mumby, Hastings, and Edwards 2007). Here,
+we find that the same principles may also mediate bleaching and recovery
+in individual corals.
+
+The model developed here can be used to explore many different dynamic
+environmental scenarios, and represents a tool that biologists and
+ecologists can use to generate hypotheses and make predictions in both
+experimental and natural settings. Moreover, parameter values can be
+modified to correspond to different genetic or functional types of coral
+hosts and \emph{Symbiodinium} partners in order to evaluate variability
+in system responses. Thus, the diversity of potential applications for
+this model is high, and we envision this work as a foundation for
+continued development, which may include more detailed treatments of
+specific modules (e.g., DIC processing), and the incorporation of more
+external forcing capabilities (e.g., external DIC, temperature).
+Importantly, open source R code allows this effort to benefit from
+contributions from the wider scientific community, including those with
+empirical and theoretical backgrounds. Ultimately, the continued
+refinement of these tools is fundamental in elucidating the mechanisms
+of symbiosis function and dysfunction, and in predicting coral responses
+to environmental change.
 

### 2017-07-14 ms/ms.Rmd
text edits/corrections

@@
 We incorporate these features into a new photo-oxidative stress module linking overreduction of the photosynthetic light reactions to downstream impacts of photoinhibition and photodamage.
-Importantly, this formulation introduces a link between CO~2~-limitation of photosynthesis and bleaching, and potential synergistic roles for heterotrophy and nutrient availability in influencing bleaching responses [@Wooldridge:2014hc]
+Importantly, this formulation introduces a link between CO~2~-limitation of photosynthesis and bleaching, and potential synergistic roles for heterotrophy and nutrient availability in influencing bleaching responses [@Wooldridge:2014hc].
 
@@
 3.
-_Provide well-documented, open source code._ In order to facilitate the continued development and application of theoretical modeling tools for coral symbioses, we provide access to the model in the form of an R package called coRal [github.com/jrcunning/coRal](github.com/jrcunning/coRal), which users may install to run and visualize model simulations.
+_Provide well-documented, open source code._ In order to facilitate the continued development and application of theoretical modeling tools for coral symbioses, we provide access to the model in the form of an R package called coRal ([github.com/jrcunning/coRal](github.com/jrcunning/coRal)), which users may install to run and visualize model simulations.
 With an accessible and modular framework, we envision this code base as a resource for further development by the scientific community to include additional complexity and problem-specific components.
@@
 
-Coral biomass formation is then specified by a parallel complementary SU (formula in Fig.
-3.7 of @Kooijman:2010vd]).
-The general form of this equation is $(m^{-1} + x^{-1} + y^{-1} - (x + y)^{-1})^{-1}$, where $m$ is the maximum production flux, and $x$ and $y$ are the input fluxes of the two substrates.
+Coral biomass formation is then specified by a parallel complementary SU [formula in Fig.
+3.7 of @Kooijman:2010vd].
+The general form for a production flux is $(m^{-1} + x^{-1} + y^{-1} - (x + y)^{-1})^{-1}$, where $m$ is the maximum production flux, and $x$ and $y$ are the input fluxes of the two substrates.
 Coral biomass is thus produced from carbon and nitrogen according to:
@@
 7).
-The parameter $y_C$ specifies the yield of biomass from organic carbon, which we take to be 0.8 to satisfy redox balance (see @Muller:2009io).
+The parameter $y_C$ specifies the yield of biomass from organic carbon, which we take to be 0.8 to satisfy redox balance [see @Muller:2009io].
 
@@
 
-where $j_{CPm}$ is the maximum specific rate of photosynthesis, and $c_{ROS}$ is the relative rate of reactive oxygen species production (see Eq.
+where $j_{CPm}$ is the maximum specific rate of photosynthesis, and $c_{ROS}$ is the relative rate ($\ge1$) of reactive oxygen species production (see Eq.
 18).
@@
 We resolved this problem by *defining the dynamical system* as the limit as a time step $\Delta t \rightarrow 0$ of a discretized system corresponding to Euler integration of the differential equations, with those fluxes that represent flows of elemental matter implemented by assuming that transfer of material between components of the system takes one time step.
-Thus, for example, CO~2~ rejected from the host SU at time $t$ arrives at the photosynthesis SU at time $t + \Delta t$.
+Thus, for example, carbon rejected from the symbiont SU at time $t$ arrives at the host SU at time $t + \Delta t$.
 
-Simulations using the discretized scheme were performed using R code developed in the coRal R package [github.com/jrcunning/coRal](github.com/jrcunning/coRal).
+Simulations using the discretized scheme were performed using R code developed in the coRal R package ([github.com/jrcunning/coRal](github.com/jrcunning/coRal)).
 By experimentation, we found that a time step of 0.1 days gave adequate precision for most simulations (including used to generate Figs.
@@
 3B), as occurs during coral bleaching.
-This reduction in symbionts consequently reduces the flux of fixed carbon to the host, resulting in increasing carbon-limitation (Fig.
-3B) and eventual cessation of growth (Fig.
+This reduction in symbionts consequently reduces the flux of fixed carbon to the host, resulting in increasing carbon-limitation and eventual cessation of growth (Fig.
 3A).
@@
 ![Hysteresis under different nutrient and feeding regimes.
-Steady state S:H biomass ratios in constant environments with (A) low DIN (1e-7 M) and low food (2e-7 M), (B) low DIN (1e-7 M) and no food, (C) high DIN (2e-6 M) and low food (2e-7 M), and (D) high DIN (2e-6 M) and high food (4e-7 M).
+Steady state S:H biomass ratios in constant environments with (A) low DIN (1e-7 mol N L^-1^) and low food (2e-7 C-molX L^-1^), (B) low DIN (1e-7 mol N L^-1^) and no food, (C) high DIN (2e-6 mol N L^-1^) and low food (2e-7 C-molX L^-1^), and (D) high DIN (2e-6 mol N L^-1^) and high food (4e-7 C-molX L^-1^).
 In each panel, steady states are shown starting from both high initial S:H biomass (1, i.e.
@@
 Fig.
-8B, 8C).
-This could explain why corals that have recovered their symbiont populations after bleaching may still exhibit energetic deficits and physiological impacts, possibly for months to years [@Levitan:2014et, @Hughes:2013hd]).
+9B,C).
+This could explain why corals that have recovered their symbiont populations after bleaching may still exhibit energetic deficits and physiological impacts, possibly for months to years [@Levitan:2014et; @Hughes:2013hd].
 These findings suggest that host acquisition of carbon from a source other than the symbiont may be extremely important for the system to recover from bleaching.

### 2017-07-14 ms/ms.tex
text edits/corrections

@@
   potential synergistic roles for heterotrophy and nutrient availability
-  in influencing bleaching responses (Wooldridge 2014b)
+  in influencing bleaching responses (Wooldridge 2014b).
 \item
@@
   the form of an R package called coRal
-  \url{github.com/jrcunning/coRal}, which users may install to run and
+  (\url{github.com/jrcunning/coRal}), which users may install to run and
   visualize model simulations. With an accessible and modular framework,
@@
 Coral biomass formation is then specified by a parallel complementary SU
-(formula in Fig. 3.7 of Kooijman (2010){]}). The general form of this
-equation is \((m^{-1} + x^{-1} + y^{-1} - (x + y)^{-1})^{-1}\), where
-\(m\) is the maximum production flux, and \(x\) and \(y\) are the input
-fluxes of the two substrates. Coral biomass is thus produced from carbon
-and nitrogen according to:
+(formula in Fig. 3.7 of Kooijman 2010). The general form for a
+production flux is \((m^{-1} + x^{-1} + y^{-1} - (x + y)^{-1})^{-1}\),
+where \(m\) is the maximum production flux, and \(x\) and \(y\) are the
+input fluxes of the two substrates. Coral biomass is thus produced from
+carbon and nitrogen according to:
 
@@
 organic carbon, which we take to be 0.8 to satisfy redox balance (see
-Muller et al. (2009)).
+Muller et al. 2009).
 
@@
 where \(j_{CPm}\) is the maximum specific rate of photosynthesis, and
-\(c_{ROS}\) is the relative rate of reactive oxygen species production
-(see Eq. 18). Dividing the photosynthetic rate by \(c_{ROS}\) causes a
-decline in response to photooxidative stress at high light levels, and
-the emergent outcome of this SU formulation demonstrates a classic
-photoinhibition response (Fig. S2).
+\(c_{ROS}\) is the relative rate (\(\ge1\)) of reactive oxygen species
+production (see Eq. 18). Dividing the photosynthetic rate by \(c_{ROS}\)
+causes a decline in response to photooxidative stress at high light
+levels, and the emergent outcome of this SU formulation demonstrates a
+classic photoinhibition response (Fig. S2).
 
@@
 between components of the system takes one time step. Thus, for example,
-CO\textsubscript{2} rejected from the host SU at time \(t\) arrives at
-the photosynthesis SU at time \(t + \Delta t\).
+carbon rejected from the symbiont SU at time \(t\) arrives at the host
+SU at time \(t + \Delta t\).
 
 Simulations using the discretized scheme were performed using R code
-developed in the coRal R package \url{github.com/jrcunning/coRal}. By
+developed in the coRal R package (\url{github.com/jrcunning/coRal}). By
 experimentation, we found that a time step of 0.1 days gave adequate
@@
 reduces the flux of fixed carbon to the host, resulting in increasing
-carbon-limitation (Fig. 3B) and eventual cessation of growth (Fig. 3A).
+carbon-limitation and eventual cessation of growth (Fig. 3A).
 
@@
 state S:H biomass ratios in constant environments with (A) low DIN (1e-7
-M) and low food (2e-7 M), (B) low DIN (1e-7 M) and no food, (C) high DIN
-(2e-6 M) and low food (2e-7 M), and (D) high DIN (2e-6 M) and high food
-(4e-7 M). In each panel, steady states are shown starting from both high
-initial S:H biomass (1, i.e.~healthy corals - closed circles) and low
-initial initial S:H biomass (0.0001, i.e.~bleached corals - open
-circles). Points are colored according to whether the host exhibits
-positive (black) or negative (red) growth at steady state. Arrows above
-points indicate values beyond the y-axis range.}
+mol N L\textsuperscript{-1}) and low food (2e-7 C-molX
+L\textsuperscript{-1}), (B) low DIN (1e-7 mol N L\textsuperscript{-1})
+and no food, (C) high DIN (2e-6 mol N L\textsuperscript{-1}) and low
+food (2e-7 C-molX L\textsuperscript{-1}), and (D) high DIN (2e-6 mol N
+L\textsuperscript{-1}) and high food (4e-7 C-molX
+L\textsuperscript{-1}). In each panel, steady states are shown starting
+from both high initial S:H biomass (1, i.e.~healthy corals - closed
+circles) and low initial initial S:H biomass (0.0001, i.e.~bleached
+corals - open circles). Points are colored according to whether the host
+exhibits positive (black) or negative (red) growth at steady state.
+Arrows above points indicate values beyond the y-axis range.}
 \end{figure}
@@
 alone does not necessarily indicate that a `healthy' state has been
-reached, since the carbon-limited state may still persist (e.g.~Fig. 8B,
-8C). This could explain why corals that have recovered their symbiont
+reached, since the carbon-limited state may still persist (e.g.~Fig.
+9B,C). This could explain why corals that have recovered their symbiont
 populations after bleaching may still exhibit energetic deficits and
 physiological impacts, possibly for months to years (Levitan et al.
-2014, Hughes and Grottoli (2013))). These findings suggest that host
+2014; Hughes and Grottoli 2013). These findings suggest that host
 acquisition of carbon from a source other than the symbiont may be
